### 2020-09-11 manuscripts/owp_fim3_2020/templates/April-30-2019-latex-templates/April 16 2019/trackchanges-0.7.0/documentation/web/screenshots/demo_simple.tex
initial commit with template as start

\documentclass[a4paper,10pt]{article}

\usepackage{url}

\usepackage{trackchanges}
\renewcommand{\initialsOne}{FS}
\renewcommand{\initialsTwo}{PS}

\begin{document}

\title{Demonstration of trackchanges.sty}
\author{Felix Salfner}

\maketitle

\begin{abstract} \noindent The intention of this text is to demonstrate the use of \texttt{trackchanges.sty}. The package defines five editing commands:
\end{abstract}

\section{Introduction}

While planning to write some complex document together with some of my colleagues the same discussion was started over and over again: should \LaTeX\ or some office suite like OpenOffice.org be used? Strengths of \LaTeX\ are its deterministic behavior, its reliable handling of split documents and unreached typesetting of formulas. On the other hand, current office suites provide the user with several features that are at least very desirable when collaboratively writing a document: they provide integrated merging facilities and are able to track changes and attach notes to the text. The merging issue can be reasonably handled by version tracking systems like SVN or CVS, but there was no acceptable solution to the issue of change tracking available. Of course, some  \LaTeX\ purists tried to convince me that all change tracking can be handled by insertion of \LaTeX\ \textit{comments}. I have tried to handle one project like this but it did not work out! The main reason was that reading and editing of large documents is mostly handled in DVI format and not on the \LaTeX\ source level -- but \LaTeX\ comments cannot be seen in DVI! Especially, if you have sent one version of the document to a colleague and you want to skim quickly over it in order to see what has been changed.

While returning from a project meeting and staring out of the train's window I had the idea how we could combine the ``best of both worlds'' for collaborative text editing: by adding change tracking and note facilities to \LaTeX ! This is the basic idea of the \texttt{trackchanges} \LaTeX\ package. But this is only one part of the change tracking convenience offered by an office suite. The second part of the story is that changes and notes need to be accepted or rejected! This is the goal of the other programs of the \textbf{trackchanges} open source project hosted on sourceforge\footnote{Please visit  \url{http://trackchanges.sourceforge.net}}

\end{document}
%This document is part of the trackchanges sourceforge project and is published under GNU Public License version 2 or above

### 2020-09-11 manuscripts/owp_fim3_2020/templates/April-30-2019-latex-templates/April 16 2019/trackchanges-0.7.0/testing/testing_old.tex
initial commit with template as start

\documentclass{report}

\usepackage{color} % Allow text colors

%\usepackage{fullpage}

%\usepackage[finalnew]{../LatexPackage/trackchanges}
%\usepackage[finalold]{../LatexPackage/trackchanges}
%\usepackage[footnotes]{../LatexPackage/trackchanges}
%\usepackage[inline]{../LatexPackage/trackchanges}
%\usepackage[margins]{../LatexPackage/trackchanges}
%\usepackage[margins, movemargins]{../LatexPackage/trackchanges}
\usepackage[margins, adjustmargins]{../LatexPackage/trackchanges}

% Setup commands for 6 editors (one more than allowed).
\addeditor{novi}
\addeditor{ym}
\addeditor{three}
\addeditor{quatro}
\addeditor{pyat}
\addeditor{one more}

\title{TrackChanges Testing Document}
\author{Monty Python's Flying Circus}
\date{April, 2009}

\begin{document}
\maketitle

\chapter{Court scene - multiple murderer}

\emph{Cut to a courtroom. Severe atmosphere.}

\noindent\\ Judge:

Michael Norman Randall, you have been found guilty of the murder of Arthur Reginald Webster, Charles Patrick Trumpington, Marcel Agnes Bernstein, Lewis Anona Rudd, John Malcolm Kerr, Nigel Sinclair Robinson, Norman Arthur Potter, Felicity Jayne Stone, Jean-Paul Reynard, Rachel Shirley Donaldson, Stephen Jay Greenblatt, Karl-Heinz Mullet, Belinda Anne Ventham, Juan-Carlos Fernandez, Thor Olaf Stensgaard, Lord Kimberley of Pretoria, Lady Kimberley of Pretoria, The Right Honourable Nigel Warmsly Kimberley, Robert Henry Noonan and Felix James Bennett, on or about the morning of the 19th December 1972. Have you anything to say before I pass sentence?

\noindent\\ Randall:

Yes, sir. I'm very sorry.

\noindent\\ Judge:

Very sorry?

\noindent\\ Randall:

Yes, sir. It was a very very bad thing to have done and I'm really very ashamed of myself. I can only say it won't happen again. To have murdered so many people in such a short space of time is really awful, and I really am very, very, very sorry that I did it, and also that I've taken up so much of the court's valuable time listening to the sordid details of these senseless killings of mine. I would particularly like to say, a very personal and sincere 'sorry' to you, m'lud, for my appalling behaviour throughout this trial. I'd also like to say sorry to the police, for putting them to so much trouble (shot of three heavily bandaged exhausted-looking policemen behind him) for the literally hours of work they've had to put in, collecting evidence and identifying corpses and so forth. You know I think sometimes we ought to realize the difficult and often dangerous work involved in tracking down violent criminals like myself and I'd just like them to know that their fine work is at least appreciated by me.

\noindent\\\emph{The policemen look embarrassed.}

\noindent\\ First Policeman:

No, no, we were only doing our job.

\noindent\\ Second Policeman:

No, no, no, no.

\noindent\\ Randall:

It's very good of you to say that, but I know what you've been through.

\noindent\\ First Policeman:

No, no, we've had worse.

\noindent\\ Third Policeman:

It was plain sailing apart from the arrest.

\noindent\\ Randall:

I know and I'm grateful. I'd like to apologize too to the prosecuting counsel for dragging him in here morning after morning in such lovely weather.

\noindent\\ Counsel:

Well, I would have had to come in anyway.

\noindent\\ Randall:

Ah good, but what a presentation of a case!

\noindent\\ Counsel:

Oh thank you.

\noindent\\ Randall:

No, no, it's a privilege to watch you in action. I never had a chance.

\noindent\\ Counsel:

Oh yes you did.

\noindent\\ Randall:

Not after that summing up. Great.

\noindent\\ Counsel:

Oh thank you. (very chuffed)

\noindent\\ Randall:

And now I must come to the jury. What can I say. I've dragged you in here, day after day, keeping you away from your homes, your jobs, your loved ones, just to hear the private details of my petty atrocities.

\noindent\\ Foreman:

No, no, it was very interesting.

\noindent\\ Randall:

But you could have had a much nicer case.

\noindent\\ Foreman:

No, no, murder's much more fun.

\noindent\\ First Juryman:

Yes and so many of them.

\noindent\\ Second Juryman:

Excellent.

\noindent\\ Third Juryman:

We've had a terrific time. (the jury applauds)

\noindent\\ Randall:

(blows his nose, does a Dickie Attenborough) I'm sorry, I'm very moved. And so, m'lud, it only remains for you to pass the most savage sentence on me that the law can provide.

\noindent\\ Judge:

Well er... not necessarily.

\noindent\\ Randall:

No, m'lud, the full penalty of the law is hardly sufficient. I insist I must be made an example of.

\noindent\\ Judge:

Well yes and no. I mean society at large...

\noindent\\ Randall:

Oh no, m'lud. Not with mass murder.

\noindent\\ Judge:

But in this case, (to court) don't you think?

\noindent\\ Court:

Yes, yes!

\noindent\\ Randall:

Oh, come on, m'lud, you've got to give me life.

\noindent\\ Court:

No, no, no, no.

\noindent\\ Randall:

(to court at large) Well, ten years at least.

\noindent\\ Judge:

Ten years!

\noindent\\ Court:

Shame. Shame!

\noindent\\ Randall:

Well five then. Be fair.

\noindent\\ Judge:

No, no. I'm giving you three months.

\noindent\\ Randall:

Oh no, that's so embarrassing. I won't hear of it. Give me six...please.

\noindent\\ Judge:

Well, all right. Six months.

\noindent\\ Randall:

Thank you, m'lud.

\noindent\\ Judge:

But suspended.

\noindent\\  Randall:

Oh no.

\noindent\\ Court:

Hooray. (they applaud)

\noindent\\ Foreman:

Three cheers for the defendant. Hip. Hip.

\noindent\\ Court:

Hooray.

\noindent\\ Foreman:

Hip. Hip.

\noindent\\ Court:

Hooray.

\noindent\\ Foreman:

Hip. Hip.

\noindent\\ Court:

Hooray.

\noindent\\ All:

For he's a jolly good fellow, For he's a jolly good fellow, For he's a jolly good fellow...

\noindent\\ Voice \emph{(off)}:

Which nobody can deny.

\end{document}

### 2020-09-17 lib/run_by_unit.sh
Squashed commit of the following:
commit c5674cffb5ea719c1cf6e713d5bc8e1813d9ee6d
Author: fernando-aristizabal <fernando-aristizabal@noaa.gov>
Date:   Thu Sep 17 15:38:33 2020 +0000

    Adds help instructions for preprocessing.
    Added preprocessing usage and USGS NHDPlusHR reference
    One last try for indentation on Readme. Need to find a tool to preview locally. My bad for extra commits
    Patches indentation on README
    Patches to Readme file

    Updates to Readme
    * Additional usage
    * Testing information
    * Credits and References

    Patches non-zero exit code in 180500

@@ -67,7 +67,7 @@ echo -e $startDiv"Convert DEM to Meters $hucNumber"$stopDiv
 date -u
 Tstart
 [ ! -f $outputHucDataDir/dem_meters.tif ] && \
-gdal_calc.py --quiet --type=Float32 --co "BLOCKXSIZE=512" --co "BLOCKYSIZE=512" --co "TILED=YES" --co "COMPRESS=LZW" --co "BIGTIFF=YES" -A $outputHucDataDir/dem.tif --outfile="$outputHucDataDir/dem_meters.tif" --calc="((float32(A)*(float32(A)>$ndv))/100)+((float32(A)<=$ndv)*$ndv)" --NoDataValue=$ndv
+gdal_calc.py --quiet --type=Float32 --co "BLOCKXSIZE=512" --co "BLOCKYSIZE=512" --co "TILED=YES" --co "COMPRESS=LZW" --co "BIGTIFF=YES" -A $outputHucDataDir/dem.tif --outfile="$outputHucDataDir/dem_meters.tif" --calc="A/100" --NoDataValue=$ndv
 Tcount
 
 ## RASTERIZE REACH BOOLEAN (1 & 0) ##
@@ -114,8 +114,8 @@ Tcount
 echo -e $startDiv"D8 Flow Directions on Burned DEM $hucNumber"$stopDiv
 date -u
 Tstart
-[ ! -f $outputHucDataDir/flowdir_d8_burned_filled.tif ] && [ ! -f $outputHucDataDir/slopes_d8_burned_filled.tif ] && \
-mpiexec -n $ncores_fd $taudemDir2/d8flowdir -fel $outputHucDataDir/dem_burned_filled.tif -p $outputHucDataDir/flowdir_d8_burned_filled.tif -sd8 $outputHucDataDir/slopes_d8_burned_filled.tif
+[ ! -f $outputHucDataDir/flowdir_d8_burned_filled.tif ] && \
+mpiexec -n $ncores_fd $taudemDir2/d8flowdir -fel $outputHucDataDir/dem_burned_filled.tif -p $outputHucDataDir/flowdir_d8_burned_filled.tif
 Tcount
 
 ## MASK BURNED DEM FOR STREAMS ONLY ###
@@ -132,6 +132,13 @@ date -u
 Tstart
 [ ! -f $outputHucDataDir/dem_thalwegCond.tif ] && \
[…]

### 2020-09-25 lib/run_by_unit.sh
Moved levee clip processes - Cleaning up run_by_unit workflow for levee - Clip processing moved to snap_and_clip_to_nhd.py - Deleted unused development code blocks

@@ -39,7 +39,7 @@ echo -e $startDiv"Get Vector Layers and Subset $hucNumber"$stopDiv
 date -u
 Tstart
 [ ! -f $outputHucDataDir/demDerived_reaches.shp ] && \
-$libDir/snap_and_clip_to_nhd.py -d $hucNumber -w $input_NWM_Flows -s $input_NHD_Flowlines -v $input_NHD_VAA -l $input_NWM_Lakes -u $outputHucDataDir/wbd.gpkg -c $outputHucDataDir/NHDPlusBurnLineEvent_subset.gpkg -a $outputHucDataDir/nwm_lakes_proj_subset.gpkg -t $outputHucDataDir/nwm_headwaters_proj_subset.gpkg -m $input_NWM_Catchments -n $outputHucDataDir/nwm_catchments_proj_subset.gpkg -e $outputHucDataDir/nhd_headwater_points_subset.gpkg -b $outputHucDataDir/nwm_subset_streams.gpkg
+$libDir/snap_and_clip_to_nhd.py -d $hucNumber -w $input_NWM_Flows -s $input_NHD_Flowlines -v $input_NHD_VAA -l $input_NWM_Lakes -r $input_NLD -u $outputHucDataDir/wbd.gpkg -c $outputHucDataDir/NHDPlusBurnLineEvent_subset.gpkg -a $outputHucDataDir/nwm_lakes_proj_subset.gpkg -z $outputHucDataDir/nld_subset_levees.gpkg -t $outputHucDataDir/nwm_headwaters_proj_subset.gpkg -m $input_NWM_Catchments -n $outputHucDataDir/nwm_catchments_proj_subset.gpkg -e $outputHucDataDir/nhd_headwater_points_subset.gpkg -b $outputHucDataDir/nwm_subset_streams.gpkg
 Tcount
 
 ## Clip WBD8 ##
@@ -50,11 +50,11 @@ ogr2ogr -f GPKG -clipsrc $outputHucDataDir/wbd_buffered.gpkg $outputHucDataDir/w
 Tcount
 
 ## Clip NLD ##
-echo -e $startDiv"Clip NLD"$stopDiv
-date -u
-Tstart
-ogr2ogr -f GPKG -nlt MULTILINESTRINGZ -clipsrc $outputHucDataDir/wbd_buffered.gpkg $outputHucDataDir/nld_clp.gpkg $input_NLD nld_preprocessed
-Tcount
+#echo -e $startDiv"Clip NLD"$stopDiv
[…]

### 2020-09-25 lib/snap_and_clip_to_nhd.py

@@ -10,7 +10,7 @@ from shapely.strtree import STRtree
 from shapely.geometry import Point,MultiLineString,LineString,mapping
 from derive_headwaters import findHeadWaterPoints
 
-def subset_vector_layers(hucCode,nwm_streams_fileName,nhd_streams_fileName,nhd_streams_vaa_fileName,nwm_lakes_fileName,nwm_catchments_fileName,wbd_fileName,subset_nhd_streams_fileName,subset_nwm_lakes_fileName,subset_nwm_headwaters_fileName,subset_nwm_catchments_fileName,subset_nwm_streams_fileName,subset_nhd_headwaters_fileName=None,dissolveLinks=False):
+def subset_vector_layers(hucCode,nwm_streams_fileName,nhd_streams_fileName,nhd_streams_vaa_fileName,nwm_lakes_fileName,nld_lines_fileName,nwm_catchments_fileName,wbd_fileName,subset_nhd_streams_fileName,subset_nwm_lakes_fileName,subset_nld_lines_fileName,subset_nwm_headwaters_fileName,subset_nwm_catchments_fileName,subset_nwm_streams_fileName,subset_nhd_headwaters_fileName=None,dissolveLinks=False):
 
     hucUnitLength = len(str(hucCode))
 
@@ -23,6 +23,13 @@ def subset_vector_layers(hucCode,nwm_streams_fileName,nhd_streams_fileName,nhd_s
     if not nwm_lakes.empty:
         nwm_lakes.to_file(subset_nwm_lakes_fileName,driver=getDriver(subset_nwm_lakes_fileName),index=False)
     del nwm_lakes
+    
+    # find intersecting levee lines
+    print("Subsetting NLD levee lines for HUC{} {}".format(hucUnitLength,hucCode),flush=True)
+    nld_lines = gpd.read_file(nld_lines_fileName, mask = wbd)
+    if not nld_lines.empty:
+        nld_lines.to_file(subset_nld_lines_fileName,driver=getDriver(subset_nld_lines_fileName),index=False)
+    del nld_lines
 
     # find intersecting nwm_catchments
     print("Subsetting NWM Catchments for HUC{} {}".format(hucUnitLength,hucCode),flush=True)
@@ -203,11 +210,13 @@ if __name__ == '__main__':
     parser.add_argument('-s','--nhd-streams',help='Basins polygons to use within project path',required=True)
[…]

### 2020-11-10 manuscripts/owp_fim3_2020/sec/abstract_plainTextSummary.tex
Split main tex file into separate files

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT and PLAIN LANGUAGE SUMMARY
%
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have 
% a background in your field.
%
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
%
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
Historically, floods have led to significant loss of life and property in the United States and around the world. 
With the expectation that climate change will only increase the intensity of extreme precipitation events, flooding will continue to play a major risk \cite{Tabari1}. 
In response to this growing issue, the National Water Model Version 2.1, released in 2021, produces forecast discharges at over 2.8 million river reaches over the continental United States, Hawaii, and Puerto Rico as well as parts of Canada and Mexico. 
The NWM is a hydrologic and hydraulic model implemented one-dimensionally to reduce computational costs inherent to continental scale modeling. 
To provide guidance to the general public on flood inundation extents, a rapid methodology is desired to convert forecast discharges to river stages and those stages to inundation extents. Relative elevation models (REM), specifically Height Above Nearest Drainage (HAND), among others, detrend digital elevation models (DEM) to the nearest thalweg elevation within the relevant drainage area. 
Generating REM's and reach level catchments enables river stages to translate to extents by assuming a consistent stage profile within each river reach. 
Stages are translated from NWM discharges via synthetic rating curves that establish stage-discharge relationships via the Manning's equation with geometric properties sampled from the REM on a per catchment basis. The National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) is currently on it's third FIM version providing version improvement
\end{abstract}

\section*{Plain Language Summary}
[ enter your Plain Language Summary here or delete this section]

### 2020-11-10 manuscripts/owp_fim3_2020/sec/keypoints.tex
Split main tex file into separate files

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item The National Water Model produces forecasts at more than 2.8 million river reaches in the United States.
\item OWP FIM produces flood inundation maps by converting discharges to stages and stages to extents and depths.
\item FIM 3 is a modified version of the Height Above Nearest Drainage technique that provides version improvement with FIM 1 and 2.
\end{keypoints}

### 2020-11-10 manuscripts/owp_fim3_2020/sec/title_authors.tex
Split main tex file into separate files

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{High Resolution, Rapid Flood Inundation Mapping for the U.S. National Water Model: From Forecast River Discharges to Forecast Inundation}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3},Trevor Grout\affil{1,3}, Nick Chadwick\affil{1,3}, Zachary Wills\affil{1,3}, Gregory Petrochenkov\affil{4}, Roland Viger\affil{4}, Jasmeet Judge\affil{2}}

\affiliation{1}{Lynker Technologies, Boulder, CO, USA}
\affiliation{2}{Center for Remote Sensing, University of Florida, Gainesville, FL, USA}
\affiliation{3}{National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
\affiliation{4}{Water Mission Area, United States Geological Survey, Boulder, CO, USA}

%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{Fernando Aristizabal}{fernando.aristizabal@noaa.gov}

### 2020-11-30 manuscripts/owp_fim3_2020/sec/abstract_plainTextSummary.tex
Various changes * added to introduction * added more references * added figures directory * deleted variables file

@@
 Historically, floods have led to significant loss of life and property in the United States and around the world. 
-With the expectation that climate change will only increase the intensity of extreme precipitation events, flooding will continue to play a major risk \cite{Tabari1}. 
+With the expectation that climate change will only increase the intensity of extreme precipitation events, flooding will continue to play a major risk. 
 In response to this growing issue, the National Water Model Version 2.1, released in 2021, produces forecast discharges at over 2.8 million river reaches over the continental United States, Hawaii, and Puerto Rico as well as parts of Canada and Mexico.

### 2020-11-30 manuscripts/owp_fim3_2020/sec/introduction.tex
Various changes * added to introduction * added more references * added figures directory * deleted variables file

\section{Introduction}

Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
In 2017 and 2019, river and flash flooding combined represented the leading cause of death among all natural disasters and in 2018 they were only second to deaths from heat wave\cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}. 
More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
Within that 10 year window, the most common activity victims were partaking in is driving year after year when compared to walking/hiking, at home, boating, fell in, working, and other \cite{us_department_of_commerce_2020}.
With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}. 
The large spike in damages for 2017 can be attributed to the Hurricane Harvey event that primarily affected Texas in August. 
Unencouragingly, the trends related to flood damages and fatalities have been steadily increasing over recent decades. \cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}. 
Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}. 
Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.

Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with approximately 3,781 forecast points across the US at typically short forecast horizons of 24 or 72 hours\cite{mcenery2005noaa}.
AHPS provides forecasting services in the form of ensemble streamflows at 3,781 forecast points illustrated in \ref{fig:all_ahps_points} and flood inundation maps (FIM) at 188 of those forecast points shown in \ref{fig:fim_ahps_points}.
AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multisensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
Despite the AHPS advances in operational flood forecasting its significant limitations are the lack of spatial coverageand short forecast time horizons.
Approximately, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.

Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
To broaden the forecasting domain, the Office of Water Prediction (OWP) at the National Water Center (NWC) in Tuscaloosa, Alabama commissioned the utilization of the National Water Model (NWM) which is an instance of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}. 
The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
The NWM enhances the spatial and temporal domain of the current AHPS capabilities at the 13 River Forecast Centers (RFC) in areas known as "hydro-blind" and is not meant to be a replacement. 
It's simply an additional model to be used in the forecasting and early warning decision making.

The National Hydrography Dataset Plus (NHDPlus) version 2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
The term hydrofabric is used within the NWM jargon to describe the collection of geospatial datasets required for modeling including but not limited to stream networks, catchments, channel properties, and elevation data. 
The routing methods used within the NWM for its stream network is Muskingam-Cunge (MC) routing \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.

\begin{figure}
\centering
\includegraphics[scale=2.0]{figs/ahps_all_forecast_points.jpg}
\label{fig:all_ahps_points}
\caption{All 3,781 forecast points in United States' Advanced Hydrologic Prediction System for lower 48 states and Alaska. No forecast points in Hawaii or territories. From: water.weather.gov/ahps}
\end{figure}

\begin{figure}
\centering
\includegraphics[scale=2.0]{figs/ahps_fim_forecast_points_conus.jpg}
\label{fig:fim_ahps_points}
\caption{All 188 forecast points in United States' Advanced Hydrologic Prediction System with flood inundation mapping capabilities. One forecast point near Juneau, Alaska not shown. No forecast points in Hawaii or remaining territories. From: water.weather.gov/ahps}
\end{figure}

### 2020-11-30 manuscripts/owp_fim3_2020/Makefile.sh

@@ -1,0 +1,14 @@
+#!/bin/bash -e
+
+fullFile=owp_fim3.tex
+baseName="${fullFile%%.*}"
+extension="${fullFile#*.}"
+
+pdflatex "$fullFile"
+bibtex "$baseName".aux
+pdflatex "$fullFile"
+pdflatex "$fullFile"
+
+
+
+

### 2020-12-01 manuscripts/owp_fim3_2020/sec/introduction.tex
Added more to intro and put figures in correct place

@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=2.0]{figs/ahps_all_forecast_points.jpg}
+\label{fig:all_ahps_points}
+\caption{All 3,781 forecast points in United States' Advanced Hydrologic Prediction System for lower 48 states and Alaska. No forecast points in Hawaii or territories. From: water.weather.gov/ahps}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=2.0]{figs/ahps_fim_forecast_points_conus.jpg}
+\label{fig:fim_ahps_points}
+\caption{All 188 forecast points in United States' Advanced Hydrologic Prediction System with flood inundation mapping capabilities.
+One forecast point near Juneau, Alaska not shown.
+No forecast points in Hawaii or remaining territories.
+From: water.weather.gov/ahps}
+\end{figure}
+
 Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
@@
 The term hydrofabric is used within the NWM jargon to describe the collection of geospatial datasets required for modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The routing methods used within the NWM for its stream network is Muskingam-Cunge (MC) routing \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
-
-\begin{figure}
-\centering
-\includegraphics[scale=2.0]{figs/ahps_all_forecast_points.jpg}
-\label{fig:all_ahps_points}
-\caption{All 3,781 forecast points in United States' Advanced Hydrologic Prediction System for lower 48 states and Alaska. No forecast points in Hawaii or territories. From: water.weather.gov/ahps}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[scale=2.0]{figs/ahps_fim_forecast_points_conus.jpg}
-\label{fig:fim_ahps_points}
-\caption{All 188 forecast points in United States' Advanced Hydrologic Prediction System with flood inundation mapping capabilities.
-One forecast point near Juneau, Alaska not shown.
-No forecast points in Hawaii or remaining territories.
-From: water.weather.gov/ahps}
-\end{figure}
+The routing methods used within the NWM for its stream network is Muskingam-Cunge (MC) routing to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
+Muskinham-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
+As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
+The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages to inundation extents.

### 2020-12-07 manuscripts/owp_fim3_2020/sec/introduction.tex
Updates * updates to introduction * updated line-endings in fig jpgs * updated gitattributes file to ignore certain extensions * removed abstract and key points for now

@@
 Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
-In 2017 and 2019, river and flash flooding combined represented the leading cause of death among all natural disasters and in 2018 they were only second to deaths from heat wave\cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
+In 2017 and 2019, river and flash flooding combined represented the leading cause of death among all natural disasters and in 2018 they were only second to deaths from heat wave \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
 More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
-Within that 10 year window, the most common activity victims were partaking in is driving year after year when compared to walking/hiking, at home, boating, fell in, working, and other \cite{us_department_of_commerce_2020}.
+Within that 10 year window, the most common activity victims were partaking in was driving when compared to other common activities leading to death \cite{us_department_of_commerce_2020}.
 With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
@@
 Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
-One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with approximately 3,781 forecast points across the US at typically short forecast horizons of 24 or 72 hours\cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble streamflows at 3,781 forecast points illustrated in \ref{fig:all_ahps_points} and flood inundation maps (FIM) at 188 of those forecast points shown in \ref{fig:fim_ahps_points}.
+One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with approximately 3,781 forecast points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
+AHPS provides forecasting services in the form of ensemble streamflows at 3,781 forecast points illustrated in Figure \ref{fig:all_ahps_points} and flood inundation maps (FIM) at 188 of those forecast points shown in Figure \ref{fig:fim_ahps_points}.
 AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multisensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
-Despite the AHPS advances in operational flood forecasting its significant limitations are the lack of spatial coverageand short forecast time horizons.
+Despite the AHPS advances in operational flood forecasting, its limitations are the lack of spatial coverage and short forecast time horizons.
 Approximately, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.
@@
 \includegraphics[scale=2.0]{figs/ahps_all_forecast_points.jpg}
+\caption{All 3,781 forecast points in United States' Advanced Hydrologic Prediction System for lower 48 states and Alaska. No forecast points in Hawaii or territories. From: water.weather.gov/ahps}
 \label{fig:all_ahps_points}
-\caption{All 3,781 forecast points in United States' Advanced Hydrologic Prediction System for lower 48 states and Alaska. No forecast points in Hawaii or territories. From: water.weather.gov/ahps}
 \end{figure}
@@
 \includegraphics[scale=2.0]{figs/ahps_fim_forecast_points_conus.jpg}
-\label{fig:fim_ahps_points}
 \caption{All 188 forecast points in United States' Advanced Hydrologic Prediction System with flood inundation mapping capabilities.
@@
 From: water.weather.gov/ahps}
+\label{fig:fim_ahps_points}
 \end{figure}
@@
 Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
-To broaden the forecasting domain, the Office of Water Prediction (OWP) at the National Water Center (NWC) in Tuscaloosa, Alabama commissioned the utilization of the National Water Model (NWM) which is an instance of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+To broaden the forecasting domain, the Office of Water Prediction (OWP) at the National Water Center (NWC) in Tuscaloosa, Alabama commissioned the development of the National Water Model (NWM) which is an instance of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
 The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
-The NWM enhances the spatial and temporal domain of the current AHPS capabilities at the 13 River Forecast Centers (RFC) in areas known as "hydro-blind" and is not meant to be a replacement. 
+The NWM enhances the spatial and temporal domain of the current AHPS capabilities at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind' and is not meant to be a replacement. 
 It's simply an additional model to be used in the forecasting and early warning decision making.
+Furthermore, the NWM as of V1.2 has implemented not only assimilation of real-time United States Geological Survey (USGS) stage discharges but also assimilation of flow forecasts from the AHPS forecast points which are in turn routed downstream and updated whenever a new point is reached.
+This configuration of the NWM is known as `Replace and Route' or `RnR' and is used to enhance the forecasting skill of the NWM with the best available regional-scale data.
 
 The National Hydrography Dataset Plus (NHDPlus) version 2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}.
-The term hydrofabric is used within the NWM jargon to describe the collection of geospatial datasets required for modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
+The term hydrofabric is used within the NWM jargon to describe the subset of hydrography comprised of geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
 The routing methods used within the NWM for its stream network is Muskingam-Cunge (MC) routing to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
@@
 The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages to inundation extents.
+
+HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and ground water dynamics \cite{renno2008hand,nobre2011height}.
+\citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages. 
+\citeA{zheng2018river} developed methodolgy for determing stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
+This collection of methods, coupling HAND with SRCs, has been experimented and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
+Many of those assesing HAND's efficacy for producing FIM have noted opportunities for improvement. 
+\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
+The colinearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
+Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
+These works motivate splitting junction to junction reaches into smaller and equidistant segments for better averaging behaviour and SRC representation. 
+Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) DEM's derived from light detection and ranging (Lidar).
+Use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
+Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
+Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
+Catchment boundary issues are noted where the interconnection of catchments are not properly accounted for \cite{zhang2018comparative,mcgehee2016modified}.
+The HAND method and its associated advances require advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).
+
+Thanks to the availability of high-performance computing (HPC) and large scale high-resolution digital elevation models (DEM) such as the National Elevation Dataset (NED) at the 10m scale, HAND has been implemented into software for large-scale, continental computation.
+HAND was initially implemented into operational software by the National Flood Interoperability Experiment (NFIE) to generate FIM hydrofabric (will be used interchangebly with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
+NFIE used open source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
+By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
+Originally, NFIE utilized the National Hydrography Dataset (NHD) Plus Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
+The original NFIE dataset was employed by the NWC to produce forecast FIM from the NWM for use within its network of RFC's for additional guidance in hydro-blind regions and tagged as OWP's FIM version 1.0.
+Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND inluding equi-distant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS and tagged as OWP FIM v2.0.
+More notably the software added the ability to derive drainage potentials on a per mainstem basis as originally conceptualized by \cite{mcgehee2016modified}.
+Mainstem is generally used here as defined in \citeA{blodgett2020mainstems} but only available in mainstems downstream of forecast AHPS points within the workflow introduced by \citeA{djokic2019arc}.
+Overall, the software package is estimated to run CONUS at the full-resolution in 0.55 CPU years in a desktop setting.
+
+To continue to innovate and implement the most advanced developments in FIM, the need arose to implement a more advanced software package with higher standards for specifications.
+
+\begin{enumerate}
+\item Use easily accesible, open-source dependencies to promote free-use, open-access, and community development and research.
+\item Excellent single-core computational and memory performance with favorable, time and space complexities.
+\item Ability to parallelize across multiple cores and nodes within HPC or cloud compute environments.
+\item Preprocessing functionality within package pipeline to acquire and process input artificats.
+\item Rapid inundation mapping capabilities to go from FIM hydrofabric datasets to forecast FIMs from NWM.
+\item Automated testing functionality to ensure modifications enhance forecast skill when compared to external sources
+\end{enumerate}

### 2020-12-07 manuscripts/owp_fim3_2020/sec/introduction.tex
More updates to introduction

@@
 Catchment boundary issues are noted where the interconnection of catchments are not properly accounted for \cite{zhang2018comparative,mcgehee2016modified}.
-The HAND method and its associated advances require advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).
+These findings suggesting improvements to HAND require advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).
 
@@
 HAND was initially implemented into operational software by the National Flood Interoperability Experiment (NFIE) to generate FIM hydrofabric (will be used interchangebly with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
-NFIE used open source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
+NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
 By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
 Originally, NFIE utilized the National Hydrography Dataset (NHD) Plus Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC to produce forecast FIM from the NWM for use within its network of RFC's for additional guidance in hydro-blind regions and tagged as OWP's FIM version 1.0.
+The original NFIE dataset was employed by the NWC to produce forecast FIM from the NWM for use within its network of RFC's for additional guidance in hydro-blind regions and tagged as OWP's FIM V1.0.
 Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND inluding equi-distant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS and tagged as OWP FIM v2.0.
@@
 To continue to innovate and implement the most advanced developments in FIM, the need arose to implement a more advanced software package with higher standards for specifications.
+The desired specifications were
 
 \begin{enumerate}
-\item Use easily accesible, open-source dependencies to promote free-use, open-access, and community development and research.
-\item Excellent single-core computational and memory performance with favorable, time and space complexities.
-\item Ability to parallelize across multiple cores and nodes within HPC or cloud compute environments.
-\item Preprocessing functionality within package pipeline to acquire and process input artificats.
-\item Rapid inundation mapping capabilities to go from FIM hydrofabric datasets to forecast FIMs from NWM.
-\item Automated testing functionality to ensure modifications enhance forecast skill when compared to external sources
+\item Use easily accesible, open-source dependencies to promote free-use and open-access community development and research.
+\item excellent single-core computational and memory performance with favorable, time and space complexities,
+\item ability to parallelize across multiple cores and nodes within HPC or cloud compute environments,
+\item preprocessing functionality within package pipeline to acquire and process input artificats,
+\item rapid inundation mapping capabilities to go from FIM hydrofabric datasets to forecast FIMs from NWM,
+\item automated testing functionality to ensure modifications enhance forecast skill when compared to external sources,
+\item ability to produce depth forecasts,
+\item full NWM domain coverage (including Hawaii and Puerto Rico),
+\item and improved FIM forecasting skill when compared to FIM v1.0 and v2.0.
 \end{enumerate}
+
+To meet these higher specifictions for the next generation of authoritative FIM capabilities, the OWP FIM V3.0 `Cahaba' was developed.
+It meets all the specifications listed above including version on version improvement of FIM skill due to the implementation of new developments in HAND techniques.

### 2020-12-08 manuscripts/owp_fim3_2020/sec/introduction.tex
Updates * slight changes to introduction * added docker file to build image with all latex dependencies * added docker run and make file command for systems with Bash * added readme.md file to guide users on how acquire depedencies and build document

@@
 Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
-In 2017 and 2019, river and flash flooding combined represented the leading cause of death among all natural disasters and in 2018 they were only second to deaths from heat wave \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
+In 2017 and 2019, river and flash flooding combined represented the leading cause of death among all natural disasters and second to deaths from heat wave in 2018 \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
 More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
@@
 One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with approximately 3,781 forecast points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble streamflows at 3,781 forecast points illustrated in Figure \ref{fig:all_ahps_points} and flood inundation maps (FIM) at 188 of those forecast points shown in Figure \ref{fig:fim_ahps_points}.
+AHPS provides forecasting services in the form of ensemble stream flows at 3,781 forecast points illustrated in Figure \ref{fig:all_ahps_points} and flood inundation maps (FIM) at 188 of those forecast points shown in Figure \ref{fig:fim_ahps_points}.
 AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multisensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
-Despite the AHPS advances in operational flood forecasting, its limitations are the lack of spatial coverage and short forecast time horizons.
-Approximately, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.
+Despite the AHPS advances in operational flood forecasting, it lacks of sufficient spatial coverage and long-range forecast horizons.
+On an approximate basis, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.
 
@@
 The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
-The NWM enhances the spatial and temporal domain of the current AHPS capabilities at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind' and is not meant to be a replacement. 
+The NWM enhances but does not replace the spatial and temporal domain of the current AHPS capabilities at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'. 
 It's simply an additional model to be used in the forecasting and early warning decision making.
@@
 The term hydrofabric is used within the NWM jargon to describe the subset of hydrography comprised of geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The routing methods used within the NWM for its stream network is Muskingam-Cunge (MC) routing to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
+The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
 Muskinham-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
@@
 This collection of methods, coupling HAND with SRCs, has been experimented and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
-Many of those assesing HAND's efficacy for producing FIM have noted opportunities for improvement. 
+
+Many of those assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
 \citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
@@
 Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
-These works motivate splitting junction to junction reaches into smaller and equidistant segments for better averaging behaviour and SRC representation. 
+These works motivate splitting junction to junction reaches into smaller and equidistant segments for better averaging behavior and SRC representation. 
 Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) DEM's derived from light detection and ranging (Lidar).
@@
 Thanks to the availability of high-performance computing (HPC) and large scale high-resolution digital elevation models (DEM) such as the National Elevation Dataset (NED) at the 10m scale, HAND has been implemented into software for large-scale, continental computation.
-HAND was initially implemented into operational software by the National Flood Interoperability Experiment (NFIE) to generate FIM hydrofabric (will be used interchangebly with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
+HAND was initially implemented into operational software by the National Flood Interoperability Experiment (NFIE) to generate FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
 NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
@@
 \begin{enumerate}
-\item Use easily accesible, open-source dependencies to promote free-use and open-access community development and research.
+\item Use easily accessible, open-source dependencies to promote free-use and open-access community development and research.
 \item excellent single-core computational and memory performance with favorable, time and space complexities,
 \item ability to parallelize across multiple cores and nodes within HPC or cloud compute environments,
-\item preprocessing functionality within package pipeline to acquire and process input artificats,
+\item preprocessing functionality within package pipeline to acquire and process input artifacts,
 \item rapid inundation mapping capabilities to go from FIM hydrofabric datasets to forecast FIMs from NWM,
@@
 
-To meet these higher specifictions for the next generation of authoritative FIM capabilities, the OWP FIM V3.0 `Cahaba' was developed.
+To meet these higher specifications for the next generation of authoritative FIM capabilities, the OWP FIM V3.0 `Cahaba' was developed.
 It meets all the specifications listed above including version on version improvement of FIM skill due to the implementation of new developments in HAND techniques.
+The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM from the NWM for operational forecast applications.
+Furthermore, it enables the contribution from the broader hydro-community as further advances are made in this developing area of research.

### 2020-12-08 manuscripts/owp_fim3_2020/Makefile.sh

@@ -1,5 +1,8 @@
 #!/bin/bash -e
 
+parent_path=$( cd "$(dirname "${BASH_SOURCE[0]}")" ; pwd -P )
+cd $parent_path
+
 fullFile=owp_fim3.tex
 baseName="${fullFile%%.*}"
 extension="${fullFile#*.}"

### 2020-12-08 manuscripts/owp_fim3_2020/docker_run_and_make.sh

@@ -1,0 +1,5 @@
+#!/bin/bash -e
+
+codeDir=$1
+
+docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex-full:dev_20201207 /foss_fim/manuscripts/owp_fim3_2020/Makefile.sh

### 2021-01-07 manuscripts/owp_fim3_2020/dependencies/trackchanges-0.7.0/documentation/web/screenshots/demo.tex
new dir names and updates to methods

--- 2020-09-11 manuscripts/owp_fim3_2020/templates/April-30-2019-latex-templates/April 16 2019/trackchanges-0.7.0/documentation/web/screenshots/demo_simple.tex
@@
 \begin{abstract} \noindent The intention of this text is to demonstrate the use of \texttt{trackchanges.sty}. The package defines five editing commands:
+\begin{itemize}
+  \item \begin{verbatim}\note[initials]{note text}\end{verbatim}
+  \item \begin{verbatim}\annote[initials]{note text}\end{verbatim}
+  \item \begin{verbatim}\add[initials]{additional text}\end{verbatim}
+  \item \begin{verbatim}\remove[initials]{removed text}\end{verbatim}
+  \item \begin{verbatim}\change[initials]{original text}{new text}\end{verbatim}
+\end{itemize}
 \end{abstract}
@@
 
-While planning to write some complex document together with some of my colleagues the same discussion was started over and over again: should \LaTeX\ or some office suite like OpenOffice.org be used?
+While planning to write some \add[FS]{complex} document together with some of my colleagues the same discussion was started over and over again: should \LaTeX\ or some office suite like \change[PS]{Microsoft Office}{OpenOffice.org} be used?
 Strengths of \LaTeX\ are its deterministic behavior, its reliable handling of split documents and unreached typesetting of formulas.
 On the other hand, current office suites provide the user with several features that are at least very desirable when collaboratively writing a document: they provide integrated merging facilities and are able to track changes and attach notes to the text.
-The merging issue can be reasonably handled by version tracking systems like SVN or CVS, but there was no acceptable solution to the issue of change tracking available.
-Of course, some  \LaTeX\ purists tried to convince me that all change tracking can be handled by insertion of \LaTeX\ \textit{comments}.
+The merging issue can be reasonably handled by version tracking systems like SVN or CVS,\note[FS]{Should we emphasize that UNIX diff only works on line basis?} but there was no acceptable solution to the issue of change tracking available.
+Of course, some \remove[PS]{militant} \LaTeX\ purists tried to convince me that all change tracking can be handled by insertion of \LaTeX\ \textit{comments}.
 I have tried to handle one project like this but it did not work out!
-The main reason was that reading and editing of large documents is mostly handled in DVI format and not on the \LaTeX\ source level -- but \LaTeX\ comments cannot be seen in DVI!
+The main reason was that reading and editing of large documents is mostly handled in \annote[PS]{DVI}{what about PDF?
+The changes can be seen in PDF as well ...} format and not on the \LaTeX\ source level -- but \LaTeX\ comments cannot be seen in DVI!
 Especially, if you have sent one version of the document to a colleague and you want to skim quickly over it in order to see what has been changed.

### 2021-01-07 manuscripts/owp_fim3_2020/dependencies/trackchanges-0.7.0/testing/testing_new.tex
new dir names and updates to methods

--- 2020-09-11 manuscripts/owp_fim3_2020/templates/April-30-2019-latex-templates/April 16 2019/trackchanges-0.7.0/testing/testing_old.tex
@@
 
-Michael Norman Randall, you have been found guilty of the murder of Arthur Reginald Webster, Charles Patrick Trumpington, Marcel Agnes Bernstein, Lewis Anona Rudd, John Malcolm Kerr, Nigel Sinclair Robinson, Norman Arthur Potter, Felicity Jayne Stone, Jean-Paul Reynard, Rachel Shirley Donaldson, Stephen Jay Greenblatt, Karl-Heinz Mullet, Belinda Anne Ventham, Juan-Carlos Fernandez, Thor Olaf Stensgaard, Lord Kimberley of Pretoria, Lady Kimberley of Pretoria, The Right Honourable Nigel Warmsly Kimberley, Robert Henry Noonan and Felix James Bennett, on or about the morning of the 19th December 1972.
+Michael Norman Randall, you have been found guilty of the murder of Arthur Reginald Webster, Charles Patrick Trumpington, Marcel Agnes Bernstein, Lewis Anona Rudd, John Malcolm Kerr, Nigel Sinclair Robinson, Harry Potter, Thing 1 \& Thing 2, Humpty Dumpty, Superman, Cold Fusion, The Lorax, Mr.
+T, The Dread Pirate Roberts, Felicity Jayne Stone, Jean-Paul Reynard, Rachel Shirley Donaldson, Stephen Jay Greenblatt, Karl-Heinz Mullet, Belinda Anne Ventham, Juan-Carlos Fernandez, Thor Olaf Stensgaard, Lord Kimberley of Pretoria, The Right Honourable Nigel Warmsly Kimberley, Robert Henry Noonan, Your Mom and Felix James Bennett, on or about the morning of the 19th December 1972.
 Have you anything to say before I pass sentence?
@@
 
-\noindent\\ Counsel:
-
-Well, I would have had to come in anyway.
-
-\noindent\\ Randall:
-
-Ah good, but what a presentation of a case!
-
-\noindent\\ Counsel:
-
-Oh thank you.
-
-\noindent\\ Randall:
-
-No, no, it's a privilege to watch you in action. I never had a chance.
-
-\noindent\\ Counsel:
-
-Oh yes you did.
-
-\noindent\\ Randall:
-
-Not after that summing up. Great.
-
-\noindent\\ Counsel:
-
-Oh thank you. (very chuffed)
-
 \noindent\\ Randall:
@@
 
-No, no, it was very interesting.
+No, no, it was very fascinating.
 

### 2021-01-07 manuscripts/owp_fim3_2020/sections/abstract_plainTextSummary.tex
new dir names and updates to methods

--- 2020-11-30 manuscripts/owp_fim3_2020/sec/abstract_plainTextSummary.tex
@@
 \begin{abstract}
-Historically, floods have led to significant loss of life and property in the United States and around the world. 
-With the expectation that climate change will only increase the intensity of extreme precipitation events, flooding will continue to play a major risk. 
-In response to this growing issue, the National Water Model Version 2.1, released in 2021, produces forecast discharges at over 2.8 million river reaches over the continental United States, Hawaii, and Puerto Rico as well as parts of Canada and Mexico.
-The NWM is a hydrologic and hydraulic model implemented one-dimensionally to reduce computational costs inherent to continental scale modeling. 
-To provide guidance to the general public on flood inundation extents, a rapid methodology is desired to convert forecast discharges to river stages and those stages to inundation extents.
-Relative elevation models (REM), specifically Height Above Nearest Drainage (HAND), among others, detrend digital elevation models (DEM) to the nearest thalweg elevation within the relevant drainage area.
-Generating REM's and reach level catchments enables river stages to translate to extents by assuming a consistent stage profile within each river reach. 
-Stages are translated from NWM discharges via synthetic rating curves that establish stage-discharge relationships via the Manning's equation with geometric properties sampled from the REM on a per catchment basis.
-The National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) is currently on it's third FIM version providing version improvement
+Abstract here
 \end{abstract}
@@
 \section*{Plain Language Summary}
-[ enter your Plain Language Summary here or delete this section]
+PLS here

### 2021-01-07 manuscripts/owp_fim3_2020/sections/introduction.tex
new dir names and updates to methods

--- 2020-12-08 manuscripts/owp_fim3_2020/sec/introduction.tex
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Operational Forecasting}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
@@
 AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multisensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
+On an approximate basis, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.
 Despite the AHPS advances in operational flood forecasting, it lacks of sufficient spatial coverage and long-range forecast horizons.
-On an approximate basis, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.
 
@@
 \centering
-\includegraphics[scale=2.0]{figs/ahps_all_forecast_points.jpg}
+\includegraphics[scale=2.0]{figures/ahps_all_forecast_points.jpg}
 \caption{All 3,781 forecast points in United States' Advanced Hydrologic Prediction System for lower 48 states and Alaska. No forecast points in Hawaii or territories. From: water.weather.gov/ahps}
@@
 \centering
-\includegraphics[scale=2.0]{figs/ahps_fim_forecast_points_conus.jpg}
+\includegraphics[scale=2.0]{figures/ahps_fim_forecast_points_conus.jpg}
 \caption{All 188 forecast points in United States' Advanced Hydrologic Prediction System with flood inundation mapping capabilities.
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{National Water Model}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-The National Hydrography Dataset Plus (NHDPlus) version 2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}.
+The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
 The term hydrofabric is used within the NWM jargon to describe the subset of hydrography comprised of geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
 The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
-Muskinham-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
+Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
 As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Height Above Nearest Drainage}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and ground water dynamics \cite{renno2008hand,nobre2011height}.
 \citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-\citeA{zheng2018river} developed methodolgy for determing stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
+The terrain index as even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
+\citeA{zheng2018river} developed methodology for determine stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
 This collection of methods, coupling HAND with SRCs, has been experimented and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
@@
 These works motivate splitting junction to junction reaches into smaller and equidistant segments for better averaging behavior and SRC representation. 
-Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) DEM's derived from light detection and ranging (Lidar).
+Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
 Use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
@@
 
-Thanks to the availability of high-performance computing (HPC) and large scale high-resolution digital elevation models (DEM) such as the National Elevation Dataset (NED) at the 10m scale, HAND has been implemented into software for large-scale, continental computation.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{HAND Implementations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEM's such as the National Elevation Dataset (NED) at the 10m scale, HAND has been implemented into software for large-scale, continental computation.
 HAND was initially implemented into operational software by the National Flood Interoperability Experiment (NFIE) to generate FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
@@
 The original NFIE dataset was employed by the NWC to produce forecast FIM from the NWM for use within its network of RFC's for additional guidance in hydro-blind regions and tagged as OWP's FIM V1.0.
-Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND inluding equi-distant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS and tagged as OWP FIM v2.0.
-More notably the software added the ability to derive drainage potentials on a per mainstem basis as originally conceptualized by \cite{mcgehee2016modified}.
+Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS and tagged as OWP FIM V2.0.
+More notably the software added the ability to derive drainage potentials on a per mainstem basis as originally conceptualized by \citeA{mcgehee2016modified}.
 Mainstem is generally used here as defined in \citeA{blodgett2020mainstems} but only available in mainstems downstream of forecast AHPS points within the workflow introduced by \citeA{djokic2019arc}.
 Overall, the software package is estimated to run CONUS at the full-resolution in 0.55 CPU years in a desktop setting.
+In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30m resolutions \cite{yamazaki2019merit,donchyts2016global}.
 
-To continue to innovate and implement the most advanced developments in FIM, the need arose to implement a more advanced software package with higher standards for specifications.
-The desired specifications were
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{OWP FIM 3.0 `Cahaba'}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A study is proposed to introduce and evaluate the development of OWP FIM V3.0 which utilizes some of the latest mapping techniques along with a few novel developments in a computationally friendly framework.
+FIM 3 was asked to meet a series of requirements to enhance its utility for continental scale FIM from the National Water Model for operational forecasting and to ensure long-term success in future developments.
+The requirements were
 
 \begin{enumerate}
-\item Use easily accessible, open-source dependencies to promote free-use and open-access community development and research.
+\item use easily accessible, open-source dependencies to promote free-use and open-access community development and research.
 \item excellent single-core computational and memory performance with favorable, time and space complexities,
@@
 \item full NWM domain coverage (including Hawaii and Puerto Rico),
-\item and improved FIM forecasting skill when compared to FIM v1.0 and v2.0.
+\item provide initial mitigation to catchment boundary issues at RnR mainstems,
+\item and improved FIM forecasting skill when compared to FIM V1.0 and V2.0.
 \end{enumerate}

### 2021-01-07 manuscripts/owp_fim3_2020/sections/keypoints.tex
new dir names and updates to methods

--- 2020-11-10 manuscripts/owp_fim3_2020/sec/keypoints.tex
@@
 \begin{keypoints}
-\item The National Water Model produces forecasts at more than 2.8 million river reaches in the United States.
-\item OWP FIM produces flood inundation maps by converting discharges to stages and stages to extents and depths.
-\item FIM 3 is a modified version of the Height Above Nearest Drainage technique that provides version improvement with FIM 1 and 2.
+\item KP1
+\item KP2
+\item KP3
 \end{keypoints}

### 2021-01-07 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
new dir names and updates to methods

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

FIM 3 is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Software Architecture and Dependencies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

FIM 3 exclusively uses free and open source software dependencies (FOSS) and only depends on a host machine having Docker installed on it. 
Within a customized Docker image, a series of dependencies are employed including but not limited to Python 3, GDAL, TauDEM, RichDEM, and Geographic Resource Analysis Support System (GRASS).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Datasets}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

All data sources used within FIM 3 are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for FIM 3 while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
The AHPS forecast points are required to determine the RnR downstream segments in order to define the RnR Mainstems (MS).
For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
Additionally, FEMA Base Level Engineering (BLE) from Region 6 (parts of Texas, Oklahoma, Arkansas, Louisiana and New Mexico) 1\% (100 year) and 0.2\% (500 year) datasets are used as validation in this study. 
These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
The full input datasets sorted by source are listed below:

\begin{enumerate}
\item{USGS - NHDPlusHR Beta}
    \begin{enumerate}
    \item{\textit{BurnLineEvents:} flowlines used by NHDPlusHR for hydro-enforcing DEM}
    \item{\textit{Value-Added Attributes:} database of additional attributes referenced to flowlines that enhance navigation, analysis, and display}
    \item{\textit{DEM:} DEM from National Elevation Dataset (NED) in 10m (or 1/3 arc-second) spatial resolution with vertical units in centimeters \cite{gesch2002national}}
    \end{enumerate}
\item{NOAA OWP - NWM V2.1 Hydrofabric}
    \begin{enumerate}
    \item{\textit{flowlines:} stream network center lines used for NWM routing and forecasting}
    \item{\textit{reach level catchments:} surface drainage area corresponding to each river reach}
    \item{\textit{waterbodies:} includes reservoirs, lakes, and others that are modeled within the NWM}
    \end{enumerate}
\item{NOAA OWP - AHPS Forecast Points}
    \begin{enumerate}
    \item{\textit{AHPS:} Collection AHPS forecast points to associate to RnR Mainstems}
    \end{enumerate}
\item{USACE - NLD}
    \begin{enumerate}
    \item{\textit{Levee elevations:} Top of levee elevations are gathered with in the NLD provided by USACE for hydro-enforcement}
    \end{enumerate}
\item{FEMA BLE}
    \begin{enumerate}
    \item{\textit{cross-sections:} cross-sections and associated data for associating 1\% and 0.2\% recurrence discharges with NWM reaches}
    \item{\textit{flood inundation maps:} FIMs produced by BLE at the two recurrence intervals for comparison with FIMs from OWP FIM versions}
    \end{enumerate}
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hydro-conditioning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The DEM from the NED is subject to a series of hydro-conditioning procedures to enhance it's suitability for riverine flood inundation mapping. 
These techniques are specific for making FIM and different than the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic. 
It can also be used to simulate more accurate bathymetry which is not accounted for in the NED 10m DEM \cite{gesch2002national}.

Specifically within the context of FIM 3, the hydro-conditioning operations that take place in sequential order are presented.
Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing unit designated at run time which can be either HUC 4, 6 or 8. 
The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reachcode's membership in the processing unit.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Headwater Seed Points}

The locations used for seeding headwater locations were enforced to provide identical stream density as the NWM V2.1 Flowlines while utilizing the exact locations of streams found in the NHDPlusHR Beta that have much higher stream densities. 
This was accomplished by finding the NHDPlusHR headwater most adjacent to every NWM headwater point. 
This point is later used for initializing the construction of a FIM 3 specific flowline network that agrees with its resulting flow directions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Stream locations and Bathymetry}

The subset of the NHDPlusHR Beta headwater seed points used in this study are then employed in deriving the remaining stream network used for enforcing the stream network and its associated bathymetry. 
All NHDPlusHR Beta BurnLineEvents downstream of the subset headwater points are selected for hydro-enforcement. 
This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.

\note[Fernando Aristizabal]{For Trevor Grout}
This trimmed stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as AGREE DEM Surface Reconditioning System. 
The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only as can be seen in \ref{fig:agree_dem_cross_section}.
In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
Using simple thalweg burning techniques as opposed to the full AGREE method helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison,}.
\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}. Some of these drawbacks are later addressed by additional conditioning techniques later on.

\begin{figure}[h!]
\centering
\includegraphics[scale=1.0]{figures/agree_dem_cross_section.jpg}
\caption{Placeholder for figure demonstrating AGREE method}
\label{fig:agree_dem_cross_section}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Levee Data}

\note[Fernando Aristizabal]{For Ryan Spies}
Levee data from the NLD was compiled and projected in order to enforce levee elevations within the NED 10m DEM.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Pit Removal}
\cite{barnes2014efficient}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Stream Thalweg Elevation Conditioning}

In order to prevent parallel streams where the steep burned streams and the original streams occur, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel. 
Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50m.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Deriving FIM Hydrofabric}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Flow directions and Flat Resolution}
\cite{survila2016scalable,tarboton1997new,wallis2009parallel}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Stream Network}
\cite{wallis2009parallel}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Reach Catchments}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{HAND}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Synthetic Rating Curves}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{RnR Mainstems Method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation and Testing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-02-02 manuscripts/owp_fim3_2020/sections/introduction.tex
more work on m&m. about to start on deriving fim network

@@
 AHPS provides forecasting services in the form of ensemble stream flows at 3,781 forecast points illustrated in Figure \ref{fig:all_ahps_points} and flood inundation maps (FIM) at 188 of those forecast points shown in Figure \ref{fig:fim_ahps_points}.
+\note{Fernando A}{Combine Figure 1 and 2 with one figure: All AHPS forecast points with no inundation extents one color and with extents other color for entire US if needed}
 AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multisensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
 On an approximate basis, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.
-Despite the AHPS advances in operational flood forecasting, it lacks of sufficient spatial coverage and long-range forecast horizons.
+Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and ground water dynamics \cite{renno2008hand,nobre2011height}.

### 2021-02-02 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
more work on m&m. about to start on deriving fim network

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Software Architecture and Dependencies}
+\subsection{Software Dependencies and Architecture}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-FIM 3 exclusively uses free and open source software dependencies (FOSS) and only depends on a host machine having Docker installed on it. 
-Within a customized Docker image, a series of dependencies are employed including but not limited to Python 3, GDAL, TauDEM, RichDEM, and Geographic Resource Analysis Support System (GRASS).
+OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
+Within the Python 3 ecosystem, many common packages are employed including but not limited RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
+To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in Docker image (\url{https://docs.docker.com/engine/install/}). 
+A user only need to install Docker on their host machine and build the image from the provided recipe.
+
+The FIM 3 pipeline is discretized into key areas that a user can interact with to reproduce the results of this study.
+Preprocessing acquires and prepares datasets for production of the FIM 3 hydrofabric.
+Producing the FIM hydrofabric produces the datasets required to make an inundation map including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
 
@@
     \end{enumerate}
-\item{NOAA OWP - AHPS Forecast Points}
-    \begin{enumerate}
-    \item{\textit{AHPS:} Collection AHPS forecast points to associate to RnR Mainstems}
-    \end{enumerate}
 \item{USACE - NLD}
@@
     \end{enumerate}
+\item{Land Sea Border}
 \item{FEMA BLE}
@@
     \item{\textit{flood inundation maps:} FIMs produced by BLE at the two recurrence intervals for comparison with FIMs from OWP FIM versions}
+    \end{enumerate}
+\item{NOAA OWP - AHPS Forecast Points}
+    \begin{enumerate}
+    \item{\textit{AHPS:} Collection AHPS forecast points to associate to RnR Mainstems}
     \end{enumerate}
@@
 The DEM from the NED is subject to a series of hydro-conditioning procedures to enhance it's suitability for riverine flood inundation mapping. 
-These techniques are specific for making FIM and different than the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
+These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
 Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
@@
 The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reachcode's membership in the processing unit.
+Hydro-conditioning raster operations take place on buffered boundary definitions to avoid edge contamination and effects \cite{lindsay2013measuring}.
 
@@
 In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
-Using simple thalweg burning techniques as opposed to the full AGREE method helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison,}.
+Using simple thalweg burning techniques as opposed to the full AGREE method helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
 \citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
@@
 Some of these drawbacks are later addressed by additional conditioning techniques later on.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/agree_dem_cross_section.jpg}
-\caption{Placeholder for figure demonstrating AGREE method}
-\label{fig:agree_dem_cross_section}
-\end{figure}
 
@@
 \note[Fernando Aristizabal]{For Ryan Spies}
-Levee data from the NLD was compiled and projected in order to enforce levee elevations within the NED 10m DEM.
+The National Elevation Dataset at 10m resolution lacks sufficient representation of fine grain features such as levees and embankments.
+In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10m DEM.
+
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[scale=1.0]{figures/levee_histogram.jpg}
+%\caption{Placeholder for figure demonstrating histogram of elevation differences. May have too many results figures to allow for one}
+%\label{fig:levee_histogram}
+%\end{figure}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Pit Removal}
-\cite{barnes2014efficient}
+\subsubsection{Depression Filling}
+
+Local depressions are a naturally occuring features of a DEM but must be addressed if a connected drainage network with continous catchments are to be derived for flood modeling purposes.
+The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
+Priority-Flood developed by \citeA{barnes2014priority} is an algorithm for filling said depressions and shown to have improved performance over early works in the field by \citeA{jenson1988extracting} implemented in \citeA{tarboton2005terrain} as well as \citeA{planchon2002fast}.
+The depression filling algorithm used in our pipeline is a Priority-Flood variant developed by \cite{zhou2016efficient} with enhanced single-thread performance and a time complexity of O(n log n) for floating point grids.
+This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
+The depression technique employed here does leave the existence of flat regions where pits existed aprior thus later requiring the need for resolving these flats.
+The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
 
@@
 
-In order to prevent parallel streams where the steep burned streams and the original streams occur, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
+Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
+In order to prevent situations where the burned thalweg and the thalweg endemic to DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
 Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50m.
+The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
+
+The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
+\citeA{garousi2019terrain} proposed a breaching algorithm that traverses stream thalweg pixels in a depth first manner  starting with adding all the headwater pixels to a queue. 
+The connectivity of the thalweg pixels is defined by the D8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
+At every thalweg pixel, the minimium elevation among itself and its upstream contributing thalweg pixels is taken as shown in equation \ref{eq:thalweg_breach},
+%
+\begin{equation}
+\label{eq:thalweg_breach}
+\textbf{D}[x] = \min_{y\ drains\ to\ x} {(\ \textbf{D}[x]\ ,\ \textbf{D}[y]\ )}
+\end{equation}
+%
+, in which \textbf{D} represents the array of thalweg adjusted elevations indexed by x and y where by y is upstream of x. 
+When a pixel's upstream neighbors are all evaluated, the downstream pixel is added into the queue thus the depth first traversal of the drainage network.
+This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in critical success index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Flow directions and Flat Resolution}
-\cite{survila2016scalable,tarboton1997new,wallis2009parallel}
+\subsubsection{Flow Directions and Flats Resolution}
+\label{ssec:flow_direction_and_flat_resolution}
+
+To faciliate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D8 flow directions.
+D8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
+The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of square root of 2.
+The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing coutner-clockwise to the Southwest direction as 8. 
+Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
+
+Flat resolution from flats endemic to the DEM or from depression filled regions is a costly,non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
+Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
+The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
+Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
+By processing flats local to each partition separatly from flats shared with other partitions, the accelarated flat resolution algorithm demonstrated an average speed up 468x when compared to prior implementations \cite{survila2016scalable}.
+OWP FIM 3 utilized a CyberGIS implementation of the D8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Network}
+\subsubsection{Deriving FIM Stream Network}
+
+D8 flow accumulation from seed points, thresholding, conversion to vector, splitting to equal lengths at junctions and lakes, globally unique reach id's based on HUC8 id
 \cite{wallis2009parallel}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{HAND}
+\subsubsection{Relative Elevation Model}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Cross-walking NWM Reaches}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{RnR Mainstems Method}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Inundation Mapping}
 

### 2021-02-02 manuscripts/owp_fim3_2020/docker_run_and_make.sh

@@ -2,4 +2,4 @@
 
 codeDir=$1
 
-docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex-full:dev_20201207 /foss_fim/manuscripts/owp_fim3_2020/Makefile.sh
+docker run --rm -v "$1":/foss_fim fim3_latex:20210131 /foss_fim/manuscripts/owp_fim3_2020/Makefile.sh

### 2021-02-03 manuscripts/owp_fim3_2020/sections/abstract_plainTextSummary.tex
Pruning figure, more m&m, wrote abstract, wrote key points, wrote plain-text summary, intialized discussion file

@@
 \begin{abstract}
-Abstract here
+The National Water Model Version 2.1 (NWM V2.1) seeks to expand upon existing systems to provide hourly forecast discharges at four time horizons (1, 3, 10, and 30 days) at over 2.7 million locations throughout the continental United States, Hawaii and Puerto Rico.
+Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D inundation extents used for forecasting purposes to protect life and property.
+The Office of Water Prediction Flood Inundation Mapping Version 3.0 `Cahaba' (OWP-FIM V3.0) was developed to provide these inundation mapping services based on the latest advancements in methodology and data sources available.
+Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, support for a two-tiered drainage density, parametrization of processing unit size, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and optimal Manning's n roughness parameter calibrated by stream order.
+OWP FIM V3 was calibrated and evaluated at 48 Base Level Engineering (BLE) sites furnished by the Federal Emergency Management Agency (FEMA).
+The modeled discharges from the BLE sections were used for the 100 and 500 year flood events as input to OWP-FIM V3 instaed of NWM discharges to isolate out factors related to hydro-meterology.
+The resulting maps are compared to the modeled extents from the BLE and binary contigency statistics are computed.
+Across the 48 sites, Critical Success Index (CSI) and Matthew's Correlation Coefficient (MCC) were found to increase across FIM versions driven by improvements to the probability of detection (POD) via reduction in errors of omission with only a slight increase in errors of commission.
+These capabalities were implemented in an open-source package available on GitHub (\url{https://github.com/NOAA-OWP/cahaba}) for continued use and collaboration with the research community.
+%
 \end{abstract}
@@
 \section*{Plain Language Summary}
-PLS here
+Flooding is one of the most impactful natural disasters on life and property.
+So having high quality flood forecasts enables people to protect lives and property.
+The United States National Water Model seeks to provide flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
+The National Water Model produces flow rates at river segments so additional work is needed to convert these flow rates to inundation extents so that flood maps can be developed.
+The Office of Water Prediction Flood Inundation Mapping Version 3 provides enhancements to existing capabilities to provide these flood maps as a service.
+The maps were compared to other, more rigorous models and shown to perform well when compared to previous versions.
+%

### 2021-02-03 manuscripts/owp_fim3_2020/sections/discussion.tex
Pruning figure, more m&m, wrote abstract, wrote key points, wrote plain-text summary, intialized discussion file

--- 2021-01-07 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-FIM 3 is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Software Architecture and Dependencies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-FIM 3 exclusively uses free and open source software dependencies (FOSS) and only depends on a host machine having Docker installed on it. 
-Within a customized Docker image, a series of dependencies are employed including but not limited to Python 3, GDAL, TauDEM, RichDEM, and Geographic Resource Analysis Support System (GRASS).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-All data sources used within FIM 3 are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for FIM 3 while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
-The AHPS forecast points are required to determine the RnR downstream segments in order to define the RnR Mainstems (MS).
-For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
-Additionally, FEMA Base Level Engineering (BLE) from Region 6 (parts of Texas, Oklahoma, Arkansas, Louisiana and New Mexico) 1\% (100 year) and 0.2\% (500 year) datasets are used as validation in this study.
-These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
-The full input datasets sorted by source are listed below:
-
-\begin{enumerate}
-\item{USGS - NHDPlusHR Beta}
-    \begin{enumerate}
-    \item{\textit{BurnLineEvents:} flowlines used by NHDPlusHR for hydro-enforcing DEM}
-    \item{\textit{Value-Added Attributes:} database of additional attributes referenced to flowlines that enhance navigation, analysis, and display}
-    \item{\textit{DEM:} DEM from National Elevation Dataset (NED) in 10m (or 1/3 arc-second) spatial resolution with vertical units in centimeters \cite{gesch2002national}}
-    \end{enumerate}
-\item{NOAA OWP - NWM V2.1 Hydrofabric}
-    \begin{enumerate}
-    \item{\textit{flowlines:} stream network center lines used for NWM routing and forecasting}
-    \item{\textit{reach level catchments:} surface drainage area corresponding to each river reach}
-    \item{\textit{waterbodies:} includes reservoirs, lakes, and others that are modeled within the NWM}
-    \end{enumerate}
-\item{NOAA OWP - AHPS Forecast Points}
-    \begin{enumerate}
-    \item{\textit{AHPS:} Collection AHPS forecast points to associate to RnR Mainstems}
-    \end{enumerate}
-\item{USACE - NLD}
-    \begin{enumerate}
-    \item{\textit{Levee elevations:} Top of levee elevations are gathered with in the NLD provided by USACE for hydro-enforcement}
-    \end{enumerate}
-\item{FEMA BLE}
-    \begin{enumerate}
-    \item{\textit{cross-sections:} cross-sections and associated data for associating 1\% and 0.2\% recurrence discharges with NWM reaches}
-    \item{\textit{flood inundation maps:} FIMs produced by BLE at the two recurrence intervals for comparison with FIMs from OWP FIM versions}
-    \end{enumerate}
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hydro-conditioning}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The DEM from the NED is subject to a series of hydro-conditioning procedures to enhance it's suitability for riverine flood inundation mapping. 
-These techniques are specific for making FIM and different than the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
-Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
-It can also be used to simulate more accurate bathymetry which is not accounted for in the NED 10m DEM \cite{gesch2002national}.
-
-Specifically within the context of FIM 3, the hydro-conditioning operations that take place in sequential order are presented.
-Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing unit designated at run time which can be either HUC 4, 6 or 8. 
-The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reachcode's membership in the processing unit.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Headwater Seed Points}
-
-The locations used for seeding headwater locations were enforced to provide identical stream density as the NWM V2.1 Flowlines while utilizing the exact locations of streams found in the NHDPlusHR Beta that have much higher stream densities.
-This was accomplished by finding the NHDPlusHR headwater most adjacent to every NWM headwater point. 
-This point is later used for initializing the construction of a FIM 3 specific flowline network that agrees with its resulting flow directions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream locations and Bathymetry}
-
-The subset of the NHDPlusHR Beta headwater seed points used in this study are then employed in deriving the remaining stream network used for enforcing the stream network and its associated bathymetry.
-All NHDPlusHR Beta BurnLineEvents downstream of the subset headwater points are selected for hydro-enforcement. 
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.
-
-\note[Fernando Aristizabal]{For Trevor Grout}
-This trimmed stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as AGREE DEM Surface Reconditioning System. 
-The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only as can be seen in \ref{fig:agree_dem_cross_section}.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
-Using simple thalweg burning techniques as opposed to the full AGREE method helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison,}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
-Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
-Some of these drawbacks are later addressed by additional conditioning techniques later on.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/agree_dem_cross_section.jpg}
-\caption{Placeholder for figure demonstrating AGREE method}
-\label{fig:agree_dem_cross_section}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Levee Data}
-
-\note[Fernando Aristizabal]{For Ryan Spies}
-Levee data from the NLD was compiled and projected in order to enforce levee elevations within the NED 10m DEM.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Pit Removal}
-\cite{barnes2014efficient}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Thalweg Elevation Conditioning}
-
-In order to prevent parallel streams where the steep burned streams and the original streams occur, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
-Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50m.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Deriving FIM Hydrofabric}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Flow directions and Flat Resolution}
-\cite{survila2016scalable,tarboton1997new,wallis2009parallel}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Network}
-\cite{wallis2009parallel}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Reach Catchments}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{HAND}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Synthetic Rating Curves}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{RnR Mainstems Method}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation and Testing}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-02-03 manuscripts/owp_fim3_2020/sections/introduction.tex
Pruning figure, more m&m, wrote abstract, wrote key points, wrote plain-text summary, intialized discussion file

--- 2021-01-07 manuscripts/owp_fim3_2020/sections/introduction.tex
@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
@@
 Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
@@
 AHPS provides forecasting services in the form of ensemble stream flows at 3,781 forecast points illustrated in Figure \ref{fig:all_ahps_points} and flood inundation maps (FIM) at 188 of those forecast points shown in Figure \ref{fig:fim_ahps_points}.
+\note{Fernando A}{Combine Figure 1 and 2 with one figure: All AHPS forecast points with no inundation extents one color and with extents other color for entire US if needed}
 AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multisensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
 On an approximate basis, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.
-Despite the AHPS advances in operational flood forecasting, it lacks of sufficient spatial coverage and long-range forecast horizons.
-
+Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons.
+%
 \begin{figure}[h!]
@@
 \end{figure}
-
+%
 \begin{figure}[h!]
@@
 \end{figure}
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
@@
 The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages to inundation extents.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and ground water dynamics \cite{renno2008hand,nobre2011height}.
@@
 These findings suggesting improvements to HAND require advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEM's such as the National Elevation Dataset (NED) at the 10m scale, HAND has been implemented into software for large-scale, continental computation.
@@
 In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30m resolutions \cite{yamazaki2019merit,donchyts2016global}.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 A study is proposed to introduce and evaluate the development of OWP FIM V3.0 which utilizes some of the latest mapping techniques along with a few novel developments in a computationally friendly framework.
@@
 The requirements were
-
+%
 \begin{enumerate}
@@
 \end{enumerate} 
-
+%
 To meet these higher specifications for the next generation of authoritative FIM capabilities, the OWP FIM V3.0 `Cahaba' was developed. 
@@
 Furthermore, it enables the contribution from the broader hydro-community as further advances are made in this developing area of research.
+%

### 2021-02-03 manuscripts/owp_fim3_2020/sections/keypoints.tex
Pruning figure, more m&m, wrote abstract, wrote key points, wrote plain-text summary, intialized discussion file

@@
 \begin{keypoints}
-\item KP1
-\item KP2
-\item KP3
+\item Office of Water Prediction Flood Inundation Mapping Version 3 `Cahaba' was developed to apply the latest mapping methodology and datasets to produce inundation maps from the United States National Water Model Version 2.1 at continental scale.
+\item OWP-FIM V3 `Cahaba' applies advanced techniques and datasets such as the latest hydrography, enhanced computational performance, channel and levee enforcement, two-tiered drainage density, and Manning's n parameter calibration by stream order.
+\item Cahaba was compared to Base Flood Elevation modeled maps and shown to provide version improvement when compared to previous OWP-FIM versions.
 \end{keypoints}

### 2021-02-03 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
Pruning figure, more m&m, wrote abstract, wrote key points, wrote plain-text summary, intialized discussion file

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 FIM 3 is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
@@
 Producing the FIM hydrofabric produces the datasets required to make an inundation map including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 All data sources used within FIM 3 are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
@@
 The full input datasets sorted by source are listed below:
-
+%
 \begin{enumerate}
@@
 \end{enumerate}
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 The DEM from the NED is subject to a series of hydro-conditioning procedures to enhance it's suitability for riverine flood inundation mapping. 
@@
 Hydro-conditioning raster operations take place on buffered boundary definitions to avoid edge contamination and effects \cite{lindsay2013measuring}. 
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Headwater Seed Points}
-
-The locations used for seeding headwater locations were enforced to provide identical stream density as the NWM V2.1 Flowlines while utilizing the exact locations of streams found in the NHDPlusHR Beta that have much higher stream densities.
-This was accomplished by finding the NHDPlusHR headwater most adjacent to every NWM headwater point. 
-This point is later used for initializing the construction of a FIM 3 specific flowline network that agrees with its resulting flow directions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream locations and Bathymetry}
-
-The subset of the NHDPlusHR Beta headwater seed points used in this study are then employed in deriving the remaining stream network used for enforcing the stream network and its associated bathymetry.
-All NHDPlusHR Beta BurnLineEvents downstream of the subset headwater points are selected for hydro-enforcement. 
+\subsubsection{Stream Network Enforcement} 
+\label{ssec:stream_network_enforcment}
+%
+\note[Fernando Aristizabal]{For Brian Avant}
+The location of the stream network are enforced to ensure general agreement with established stream networks.
+The NHDPlus HR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
+However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
+Pruning the NHD network is completed at for every NWM V2.1 headwater point, the nearest NHDPlusHR Burline Event feature is selected and downstream features are also selected.
+The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
+This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
+Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees withe DEM after all hydro-conditioning takes place.
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/headwaters.jpg}
+\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection.
+Resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR Burnline Events.}
+\label{fig:stream_density_pruning}
+\end{figure}
+%
 This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.
@@
 \note[Fernando Aristizabal]{For Trevor Grout}
-This trimmed stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as AGREE DEM Surface Reconditioning System. 
+The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
 The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
 The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only as can be seen in \ref{fig:agree_dem_cross_section}.
+The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only as can be seen in Figure \ref{fig:agree_dem_cross_section}.
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/agree_dem_cross_section.jpg}
+\caption{Placeholder for AGREE DEM illustration}
+\label{fig:agree_dem_cross_section}
+\end{figure}
+%
 In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
@@
 Some of these drawbacks are later addressed by additional conditioning techniques later on.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Levee Data}
-
+\subsubsection{Levee Enforcement}
+%
 \note[Fernando Aristizabal]{For Ryan Spies}
@@
 In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10m DEM.
-
+%
 %\begin{figure}[h!]
@@
 %\end{figure}
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Depression Filling}
-
-Local depressions are a naturally occuring features of a DEM but must be addressed if a connected drainage network with continous catchments are to be derived for flood modeling purposes.
+%
+Local depressions are a naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
 The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
@@
 The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Stream Thalweg Elevation Conditioning}
-
+%
 Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
@@
 The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
-\citeA{garousi2019terrain} proposed a breaching algorithm that traverses stream thalweg pixels in a depth first manner  starting with adding all the headwater pixels to a queue. 
-The connectivity of the thalweg pixels is defined by the D8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
-At every thalweg pixel, the minimium elevation among itself and its upstream contributing thalweg pixels is taken as shown in equation \ref{eq:thalweg_breach},
+\citeA{garousi2019terrain} proposed an algorithm that traverses stream thalweg pixels in a depth first manner starting with adding all the headwater pixels to a queue. 
+The connectivity of the thalweg pixels is defined by the D-8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
+At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in equation \ref{eq:thalweg_breach},
 %
@@
 This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in critical success index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{ssec:flow_direction_and_flat_resolution}
-
-To faciliate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D8 flow directions.
-D8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
+%
+To facilitate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D-8 flow directions.
+D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
 The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of square root of 2.
-The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing coutner-clockwise to the Southwest direction as 8. 
+The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southwest direction as 8. 
 Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
@@
 Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
-By processing flats local to each partition separatly from flats shared with other partitions, the accelarated flat resolution algorithm demonstrated an average speed up 468x when compared to prior implementations \cite{survila2016scalable}.
-OWP FIM 3 utilized a CyberGIS implementation of the D8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
-
+By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up 468x when compared to prior implementations \cite{survila2016scalable}.
+OWP FIM 3 utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Deriving FIM Stream Network}
+%
+\note[Fernando Aristizabal]{For Brian Avant}
+The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
+The FIM stream network is of the same density as the NWM V2.1 network and fully converges at all junctions leaving no divergences in the network.
+This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{sec:stream_network_enforcment}).
+These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to theheadwater definitions in the stream network of the NWM V2.1.
 
-D8 flow accumulation from seed points, thresholding, conversion to vector, splitting to equal lengths at junctions and lakes, globally unique reach id's based on HUC8 id
+Additionally, each stream link in the FIM stream network is split into equidistant reaches
+Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments or NWM lakes.
+D-8 flow accumulation from seed points, thresholding, conversion to vector, splitting to equal lengths at junctions and lakes, globally unique reach id's based on HUC 8 id, cross-walking
 \cite{wallis2009parallel}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Catchments}
+%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Reach Catchments}
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Relative Elevation Model}
+%
 
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Synthetic Rating Curves}
+%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cross-walking NWM Reaches}
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{RnR Mainstems Method}
+%
 
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Inundation Mapping}
+%
 
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+
+%

### 2021-02-03 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
More m&m

@@
 The FIM stream network is of the same density as the NWM V2.1 network and fully converges at all junctions leaving no divergences in the network.
-This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{sec:stream_network_enforcment}).
+This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
 These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to theheadwater definitions in the stream network of the NWM V2.1.
+Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
+Each stream link in this derived FIM stream network is split into equidistant reaches.
+Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments or NWM lakes.
+Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
+Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
 
-Additionally, each stream link in the FIM stream network is split into equidistant reaches
-Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments or NWM lakes.
-D-8 flow accumulation from seed points, thresholding, conversion to vector, splitting to equal lengths at junctions and lakes, globally unique reach id's based on HUC 8 id, cross-walking
-\cite{wallis2009parallel}
+Now that the stream network has been re-derived using the hydro-conditioning techniques previously discussed, the split stream reaches need to cross-walked to the forecast stream network which is that of the National Water Model's.
 %

### 2021-02-11 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
Updated inline equation

@@
 D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of square root of 2.
+The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
 The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southwest direction as 8. 
@@
 Each stream link in this derived FIM stream network is split into equidistant reaches.
-Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments or NWM lakes.
+Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments.
+Stream links are then further segmented at NWM lakes and HUC8 boundaries.
 Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.

### 2021-04-13 manuscripts/owp_fim3_2020/sections/introduction.tex
added to methods

@@
 These works motivate splitting junction to junction reaches into smaller and equidistant segments for better averaging behavior and SRC representation. 
-Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
+Furthermore, \citeAgarousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
 Use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
@@
 Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
-Catchment boundary issues are noted where the interconnection of catchments are not properly accounted for \cite{zhang2018comparative,mcgehee2016modified}.
+Catchment boundary issues are noted where the interconnection of catchments are not properly accounted for \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
 These findings suggesting improvements to HAND require advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).

### 2021-04-13 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
added to methods

@@
 \subsubsection{Stream Thalweg Elevation Conditioning}
+\label{sssec:stream_thalweg_elevation_conditioning}
 %
@@
 %
+The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsubsection{Deriving FIM Stream Network}
+\label{sssec:deriving_fim_stream_network}
 %
@@
 Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
+Based on previous research, splitting each remaining stream link into equi-distant reaches not to exceed a parameterized value of 1.5km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
+Small reaches can lead to unrealistic variances in channel geometries while too large reaches can lead to grouping too much geometry variance into one parametrized unit.
 Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
-
-Now that the stream network has been re-derived using the hydro-conditioning techniques previously discussed, the split stream reaches need to cross-walked to the forecast stream network which is that of the National Water Model's.
+This stream network is important since it drives the HAND calculation and derivation of catchments.
 %
@@
 %
-
+Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
+Gages set at the pixel center points of the delineated stream lines explained in section \ref{sssec:deriving_fim_stream_network}.
+The gages act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
+Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
+The other catchments is derived for the contributing area for each stream reach as defined in section \ref{sssec:deriving_fim_stream_network}. 
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Relative Elevation Model}
+\subsubsection{Height Above Nearest Drainage}
 %
-
+Once the pixel level catchments are derived the final relative elevations can be computed.
+Every non-thalweg elevation is substracted from the thalweg elevation within the same pixel-level catchment.
+The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in section \ref{sssec:stream_thalweg_elevation_conditioning}.
+Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling. 
 %

### 2021-04-15 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
more m and m

@@
 %
+ A method for converting forecast river discharges from the NWM to stages or river depths is necessary for producing FIMs with HAND. 
+For one-dimensional models such as the NWM, the typical procedure for establishing this conversion is to determine the stage-discharge relationship sampling data from the DEM known as a synthetic rating curve with cross-sections \cite{quintero2021development,di2011hydraulic}.
+For this application, we utilized the reach averaged approach for developing synthetic rating curves (SRC) \cite{zheng2018river}.
+The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
+The reach average Manning's formula is derived to be
+
+\begin{equation}
+\label{eq:reach_averaged_mannings_equation}
+\textbf{Q}[y] = \frac{1}{n} \frac{V[y]^{5/3}S^{1/2}}{L B[y]^{2/3}} 
+\end{equation}
+
+where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
+All units are international given the 1 numerator above n.
+The reach averaged method has been evaluated to ...
+The reach averaged parameters are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
+
+The Manning's n values are tuned ...
 

### 2021-04-15 gms_run.sh
still working through batch

@@ -167,7 +167,7 @@ ogrmerge.py -f GPKG -single -overwrite_ds -o $outputRunDataDir/aggregate_fim_out
 
 ## DERIVE LEVELPATH  ##
 echo -e $startDiv"Generating Level Paths"$stopDiv
-$srcDir/gms/derive_level_paths.py -i $outputRunDataDir/aggregate_fim_outputs/demDerived_merged.gpkg -b $branch_id_attribute -o $outputGmsDataDir/demDerived_reaches_levelPaths.gpkg -d $outputGmsDataDir/demDerived_reaches_levelPaths_dissolved.gpkg -v
+$srcDir/gms/derive_level_paths.py -i $outputRunDataDir/aggregate_fim_outputs/demDerived_merged.gpkg -b $branch_id_attribute -o $outputRunDataDir/aggregate_fim_outputs/demDerived_reaches_levelPaths.gpkg -d $outputRunDataDir/aggregate_fim_outputs/demDerived_reaches_levelPaths_dissolved.gpkg -v
 
 ## RUN GMS ##
 if [ -f "$hucList" ]; then

### 2021-04-15 src/gms/run_by_branch.sh

@@ -7,15 +7,6 @@ T_total_start
 ## SOURCE BASH FUNCTIONS
 source $srcDir/bash_functions.env
 
-# make outputs directory
-if [ ! -d "$outputGmsDataDir" ]; then
-    mkdir -p $outputGmsDataDir
-else # remove contents if already exists
-    rm -rf $outputGmsDataDir
-    mkdir -p $outputGmsDataDir
-fi
-
-## TEMP ##
 ## SET VARIABLES AND FILE INPUTS ##
 source /foss_fim/config/params_calibrated.env
 hucNumber="$1"
@@ -33,6 +24,15 @@ input_src_full=$outputHucDataDir/src_full_crosswalked.csv
 ## SET OUTPUT DIRECTORY FOR UNIT ##
 outputHucDataDir=$outputRunDataDir/$hucNumber
 outputGmsDataDir=$outputHucDataDir/gms
+
+# make outputs directory
+if [ ! -d "$outputGmsDataDir" ]; then
+    mkdir -p $outputGmsDataDir
+else # remove contents if already exists
+    rm -rf $outputGmsDataDir
+    mkdir -p $outputGmsDataDir
+fi
+
 
 ## ECHO PARAMETERS
 echo -e $startDiv"Parameter Values"
@@ -55,18 +55,21 @@ echo -e "memfree=$memfree"
 echo -e "branch_id_attribute=$branch_id_attribute"
 echo -e "branch_buffer_distance_meters=$branch_buffer_distance_meters"$stopDiv
 
+
+##### SUBSET Levelpaths to HUC
+
 ## STREAM BRANCH POLYGONS
 echo -e $startDiv"Generating Stream Branch Polygons for $hucNumber"$stopDiv
 date -u
 Tstart
-$srcDir/gms/buffer_stream_branches.py -s $outputGmsDataDir/demDerived_reaches_levelPaths_dissolved.gpkg -i $branch_id_attribute -d $branch_buffer_distance_meters -b $outputGmsDataDir/polygons.gpkg -v 
+$srcDir/gms/buffer_stream_branches.py -s $outputRunDataDir/aggregate_fim_outputs/demDerived_reaches_levelPaths_dissolved.gpkg -i $branch_id_attribute -d $branch_buffer_distance_meters -b $outputRunDataDir/aggregate_fim_outputs/polygons.gpkg -v 
 Tcount
 
 ## CLIP RASTERS
[…]

### 2021-04-15 src/run_by_unit.sh

@@ -436,7 +436,7 @@ Tcount
 echo -e $startDiv"USGS Crosswalk $hucNumber"$stopDiv
 date -u
 Tstart
-$srcDir/usgs_gage_crosswalk.py -gages $inputDataDir/usgs_gages/usgs_gages.gpkg -dem $outputHucDataDir/dem_meters.tif -flows $outputHucDataDir/demDerived_reaches_split_filtered_addedAttributes_crosswalked.gpkg -cat $outputHucDataDir/gw_catchments_reaches_filtered_addedAttributes_crosswalked.gpkg -wbd $outputHucDataDir/wbd_buffered.gpkg -dem_adj $dem_thalwegCond -outtable $outputHucDataDir/usgs_elev_table.csv -e $extent
+#$srcDir/usgs_gage_crosswalk.py -gages $inputDataDir/usgs_gages/usgs_gages.gpkg -dem $outputHucDataDir/dem_meters.tif -flows $outputHucDataDir/demDerived_reaches_split_filtered_addedAttributes_crosswalked.gpkg -cat $outputHucDataDir/gw_catchments_reaches_filtered_addedAttributes_crosswalked.gpkg -wbd $outputHucDataDir/wbd_buffered.gpkg -dem_adj $dem_thalwegCond -outtable $outputHucDataDir/usgs_elev_table.csv -e $extent
 Tcount
 
 ## CLEANUP OUTPUTS ##

### 2021-05-16 manuscripts/owp_fim3_2020/sections/abstract_plainTextSummary.tex
more updates

@@
 The National Water Model Version 2.1 (NWM V2.1) seeks to expand upon existing systems to provide hourly forecast discharges at four time horizons (1, 3, 10, and 30 days) at over 2.7 million locations throughout the continental United States, Hawaii and Puerto Rico.
-Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D inundation extents used for forecasting purposes to protect life and property.
-The Office of Water Prediction Flood Inundation Mapping Version 3.0 `Cahaba' (OWP-FIM V3.0) was developed to provide these inundation mapping services based on the latest advancements in methodology and data sources available.
+Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D fluvial inundation extents used for forecasting purposes to protect life and property.
+The Office of Water Prediction Flood Inundation Mapping Version 3.0 `Cahaba' (OWP-FIM V3.0) is an enhancemant of the Height Above Nearest Drainage (HAND) method developed to improve upon some of HAND's limitations.
 Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, support for a two-tiered drainage density, parametrization of processing unit size, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and optimal Manning's n roughness parameter calibrated by stream order.

### 2021-05-16 manuscripts/owp_fim3_2020/sections/introduction.tex
more updates

--- 2021-02-03 manuscripts/owp_fim3_2020/sections/introduction.tex
@@
 This collection of methods, coupling HAND with SRCs, has been experimented and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
-
-Many of those assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
-\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
-The colinearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
-Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
-These works motivate splitting junction to junction reaches into smaller and equidistant segments for better averaging behavior and SRC representation. 
-Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
-Use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
-Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
-Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
-Catchment boundary issues are noted where the interconnection of catchments are not properly accounted for \cite{zhang2018comparative,mcgehee2016modified}.
-These findings suggesting improvements to HAND require advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).
 %
@@
 %
-A study is proposed to introduce and evaluate the development of OWP FIM V3.0 which utilizes some of the latest mapping techniques along with a few novel developments in a computationally friendly framework.
-FIM 3 was asked to meet a series of requirements to enhance its utility for continental scale FIM from the National Water Model for operational forecasting and to ensure long-term success in future developments.
-The requirements were
+Many of those assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
+\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
+The colinearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
+Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
+These works motivate splitting junction to junction reaches into smaller and equidistant segments for better averaging behavior and SRC representation. 
+Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
+Use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
+Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
+Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
+Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloaquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
+These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).
+
+A study is proposed to introduce and evaluate the development of OWP FIM V3.0 `Cahaba' which utilizes some of the latest mapping techniques along with a few novel developments in a computationally friendly framework.
+The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM from the NWM for operational forecast applications. 
+Furthermore, it enables the contribution from the broader hydro-community through its use of open-source code as further advances are made in this developing area of research.
 %
-\begin{enumerate}
-\item use easily accessible, open-source dependencies to promote free-use and open-access community development and research.
-\item excellent single-core computational and memory performance with favorable, time and space complexities,
-\item ability to parallelize across multiple cores and nodes within HPC or cloud compute environments,
-\item preprocessing functionality within package pipeline to acquire and process input artifacts,
-\item rapid inundation mapping capabilities to go from FIM hydrofabric datasets to forecast FIMs from NWM,
-\item automated testing functionality to ensure modifications enhance forecast skill when compared to external sources,
-\item ability to produce depth forecasts,
-\item full NWM domain coverage (including Hawaii and Puerto Rico),
-\item provide initial mitigation to catchment boundary issues at RnR mainstems,
-\item and improved FIM forecasting skill when compared to FIM V1.0 and V2.0.
-\end{enumerate} 
-%
-To meet these higher specifications for the next generation of authoritative FIM capabilities, the OWP FIM V3.0 `Cahaba' was developed. 
-It meets all the specifications listed above including version on version improvement of FIM skill due to the implementation of new developments in HAND techniques. 
-The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM from the NWM for operational forecast applications. 
-Furthermore, it enables the contribution from the broader hydro-community as further advances are made in this developing area of research.
-%

### 2021-05-16 manuscripts/owp_fim3_2020/sections/keypoints.tex
more updates

--- 2021-01-07 manuscripts/owp_fim3_2020/sections/keypoints.tex
@@
 \begin{keypoints}
-\item KP1
-\item KP2
-\item KP3
+\item Height Above Nearest Drainage (HAND) is a terrain index that detrends elevation data to the nearest relevant drainage path and can be used for rapid generation of flood inundation maps (FIM) across large spatial scales and high spatial resolutions specifically with applications with the United States National Water Model.
+\item Despite its advantages, FIM generated from HAND suffers from a number of limitations most notably its dependence on drainage density of the stream network employed.
+\item The Office of Water Prediction at the National Oceanic and Atmospheric Administration presents a means of resolving stream density by reducing drainage density to a singular flow path for HAND based FIM.
+More recent developments in FIM based HAND are also employed in this continental scale model.
 \end{keypoints}

### 2021-05-16 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
more updates

@@
 OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
-Within the Python 3 ecosystem, many common packages are employed including but not limited RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
+Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
 To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in Docker image (\url{https://docs.docker.com/engine/install/}). 
@@
 \subsection{Datasets}
+\label{ssec:datasets}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-All data sources used within FIM 3 are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
+Most data sources used within FIM 3 are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
 It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for FIM 3 while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
-The AHPS forecast points are required to determine the RnR downstream segments in order to define the RnR Mainstems (MS).
 For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
-Additionally, FEMA Base Level Engineering (BLE) from Region 6 (parts of Texas, Oklahoma, Arkansas, Louisiana and New Mexico) 1\% (100 year) and 0.2\% (500 year) datasets are used as validation in this study.
+Additionally, FEMA Base Level Engineering (BLE) from Region 6 (parts of Texas, Oklahoma, Arkansas, Louisiana and New Mexico) 1\% (100 year) and 0.2\% (500 year) datasets are used as validation in this study \cite{fema2021base,fema2021estimated}.
 These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
-The full input datasets sorted by source are listed below:
+The full input datasets presented by source are list in Table \ref{tab:data}.
+\note[Fernando Aristizabal]{Check Brian Avant, Trevor Grout, Bradford Bates, Ryan Spies}
 %
-\begin{enumerate}
-\item{USGS - NHDPlusHR Beta}
-    \begin{enumerate}
-    \item{\textit{BurnLineEvents:} flowlines used by NHDPlusHR for hydro-enforcing DEM}
-    \item{\textit{Value-Added Attributes:} database of additional attributes referenced to flowlines that enhance navigation, analysis, and display}
-    \item{\textit{DEM:} DEM from National Elevation Dataset (NED) in 10m (or 1/3 arc-second) spatial resolution with vertical units in centimeters \cite{gesch2002national}}
-    \end{enumerate}
-\item{NOAA OWP - NWM V2.1 Hydrofabric}
-    \begin{enumerate}
-    \item{\textit{flowlines:} stream network center lines used for NWM routing and forecasting}
-    \item{\textit{reach level catchments:} surface drainage area corresponding to each river reach}
-    \item{\textit{waterbodies:} includes reservoirs, lakes, and others that are modeled within the NWM}
-    \end{enumerate}
-\item{USACE - NLD}
-    \begin{enumerate}
-    \item{\textit{Levee elevations:} Top of levee elevations are gathered with in the NLD provided by USACE for hydro-enforcement}
-    \end{enumerate}
-\item{Land Sea Border}
-\item{FEMA BLE}
-    \begin{enumerate}
-    \item{\textit{cross-sections:} cross-sections and associated data for associating 1\% and 0.2\% recurrence discharges with NWM reaches}
-    \item{\textit{flood inundation maps:} FIMs produced by BLE at the two recurrence intervals for comparison with FIMs from OWP FIM versions}
-    \end{enumerate}
-\item{NOAA OWP - AHPS Forecast Points}
-    \begin{enumerate}
-    \item{\textit{AHPS:} Collection AHPS forecast points to associate to RnR Mainstems}
-    \end{enumerate}
-\end{enumerate}
+
+\begin{table}
+\caption{Data sources, names, and descriptions.}
+\label{tab:data}
+\centering
+\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
+%\begin{tabular}{l c c}
+\hline
+Source & Name & Description \\
+\hline
+USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
+\hline
+USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
+\hline
+USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10m) spatial resolution and vertical units in centimeters \\
+\hline
+NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting \\
+\hline
+NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM \\
+\hline
+NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as resevoirs or lakes \\
+\hline
+USACE & NLD & Levee database of locations and elevations  \\
+\hline
+WHERE? & Land-Sea Border & Border of land and sea and land and Great Lakes \\
+\hline
+FEMA & Cross-Sections & HEC-RAS cross-sections used for modeling in BLE datasets. Include discharges for 1\% and 0.2\% recurrence interval events \\
+\hline
+FEMA & Flood Inundation Extents & Extents produced by FEMA BLE for 1\% and 0.2\% recurrence interval events \\
+\hline
+\multicolumn{2}{l}{$^{a}$Footnote text here.}
+\end{tabular}
+\end{table}
+
+Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain. 
+\note[Fernando Aristizabal]{For Nick Chadwich} This includes XXXX HUC4's, XXX HUC'6s, and XX HUC8's.
+An enhancement of OWP FIM v2 `Cahaba' over previous FIM versions is the support for Hawaii and Puerto Rico which the NWM v2.1 will cover.
 %
@@
 Based on previous research, splitting each remaining stream link into equi-distant reaches not to exceed a parameterized value of 1.5km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while too large reaches can lead to grouping too much geometry variance into one parametrized unit.
+Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much geometry variance into one parametrized unit.
+Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
 Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
@@
 \subsubsection{Synthetic Rating Curves}
+\label{sssec:synthetic_rating_curve}
 %
@@
 All units are international given the 1 numerator above n.
-The reach averaged method has been evaluated to ...
-The reach averaged parameters are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-
-The Manning's n values are tuned ...
-
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{RnR Mainstems Method}
-%
-
+The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
+The reach averaged geometry parameters including number of cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
+Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to the channel length.
+Calibration of the Manning's n roughness coefficient is conducted by optimizing inundation extent performance to a benchmark dataset best described in Section \ref{ssec:evaluation_and_calibration}. 
 %
@@
 \subsection{Inundation Mapping}
+\label{ssec:inundation_mapping}
 %
-
+The FIM hydrofabric consisting of the relative elevations grid, catchments grid, catchment polygons, rating curve, and cross-walking data are all used to convert forecasts from the NWM into forecasts extents.
+For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
+From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 4, 6, or 8 spatial processing units in a rapid, parallel operation.
+The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
+Once associated to a mapping stream network, 
+the catchments grid encoded with the reach identifiers are used to map
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation and Testing}
+\subsection{Evaluation and Calibration}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\label{ssec:evaluation_and_calibration}
 %
+Evaluation of the OWP-FIM V3 Cahaba is conducted by comparison to the HEC-RAS 1D derived models produced by FEMA region 6 \cite{fema2021base,fema2021estimated}.
+The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by FEMA so we used those corresponding discharges and mapping extents for evaluation.
+By not using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors imputed by hydrology, forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
+We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates with each NWM reach. 
+To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
+This partially handles the influence of neighboring cross sections that could cause flow discontinuties and mass conservation issues.
+While the limitions are noted, this method does best detangle the influence of exogeneous variables that we do not wish to study in this comparison.
 
+Calibrating the Manning's n values is very important for optimizing the skill of inundation extent forecasts \cite{zheng2018river,garousi2019terrain}.
+\note[Fernando Aristizabal]{For Brian Avant} In this study we propose a simplified parameterization of roughness by stream order and tune inundation performance relative to the validation data set 
 %

### 2021-05-16 manuscripts/owp_fim3_2020/sections/results.tex
more updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computational Performance}
\label{ssec:compuational_performance}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
All computational experiments were conducted on an Amazon Web Services (AWS) Elastic Cloud Compute (EC2) T3 instance (2x-large size) with 4 cores, 2 threads per core, and 32GB of random access memory (RAM).
An Elastic Block Storage (EBS) solid state drive (SSD) of the general purpose 2 (gp2) type was used for 2TB of storage.
All tests took place within the project's Docker image providing the operating system and software dependencies.
Table \ref{tab:comp_results} details the computational metrics including CPU time, RAM usage, and data storage footprint for both the hydrofabric production and NRT generation of FIMs.
For the hydrofabric production, computational results were produced for the entire NWM FIM domain noted in \ref{ssec:datasets}.
For the NRT production of FIM from the hydrofabric, computational results were aggregated for the 48 HUC study area.
Computation of the depths and inundation rasters as well as the inundation in polygon form are included in this anaylsis.
Only a subset of NWM FIM domain ever gets mapped for operational purposes to avoid computational burdens by applying a 1.5\% NWM recurrence flow threshold.

\begin{table}
\caption{Descriptive computational statistics for `Cahaba' hydrofabric and NRT FIM production.}
\label{tab:comp_results}
\centering
\begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
%\hline
%"" & \multicolumn{7}{l}{CPU Time (Hrs)} & \multicolumn{7}{l}{RAM (GB)} \\
\hline
Variable & Statistic & Hydrofabric & FIM \\
\hline
\multirow{7}{*}{CPU Time (hrs)} &
n$^{a}$ & XX & XX \\ 
\cline{2-4}
& Sum  & XX & XX \\ 
\cline{2-4}
& Mean & XX & XX \\ 
\cline{2-4}
& Median & XX & XX \\ 
\cline{2-4}
& SD$^{b}$ & XX & XX \\ 
\cline{2-4}
& Max & XX & XX \\ 
\cline{2-4}
& Min & XX & XX \\ 
\hline
\multirow{7}{*}{RAM (GB)} &
n$^{a}$ & XX & XX \\
\cline{2-4}
& Sum  & XX & XX \\
\cline{2-4}
& Mean & XX & XX \\
\cline{2-4}
& Median & XX & XX \\
\cline{2-4}
& SD$^{b}$ & XX & XX \\
\cline{2-4}
& Max & XX & XX \\
\cline{2-4}
& Min & XX & XX \\
\hline
\multirow{7}{*}{Data Storage (GB)} &
n$^{a}$ & XX & XX \\
\cline{2-4}
& Sum  & XX & XX \\
\cline{2-4}
& Mean & XX & XX \\
\cline{2-4}
& Median & XX & XX \\
\cline{2-4}
& Std.Dev.$^{b}$ & XX & XX \\
\cline{2-4}
& Max & XX & XX \\
\cline{2-4}
& Min & XX & XX \\
\hline
\multicolumn{2}{l}{$^{a}$Sample size. $^{b}$Standard Deviation}
\end{tabular}
\end{table}

For comparison purposes, OWP FIM v1 measured that the entire CONUS area excepting a few HUC6's took 1.34 CPU years \cite{liu2016cybergis}. 
OWP FIM v2, developed by \citeA{djokic2019arc}, was estimated to take 0.55 CPU years for the entire NWM stream resolution but lacked the ability to run in parallel on HPC environments.
By utilizing solid programming practices and optimal algorithm selection, we were able to reduce computational requirements while enhancing hydrologic performance.

% TEST SCRIPT: 
% for huc in $(cat /data/temp/ble_test_cases_fa_20210418.lst); do echo HUC: $huc; /usr/bin/time -v /foss_fim/tools/inundation.py -r /data/outputs/first_gms_batch_test/"$huc"/rem_zeroed_masked.tif -c /data/outputs/first_gms_batch_test/"$huc"/gw_catchments_reaches_filtered_addedAttributes.tif -b /data/outputs/first_gms_batch_test/"$huc"/gw_catchments_reaches_filtered_addedAttributes_crosswalked.gpkg -t /data/outputs/first_gms_batch_test/"$huc"/hydroTable.csv -f /data/test_cases/ble_test_cases/validation_data_ble/"$huc"/100yr/ble_huc_"$huc"_flows_100yr.csv -p /data/temp/inundation_py_benchmarking_for_manuscript/inundation_"$huc".gpkg -m filter -i /data/temp/inundation_py_benchmarking_for_manuscript/inundation_"$huc".tif -d /data/temp/inundation_py_benchmarking_for_manuscript/depths_"$huc".tif -u /data/inputs/wbd/WBD_National.gpkg -l WBDHU8 -q | tee /data/temp/inundation_py_benchmarking_for_manuscript/test_1.log; done

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mapping Performance}
\label{ssec:mapping_performance}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-05-16 manuscripts/owp_fim3_2020/sections/abstract_plainTextSummary.tex
slight changes

@@
 Flooding is one of the most impactful natural disasters on life and property.
-So having high quality flood forecasts enables people to protect lives and property.
-The United States National Water Model seeks to provide flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
-The National Water Model produces flow rates at river segments so additional work is needed to convert these flow rates to inundation extents so that flood maps can be developed.
-The Office of Water Prediction Flood Inundation Mapping Version 3 provides enhancements to existing capabilities to provide these flood maps as a service.
+The United States National Water Model (NWM) seeks to provide flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
+In order to convert forecasts flow rates from the NWM to inundation extents, a model is used that converts elevations relative to sea-level to elevations relative to the nearest river bottom elevation.
+This model known as Height Above Nearest Drainage (HAND) suffers fro
 The maps were compared to other, more rigorous models and shown to perform well when compared to previous versions.

### 2021-05-16 manuscripts/owp_fim3_2020/sections/introduction.tex
Updates * removed fluff from intro * edited authors * more M&M

@@
 Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
-In 2017 and 2019, river and flash flooding combined represented the leading cause of death among all natural disasters and second to deaths from heat wave in 2018 \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
+In 2017 and 2019, river and flash flooding combined represented the leading cause of death among all natural disasters and second in 2018 \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
 More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
-Within that 10 year window, the most common activity victims were partaking in was driving when compared to other common activities leading to death \cite{us_department_of_commerce_2020}.
-With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
-The large spike in damages for 2017 can be attributed to the Hurricane Harvey event that primarily affected Texas in August. 
+With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
 Unencouragingly, the trends related to flood damages and fatalities have been steadily increasing over recent decades.
@@
 Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
-These works motivate splitting junction to junction reaches into smaller and equidistant segments for better averaging behavior and SRC representation. 
 Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
@@
 A study is proposed to introduce and evaluate the development of OWP FIM V3.0 `Cahaba' which utilizes some of the latest mapping techniques along with a few novel developments in a computationally friendly framework.
-The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM from the NWM for operational forecast applications. 
+Most notably this study demonstrates how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
+The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM for the NWM for operational forecast applications. 
 Furthermore, it enables the contribution from the broader hydro-community through its use of open-source code as further advances are made in this developing area of research.

### 2021-05-16 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
Updates * removed fluff from intro * edited authors * more M&M

@@
 The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only as can be seen in Figure \ref{fig:agree_dem_cross_section}.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/agree_dem_cross_section.jpg}
-\caption{Placeholder for AGREE DEM illustration}
-\label{fig:agree_dem_cross_section}
-\end{figure}
-%
+The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only.
 In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
@@
 %
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[scale=1.0]{figures/levee_histogram.jpg}
-%\caption{Placeholder for figure demonstrating histogram of elevation differences. May have too many results figures to allow for one}
-%\label{fig:levee_histogram}
-%\end{figure}
-%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Depression Filling}
+\label{sssec:depression_filling}
 %
-Local depressions are a naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
+Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
 The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Unary Stream Order}
+\label{ssec:unary_stream_order}
+%
+FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
+Figure \ref{fig:confluence_problem_statement} illustrates this exact situation where two tributatries converge with a higher order stream segment. 
+A `Cahaba' FIM is generated and compared with a FEMA 100 year extent (please see sections \ref{ssec:evaluation_and_calibration} for more details) showing significant under-prediction in inundation extent
+The higher discharge along the mainstem of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
+This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
+As a parallel problem, there is excess water accumulated along the mainstem that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the mainstem catchments.
+
+\begin{figure}[h!]
+\centering
+%\includegraphics[scale=1.0]{figures/headwaters.jpg}
+\caption{}
+\label{fig:confluence_problem_statement}
+\end{figure}
+%
+The limitation imposed by using stream networks of Horton-Strahler stream order \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} greater than one for HAND can be described mathematically with Graph Theory.
+Assume a drainage area denoted by special case of a fully connected, directed acyclic graph known as a tree, G, with all nodes (or reaches) directed from the leaf nodes (or headwater reaches) toward a singular root node $r_0$ known as the outlet in the application's field.
+Assume a singluar subgraph of G where by a new root node,$r_n$, is introduced that represents another stream location.
+n is used here to index possible root nodes which are a subset of G.
+Ensuring $r_n$ is itself an inner node (has upstream drainage) disconnects G into $G_0$ and $G_n$ thus removing influence of conditions at $r_0$ upon $G_n$.
+HAND seeks to divide drainage areas into reach scale catchments that are independent of one another.
+We intend to demonstrate how introduction of stream orders greater than one can generate errors at confluences.
+We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{NWS Replace and Route Mainstems}
+\label{ssec:unary_stream_order}
+%
+
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Generalized Mainstems}
+\label{ssec:unary_stream_order}
+%
+
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Inundation Mapping}

### 2021-05-16 manuscripts/owp_fim3_2020/sections/title_authors.tex
Updates * removed fluff from intro * edited authors * more M&M

--- 2020-11-10 manuscripts/owp_fim3_2020/sec/title_authors.tex
@@
 
-\title{High Resolution, Rapid Flood Inundation Mapping for the U.S. National Water Model: From Forecast River Discharges to Forecast Inundation}
+\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
 
@@
 
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3},Trevor Grout\affil{1,3}, Nick Chadwick\affil{1,3}, Zachary Wills\affil{1,3}, Gregory Petrochenkov\affil{4}, Roland Viger\affil{4}, Jasmeet Judge\affil{2}}
+\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3},Trevor Grout\affil{1,3}, Gregory Petrochenkov\affil{4}, Nick Chadwick\affil{1,3}, Zachary Wills\affil{1,3}, Fred Ogden \affil{3}, Roland Viger\affil{4}, Jasmeet Judge\affil{2}}
 
-\affiliation{1}{Lynker Technologies, Boulder, CO, USA}
+\affiliation{1}{Lynker Technologies, Leesburg, VA, USA}
 \affiliation{2}{Center for Remote Sensing, University of Florida, Gainesville, FL, USA}

### 2021-05-16 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
added m&m

@@
 \subsection{Hydro-conditioning}
+\label{ssec:hydro_conditioning}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Deriving FIM Hydrofabric}
+\label{ssec:deriving_fim_hydrofabric}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-
+The RnR configuration of the NWM seeks to replace NWM forecasts by assimilating best available AHPS forecasts and routing them downstream.
+The RnR Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points.
+The RnR MS network, or simply MS, comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
+HAND was originally derived for this stream network to enhance mapping skill along these critical MS \cite{djokic2019arc}. 
+The inundation derived from this stream network is mosaiced or composited with the inundation from the FR network to form composite (C) FIMs. 
+Within each HAND processing unit (HUC 4, 6, 8), you'll typically only find a MS stream network of stream order 1.
 %
@@
 %
+To further the efforts implmented by MS, we sought to derive HAND at a level-path scale.
+Level-path groups flow paths by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain. 
+Starting at the outlet, a unique level-path is propogated upstream. 
+At every confluence, the direction of maximum flow path length is sought to propogate the current level-path identifier.
+For the remaining parent reaches of the given junction, a new level-path identifier is assigned and the process recursively continues with them.
+Each HUC 4,6,or 8 is discretized into level-paths indepedently and relevant inputs are assigned to each level-path processing unit given a buffer of 7 km.
+At the level-path scale, the methods in Section \ref{ssec:deriving_fim_hydrofabric} are executed while the methods in Section \ref{ssec:hydro_conditioning} are still completed at the HUC scale.
 
@@
 This partially handles the influence of neighboring cross sections that could cause flow discontinuties and mass conservation issues.
+Additionally, the stream network of the FEMA furnished models are of higher stream densities and bifurcation ratios leading to a signicant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
 While the limitions are noted, this method does best detangle the influence of exogeneous variables that we do not wish to study in this comparison.
 
+The metrics employed in this study to evaluate inundation extents include critical success index (CSI), probability of detection (POD), and false alarm ratio (FAR) and presented in equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
+To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.
+The two types of errors consist of false positives (FP), or type I errors, which is dry in benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in benchmark by predicted dry.
+Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
+Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
+The NWS denotes CSI a good proxy for measuring a forecasting systems utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$ \cite{gerapetritis2004behavior}.
+While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
+Thus frequency dependent statistics should be used with caution. 
+%
+\begin{equation}
+\label{eq:csi}
+CSI = \frac{TP}{TP + FN + FP}
+\end{equation}
+%
+\begin{equation}
+\label{eq:pod}
+POD = \frac{TP}{TP + FN}
+\end{equation}
+%
+\begin{equation}
+\label{eq:far}
+FAR = \frac{FP}{TP + FP}
+\end{equation}
+%
 Calibrating the Manning's n values is very important for optimizing the skill of inundation extent forecasts \cite{zheng2018river,garousi2019terrain}.
-\note[Fernando Aristizabal]{For Brian Avant} In this study we propose a simplified parameterization of roughness by stream order and tune inundation performance relative to the validation data set 
+\note[Fernando Aristizabal]{For Brian Avant} In this study we propose a simplified parameterization of roughness by stream order and tune inundation performance relative to the validation data set.
+For every combination of stream order (from 1 to 14) and Manning's n value (from X.XX to X.XX), synthetic rating curves are generated and compared to the FEMA inundation extents.
+The optimal C
 %

### 2021-05-18 manuscripts/owp_fim3_2020/sections/introduction.tex
Updates to figures

@@
 Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
-In 2017 and 2019, river and flash flooding combined represented the leading cause of death among all natural disasters and second in 2018 \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
+In 2017 and 2019, river and flash flooding combined represented the leading cause of death and the second leading cause in 2018 among all natural disasters in the US \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
 More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
@@
 Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
-One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with approximately 3,781 forecast points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble stream flows at 3,781 forecast points illustrated in Figure \ref{fig:all_ahps_points} and flood inundation maps (FIM) at 188 of those forecast points shown in Figure \ref{fig:fim_ahps_points}.
-\note{Fernando A}{Combine Figure 1 and 2 with one figure: All AHPS forecast points with no inundation extents one color and with extents other color for entire US if needed}
+One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with more than 3,000 forecast points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
+AHPS provides forecasting services in the form of ensemble stream flows at more than 3,000 forecast points and flood inundation maps (FIM) at more than 150 of those locations shown in Figure \ref{fig:forecast_points}.
+Besides the AHPS forecast points, we display in Figure \ref{fig:forecast_points} two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model}.
 AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multisensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
-On an approximate basis, there is only one forecast point every 1,455 km of river and one forecast point with FIM every 29,261 km.
+On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
 Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons.
@@
 \centering
-\includegraphics[scale=2.0]{figures/ahps_all_forecast_points.jpg}
-\caption{All 3,781 forecast points in United States' Advanced Hydrologic Prediction System for lower 48 states and Alaska. No forecast points in Hawaii or territories. From: water.weather.gov/ahps}
-\label{fig:all_ahps_points}
-\end{figure}
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=2.0]{figures/ahps_fim_forecast_points_conus.jpg}
-\caption{All 188 forecast points in United States' Advanced Hydrologic Prediction System with flood inundation mapping capabilities.
-One forecast point near Juneau, Alaska not shown.
-No forecast points in Hawaii or remaining territories.
-From: water.weather.gov/ahps}
-\label{fig:fim_ahps_points}
+\includegraphics[scale=1.0]{figures/forecast_points.jpg}
+\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
+Also show are the National Water Model stream networks at the full resolution (FR) and Replace and Reroute (RnR) Mainstems (MS) resolution.}
+\label{fig:forecast_points}
 \end{figure}
@@
 \subsection{National Water Model}
+\label{ssec:national_water_model}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 This configuration of the NWM is known as `Replace and Route' or `RnR' and is used to enhance the forecasting skill of the NWM with the best available regional-scale data.
+Figure \ref{ssec:national_water_model} shows Full-Resolution (FR) modeling stream network as well as the RnR mainstems (MS) network.
 
@@
 The terrain index as even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
-\citeA{zheng2018river} developed methodology for determine stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
+\citeA{zheng2018river} developed methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
 This collection of methods, coupling HAND with SRCs, has been experimented and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
@@
 Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS and tagged as OWP FIM V2.0.
-More notably the software added the ability to derive drainage potentials on a per mainstem basis as originally conceptualized by \citeA{mcgehee2016modified}.
-Mainstem is generally used here as defined in \citeA{blodgett2020mainstems} but only available in mainstems downstream of forecast AHPS points within the workflow introduced by \citeA{djokic2019arc}.
+More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
 Overall, the software package is estimated to run CONUS at the full-resolution in 0.55 CPU years in a desktop setting.
@@
 This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
-These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).
+These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.
 

### 2021-05-18 manuscripts/owp_fim3_2020/sections/keypoints.tex
Updates to figures

--- 2021-01-07 manuscripts/owp_fim3_2020/sections/keypoints.tex
@@
 \begin{keypoints}
-\item KP1
-\item KP2
-\item KP3
+\item Height Above Nearest Drainage (HAND) is a terrain index used for rapid generation of flood inundation maps (FIM) across large scales.
+\item Despite its advantages, FIM generated from HAND suffers from a number of limitations most notably its dependence on drainage density.
+\item Among other developments, here we present a means of resolving stream density by reducing drainage density to a singular flow path. 
 \end{keypoints}

### 2021-05-18 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
Updates to figures

@@
 \hline
-\multicolumn{2}{l}{$^{a}$Footnote text here.}
 \end{tabular}
@@
 \note[Fernando Aristizabal]{For Nick Chadwich} This includes XXXX HUC4's, XXX HUC'6s, and XX HUC8's.
-An enhancement of OWP FIM v2 `Cahaba' over previous FIM versions is the support for Hawaii and Puerto Rico which the NWM v2.1 will cover.
+An enhancement of OWP FIM v3 `Cahaba' over previous FIM versions is the support for Hawaii and Puerto Rico which the NWM v2.1 will cover.
 %
@@
 However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-Pruning the NHD network is completed at for every NWM V2.1 headwater point, the nearest NHDPlusHR Burline Event feature is selected and downstream features are also selected.
+Pruning the NHD network is completed at for every NWM V2.1 headwater point where the nearest NHDPlusHR Burline Event feature is selected and downstream features are also selected.
 The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
 This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees withe DEM after all hydro-conditioning takes place.
+Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
 %
@@
 Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
-By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up 468x when compared to prior implementations \cite{survila2016scalable}.
+By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
 OWP FIM 3 utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
@@
 This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to theheadwater definitions in the stream network of the NWM V2.1.
+These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
 Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
@@
 In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
-Figure \ref{fig:confluence_problem_statement} illustrates this exact situation where two tributatries converge with a higher order stream segment. 
+Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributatries converge with a higher order stream segment. 
 A `Cahaba' FIM is generated and compared with a FEMA 100 year extent (please see sections \ref{ssec:evaluation_and_calibration} for more details) showing significant under-prediction in inundation extent
@@
 \centering
-%\includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{}
-\label{fig:confluence_problem_statement}
+\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
+\caption{False negatives associated with confluence of tributaries with mainstems.
+Integers represent flow values from BLE 100 year for the associated areas.
+No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
+\label{fig:catchment_boundaries_issue}
 \end{figure}
@@
 Assume a drainage area denoted by special case of a fully connected, directed acyclic graph known as a tree, G, with all nodes (or reaches) directed from the leaf nodes (or headwater reaches) toward a singular root node $r_0$ known as the outlet in the application's field.
-Assume a singluar subgraph of G where by a new root node,$r_n$, is introduced that represents another stream location.
+Assume a singluar subgraph of G where by a new root node, $r_n$, is introduced that represents another stream location.
 n is used here to index possible root nodes which are a subset of G.
@@
 \subsubsection{NWS Replace and Route Mainstems}
-\label{ssec:unary_stream_order}
+\label{sssec:replace_and_route_mainstems}
 %
@@
 \subsubsection{Generalized Mainstems}
-\label{ssec:unary_stream_order}
+\label{sssec:generalized_mainstems}
 %
-To further the efforts implmented by MS, we sought to derive HAND at a level-path scale.
+To further the efforts implmented by MS, we sought to derive HAND at a level-path scale which we call Generalized Mainstems (GMS).
 Level-path groups flow paths by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain. 
@@
 At the level-path scale, the methods in Section \ref{ssec:deriving_fim_hydrofabric} are executed while the methods in Section \ref{ssec:hydro_conditioning} are still completed at the HUC scale.
-
 %
@@
 The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
-Once associated to a mapping stream network, 
-the catchments grid encoded with the reach identifiers are used to map
+Once associated to a mapping stream network, the catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
+Mathematically, the catchment pixels, $C_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
+For each forcast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in equation \ref{eq:hand_fim}.
+%
+\begin{equation}
+\label{eq:hand_fim}
+    F_{ij} = 
+    \begin{cases}
+        1 & if S_i > C_{ij} \\
+        0 & else
+    \end{cases}
+\end{equation}
+%
+For the cases of MS and GMS, the inundations must be mosaiced together to form a seamless forecast.
+For MS, the inundation from MS stream network is unioned with the FR inundation extents while all the individual level-path inundation extents are unioned to form the composite inundation, C, for GMS.
+Equation \ref{eq:comp_fim} illustrates how the composite inundation, C, can be determined from n sources,F, all indexed by i.
+%
+\begin{equation}
+\label{eq:comp_fim}
+    C = \cup_{i=1}^{n}F_i
+\end{equation}
+In the case of MS, the n value is 2 for the FR and MS inundation sources while for GMS n is equivalent to the number of unique level-paths in the domain of concern.
 %
@@
 \subsection{Evaluation and Calibration}
+\label{ssec:evaluation_and_calibration}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\label{ssec:evaluation_and_calibration}
 %
@@
 For every combination of stream order (from 1 to 14) and Manning's n value (from X.XX to X.XX), synthetic rating curves are generated and compared to the FEMA inundation extents.
-The optimal C
+The optimal roughness for each stream order is determined by maximizing CSI.
+A clear limitation of this approach is over-fitting to the BLE domain at 1\% and 0.2\% recurrence discharges.
+On-going work is focused on applying more advanced techniques through use of USGS stream gage rating curves.
 %

### 2021-05-18 manuscripts/owp_fim3_2020/sections/title_authors.tex
Updates to figures

@@
 
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3},Trevor Grout\affil{1,3}, Gregory Petrochenkov\affil{4}, Nick Chadwick\affil{1,3}, Zachary Wills\affil{1,3}, Fred Ogden \affil{3}, Roland Viger\affil{4}, Jasmeet Judge\affil{2}}
+\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Trevor Grout\affil{1,3}, Gregory Petrochenkov\affil{4}, Nick Chadwick\affil{1,3}, Zachary Wills\affil{3,5}, Fred Ogden \affil{3}, Jasmeet Judge\affil{2}}
 
@@
 \affiliation{3}{National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
-\affiliation{4}{Water Mission Area, United States Geological Survey, Boulder, CO, USA}
+\affiliation{4}{New York Water Science Center, United States Geological Survey, Troy, NY, USA}
+\affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Alabama, Tuscaloosa, AL, USA}
 

### 2021-05-18 api/frontend/output_handler/output_handler.py
Squashed commit of the following:
commit cbc360f1f27ea15e87510024ae340b5f53fe1638
Author: NickChadwick-NOAA <nick.chadwick@noaa.gov>
Date:   Tue May 11 10:42:27 2021 -0500

    Updated Eventlet package to 0.31.0 to patch a vulnerability

    The current version of Eventlet used in the Connector module of the FIM API is outdated and vulnerable. This update bumps the version to the patched version.

    Using new nhd inputs.

commit b5da4b4869ccff2d87b2326c29d378990daf1c3e
Author: NickChadwick-NOAA <nick.chadwick@noaa.gov>
Date:   Fri May 7 14:45:39 2021 -0500

    Added fuctionality for new "Re

@@ -5,83 +5,35 @@ import socketio
 
 SOCKET_URL = os.environ.get('SOCKET_URL')
 
-pending_files = {}
-
 def handle_outputs(data):
-    name = f"{data['job_name']}_{data['file_name']}"
-    if name not in pending_files:
-        pending_files[name] = {
-            'locked': False,
-            'current_index': 0,
-            'nice_name': data['nice_name'],
-            'job_name': data['job_name'],
-            'directory_path': data['directory_path'],
-            'file_name': data['file_name']
-        }
+    job_name = data['job_name']
+    directory_path = data['directory_path']
+    file_name = data['file_name']
+    file_chunk = data['file_chunk']
+    chunk_index = data['chunk_index']
 
-    pending_files[name][data['chunk_index']] = data['file_chunk']
+    # Create folder if it doesn't yet exist and set writing mode
+    mode = 'ab'
+    if chunk_index == 0:
+        mode = 'wb'
+        try:
+            os.makedirs(directory_path)
+        except:
+            pass
+        
+    # Write binary data to file
+    with open(f"{directory_path}/{file_name}", mode) as binary_file:
+        print(f"Writing chunk {chunk_index} for file {directory_path}/{file_name}")
+        binary_file.write(file_chunk)
 
+    sio.emit('output_handler_finished_file_chunk', {'job_name': job_name, 'file_path': f"{directory_path}/{file_name}"})
 
-    work_to_do = True
-    while work_to_do:
-        work_to_do = False
-
-        nice_name = pending_files[name]['nice_name']
-        job_name = pending_files[name]['job_name']
-        directory_path = pending_files[name]['directory_path']
-        file_name = pending_files[name]['file_name']
-
-        # If the last chunk just got added, waiting to write any potentially missing data to file
-        if data['file_chunk'] == None and pending_files[name]['locked']:
-            while name in pending_files and pending_files[name]['locked']:
-                print("EOF, waiting till not locked")
-                sio.sleep(0.5)
-            if not name in pending_files:
-                return
-
-        # To ensure that the files are being written in the correct order, use current_index
-        # to write the correct file chunk.
-        if not pending_files[name]['locked'] and pending_files[name]['current_index'] in pending_files[name]:
-            pending_files[name]['locked'] = True
-            file_chunk = pending_files[name].pop(pending_files[name]['current_index'])
-
-            # End of file
-            if file_chunk == None:
-                if sio.connected:
[…]

### 2021-05-18 api/node/connector/connector.py

@@ -3,6 +3,7 @@ eventlet.monkey_patch()
 
 import os
 import re
+import time
 import random
 import logging
 import subprocess
@@ -40,8 +41,8 @@ def ws_disconn():
     emit('is_connected', False)
 
 @socketio.on('update')
-def ws_update(current_jobs):
-    emit('client_update', current_jobs, broadcast=True)
+def ws_update(data):
+    emit('client_update', data, broadcast=True)
 
 @socketio.on('output_handler_connected')
 def ws_output_handler_connected():
@@ -57,11 +58,9 @@ def ws_updater_connected():
 
 @socketio.on('ready_for_output_handler')
 def ws_ready_for_output_handler(data):
-    nice_name = data['nice_name']
     job_name = data['job_name']
     path = data['path']
-
-    print(f"handler_sid: {shared_data['handler_sid']}")
+    chunk_index = data['chunk_index']
 
     if shared_data['handler_sid'] == None:
         print("output handler not connected!")
@@ -69,101 +68,160 @@ def ws_ready_for_output_handler(data):
         return
 
     # Split up path into parts for the output handler
-    path_parts = re.search(rf"/data/outputs/{job_name}/(.+)/(.+)", path)
+    path_parts = re.search(rf"(.+)/(.+)", path)
     directory_path = path_parts.group(1)
     file_name = path_parts.group(2)
 
+    file_read_start = time.time()
     with open(path, "rb") as binary_file:
-        print("Sending to output handler", path)
-        
         # Read and emit file chunk by chunk (50MB at a time)
-        chunk_index = 0
+        binary_file.seek(chunk_index * 52428800)
         file_chunk = binary_file.read(52428800)
-        # file_chunk = binary_file.read(104857600)
-        while file_chunk:
-            print("Sending to output handler", path, "Chunk:", chunk_index)
-            emit('new_job_outputs', {
-                'nice_name': nice_name,
+
+        if len(file_chunk) == 0:
+            print('End of File')
+            emit('file_saved', {
                 'job_name': job_name,
-                'directory_path': directory_path,
-                'file_name': file_name,
-                'file_chunk': file_chunk,
-                'chunk_index': chunk_index
-            }, room=shared_data['handler_sid'])
+                'file_path': path
+            }, room=shared_data['updater_sid'])
+            return
 
-            chunk_index += 1
-            file_chunk = binary_file.read(52428800)
[…]

### 2021-05-18 fim_run.sh

@@ -94,7 +94,7 @@ fi
 
 ## SOURCE ENV FILE AND FUNCTIONS ##
 source $envFile
-source $libDir/bash_functions.env
+source $srcDir/bash_functions.env
 
 # default values
 if [ "$jobLimit" = "" ] ; then
@@ -111,19 +111,14 @@ logFile=$outputRunDataDir/logs/summary.log
 
 ## Define inputs
 export input_WBD_gdb=$inputDataDir/wbd/WBD_National.gpkg
-export input_NWM_Lakes=$inputDataDir/nwm_hydrofabric/nwm_lakes.gpkg
-export input_NWM_Catchments_fr=$inputDataDir/nwm_hydrofabric/nwm_catchments.gpkg
-export input_NWM_Catchments_ms=$inputDataDir/nwm_hydrofabric/nwm_catchments_ms.gpkg
-export input_NWM_Flows_fr=$inputDataDir/nwm_hydrofabric/nwm_flows.gpkg
-export input_NWM_Flows_ms=$inputDataDir/nwm_hydrofabric/nwm_flows_ms.gpkg
-export input_NWM_Headwaters=$inputDataDir/nwm_hydrofabric/nwm_headwaters.gpkg
-export input_nhd_flowlines_fr=$inputDataDir/nhdplus_vectors_aggregate/NHDPlusBurnLineEvent_fr_adjusted.gpkg
-export input_nhd_flowlines_ms=$inputDataDir/nhdplus_vectors_aggregate/NHDPlusBurnLineEvent_ms_adjusted.gpkg
-export input_nhd_headwaters_fr=$inputDataDir/nhdplus_vectors_aggregate/nhd_headwaters_adjusted_fr.gpkg
-export input_nhd_headwaters_ms=$inputDataDir/nhdplus_vectors_aggregate/nhd_headwaters_adjusted_ms.gpkg
-
+export input_nwm_lakes=$inputDataDir/nwm_hydrofabric/nwm_lakes.gpkg
+export input_nwm_catchments=$inputDataDir/nwm_hydrofabric/nwm_catchments.gpkg
+export input_nwm_flows=$inputDataDir/nwm_hydrofabric/nwm_flows.gpkg
+export input_nhd_flowlines=$inputDataDir/nhdplus_vectors_aggregate/agg_nhd_streams_adj.gpkg
+export input_nhd_headwaters=$inputDataDir/nhdplus_vectors_aggregate/agg_nhd_headwaters_adj.gpkg
[…]

### 2021-05-18 lib/add_crosswalk.py

@@ -1,302 +1,0 @@
-#!/usr/bin/env python3
-
-import os
-import geopandas as gpd
-import pandas as pd
-from numpy import unique
-from rasterstats import zonal_stats
-import json
-import argparse
-import sys
-from utils.shared_functions import getDriver
-from utils.shared_variables import FIM_ID
-
-def add_crosswalk(input_catchments_fileName,input_flows_fileName,input_srcbase_fileName,output_catchments_fileName,output_flows_fileName,output_src_fileName,output_src_json_fileName,output_crosswalk_fileName,output_hydro_table_fileName,input_huc_fileName,input_nwmflows_fileName,input_nwmcatras_fileName,mannings_n,input_nwmcat_fileName,extent,small_segments_filename,calibration_mode=False):
-
-    input_catchments = gpd.read_file(input_catchments_fileName)
-    input_flows = gpd.read_file(input_flows_fileName)
-    input_huc = gpd.read_file(input_huc_fileName)
-    input_nwmflows = gpd.read_file(input_nwmflows_fileName)
-    min_catchment_area = float(os.environ['min_catchment_area'])
-    min_stream_length = float(os.environ['min_stream_length'])
-
-    if extent == 'FR':
-        ## crosswalk using majority catchment method
-
-        # calculate majority catchments
-        majority_calc = zonal_stats(input_catchments, input_nwmcatras_fileName, stats=['majority'], geojson_out=True)
-        input_majorities = gpd.GeoDataFrame.from_features(majority_calc)
-        input_majorities = input_majorities.rename(columns={'majority' : 'feature_id'})
-
-        input_majorities = input_majorities[:][input_majorities['feature_id'].notna()]
-        if input_majorities.feature_id.dtype != 'int': input_majorities.feature_id = input_majorities.feature_id.astype(int)
-        if input_majorities.HydroID.dtype != 'int': input_majorities.HydroID = input_majorities.HydroID.astype(int)
-
-        input_nwmflows = input_nwmflows.rename(columns={'ID':'feature_id'})
-        if input_nwmflows.feature_id.dtype != 'int': input_nwmflows.feature_id = input_nwmflows.feature_id.astype(int)
-        relevant_input_nwmflows = input_nwmflows[input_nwmflows['feature_id'].isin(input_majorities['feature_id'])]
-        relevant_input_nwmflows = relevant_input_nwmflows.filter(items=['feature_id','order_'])
-
[…]

### 2021-05-18 lib/adjust_headwater_streams.py

@@ -1,142 +1,0 @@
-#!/usr/bin/env python3
-
-import geopandas as gpd
-import pandas as pd
-import numpy as np
-from os.path import splitext
-from tqdm import tqdm
-import argparse
-import pygeos
-from shapely.geometry import Point,LineString
-from shapely.ops import split
-from shapely.wkb import dumps, loads
-from utils.shared_variables import PREP_PROJECTION
-from utils.shared_functions import getDriver
-
-def adjust_headwaters(huc,nhd_streams,headwaters,headwater_id):
-
-    # identify true headwater segments
-    if nhd_streams['headwaters_id'].dtype=='int':
-        nhd_streams_adj = nhd_streams.loc[(nhd_streams.headwaters_id > 0) & (nhd_streams.downstream_of_headwater == False),:].copy()
-        if headwaters[headwater_id].dtype != 'int': headwaters[headwater_id] = headwaters[headwater_id].astype(int)
-    else:
-        nhd_streams_adj = nhd_streams.loc[(nhd_streams.headwaters_id.notna()) & (nhd_streams.downstream_of_headwater == False),:].copy()
-
-    nhd_streams_adj = nhd_streams_adj.explode()
-    nhd_streams_adj = nhd_streams_adj.reset_index(drop=True)
-
-    headwater_limited = headwaters.merge(nhd_streams_adj["headwaters_id"],left_on=headwater_id, right_on="headwaters_id",how='right')
-
-    headwaterstreams = []
-    referencedpoints = []
-
-    for index, point in headwater_limited.iterrows():
-
-        # convert headwaterpoint geometries to WKB representation
-        wkb_points = dumps(point.geometry)
-
-        # create pygeos headwaterpoint geometries from WKB representation
-        pointbin_geom = pygeos.io.from_wkb(wkb_points)
-
-        # Closest segment to headwater
-        closest_stream = nhd_streams_adj.loc[nhd_streams_adj["headwaters_id"]==point[headwater_id]]
-
-        try: # seeing inconsistent geometry objects even after exploding nhd_streams_adj; not sure why this is
-            closest_stream =closest_stream.explode()
-        except:
-            pass
-        try:
-            wkb_closest_stream = dumps(closest_stream.geometry[0])
-        except:
-            wkb_closest_stream = dumps(closest_stream.geometry[0][0])
-
-        streambin_geom = pygeos.io.from_wkb(wkb_closest_stream)
-
-        # Linear reference headwater to closest stream segment
[…]

### 2021-05-18 lib/aggregate_vector_inputs.py

@@ -1,412 +1,0 @@
-#!/usr/bin/env python3
-
-import os
-import geopandas as gpd
-from utils.shared_variables import PREP_PROJECTION
-from utils.shared_functions import getDriver
-from derive_headwaters import findHeadWaterPoints
-from reduce_nhd_stream_density import subset_nhd_network
-from adjust_headwater_streams import adjust_headwaters
-from tqdm import tqdm
-from os.path import splitext
-from shapely.geometry import Point
-from concurrent.futures import ProcessPoolExecutor,as_completed
-from collections import deque
-import numpy as np
-from shapely.wkb import dumps, loads
-import pygeos
-
-in_dir ='data/inputs/nhdplus_vectors'
-nwm_dir = 'data/inputs/nwm_hydrofabric'
-wbd_dir = 'data/inputs/wbd'
-ahps_dir = 'data/inputs/ahp_sites'
-agg_dir = 'data/inputs/nhdplus_vectors_aggregate'
-
-wbd_filename = os.path.join(wbd_dir, 'WBD_National.gpkg')
-nwm_streams_fr_filename = os.path.join(nwm_dir,'nwm_flows.gpkg')
-nwm_streams_ms_filename = os.path.join(nwm_dir,'nwm_flows_ms.gpkg')
-nwm_headwaters_filename = os.path.join(nwm_dir,'nwm_headwaters.gpkg')
-nwm_huc4_intersections_ms_filename = os.path.join(nwm_dir,'nwm_ms_huc4_intersections.gpkg')
-nwm_huc4_intersections_fr_filename = os.path.join(nwm_dir,'nwm_fr_huc4_intersections.gpkg')
-
-def subset_nwm_ms_streams(args):
-    nwm_streams_filename    = args[0]
-    in_dir                  = args[1]
-    ahps_dir                = args[2]
-    output_filename         = args[3]
-
-    # subset nwm network to ms
-    ahps_headwaters_filename = os.path.join(ahps_dir,'bed_lids.gpkg')
-    ahps_headwaters = gpd.read_file(ahps_headwaters_filename)
-
-    nwm_streams = gpd.read_file(nwm_streams_filename)
-
-    nwm_streams['is_headwater'] = False
-    nwm_streams['downstream_of_headwater'] = False
-
-    nwm_streams.loc[nwm_streams.ID.isin(list(ahps_headwaters.nwm_featur)),'is_headwater'] = True
-
-    ## subset NHDPlus HR
[…]

### 2021-05-18 lib/output_cleanup.py

@@ -1,87 +1,0 @@
-#!/usr/bin/env python3
-import os
-import csv
-import json
-import shutil
-import argparse
-
-def output_cleanup(huc_number, output_folder_path, additional_whitelist, is_production, viz_post_processing):
-    '''
-    Processes all the final output files to cleanup and add post-processing
-
-    Parameters
-    ----------
-    huc_number : STR
-        The HUC
-    output_folder_path : STR
-        Path to the outputs for the specific huc
-    additional_whitelist : STR
-        Additional list of files to keep during a production run
-    is_production : BOOL
-        Determine whether or not to only keep whitelisted production files
-    is_viz_post_processing : BOOL
-        Determine whether or not to process outputs for Viz
-    '''
-
-    # List of files that will be saved during a production run
-    production_whitelist = [
-        'rem_zeroed_masked.tif',
-        'gw_catchments_reaches_filtered_addedAttributes_crosswalked.gpkg',
-        'demDerived_reaches_split_filtered_addedAttributes_crosswalked.gpkg',
-        'gw_catchments_reaches_filtered_addedAttributes.tif',
-        'hydroTable.csv',
-        'src.json',
-        'small_segments.csv'
-    ]
-
-    # List of files that will be saved during a viz run
-    viz_whitelist = [
-        'rem_zeroed_masked.tif',
-        'gw_catchments_reaches_filtered_addedAttributes.tif',
-        'hydroTable.csv',
-        'src.json'
-    ]
-
-    # If "production" run, only keep whitelisted files
-    if is_production and not is_viz_post_processing:
-        whitelist_directory(output_folder_path, production_whitelist, additional_whitelist)
-
-    # If Viz post-processing is enabled, form output files to Viz specifications
-    if is_viz_post_processing:
-        # Step 1, keep only files that Viz needs
-        whitelist_directory(output_folder_path, viz_whitelist, additional_whitelist)
-
-def whitelist_directory(directory_path, whitelist, additional_whitelist):
-    # Add any additional files to the whitelist that the user wanted to keep
-    if additional_whitelist:
-        whitelist = whitelist + [filename for filename in additional_whitelist.split(',')]
-
-    # Delete any non-whitelisted files
-    directory = os.fsencode(directory_path)
-    for file in os.listdir(directory_path):
-        filename = os.fsdecode(file)
-        if filename not in whitelist:
-            os.remove(os.path.join(directory_path, filename))
-
-
-if __name__ == '__main__':
-    #Parse arguments
[…]

### 2021-05-18 lib/reduce_nhd_stream_density.py

@@ -1,165 +1,0 @@
-#!/usr/bin/env python3
-
-import geopandas as gpd
-import pandas as pd
-import numpy as np
-from os.path import splitext
-from tqdm import tqdm
-from collections import deque
-import argparse
-import pygeos
-from shapely.wkb import dumps
-from utils.shared_functions import getDriver
-
-def subset_nhd_network(huc4,huc4_mask,selected_wbd8,nhd_streams_filename,headwaters_filename,headwater_id,nwm_intersections_filename):
-
-    headwater_streams = pd.DataFrame()
-
-    nhd_streams = gpd.read_file(nhd_streams_filename)
-
-    for index, row in selected_wbd8.iterrows():
-        huc = row["HUC8"]
-
-        if huc.startswith(str(huc4)):
-            huc8_mask = selected_wbd8.loc[selected_wbd8.HUC8.str.startswith(huc)]
-            huc8_mask = huc8_mask.reset_index(drop=True)
-
-            headwaters_mask = gpd.read_file(headwaters_filename, mask = huc8_mask)
-            headwaters_mask = headwaters_mask.reset_index(drop=True)
-
-            streams_subset = gpd.read_file(nhd_streams_filename, mask = huc8_mask)
-
-            if not streams_subset.empty:
-                streams_subset.loc[:,'is_headwater'] = False
-                streams_subset = streams_subset.reset_index(drop=True)
-
-                streams_subset['b_geom'] = None
-                for index, linestring in enumerate(streams_subset.geometry):
-                    streams_subset.at[index, 'b_geom'] = dumps(linestring)
-
-                # create pygeos nhd stream geometries from WKB representation
-                streambin_geom = pygeos.io.from_wkb(streams_subset['b_geom'])
-
-                streams_subset.loc[:,'HUC8'] = str(huc)
-
-                if headwaters_mask[headwater_id].dtype=='int':
-                    n = -1
-                else:
-                    n = ''
-
-                streams_subset.loc[:,'headwaters_id'] = n
-
-                # find stream segment closest to headwater point
-                for index, point in headwaters_mask.iterrows():
-
-                    # convert headwaterpoint geometries to WKB representation
-                    wkb_points = dumps(point.geometry)
-
-                    # create pygeos headwaterpoint geometries from WKB representation
-                    pointbin_geom = pygeos.io.from_wkb(wkb_points)
-
[…]

### 2021-05-18 lib/rem.py

@@ -1,190 +1,0 @@
-#!/usr/bin/env python3
-
-from numba import njit, typeof, typed, types
-import rasterio
-import numpy as np
-import argparse
-import os
-import pandas as pd
-from osgeo import ogr, gdal
-import geopandas as gpd
-from utils.shared_functions import getDriver
-
-
-def rel_dem(dem_fileName, pixel_watersheds_fileName, rem_fileName, thalweg_raster, hydroid_fileName, hand_ref_elev_fileName, dem_reaches_filename):
-    """
-        Calculates REM/HAND/Detrended DEM
-
-        Parameters
-        ----------
-        dem_fileName : str
-            File name of pit filled DEM raster.
-        pixel_watersheds_fileName : str
-            File name of stream pixel watersheds raster.
-        rem_fileName : str
-            File name of output relative elevation raster.
-        hydroid_fileName : str
-            File name of the hydroid raster (i.e. gw_catchments_reaches.tif)
-        hand_ref_elev_fileName
-            File name of the output csv containing list of hydroid values and HAND zero/reference elev
-        dem_reaches_filename
-            File name of the reaches layer to populate HAND elevation attribute values and overwrite as output
-
-    """
-
-    # ------------------------------------------- Get catchment_hydroid_dict --------------------------------------------------- #
-    # The following creates a dictionary of the catchment ids (key) and their hydroid along the thalweg (value).
-    # This is needed to produce a HAND zero reference elevation by hydroid dataframe (helpful for evaluating rating curves & bathy properties)
-    @njit
-    def make_catchment_hydroid_dict(flat_value_raster, catchment_hydroid_dict, flat_catchments, thalweg_window):
-
-        for i,cm in enumerate(flat_catchments):
-            if thalweg_window[i] == 1:  # Only allow reference hydroid to be within thalweg.
-                catchment_hydroid_dict[cm] = flat_value_raster[i]
-        return(catchment_hydroid_dict)
-
-    # Open the masked gw_catchments_pixels_masked, hydroid_raster, and dem_thalwegCond_masked.
-    gw_catchments_pixels_masked_object = rasterio.open(pixel_watersheds_fileName)
-    hydroid_pixels_object = rasterio.open(hydroid_fileName)
-    thalweg_raster_object = rasterio.open(thalweg_raster)
-
-    # Specify raster object metadata.
-    meta = hydroid_pixels_object.meta.copy()
[…]

### 2021-05-18 lib/run_by_unit.sh

@@ -1,444 +1,0 @@
-#!/bin/bash -e
-
-## INITIALIZE TOTAL TIME TIMER ##
-T_total_start
-
-echo -e $startDiv"Parameter Values"
-echo -e "extent=$extent"
-echo -e "agree_DEM_buffer=$agree_DEM_buffer"
-echo -e "wbd_buffer=$wbd_buffer"
-echo -e "ms_buffer_dist=$ms_buffer_dist"
-echo -e "lakes_buffer_dist_meters=$lakes_buffer_dist_meters"
-echo -e "negative_burn_value=$negative_burn_value"
-echo -e "max_split_distance_meters=$max_split_distance_meters"
-echo -e "mannings_n=$manning_n"
-echo -e "stage_min_meters=$stage_min_meters"
-echo -e "stage_interval_meters=$stage_interval_meters"
-echo -e "stage_max_meters=$stage_max_meters"
-echo -e "slope_min=$slope_min"
-echo -e "ms_buffer_dist=$ms_buffer_dist"
-echo -e "ncores_gw=$ncores_gw"
-echo -e "ncores_fd=$ncores_fd"
-echo -e "default_max_jobs=$default_max_jobs"
-echo -e "memfree=$memfree"$stopDiv
-
-## SET OUTPUT DIRECTORY FOR UNIT ##
-hucNumber="$1"
-outputHucDataDir=$outputRunDataDir/$hucNumber
-mkdir $outputHucDataDir
-
-## SET VARIABLES AND FILE INPUTS ##
-hucUnitLength=${#hucNumber}
-huc4Identifier=${hucNumber:0:4}
-huc2Identifier=${hucNumber:0:2}
-input_NHD_WBHD_layer=WBDHU$hucUnitLength
-input_DEM=$inputDataDir/nhdplus_rasters/HRNHDPlusRasters"$huc4Identifier"/elev_cm.tif
-input_NLD=$inputDataDir/nld_vectors/huc2_levee_lines/nld_preprocessed_"$huc2Identifier".gpkg
-
-# Define the landsea water body mask using either Great Lakes or Ocean polygon input #
-if [[ $huc2Identifier == "04" ]] ; then
-  input_LANDSEA=$inputDataDir/landsea/gl_water_polygons.gpkg
-  echo -e "Using $input_LANDSEA for water body mask (Great Lakes)"
-else
[…]

### 2021-05-18 lib/utils/shared_functions.py

@@ -1,85 +1,0 @@
-#!/usr/bin/env python3
-
-import os
-from os.path import splitext
-
-def getDriver(fileName):
-
-    driverDictionary = {'.gpkg' : 'GPKG','.geojson' : 'GeoJSON','.shp' : 'ESRI Shapefile'}
-    driver = driverDictionary[splitext(fileName)[1]]
-
-    return(driver)
-
-def pull_file(url, full_pulled_filepath):
-    """
-    This helper function pulls a file and saves it to a specified path.
-
-    Args:
-        url (str): The full URL to the file to download.
-        full_pulled_filepath (str): The full system path where the downloaded file will be saved.
-    """
-    import urllib.request
-
-    print("Pulling " + url)
-    urllib.request.urlretrieve(url, full_pulled_filepath)
-
-
-def delete_file(file_path):
-    """
-    This helper function deletes a file.
-
-    Args:
-        file_path (str): System path to a file to be deleted.
-    """
-
-    try:
-        os.remove(file_path)
-    except FileNotFoundError:
-        pass
-
-
-def run_system_command(args):
-    """
-    This helper function takes a system command and runs it. This function is designed for use
-    in multiprocessing.
-
-    Args:
-        args (list): A single-item list, the first and only item being a system command string.
-    """
-
-    # Parse system command.
-    command = args[0]
-
-    # Run system command.
-    os.system(command)
-
-
-def subset_wbd_gpkg(wbd_gpkg, multilayer_wbd_geopackage):
-
-    import geopandas as gp
-    from utils.shared_variables import CONUS_STATE_LIST, PREP_PROJECTION
-
-    print("Subsetting " + wbd_gpkg + "...")
-    # Read geopackage into dataframe.
-    wbd = gp.read_file(wbd_gpkg)
-    gdf = gp.GeoDataFrame(wbd)
-
-    for index, row in gdf.iterrows():
-        state = row["STATES"]
-        if state != None:  # Some polygons are empty in the STATES field.
-            keep_flag = False  # Default to Fault, i.e. to delete the polygon.
-            if state in CONUS_STATE_LIST:
-                keep_flag = True
-            # Only split if multiple states present. More efficient this way.
-            elif len(state) > 2:
-                for wbd_state in state.split(","):  # Some polygons have multiple states, separated by a comma.
[…]

### 2021-05-18 lib/utils/shared_variables.py

@@ -1,35 +1,0 @@
-#!/usr/bin/env python3
-
-# Projections.
-#PREP_PROJECTION = "+proj=aea +datum=NAD83 +x_0=0.0 +y_0=0.0 +lon_0=96dW +lat_0=23dN +lat_1=29d30'N +lat_2=45d30'N +towgs84=-0.9956000824677655,1.901299877314078,0.5215002840524426,0.02591500053005733,0.009425998542707753,0.01159900118427752,-0.00062000005129903 +no_defs +units=m"
-PREP_PROJECTION = 'PROJCS["USA_Contiguous_Albers_Equal_Area_Conic_USGS_version",GEOGCS["NAD83",DATUM["North_American_Datum_1983",SPHEROID["GRS 1980",6378137,298.2572221010042,AUTHORITY["EPSG","7019"]],AUTHORITY["EPSG","6269"]],PRIMEM["Greenwich",0],UNIT["degree",0.0174532925199433],AUTHORITY["EPSG","4269"]],PROJECTION["Albers_Conic_Equal_Area"],PARAMETER["standard_parallel_1",29.5],PARAMETER["standard_parallel_2",45.5],PARAMETER["latitude_of_center",23],PARAMETER["longitude_of_center",-96],PARAMETER["false_easting",0],PARAMETER["false_northing",0],UNIT["metre",1,AUTHORITY["EPSG","9001"]]]'
-
-# -- Data URLs-- #
-NHD_URL_PARENT = r'https://prd-tnm.s3.amazonaws.com/StagedProducts/Hydrography/NHDPlusHR/Beta/GDB/'
-NWM_HYDROFABRIC_URL = r'http://www.nohrsc.noaa.gov/pub/staff/keicher/NWM_live/web/data_tools/NWM_channel_hydrofabric.tar.gz'  # Temporary
-WBD_NATIONAL_URL = r'https://prd-tnm.s3.amazonaws.com/StagedProducts/Hydrography/WBD/National/GDB/WBD_National_GDB.zip'
-WBD_HU2_URL_PARENT = r'http://prd-tnm.s3-website-us-west-2.amazonaws.com/?prefix=StagedProducts/Hydrography/WBD/HU2/GDB'
-
-# -- Prefixes and Suffixes -- #
[…]

### 2021-05-18 src/acquire_and_preprocess_inputs.py

@@ -1,0 +1,382 @@
+#!/usr/bin/env python3
+
+import os
+import argparse
+import csv
+import sys
+sys.path.append('/foss_fim/src')
+import shutil
+from multiprocessing import Pool
+import geopandas as gpd
+from urllib.error import HTTPError
+from tqdm import tqdm
+
+from utils.shared_variables import (NHD_URL_PARENT,
+                                    NHD_URL_PREFIX,
+                                    NHD_RASTER_URL_SUFFIX,
+                                    NHD_VECTOR_URL_SUFFIX,
+                                    NHD_VECTOR_EXTRACTION_PREFIX,
+                                    NHD_VECTOR_EXTRACTION_SUFFIX,
+                                    PREP_PROJECTION,
+                                    WBD_NATIONAL_URL,
+                                    FIM_ID,
+                                    OVERWRITE_WBD,
+                                    OVERWRITE_NHD,
+                                    OVERWRITE_ALL)
+
+from utils.shared_functions import (pull_file, run_system_command,
+                                    subset_wbd_gpkg, delete_file,
+                                    getDriver)
+
+NHDPLUS_VECTORS_DIRNAME = 'nhdplus_vectors'
+NHDPLUS_RASTERS_DIRNAME = 'nhdplus_rasters'
+NWM_HYDROFABRIC_DIRNAME = 'nwm_hydrofabric'
+NWM_FILE_TO_SUBSET_WITH = 'nwm_flows.gpkg'
+
+def subset_wbd_to_nwm_domain(wbd,nwm_file_to_use):
+
+    intersecting_indices = [not (gpd.read_file(nwm_file_to_use,mask=b).empty) for b in wbd.geometry]
+
+    return(wbd[intersecting_indices])
+
+def pull_and_prepare_wbd(path_to_saved_data_parent_dir,nwm_dir_name,nwm_file_to_use,overwrite_wbd,num_workers):
+    """
+    This helper function pulls and unzips Watershed Boundary Dataset (WBD) data. It uses the WBD URL defined by WBD_NATIONAL_URL.
+    This function also subsets the WBD layers (HU4, HU6, HU8) to CONUS and converts to geopkacage layers.
+
+    Args:
[…]

### 2021-05-18 src/add_crosswalk.py

@@ -1,0 +1,320 @@
+#!/usr/bin/env python3
+
+import os
+import geopandas as gpd
+import pandas as pd
+from numpy import unique
+from rasterstats import zonal_stats
+import json
+import argparse
+import sys
+from bathy_rc_adjust import bathy_rc_lookup
+from utils.shared_functions import getDriver
+from utils.shared_variables import FIM_ID
+
+def add_crosswalk(input_catchments_fileName,input_flows_fileName,input_srcbase_fileName,input_bathy_fileName,output_bathy_fileName,output_bathy_streamorder_fileName,output_bathy_thalweg_fileName,output_bathy_xs_lookup_fileName,output_catchments_fileName,output_flows_fileName,output_src_fileName,output_src_json_fileName,output_crosswalk_fileName,output_hydro_table_fileName,input_huc_fileName,input_nwmflows_fileName,input_nwmcatras_fileName,mannings_n,input_nwmcat_fileName,extent,small_segments_filename,calibration_mode=False):
+
+    input_catchments = gpd.read_file(input_catchments_fileName)
+    input_flows = gpd.read_file(input_flows_fileName)
+    input_huc = gpd.read_file(input_huc_fileName)
+    input_nwmflows = gpd.read_file(input_nwmflows_fileName)
+    min_catchment_area = float(os.environ['min_catchment_area']) #0.25#
+    min_stream_length = float(os.environ['min_stream_length']) #0.5#
+    bathy_src_calc = os.environ['bathy_src_modification'] == "True" # env variable to toggle on/off the bathy calc and src modifications
+
+    if extent == 'FR':
+        ## crosswalk using majority catchment method
+
+        # calculate majority catchments
+        majority_calc = zonal_stats(input_catchments, input_nwmcatras_fileName, stats=['majority'], geojson_out=True)
+        input_majorities = gpd.GeoDataFrame.from_features(majority_calc)
+        input_majorities = input_majorities.rename(columns={'majority' : 'feature_id'})
+
+        input_majorities = input_majorities[:][input_majorities['feature_id'].notna()]
+        if input_majorities.feature_id.dtype != 'int': input_majorities.feature_id = input_majorities.feature_id.astype(int)
+        if input_majorities.HydroID.dtype != 'int': input_majorities.HydroID = input_majorities.HydroID.astype(int)
+
+        input_nwmflows = input_nwmflows.rename(columns={'ID':'feature_id'})
[…]

### 2021-05-18 src/adjust_headwater_streams.py

@@ -1,0 +1,185 @@
+#!/usr/bin/env python3
+
+import geopandas as gpd
+import pandas as pd
+import numpy as np
+import argparse
+import pygeos
+from shapely.geometry import Point,LineString
+from shapely.ops import split
+from shapely.wkb import dumps, loads
+from utils.shared_variables import PREP_PROJECTION
+from utils.shared_functions import getDriver
+import warnings
+warnings.simplefilter("ignore")
+
+
+def adjust_headwaters(huc,nhd_streams,nwm_headwaters,nws_lids,headwater_id):
+
+    # Identify true headwater segments
+    nhd_streams_adj = nhd_streams.loc[(nhd_streams.headwaters_id > 0) & (nhd_streams.downstream_of_headwater == False),:].copy()
+    nhd_streams_adj = nhd_streams_adj.explode()
+    nhd_streams_adj = nhd_streams_adj.reset_index(drop=True)
+
+    if nwm_headwaters["site_id"].dtype != 'int': nwm_headwaters["site_id"] = nwm_headwaters["site_id"].astype(int)
+    headwater_limited = nwm_headwaters.merge(nhd_streams_adj[["headwaters_id","mainstem"]],left_on="site_id", right_on="headwaters_id",how='right')
+    headwater_limited = headwater_limited.drop(columns=['headwaters_id'])
+
+    nws_lid_limited = nws_lids.merge(nhd_streams[["nws_lid"]],left_on="site_id", right_on="nws_lid",how='right')
+    nws_lid_limited = nws_lid_limited.loc[nws_lid_limited.nws_lid!='']
+    nws_lid_limited = nws_lid_limited.drop(columns=['nws_lid'])
+
+    # Check for issues in nws_lid layer
+    if len(nws_lid_limited) < len(nws_lids):
+        missing_nws_lids = list(set(nws_lids.site_id) - set(nws_lid_limited.site_id))
+        print (f"nws lid(s) {missing_nws_lids} missing from aggregate dataset in huc {huc}")
+
+    # Combine NWM headwaters and AHPS sites to be snapped to NHDPlus HR segments
+    headwater_pts = headwater_limited.append(nws_lid_limited)
+    headwater_pts = headwater_pts.reset_index(drop=True)
+
+    if headwater_pts is not None:
+
+        headwaterstreams = []
+        referencedpoints = []
+        snapped_ahps = []
+        nws_lid = []
+        for index, point in headwater_pts.iterrows():
+
[…]

### 2021-05-18 src/aggregate_fim_outputs.py

@@ -1,0 +1,222 @@
+#!/usr/bin/env python3
+
+import os
+import argparse
+from multiprocessing import Pool
+import pandas as pd
+import json
+import rasterio
+from rasterio.merge import merge
+from rasterio.warp import calculate_default_transform, reproject, Resampling
+import shutil
+import csv
+from utils.shared_variables import PREP_PROJECTION,VIZ_PROJECTION
+
+def aggregate_fim_outputs(args):
+
+    fim_out_dir   = args[0]
+    huc6          = args[1]
+    huc_list      = args[2]
+
+    print(f"aggregating {huc6}")
+
+    huc6_dir = os.path.join(fim_out_dir,'aggregate_fim_outputs',str(huc6))
+    os.makedirs(huc6_dir, exist_ok=True)
+
+    # aggregate file name paths
+    aggregate_hydrotable = os.path.join(fim_out_dir,'aggregate_fim_outputs',str(huc6),'hydroTable.csv')
+    aggregate_src = os.path.join(fim_out_dir,'aggregate_fim_outputs',str(huc6),f'rating_curves_{huc6}.json')
+
+    for huc in huc_list:
+
+        # original file paths
+        hydrotable_filename = os.path.join(fim_out_dir,huc,'hydroTable.csv')
+        src_filename = os.path.join(fim_out_dir,huc,'src.json')
+
+        if len(huc)> 6:
+
+            # open hydrotable
+            hydrotable = pd.read_csv(hydrotable_filename)
+
+            # write/append aggregate hydrotable
+            if os.path.isfile(aggregate_hydrotable):
+                hydrotable.to_csv(aggregate_hydrotable,index=False, mode='a',header=False)
+            else:
+                hydrotable.to_csv(aggregate_hydrotable,index=False)
+
+            del hydrotable
+
+            # open src
+            src = open(src_filename)
+            src = json.load(src)
+
+            # write/append aggregate src
+            if os.path.isfile(aggregate_src):
+
+                with open(aggregate_src, "r+") as file:
+                    data = json.load(file)
+                    data.update(src)
+
+                with open(aggregate_src, 'w') as outfile:
+                    json.dump(data, outfile)
+            else:
+                with open(aggregate_src, 'w') as outfile:
+                        json.dump(src, outfile)
+
+            del src
+
+        else:
[…]

### 2021-05-18 src/aggregate_vector_inputs.py

@@ -1,0 +1,403 @@
+#!/usr/bin/env python3
+
+import os
+import sys
+# sys.path.append('/foss_fim/src')
+import geopandas as gpd
+from utils.shared_variables import PREP_PROJECTION
+from utils.shared_functions import getDriver
+from derive_headwaters import findHeadWaterPoints
+from reduce_nhd_stream_density import subset_nhd_network
+from adjust_headwater_streams import adjust_headwaters
+from shapely.geometry import Point
+from concurrent.futures import ProcessPoolExecutor
+from collections import deque
+import numpy as np
+from shapely.wkb import dumps, loads
+import pygeos
+
+nhdplus_vectors_dir = os.environ.get('nhdplus_vectors_dir')
+wbd_filename = os.environ.get('wbd_filename')
+nwm_streams_orig_filename = os.environ.get('nwm_streams_orig_filename')
+nwm_streams_all_filename = os.environ.get('nwm_streams_all_filename')
+nwm_headwaters_filename = os.environ.get('nwm_headwaters_filename')
+nwm_catchments_orig_filename = os.environ.get('nwm_catchments_orig_filename')
+nwm_catchments_all_filename = os.environ.get('nwm_catchments_all_filename')
+ahps_filename = os.environ.get('ahps_filename')
+nwm_huc4_intersections_filename = os.environ.get('nwm_huc4_intersections_filename')
+nhd_huc8_intersections_filename = os.environ.get('nhd_huc8_intersections_filename')
+agg_nhd_headwaters_adj_fileName = os.environ['agg_nhd_headwaters_adj_fileName']
+agg_nhd_streams_adj_fileName = os.environ['agg_nhd_streams_adj_fileName']
+
+
+def identify_nwm_ms_streams(nwm_streams_filename,ahps_filename,nwm_streams_all_filename):
+
+    # Subset nwm network to ms
+    ahps_headwaters = gpd.read_file(ahps_filename)
+
+    nwm_streams = gpd.read_file(nwm_streams_filename)
+
+    # Remove mainstem column if it already exists
+    nwm_streams = nwm_streams.drop(['mainstem'], axis=1, errors='ignore')
+
+    nwm_streams['is_headwater'] = False
+
+    nwm_streams.loc[nwm_streams.ID.isin(list(ahps_headwaters.nwm_featur)),'is_headwater'] = True
+
+    # Subset NHDPlus HR
+    nwm_streams['is_relevant_stream'] = nwm_streams['is_headwater'].copy()
+
[…]

### 2021-05-18 src/agreedem.py

@@ -1,0 +1,261 @@
+#!/usr/bin/env python3
+import rasterio
+import numpy as np
+import os
+import argparse
+from r_grow_distance import r_grow_distance
+
+
+def agreedem(rivers_raster, dem, output_raster, workspace, grass_workspace, buffer_dist, smooth_drop, sharp_drop, delete_intermediate_data):
+    '''
+    Produces a hydroconditioned raster using the AGREE DEM methodology as described by Ferdi Hellweger (https://www.caee.utexas.edu/prof/maidment/gishydro/ferdi/research/agree/agree.html). The GRASS gis tool r.grow.distance is used to calculate intermediate allocation and proximity rasters.
+
+    Parameters
+    ----------
+    rivers_raster : STR
+        Path to the river raster. River cells = 1 and non-river cells = 0. For example, see flows_grid_boolean.tif.
+    dem : STR
+        Elevation DEM (units assumed to be in meters). For example, see dem_meters.tif.
+    output_raster : STR
+        Path to output raster. For example, dem_burned.tif
+    workspace : STR
+        Path to workspace to save all intermediate files.
+    grass_workspace : STR
+        Path to the temporary workspace for grass inputs. This temporary workspace is deleted once grass datasets are produced and exported to tif files.
+    buffer_dist : FLOAT
+        AGREE stream buffer distance (in meters) on either side of stream.
+    smooth_drop : FLOAT
+        Smooth drop distance (in meters). Typically this has been 10m.
+    sharp_drop : FLOAT
+        Sharp drop distance (in meters). Typically this has been 1000m.
+    delete_intermediate_data: BOOL
+        If True all intermediate data is deleted, if False (default) no intermediate datasets are deleted.
+
+    Returns
+    -------
+    None.
+
+    '''
+    #------------------------------------------------------------------
+    # 1. From Hellweger documentation: Compute the vector grid
+    # (vectgrid). The cells in the vector grid corresponding to the
+    # lines in the vector coverage have data. All other cells have no
+    # data.
+
+    # Import dem layer and river layer and get dem profile.
+    elev = rasterio.open(dem)
+    dem_profile = elev.profile
+
+    rivers = rasterio.open(rivers_raster)
+
+    # Define smogrid profile and output file
+    smo_profile = dem_profile.copy()
+    smo_profile.update(nodata = 0)
+    smo_profile.update(dtype = 'float32')
[…]

### 2021-05-18 src/clip_vectors_to_wbd.py

@@ -1,0 +1,179 @@
+#!/usr/bin/env python3
+
+import sys
+import geopandas as gpd
+import argparse
+from os.path import splitext
+from shapely.geometry import MultiPolygon,Polygon,Point
+from utils.shared_functions import getDriver
+
+def subset_vector_layers(hucCode,nwm_streams_filename,nhd_streams_filename,nwm_lakes_filename,nld_lines_filename,nwm_catchments_filename,nhd_headwaters_filename,landsea_filename,wbd_filename,wbd_buffer_filename,subset_nhd_streams_filename,subset_nld_lines_filename,subset_nwm_lakes_filename,subset_nwm_catchments_filename,subset_nhd_headwaters_filename,subset_nwm_streams_filename,subset_landsea_filename,extent,great_lakes_filename,wbd_buffer_distance,lake_buffer_distance,dissolveLinks=False):
+
+    hucUnitLength = len(str(hucCode))
+
+    # Get wbd buffer
+    wbd = gpd.read_file(wbd_filename)
+    wbd_buffer = wbd.copy()
+    wbd_buffer.geometry = wbd.geometry.buffer(wbd_buffer_distance,resolution=32)
+    projection = wbd_buffer.crs
+
+    great_lakes = gpd.read_file(great_lakes_filename, mask = wbd_buffer).reset_index(drop=True)
+
+    if not great_lakes.empty:
+        print("Masking Great Lakes for HUC{} {}".format(hucUnitLength,hucCode),flush=True)
+
+        # Clip excess lake area
+        great_lakes = gpd.clip(great_lakes, wbd_buffer)
+
+        # Buffer remaining lake area
+        great_lakes.geometry = great_lakes.buffer(lake_buffer_distance)
+
+        # Removed buffered GL from WBD buffer
+        wbd_buffer = gpd.overlay(wbd_buffer, great_lakes, how='difference')
+        wbd_buffer = wbd_buffer[['geometry']]
+        wbd_buffer.to_file(wbd_buffer_filename,driver=getDriver(wbd_buffer_filename),index=False)
+
+    else:
+        wbd_buffer = wbd_buffer[['geometry']]
+        wbd_buffer.to_file(wbd_buffer_filename,driver=getDriver(wbd_buffer_filename),index=False)
+
+    del great_lakes
+
+    # Clip ocean water polygon for future masking ocean areas (where applicable)
+    landsea = gpd.read_file(landsea_filename, mask = wbd_buffer)
+    if not landsea.empty:
[…]

### 2021-05-18 src/output_cleanup.py

@@ -1,0 +1,97 @@
+#!/usr/bin/env python3
+import os
+import csv
+import json
+import shutil
+import argparse
+
+def output_cleanup(huc_number, output_folder_path, additional_whitelist, is_production, viz_post_processing):
+    '''
+    Processes all the final output files to cleanup and add post-processing
+
+    Parameters
+    ----------
+    huc_number : STR
+        The HUC
+    output_folder_path : STR
+        Path to the outputs for the specific huc
+    additional_whitelist : STR
+        Additional list of files to keep during a production run
+    is_production : BOOL
+        Determine whether or not to only keep whitelisted production files
+    is_viz_post_processing : BOOL
+        Determine whether or not to process outputs for Viz
+    '''
+
+    # List of files that will be saved during a production run
+    production_whitelist = [
+        'rem_zeroed_masked.tif',
+        'gw_catchments_reaches_filtered_addedAttributes_crosswalked.gpkg',
+        'demDerived_reaches_split_filtered_addedAttributes_crosswalked.gpkg',
+        'gw_catchments_reaches_filtered_addedAttributes.tif',
+        'hydroTable.csv',
+        'src.json',
+        'small_segments.csv',
+        'bathy_crosswalk_calcs.csv',
+        'bathy_stream_order_calcs.csv',
+        'bathy_thalweg_flag.csv',
+        'bathy_xs_area_hydroid_lookup.csv',
+        'src_full_crosswalked.csv',
+        'usgs_elev_table.csv',
+        'hand_ref_elev_table.csv'
+    ]
+
+    # List of files that will be saved during a viz run
+    viz_whitelist = [
+        'rem_zeroed_masked.tif',
+        'gw_catchments_reaches_filtered_addedAttributes_crosswalked.gpkg',
+        'demDerived_reaches_split_filtered_addedAttributes_crosswalked.gpkg',
+        'gw_catchments_reaches_filtered_addedAttributes.tif',
+        'hydroTable.csv',
+        'src.json',
+        'small_segments.csv'
+    ]
+
+    # If "production" run, only keep whitelisted files
+    if is_production and not is_viz_post_processing:
+        whitelist_directory(output_folder_path, production_whitelist, additional_whitelist)
+
+    # If Viz post-processing is enabled, form output files to Viz specifications
+    if is_viz_post_processing:
+        # Step 1, keep only files that Viz needs
+        whitelist_directory(output_folder_path, viz_whitelist, additional_whitelist)
[…]

### 2021-05-18 src/reduce_nhd_stream_density.py

@@ -1,0 +1,204 @@
+#!/usr/bin/env python3
+
+import geopandas as gpd
+import pandas as pd
+import numpy as np
+from collections import deque
+import argparse
+import pygeos
+from shapely.wkb import dumps
+from utils.shared_functions import getDriver
+
+
+def subset_nhd_network(huc4,huc4_mask,selected_wbd8,nhd_streams_,headwaters_filename,headwater_id,nwm_intersections_filename,mainstem_flag=False):
+
+    headwater_streams = pd.DataFrame()
+
+    if mainstem_flag == False:
+        nhd_streams = gpd.read_file(nhd_streams_)
+        headwater_col = 'is_headwater'
+        id_col = 'headwaters_id'
+        n = -1
+    else:
+        nhd_streams = nhd_streams_.copy()
+        headwater_col = 'mainstem'
+        id_col = 'nws_lid'
+        n = ''
+
+    # Locate the closest NHDPlus HR stream segment to NWM headwater points. Done by HUC8 to reduce processing time and to contain NWM headwater in the same HUC
+    for index, row in selected_wbd8.iterrows():
+        huc = row["HUC8"]
+
+        # Double check that this is a nested HUC (probably overkill)
+        if huc.startswith(str(huc4)):
+
+            huc8_mask = selected_wbd8.loc[selected_wbd8.HUC8==huc]
+            huc8_mask = huc8_mask.reset_index(drop=True)
+
+            # Masking headwaters by HUC8
+            headwaters_mask = gpd.read_file(headwaters_filename, mask = huc8_mask)
+            headwaters_mask = headwaters_mask.reset_index(drop=True)
+
+            # Masking subset streams by HUC8
+            if mainstem_flag == False:
+                streams_subset = gpd.read_file(nhd_streams_, mask = huc8_mask)
+            else:
+                streams_subset = nhd_streams.loc[nhd_streams.HUC8==huc].copy()
+
+            if not streams_subset.empty:
+                streams_subset[headwater_col] = False
+                streams_subset = streams_subset.reset_index(drop=True)
+
+                # Create WKB geometry column
+                streams_subset['b_geom'] = None
+                for index, linestring in enumerate(streams_subset.geometry):
+                    streams_subset.at[index, 'b_geom'] = dumps(linestring)
+
+                # Create pygeos nhd stream geometries from WKB representation
+                streambin_geom = pygeos.io.from_wkb(streams_subset['b_geom'])
+
[…]

### 2021-05-18 src/rem.py

@@ -1,0 +1,191 @@
+#!/usr/bin/env python3
+
+from numba import njit, typeof, typed, types
+import rasterio
+import numpy as np
+import argparse
+import os
+import pandas as pd
+from osgeo import ogr, gdal
+import geopandas as gpd
+from utils.shared_functions import getDriver
+
+
+def rel_dem(dem_fileName, pixel_watersheds_fileName, rem_fileName, thalweg_raster, hydroid_fileName, dem_reaches_filename):
+    """
+        Calculates REM/HAND/Detrended DEM
+
+        Parameters
+        ----------
+        dem_fileName : str
+            File name of pit filled DEM raster.
+        pixel_watersheds_fileName : str
+            File name of stream pixel watersheds raster.
+        rem_fileName : str
+            File name of output relative elevation raster.
+        hydroid_fileName : str
+            File name of the hydroid raster (i.e. gw_catchments_reaches.tif)
+        dem_reaches_filename
+            File name of the reaches layer to populate HAND elevation attribute values and overwrite as output
+
+    """
+
+    # ------------------------------------------- Get catchment_hydroid_dict --------------------------------------------------- #
+    # The following creates a dictionary of the catchment ids (key) and their hydroid along the thalweg (value).
+    # This is needed to produce a HAND zero reference elevation by hydroid dataframe (helpful for evaluating rating curves & bathy properties)
+    @njit
+    def make_catchment_hydroid_dict(flat_value_raster, catchment_hydroid_dict, flat_catchments, thalweg_window):
+
+        for i,cm in enumerate(flat_catchments):
+            if thalweg_window[i] == 1:  # Only allow reference hydroid to be within thalweg.
+                catchment_hydroid_dict[cm] = flat_value_raster[i]
+        return(catchment_hydroid_dict)
+
+    # Open the masked gw_catchments_pixels_masked, hydroid_raster, and dem_thalwegCond_masked.
+    gw_catchments_pixels_masked_object = rasterio.open(pixel_watersheds_fileName)
+    hydroid_pixels_object = rasterio.open(hydroid_fileName)
+    thalweg_raster_object = rasterio.open(thalweg_raster)
+
+    # Specify raster object metadata.
+    meta = hydroid_pixels_object.meta.copy()
+    meta['tiled'], meta['compress'] = True, 'lzw'
+
+    # -- Create catchment_hydroid_dict -- #
[…]

### 2021-05-18 src/usgs_gage_crosswalk.py

@@ -1,0 +1,134 @@
+#!/usr/bin/env python3
+
+import os
+import geopandas as gpd
+import pandas as pd
+import numpy as np
+import rasterio
+import argparse
+import pygeos
+from shapely.wkb import dumps, loads
+import warnings
+warnings.simplefilter("ignore")
+
+''' Get elevation at adjusted USGS gages locations
+
+    Parameters
+    ----------
+    usgs_gages_filename : str
+        File name of USGS stations layer.
+    dem_filename : str
+        File name of original DEM.
+    input_flows_filename : str
+        File name of FIM streams layer.
+    input_catchment_filename : str
+        File name of FIM catchment layer.
+    wbd_buffer_filename : str
+        File name of buffered wbd.
+    dem_adj_filename : str
+        File name of thalweg adjusted DEM.
+    output_table_filename : str
+        File name of output table.
+'''
+
+
+def crosswalk_usgs_gage(usgs_gages_filename,dem_filename,input_flows_filename,input_catchment_filename,wbd_buffer_filename,dem_adj_filename,output_table_filename,extent):
+
+    wbd_buffer = gpd.read_file(wbd_buffer_filename)
+    usgs_gages = gpd.read_file(usgs_gages_filename, mask=wbd_buffer)
+    dem_m = rasterio.open(dem_filename,'r')
+    input_flows = gpd.read_file(input_flows_filename)
+    input_catchment = gpd.read_file(input_catchment_filename)
+    dem_adj = rasterio.open(dem_adj_filename,'r')
+
+
+    #MS extent use gages that are mainstem
+    if extent == "MS":
+        usgs_gages = usgs_gages.query('curve == "yes" & mainstem == "yes"')
+    #FR extent use gages that are not mainstem
+    if extent == "FR":
+        usgs_gages = usgs_gages.query('curve == "yes" & mainstem == "no"')
+
+    if input_flows.HydroID.dtype != 'int': input_flows.HydroID = input_flows.HydroID.astype(int)
+
+    # Identify closest HydroID
+    closest_catchment = gpd.sjoin(usgs_gages, input_catchment, how='left', op='within').reset_index(drop=True)
+    closest_hydro_id = closest_catchment.filter(items=['location_id','HydroID','min_thal_elev','med_thal_elev','max_thal_elev', 'order_'])
+    closest_hydro_id = closest_hydro_id.dropna()
+
+    # Get USGS gages that are within catchment boundaries
[…]

### 2021-05-18 src/utils/reproject_dem.py

@@ -1,0 +1,50 @@
+#!/usr/bin/env python3
+
+import os
+from osgeo import gdal
+import sys
+sys.path.append('/foss_fim/src')
+from utils.shared_variables import PREP_PROJECTION_CM
+import shutil
+from multiprocessing import Pool
+import argparse
+
+
+def reproject_dem(args):
+
+    raster_dir            = args[0]
+    elev_cm               = args[1]
+    elev_cm_proj          = args[2]
+    reprojection          = args[3]
+
+    if os.path.exists(elev_cm_proj):
+        os.remove(elev_cm_proj)
+
+    shutil.copy(elev_cm, elev_cm_proj)
+
+    print(f"Reprojecting {elev_cm_proj}")
+    gdal.Warp(elev_cm_proj,elev_cm_proj,dstSRS=reprojection)
+
+
+if __name__ == '__main__':
+
+    parser = argparse.ArgumentParser(description='Burn in NLD elevations')
+    parser.add_argument('-dem_dir','--dem-dir', help='DEM filename', required=True,type=str)
+    parser.add_argument('-j','--number-of-jobs',help='Number of processes to use. Default is 1.',required=False, default="1",type=int)
+
+    args = vars(parser.parse_args())
+
+    dem_dir            = args['dem_dir']
+    number_of_jobs     = args['number_of_jobs']
+
+    reproject_procs_list = []
+
+    for huc in os.listdir(dem_dir):
+        raster_dir = os.path.join(dem_dir,huc)
+        elev_cm = os.path.join(raster_dir, 'elev_cm.tif')
+        elev_cm_proj = os.path.join(raster_dir, 'elev_cm_proj.tif')
+        reproject_procs_list.append([raster_dir, elev_cm, elev_cm_proj, PREP_PROJECTION_CM])
+
+    # Multiprocess reprojection
+    with Pool(processes=number_of_jobs) as pool:
+        pool.map(reproject_dem, reproject_procs_list)

### 2021-05-18 src/utils/shared_functions.py

@@ -1,0 +1,210 @@
+#!/usr/bin/env python3
+
+import os
+from os.path import splitext
+import rasterio
+import numpy as np
+from rasterio.warp import calculate_default_transform, reproject, Resampling
+from pyproj.crs import CRS
+
+def getDriver(fileName):
+
+    driverDictionary = {'.gpkg' : 'GPKG','.geojson' : 'GeoJSON','.shp' : 'ESRI Shapefile'}
+    driver = driverDictionary[splitext(fileName)[1]]
+
+    return(driver)
+
+def pull_file(url, full_pulled_filepath):
+    """
+    This helper function pulls a file and saves it to a specified path.
+
+    Args:
+        url (str): The full URL to the file to download.
+        full_pulled_filepath (str): The full system path where the downloaded file will be saved.
+    """
+    import urllib.request
+
+    print("Pulling " + url)
+    urllib.request.urlretrieve(url, full_pulled_filepath)
+
+
+def delete_file(file_path):
+    """
+    This helper function deletes a file.
+
+    Args:
+        file_path (str): System path to a file to be deleted.
+    """
+
+    try:
+        os.remove(file_path)
+    except FileNotFoundError:
+        pass
+
+
+def run_system_command(args):
+    """
+    This helper function takes a system command and runs it. This function is designed for use
+    in multiprocessing.
+
+    Args:
+        args (list): A single-item list, the first and only item being a system command string.
+    """
+
+    # Parse system command.
+    command = args[0]
+
+    # Run system command.
+    os.system(command)
+
+
+def subset_wbd_gpkg(wbd_gpkg, multilayer_wbd_geopackage):
+
+    import geopandas as gp
+    from utils.shared_variables import CONUS_STATE_LIST, PREP_PROJECTION
+
+    print("Subsetting " + wbd_gpkg + "...")
+    # Read geopackage into dataframe.
+    wbd = gp.read_file(wbd_gpkg)
+    gdf = gp.GeoDataFrame(wbd)
+
+    for index, row in gdf.iterrows():
+        state = row["STATES"]
+        if state != None:  # Some polygons are empty in the STATES field.
+            keep_flag = False  # Default to Fault, i.e. to delete the polygon.
+            if state in CONUS_STATE_LIST:
+                keep_flag = True
+            # Only split if multiple states present. More efficient this way.
+            elif len(state) > 2:
+                for wbd_state in state.split(","):  # Some polygons have multiple states, separated by a comma.
[…]

### 2021-05-18 src/utils/shared_variables.py

@@ -1,0 +1,63 @@
+#!/usr/bin/env python3
+
+import os
+
+# Projections.
+#PREP_PROJECTION = "+proj=aea +datum=NAD83 +x_0=0.0 +y_0=0.0 +lon_0=96dW +lat_0=23dN +lat_1=29d30'N +lat_2=45d30'N +towgs84=-0.9956000824677655,1.901299877314078,0.5215002840524426,0.02591500053005733,0.009425998542707753,0.01159900118427752,-0.00062000005129903 +no_defs +units=m"
+PREP_PROJECTION_CM = 'PROJCS["USA_Contiguous_Albers_Equal_Area_Conic_USGS_version",GEOGCS["GCS_North_American_1983",DATUM["D_North_American_1983",SPHEROID["GRS_1980",6378137.0,298.257222101]],PRIMEM["Greenwich",0.0],UNIT["Degree",0.0174532925199433]],PROJECTION["Albers"],PARAMETER["false_easting",0.0],PARAMETER["false_northing",0.0],PARAMETER["central_meridian",-96.0],PARAMETER["standard_parallel_1",29.5],PARAMETER["standard_parallel_2",45.5],PARAMETER["latitude_of_origin",23.0],UNIT["Meter",1.0],VERTCS["NAVD_1988",VDATUM["North_American_Vertical_Datum_1988"],PARAMETER["Vertical_Shift",0.0],PARAMETER["Direction",1.0],UNIT["Centimeter",0.01]]]'
+PREP_PROJECTION = 'PROJCS["USA_Contiguous_Albers_Equal_Area_Conic_USGS_version",GEOGCS["NAD83",DATUM["North_American_Datum_1983",SPHEROID["GRS 1980",6378137,298.2572221010042,AUTHORITY["EPSG","7019"]],AUTHORITY["EPSG","6269"]],PRIMEM["Greenwich",0],UNIT["degree",0.0174532925199433],AUTHORITY["EPSG","4269"]],PROJECTION["Albers_Conic_Equal_Area"],PARAMETER["standard_parallel_1",29.5],PARAMETER["standard_parallel_2",45.5],PARAMETER["latitude_of_center",23],PARAMETER["longitude_of_center",-96],PARAMETER["false_easting",0],PARAMETER["false_northing",0],UNIT["metre",1,AUTHORITY["EPSG","9001"]]]'
[…]

### 2021-05-18 tests/comparing_src.py

@@ -1,391 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import json
-import geopandas as gpd
-import pandas as pd
-from raster import Raster
-import os
-from shapely.geometry import Point
-
-projectDirectory = os.path.join(os.path.expanduser('~'),'projects','foss_fim')
-dataDirectory = os.path.join(projectDirectory,'data')
-
-# nwm_catchments_fileName = os.path.join(dataDirectory,'nwm','NWMCatchment.shp')
-# nwm_flows_fileName = os.path.join(dataDirectory,'test2','inputs','nwm_flows_proj_120903_v2.gpkg')
-#
-# esri_catchments_fileName = os.path.join(projectDirectory,'tests','CatchmentH.shp')
-esri_flows_fileName = os.path.join(projectDirectory,'tests','eval_1','final_esri_hand_outputs','FPRiver.gpkg')
-
-# foss_catchments_fileName = os.path.join(dataDirectory,'test2','outputs','gw_catchments_reaches_clipped_addedAttributes_crosswalked.gpkg')
-# foss_raster_catchments_fileName = os.path.join(dataDirectory,'test2','outputs','gw_catchments_reaches_clipped_addedAttributes.tif')
-foss_flows_fileName = os.path.join(dataDirectory,'test2','outputs','demDerived_reaches_split_clipped_addedAttributes_crosswalked.gpkg')
-foss_flows_fileName = os.path.join(dataDirectory,'test2','outputs','NHDPlusBurnLineEvent_subset_split_clipped_addedAttributes_crosswalked.gpkg')
-# foss_flows_fileName = os.path.join(dataDirectory,'test2','outputs_v32','demDerived_reaches_split_clipped_addedAttributes_crosswalked.gpkg')
-
-esri_src_fileName = os.path.join(projectDirectory,'tests','eval_1','final_esri_hand_outputs','120903_channel_properties.json')
-foss_src_fileName = os.path.join(dataDirectory,'test2','outputs','src.json')
-# foss_src_fileName = os.path.join(dataDirectory,'test2','outputs_v32','src.json')
-
[…]

### 2021-05-18 tests/inundation.py

@@ -1,557 +1,0 @@
-#!/usr/bin/env python3
-
-import numpy as np
-import pandas as pd
-from numba import njit, typed, types
-from concurrent.futures import ThreadPoolExecutor,as_completed
-from subprocess import run
-from os.path import splitext
-import rasterio
-import fiona
-from shapely.geometry import shape
-from rasterio.mask import mask
-from rasterio.io import DatasetReader,DatasetWriter
-from collections import OrderedDict
-import argparse
-from warnings import warn
-from gdal import BuildVRT
-import geopandas as gpd
-
-
-def inundate(
-             rem,catchments,catchment_poly,hydro_table,forecast,mask_type,hucs=None,hucs_layerName=None,
-             subset_hucs=None,num_workers=1,aggregate=False,inundation_raster=None,inundation_polygon=None,
-             depths=None,out_raster_profile=None,out_vector_profile=None,quiet=False
-            ):
-    """
-
-    Run inundation on FIM >=3.0 outputs at job-level scale or aggregated scale
-
-    Generate depths raster, inundation raster, and inundation polygon from FIM >=3.0 outputs. Can use the FIM 3.0 outputs at native HUC level or the aggregated products. Be sure to pass a HUCs file to process in batch mode if passing aggregated products.
-
-    Parameters
-    ----------
-    rem : str or rasterio.DatasetReader
-        File path to or rasterio dataset reader of Relative Elevation Model raster. Must have the same CRS as catchments raster.
-    catchments : str or rasterio.DatasetReader
-        File path to or rasterio dataset reader of Catchments raster. Must have the same CRS as REM raster
-    hydro_table : str or pandas.DataFrame
-        File path to hydro-table csv or Pandas DataFrame object with correct indices and columns.
-    forecast : str or pandas.DataFrame
-        File path to forecast csv or Pandas DataFrame with correct column names.
-    hucs : str or fiona.Collection, optional
-        Batch mode only. File path or fiona collection of vector polygons in HUC 4,6,or 8's to inundate on. Must have an attribute named as either "HUC4","HUC6", or "HUC8" with the associated values.
-    hucs_layerName : str, optional
-        Batch mode only. Layer name in hucs to use if multi-layer file is passed.
-    subset_hucs : str or list of str, optional
-        Batch mode only. File path to line delimited file, HUC string, or list of HUC strings to further subset hucs file for inundating.
-    num_workers : int, optional
[…]

### 2021-05-18 tests/mannings_calibration_run.sh

@@ -1,123 +1,0 @@
-#!/bin/bash -e
-:
-usage ()
-{
-    echo "Calibrate FIM based on Manning's n values"
-    echo 'Usage : fim_run.sh [REQ: -d <fimdir> -t <huclist> -g <outdir> -n <paramfile>] [OPT: -h]'
-    echo ''
-    echo 'REQUIRED:'
-    echo '  -d/--fimdir    : initial run directory with default mannings values'
-    echo '  -t/--huclist    : huc or list of hucs'
-    echo '  -g/--outdir     : output directory for mannings parameter adjustment files'
-    echo '  -n/--paramfile    : parameter set file'
-    echo ''
-    echo 'OPTIONS:'
-    echo '  -h/--help       : help file'
-    echo '  -o/--overwrite  : overwrite outputs if already exist'
-    echo '  -j/--jobLimit   : max number of concurrent jobs to run. Default 1 job at time. 1 outputs'
-    exit
-}
-
-if [ "$#" -lt 7 ]
-then
-  usage
-fi
-
-while [ "$1" != "" ]; do
-case $1
-in
-    -d|--fimdir)
-        shift
-        fimdir="$1"
-        ;;
-    -t|--huclist )
-        shift
-        huclist=$1
-        ;;
-    -g|--outdir )
-        shift
-        outdir=$1
-        ;;
-    -n|--paramfile)
-        shift
-        paramfile=$1
-        ;;
-    -h|--help)
-        shift
-        usage
-        ;;
-    -o|--overwrite)
-        overwrite=1
-        ;;
-    -j|--jobLimit)
-        shift
-        jobLimit=$1
-        ;;
-    *) ;;
-    esac
-    shift
-done
-
-# print usage if arguments empty
-if [ "$fimdir" = "" ]
-then
-    usage
-fi
-
-# default values
-if [ "$jobLimit" = "" ] ; then
-    jobLimit=1
-fi
-
-export input_NWM_Catchments=$inputDataDir/nwm_hydrofabric/nwm_catchments.gpkg
-export outdir=$outdir
-export testdir="/foss_fim/tests"
-
-if [ -f "$huclist" ]; then
-
-  while read huc; do
-
-    export huc=$huc
-    export fimdir=$fimdir
-    export hucdir="/data/outputs/"$fimdir/$huc
-
-    ## RUN ##
-    if [ -f "$paramfile" ]; then
-        if [ "$jobLimit" -eq 1 ]; then
[…]

### 2021-05-18 tests/preprocess/preprocess_benchmark.py

@@ -1,90 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Jul 23 15:17:04 2020
-
-@author: trevor.grout
-"""
-
-import rasterio
-from rasterio.warp import calculate_default_transform, reproject, Resampling
-import rasterio.mask
-import numpy as np
-import argparse
-
-def preprocess_benchmark_static(benchmark_raster, reference_raster, out_raster_path = None):
-    '''
-    This function will preprocess a benchmark dataset for purposes of evaluating FIM output. A benchmark dataset will be transformed using properties (CRS, resolution) from an input reference dataset. The benchmark raster will also be converted to a boolean (True/False) raster with inundated areas (True or 1) and dry areas (False or 0). 
-
-    Parameters
-    ----------
-    benchmark_raster : STRING
-        Path to benchmark raster (e.g. BLE elevation or WSE elevations).
-    reference_raster : STRING
-        Path to reference raster (e.g. output from inundation.py).
-    out_raster_path : STRING, optional
-        Path to the output raster. The default is None.
-
-    Returns
-    -------
-    boolean_benchmark : Numpy Array
-        The preprocessed benchmark array.
-    profile : STRING
-        Raster profile information for the preprocessed benchmark array (required for writing to output dataset).
-
-    '''
-    #Open and read raster and benchmark rasters
-    reference = rasterio.open(reference_raster)
-    benchmark = rasterio.open(benchmark_raster)
-    benchmark_arr = benchmark.read(1)    
-
-    #Set arbitrary no data value that is not possible value of the benchmark dataset. This will be reassigned later.
-    nodata_value = -2147483648
-    
-    #Determine the new transform and dimensions of reprojected/resampled raster.
-    new_transform, new_width, new_height = calculate_default_transform(benchmark.crs, reference.crs, benchmark.width, benchmark.height, *benchmark.bounds, resolution = reference.res)
-
-    #Define an empty array that is same dimensions as output by the "calculate_default_transform" command. 
-    benchmark_projected = np.empty((new_height,new_width), dtype=np.int32)
-
-    #Reproject and resample the benchmark dataset. Bilinear resampling due to continuous depth data.
-    reproject(benchmark_arr, 
-              destination = benchmark_projected,
-              src_transform = benchmark.transform, 
-              src_crs = benchmark.crs,
-              src_nodata = benchmark.nodata,
-              dst_transform = new_transform, 
[…]

### 2021-05-18 tests/preprocess/preprocess_fimx.py

@@ -1,129 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Fri Jul 24 13:50:59 2020
-
-@author: trevor.grout
-"""
-import rasterio
-from rasterio.warp import calculate_default_transform, reproject, Resampling
-from rasterio import features
-import numpy as np
-import geopandas as gpd
-import os
-import argparse
-
-def fimx_to_fim3(catchments_path, raster_value_field, hand_raster_path, template_raster, out_hand_path = None, out_catchment_path = None):
-    '''
-    This function will produce a hand and catchment grid from fim1/fim2 for use in the fim3 inundation.py script. To accomplish this it:
-        1) reprojects the hand raster to the template raster crs.
-        2) It reprojects the catchment polygons to the template raster crs and then converts the polygons to a raster. The reprojected HAND raster properties are applied (extent, resolution)
-        3) Performs the intersection of the two rasters (HAND/catchments) and applies NODATA if either dataset has NODATA
-        4) Writes out the preprocessed HAND and Catchments raster if a path is specified.
-
-    Parameters
-    ----------
-    catchments_path : STRING
-        The path to the catchments vector data (assumes it is in a database).
-    raster_value_field : STRING
-        Attribute field in catchments layer whos values will be used for raster creation.
-    hand_raster_path : STRING
-        The path to the HAND raster dataset (ARC GRID is acceptable).
-    template_raster : STRING
-        Path to the template raster so that a CRS can be applied to output.
-    out_hand_path : STRING, optional
-        Path to the output HAND grid (Gtif format). The default is None.
-    out_catchment_path : STRING, optional
-        Path to the output catchment grid (Gtif format). The default is None.
-
-    Returns
-    -------
-    hand_masked : Numpy Array
-        Preprocessed HAND raster array.
-    hand_profile : LIST
-        Preprocessed HAND profile.
-    catchment_masked : Numpy Array
-        Preprocessed Catchment raster array.
-    catchment_profile : LIST
-        Preprocessed catchment raster profile.
-
-    '''
-    
-    
-    #Read in template raster as band object.
-    reference = rasterio.open(template_raster)
-    
-    #Step 1: Convert HAND grid
-    #Read in the hand raster     
-    hand = rasterio.open(hand_raster_path)
-    hand_arr = hand.read(1)
[…]

### 2021-05-18 tools/aggregate_mannings_calibration.py

@@ -1,0 +1,116 @@
+#!/usr/bin/env python3
+
+import os
+import pandas as pd
+import csv
+import json
+import argparse
+
+
+def aggregate_parameter_sets(huc_list_path,calibration_stat_folder,summary_file, mannings_json):
+
+    outfolder = os.path.dirname(summary_file)
+    aggregate_output_dir = os.path.join(outfolder, 'aggregate_metrics')
+
+    if not os.path.exists(aggregate_output_dir):
+        os.makedirs(aggregate_output_dir)
+
+    mannings_summary_table = pd.DataFrame(columns = ['metric', 'value', 'stream_order', 'mannings_n', 'huc', 'interval'])
+
+    with open(huc_list_path) as f:
+        huc_list = [huc.rstrip() for huc in f]
+
+    for huc in huc_list:
+        branch_dir = os.path.join('data','test_cases',str(huc) + '_ble','performance_archive','development_versions', calibration_stat_folder)
+        for stream_order in os.listdir(branch_dir):
+            stream_order_dir = os.path.join(branch_dir, stream_order)
+            for mannings_value in os.listdir(stream_order_dir):
+                mannings_value_dir = os.path.join(stream_order_dir, mannings_value)
+                for flood_recurrence in os.listdir(mannings_value_dir):
+                    flood_recurrence_dir = os.path.join(mannings_value_dir, flood_recurrence)
+                    total_area_stats = pd.read_csv(os.path.join(flood_recurrence_dir, 'total_area_stats.csv'), index_col=0)
+                    total_area_stats = total_area_stats.loc[['true_positives_count', 'true_negatives_count', 'false_positives_count', 'false_negatives_count','masked_count', 'cell_area_m2', 'CSI'],:]
+                    total_area_stats = total_area_stats.reset_index()
+                    total_area_stats_table = pd.DataFrame({'metric': total_area_stats.iloc[:,0], 'value': total_area_stats.iloc[:,1], 'stream_order': stream_order, 'mannings_n': mannings_value, 'huc': huc, 'interval': flood_recurrence})
+                    mannings_summary_table = mannings_summary_table.append(total_area_stats_table, ignore_index=True)
+
+    mannings_summary_table.to_csv(summary_file,index=False)
+
+    ## calculate optimal parameter set
+    from utils.shared_functions import compute_stats_from_contingency_table
+
+    true_positives, true_negatives, false_positives, false_negatives, cell_area, masked_count = 0, 0, 0, 0, 0, 0
+
[…]

### 2021-05-18 tools/cache_metrics.py

@@ -1,0 +1,118 @@
+#!/usr/bin/env python3
+
+import os
+import argparse
+import traceback
+
+from run_test_case import run_alpha_test
+from multiprocessing import Pool
+
+TEST_CASES_DIR = r'/data/test_cases_new/'  #TODO remove "_new"
+PREVIOUS_FIM_DIR = r'/data/previous_fim'
+OUTPUTS_DIR = r'/data/outputs'
+
+
+def process_alpha_test(args):
+
+    fim_run_dir = args[0]
+    version = args[1]
+    test_id = args[2]
+    magnitude = args[3]
+    archive_results = args[4]
+
+    mask_type = 'huc'
+
+    if archive_results == False:
+        compare_to_previous = True
+    else:
+        compare_to_previous = False
+
+    try:
+        run_alpha_test(fim_run_dir, version, test_id, magnitude, compare_to_previous=compare_to_previous, archive_results=archive_results, mask_type=mask_type)
+    except Exception:
+        traceback.print_exc()
+
+
+if __name__ == '__main__':
+
+    # Parse arguments.
+    parser = argparse.ArgumentParser(description='Caches metrics from previous versions of HAND.')
+    parser.add_argument('-c','--config',help='Save outputs to development_versions or previous_versions? Options: "DEV" or "PREV"',required=True)
+    parser.add_argument('-v','--fim-version',help='Name of fim version to cache.',required=False, default="all")
+    parser.add_argument('-j','--job-number',help='Number of processes to use. Default is 1.',required=False, default="1")
+    parser.add_argument('-s','--special-string',help='Add a special name to the end of the branch.',required=False, default="")
+    parser.add_argument('-b','--benchmark-category',help='Options include ble or ahps. Defaults to process both.',required=False, default=None)
+
+    test_cases_dir_list = os.listdir(TEST_CASES_DIR)
+
+    args = vars(parser.parse_args())
+
+    config = args['config']
+    fim_version = args['fim_version']
+    job_number = int(args['job_number'])
+    special_string = args['special_string']
+    benchmark_category = args['benchmark_category']
+
+    if fim_version != "all":
+        previous_fim_list = [fim_version]
+    else:
+        previous_fim_list = os.listdir(PREVIOUS_FIM_DIR)
+
+    if config == 'PREV':
+        archive_results = True
+    elif config == 'DEV':
+        archive_results = False
+    else:
[…]

### 2021-05-18 tools/eval_plots.py

@@ -1,0 +1,686 @@
+#!/usr/bin/env python3
+import pandas as pd
+from pathlib import Path
+import argparse
+from natsort import natsorted
+import geopandas as gpd
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+import re
+import os
+import sys
+sys.path.append('/foss_fim/src')
+from utils.shared_variables import VIZ_PROJECTION
+from dotenv import load_dotenv
+from tools_shared_functions import aggregate_wbd_hucs, get_metadata
+
+#Get variables from .env file.
+load_dotenv()
+WBD_LAYER = os.getenv("WBD_LAYER")
+API_BASE_URL = os.getenv("API_BASE_URL")
+
+#########################################################################
+#Create boxplot
+#########################################################################
+def boxplot(dataframe, x_field, x_order, y_field, hue_field, ordered_hue, title_text, fim_configuration, textbox_str = False, simplify_legend = False, dest_file = False):
+    '''
+    Create boxplots.
+
+    Parameters
+    ----------
+    dataframe : DataFrame
+        Pandas dataframe data to be plotted.
+    x_field : STR
+        Field to use for x-axis
+    x_order : List
+        Order to arrange the x-axis.
+    y_field : STR
+        Field to use for the y-axis
+    hue_field : STR
+        Field to use for hue (typically FIM version)
+    title_text : STR
+        Text for plot title.
+    fim_configuration: STR
+        Configuration of FIM (FR or MS or Composite).
+    simplify_legend : BOOL, optional
+        If True, it will simplify legend to FIM 1, FIM 2, FIM 3.
+        The default is False.
+    dest_file : STR or BOOL, optional
+        If STR provide the full path to the figure to be saved. If False
+        no plot is saved to disk. The default is False.
+
+    Returns
+    -------
+    fig : MATPLOTLIB
+        Plot.
+
+    '''
+
+    #initialize plot
+    fig, axes = plt.subplots(nrows=1, ncols=1, figsize=(15, 10))
+    #Use seaborn to plot the boxplot
+    axes=sns.boxplot(x=x_field, y=y_field, order=x_order, hue=hue_field, hue_order = ordered_hue, data=dataframe, palette='bright')
+    #set title of plot
+    axes.set_title(f'{title_text} ({y_field})',fontsize=20, weight = 'bold')
+    #Set yticks and background horizontal line.
+    axes.set(ylim=(0.0,1.0),yticks = np.arange(0,1.1,0.1))
[…]

### 2021-05-18 tools/generate_categorical_fim.py

@@ -1,155 +1,108 @@
+#!/usr/bin/env python3
+
 import os
-from multiprocessing import Pool
+import subprocess
 import argparse
-import traceback
-import sys
+import time
+from pathlib import Path
+import geopandas as gpd
+import pandas as pd
+from datetime import date
 
-sys.path.insert(1, 'foss_fim/tests')
-from inundation import inundate
+def update_mapping_status(output_mapping_dir, output_flows_dir):
+    '''
+    Updates the status for nws_lids from the flows subdirectory. Status
+    is updated for sites where the inundation.py routine was not able to
+    produce inundation for the supplied flow files. It is assumed that if
+    an error occured in inundation.py that all flow files for a given site
+    experienced the error as they all would have the same nwm segments.
 
-INPUTS_DIR = r'/data/inputs'
+    Parameters
+    ----------
+    output_mapping_dir : STR
+        Path to the output directory of all inundation maps.
+    output_flows_dir : STR
+        Path to the directory containing all flows.
 
-# Define necessary variables for inundation().
-hucs, hucs_layerName = os.path.join(INPUTS_DIR, 'wbd', 'WBD_National.gpkg'), 'WBDHU8'
-mask_type, catchment_poly = 'huc', ''
-    
+    Returns
+    -------
+    None.
 
-def generate_categorical_fim(fim_run_dir, source_flow_dir, output_cat_fim_dir, job_number, gpkg, extif, depthtif):
-    
-    # Create output directory and log directory.
-    if not os.path.exists(output_cat_fim_dir):
-        os.mkdir(output_cat_fim_dir)
-    log_dir = os.path.join(output_cat_fim_dir, 'logs')
-    if not os.path.exists(log_dir):
-        os.mkdir(log_dir)
-    
-    no_data_list = []
-    procs_list = []
-       
-    # Loop through huc directories in the source_flow directory.
-    source_flow_dir_list = os.listdir(source_flow_dir)
-    for huc in source_flow_dir_list:
-        if "." not in huc:
-            
-            # Get list of AHPS site directories.
-            ahps_site_dir = os.path.join(source_flow_dir, huc)
-            ahps_site_dir_list = os.listdir(ahps_site_dir)
-            
-            # Map paths to HAND files needed for inundation().
-            fim_run_huc_dir = os.path.join(fim_run_dir, huc)
-            rem = os.path.join(fim_run_huc_dir, 'rem_zeroed_masked.tif')
[…]

### 2021-05-18 tools/inundation.py

@@ -1,0 +1,570 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import pandas as pd
+from numba import njit, typed, types
+from concurrent.futures import ThreadPoolExecutor,as_completed
+from subprocess import run
+from os.path import splitext
+import rasterio
+import fiona
+from shapely.geometry import shape
+from rasterio.mask import mask
+from rasterio.io import DatasetReader,DatasetWriter
+from rasterio.features import shapes
+from collections import OrderedDict
+import argparse
+from warnings import warn
+from gdal import BuildVRT
+import geopandas as gpd
+import sys
+
+
+def inundate(
+             rem,catchments,catchment_poly,hydro_table,forecast,mask_type,hucs=None,hucs_layerName=None,
+             subset_hucs=None,num_workers=1,aggregate=False,inundation_raster=None,inundation_polygon=None,
+             depths=None,out_raster_profile=None,out_vector_profile=None,quiet=False
+            ):
+    """
+
+    Run inundation on FIM >=3.0 outputs at job-level scale or aggregated scale
+
+    Generate depths raster, inundation raster, and inundation polygon from FIM >=3.0 outputs. Can use the FIM 3.0 outputs at native HUC level or the aggregated products. Be sure to pass a HUCs file to process in batch mode if passing aggregated products.
+
+    Parameters
+    ----------
+    rem : str or rasterio.DatasetReader
+        File path to or rasterio dataset reader of Relative Elevation Model raster. Must have the same CRS as catchments raster.
+    catchments : str or rasterio.DatasetReader
+        File path to or rasterio dataset reader of Catchments raster. Must have the same CRS as REM raster
+    hydro_table : str or pandas.DataFrame
+        File path to hydro-table csv or Pandas DataFrame object with correct indices and columns.
+    forecast : str or pandas.DataFrame
+        File path to forecast csv or Pandas DataFrame with correct column names.
+    hucs : str or fiona.Collection, optional
+        Batch mode only. File path or fiona collection of vector polygons in HUC 4,6,or 8's to inundate on. Must have an attribute named as either "HUC4","HUC6", or "HUC8" with the associated values.
+    hucs_layerName : str, optional
+        Batch mode only. Layer name in hucs to use if multi-layer file is passed.
+    subset_hucs : str or list of str, optional
+        Batch mode only. File path to line delimited file, HUC string, or list of HUC strings to further subset hucs file for inundating.
[…]

### 2021-05-18 tools/mannings_calibration_run.sh

@@ -1,0 +1,123 @@
+#!/bin/bash -e
+:
+usage ()
+{
+    echo "Calibrate FIM based on Manning's n values"
+    echo 'Usage : fim_run.sh [REQ: -d <fimdir> -t <huclist> -g <outdir> -n <paramfile>] [OPT: -h]'
+    echo ''
+    echo 'REQUIRED:'
+    echo '  -d/--fimdir    : initial run directory with default mannings values'
+    echo '  -t/--huclist    : huc or list of hucs'
+    echo '  -g/--outdir     : output directory for mannings parameter adjustment files'
+    echo '  -n/--paramfile    : parameter set file'
+    echo ''
+    echo 'OPTIONS:'
+    echo '  -h/--help       : help file'
+    echo '  -o/--overwrite  : overwrite outputs if already exist'
+    echo '  -j/--jobLimit   : max number of concurrent jobs to run. Default 1 job at time. 1 outputs'
+    exit
+}
+
+if [ "$#" -lt 7 ]
+then
+  usage
+fi
+
+while [ "$1" != "" ]; do
+case $1
+in
+    -d|--fimdir)
+        shift
+        fimdir="$1"
+        ;;
+    -t|--huclist )
+        shift
+        huclist=$1
+        ;;
+    -g|--outdir )
+        shift
+        outdir=$1
+        ;;
+    -n|--paramfile)
+        shift
+        paramfile=$1
+        ;;
+    -h|--help)
+        shift
+        usage
+        ;;
+    -o|--overwrite)
+        overwrite=1
+        ;;
+    -j|--jobLimit)
+        shift
+        jobLimit=$1
+        ;;
+    *) ;;
+    esac
+    shift
+done
+
+# print usage if arguments empty
+if [ "$fimdir" = "" ]
+then
+    usage
+fi
+
+# default values
+if [ "$jobLimit" = "" ] ; then
+    jobLimit=1
+fi
+
+export input_NWM_Catchments=$inputDataDir/nwm_hydrofabric/nwm_catchments.gpkg
+export outdir=$outdir
+export toolsdir="/foss_fim/tools"
+
+if [ -f "$huclist" ]; then
+
+  while read huc; do
+
+    export huc=$huc
+    export fimdir=$fimdir
+    export hucdir="/data/outputs/"$fimdir/$huc
+
+    ## RUN ##
+    if [ -f "$paramfile" ]; then
+        if [ "$jobLimit" -eq 1 ]; then
[…]

### 2021-05-18 tools/mannings_run_by_set.sh

@@ -1,0 +1,15 @@
+#!/bin/bash -e
+
+param_set="$1"
+IFS=',' read -r -a array <<< $param_set
+
+strorder="${array[1]}"
+mannings_row=1+"$strorder"
+mannings_value="${array[$mannings_row]}"
+
+subdir=$outdir/$huc"_"$strorder"_"$mannings_value
+mkdir -p $subdir
+
+$srcDir/add_crosswalk.py -d $hucdir/gw_catchments_reaches_filtered_addedAttributes.gpkg -a $hucdir/demDerived_reaches_split_filtered.gpkg -s $hucdir/src_base.csv -l $subdir/gw_catchments_reaches_filtered_addedAttributes_crosswalked.gpkg -f $subdir/demDerived_reaches_split_filtered_addedAttributes_crosswalked.gpkg -r $subdir/src_full_crosswalked.csv -j $subdir/src.json -x $subdir/crosswalk_table.csv -t $subdir/hydroTable.csv -w $hucdir/wbd8_clp.gpkg -b $hucdir/nwm_subset_streams.gpkg -y $hucdir/nwm_catchments_proj_subset.tif -m $param_set -z $input_NWM_Catchments -p FR -c
+
+python3 foss_fim/tools/run_test_case_calibration.py -r $fimdir/$huc -d $subdir -t $huc"_ble" -b "mannings_calibration"/$strorder/$mannings_value

### 2021-05-18 tools/rating_curve_comparison.py

@@ -1,0 +1,443 @@
+#!/usr/bin/env python3
+
+import os
+import sys
+import geopandas as gpd
+import pandas as pd
+import numpy as np
+import argparse
+import matplotlib.pyplot as plt
+import seaborn as sns
+from functools import reduce
+from multiprocessing import Pool
+from os.path import isfile, join, dirname
+import shutil
+import warnings
+from pathlib import Path
+import time
+warnings.simplefilter(action='ignore', category=FutureWarning)
+
+"""
+    Plot Rating Curves and Compare to USGS Gages
+
+    Parameters
+    ----------
+    fim_dir : str
+        Directory containing FIM output folders.
+    output_dir : str
+        Directory containing rating curve plots and tables.
+    usgs_gages_filename : str
+        File name of USGS rating curves.
+    nwm_flow_dir : str
+        Directory containing NWM recurrence flows files.
+    number_of_jobs : str
+        Number of jobs.
+    stat_groups : str
+        string of columns to group eval metrics.
+"""
+
+def check_file_age(file):
+    '''
+    Checks if file exists, determines the file age, and recommends
+    updating if older than 1 month.
+
+    Returns
+    -------
+    None.
+
+    '''
+    file = Path(file)
+    if file.is_file():
+        modification_time = file.stat().st_mtime
+        current_time = time.time()
+        file_age_days = (current_time - modification_time)/86400
+        if file_age_days > 30:
+            check = f'{file.name} is {int(file_age_days)} days old, consider updating.\nUpdate with rating_curve_get_usgs_curves.py'
+        else:
+            check = f'{file.name} is {int(file_age_days)} days old.'
+
+        return check
+
+# recurr_intervals = ['recurr_1_5_cms.csv','recurr_5_0_cms.csv','recurr_10_0_cms.csv']
+
+def generate_rating_curve_metrics(args):
+
+    elev_table_filename         = args[0]
+    hydrotable_filename         = args[1]
+    usgs_gages_filename         = args[2]
+    usgs_recurr_stats_filename  = args[3]
+    nwm_recurr_data_filename    = args[4]
+    rc_comparison_plot_filename = args[5]
+    nwm_flow_dir                = args[6]
+    catfim_flows_filename       = args[7]
[…]

### 2021-05-18 tools/rating_curve_get_usgs_curves.py

@@ -1,0 +1,370 @@
+#!/usr/bin/env python3
+import time
+import pandas as pd
+import geopandas as gpd
+from pathlib import Path
+from tools_shared_functions import get_metadata, get_datum, ngvd_to_navd_ft, get_rating_curve, aggregate_wbd_hucs, get_thresholds, flow_data
+from dotenv import load_dotenv
+import os
+import argparse
+import sys
+sys.path.append('/foss_fim/src')
+from utils.shared_variables import PREP_PROJECTION,VIZ_PROJECTION
+
+'''
+This script calls the NOAA Tidal API for datum conversions. Experience shows that
+running script outside of business hours seems to be most consistent way
+to avoid API errors. Currently configured to get rating curve data within
+CONUS. Tidal API call may need to be modified to get datum conversions for 
+AK, HI, PR/VI.
+'''
+
+#import variables from .env file
+load_dotenv()
+API_BASE_URL = os.getenv("API_BASE_URL")
+WBD_LAYER = os.getenv("WBD_LAYER")
+EVALUATED_SITES_CSV = os.getenv("EVALUATED_SITES_CSV")
+NWM_FLOWS_MS = os.getenv("NWM_FLOWS_MS")
+
+def get_all_active_usgs_sites():
+    '''
+    Compile a list of all active usgs gage sites that meet certain criteria. 
+    Return a GeoDataFrame of all sites.
+
+    Returns
+    -------
+    None.
+
+    '''
+    #Get metadata for all usgs_site_codes that are active in the U.S.
+    metadata_url = f'{API_BASE_URL}/metadata' 
+    #Define arguments to retrieve metadata and then get metadata from WRDS
+    select_by = 'usgs_site_code'
+    selector = ['all']
+    must_include = 'usgs_data.active'
+    metadata_list, metadata_df = get_metadata(metadata_url, select_by, selector, must_include = must_include, upstream_trace_distance = None, downstream_trace_distance = None )
+
+    #Filter out sites based quality of site. These acceptable codes were initially
+    #decided upon and may need fine tuning. A link where more information
+    #regarding the USGS attributes is provided. 
+    
+    #https://help.waterdata.usgs.gov/code/coord_acy_cd_query?fmt=html
+    acceptable_coord_acc_code = ['H','1','5','S','R','B','C','D','E']
+    #https://help.waterdata.usgs.gov/code/coord_meth_cd_query?fmt=html
[…]

### 2021-05-18 tools/run_test_case.py

@@ -1,0 +1,248 @@
+#!/usr/bin/env python3
+
+import os
+import sys
+import shutil
+import argparse
+
+from tools_shared_functions import compute_contingency_stats_from_rasters
+from tools_shared_variables import (TEST_CASES_DIR, INPUTS_DIR, ENDC, TRED_BOLD, WHITE_BOLD, CYAN_BOLD, AHPS_BENCHMARK_CATEGORIES)
+from inundation import inundate
+
+
+def run_alpha_test(fim_run_dir, version, test_id, magnitude, compare_to_previous=False, archive_results=False, mask_type='huc', inclusion_area='', inclusion_area_buffer=0, light_run=False, overwrite=True):
+
+    benchmark_category = test_id.split('_')[1] # Parse benchmark_category from test_id.
+    current_huc = test_id.split('_')[0]  # Break off HUC ID and assign to variable.
+
+    # Construct paths to development test results if not existent.
+    if archive_results:
+        version_test_case_dir_parent = os.path.join(TEST_CASES_DIR, benchmark_category + '_test_cases', test_id, 'official_versions', version)
+    else:
+        version_test_case_dir_parent = os.path.join(TEST_CASES_DIR, benchmark_category + '_test_cases', test_id, 'testing_versions', version)
+
+    # Delete the entire directory if it already exists.
+    if os.path.exists(version_test_case_dir_parent):
+        if overwrite == True:
+            shutil.rmtree(version_test_case_dir_parent)
+        else:
+            print("Metrics for ({version}: {test_id}) already exist. Use overwrite flag (-o) to overwrite metrics.".format(version=version, test_id=test_id))
+            return
+
+    os.mkdir(version_test_case_dir_parent)
+
+    print("Running the alpha test for test_id: " + test_id + ", " + version + "...")
+    stats_modes_list = ['total_area']
+
+    fim_run_parent = os.path.join(os.environ['outputDataDir'], fim_run_dir)
+    assert os.path.exists(fim_run_parent), "Cannot locate " + fim_run_parent
+
+    # Create paths to fim_run outputs for use in inundate().
+    rem = os.path.join(fim_run_parent, 'rem_zeroed_masked.tif')
+    if not os.path.exists(rem):
+        rem = os.path.join(fim_run_parent, 'rem_clipped_zeroed_masked.tif')
+    catchments = os.path.join(fim_run_parent, 'gw_catchments_reaches_filtered_addedAttributes.tif')
[…]

### 2021-05-18 tools/run_test_case_calibration.py

@@ -1,0 +1,456 @@
+#!/usr/bin/env python3
+
+import os
+import sys
+import pandas as pd
+import rasterio
+import json
+import csv
+import argparse
+import shutil
+
+from tools_shared_functions import get_contingency_table_from_binary_rasters, compute_stats_from_contingency_table
+from inundation import inundate
+
+TEST_CASES_DIR = r'/data/test_cases/'  # Will update.
+INPUTS_DIR = r'/data/inputs'
+PRINTWORTHY_STATS = ['CSI', 'TPR', 'TNR', 'FAR', 'MCC', 'TP_area_km2', 'FP_area_km2', 'TN_area_km2', 'FN_area_km2', 'contingency_tot_area_km2', 'TP_perc', 'FP_perc', 'TN_perc', 'FN_perc']
+GO_UP_STATS = ['CSI', 'TPR', 'MCC', 'TN_area_km2', 'TP_area_km2', 'TN_perc', 'TP_perc', 'TNR']
+GO_DOWN_STATS = ['FAR', 'FN_area_km2', 'FP_area_km2', 'FP_perc', 'FN_perc']
+OUTPUTS_DIR = os.environ['outputDataDir']
+
+ENDC = '\033[m'
+TGREEN_BOLD = '\033[32;1m'
+TGREEN = '\033[32m'
+TRED_BOLD = '\033[31;1m'
+TWHITE = '\033[37m'
+WHITE_BOLD = '\033[37;1m'
+CYAN_BOLD = '\033[36;1m'
+
+
+def profile_test_case_archive(archive_to_check, return_interval, stats_mode):
+    """
+    This function searches multiple directories and locates previously produced performance statistics.
+
+    Args:
+        archive_to_check (str): The directory path to search.
+        return_interval (str): Because a benchmark dataset may have multiple return intervals, this argument defines
+                               which return interval is to be used when searching for previous statistics.
+    Returns:
+        archive_dictionary (dict): A dictionary of available statistics for previous versions of the domain and return interval.
+                                  {version: {agreement_raster: agreement_raster_path, stats_csv: stats_csv_path, stats_json: stats_json_path}}
+                                  *Will only add the paths to files that exist.
+
+    """
+
+    archive_dictionary = {}
+
+    # List through previous version and check for available stats and maps. If available, add to dictionary.
+    available_versions_list = os.listdir(archive_to_check)
+
+    if len(available_versions_list) == 0:
[…]

### 2021-05-18 tools/shapefile_to_raster.py

@@ -1,0 +1,41 @@
+#!/usr/bin/env python3
+
+# A script to rasterise a shapefile to the same projection & pixel resolution as a reference image.
+from osgeo import ogr, gdal
+import subprocess
+
+InputVector = r'/data/misc/msft_building_footprints/MSFP_12090301.shp'
+OutputImage = r'/data/misc/msft_building_footprints/Result.tif'
+
+RefImage = r'/data/test_cases/12090301_ble/validation_data/100yr/ble_huc_12090301_inundation_extent_100yr.tif'
+
+gdalformat = 'GTiff'
+datatype = gdal.GDT_Byte
+burnVal = 1 #value for the output image pixels
+# Get projection info from reference image
+Image = gdal.Open(RefImage, gdal.GA_ReadOnly)
+
+# Open Shapefile
+Shapefile = ogr.Open(InputVector)
+Shapefile_layer = Shapefile.GetLayer()
+
+# Rasterise
+print("Rasterising shapefile...")
+Output = gdal.GetDriverByName(gdalformat).Create(OutputImage, Image.RasterXSize, Image.RasterYSize, 1, datatype, options=['COMPRESS=DEFLATE'])
+Output.SetProjection(Image.GetProjectionRef())
+Output.SetGeoTransform(Image.GetGeoTransform())
+
+# Write data to band 1
+Band = Output.GetRasterBand(1)
+Band.SetNoDataValue(0)
+gdal.RasterizeLayer(Output, [1], Shapefile_layer, burn_values=[burnVal])
+
+# Close datasets
+Band = None
+Output = None
+Image = None
+Shapefile = None
+
+# Build image overviews
+subprocess.call("gdaladdo --config COMPRESS_OVERVIEW DEFLATE "+OutputImage+" 2 4 8 16 32 64", shell=True)
+print("Done.")

### 2021-05-18 tools/synthesize_test_cases.py

@@ -1,0 +1,334 @@
+#!/usr/bin/env python3
+
+import os
+import argparse
+from multiprocessing import Pool
+import json
+import csv
+
+from run_test_case import run_alpha_test
+from tools_shared_variables import TEST_CASES_DIR, PREVIOUS_FIM_DIR, OUTPUTS_DIR, AHPS_BENCHMARK_CATEGORIES
+
+
+def create_master_metrics_csv(master_metrics_csv_output, dev_comparison):
+
+    # Construct header
+    metrics_to_write = ['true_negatives_count',
+                        'false_negatives_count',
+                        'true_positives_count',
+                        'false_positives_count',
+                        'contingency_tot_count',
+                        'cell_area_m2',
+                        'TP_area_km2',
+                        'FP_area_km2',
+                        'TN_area_km2',
+                        'FN_area_km2',
+                        'contingency_tot_area_km2',
+                        'predPositive_area_km2',
+                        'predNegative_area_km2',
+                        'obsPositive_area_km2',
+                        'obsNegative_area_km2',
+                        'positiveDiff_area_km2',
+                        'CSI',
+                        'FAR',
+                        'TPR',
+                        'TNR',
+                        'PPV',
+                        'NPV',
+                        'ACC',
+                        'Bal_ACC',
+                        'MCC',
+                        'EQUITABLE_THREAT_SCORE',
+                        'PREVALENCE',
+                        'BIAS',
+                        'F1_SCORE',
+                        'TP_perc',
+                        'FP_perc',
+                        'TN_perc',
+                        'FN_perc',
+                        'predPositive_perc',
+                        'predNegative_perc',
+                        'obsPositive_perc',
+                        'obsNegative_perc',
+                        'positiveDiff_perc',
+                        'masked_count',
+                        'masked_perc',
+                        'masked_area_km2'
+                        ]
+
+    additional_header_info_prefix = ['version', 'nws_lid', 'magnitude', 'huc']
+    list_to_write = [additional_header_info_prefix + metrics_to_write + ['full_json_path'] + ['flow'] + ['benchmark_source'] + ['extent_config'] + ["calibrated"]]
+
+    versions_to_aggregate = os.listdir(PREVIOUS_FIM_DIR)
+
+    if dev_comparison != None:
+        iteration_list = ['official', 'comparison']
+    if dev_comparison == None:
+        iteration_list = ['official']
+
[…]

### 2021-05-31 tools/eval_plots.py
M&M and eval_plots updates

@@ -69,7 +69,7 @@ def boxplot(dataframe, x_field, x_order, y_field, hue_field, ordered_hue, title_
         plt.axhline(y=ytick,color='black',linestyle = '--',linewidth = 1,alpha = 0.1)
     #Define y axis label and x axis label.
     axes.set_ylabel(f'{y_field}',fontsize='xx-large',weight = 'bold')
-    axes.set_xlabel('',fontsize=0,weight = 'bold')
+    axes.set_xlabel('Flood Magnitude',fontsize=0,weight = 'bold')
     #Set sizes of ticks and legend.
     axes.tick_params(labelsize = 'xx-large')
     axes.legend(markerscale = 2, fontsize =20, loc = 'lower left')
@@ -237,7 +237,7 @@ def barplot(dataframe, x_field, x_order, y_field, hue_field, ordered_hue, title_
         plt.axhline(y=ytick,color='black',linestyle = '--',linewidth = 1,alpha = 0.1)
     #Define y axis label and x axis label.
     axes.set_ylabel(f'{y_field.upper()}',fontsize='xx-large',weight = 'bold')
-    axes.set_xlabel('',fontsize=0,weight = 'bold')
+    axes.set_xlabel('Flood Magnitude',fontsize=0,weight = 'bold')
     #Set sizes of ticks and legend.
     axes.tick_params(labelsize = 'xx-large')
     axes.legend(markerscale = 2, fontsize =20, loc = 'upper right')

### 2021-10-19 manuscripts/owp_fim3_2020/sections/abstract_plainTextSummary.tex
Updates * first draft complete * version sent out to JJ

@@
 Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D fluvial inundation extents used for forecasting purposes to protect life and property.
-The Office of Water Prediction Flood Inundation Mapping Version 3.0 `Cahaba' (OWP-FIM V3.0) is an enhancemant of the Height Above Nearest Drainage (HAND) method developed to improve upon some of HAND's limitations.
-Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, support for a two-tiered drainage density, parametrization of processing unit size, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and optimal Manning's n roughness parameter calibrated by stream order.
-OWP FIM V3 was calibrated and evaluated at 48 Base Level Engineering (BLE) sites furnished by the Federal Emergency Management Agency (FEMA).
-The modeled discharges from the BLE sections were used for the 100 and 500 year flood events as input to OWP-FIM V3 instaed of NWM discharges to isolate out factors related to hydro-meterology.
+The Office of Water Prediction Flood Inundation Mapping (OWP-FIM) model `Cahaba' is an enhancemant of the Height Above Nearest Drainage (HAND) method developed to improve upon some of HAND's limitations.
+Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and simpilified Manning's N variation.
+Most importantly we detail an issue previously mentioned with HAND as its FIM skill is dependent on drainage threshold.
+Since our modeling assumptions our different, we illustrate the enhancement of HAND based FIMs skill as the Horton-Strahler stream orders of the stream networks of the HAND processing units are reduced all the way down to unary.
+This is accomplished by deriving level paths for the National Water Model stream network and utilizing those definitions for deriving HAND datasets independetly from other level paths.
+`Cahaba' was evaluated at 48 Base Level Engineering (BLE) sites furnished by the Federal Emergency Management Agency (FEMA).
+The modeled discharges from the BLE sections were used for the 100 and 500 year flood events as input to OWP FIM instead of NWM discharges to isolate out factors related to hydro-meterology.
 The resulting maps are compared to the modeled extents from the BLE and binary contigency statistics are computed.
-Across the 48 sites, Critical Success Index (CSI) and Matthew's Correlation Coefficient (MCC) were found to increase across FIM versions driven by improvements to the probability of detection (POD) via reduction in errors of omission with only a slight increase in errors of commission.
+Across the 49 sites, Critical Success Index (CSI) was found to increase across FIM versions driven by improvements to the probability of detection (POD) and slight reduction in False Alarm Ratio (FAR).
 These capabalities were implemented in an open-source package available on GitHub (\url{https://github.com/NOAA-OWP/cahaba}) for continued use and collaboration with the research community.
@@
 \end{abstract}
-
+%
 \section*{Plain Language Summary}
@@
 In order to convert forecasts flow rates from the NWM to inundation extents, a model is used that converts elevations relative to sea-level to elevations relative to the nearest river bottom elevation.
-This model known as Height Above Nearest Drainage (HAND) suffers fro
-The maps were compared to other, more rigorous models and shown to perform well when compared to previous versions.
+This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
+The maps were compared to other, more rigorous models and shown to perform well when compared to previous versions given certain enhancements to the way HAND to computed.
 %

### 2021-10-19 manuscripts/owp_fim3_2020/sections/conclusions.tex
Updates * first draft complete * version sent out to JJ

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Overall, we have produced a product that is able to produce FIMs for the entire US using stream flows whether retrospective or forecast.
Using the HAND method, we can produce this product at very large domains and high resolutions in a relatively short amount of time which enables creating forecast maps at high temporal frequencies.
This yields a significant amount of information that was previously unavailable to forecasters before enabling for higher resolution and higher accuracy forecasts across areas of the country previously unsupported.
All of this is presented in a open-source framework that utilizes some of the latest enhancements to HAND as well as some of the latest data sources available. 
Additionally, a significant issue previously documented with HAND was identified and addressed via the reduction of stream order down to the unary level.
We illustrate how reducing a HAND processing unit to a stream network with unit Horton-Strahler stream order improves FIM skill especially at junctions of higher order and higher flow magnitude rivers.
We additionally illustrate the sensitivity to rating curve bias and implications of stream order reduction.
%

### 2021-10-19 manuscripts/owp_fim3_2020/sections/discussion.tex
added m&m
Updates * first draft complete * version sent out to JJ

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
+\label{sec:discussion}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Overall, we note a positive relationship between FIM skill and a reduction of the stream order of the stream network we use to derive the HAND datasets.
+Most of this change is accounted for by increasing POD thus reducing FNs especially along higher order rivers with higher flow magnitudes.
+We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
+This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
+One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
+
+In analyzing the data, we found a slight reduction in FAR was detected and more digging pointed to a bias in rating curves introduced by stream order reduction.
+Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
+Sub-figure \ref{fig:rating_curve_comparison}a shows how the average rating curves for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to bias down (and to the right) with ever increasing stream order reduction (FR to MS to GMS).
+This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
+Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
+Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's N values.
+While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
+Again, volume (V[y] or simply V) is synomous to reach-averaged cross-sectional area and bed area (B[y] or B) is analogous to reach-averaged hydraulic radius.
+Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively.
+Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significant higher values than MS which has higher values than FR.
+Again the relative discrepency between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
+Both V and B values increase but since both are weighed differently by their respective exponents and pull Q in different directions.
+We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ is plotted against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
+In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
+Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as show in Figure \ref{fig:rating_curve_comparison}e.
+Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments than encompass neighboring tributaries.
+These tributaries have a significant amount of bathymetry that is low-lying thus easily including the SRC derivation. 
+They also contribute volume and bed area that is technically not perpendicular to the flux of streamflow being accounted for in the stream in question. 
+Careful examination of Figure \ref{fig:gms_enhancement}b shows how much larger catchments include neighboring tributaries and the geometry associated with those tributaries. 
+This geometry is not perpendicular to the flow that is associated with the main reach thus leading to biases in the SRC.
+We consider this fact to have a nuanced effect on skill while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
+
+Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
+While the cause of these errors can be varied, we assert here that conflating 4 networks for use in evaluations leads to significant error.
+As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
+One of the issues is when a reach of given stream order accidently conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
+The utilization of MS and GMS only conflates to NWM catchments directly associated with the level path in question which is inherently easy to do with those methods. 
+Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
+The NWM stream network was derived using the NHD medium resolution dataset which was derived from coarser DEMs than those used here. 
+Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
+Until a singular stream network is used for the NWM, BLE benchmark, and for HAND based FIM, conflation will continue being a source of error.
+
+For future development, we assert using experience from qualitative analysis that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM.
+The bathymetry of the NED 10m DEM is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
+Manning's N which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
+Some adjusting parameter must be introduced to enhance the estimation of the bathymetric representation.
+Lidar DEMs from the USGS at 3m and 1m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
+
+Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
+For HAND to be derived and thus create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
+In other words, an entire area eligible for flooding must monotonically decrease in elevation. 
+DEM's naturally don't do this and the dynamics of true flood events don't follow drainage patterns.
+Enforcing this assumption for HAND leads to significant amount of DEM manipulations that introduce basic errors.
+These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
+Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large scales.
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/rating_curve_comparison.jpg}
+\caption{Illustrates average quantities for the three methods, FR, MS, and GMS, for each stage value (m). 
+The values are (a) Discharge $m^3s^{-1}$, (b) Volume $m^3$, (c) Bed Area $m^2$, (d) a function of Volume and Bed Area, and (e) number of pixels.
+}
+\label{fig:rating_curve_comparison}
+\end{figure}
+%
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
+\caption{OWP FIM `Cahaba' inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105.
+Catchment boundaries and stream lines are shown in white and dotted green, respectively.
+Sub-figure (a) shows agreement of FR HAND denoting significant areas of under-prediction due to junctions and catchment boundaries.
+Meanwhile, (b) shows the agreement for GMS and much larger catchments leading to much better inundation agreement for this given reach. 
+Overall, this illustrates the benefits of stream order reduction for deriving HAND datasets.
+}
+\label{fig:gms_enhancement}
+\end{figure}
+%

### 2021-10-19 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
M&M and eval_plots updates
Updates * first draft complete * version sent out to JJ

--- 2021-05-16 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A user only need to install Docker on their host machine and build the image from the provided recipe. 
+Source code is made available for this project on GitHub (https://github.com/NOAA-OWP/cahaba).
 
@@
 Producing the FIM hydrofabric produces the datasets required to make an inundation map including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
+A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
 %
@@
 Most data sources used within FIM 3 are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
+Instructions for accessing data are provided on the projects GitHub page via an Amazon Web Services (AWS) S3 bucket.
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
@@
 For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
-Additionally, FEMA Base Level Engineering (BLE) from Region 6 (parts of Texas, Oklahoma, Arkansas, Louisiana and New Mexico) 1\% (100 year) and 0.2\% (500 year) datasets are used as validation in this study \cite{fema2021base,fema2021estimated}.
+Additionally, FEMA Base Level Engineering (BLE) from Region 6 (parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi) 1\% (100 year) and 0.2\% (500 year) recurrence flow datasets are used as validation in this study \cite{fema2021base,fema2021estimated}.
 These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
@@
 %
-
 \begin{table}
@@
 \hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as resevoirs or lakes \\
+NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes \\
 \hline
@@
 \hline
-\multicolumn{2}{l}{$^{a}$Footnote text here.}
 \end{tabular}
 \end{table}
-
-Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain. 
-\note[Fernando Aristizabal]{For Nick Chadwich} This includes XXXX HUC4's, XXX HUC'6s, and XX HUC8's.
-An enhancement of OWP FIM v2 `Cahaba' over previous FIM versions is the support for Hawaii and Puerto Rico which the NWM v2.1 will cover.
+%
+Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and Generalized Mainstems (GMS) networks and 1,604 HUC8s for the MS method.
+These methods will be explained more later.
+An enhancement of OWP FIM `Cahaba' over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM v2.1 will cover.
 %
@@
 These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
+HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
 Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
 It can also be used to simulate more accurate bathymetry which is not accounted for in the NED 10m DEM \cite{gesch2002national}.
+HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
 
 Specifically within the context of FIM 3, the hydro-conditioning operations that take place in sequential order are presented. 
-Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing unit designated at run time which can be either HUC 4, 6 or 8. 
-The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reachcode's membership in the processing unit.
+Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing units of size HUC8. 
+The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reach code's membership in the processing unit.
 Hydro-conditioning raster operations take place on buffered boundary definitions to avoid edge contamination and effects \cite{lindsay2013measuring}. 
@@
 \note[Fernando Aristizabal]{For Brian Avant}
-The location of the stream network are enforced to ensure general agreement with established stream networks.
+The location of the stream network is enforced to ensure general agreement with established stream networks.
 The NHDPlus HR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
 However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-Pruning the NHD network is completed at for every NWM V2.1 headwater point, the nearest NHDPlusHR Burline Event feature is selected and downstream features are also selected.
+Pruning the NHD network is completed at for every NWM V2.1 headwater point where the nearest NHDPlusHR Burnline Event feature is selected and downstream features are also selected.
 The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
 This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees withe DEM after all hydro-conditioning takes place.
+Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
 %
@@
 In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10m DEM.
+The elevations found in the NLD are burned into the DEM if those elevations were found to exceed those already in place.
 %
@@
 This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
-The depression technique employed here does leave the existence of flat regions where pits existed aprior thus later requiring the need for resolving these flats.
+The depression technique employed here does leave the existence of flat regions where pits existed a prior thus later requiring the need for resolving these flats.
 The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
@@
 Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
-In order to prevent situations where the burned thalweg and the thalweg endemic to DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
+In order to prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
 Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50m.
@@
 The connectivity of the thalweg pixels is defined by the D-8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
-At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in equation \ref{eq:thalweg_breach},
+At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in Equation \ref{eq:thalweg_breach},
 %
+\begin{linenomath*}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath*}
 %
@@
 The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
-The FIM stream network is of the same density as the NWM V2.1 network and fully converges at all junctions leaving no divergences in the network.
+The FIM stream network is of similar drainage density as the NWM V2.1 network but fully converges at all junctions leaving no divergences in the network.
 This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to theheadwater definitions in the stream network of the NWM V2.1.
+These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
 Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
@@
 Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equi-distant reaches not to exceed a parameterized value of 1.5km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much geometry variance into one parametrized unit.
+Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
+Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much geometry variance into one parameterized unit.
 Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
@@
 \subsubsection{Catchments}
+\label{sssec:catchments}
 %
@@
 Once the pixel level catchments are derived the final relative elevations can be computed.
-Every non-thalweg elevation is substracted from the thalweg elevation within the same pixel-level catchment.
+Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment.
 The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in section \ref{sssec:stream_thalweg_elevation_conditioning}.
@@
 The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-The reach average Manning's formula is derived to be
-
+The reach averaged Manning's formula is derived to be 
+%
+\begin{linenomath*}
 \begin{equation}
@@
 \end{equation}
-
-where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
+\end{linenomath*}
+%
+where Q is discharge, y indicates the stage, n is the Manning's N roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
 All units are international given the 1 numerator above n.
@@
 Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to the channel length.
-Calibration of the Manning's n roughness coefficient is conducted by optimizing inundation extent performance to a benchmark dataset best described in Section \ref{ssec:evaluation_and_calibration}. 
+
+Setting of the Manning's N roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
+These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
+We note significant opportunity to enhance CFIM skill by better parameterizing Manning's N according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
+For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's N set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
+After all the parameters to the Manning's equation have been determined with either through hydrofabric sampling or user parameterization, we select stage values of 0 to 25 meters in third of a meter increments to calculate the discharge values for each stage value.
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Unary Stream Order}
-\label{ssec:unary_stream_order}
+\subsubsection{Cross-walking NWM Networks}
+\label{sssec:cross_walking_networks}
+%
+The stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with a NWM reach identifier so that a discharge can be converted to stage and later inundation extent.
+For the methods already discussed, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
+For two subsequent methods discussed in Sections \ref{sssec:replace_and_route_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
+Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:replace_and_route_mainstems} and \ref{sssec:generalized_mainstems}.
+While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
+More discussion on this will follow in Section \ref{sec:dicussion}.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Stream Order Reduction}
+\label{ssec:stream_order_reduction}
 %
@@
 In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
-Figure \ref{fig:confluence_problem_statement} illustrates this exact situation where two tributatries converge with a higher order stream segment. 
-A `Cahaba' FIM is generated and compared with a FEMA 100 year extent (please see sections \ref{ssec:evaluation_and_calibration} for more details) showing significant under-prediction in inundation extent
+Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
+A OWP FIM `Cahaba' FIM is generated and compared with a FEMA 100 year extent (please see sections \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
 The higher discharge along the mainstem of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
@@
 As a parallel problem, there is excess water accumulated along the mainstem that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the mainstem catchments.
-
+%
 \begin{figure}[h!]
 \centering
-%\includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{}
-\label{fig:confluence_problem_statement}
+\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
+\caption{False negatives associated with confluence of tributaries with mainstems.
+Integers represent flow values from BLE 100 year for the associated areas.
+No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
+\label{fig:catchment_boundaries_issue}
 \end{figure}
 %
-The limitation imposed by using stream networks of Horton-Strahler stream order \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} greater than one for HAND can be described mathematically with Graph Theory.
-Assume a drainage area denoted by special case of a fully connected, directed acyclic graph known as a tree, G, with all nodes (or reaches) directed from the leaf nodes (or headwater reaches) toward a singular root node $r_0$ known as the outlet in the application's field.
-Assume a singluar subgraph of G where by a new root node,$r_n$, is introduced that represents another stream location.
-n is used here to index possible root nodes which are a subset of G.
-Ensuring $r_n$ is itself an inner node (has upstream drainage) disconnects G into $G_0$ and $G_n$ thus removing influence of conditions at $r_0$ upon $G_n$.
-HAND seeks to divide drainage areas into reach scale catchments that are independent of one another.
-We intend to demonstrate how introduction of stream orders greater than one can generate errors at confluences.
+We seek to resolve this problem by deriving HAND for processing units with stream networks of reduced stream order. 
 We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
+The resulting FIMs from the overlapping HAND processing units are mosaiced together taking an inundated area to be so.
 %
@@
 \subsubsection{NWS Replace and Route Mainstems}
-\label{ssec:unary_stream_order}
+\label{sssec:replace_and_route_mainstems}
 %
-
+The RnR configuration of the NWM seeks to replace NWM forecasts by assimilating best available AHPS forecasts and routing them downstream.
+The RnR Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
+The RnR MS network, or simply MS, comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
+It also spans 121,724 reaches across 1,608 HUC8s.
+HAND was originally derived for this stream network to enhance mapping skill along these critical MS segments \cite{djokic2019arc}. 
+The inundation derived from this stream network is mosaiced with the inundation from the FR network to form the MS FIMs. 
+Within each HUC, you'll typically only find a MS stream network of stream order 1 but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
 %
@@
 \subsubsection{Generalized Mainstems}
-\label{ssec:unary_stream_order}
+\label{sssec:generalized_mainstems}
 %
+To further the efforts implemented by MS, we sought to derive HAND at a level-path scale which we call Generalized Mainstems (GMS).
+Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
+Level-paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
+Starting at the outlet, a unique level-path is propagated upstream. 
+At every confluence, the direction of maximum flow path length is sought to propagate the current level-path identifier.
+For the remaining parent reaches of the given junction, a new level-path identifier is assigned and the process recursively continues with them.
+Figure \ref{fig:level_path_methods} illustrate how level-paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
+Each HUC8 is discretized into level-paths independently and relevant inputs are assigned to each level-path processing unit given a buffer of 7 km.
+At the level-path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
+The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level-paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
 
+To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
+In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
+For each one of these lines, we derive HAND and it's associated datasets including catchments, crosswalks, and rating curves.
+Each level path is buffered to a polygon with a user-available distance parameter of 7km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
+We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
+Once the FIM hydrofabrics for each level path, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
+Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and show in Figure \ref{fig:gms_methods}c.
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/level_path_methods.jpg}
+\caption{Illustrates how level paths for the NWM are derived.
+Level paths symbolized by lines of unique colors are propagated upstream following the direction of maximum arbolate sum at each junction.
+}
+\label{fig:level_path_methods}
+\end{figure}
+%
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/gms_methods.jpg}
+\caption{Overall procedure for GMS HAND.
+In (a), we illustrate all NWM stream lines symbolized by their level path.
+Meanwhile (b), demonstrates the NED DEM clipped to a 7km buffer around two randomly chosen level paths.
+In (c), we show how HAND can be computed just for each one of these two level paths independently. 
+We also show inundation maps that be created for these two level paths in (c). 
+In (d), we show all the inundation maps for all the level paths mosaiced together. }
+\label{fig:gms_methods}
+\end{figure}
 %
@@
 For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
-From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 4, 6, or 8 spatial processing units in a rapid, parallel operation.
+From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 8 spatial processing units in a rapid, parallel operation.
 The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
-Once associated to a mapping stream network, 
-the catchments grid encoded with the reach identifiers are used to map
+
+Utilizing the stage-discharge relationships in the synthetic rating curves, each forecast for each catchment identifier is assigned a stage value. 
+The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
+We use the basic logic already established in previous works to conduct this \cite{nobre2016hand,liu2016cybergis,maidment2017conceptual}.
+Mathematically, the HAND values, $H_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
+For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
+For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
+%
+\begin{linenomath*}
+\begin{equation}
+\label{eq:hand_fim_depth}
+    D_{ij} = S_i - H_{ij}
+\end{equation}
+\end{linenomath*}
+%
+\begin{linenomath*}
+\begin{equation}
+\label{eq:hand_fim}
+    F_{ij} = D_{ij} > 0
+\end{equation}
+\end{linenomath*}
+%
+For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast.
+For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
+Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
+Equation \ref{eq:comp_fim} illustrates the same process but for the mosaic the binary inundation.
+%
+\begin{linenomath*}
+\begin{equation}
+\label{eq:comp_fim_depths}
+    D_{ij} = \max_{k=[1,...,K]} D_{ijk}
+\end{equation}
+\end{linenomath*}
+%
+For the MS and GMS methods, the contributing areas are defined differently.
+For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to 2 for that case.
+For GMS, all FIMs from all the level paths in a given area are mosaiced together then K is set to this number of level paths.
+Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
+%
+\begin{linenomath*}
+\begin{equation}
+\label{eq:comp_fim}
+    F_{ij} = \max_{k=[1,...,K]} F_{ijk}
+\end{equation}
+\end{linenomath*}
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation and Calibration}
+\subsection{Evaluation}
+\label{ssec:evaluation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\label{ssec:evaluation_and_calibration}
 %
-Evaluation of the OWP-FIM V3 Cahaba is conducted by comparison to the HEC-RAS 1D derived models produced by FEMA region 6 \cite{fema2021base,fema2021estimated}.
+\note[Fernando Aristizabal]{For Trevor Grout and Brad Bates}
+Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced by FEMA region 6 \cite{fema2021base,fema2021estimated}.
+49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
 The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by FEMA so we used those corresponding discharges and mapping extents for evaluation.
-By not using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors imputed by hydrology, forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
+We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
+By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
+Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
 We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates with each NWM reach. 
 To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
-This partially handles the influence of neighboring cross sections that could cause flow discontinuties and mass conservation issues.
-While the limitions are noted, this method does best detangle the influence of exogeneous variables that we do not wish to study in this comparison.
-
-Calibrating the Manning's n values is very important for optimizing the skill of inundation extent forecasts \cite{zheng2018river,garousi2019terrain}.
-\note[Fernando Aristizabal]{For Brian Avant} In this study we propose a simplified parameterization of roughness by stream order and tune inundation performance relative to the validation data set 
+This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
+Additionally, the stream network of the FEMA furnished models are of higher stream densities and bifurcation ratios, as evident in \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
+While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
 %
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
+\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states. This dataset for 1\% and 0.2\% recurrence flows were used as benchmarks.}
+\label{fig:all_ble_maps}
+\end{figure}
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
+\caption{Illustrates Base Level Engineering (BLE) cross sections and stream lines.
+The BLE stream network, which is denser than the NWM V2.1 stream lines, is also shown.
+BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM reach.
+}
+\label{fig:ble_evaluation_method}
+\end{figure}
+%
+The metrics employed in this study to evaluate inundation extents include critical success index (CSI), probability of detection (POD), and false alarm ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
+To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.
+The two types of errors consist of false positives (FP), or type I errors, which is dry in benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in benchmark by predicted dry.
+Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
+Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
+Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting systems utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
+While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
+Thus, frequency dependent statistics should be used with caution when comparing across sites with varying frequencies. 
+Lastly, approximately 6 HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability.
+%
+\begin{linenomath*}
+\begin{equation}
+\label{eq:csi}
+CSI = \frac{TP}{TP + FN + FP}
+\end{equation}
+\end{linenomath*}
+%
+\begin{linenomath*}
+\begin{equation}
+\label{eq:pod}
+POD = \frac{TP}{TP + FN}
+\end{equation}
+\end{linenomath*}
+%
+\begin{linenomath*}
+\begin{equation}
+\label{eq:far}
+FAR = \frac{FP}{TP + FP}
+\end{equation}
+\end{linenomath*}
+%

### 2021-10-19 manuscripts/owp_fim3_2020/sections/results.tex
Updates * first draft complete * version sent out to JJ

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Mapping Performance}
+\label{ssec:mapping_performance}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+We produced FIMs for the entire BLE domain within the 49 HUC8 area across several states in the south central US. 
+The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
+We computed the statistics (CSI, POD, and FAR) for both 100 and 500 year events for Mannings N set to 0.06 and 0.12. 
+The distribution of these statistics can be examined in Figure \ref{fig:violin_plot} as violin plots.
+Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's N value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
+We also denote trend lines for each metric and Manning's n setting as well as their respective slope estimate and one-tailed p-value denoting the level of significance of the trend.
 
+Aggregating the metrics in the method above treats each HUC8 as it's own unit and does little to consider the size differences of the HUCs. 
+In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics} the sum of all the TPs, FPs, and FNs to recompute CSI, POD, and FAR across the entire domain. 
+%
+\begin{table}[h!]
+\caption{Primary metrics, TPs, FPs, and FNs, aggregated for BLE domain to recompute CSI, POD, and FAR.}
+\label{tab:aggregate_metrics}
+\centering
+%\begin{tabular}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
+\begin{tabular}{|c|c||c|c|c|c|c|c|}
+\hline
+\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's N} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{MS} & \multicolumn{2}{|c|}{GMS} \\
+\cline{3-8}
+  &  & 100yr & 500yr & 100yr & 500yr & 100yr & 500yr \\
+\hline
+\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & 0.5796 & 0.6075 \\
+\cline{2-8}
+  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & 0.6182 & 0.6435 \\
+\hline
+\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & 0.6633 & 0.6863 \\
+\cline{2-8}
+  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & 0.7606 & 0.7810 \\
+\hline
+\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1778 & 0.1589 & 0.1787 & 0.1589 \\
+\cline{2-8}
+  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & 0.2324 & 0.2148 \\
+\hline
+\end{tabular}
+\end{table}
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.9]{figures/violin_plots.jpg}
+\caption{Shows kernel density estimation of the distributions (N = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
+The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's N (0.06 and 0.12).
+Trend lines for each metric / Mannings combination are shown (N = 294) along with associated slope and p-value of slope testing one-tailed significance.}
+\label{fig:violin_plot}
+\end{figure}
+%
+% Interpretation of metrics
+From Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, we denote several meaningful trends. 
+Using CSI as an overall proxy for skill of the FIM, we note that generally speaking the skill is correlated with a reduction of the stream orders of the processing units used for HAND.
+In other words, the more you derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
+While this trend is evident for both sets of Manning's N values, the trend is slightly more significant for the higher value of 0.12.
+Other trends related to this figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
+We also note how the higher Manning's N value enhances performance for both of these recurrence intervals across all models.
+
+Dissecting the improvements and trends presented in the previous paragraph comes down mostly to improvement in POD or a reduction in absolute amount of FNs.
+POD being the primary driver in skill enhancement is evident across models by comparing the slope of the POD lines with the slope of the FAR lines.
+Even though aggregating metrics by HUC8 yields a statistically zero trend, one does see a slight reduction in FAR across models that reduce HAND's maximum stream order.
+Additionally, we note that POD is a primary driver in enhancing performance across Manning's N values as well.
+This significant improvement comes at a cost of false alarms as the FAR increases significantly across Manning's N values.
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-All computational experiments were conducted on an Amazon Web Services (AWS) Elastic Cloud Compute (EC2) T3 instance (2x-large size) with 4 cores, 2 threads per core, and 32GB of random access memory (RAM).
-An Elastic Block Storage (EBS) solid state drive (SSD) of the general purpose 2 (gp2) type was used for 2TB of storage.
-All tests took place within the project's Docker image providing the operating system and software dependencies.
-Table \ref{tab:comp_results} details the computational metrics including CPU time, RAM usage, and data storage footprint for both the hydrofabric production and NRT generation of FIMs.
-For the hydrofabric production, computational results were produced for the entire NWM FIM domain noted in \ref{ssec:datasets}.
-For the NRT production of FIM from the hydrofabric, computational results were aggregated for the 48 HUC study area.
-Computation of the depths and inundation rasters as well as the inundation in polygon form are included in this anaylsis.
-Only a subset of NWM FIM domain ever gets mapped for operational purposes to avoid computational burdens by applying a 1.5\% NWM recurrence flow threshold.
-
-\begin{table}
-\caption{Descriptive computational statistics for `Cahaba' hydrofabric and NRT FIM production.}
-\label{tab:comp_results}
-\centering
-\begin{tabular}{|p{3cm}|p{3cm}|p{3cm}|p{3cm}|}
-%\hline
-%"" & \multicolumn{7}{l}{CPU Time (Hrs)} & \multicolumn{7}{l}{RAM (GB)} \\
-\hline
-Variable & Statistic & Hydrofabric & FIM \\
-\hline
-\multirow{7}{*}{CPU Time (hrs)} &
-n$^{a}$ & XX & XX \\ 
-\cline{2-4}
-& Sum  & XX & XX \\ 
-\cline{2-4}
-& Mean & XX & XX \\ 
-\cline{2-4}
-& Median & XX & XX \\ 
-\cline{2-4}
-& SD$^{b}$ & XX & XX \\ 
-\cline{2-4}
-& Max & XX & XX \\ 
-\cline{2-4}
-& Min & XX & XX \\ 
-\hline
-\multirow{7}{*}{RAM (GB)} &
-n$^{a}$ & XX & XX \\
-\cline{2-4}
-& Sum  & XX & XX \\
-\cline{2-4}
-& Mean & XX & XX \\
-\cline{2-4}
-& Median & XX & XX \\
-\cline{2-4}
-& SD$^{b}$ & XX & XX \\
-\cline{2-4}
-& Max & XX & XX \\
-\cline{2-4}
-& Min & XX & XX \\
-\hline
-\multirow{7}{*}{Data Storage (GB)} &
-n$^{a}$ & XX & XX \\
-\cline{2-4}
-& Sum  & XX & XX \\
-\cline{2-4}
-& Mean & XX & XX \\
-\cline{2-4}
-& Median & XX & XX \\
-\cline{2-4}
-& Std.Dev.$^{b}$ & XX & XX \\
-\cline{2-4}
-& Max & XX & XX \\
-\cline{2-4}
-& Min & XX & XX \\
-\hline
-\multicolumn{2}{l}{$^{a}$Sample size. $^{b}$Standard Deviation}
-\end{tabular}
-\end{table}
-
-For comparison purposes, OWP FIM v1 measured that the entire CONUS area excepting a few HUC6's took 1.34 CPU years \cite{liu2016cybergis}. 
-OWP FIM v2, developed by \citeA{djokic2019arc}, was estimated to take 0.55 CPU years for the entire NWM stream resolution but lacked the ability to run in parallel on HPC environments.
-By utilizing solid programming practices and optimal algorithm selection, we were able to reduce computational requirements while enhancing hydrologic performance.
-
-% TEST SCRIPT: 
-% for huc in $(cat /data/temp/ble_test_cases_fa_20210418.lst); do echo HUC: $huc; /usr/bin/time -v /foss_fim/tools/inundation.py -r /data/outputs/first_gms_batch_test/"$huc"/rem_zeroed_masked.tif -c /data/outputs/first_gms_batch_test/"$huc"/gw_catchments_reaches_filtered_addedAttributes.tif -b /data/outputs/first_gms_batch_test/"$huc"/gw_catchments_reaches_filtered_addedAttributes_crosswalked.gpkg -t /data/outputs/first_gms_batch_test/"$huc"/hydroTable.csv -f /data/test_cases/ble_test_cases/validation_data_ble/"$huc"/100yr/ble_huc_"$huc"_flows_100yr.csv -p /data/temp/inundation_py_benchmarking_for_manuscript/inundation_"$huc".gpkg -m filter -i /data/temp/inundation_py_benchmarking_for_manuscript/inundation_"$huc".tif -d /data/temp/inundation_py_benchmarking_for_manuscript/depths_"$huc".tif -u /data/inputs/wbd/WBD_National.gpkg -l WBDHU8 -q | tee /data/temp/inundation_py_benchmarking_for_manuscript/test_1.log; done
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Mapping Performance}
-\label{ssec:mapping_performance}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The NFIE experiments were able to produce HAND for 331 HUC6's in 1.34 CPU years and estimates using work from \citeA{djokic2019arc} put producing HAND at the FR NWM at 0.55 CPU years. 
+For our work, we were able to produce HAND at the full NWM resolution in 0.13 CPU years which represents a substantial speed-up compared to previous works.
+For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
+GMS which generalizes HAND production to level path scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years \note[Fernando Aristizabal]{Fernando A: Review}.
+%

### 2021-10-19 manuscripts/owp_fim3_2020/sections/title_authors.tex
Updates * first draft complete * version sent out to JJ

@@
 % Example: \title{This is a test title}
-
+%
 \title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
-
+%
 %% ------------------------------------------------------------------------ %%
@@
 % Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Trevor Grout\affil{1,3}, Gregory Petrochenkov\affil{4}, Nick Chadwick\affil{1,3}, Zachary Wills\affil{3,5}, Fred Ogden \affil{3}, Jasmeet Judge\affil{2}}
-
+%
+%
+\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{1,3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
+%
+%
 \affiliation{1}{Lynker Technologies, Leesburg, VA, USA}
@@
 \affiliation{3}{National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
-\affiliation{4}{New York Water Science Center, United States Geological Survey, Troy, NY, USA}
+\affiliation{4}{New York Water Science Center, Hydrologic Applied Innovations Lab, United States Geological Survey, Troy, NY, USA}
 \affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Alabama, Tuscaloosa, AL, USA}
-
+%
 %(repeat as many times as is necessary)
-
+%
 %% Corresponding Author:
@@
 % editorial system.)
-
+%
 % Example: \correspondingauthor{First and Last Name}{email@address.edu}
-
+%
 \correspondingauthor{Fernando Aristizabal}{fernando.aristizabal@noaa.gov}

### 2021-10-19 manuscripts/owp_fim4_2021/sections/introduction.tex
Changed all references to FIM3 to FIM3

--- 2021-05-18 manuscripts/owp_fim3_2020/sections/introduction.tex
@@
 Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
-One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with more than 3,000 forecast points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble stream flows at more than 3,000 forecast points and flood inundation maps (FIM) at more than 150 of those locations shown in Figure \ref{fig:forecast_points}.
+One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
+AHPS provides forecasting services in the form of ensemble stream flows at more than 3,000 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
 Besides the AHPS forecast points, we display in Figure \ref{fig:forecast_points} two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model}.
-AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multisensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
+AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
 On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
@@
 Figure \ref{ssec:national_water_model} shows Full-Resolution (FR) modeling stream network as well as the RnR mainstems (MS) network.
+The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
 
@@
 As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
-The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages to inundation extents.
+The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
 %
@@
 \citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
+A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant DEM manipulations to make a reality.
 The terrain index as even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
@@
 Originally, NFIE utilized the National Hydrography Dataset (NHD) Plus Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC to produce forecast FIM from the NWM for use within its network of RFC's for additional guidance in hydro-blind regions and tagged as OWP's FIM V1.0.
-Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS and tagged as OWP FIM V2.0.
+The original NFIE dataset was employed by the NWC to produce forecast FIM from the NWM for use within its network of RFC's for additional guidance in hydro-blind regions.
+Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
 More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{OWP FIM 3.0 `Cahaba'}
+\subsection{OWP FIM `Cahaba'}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-Many of those assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
+Many of those in the academic community assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
 \citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
-The colinearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
+The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
 Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
@@
 Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
-Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloaquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
 This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
@@
 
-A study is proposed to introduce and evaluate the development of OWP FIM V3.0 `Cahaba' which utilizes some of the latest mapping techniques along with a few novel developments in a computationally friendly framework.
-Most notably this study demonstrates how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
-The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM for the NWM for operational forecast applications. 
-Furthermore, it enables the contribution from the broader hydro-community through its use of open-source code as further advances are made in this developing area of research.
+With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operational pipeline is required.
+Here we introduce OWP FIM `Cahaba' that utilizes a few of the latest techniques in HAND based FIM oriented for use with NWM in continental scale operational forecasting settings. 
+This framework is open-source and modular enabling and accelerating the research to operational development pipeline.
+Automated evaluation tooling and processed test case data enable for the rapid testing and evaluation of new techniques in consistent contexts.
+In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
+Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
+Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order by assigning level paths to the target stream network and reducing the HAND processing unit to that singular level path.
+The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM for the NWM with applications.
 %

### 2021-10-19 manuscripts/owp_fim4_2021/sections/keypoints.tex
Changed all references to FIM3 to FIM3

--- 2021-05-18 manuscripts/owp_fim3_2020/sections/keypoints.tex
@@
 % \end{keypoints}
-
+%
 \begin{keypoints}
-\item Height Above Nearest Drainage (HAND) is a terrain index used for rapid generation of flood inundation maps (FIM) across large scales.
-\item Despite its advantages, FIM generated from HAND suffers from a number of limitations most notably its dependence on drainage density.
-\item Among other developments, here we present a means of resolving stream density by reducing drainage density to a singular flow path. 
+\item Height Above Nearest Drainage (HAND) is a terrain index used for rapid generation of large scale flood inundation maps at high resolutions.
+\item Despite its advantages, FIM derived from HAND and synthetic rating curves is subject to limitations at river junctions.
+\item A means of resolving this limitation is provided by reducing HAND processing units to level-paths of unit stream order.
 \end{keypoints}
+%

### 2021-10-19 manuscripts/owp_fim3_2020/docker_run_and_make.sh

@@ -1,5 +1,0 @@
-#!/bin/bash -e
-
-codeDir=$1
-
-docker run --rm -v "$1":/foss_fim fim3_latex:20210131 /foss_fim/manuscripts/owp_fim3_2020/Makefile.sh

### 2021-10-19 manuscripts/owp_fim4_2021/Makefile.sh

@@ -3,7 +3,7 @@
 parent_path=$( cd "$(dirname "${BASH_SOURCE[0]}")" ; pwd -P )
 cd $parent_path
 
-fullFile=owp_fim3.tex
+fullFile=owp_fim4.tex
 baseName="${fullFile%%.*}"
 extension="${fullFile#*.}"
 

### 2021-10-19 manuscripts/owp_fim4_2021/docker_run_and_make.sh

@@ -1,0 +1,5 @@
+#!/bin/bash -e
+
+codeDir=$1
+
+docker run --rm -v "$1":/foss_fim fim4_latex:20210131 /foss_fim/manuscripts/owp_fim4_2021/Makefile.sh

### 2021-11-03 manuscripts/owp_fim4_2021/sections/abstract_plainTextSummary.tex
latest changes

--- 2021-10-19 manuscripts/owp_fim3_2020/sections/abstract_plainTextSummary.tex
@@
-%% ------------------------------------------------------------------------ %%
+% ------------------------------------------------------------------------ %%
 %
@@
 Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D fluvial inundation extents used for forecasting purposes to protect life and property.
-The Office of Water Prediction Flood Inundation Mapping (OWP-FIM) model `Cahaba' is an enhancemant of the Height Above Nearest Drainage (HAND) method developed to improve upon some of HAND's limitations.
-Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and simpilified Manning's N variation.
-Most importantly we detail an issue previously mentioned with HAND as its FIM skill is dependent on drainage threshold.
-Since our modeling assumptions our different, we illustrate the enhancement of HAND based FIMs skill as the Horton-Strahler stream orders of the stream networks of the HAND processing units are reduced all the way down to unary.
-This is accomplished by deriving level paths for the National Water Model stream network and utilizing those definitions for deriving HAND datasets independetly from other level paths.
-`Cahaba' was evaluated at 48 Base Level Engineering (BLE) sites furnished by the Federal Emergency Management Agency (FEMA).
-The modeled discharges from the BLE sections were used for the 100 and 500 year flood events as input to OWP FIM instead of NWM discharges to isolate out factors related to hydro-meterology.
-The resulting maps are compared to the modeled extents from the BLE and binary contigency statistics are computed.
+Height Above Nearest Drainage (HAND) coupled with the use of synthetic rating curves can convert discharges into stages and stages to inundation extents in a rapid manner across large spatial domains and at high resolutions.
+The Office of Water Prediction Flood Inundation Mapping (OWP FIM) model `Cahaba' is an enhancement of the HAND method developed to improve upon some of HAND's limitations.
+Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and simplified Manning's N variation.
+Most importantly, we detail an issue previously mentioned with HAND as its FIM skill is dependent on a drainage threshold parameter used to delineate stream lines.
+Since our modeling assumptions are different, we illustrate how reducing Horton-Strahler stream order of the stream networks used within each HAND processing unit down to unary can improve FIM skill.
+This is accomplished by deriving level paths for the National Water Model stream network and utilizing those definitions for deriving HAND datasets independently from other level paths.
+`Cahaba' was evaluated at 49 Base Level Engineering (BLE) sites furnished by the Federal Emergency Management Agency (FEMA).
+The modeled discharges from the BLE sections were used for the 100 and 500 year flood events as input to OWP FIM instead of NWM discharges to isolate out factors related to hydro-meteorology.
+The resulting maps are compared to the modeled extents from the BLE and binary contingency statistics are computed.
 Across the 49 sites, Critical Success Index (CSI) was found to increase across FIM versions driven by improvements to the probability of detection (POD) and slight reduction in False Alarm Ratio (FAR).
-These capabalities were implemented in an open-source package available on GitHub (\url{https://github.com/NOAA-OWP/cahaba}) for continued use and collaboration with the research community.
+Each successive version of HAND here utilizes greater stream order reduction illustrating a key limitation of HAND and enhanced agreement when utilizing stream networks of lower stream orders as a datum for HAND.
+Additionally we detail the effects of two different Manning's N values and how order reduction drives rating curve bias downward due to the nature of reach averaged synthetic rating curves.
+These capabilities were implemented in an open-source package available on GitHub (\url{https://github.com/NOAA-OWP/cahaba}) for continued use and collaboration with the research community.
 %

### 2021-11-03 manuscripts/owp_fim4_2021/sections/discussion.tex
latest changes

--- 2021-10-19 manuscripts/owp_fim3_2020/sections/discussion.tex
@@
 While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
-Again, volume (V[y] or simply V) is synomous to reach-averaged cross-sectional area and bed area (B[y] or B) is analogous to reach-averaged hydraulic radius.
+Again, volume (V[y] or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B[y] or B) is analogous to reach-averaged hydraulic radius.
 Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively.
 Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significant higher values than MS which has higher values than FR.
-Again the relative discrepency between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
+Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
 Both V and B values increase but since both are weighed differently by their respective exponents and pull Q in different directions.
@@
 In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
-Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as show in Figure \ref{fig:rating_curve_comparison}e.
+Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
 Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments than encompass neighboring tributaries.
@@
 This geometry is not perpendicular to the flow that is associated with the main reach thus leading to biases in the SRC.
-We consider this fact to have a nuanced effect on skill while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
+We consider this fact to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
 
@@
 As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
-One of the issues is when a reach of given stream order accidently conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
+One of the issues is when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
 The utilization of MS and GMS only conflates to NWM catchments directly associated with the level path in question which is inherently easy to do with those methods. 

### 2021-11-03 manuscripts/owp_fim4_2021/sections/introduction.tex
latest changes

@@
 With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
-Unencouragingly, the trends related to flood damages and fatalities have been steadily increasing over recent decades.
+Trends related to flood damages and fatalities have been steadily increasing over recent decades.
 \cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
@@
 AHPS provides forecasting services in the form of ensemble stream flows at more than 3,000 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
-Besides the AHPS forecast points, we display in Figure \ref{fig:forecast_points} two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model}.
+Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
 AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
@@
 Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
-To broaden the forecasting domain, the Office of Water Prediction (OWP) at the National Water Center (NWC) in Tuscaloosa, Alabama commissioned the development of the National Water Model (NWM) which is an instance of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+To broaden the forecasting domain, the Office of Water Prediction (OWP) at the National Water Center (NWC) in Tuscaloosa, Alabama commissioned the development of the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
 The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
 The NWM enhances but does not replace the spatial and temporal domain of the current AHPS capabilities at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'. 
-It's simply an additional model to be used in the forecasting and early warning decision making.
 Furthermore, the NWM as of V1.2 has implemented not only assimilation of real-time United States Geological Survey (USGS) stage discharges but also assimilation of flow forecasts from the AHPS forecast points which are in turn routed downstream and updated whenever a new point is reached.
@@
 Overall, the software package is estimated to run CONUS at the full-resolution in 0.55 CPU years in a desktop setting.
+
+Related to these, the USGS has invested in relative elevation HAND like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
+Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
 In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30m resolutions \cite{yamazaki2019merit,donchyts2016global}.
@@
 Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order by assigning level paths to the target stream network and reducing the HAND processing unit to that singular level path.
+Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
 The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM for the NWM with applications.

### 2021-11-03 manuscripts/owp_fim4_2021/sections/materials_and_methods.tex
latest changes

--- 2021-10-19 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
@@
 %
-FIM 3 is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
+OWP FIM `Cahaba' is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
 %
@@
 %
-OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
+OWP FIM `Cahaba' exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
 Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in Docker image (\url{https://docs.docker.com/engine/install/}). 
+To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
 A user only need to install Docker on their host machine and build the image from the provided recipe. 
@@
 
-The FIM 3 pipeline is discretized into key areas that a user can interact with to reproduce the results of this study.
-Preprocessing acquires and prepares datasets for production of the FIM 3 hydrofabric.
-Producing the FIM hydrofabric produces the datasets required to make an inundation map including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
+The `Cahaba' pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
+Preprocessing acquires and prepares datasets for production of the FIM hydrofabric. 
+The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
 A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
@@
 %
-Most data sources used within FIM 3 are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
+Data sources used within `Cahaba' are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
 Instructions for accessing data are provided on the projects GitHub page via an Amazon Web Services (AWS) S3 bucket.
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for FIM 3 while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
+It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for `Cahaba' while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
 For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
@@
 HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
+In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligble for flooding.
 Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
 It can also be used to simulate more accurate bathymetry which is not accounted for in the NED 10m DEM \cite{gesch2002national}.
-HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
 
-Specifically within the context of FIM 3, the hydro-conditioning operations that take place in sequential order are presented. 
+Specifically within the context of `Cahaba', the hydro-conditioning operations that take place in sequential order are presented. 
 Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing units of size HUC8. 
@@
 However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-Pruning the NHD network is completed at for every NWM V2.1 headwater point where the nearest NHDPlusHR Burnline Event feature is selected and downstream features are also selected.
+Pruning the NHD network is completed for every NWM V2.1 headwater point where the nearest NHDPlusHR Burnline Event feature is selected and downstream features are also selected.
 The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
@@
 The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
-The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southwest direction as 8. 
+The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southeast direction as 8. 
 Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
 
-Flat resolution from flats endemic to the DEM or from depression filled regions is a costly,non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
+Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
 Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
@@
 By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
-OWP FIM 3 utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
+`Cahaba' utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
 %
@@
 Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
-Gages set at the pixel center points of the delineated stream lines explained in section \ref{sssec:deriving_fim_stream_network}.
+Gages are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
 The gages act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
 Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
-The other catchments is derived for the contributing area for each stream reach as defined in section \ref{sssec:deriving_fim_stream_network}. 
+The other catchments is derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
 %
@@
 %
-Once the pixel level catchments are derived the final relative elevations can be computed.
-Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment.
-The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in section \ref{sssec:stream_thalweg_elevation_conditioning}.
+Once the pixel level catchments are derived, the final relative elevations can be computed.
+Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment described in Section \ref{sssec:catchments}.
+The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
 Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling. 
@@
 For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's N set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either through hydrofabric sampling or user parameterization, we select stage values of 0 to 25 meters in third of a meter increments to calculate the discharge values for each stage value.
+After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select stage values of 0 to 25 meters in third of a meter increments to calculate the discharge values for each stage value.
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cross-walking NWM Networks}
+\subsubsection{Cross-walking with NWM Stream Network}
 \label{sssec:cross_walking_networks}
@@
 While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
-More discussion on this will follow in Section \ref{sec:dicussion}.
+More discussion on this will follow in Section \ref{sec:discussion}.
 %
@@
 Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
-A OWP FIM `Cahaba' FIM is generated and compared with a FEMA 100 year extent (please see sections \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
+A `Cahaba' FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (please see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
 The higher discharge along the mainstem of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
@@
 \includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{False negatives associated with confluence of tributaries with mainstems.
-Integers represent flow values from BLE 100 year for the associated areas.
+\caption{False negatives associated with confluence of tributaries with mainstems. Integers represent flow values from BLE 100 year event for the associated areas. 
 No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
@@
 We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
-The resulting FIMs from the overlapping HAND processing units are mosaiced together taking an inundated area to be so.
+The resulting FIMs from the overlapping HAND processing units are mosaiced together taking any inundated area to be inundated but more will be explained in Section \ref{ssec:inundation_mapping}.
 %
@@
 %
-To further the efforts implemented by MS, we sought to derive HAND at a level-path scale which we call Generalized Mainstems (GMS).
+To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call Generalized Mainstems (GMS).
 Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
-Level-paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-Starting at the outlet, a unique level-path is propagated upstream. 
-At every confluence, the direction of maximum flow path length is sought to propagate the current level-path identifier.
-For the remaining parent reaches of the given junction, a new level-path identifier is assigned and the process recursively continues with them.
-Figure \ref{fig:level_path_methods} illustrate how level-paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-Each HUC8 is discretized into level-paths independently and relevant inputs are assigned to each level-path processing unit given a buffer of 7 km.
-At the level-path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level-paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
+Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
+Starting at the outlet, a unique level path is propagated upstream. 
+At every confluence, the direction of maximum flow path length is sought to propagate the current level path identifier.
+For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
+Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
+Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
+At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
+The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
 
@@
 In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
-For each one of these lines, we derive HAND and it's associated datasets including catchments, crosswalks, and rating curves.
+For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
 Each level path is buffered to a polygon with a user-available distance parameter of 7km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
@@
 Once the FIM hydrofabrics for each level path, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
-Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and show in Figure \ref{fig:gms_methods}c.
+Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}c.
 %
@@
 In (a), we illustrate all NWM stream lines symbolized by their level path.
-Meanwhile (b), demonstrates the NED DEM clipped to a 7km buffer around two randomly chosen level paths.
+Meanwhile (b), demonstrates the NED DEM clipped to a 7km buffer around two selected level paths.
 In (c), we show how HAND can be computed just for each one of these two level paths independently. 
-We also show inundation maps that be created for these two level paths in (c). 
+We also show inundation maps created for these two level paths in (c). 
 In (d), we show all the inundation maps for all the level paths mosaiced together. }
@@
 Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting systems utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
+Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
 While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.

### 2021-11-03 manuscripts/owp_fim4_2021/sections/results.tex
latest changes

--- 2021-10-19 manuscripts/owp_fim3_2020/sections/results.tex
@@
 In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics} the sum of all the TPs, FPs, and FNs to recompute CSI, POD, and FAR across the entire domain. 
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.9]{figures/violin_plots.jpg}
+\caption{Shows kernel density estimation of the distributions (sample size = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
+The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's N (0.06 and 0.12).
+Trend lines for each metric - Mannings combination are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}
+\label{fig:violin_plot}
+\end{figure}
 %
@@
 %
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.9]{figures/violin_plots.jpg}
-\caption{Shows kernel density estimation of the distributions (N = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
-The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's N (0.06 and 0.12).
-Trend lines for each metric / Mannings combination are shown (N = 294) along with associated slope and p-value of slope testing one-tailed significance.}
-\label{fig:violin_plot}
-\end{figure}
-%
 % Interpretation of metrics
@@
 While this trend is evident for both sets of Manning's N values, the trend is slightly more significant for the higher value of 0.12.
-Other trends related to this figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
+Other trends related to this Figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
 We also note how the higher Manning's N value enhances performance for both of these recurrence intervals across all models.
@@
 %
-The NFIE experiments were able to produce HAND for 331 HUC6's in 1.34 CPU years and estimates using work from \citeA{djokic2019arc} put producing HAND at the FR NWM at 0.55 CPU years. 
+The NFIE experiments were able to produce HAND for 331 HUC6's in 1.34 CPU years \cite{liu2016cybergis} and estimates using work from \citeA{djokic2019arc} put producing HAND at the FR NWM at 0.55 CPU years.
 For our work, we were able to produce HAND at the full NWM resolution in 0.13 CPU years which represents a substantial speed-up compared to previous works.

### 2021-12-01 manuscripts/owp_fim4_2021/sections/title_authors.tex
latest changes
Update author info

--- 2021-10-19 manuscripts/owp_fim3_2020/sections/title_authors.tex
@@
 \affiliation{1}{Lynker Technologies, Leesburg, VA, USA}
-\affiliation{2}{Center for Remote Sensing, University of Florida, Gainesville, FL, USA}
+\affiliation{2}{Center for Remote Sensing, Agricultural and Biological Engineering Department, University of Florida, Gainesville, FL, USA}
 \affiliation{3}{National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
 \affiliation{4}{New York Water Science Center, Hydrologic Applied Innovations Lab, United States Geological Survey, Troy, NY, USA}
-\affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Alabama, Tuscaloosa, AL, USA}
+\affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Maryland, College Park, MD, USA}
 %

### 2021-12-22 manuscripts/owp_fim4_2021/sections/abstract_plainTextSummary.tex
Manuscript edits made after feedback from co-authors

@@
 The Office of Water Prediction Flood Inundation Mapping (OWP FIM) model `Cahaba' is an enhancement of the HAND method developed to improve upon some of HAND's limitations.
-Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and simplified Manning's N variation.
+Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, enforcement of levee elevation data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and simplified Manning's N variation.
 Most importantly, we detail an issue previously mentioned with HAND as its FIM skill is dependent on a drainage threshold parameter used to delineate stream lines.
@@
 The resulting maps are compared to the modeled extents from the BLE and binary contingency statistics are computed.
-Across the 49 sites, Critical Success Index (CSI) was found to increase across FIM versions driven by improvements to the probability of detection (POD) and slight reduction in False Alarm Ratio (FAR).
+Across the 49 sites, Critical Success Index (CSI) was found to increase across FIM versions driven by improvements to the Probability of Detection (POD) and slight reduction in False Alarm Ratio (FAR).
 Each successive version of HAND here utilizes greater stream order reduction illustrating a key limitation of HAND and enhanced agreement when utilizing stream networks of lower stream orders as a datum for HAND.
@@
 The United States National Water Model (NWM) seeks to provide flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
-In order to convert forecasts flow rates from the NWM to inundation extents, a model is used that converts elevations relative to sea-level to elevations relative to the nearest river bottom elevation.
+In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
 This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
-The maps were compared to other, more rigorous models and shown to perform well when compared to previous versions given certain enhancements to the way HAND to computed.
+We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
+We compared these HAND derived FIMs to maps from more realistic models and found and improvement in mapping performance.
 %

### 2021-12-22 manuscripts/owp_fim4_2021/sections/conclusions.tex
Manuscript edits made after feedback from co-authors

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Overall, we have produced a product that is able to produce FIMs for the entire US using stream flows whether retrospective or forecast.
Using the HAND method, we can produce this product at very large domains and high resolutions in a relatively short amount of time which enables creating forecast maps at high temporal frequencies.
This yields a significant amount of intelligence that was previously unavailable to forecasters.
All of this is presented in an open-source framework that utilizes some of the latest enhancements to HAND as well as some of the latest data sources available. 
HAND is a terrain index that normalizes height above mean sea level to height above the nearest, relevant drainage line thus detrending elevations.
It is dependent on a series of drainage enforcing operations and suffers from limitations at stream junctions that partially result from independent catchments and multi-order stream networks.
This issue is addressed here by reducing the scale at which HAND is derived from a HUC level to one of a levelpath that reduces the stream order down to a unary level.
We illustrated how this technique improves FIM skill especially at junctions of higher order and higher flow magnitude rivers.
We additionally illustrated the sensitivity to rating curve bias and implications of stream order reduction.
Lastly, continental scale flood forecasting still has much progress left to make so we proposed a path forward for improving FIM via the enhancement of terrain data, roughness estimates, and physics constraints. 
%

### 2021-12-22 manuscripts/owp_fim4_2021/sections/discussion.tex
Manuscript edits made after feedback from co-authors

@@
 Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively.
-Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significant higher values than MS which has higher values than FR.
+Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
 Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
@@
 Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
-Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments than encompass neighboring tributaries.
+Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments that encompass neighboring tributaries.
 These tributaries have a significant amount of bathymetry that is low-lying thus easily including the SRC derivation. 
@@
 
-For future development, we assert using experience from qualitative analysis that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM.
+Our qualitative analysis suggests that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM for future development.
 The bathymetry of the NED 10m DEM is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
@@
 These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
-Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large scales.
+Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
 %

### 2021-12-22 manuscripts/owp_fim4_2021/sections/introduction.tex
Manuscript edits made after feedback from co-authors

@@
 \caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
-Also show are the National Water Model stream networks at the full resolution (FR) and Replace and Reroute (RnR) Mainstems (MS) resolution.}
+Also show are the National Water Model stream networks at the full resolution (FR) and Replace and Route (RnR) Mainstems (MS) resolution.}
 \label{fig:forecast_points}
@@
 The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
-The term hydrofabric is used within the NWM jargon to describe the subset of hydrography comprised of geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
+The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
 The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
@@
 %
-HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and ground water dynamics \cite{renno2008hand,nobre2011height}.
+HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
 \citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
 A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant DEM manipulations to make a reality.
-The terrain index as even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
-\citeA{zheng2018river} developed methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
-This collection of methods, coupling HAND with SRCs, has been experimented and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
+The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
+\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
+This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
 %
@@
 %
-Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEM's such as the National Elevation Dataset (NED) at the 10m scale, HAND has been implemented into software for large-scale, continental computation.
+Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the National Elevation Dataset (NED) at the 10m scale, HAND has been implemented into software for large-scale, continental computation.
 HAND was initially implemented into operational software by the National Flood Interoperability Experiment (NFIE) to generate FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
@@
 Originally, NFIE utilized the National Hydrography Dataset (NHD) Plus Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC to produce forecast FIM from the NWM for use within its network of RFC's for additional guidance in hydro-blind regions.
+The original NFIE dataset was employed by the NWC to produce forecast FIM from the NWM for use within its network of RFCs for additional guidance in hydro-blind regions.
 Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
@@
 
-Related to these, the USGS has invested in relative elevation HAND like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
+Related to these, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
 Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
@@
 Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
-Use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
+The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
 Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
@@
 
-With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operational pipeline is required.
+With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
 Here we introduce OWP FIM `Cahaba' that utilizes a few of the latest techniques in HAND based FIM oriented for use with NWM in continental scale operational forecasting settings. 
 This framework is open-source and modular enabling and accelerating the research to operational development pipeline.
-Automated evaluation tooling and processed test case data enable for the rapid testing and evaluation of new techniques in consistent contexts.
+Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
 In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
@@
 Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
-The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM for the NWM with applications.
+The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
 %

### 2021-12-22 manuscripts/owp_fim4_2021/sections/materials_and_methods.tex
Manuscript edits made after feedback from co-authors

@@
 Data sources used within `Cahaba' are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data are provided on the projects GitHub page via an Amazon Web Services (AWS) S3 bucket.
+Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket.
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
@@
 For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
-Additionally, FEMA Base Level Engineering (BLE) from Region 6 (parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi) 1\% (100 year) and 0.2\% (500 year) recurrence flow datasets are used as validation in this study \cite{fema2021base,fema2021estimated}.
+Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Aquatic Habitat Framework (GLAHF) to exclude those areas from production of FIMs.
+Additionally, the Base Level Engineering (BLE) datasets spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by Region 6 of FEMA \cite{fema2021base,fema2021estimated}.
 These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
-The full input datasets presented by source are list in Table \ref{tab:data}.
-\note[Fernando Aristizabal]{Check Brian Avant, Trevor Grout, Bradford Bates, Ryan Spies}
+The full input datasets presented by source are listed in Table \ref{tab:data}.
 %
@@
 \hline
-WHERE? & Land-Sea Border & Border of land and sea and land and Great Lakes \\
+OSM & Land-Sea Border & Border of land and sea. \\
 \hline
-FEMA & Cross-Sections & HEC-RAS cross-sections used for modeling in BLE datasets. Include discharges for 1\% and 0.2\% recurrence interval events \\
+GLAHF & Land-Great Lakes Border & Border of land and Great Lakes \\
 \hline
-FEMA & Flood Inundation Extents & Extents produced by FEMA BLE for 1\% and 0.2\% recurrence interval events \\
+FEMA & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Include discharges for 1\% and 0.2\% recurrence interval events \\
+\hline
+FEMA & Flood Inundation Extents & Inundation depths produced by FEMA BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events \\
 \hline
@@
 %
-\note[Fernando Aristizabal]{For Brian Avant}
 The location of the stream network is enforced to ensure general agreement with established stream networks.
@@
 However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-Pruning the NHD network is completed for every NWM V2.1 headwater point where the nearest NHDPlusHR Burnline Event feature is selected and downstream features are also selected.
+For every NWM headwater segment a headwater point is derived and linearly interpolated to the nearest Burnline Event segment.
+Burnline Event headwater segments are split at the adjusted headwater point to match NWM flowlines.
 The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
@@
 \includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection.
+\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection and linear interpolation.
 Resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR Burnline Events.}
@@
 
-\note[Fernando Aristizabal]{For Trevor Grout}
 The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
@@
 In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
-Using simple thalweg burning techniques as opposed to the full AGREE method helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
+Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
 \citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
@@
 %
-\note[Fernando Aristizabal]{For Ryan Spies}
-The National Elevation Dataset at 10m resolution lacks sufficient representation of fine grain features such as levees and embankments.
+The National Elevation Dataset at 10m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
 In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10m DEM.
@@
 When a pixel's upstream neighbors are all evaluated, the downstream pixel is added into the queue thus the depth first traversal of the drainage network.
-This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in critical success index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
+This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in Critical Success Index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
 %
@@
 %
-\note[Fernando Aristizabal]{For Brian Avant}
 The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
@@
 Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much geometry variance into one parameterized unit.
+Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherent the synthetic rating curve properties of the upstream or downstream segment, depending on the topology.
 Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
@@
  A method for converting forecast river discharges from the NWM to stages or river depths is necessary for producing FIMs with HAND. 
-For one-dimensional models such as the NWM, the typical procedure for establishing this conversion is to determine the stage-discharge relationship sampling data from the DEM known as a synthetic rating curve with cross-sections \cite{quintero2021development,di2011hydraulic}.
+For one-dimensional models such as the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a synthetic rating curve at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
 For this application, we utilized the reach averaged approach for developing synthetic rating curves (SRC) \cite{zheng2018river}.
@@
 The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
-The reach averaged geometry parameters including number of cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to the channel length.
+The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
+Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the channel length.
 
@@
 For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's N set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select stage values of 0 to 25 meters in third of a meter increments to calculate the discharge values for each stage value.
+After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
 %
@@
 %
-For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast.
+For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast in the form of a single raster file.
 For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
@@
 %
-\note[Fernando Aristizabal]{For Trevor Grout and Brad Bates}
 Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced by FEMA region 6 \cite{fema2021base,fema2021estimated}.
@@
 This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the FEMA furnished models are of higher stream densities and bifurcation ratios, as evident in \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
+Additionally, the stream network of the FEMA furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
 While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
@@
 %
-The metrics employed in this study to evaluate inundation extents include critical success index (CSI), probability of detection (POD), and false alarm ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
+The metrics employed in this study to evaluate inundation extents include Critical Success Index (CSI), Probability of Detection (POD), and False Alarm Ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
 To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.

### 2021-12-22 manuscripts/owp_fim4_2021/sections/results.tex
Manuscript edits made after feedback from co-authors

@@
 For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
-GMS which generalizes HAND production to level path scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years \note[Fernando Aristizabal]{Fernando A: Review}.
+GMS which generalizes HAND production to level path scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
 %

### 2021-12-22 manuscripts/owp_fim4_2021/docker_run_and_make.sh

@@ -2,4 +2,4 @@
 
 codeDir=$1
 
-docker run --rm -v "$1":/foss_fim fim4_latex:20210131 /foss_fim/manuscripts/owp_fim4_2021/Makefile.sh
+docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex:20211220 /foss_fim/manuscripts/owp_fim4_2021/Makefile.sh

### 2021-12-22 manuscripts/owp_fim4_2021/sections/abstract_plainTextSummary.tex
Minor edits to manuscript

@@
 The Office of Water Prediction Flood Inundation Mapping (OWP FIM) model `Cahaba' is an enhancement of the HAND method developed to improve upon some of HAND's limitations.
-Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, enforcement of levee elevation data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and simplified Manning's N variation.
+Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, enforcement of levee elevation data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and simplified Manning's n variation.
 Most importantly, we detail an issue previously mentioned with HAND as its FIM skill is dependent on a drainage threshold parameter used to delineate stream lines.
@@
 Each successive version of HAND here utilizes greater stream order reduction illustrating a key limitation of HAND and enhanced agreement when utilizing stream networks of lower stream orders as a datum for HAND.
-Additionally we detail the effects of two different Manning's N values and how order reduction drives rating curve bias downward due to the nature of reach averaged synthetic rating curves.
+Additionally we detail the effects of two different Manning's n values and how order reduction drives rating curve bias downward due to the nature of reach averaged synthetic rating curves.
 These capabilities were implemented in an open-source package available on GitHub (\url{https://github.com/NOAA-OWP/cahaba}) for continued use and collaboration with the research community.

### 2021-12-22 manuscripts/owp_fim4_2021/sections/introduction.tex
Minor edits to manuscript

@@
 The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
-Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
+Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's N coefficient have been observed.
 Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).

### 2021-12-22 manuscripts/owp_fim4_2021/sections/materials_and_methods.tex
Minor edits to manuscript

@@
 For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Aquatic Habitat Framework (GLAHF) to exclude those areas from production of FIMs.
+Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{OpenStreetMap,GreatLakesHydrographyDataset}.
 Additionally, the Base Level Engineering (BLE) datasets spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by Region 6 of FEMA \cite{fema2021base,fema2021estimated}.
@@
 \hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting \\
+NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. \\
 \hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM \\
+NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. \\
 \hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes \\
+NOAA-OWP & Waterbodies considered by the NWM as reservoirs or lakes. \\
 \hline
@@
 \hline
-GLAHF & Land-Great Lakes Border & Border of land and Great Lakes \\
+GLHD & Land-Great Lakes Border & Border of land and Great Lakes. \\
 \hline
-FEMA & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Include discharges for 1\% and 0.2\% recurrence interval events \\
+FEMA & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
 \hline
-FEMA & Flood Inundation Extents & Inundation depths produced by FEMA BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events \\
+FEMA & Flood Inundation Extents & Inundation depths produced by FEMA BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
 \hline
@@
 %
-The DEM from the NED is subject to a series of hydro-conditioning procedures to enhance it's suitability for riverine flood inundation mapping. 
+The DEM from the NED is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
 These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
@@
 Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
-The other catchments is derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
+The other catchments are derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
 %

### 2021-12-23 manuscripts/owp_fim4_2021/sections/discussion.tex
minor edits

@@
 Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
-Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's N values.
+Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
 While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
@@
 The bathymetry of the NED 10m DEM is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
-Manning's N which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
+Manning's n which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
 Some adjusting parameter must be introduced to enhance the estimation of the bathymetric representation.

### 2021-12-23 manuscripts/owp_fim4_2021/sections/materials_and_methods.tex
minor edits

@@
 %
-where Q is discharge, y indicates the stage, n is the Manning's N roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
+where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
 All units are international given the 1 numerator above n.
@@
 
-Setting of the Manning's N roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
+Setting of the Manning's n roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
 These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's N according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's N set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
+We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
+For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
 After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.

### 2021-12-23 manuscripts/owp_fim4_2021/sections/results.tex
minor edits

@@
 The distribution of these statistics can be examined in Figure \ref{fig:violin_plot} as violin plots.
-Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's N value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
+Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
 We also denote trend lines for each metric and Manning's n setting as well as their respective slope estimate and one-tailed p-value denoting the level of significance of the trend.
@@
 \caption{Shows kernel density estimation of the distributions (sample size = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
-The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's N (0.06 and 0.12).
+The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's n (0.06 and 0.12).
 Trend lines for each metric - Mannings combination are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}
@@
 \hline
-\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's N} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{MS} & \multicolumn{2}{|c|}{GMS} \\
+\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{MS} & \multicolumn{2}{|c|}{GMS} \\
 \cline{3-8}
@@
 In other words, the more you derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
-While this trend is evident for both sets of Manning's N values, the trend is slightly more significant for the higher value of 0.12.
+While this trend is evident for both sets of Manning's n values, the trend is slightly more significant for the higher value of 0.12.
 Other trends related to this Figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
-We also note how the higher Manning's N value enhances performance for both of these recurrence intervals across all models.
+We also note how the higher Manning's n value enhances performance for both of these recurrence intervals across all models.
 
@@
 Even though aggregating metrics by HUC8 yields a statistically zero trend, one does see a slight reduction in FAR across models that reduce HAND's maximum stream order.
-Additionally, we note that POD is a primary driver in enhancing performance across Manning's N values as well.
-This significant improvement comes at a cost of false alarms as the FAR increases significantly across Manning's N values.
+Additionally, we note that POD is a primary driver in enhancing performance across Manning's n values as well.
+This significant improvement comes at a cost of false alarms as the FAR increases significantly across Manning's n values.
 %

### 2022-01-10 manuscripts/owp_fim4_2021/sections/abstract_plainTextSummary.tex
edits after odgen/salas comments

@@
 \begin{abstract}
-The National Water Model Version 2.1 (NWM V2.1) seeks to expand upon existing systems to provide hourly forecast discharges at four time horizons (1, 3, 10, and 30 days) at over 2.7 million locations throughout the continental United States, Hawaii and Puerto Rico.
-Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D fluvial inundation extents used for forecasting purposes to protect life and property.
+The National Water Model Version 2.1 (NWM V2.1) seeks to expand upon existing systems to provide up to hourly forecast discharges for four time horizons (1, 3, 10, and 30 days) at over 2.7 million locations throughout the continental United States, Hawaii and Puerto Rico.
+Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D fluvial inundation extents used for forecasting purposes to aid in the protection of life and property.
 Height Above Nearest Drainage (HAND) coupled with the use of synthetic rating curves can convert discharges into stages and stages to inundation extents in a rapid manner across large spatial domains and at high resolutions.
@@
 This is accomplished by deriving level paths for the National Water Model stream network and utilizing those definitions for deriving HAND datasets independently from other level paths.
-`Cahaba' was evaluated at 49 Base Level Engineering (BLE) sites furnished by the Federal Emergency Management Agency (FEMA).
-The modeled discharges from the BLE sections were used for the 100 and 500 year flood events as input to OWP FIM instead of NWM discharges to isolate out factors related to hydro-meteorology.
+`Cahaba' was evaluated at 49 Base Level Engineering (BLE) sites furnished by the Interagency Flood Risk Management (InFIRM).
+The modeled discharges for 100 and 500 year events from the BLE cross-sections were used for input to OWP FIM instead of NWM discharges to isolate out factors related to hydro-meteorology.
 The resulting maps are compared to the modeled extents from the BLE and binary contingency statistics are computed.
@@
 Flooding is one of the most impactful natural disasters on life and property.
-The United States National Water Model (NWM) seeks to provide flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
+The United States National Water Model (NWM) provides flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
 In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.

### 2022-01-10 manuscripts/owp_fim4_2021/sections/conclusions.tex
edits after odgen/salas comments

@@
 %
-Overall, we have produced a product that is able to produce FIMs for the entire US using stream flows whether retrospective or forecast.
-Using the HAND method, we can produce this product at very large domains and high resolutions in a relatively short amount of time which enables creating forecast maps at high temporal frequencies.
-This yields a significant amount of intelligence that was previously unavailable to forecasters.
-All of this is presented in an open-source framework that utilizes some of the latest enhancements to HAND as well as some of the latest data sources available. 
-HAND is a terrain index that normalizes height above mean sea level to height above the nearest, relevant drainage line thus detrending elevations.
-It is dependent on a series of drainage enforcing operations and suffers from limitations at stream junctions that partially result from independent catchments and multi-order stream networks.
-This issue is addressed here by reducing the scale at which HAND is derived from a HUC level to one of a levelpath that reduces the stream order down to a unary level.
-We illustrated how this technique improves FIM skill especially at junctions of higher order and higher flow magnitude rivers.
-We additionally illustrated the sensitivity to rating curve bias and implications of stream order reduction.
-Lastly, continental scale flood forecasting still has much progress left to make so we proposed a path forward for improving FIM via the enhancement of terrain data, roughness estimates, and physics constraints.
+Floods present a significant, underserved, and expanding risk to life, property, and resources.
+Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
+The National Water Model (NWM) developed by the National Oceanic and Atmospheric Administration's Office of Water Prediction, along with partners, provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
+Additional processing is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
+Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
+HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
+Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico (https://github.com/NOAA-OWP/cahaba).
+HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
+Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
+We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
+This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
+FIM skill was evaluated over large spatial scales by comparsion to HEC-RAS 1D models.
+Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.
+Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
+Due to inherent limitations with HAND, scalable physics-based methods will need to be worked on to provide better representation of flood extent dynamics in steady and unsteady conditions. 
 %

### 2022-01-10 manuscripts/owp_fim4_2021/sections/discussion.tex
edits after odgen/salas comments

@@
 While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
-Again, volume (V[y] or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B[y] or B) is analogous to reach-averaged hydraulic radius.
-Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively.
+Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius.
+Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively (see Equation \ref{eq:reach_averaged_mannings_equation}).
 Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
 Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
-Both V and B values increase but since both are weighed differently by their respective exponents and pull Q in different directions.
-We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ is plotted against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
+Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
+We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
 In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
@@
 Our qualitative analysis suggests that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM for future development.
-The bathymetry of the NED 10m DEM is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
+The bathymetry of the 10 m DEM from 3DEP is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
 Manning's n which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
 Some adjusting parameter must be introduced to enhance the estimation of the bathymetric representation.
-Lidar DEMs from the USGS at 3m and 1m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
+Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
 

### 2022-01-10 manuscripts/owp_fim4_2021/sections/introduction.tex
edits after odgen/salas comments

--- 2021-12-22 manuscripts/owp_fim4_2021/sections/introduction.tex
@@
 Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
-One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble stream flows at more than 3,000 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
+One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
+AHPS provides forecasting services in the form of ensemble stream flows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
 Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
@@
 On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
-Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons.
+Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons to address the increasingly complex water challenges facing the US.
 %
@@
 \caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
-Also show are the National Water Model stream networks at the full resolution (FR) and Replace and Route (RnR) Mainstems (MS) resolution.}
+Also show are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
 \label{fig:forecast_points}
@@
 Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
-To broaden the forecasting domain, the Office of Water Prediction (OWP) at the National Water Center (NWC) in Tuscaloosa, Alabama commissioned the development of the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
 The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
-The NWM enhances but does not replace the spatial and temporal domain of the current AHPS capabilities at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'. 
-Furthermore, the NWM as of V1.2 has implemented not only assimilation of real-time United States Geological Survey (USGS) stage discharges but also assimilation of flow forecasts from the AHPS forecast points which are in turn routed downstream and updated whenever a new point is reached.
-This configuration of the NWM is known as `Replace and Route' or `RnR' and is used to enhance the forecasting skill of the NWM with the best available regional-scale data.
-Figure \ref{ssec:national_water_model} shows Full-Resolution (FR) modeling stream network as well as the RnR mainstems (MS) network.
+The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
+As a complement to the operational NWM, OWP has also developed a configuration of the NWM that extends RFC forecasts downstream by assimilating and routing forecast streamflows to the next downstream AHPS forecast point.
+This configuration of the NWM is used to enhance forecasting skill by leveraging best available regional-scale data.
+The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
+Figure \ref{ssec:national_water_model} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
 The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
@@
 Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
-As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
+As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
 The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
@@
 \citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant DEM manipulations to make a reality.
+A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
 The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
@@
 %
-Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the National Elevation Dataset (NED) at the 10m scale, HAND has been implemented into software for large-scale, continental computation.
-HAND was initially implemented into operational software by the National Flood Interoperability Experiment (NFIE) to generate FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
+Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
+As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
 NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
 By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
-Originally, NFIE utilized the National Hydrography Dataset (NHD) Plus Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC to produce forecast FIM from the NWM for use within its network of RFCs for additional guidance in hydro-blind regions.
+Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
+The original NFIE dataset was employed by the NWC as an unofficial demonstrations to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
 Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
 More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
-Overall, the software package is estimated to run CONUS at the full-resolution in 0.55 CPU years in a desktop setting.
 
-Related to these, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
+Related to these efforts, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
 Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
-In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30m resolutions \cite{yamazaki2019merit,donchyts2016global}.
+In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30 m resolutions \cite{yamazaki2019merit,donchyts2016global}.
 %
@@
 Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
-Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
+Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3 m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
 The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
@@
 With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
-Here we introduce OWP FIM `Cahaba' that utilizes a few of the latest techniques in HAND based FIM oriented for use with NWM in continental scale operational forecasting settings. 
-This framework is open-source and modular enabling and accelerating the research to operational development pipeline.
+Here we introduce OWP FIM `Cahaba' that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
+This framework is open-source and modular enabling and accelerating the research to operations development pipeline.
 Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
@@
 Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
+This capability will be referred to as Generalized Mainstems (GMS).
 The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.

### 2022-01-10 manuscripts/owp_fim4_2021/sections/materials_and_methods.tex
edits after odgen/salas comments

@@
 To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
-A user only need to install Docker on their host machine and build the image from the provided recipe. 
+A user only needs to install Docker on their host machine and build the image from the provided recipe. 
 Source code is made available for this project on GitHub (https://github.com/NOAA-OWP/cahaba).
@@
 The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
+Functionality is included to turn FIM hydrofabric and streamflows into inundation maps represented in both binary and depths.
+Lastly, a test suite includes means of calculating evaluation metrics compared to a variety of preprocessed test case data.
 A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
@@
 Data sources used within `Cahaba' are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket.
+Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP).
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
 It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for `Cahaba' while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
-For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
+For enforcing levee data into the DEMs, the USACE National Levee Database (NLD) is used to burn feature elevations \cite{engineers2016national}.
 Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{OpenStreetMap,GreatLakesHydrographyDataset}.
-Additionally, the Base Level Engineering (BLE) datasets spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by Region 6 of FEMA \cite{fema2021base,fema2021estimated}.
+Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
 These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
@@
 \hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10m) spatial resolution and vertical units in centimeters \\
+USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters \\
 \hline
@@
 \hline
-NOAA-OWP & Waterbodies considered by the NWM as reservoirs or lakes. \\
+NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. \\
 \hline
@@
 \hline
-FEMA & Flood Inundation Extents & Inundation depths produced by FEMA BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
+FEMA & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
 \hline
@@
 %
-Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and Generalized Mainstems (GMS) networks and 1,604 HUC8s for the MS method.
+Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and GMS networks and 1,604 HUC8s for the MS method. 
 These methods will be explained more later.
-An enhancement of OWP FIM `Cahaba' over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM v2.1 will cover.
+An enhancement of OWP FIM `Cahaba' over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
 %
@@
 %
-The DEM from the NED is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
+The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
 These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
@@
 Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
-It can also be used to simulate more accurate bathymetry which is not accounted for in the NED 10m DEM \cite{gesch2002national}.
+It can also be used to simulate more accurate bathymetry which is not accounted for in the 10 m DEM \cite{gesch2002national}.
 
@@
 The location of the stream network is enforced to ensure general agreement with established stream networks.
-The NHDPlus HR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
+The NHDPlusHR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
 However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
@@
 %
-The National Elevation Dataset at 10m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
-In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10m DEM.
+The DEM at 10 m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
+In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10 m DEM.
 The elevations found in the NLD are burned into the DEM if those elevations were found to exceed those already in place.
@@
 In order to prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
-Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50m.
+Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50 m.
 The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
@@
 Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much geometry variance into one parameterized unit.
+Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
+Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretation of the stream network.
 Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherent the synthetic rating curve properties of the upstream or downstream segment, depending on the topology.
@@
 Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
-Gages are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
-The gages act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
+Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
+The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
 Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
@@
 \label{eq:reach_averaged_mannings_equation}
-\textbf{Q}[y] = \frac{1}{n} \frac{V[y]^{5/3}S^{1/2}}{L B[y]^{2/3}} 
+\textbf{Q} = \frac{1}{n} \frac{V^{5/3}S^{1/2}}{L B^{2/3}} 
 \end{equation}
@@
 where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
+Q, V, and B are taken a specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
 All units are international given the 1 numerator above n.
@@
 For the methods already discussed, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent methods discussed in Sections \ref{sssec:replace_and_route_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
-Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:replace_and_route_mainstems} and \ref{sssec:generalized_mainstems}.
+For two subsequent methods discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
+Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
 While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
@@
 Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
-A `Cahaba' FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (please see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the mainstem of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
+A `Cahaba' FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
+The higher discharge along the MS of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
 This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the mainstem that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the mainstem catchments.
+As a parallel problem, there is excess water accumulated along the MS that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the MS catchments.
 %
@@
 \includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{False negatives associated with confluence of tributaries with mainstems. Integers represent flow values from BLE 100 year event for the associated areas. 
+\caption{False negatives associated with confluence of tributaries with MS. Integers represent flow values from BLE 100 year event for the associated areas. 
 No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Replace and Route Mainstems}
-\label{sssec:replace_and_route_mainstems}
+\subsubsection{NWS Main-stems}
+\label{sssec:nws_mainstems}
 %
-The RnR configuration of the NWM seeks to replace NWM forecasts by assimilating best available AHPS forecasts and routing them downstream.
-The RnR Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
-The RnR MS network, or simply MS, comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
+The Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
+The MS network comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
 It also spans 121,724 reaches across 1,608 HUC8s.
@@
 The inundation derived from this stream network is mosaiced with the inundation from the FR network to form the MS FIMs. 
-Within each HUC, you'll typically only find a MS stream network of stream order 1 but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
+Within each HUC, you'll typically only find a MS stream network of stream order 1 (i.e.
+headwater) but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
 %
@@
 %
-To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call Generalized Mainstems (GMS).
+To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call GMS.
 Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
@@
 For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each level path is buffered to a polygon with a user-available distance parameter of 7km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
+Each level path is buffered to a polygon with a user-available distance parameter of 7 km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
 We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each level path, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
+Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
 Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}c.
@@
 In (a), we illustrate all NWM stream lines symbolized by their level path.
-Meanwhile (b), demonstrates the NED DEM clipped to a 7km buffer around two selected level paths.
+Meanwhile (b), demonstrates the DEM clipped to a 7 km buffer around two selected level paths.
 In (c), we show how HAND can be computed just for each one of these two level paths independently. 
@@
 Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
-Equation \ref{eq:comp_fim} illustrates the same process but for the mosaic the binary inundation.
+Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps.
 %
@@
 %
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced by FEMA region 6 \cite{fema2021base,fema2021estimated}.
+Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
 49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by FEMA so we used those corresponding discharges and mapping extents for evaluation.
+The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM so we used those corresponding discharges and mapping extents for evaluation.
 We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
+By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
 Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
-We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates with each NWM reach. 
+We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
 To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
 This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the FEMA furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
+Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
 While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.

### 2022-01-10 manuscripts/owp_fim4_2021/sections/results.tex
edits after odgen/salas comments

@@
 %
-We produced FIMs for the entire BLE domain within the 49 HUC8 area across several states in the south central US. 
+We produced FIMs for the entire BLE domain within the 49 HUC8 areas across several states in the south central US. 
 The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
@@
 Aggregating the metrics in the method above treats each HUC8 as it's own unit and does little to consider the size differences of the HUCs. 
-In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics} the sum of all the TPs, FPs, and FNs to recompute CSI, POD, and FAR across the entire domain. 
+In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics}  the CSI, POD, and FAR recomputed for the entire domain using the sum of all the TPs, FPs, and FNs. 
 %
@@
 \begin{table}[h!]
-\caption{Primary metrics, TPs, FPs, and FNs, aggregated for BLE domain to recompute CSI, POD, and FAR.}
+\caption{Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated for BLE domain.
+         Best value for across models is highlighted in bold.}
 \label{tab:aggregate_metrics}
@@
 \hline
-\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & 0.5796 & 0.6075 \\
+\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & \textbf{0.5796} & \textbf{0.6075} \\
 \cline{2-8}
-  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & 0.6182 & 0.6435 \\
+  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & \textbf{0.6182} & \textbf{0.6435} \\
 \hline
-\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & 0.6633 & 0.6863 \\
+\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & \textbf{0.6633} & \textbf{0.6863} \\
 \cline{2-8}
-  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & 0.7606 & 0.7810 \\
+  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & \textbf{0.7606} & \textbf{0.7810} \\
 \hline
-\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1778 & 0.1589 & 0.1787 & 0.1589 \\
+\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1787 & \textbf{0.1589} & \textbf{0.1778} & \textbf{0.1589} \\
 \cline{2-8}
-  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & 0.2324 & 0.2148 \\
+  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & \textbf{0.2324} & \textbf{0.2148} \\
 \hline
@@
 Using CSI as an overall proxy for skill of the FIM, we note that generally speaking the skill is correlated with a reduction of the stream orders of the processing units used for HAND.
-In other words, the more you derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
+In other words, the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
 While this trend is evident for both sets of Manning's n values, the trend is slightly more significant for the higher value of 0.12.

### 2022-01-14 manuscripts/owp_fim4_2021/sections/abstract_plainTextSummary.tex
abstract rewrite

@@
 \begin{abstract}
-The National Water Model Version 2.1 (NWM V2.1) seeks to expand upon existing systems to provide up to hourly forecast discharges for four time horizons (1, 3, 10, and 30 days) at over 2.7 million locations throughout the continental United States, Hawaii and Puerto Rico.
-Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D fluvial inundation extents used for forecasting purposes to aid in the protection of life and property.
-Height Above Nearest Drainage (HAND) coupled with the use of synthetic rating curves can convert discharges into stages and stages to inundation extents in a rapid manner across large spatial domains and at high resolutions.
-The Office of Water Prediction Flood Inundation Mapping (OWP FIM) model `Cahaba' is an enhancement of the HAND method developed to improve upon some of HAND's limitations.
-Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, enforcement of levee elevation data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and simplified Manning's n variation.
-Most importantly, we detail an issue previously mentioned with HAND as its FIM skill is dependent on a drainage threshold parameter used to delineate stream lines.
-Since our modeling assumptions are different, we illustrate how reducing Horton-Strahler stream order of the stream networks used within each HAND processing unit down to unary can improve FIM skill.
-This is accomplished by deriving level paths for the National Water Model stream network and utilizing those definitions for deriving HAND datasets independently from other level paths.
-`Cahaba' was evaluated at 49 Base Level Engineering (BLE) sites furnished by the Interagency Flood Risk Management (InFIRM).
-The modeled discharges for 100 and 500 year events from the BLE cross-sections were used for input to OWP FIM instead of NWM discharges to isolate out factors related to hydro-meteorology.
-The resulting maps are compared to the modeled extents from the BLE and binary contingency statistics are computed.
-Across the 49 sites, Critical Success Index (CSI) was found to increase across FIM versions driven by improvements to the Probability of Detection (POD) and slight reduction in False Alarm Ratio (FAR).
-Each successive version of HAND here utilizes greater stream order reduction illustrating a key limitation of HAND and enhanced agreement when utilizing stream networks of lower stream orders as a datum for HAND.
-Additionally we detail the effects of two different Manning's n values and how order reduction drives rating curve bias downward due to the nature of reach averaged synthetic rating curves.
-These capabilities were implemented in an open-source package available on GitHub (\url{https://github.com/NOAA-OWP/cahaba}) for continued use and collaboration with the research community.
-%
+The National Water Model (NWM) currently requires the post-processing of forecast discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
+Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is worthy of producing high-resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
+However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
+To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that reduces the Horton-Strahler stream order of the stream network.
+GMS contains all segments within the NWM stream network but instead of deriving HAND by accounting for all river segments at once, it is derived independently at the level path (LP) scale.
+LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
+These FIMs are then mosaiced together, effectively turning the stream network into discrete groups of homogenous unit stream order by removing the influence of neighboring tributaries.
+Improvement in mapping skill is observed when compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
+A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
 \end{abstract}

### 2022-01-14 manuscripts/owp_fim4_2021/sections/keypoints.tex
latest changes
abstract rewrite

@@
 \begin{keypoints}
-\item Height Above Nearest Drainage (HAND) is a terrain index used for rapid generation of large scale flood inundation maps at high resolutions.
-\item Despite its advantages, FIM derived from HAND and synthetic rating curves is subject to limitations at river junctions.
-\item A means of resolving this limitation is provided by reducing HAND processing units to level-paths of unit stream order.
+\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
+\item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
+\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation.
 \end{keypoints}

### 2022-01-14 manuscripts/owp_fim4_2021/sections/title_authors.tex
Manuscript edits made after feedback from co-authors
edits after odgen/salas comments
abstract rewrite

@@
 %
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{1,3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
+\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
 %
 %
-\affiliation{1}{Lynker Technologies, Leesburg, VA, USA}
+\affiliation{1}{Lynker, Leesburg, VA, USA}
 \affiliation{2}{Center for Remote Sensing, Agricultural and Biological Engineering Department, University of Florida, Gainesville, FL, USA}
-\affiliation{3}{National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
+\affiliation{3}{National Water Center, Office of Water Prediction, National Weather Service, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
 \affiliation{4}{New York Water Science Center, Hydrologic Applied Innovations Lab, United States Geological Survey, Troy, NY, USA}

### 2022-01-20 manuscripts/owp_fim4_2021/sections/introduction.tex
Removed references to cahaba

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{OWP FIM `Cahaba'}
+\subsection{OWP FIM}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
-Here we introduce OWP FIM `Cahaba' that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
+Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
 This framework is open-source and modular enabling and accelerating the research to operations development pipeline.

### 2022-01-20 manuscripts/owp_fim4_2021/sections/materials_and_methods.tex
Removed references to cahaba

@@
 %
-OWP FIM `Cahaba' is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
+OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
 %
@@
 %
-OWP FIM `Cahaba' exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
+OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
 Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
@@
 
-The `Cahaba' pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
+The pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
 Preprocessing acquires and prepares datasets for production of the FIM hydrofabric. 
@@
 %
-Data sources used within `Cahaba' are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
+Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
 Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP).
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for `Cahaba' while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
+It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
 For enforcing levee data into the DEMs, the USACE National Levee Database (NLD) is used to burn feature elevations \cite{engineers2016national}.
@@
 These methods will be explained more later.
-An enhancement of OWP FIM `Cahaba' over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
+An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
 %
@@
 
-Specifically within the context of `Cahaba', the hydro-conditioning operations that take place in sequential order are presented. 
+Specifically within the context of OWP FIM, the hydro-conditioning operations that take place in sequential order are presented. 
 Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing units of size HUC8. 
@@
 By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
-`Cahaba' utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
+OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
 %
@@
 Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
-A `Cahaba' FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
+An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
 The higher discharge along the MS of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 

### 2022-01-20 manuscripts/owp_fim4_2021/dependencies/texcount.pl

@@ -1,0 +1,4247 @@
+#! /usr/bin/env perl
+use strict;
+use warnings;
+use utf8; # Because the script itself is UTF-8 encoded
+use Encode;
+use Text::Wrap;
+use Term::ANSIColor;
+
+##### Version information
+my $versionnumber="3.2";
+my $versiondate="2020 Aug 04";
+
+
+# System variables
+my $terminalwidth;
+
+# Conditional package inclusion
+if ($^O=~/^MSWin/) {
+  eval {
+    require Win32::Console::ANSI;
+    import Win32::Console::ANSI;
+  };
+  if ($@) {
+    option_ansi_colours(0);
+    print STDERR "NOTE: Package Win32::Console::ANSI required for colour coded output.\n";
+  }
+}
+
+# Terminal or not
+if (-t STDOUT) { # If in terminal
+  eval {
+    require Term::ReadKey;
+    import Term::ReadKey;
+    ($terminalwidth)=GetTerminalSize();
+  };
+} else {
+  option_ansi_colours(0);
+}
+
+if (!defined $terminalwidth) {$terminalwidth=76;}
+elsif ($terminalwidth<60) {$terminalwidth=60;}
+elsif ($terminalwidth>120) {$terminalwidth=120;}
+
+
+###### Set global constants and names
+
+### Global data about TeXcount
+my %GLOBALDATA=
+   ('versionnumber'  => $versionnumber
+   ,'versiondate'    => $versiondate
+   ,'maintainer'     => 'Einar Andreas Rodland'
+   ,'copyrightyears' => '2008-2020'
+   ,'website'        => 'https://app.uio.no/ifi/texcount/'
+   );
+
+### Named constants
+my ($INCLUDE_NONE,$INCLUDE_AFTER,$INCLUDE_MERGE)=(0,1,2); # Value of $includeTeX
+my ($HTML_NONE,$HTML_CORE,$HTML_FULL)=(0,1,2); # Value of $htmlstyle
+
+
+
+###### Set global settings and variables
+
+### Options and states
+
+# Outer object (for error reports not added by a TeX object)
+my $Main=getMain();
+
+# Global options and settings
+my $htmlstyle=$HTML_NONE; # Flag to print HTML
+my $texcodeoutput=0; # Flag to convert output to valid TeX text
+my $encoding=undef; # Selected input encoding (default will be guess)
[…]

### 2022-01-23 manuscripts/owp_fim4_2022/dependencies/trackchanges-0.7.0/testing/testing_edit.tex
minor changes

--- 2020-09-11 manuscripts/owp_fim3_2020/templates/April-30-2019-latex-templates/April 16 2019/trackchanges-0.7.0/testing/testing_old.tex
@@
 
-\emph{Cut to a courtroom. Severe atmosphere.}
+\annote[novi]{\emph{Cut to a courtroom. Severe atmosphere.}}{In the Flying Circus there is never a ``severe'' atmosphere.}
 
@@
 
-Michael Norman Randall, you have been found guilty of the murder of Arthur Reginald Webster, Charles Patrick Trumpington, Marcel Agnes Bernstein, Lewis Anona Rudd, John Malcolm Kerr, Nigel Sinclair Robinson, Norman Arthur Potter, Felicity Jayne Stone, Jean-Paul Reynard, Rachel Shirley Donaldson, Stephen Jay Greenblatt, Karl-Heinz Mullet, Belinda Anne Ventham, Juan-Carlos Fernandez, Thor Olaf Stensgaard, Lord Kimberley of Pretoria, Lady Kimberley of Pretoria, The Right Honourable Nigel Warmsly Kimberley, Robert Henry Noonan and Felix James Bennett, on or about the morning of the 19th December 1972.
+Michael Norman Randall, you have been found guilty of the murder of Arthur Reginald Webster, Charles Patrick Trumpington, Marcel Agnes Bernstein, Lewis Anona Rudd, John Malcolm Kerr, Nigel Sinclair Robinson, \change[ym]{Norman Arthur Potter}{Harry Potter}, \add[novi]{Thing 1 \& Thing 2, }\add[ym]{Humpty Dumpty, }\add[three]{Superman, }\add[quatro]{Cold Fusion, }\add[pyat]{The Lorax, }\add[one more]{Mr.
+T, }\add{The Dread Pirate Roberts, }Felicity Jayne Stone, Jean-Paul Reynard, Rachel Shirley Donaldson, Stephen Jay Greenblatt, Karl-Heinz Mullet, Belinda Anne Ventham, Juan-Carlos Fernandez, Thor Olaf Stensgaard, Lord Kimberley of Pretoria\note{Isn't Pretoria in South Africa?}, \remove{Lady Kimberley of Pretoria, }The Right Honourable Nigel Warmsly Kimberley, Robert Henry Noonan\add[novi]{, Your Mom} and Felix James Bennett, on or about the morning of the 19th December 1972\refneeded[novi]{}.
 Have you anything to say before I pass sentence?
@@
 
-Yes, sir. I'm very sorry.
+\annote[ym]{Yes, sir. I'm very sorry.}{That was short.}
 
@@
 
-\noindent\\ Second Policeman:
+\noindent\\ \annote{Second Policeman:
 
@@
 
-It was plain sailing apart from the arrest.
+It was plain sailing apart from the arrest.}{That could have been left out.}
 
@@
 
-\noindent\\ Counsel:
+\remove[ym]{\noindent\\ Counsel:
 
@@
 
-Oh thank you. (very chuffed)
+Oh thank you. (very chuffed)}
 
@@
 
-No, no, it was very interesting.
+No, no, it was very \change{interesting}{fascinating}.
 

### 2022-01-23 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
minor changes

--- 2022-01-20 manuscripts/owp_fim4_2021/sections/materials_and_methods.tex
@@
 \hline
-FEMA & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
+InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
 \hline
-FEMA & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
+InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
 \hline

### 2022-01-23 manuscripts/owp_fim4_2021/docker_run_and_make.sh

@@ -1,5 +1,0 @@
-#!/bin/bash -e
-
-codeDir=$1
-
-docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex:20211220 /foss_fim/manuscripts/owp_fim4_2021/Makefile.sh

### 2022-01-23 manuscripts/owp_fim4_2022/docker_run_and_make.sh

@@ -1,0 +1,5 @@
+#!/bin/bash -e
+
+codeDir=$1
+
+docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex:20211220 /foss_fim/manuscripts/owp_fim4_2022/Makefile.sh

### 2022-02-06 manuscripts/owp_fim4_2022/sections/conclusions.tex
conclusion update

--- 2022-01-10 manuscripts/owp_fim4_2021/sections/conclusions.tex
@@
 Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
-Due to inherent limitations with HAND, scalable physics-based methods will need to be worked on to provide better representation of flood extent dynamics in steady and unsteady conditions. 
+Due to inherent limitations with HAND, scalable, physics-based methods are necessary to consider to provide a better representation of flood extent dynamics in steady and unsteady conditions. 
 %

### 2022-02-09 manuscripts/owp_fim4_2022/sections/conclusions.tex
wrr initial submission

@@
 %
-Floods present a significant, underserved, and expanding risk to life, property, and resources.
+Floods present a significant, under-served, and expanding risk to life, property, and resources.
 Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
@@
 HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
-Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico (https://github.com/NOAA-OWP/cahaba).
+Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
 HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.

### 2022-02-09 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
wrr initial submission

@@
 A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub (https://github.com/NOAA-OWP/cahaba).
+Source code is made available for this project on GitHub \cite{inundationMapping2022}.
 
@@
 Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP).
+Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP) \cite{esipData2022}.
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 

### 2022-02-09 manuscripts/owp_fim4_2022/Makefile.sh

@@ -12,6 +12,9 @@ bibtex "$baseName".aux
 pdflatex "$fullFile"
 pdflatex "$fullFile"
 
+# expands all tex files into one file
+echo 'Lat-Expand to Create One Tex File ...'
+/usr/bin/latexpand/latexpand --keep-comments --biber bibliography/owp_fim4_2021.bib /foss_fim/manuscripts/owp_fim4_2022/owp_fim4.tex -o owp_fim4_expanded.tex
 
 
 

### 2022-02-09 manuscripts/owp_fim4_2022/docker_run_and_make.sh

@@ -2,4 +2,4 @@
 
 codeDir=$1
 
-docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex:20211220 /foss_fim/manuscripts/owp_fim4_2022/Makefile.sh
+docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex:20220201 /foss_fim/manuscripts/owp_fim4_2022/Makefile.sh

### 2022-02-14 manuscripts/owp_fim4_2022/sections/abstract_plainTextSummary.tex
spelling updates

--- 2022-01-14 manuscripts/owp_fim4_2021/sections/abstract_plainTextSummary.tex
@@
 LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
-These FIMs are then mosaiced together, effectively turning the stream network into discrete groups of homogenous unit stream order by removing the influence of neighboring tributaries.
+These FIMs are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
 Improvement in mapping skill is observed when compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
@@
 We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-We compared these HAND derived FIMs to maps from more realistic models and found and improvement in mapping performance.
+We compared these HAND derived FIMs to maps from more realistic models and found improvement in mapping performance.
 %

### 2022-02-14 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
adding expanded latex file from last change

--- 2022-02-09 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
+%
+% This file includes commands and instructions
+% given in the order necessary to produce a final output that will
+% satisfy AGU requirements, including customized APA reference formatting.
+%
+% You may copy this file and give it your
+% article name, and enter your text.
+%
+%
+% Step 1: Set the \documentclass
+%
+%
+
+%% To submit your paper:
+\documentclass[draft]{dependencies/agujournal2019}
+\usepackage{url} %this package should fix any errors with URLs in refs.
+\usepackage{lineno}
+\usepackage[inline]{dependencies/trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
+\usepackage{soul}
+
+%% added
+\usepackage{multirow}
+\usepackage{amsmath}
+
+\linenumbers
+
+%%%%%%%
+% As of 2018 we recommend use of the TrackChanges package to mark revisions.
+% Add track changes editors
+\addeditor{Fernando Aristizabal}
+
+% The trackchanges package adds five new LaTeX commands:
+%
+%  \note[editor]{The note}
+%  \annote[editor]{Text to annotate}{The note}
+%  \add[editor]{Text to add}
+%  \remove[editor]{Text to remove}
+%  \change[editor]{Text to remove}{Text to add}
+%
+% complete documentation is here: http://trackchanges.sourceforge.net/
+%%%%%%%
+
+\drafttrue
+
+%% Enter journal name below.
+%% Choose from this list of Journals:
+%
+% JGR: Atmospheres
+% JGR: Biogeosciences
+% JGR: Earth Surface
+% JGR: Oceans
+% JGR: Planets
+% JGR: Solid Earth
+% JGR: Space Physics
+% Global Biogeochemical Cycles
+% Geophysical Research Letters
+% Paleoceanography and Paleoclimatology
+% Radio Science
+% Reviews of Geophysics
+% Tectonics
+% Space Weather
+% Water Resources Research
+% Geochemistry, Geophysics, Geosystems
+% Journal of Advances in Modeling Earth Systems (JAMES)
+% Earth's Future
+% Earth and Space Science
+% Geohealth
+%
+% ie, \journalname{Water Resources Research}
+
+\journalname{Water Resources Research}
+
+% starts document
+\begin{document}
+
+%% Title and Authors
+%% ------------------------------------------------------------------------ %%
+%  Title
+%
+% (A title should be specific, informative, and brief. Use
+% abbreviations only if they are defined in the abstract. Titles that
+% start with general keywords then specific terms are optimized in
+% searches)
+%
+%% ------------------------------------------------------------------------ %%
+
+% Example: \title{This is a test title}
+%
+\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  AUTHORS AND AFFILIATIONS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Authors are individuals who have significantly contributed to the
+% research and preparation of the article. Group authors are allowed, if
+% each author in the group is separately identified in an appendix.)
+
+% List authors by first name or initial followed by last name and
+% separated by commas. Use \affil{} to number affiliations, and
+% \thanks{} for author notes.
+% Additional author notes should be indicated with \thanks{} (for
+% example, for current addresses).
+
+% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
+% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
+%
+%
+\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
+%
+%
+\affiliation{1}{Lynker, Leesburg, VA, USA}
+\affiliation{2}{Center for Remote Sensing, Agricultural and Biological Engineering Department, University of Florida, Gainesville, FL, USA}
+\affiliation{3}{National Water Center, Office of Water Prediction, National Weather Service, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
+\affiliation{4}{New York Water Science Center, Hydrologic Applied Innovations Lab, United States Geological Survey, Troy, NY, USA}
+\affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Maryland, College Park, MD, USA}
+%
+%(repeat as many times as is necessary)
+%
+%% Corresponding Author:
+% Corresponding author mailing address and e-mail address:
+
+% (include name and email addresses of the corresponding author.  More
+% than one corresponding author is allowed in this LaTeX file and for
+% publication; but only one corresponding author is allowed in our
+% editorial system.)
+%
+% Example: \correspondingauthor{First and Last Name}{email@address.edu}
+%
+\correspondingauthor{Fernando Aristizabal}{fernando.aristizabal@noaa.gov}
+
+%% Keypoints
+%% Keypoints, final entry on title page.
+
+%  List up to three key points (at least one is required)
+%  Key Points summarize the main points and conclusions of the article
+%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
+
+% Example:
+% \begin{keypoints}
+% \item	List up to three key points (at least one is required)
+% \item	Key Points summarize the main points and conclusions of the article
+% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
+% \end{keypoints}
+%
+\begin{keypoints}
+\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
+\item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
+\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation.
+\end{keypoints}
+%
+
+%% Abstract
+% ------------------------------------------------------------------------ %%
+%
+%  ABSTRACT and PLAIN LANGUAGE SUMMARY
+%
+% A good Abstract will begin with a short description of the problem
+% being addressed, briefly describe the new data or analyses, then
+% briefly states the main conclusion(s) and how they are supported and
+% uncertainties.
+
+% The Plain Language Summary should be written for a broad audience,
+% including journalists and the science-interested public, that will not have 
+% a background in your field.
+%
+% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
+% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
+% see http://sharingscience.agu.org/creating-plain-language-summary/)
+%
+%% ------------------------------------------------------------------------ %%
+
+%% \begin{abstract} starts the second page
+
+\begin{abstract}
+The National Water Model (NWM) currently requires the post-processing of forecast discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
+Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is worthy of producing high-resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
+However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
+To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that reduces the Horton-Strahler stream order of the stream network.
+GMS contains all segments within the NWM stream network but instead of deriving HAND by accounting for all river segments at once, it is derived independently at the level path (LP) scale.
+LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
+These FIMs are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
+Improvement in mapping skill is observed when compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
+A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
+\end{abstract}
+%
+\section*{Plain Language Summary}
+Flooding is one of the most impactful natural disasters on life and property.
+The United States National Water Model (NWM) provides flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
+In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
+This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
+We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
+We compared these HAND derived FIMs to maps from more realistic models and found improvement in mapping performance.
+%
+
+%%% Suggested section heads:
+% \section{Introduction}
+%
+% The main text should start with an introduction. Except for short
+% manuscripts (such as comments and replies), the text should be divided
+% into sections, each with its own heading.
+
+% Headings should be sentence fragments and do not begin with a
+% lowercase letter or number. Examples of good headings are:
+
+% \section{Materials and Methods}
+% Here is text on Materials and Methods.
+%
+% \subsection{A descriptive heading about methods}
+% More about Methods.
+%
+% \section{Data} (Or section title might be a descriptive heading about data)
+%
+% \section{Results} (Or section title might be a descriptive heading about the
+% results)
+%
+% \section{Conclusions}
+
+% Introduction
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
+In 2017 and 2019, river and flash flooding combined represented the leading cause of death and the second leading cause in 2018 among all natural disasters in the US \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
+More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
+With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
+Trends related to flood damages and fatalities have been steadily increasing over recent decades.
+\cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
+Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
+Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Operational Forecasting}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
+One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
+AHPS provides forecasting services in the form of ensemble stream flows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
+Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
+AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
+On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
+Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons to address the increasingly complex water challenges facing the US.
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/forecast_points.jpg}
+\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
+Also show are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
+\label{fig:forecast_points}
+\end{figure}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{National Water Model}
+\label{ssec:national_water_model}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
+In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
+The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
+As a complement to the operational NWM, OWP has also developed a configuration of the NWM that extends RFC forecasts downstream by assimilating and routing forecast streamflows to the next downstream AHPS forecast point.
+This configuration of the NWM is used to enhance forecasting skill by leveraging best available regional-scale data.
+The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
+Figure \ref{ssec:national_water_model} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
+The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
+
+The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
+The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
+The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
+Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
+As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
+The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Height Above Nearest Drainage}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
+\citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
+A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
+The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
+\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
+This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{HAND Implementations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
+As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
+NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
+By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
+Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
+The original NFIE dataset was employed by the NWC as an unofficial demonstrations to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
+Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
+More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
+
+Related to these efforts, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
+Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
+In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30 m resolutions \cite{yamazaki2019merit,donchyts2016global}.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{OWP FIM}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Many of those in the academic community assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
+\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
+The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
+Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
+Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3 m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
+The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
+Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
+Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
+Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
+These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.
+
+With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
+Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
+This framework is open-source and modular enabling and accelerating the research to operations development pipeline.
+Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
+In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
+Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
+Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
+This capability will be referred to as Generalized Mainstems (GMS).
+The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
+%
+ %
+% Materials and Methods
+\clearpage % this clears figures before references
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
+ %
+% Results
+\clearpage % this clears figures before references
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+\label{sec:results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Mapping Performance}
+\label{ssec:mapping_performance}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+We produced FIMs for the entire BLE domain within the 49 HUC8 areas across several states in the south central US. 
+The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
+We computed the statistics (CSI, POD, and FAR) for both 100 and 500 year events for Mannings N set to 0.06 and 0.12. 
+The distribution of these statistics can be examined in Figure \ref{fig:violin_plot} as violin plots.
+Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
+We also denote trend lines for each metric and Manning's n setting as well as their respective slope estimate and one-tailed p-value denoting the level of significance of the trend.
+
+Aggregating the metrics in the method above treats each HUC8 as it's own unit and does little to consider the size differences of the HUCs. 
+In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics}  the CSI, POD, and FAR recomputed for the entire domain using the sum of all the TPs, FPs, and FNs. 
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.9]{figures/violin_plots.jpg}
+\caption{Shows kernel density estimation of the distributions (sample size = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
+The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's n (0.06 and 0.12).
+Trend lines for each metric - Mannings combination are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}
+\label{fig:violin_plot}
+\end{figure}
+%
+\begin{table}[h!]
+\caption{Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated for BLE domain.
+         The best value across models is highlighted in bold.}
+\label{tab:aggregate_metrics}
+\centering
+%\begin{tabular}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
+\begin{tabular}{|c|c||c|c|c|c|c|c|}
+\hline
+\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{MS} & \multicolumn{2}{|c|}{GMS} \\
+\cline{3-8}
+  &  & 100yr & 500yr & 100yr & 500yr & 100yr & 500yr \\
+\hline
+\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & \textbf{0.5796} & \textbf{0.6075} \\
+\cline{2-8}
+  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & \textbf{0.6182} & \textbf{0.6435} \\
+\hline
+\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & \textbf{0.6633} & \textbf{0.6863} \\
+\cline{2-8}
+  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & \textbf{0.7606} & \textbf{0.7810} \\
+\hline
+\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1787 & \textbf{0.1589} & \textbf{0.1778} & \textbf{0.1589} \\
+\cline{2-8}
+  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & \textbf{0.2324} & \textbf{0.2148} \\
+\hline
+\end{tabular}
+\end{table}
+%
+% Interpretation of metrics
+From Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, we denote several meaningful trends. 
+Using CSI as an overall proxy for skill of the FIM, we note that generally speaking the skill is correlated with a reduction of the stream orders of the processing units used for HAND.
+In other words, the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
+While this trend is evident for both sets of Manning's n values, the trend is slightly more significant for the higher value of 0.12.
+Other trends related to this Figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
+We also note how the higher Manning's n value enhances performance for both of these recurrence intervals across all models.
+
+Dissecting the improvements and trends presented in the previous paragraph comes down mostly to improvement in POD or a reduction in absolute amount of FNs.
+POD being the primary driver in skill enhancement is evident across models by comparing the slope of the POD lines with the slope of the FAR lines.
+Even though aggregating metrics by HUC8 yields a statistically zero trend, one does see a slight reduction in FAR across models that reduce HAND's maximum stream order.
+Additionally, we note that POD is a primary driver in enhancing performance across Manning's n values as well.
+This significant improvement comes at a cost of false alarms as the FAR increases significantly across Manning's n values.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Computational Performance}
+\label{ssec:compuational_performance}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+The NFIE experiments were able to produce HAND for 331 HUC6's in 1.34 CPU years \cite{liu2016cybergis} and estimates using work from \citeA{djokic2019arc} put producing HAND at the FR NWM at 0.55 CPU years.
+For our work, we were able to produce HAND at the full NWM resolution in 0.13 CPU years which represents a substantial speed-up compared to previous works.
+For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
+GMS which generalizes HAND production to level path scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
+%
+ %
+% Discussion
+\clearpage % this clears figures before references
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+\label{sec:discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Overall, we note a positive relationship between FIM skill and a reduction of the stream order of the stream network we use to derive the HAND datasets.
+Most of this change is accounted for by increasing POD thus reducing FNs especially along higher order rivers with higher flow magnitudes.
+We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
+This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
+One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
+
+In analyzing the data, we found a slight reduction in FAR was detected and more digging pointed to a bias in rating curves introduced by stream order reduction.
+Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
+Sub-figure \ref{fig:rating_curve_comparison}a shows how the average rating curves for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to bias down (and to the right) with ever increasing stream order reduction (FR to MS to GMS).
+This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
+Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
+Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
+While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
+Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius.
+Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively (see Equation \ref{eq:reach_averaged_mannings_equation}).
+Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
+Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
+Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
+We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
+In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
+Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
+Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments that encompass neighboring tributaries.
+These tributaries have a significant amount of bathymetry that is low-lying thus easily including the SRC derivation. 
+They also contribute volume and bed area that is technically not perpendicular to the flux of streamflow being accounted for in the stream in question. 
+Careful examination of Figure \ref{fig:gms_enhancement}b shows how much larger catchments include neighboring tributaries and the geometry associated with those tributaries. 
+This geometry is not perpendicular to the flow that is associated with the main reach thus leading to biases in the SRC.
+We consider this fact to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
+
+Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
+While the cause of these errors can be varied, we assert here that conflating 4 networks for use in evaluations leads to significant error.
+As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
+One of the issues is when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
+The utilization of MS and GMS only conflates to NWM catchments directly associated with the level path in question which is inherently easy to do with those methods. 
+Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
+The NWM stream network was derived using the NHD medium resolution dataset which was derived from coarser DEMs than those used here. 
+Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
+Until a singular stream network is used for the NWM, BLE benchmark, and for HAND based FIM, conflation will continue being a source of error.
+
+Our qualitative analysis suggests that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM for future development.
+The bathymetry of the 10 m DEM from 3DEP is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
+Manning's n which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
+Some adjusting parameter must be introduced to enhance the estimation of the bathymetric representation.
+Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
+
+Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
+For HAND to be derived and thus create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
+In other words, an entire area eligible for flooding must monotonically decrease in elevation. 
+DEM's naturally don't do this and the dynamics of true flood events don't follow drainage patterns.
+Enforcing this assumption for HAND leads to significant amount of DEM manipulations that introduce basic errors.
+These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
+Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/rating_curve_comparison.jpg}
+\caption{Illustrates average quantities for the three methods, FR, MS, and GMS, for each stage value (m). 
+The values are (a) Discharge $m^3s^{-1}$, (b) Volume $m^3$, (c) Bed Area $m^2$, (d) a function of Volume and Bed Area, and (e) number of pixels.
+}
+\label{fig:rating_curve_comparison}
+\end{figure}
+%
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
+\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105.
+Catchment boundaries and stream lines are shown in white and dotted green, respectively.
+Sub-figure (a) shows agreement of FR HAND denoting significant areas of under-prediction due to junctions and catchment boundaries.
+Meanwhile, (b) shows the agreement for GMS and much larger catchments leading to much better inundation agreement for this given reach. 
+Overall, this illustrates the benefits of stream order reduction for deriving HAND datasets.
+}
+\label{fig:gms_enhancement}
+\end{figure}
+%
+ %
+% Conclusions
+\clearpage % this clears figures before references
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+\label{sec:conclusions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Floods present a significant, under-served, and expanding risk to life, property, and resources.
+Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
+The National Water Model (NWM) developed by the National Oceanic and Atmospheric Administration's Office of Water Prediction, along with partners, provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
+Additional processing is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
+Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
+HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
+Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
+HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
+Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
+We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
+This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
+FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
+Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.
+Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
+Due to inherent limitations with HAND, scalable, physics-based methods are necessary to consider to provide a better representation of flood extent dynamics in steady and unsteady conditions. 
+%
+ %
+%%
+
+%  Numbered lines in equations:
+%  To add line numbers to lines in equations,
+%  \begin{linenomath*}
+%  \begin{equation}
+%  \end{equation}
+%  \end{linenomath*}
+
+%% Enter Figures and Tables near as possible to where they are first mentioned:
+%
+% DO NOT USE \psfrag or \subfigure commands.
+%
+% Figure captions go below the figure.
+%\section{= enter section title =}
+%Text here ===>>>
+%\section{= enter section title =}
+%Text here ===>>>
+%%
+
+%  Numbered lines in equations:
+%  To add line numbers to lines in equations,
+%  \begin{linenomath*}
+%  \begin{equation}
+%  \end{equation}
+%  \end{linenomath*}
+
+%% Enter Figures and Tables near as possible to where they are first mentioned:
+%
+% DO NOT USE \psfrag or \subfigure commands.
+%
+% Figure captions go below the figure.
+% Table titles go above tables;  other caption information
+%  should be placed in last line of the table, using
+% \multicolumn2l{$^a$ This is a table note.}
+%
+%----------------
+% EXAMPLE FIGURES
+%
+% \begin{figure}
+% \includegraphics{example.png}
+% \caption{caption}
+% \end{figure}
+%
+% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
+% \begin{figure}
+% \noindent\includegraphics[width=\textwidth]{anothersample.png}
+%\caption{caption}
+%\label{pngfiguresample}
+%\end{figure}
+%
+%
+% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
+This is common when trying to add a PDF figure without pdflatex.
+% \begin{figure}
+% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
+%\caption{caption}
+%\label{pdffiguresample}
+%\end{figure}
+%
+%
+% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
+%
+
+%
+% ---------------
+% EXAMPLE TABLE
+%
+% \begin{table}
+% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
+% \centering
+% \begin{tabular}{l c}
+% \hline
+%  Run  & Time (min)  \\
+% \hline
+%   $l1$  & 260   \\
+%   $l2$  & 300   \\
+%   $l3$  & 340   \\
+%   $h1$  & 270   \\
+%   $h2$  & 250   \\
+%   $h3$  & 380   \\
+%   $r1$  & 370   \\
+%   $r2$  & 390   \\
+% \hline
+% \multicolumn{2}{l}{$^{a}$Footnote text here.}
+% \end{tabular}
+% \end{table}
+
+%% SIDEWAYS FIGURE and TABLE
+% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
+%
+% \begin{sidewaysfigure}
+% \includegraphics[width=20pc]{figsamp}
+% \caption{caption here}
+% \label{newfig}
+% \end{sidewaysfigure}
+%
+%  \begin{sidewaystable}
+%  \caption{Caption here}
+% \label{tab:signif_gap_clos}
+%  \begin{tabular}{ccc}
+% one&two&three\\
+% four&five&six
+%  \end{tabular}
+%  \end{sidewaystable}
+
+%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
+%\begin{linenomath*}
+%\begin{equation}
+%y|{f} \sim g(m, \sigma),
+%\end{equation}
+%\end{linenomath*}
+
+%%% End of body of article
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional Appendix goes here
+%
+% The \appendix command resets counters and redefines section heads
+%
+% After typing \appendix
+%
+%\section{Here Is Appendix Title}
+% will show
+% A: Here Is Appendix Title
+%
+%\appendix
+%\section{Here is a sample appendix}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Optional Glossary, Notation or Acronym section goes here:
+%
+%%%%%%%%%%%%%%
+% Glossary is only allowed in Reviews of Geophysics
+%  \begin{glossary}
+%  \term{Term}
+%   Term Definition here
+%  \term{Term}
+%   Term Definition here
+%  \term{Term}
+%   Term Definition here
+%  \end{glossary}
+
+%
+%%%%%%%%%%%%%%
+% Acronyms
+%   \begin{acronyms}
+%   \acro{Acronym}
+%   Definition here
+%   \acro{EMOS}
+%   Ensemble model output statistics
+%   \acro{ECMWF}
+%   Centre for Medium-Range Weather Forecasts
+%   \end{acronyms}
+
+%
+%%%%%%%%%%%%%%
+% Notation
+%   \begin{notation}
+%   \notation{$a+b$} Notation Definition here
+%   \notation{$e=mc^2$}
+%   Equation in German-born physicist Albert Einstein's theory of special
+%  relativity that showed that the increased relativistic mass ($m$) of a
+%  body comes from the energy of motion of the body—that is, its kinetic
+%  energy ($E$)—divided by the speed of light squared ($c^2$).
+%   \end{notation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%  ACKNOWLEDGMENTS
+%
+% The acknowledgments must list:
+%
+% >>>>	A statement that indicates to the reader where the data
+% 	supporting the conclusions can be obtained (for example, in the
+% 	references, tables, supporting information, and other databases).
+%
+% 	All funding sources related to this work from all authors
+%
+% 	Any real or perceived financial conflicts of interests for any
+%	author
+%
+% 	Other affiliations for any author that may be perceived as
+% 	having a conflict of interest with respect to the results of this
+% 	paper.
+%
+%
+% It is also the appropriate place to thank colleagues and other contributors.
+% AGU does not normally allow dedications.
+
+\acknowledgments
+This work was funded by the OWP which is part of NOAA's National Weather Service.
+Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
+We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
+Additionally, David Blodgett from the USGS Water Mission Area was instrumental in helping define level paths and other hydrographic work.
+More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
+Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
+%The [type of data] data used for [brief context, description] in the study are available at [repository, source name] via [DOI, persistent identifier link] with  [license, access conditions] [optional in-text citation in References]
+%[Version number] of the [software name] used for [brief context, description of what the software was used for] is preserved at [DOI, persistent identifier link], available via [license type, access conditions] and developed openly at [software development platform link].* [optional in-text citation in References].
+
+%% ------------------------------------------------------------------------ %%
+%% References and Citations
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% \bibliography{<name of your .bib file>} don't specify the file extension
+%
+% don't specify bibliography style
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\clearpage % this clears figures before references
+\bibliography{bibliography/owp_fim4_2022}
+%
+%Reference citation instructions and examples:
+%
+% Please use ONLY \cite and \citeA for reference citations.
+% \cite for parenthetical references
+% ...as shown in recent studies (Simpson et al., 2019)
+% \citeA for in-text citations
+% ...Simpson et al. (2019) have shown...
+%
+%
+%...as shown by \citeA{jskilby}.
+%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
+%...has been shown \cite{jskilbye}.
+%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
+%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
+%...has been shown by \cite <e.g.,>[and others]{lewin76}.
+%
+% apacite uses < > for prenotes and [ ] for postnotes
+% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
+%
+
+\end{document}
+
+%%%% More Information and Advice:
+
+%% ------------------------------------------------------------------------ %%
+%
+%  SECTION HEADS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Capitalize the first letter of each word (except for
+% prepositions, conjunctions, and articles that are
+% three or fewer letters).
+
+% AGU follows standard outline style; therefore, there cannot be a section 1 without
+% a section 2, or a section 2.3.1 without a section 2.3.2.
+% Please make sure your section numbers are balanced.
+% ---------------
+% Level 1 head
+%
+% Use the \section{} command to identify level 1 heads;
+% type the appropriate head wording between the curly
+% brackets, as shown below.
+%
+%An example:
+%\section{Level 1 Head: Introduction}
+%
+% ---------------
+% Level 2 head
+%
+% Use the \subsection{} command to identify level 2 heads.
+%An example:
+%\subsection{Level 2 Head}
+%
+% ---------------
+% Level 3 head
+%
+% Use the \subsubsection{} command to identify level 3 heads
+%An example:
+%\subsubsection{Level 3 Head}
+%
+%---------------
+% Level 4 head
+%
+% Use the \subsubsubsection{} command to identify level 3 heads
+% An example:
+%\subsubsubsection{Level 4 Head} An example.
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  IN-TEXT LISTS
+%
+%% ------------------------------------------------------------------------ %%
+%
+% Do not use bulleted lists; enumerated lists are okay.
+% \begin{enumerate}
+% \item
+% \item
+% \item
+% \end{enumerate}
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  EQUATIONS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Single-line equations are centered.
+% Equation arrays will appear left-aligned.
+
+Math coded inside display math mode \[ ...\]
+ will not be numbered, e.g.,:
+ \[ x^2=y^2 + z^2\]
+
+ Math coded inside \begin{equation} and \end{equation} will
+ be automatically numbered, e.g.,:
+ \begin{equation}
+ x^2=y^2 + z^2
+ \end{equation}
+
+% To create multiline equations, use the
+% \begin{eqnarray} and \end{eqnarray} environment
+% as demonstrated below.
+\begin{eqnarray}
+  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
+        && + (y - y_{0}) \sin \Theta  \nonumber \\
+  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
+        && + (y - y_{0}) \cos \Theta.
+\end{eqnarray}
+
+%If you don't want an equation number, use the star form:
+%\begin{eqnarray*}...\end{eqnarray*}
+
+% Break each line at a sign of operation
+% (+, -, etc.) if possible, with the sign of operation
+% on the new line.
+
+% Indent second and subsequent lines to align with
+% the first character following the equal sign on the
+% first line.
+
+% Use an \hspace{} command to insert horizontal space
+% into your equation if necessary. Place an appropriate
+% unit of measure between the curly braces, e.g.
+% \hspace{1in}; you may have to experiment to achieve
+% the correct amount of space.
+
+%% ------------------------------------------------------------------------ %%
+%
+%  EQUATION NUMBERING: COUNTER
+%
+%% ------------------------------------------------------------------------ %%
+
+% You may change equation numbering by resetting
+% the equation counter or by explicitly numbering
+% an equation.
+
+% To explicitly number an equation, type \eqnum{}
+% (with the desired number between the brackets)
+% after the \begin{equation} or \begin{eqnarray}
+% command.  The \eqnum{} command will affect only
+% the equation it appears with; LaTeX will number
+% any equations appearing later in the manuscript
+% according to the equation counter.
+%
+
+% If you have a multiline equation that needs only
+% one equation number, use a \nonumber command in
+% front of the double backslashes (\\) as shown in
+% the multiline equation above.
+
+% If you are using line numbers, remember to surround
+% equations with \begin{linenomath*}...\end{linenomath*}
+
+%  To add line numbers to lines in equations:
+%  \begin{linenomath*}
+%  \begin{equation}
+%  \end{equation}
+%  \end{linenomath*}

### 2022-02-24 manuscripts/owp_fim4_2022/sections/introduction.tex
Typo fix

--- 2022-01-23 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
+Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
+In 2017 and 2019, river and flash flooding combined represented the leading cause of death and the second leading cause in 2018 among all natural disasters in the US \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
+More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
+With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
+Trends related to flood damages and fatalities have been steadily increasing over recent decades.
+\cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
+Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
+Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Software Dependencies and Architecture}
+\subsection{Operational Forecasting}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
-Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
-A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub (https://github.com/NOAA-OWP/cahaba).
-
-The pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
-Preprocessing acquires and prepares datasets for production of the FIM hydrofabric. 
-The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
-Functionality is included to turn FIM hydrofabric and streamflows into inundation maps represented in both binary and depths.
-Lastly, a test suite includes means of calculating evaluation metrics compared to a variety of preprocessed test case data.
-A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}
-\label{ssec:datasets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP).
-The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
-For enforcing levee data into the DEMs, the USACE National Levee Database (NLD) is used to burn feature elevations \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{OpenStreetMap,GreatLakesHydrographyDataset}.
-Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
-These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
-The full input datasets presented by source are listed in Table \ref{tab:data}.
-%
-\begin{table}
-\caption{Data sources, names, and descriptions.}
-\label{tab:data}
-\centering
-\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
-%\begin{tabular}{l c c}
-\hline
-Source & Name & Description \\
-\hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
-\hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
-\hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters \\
-\hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. \\
-\hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. \\
-\hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. \\
-\hline
-USACE & NLD & Levee database of locations and elevations  \\
-\hline
-OSM & Land-Sea Border & Border of land and sea. \\
-\hline
-GLHD & Land-Great Lakes Border & Border of land and Great Lakes. \\
-\hline
-InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
-\hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
-\hline
-\end{tabular}
-\end{table}
-%
-Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and GMS networks and 1,604 HUC8s for the MS method. 
-These methods will be explained more later.
-An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hydro-conditioning}
-\label{ssec:hydro_conditioning}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
-These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
-HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
-In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligble for flooding.
-Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
-It can also be used to simulate more accurate bathymetry which is not accounted for in the 10 m DEM \cite{gesch2002national}.
-
-Specifically within the context of OWP FIM, the hydro-conditioning operations that take place in sequential order are presented. 
-Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing units of size HUC8. 
-The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reach code's membership in the processing unit.
-Hydro-conditioning raster operations take place on buffered boundary definitions to avoid edge contamination and effects \cite{lindsay2013measuring}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Network Enforcement} 
-\label{ssec:stream_network_enforcment}
-%
-The location of the stream network is enforced to ensure general agreement with established stream networks.
-The NHDPlusHR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
-However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-For every NWM headwater segment a headwater point is derived and linearly interpolated to the nearest Burnline Event segment.
-Burnline Event headwater segments are split at the adjusted headwater point to match NWM flowlines.
-The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
+Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
+One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
+AHPS provides forecasting services in the form of ensemble stream flows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
+Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
+AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
+On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
+Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons to address the increasingly complex water challenges facing the US.
 %
@@
 \centering
-\includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection and linear interpolation.
-Resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR Burnline Events.}
-\label{fig:stream_density_pruning}
-\end{figure}
-%
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.
-
-The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
-The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
-Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
-Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
-Some of these drawbacks are later addressed by additional conditioning techniques later on.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Levee Enforcement}
-%
-The DEM at 10 m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
-In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10 m DEM.
-The elevations found in the NLD are burned into the DEM if those elevations were found to exceed those already in place.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Depression Filling}
-\label{sssec:depression_filling}
-%
-Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
-The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
-Priority-Flood developed by \citeA{barnes2014priority} is an algorithm for filling said depressions and shown to have improved performance over early works in the field by \citeA{jenson1988extracting} implemented in \citeA{tarboton2005terrain} as well as \citeA{planchon2002fast}.
-The depression filling algorithm used in our pipeline is a Priority-Flood variant developed by \cite{zhou2016efficient} with enhanced single-thread performance and a time complexity of O(n log n) for floating point grids.
-This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
-The depression technique employed here does leave the existence of flat regions where pits existed a prior thus later requiring the need for resolving these flats.
-The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Thalweg Elevation Conditioning}
-\label{sssec:stream_thalweg_elevation_conditioning}
-%
-Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
-In order to prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
-Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50 m.
-The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
-
-The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
-\citeA{garousi2019terrain} proposed an algorithm that traverses stream thalweg pixels in a depth first manner starting with adding all the headwater pixels to a queue. 
-The connectivity of the thalweg pixels is defined by the D-8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
-At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in Equation \ref{eq:thalweg_breach},
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:thalweg_breach}
-\textbf{D}[x] = \min_{y\ drains\ to\ x} {(\ \textbf{D}[x]\ ,\ \textbf{D}[y]\ )}
-\end{equation}
-\end{linenomath*}
-%
-, in which \textbf{D} represents the array of thalweg adjusted elevations indexed by x and y where by y is upstream of x. 
-When a pixel's upstream neighbors are all evaluated, the downstream pixel is added into the queue thus the depth first traversal of the drainage network.
-This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in Critical Success Index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Deriving FIM Hydrofabric}
-\label{ssec:deriving_fim_hydrofabric}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Flow Directions and Flats Resolution}
-\label{ssec:flow_direction_and_flat_resolution}
-%
-To facilitate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D-8 flow directions.
-D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
-The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southeast direction as 8. 
-Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
-
-Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
-Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
-The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
-Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
-By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Deriving FIM Stream Network}
-\label{sssec:deriving_fim_stream_network}
-%
-The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
-The FIM stream network is of similar drainage density as the NWM V2.1 network but fully converges at all junctions leaving no divergences in the network.
-This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
-Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
-Each stream link in this derived FIM stream network is split into equidistant reaches.
-Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments.
-Stream links are then further segmented at NWM lakes and HUC8 boundaries.
-Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretation of the stream network.
-Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherent the synthetic rating curve properties of the upstream or downstream segment, depending on the topology.
-Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
-Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
-This stream network is important since it drives the HAND calculation and derivation of catchments.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Catchments}
-\label{sssec:catchments}
-%
-Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
-Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
-The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
-Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
-The other catchments are derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Height Above Nearest Drainage}
-%
-Once the pixel level catchments are derived, the final relative elevations can be computed.
-Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment described in Section \ref{sssec:catchments}.
-The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
-Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Synthetic Rating Curves}
-\label{sssec:synthetic_rating_curve}
-%
- A method for converting forecast river discharges from the NWM to stages or river depths is necessary for producing FIMs with HAND. 
-For one-dimensional models such as the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a synthetic rating curve at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
-For this application, we utilized the reach averaged approach for developing synthetic rating curves (SRC) \cite{zheng2018river}.
-The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-The reach averaged Manning's formula is derived to be 
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:reach_averaged_mannings_equation}
-\textbf{Q} = \frac{1}{n} \frac{V^{5/3}S^{1/2}}{L B^{2/3}} 
-\end{equation}
-\end{linenomath*}
-%
-where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
-Q, V, and B are taken a specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
-All units are international given the 1 numerator above n.
-The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
-The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the channel length.
-
-Setting of the Manning's n roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
-These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cross-walking with NWM Stream Network}
-\label{sssec:cross_walking_networks}
-%
-The stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with a NWM reach identifier so that a discharge can be converted to stage and later inundation extent.
-For the methods already discussed, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent methods discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
-Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
-While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
-More discussion on this will follow in Section \ref{sec:discussion}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stream Order Reduction}
-\label{ssec:stream_order_reduction}
-%
-FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
-Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
-An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the MS of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
-This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the MS that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the MS catchments.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{False negatives associated with confluence of tributaries with MS. Integers represent flow values from BLE 100 year event for the associated areas. 
-No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
-\label{fig:catchment_boundaries_issue}
-\end{figure}
-%
-We seek to resolve this problem by deriving HAND for processing units with stream networks of reduced stream order. 
-We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
-The resulting FIMs from the overlapping HAND processing units are mosaiced together taking any inundated area to be inundated but more will be explained in Section \ref{ssec:inundation_mapping}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Main-stems}
-\label{sssec:nws_mainstems}
-%
-The Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
-The MS network comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
-It also spans 121,724 reaches across 1,608 HUC8s.
-HAND was originally derived for this stream network to enhance mapping skill along these critical MS segments \cite{djokic2019arc}. 
-The inundation derived from this stream network is mosaiced with the inundation from the FR network to form the MS FIMs. 
-Within each HUC, you'll typically only find a MS stream network of stream order 1 (i.e.
-headwater) but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Generalized Mainstems}
-\label{sssec:generalized_mainstems}
-%
-To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call GMS.
-Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
-Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-Starting at the outlet, a unique level path is propagated upstream. 
-At every confluence, the direction of maximum flow path length is sought to propagate the current level path identifier.
-For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
-Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
-At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
-
-To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
-In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
-For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each level path is buffered to a polygon with a user-available distance parameter of 7 km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
-We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
-Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}c.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates how level paths for the NWM are derived.
-Level paths symbolized by lines of unique colors are propagated upstream following the direction of maximum arbolate sum at each junction.
-}
-\label{fig:level_path_methods}
-\end{figure}
-%
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/gms_methods.jpg}
-\caption{Overall procedure for GMS HAND.
-In (a), we illustrate all NWM stream lines symbolized by their level path.
-Meanwhile (b), demonstrates the DEM clipped to a 7 km buffer around two selected level paths.
-In (c), we show how HAND can be computed just for each one of these two level paths independently. 
-We also show inundation maps created for these two level paths in (c). 
-In (d), we show all the inundation maps for all the level paths mosaiced together. }
-\label{fig:gms_methods}
+\includegraphics[scale=1.0]{figures/forecast_points.jpg}
+\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
+Also shown are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
+\label{fig:forecast_points}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Inundation Mapping}
-\label{ssec:inundation_mapping}
+\subsection{National Water Model}
+\label{ssec:national_water_model}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-The FIM hydrofabric consisting of the relative elevations grid, catchments grid, catchment polygons, rating curve, and cross-walking data are all used to convert forecasts from the NWM into forecasts extents.
-For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
-From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 8 spatial processing units in a rapid, parallel operation.
-The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
+Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
+In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
+The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
+As a complement to the operational NWM, OWP has also developed a configuration of the NWM that extends RFC forecasts downstream by assimilating and routing forecast streamflows to the next downstream AHPS forecast point.
+This configuration of the NWM is used to enhance forecasting skill by leveraging best available regional-scale data.
+The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
+Figure \ref{ssec:national_water_model} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
+The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
 
-Utilizing the stage-discharge relationships in the synthetic rating curves, each forecast for each catchment identifier is assigned a stage value. 
-The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
-We use the basic logic already established in previous works to conduct this \cite{nobre2016hand,liu2016cybergis,maidment2017conceptual}.
-Mathematically, the HAND values, $H_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
-For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
-For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim_depth}
-    D_{ij} = S_i - H_{ij}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim}
-    F_{ij} = D_{ij} > 0
-\end{equation}
-\end{linenomath*}
-%
-For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast in the form of a single raster file.
-For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
-Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
-Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim_depths}
-    D_{ij} = \max_{k=[1,...,K]} D_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-For the MS and GMS methods, the contributing areas are defined differently.
-For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to 2 for that case.
-For GMS, all FIMs from all the level paths in a given area are mosaiced together then K is set to this number of level paths.
-Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim}
-    F_{ij} = \max_{k=[1,...,K]} F_{ijk}
-\end{equation}
-\end{linenomath*}
+The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
+The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
+The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
+Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
+As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
+The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}
-\label{ssec:evaluation}
+\subsection{Height Above Nearest Drainage}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
-49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM so we used those corresponding discharges and mapping extents for evaluation.
-We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
-Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
-We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
-To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
-This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
-While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
+HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
+\citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
+A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
+The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
+\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
+This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
 %
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
-\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states. This dataset for 1\% and 0.2\% recurrence flows were used as benchmarks.}
-\label{fig:all_ble_maps}
-\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{HAND Implementations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
-\caption{Illustrates Base Level Engineering (BLE) cross sections and stream lines.
-The BLE stream network, which is denser than the NWM V2.1 stream lines, is also shown.
-BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM reach.
-}
-\label{fig:ble_evaluation_method}
-\end{figure}
+Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
+As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
+NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
+By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
+Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
+The original NFIE dataset was employed by the NWC as an unofficial demonstrations to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
+Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
+More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
+
+Related to these efforts, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
+Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
+In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30 m resolutions \cite{yamazaki2019merit,donchyts2016global}.
 %
-The metrics employed in this study to evaluate inundation extents include Critical Success Index (CSI), Probability of Detection (POD), and False Alarm Ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
-To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.
-The two types of errors consist of false positives (FP), or type I errors, which is dry in benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in benchmark by predicted dry.
-Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
-Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
-While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
-Thus, frequency dependent statistics should be used with caution when comparing across sites with varying frequencies. 
-Lastly, approximately 6 HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{OWP FIM}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-\begin{linenomath*}
-\begin{equation}
-\label{eq:csi}
-CSI = \frac{TP}{TP + FN + FP}
-\end{equation}
-\end{linenomath*}
+Many of those in the academic community assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
+\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
+The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
+Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
+Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3 m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
+The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
+Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
+Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
+Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
+These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.
+
+With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
+Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
+This framework is open-source and modular enabling and accelerating the research to operations development pipeline.
+Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
+In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
+Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
+Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
+This capability will be referred to as Generalized Mainstems (GMS).
+The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
 %
-\begin{linenomath*}
-\begin{equation}
-\label{eq:pod}
-POD = \frac{TP}{TP + FN}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:far}
-FAR = \frac{FP}{TP + FP}
-\end{equation}
-\end{linenomath*}
-%

### 2022-07-07 manuscripts/owp_fim4_2022/reviews/first_review/rebuttal.tex
* adding initial rebuttal.tex file with some changes * adding .gitignore with rebuttal.pdf on it

% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from 
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
% 
% Licensed under cc by-sa 3.0 with attribution required.
% See https://creativecommons.org/licenses/by-sa/3.0/
% and https://stackoverflow.blog/2009/06/25/attribution-required/

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}

% import Eq and Section references from the main manuscript where needed
% \usepackage{xr}
% \externaldocument{manuscript}

% package needed for optional arguments
\usepackage{xifthen}
% define counters for reviewers and their points
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% This refines the format of how the reviewer/point reference will appear.
\renewcommand{\thepoint}{P\,\thereviewer.\arabic{point}}

% command declarations for reviewer points and our responses
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
                  \section*{Reviewer \thereviewer}}

\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Reviewer~Point~\thepoint} } ---\ }
   {\par }

\newcommand{\shortpoint}[1]{\refstepcounter{point}  \bigskip \noindent 
	{\textbf{Reviewer~Point~\thepoint} } ---~#1\par }

\newenvironment{reply}
   {\medskip \noindent \begin{sf}\textbf{Reply}:\  }
   {\medskip \end{sf}}

\newcommand{\shortreply}[2][]{\medskip \noindent \begin{sf}\textbf{Reply}:\  #2
	\ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
	\medskip \end{sf}}

\begin{document}

\section*{Response to the reviewers}

\subsection*{Editor Response}

Associate Editor Evaluations: \\
Recommendation (Required): Return to author for major revisions \\
Accurate Key Points: Yes \\
\noindent
Associate Editor (Remarks to Author):

Dear authors,

We have received three very detailed reviews for your manuscript, which suggest that major revisions are needed before your manuscript can be reconsidered for publication in WRR.
All reviewers agree that the manuscript's structure has to be improved and that the method description needs clarification.
Reviewers 1 and 2 highlight the need to better work out the novel contribution of this study, which is hidden behind the report-like structure of this manuscript.
I agree with the reviewers that the manuscript needs a clearer problem statement, more focus on the major contribution, results and discussion.
Some clarification regarding methodology seems to be needed in Section 2.5.
In addition, Reviewer 3 points out that substantial editing is needed before re-submission.
I am looking forward to reading a substantially revised and edited version of this manuscript.\\

Best regards

\begin{reply}
We thank the reviewers for carefully reading the manuscript and for providing valuable feedback.
We have gone through the reviews and made every effort to address the concerns while staying within the scope of the paper.
We feel like we have addressed the concerns around not highlighting the primary contribution of the paper while strengthening the explaination of the contribution's methods in Section 2.5.
\end{reply}

% Let's start point-by-point with Reviewer 1
\reviewersection

\subsection*{Introductory Comment}

% Point one description 
\begin{point}
    This work improved the accuracy of the height above nearest drainage (HAND) based flood inundation mapping.
    This study first compared the original HAND-based flood inundation mapping approach with HEC-RAS 1D results and showed that HAND suffers from a limitation caused by independent neighboring catchments that cannot cross catchment boundaries.
    Then, the study proposed a series of terrain analysis steps to resolve this issue.
    The most crucial new method compared to other steps that had already been used in previous studies was to reduce the Horton-Strahler stream order of a HAND processing unit down to one.
    Overall, I support this work and find it a valuable research area.
    The paper was well written, and the work is original in its effort to resolve some issues associated with the HAND-based flood mapping method.
    However, I sometimes found it more like a technical report than a research paper.
    I appreciate the effort authors took to explain every step of the methodology in detail; however, I hoped to see more results and discussion than learning every detail mentioned in the methodology section.
    Also, the main part of the methodology section (section 2.5) requires modification as it does not clearly present the proposed method (please see my major comments below).
    I have some major, minor, and editorial points that I hope authors find helpful in revising this manuscript.
    \label{pt:foo}
\end{point}

% Our reply
\begin{reply}
Thank you for providing an overall assessment of the paper.
We have carefully reviewed your comments and made every effort to clarify or correct any issues you've brought up.
Any references to line numbers, paragraphs, sections, figures, or tables in the proceeding responses to your comments will refer to those in the original PDF document unless otherwise noted.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Major Comments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
My first comment is regarding comparing results with the HEC-RAS model results.
In this comparison, you evaluated the flood inundation extent and not depth.
My question is, how the proposed method affects the estimated flood stage (depth)?
Figure 9a shows that the proposed method resulted in a smaller estimated stage value.
This should then result in getting a smaller HAND-based inundation extent.
But you showed a larger inundated extent in Figure 10.
How do you separate these two effects?
This needs to be at least discussed in your discussion section.
    \label{pt:bar}
\end{point}

\begin{reply}
We thank you very much for making this important point as one of the main purposes of the paper is to introduce how exactly drainage reduction techniques change HAND derived inundation performance.
After reading your comment and our initial discussion section submission, we did notice a missing portion that properly explains how drainage order reduction increases catchment boundary extents and how that drives increased inundation extents and Probability of Detection (POD).
We noted that a fair portion of the discussion (second paragraph in Discussion) was dedicated to how reducing drainage order promotes stage reduction and an associated marginal reduction in False Alarm Rate (FAR).
This, as you correctly noted, is a competing effect to the increased catchment size and associated increase in inundation area and POD.

These competing effects are very difficult to detangle since catchment definitions are used to derive synthetic rating curves (SRCs).
When viewed from the perspective of change in the total inundation area, these two effects can correctly be described as competing since one tends to increase inundation area while the other generally decreases it.
However, when viewed in the paradigm of overall FIM performance, we believe they are complimentary in nature as hinted by the concurrent increase in POD and slight decrease in FAR.

Motivated by this comment, we have inserted a paragraph that better describes the increase POD as driven by catchment size enlargement especially around the critical areas described in the paper.
Additionally due to your comment on the competing effects, we have inserted another paragraph that gives the audience a better conceptual idea as to how these effects both compete in inundation area but complement each other in terms of performance.

Limited catchment boundaries drive pools of water that are artificially built up without room for the water to naturall spread. 
This can be conceptually visualized as a column of water that extends beyond its containers top edge.
This inherent limitation in HAND skews the stage-discharge relationship in addition to limiting the inunduation extents.
Thanks to your comment we have contributed a significant addition that better highlights some of the points made in this reply.

Thank you for this important question as one of the main discussion points proposed in the paper was introducing how this demonstrated drainage order reduction technique for HAND computation affects FIM skill.
In this study, we noted that FIM skill was enhanced with the increased use of drainage order reduction via two interconnected avenues: catchment boundary modifications and river stage reduction.
For inundation mapping purporses, HAND works on the principle of independent catchments that translate reach-level stream flow values to stages with the use of reach-level synthetic rating curves.
This relationship between stage and discharge was derived using geometric values sampled from the HAND and catchment rasters that are inputted into the reach-averaged Manning' N equation (Equation 2).

The critical argument here is conveying that a given's reach stage value can only be propogated as far as the given's reach catchment boundary. 
The term "nearest" in HAND is a limiting factor as higher order stream networks have locations with multiple contributing reaches that are not being considered.
When delineating HAND with unit-order streams, catchments are extended to their more natural extents thus making them more expansive.
These catchments can overlap considerably allowing for multi-source flooding effects.

The phenomena of larger catchments is illustrated in Figure 10 where catchments delineated at full stream resolution in Figure 10a are shown in white.
When compared to the catchments also symbolized in white in Figure 10b for the unit-stream order network, one observes a sizeable increase in the catchment size for the horizontally oriented mainstem illustrated in the image.
By expanding the catchment sizes by unconstraing catchment lines to only their nearest drainage line, catchments begin to include more expansive drainage areas that would potentially include those of upstream tributaries.
This size increase enables the inundation depth to increase considerably.

As an opposing effect, larger catchments influence the geometric properties that are used in the reach-averaged Manning's equation including bed area and volume.
Figure 9 attempts to illustrate these changes where we show how the more dependency on lower resolution stream networks the lower the rating curve biases.
These changes are influenced by the shape of catchments by making more volume and bed area for water to settle into.
While volume and bed area denote competing forces on discharge, Figure 9d illustrates how the net affect increases discharge with respect to a given stage value.

So in summary, yes stage does decrease with increased reliance of unit drainage order networks which is due to the larger, overlapping catchments that are created.
This reduction in stage, itself, likely would create a reduction in total inundation area if it weren't for the fact that these new larger catchments 
\end{reply}

\begin{point}
My second question is that it seems the proposed method reduces the false alarm rate.
This means the underestimation of the original HAND-based flood mapping compared to the FEMA maps is resolved.
However, we know there are uncertainties associated with the FEMA flood extent.
How would you say this reduction of false negatives is really in a right direction when you did not compare results with ground truth data?
    \label{pt:bar}
\end{point}

\begin{reply}
And our reply to it.
\end{reply}

\begin{point}
My second major comment is that the methodology section tends to be very long, particularly sections 2.1-2.4.
The main contribution of this paper is section 2.5, but it took me as a reader a long time to get to this point and figure out what the paper presents.
Also, section 2.5 is somehow confusing and could be revised and simplified.
For example, in L531, you mentioned you present two successive methods.
Is the first one sub-setting MS?
The NWM main-stem part (2.5.1) seems (at least to me) like an introduction, not something that you did.
Please clarify and revise this part such that a reader can follow the story that connects 2.5.1 and 2.5.2 with the main paragraph in Section 2.5.
Again, this is the most important part of you work. 
\label{pt:bar}
\end{point}

\begin{reply}
And our reply to it.
\end{reply}

\begin{point}
The limitation that the authors attempted to resolve is introduced as a known limitation in the abstract.
Is that really a known limitation?
If yes, did you cite it properly in your literature review?
I believe this was not introduced well as a limitation.
You only pointed out to this in your results (Figure 10).
This is somehow explained in L203-210.
But not very well explained.
This is an important part and the gap your research is trying to resolve so be generous and explain more.
Could you also clarify whether this limitation always results in more overestimation or underestimation? 
\label{pt:bar}
\end{point}

\begin{reply}
And our reply to it.
\end{reply}

\begin{point}
Links to all datasets introduced in section 2.2 should be included in your manuscript either in the suggested table or within the "Open Research" section at the end of your manuscript. 
\label{pt:bar}
\end{point}

\begin{reply}
And our reply to it.
\end{reply}

\begin{point}
Line 331-332: what were the values for buffer distance, smooth drop, and sharp drop you used?
Were they fixed numbers or differ?
How did you choose these values? 
\label{pt:bar}
\end{point}

\begin{reply}
And our reply to it.
\end{reply}

\begin{point}
I believe a diagram showing all the steps and methods you took to prepare your data and method could be helpful.
Your material and data section has about 14 sub-sections collectively.
It is sometimes hard to follow along. 
\label{pt:bar}
\end{point}

\begin{reply}
And our reply to it.
\end{reply}

\begin{point}
LINE 462-464. Could you explain more about why this is required?
Is it to improve flat areas?
Can't we see any results associated with this error improvement?
Any other studies that observed this that you could cite?
\label{pt:bar}
\end{point}

\begin{reply}
And our reply to it.
\end{reply}

\begin{point}
Fig 4. You need to explain this figure in the content more thoroughly.
What are the numbers?
How does one dominate the other?
You mention that the level path method starts from an outlet, so it is probably better to show an outlet in Figure 4.
Please clarify what you mean by "arbolate sum"? I cannot understand what this summation is.
\label{pt:bar}
\end{point}

\begin{reply}
And our reply to it.
\end{reply}

\begin{point}
In equations 5 and 6, When you apply the mosaic method, what is the reason to choose the max?
Is that because you always see the original HAND approach, as shown in Figure 3, underestimating the flood extent?
This may not be true everywhere, or it might be.
Please provide your thoughts on this and add it to the manuscript. 
\label{pt:bar}
\end{point}

\begin{reply}
And our reply to it.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Minor Comments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the short-hand macros for one-liners.
\begin{point}
In the third key point, what do you mean by higher skill inundation?
Please be more specific.
Do you mean more accurate inundation depth or inundation extent?
Same for L39.
Mapping skill itself is not clear.
Does that refer to flood depth or flood extent?
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
The last line of your plain language summary: Philosophically speaking, if you compared the HAND to a more realistic model and found improvements, what is the point you offer using HAND instead of a more realistic model?
If HAND is not a realistic model, the logic is that we should not use an unrealistic model.
Please revise this.
I agree HAND is a simplified model, but that does not make it unrealistic.
It is realistic because it is based on some terrain physics.
You may want to revise this sentence and use "physically-based" rather than "realistic".
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
L121. There is a gap between the geofabric concept you introduced and the Muskingam-Cunge routing method.
You might want to start this sentence like this: The NWM provides stream forecasts at these geofabric segments using the Muskingham-Cunge method. 
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Page 29: It could be better to change the title of subsection 3.1 to "Flood mapping performance". 
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
L671. Please add "(Table 2)" at the end of the sentence.
This guides the reader to know what you are referring to immediately. 
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Line 315-318: Please clarify what NHD is?
Does that refer to the NHDPlusHR or NHDPlus medium resolution as used in the NWM v2.1?
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Line 344-345: This statement requires citation(s).
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
I would suggest removing lines 534-535. 
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
I believe section 2.3.1 is not a DEM hydro conditioning step.
It is more like creating an input (as seed points) that you need when delineating the stream network.
The hydro conditioning of the DEM section should start with levee enforcement (section 2.3.2).
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Line 570. What is the logic behind using a 7 km buffer?
Did you first come with this number or after some test?
What is your recommendation or suggestion?
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Maybe reporting the percentage of improvements in Table 2 would help better understanding the contribution of your work.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Figure 8 contains a lot of information, and readers may require a substantial explanation to understand what these graphs and numbers all shown in a figure mean.
You introduced the figure in the first paragraph of the result section (L672-677).
I suggest you take one example from this figure and walk the reader through the numbers.
For example, what does it mean when the KDE of CSI for the GMS model has the most left-skewed graph compared to the other models?
Is it considered good or bad?
Is it something you had expected?
Another question that arises here is why using a higher Manning's N shifts the KDE graphs up and whether this means results get better or worse.
You discuss this figure on Line 686-690, but I think it could be better to bring this part further up where you first mention Figure 8.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{FIGURES}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
Fig 1. The NWM FR streams line is hardly visible in the legend. Please revise. 
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Fig 3. Please define MS in the figure caption (Is it mainstem?).
Also, TN, FN, FP, and TP should be described here.
You introduced these terms later in the evaluation section.
A reader has no idea what these mean unless they first read your evaluation section (section 2.7).
Please add flow direction to this map using arrows.
It can help a reader quickly understand the flow numbers on the figure as you talk about in the content.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Line 575. This should be Figure 5d not 5c.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Fig 5. How many different color codes (level paths) are there?
Is fig 5a one HUC 8?
Please mention the HUC8 identifier.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Fig 6. It seems that the blue areas are inundated areas for the 0.2\% recurrence flow.
How about the inundated areas for the 0.1\% recurrence flow?
I am asking this because the caption mentions 0.2\% and 0.1\%.
Please clarify and revise the figure and caption accordingly. 
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Fig 7. It could be good to add some labels on the figure indicating where this location is within your study area.
What is the mainstem's name?
Would it be possible to add NWM v2.1 catchments to the map?
I understand this might make the figure busy, but I would like to know how many cross sections falls within a NWM catchment?
Is this the FR or MR NWM v2.1 stream or the one you created (GSM)?
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Fig 10. Which recurrence event was used to create this inundation map?
Please add this to the caption and mention it in the content where you explain results in Figure 10.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{EDITORIAL COMMENTS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
L66. Remove the period before the parenthesis, followed by the citations.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
L101. NWM is already defined. Just use the abbreviation here. 
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Title of section 1.5: It could be better to use the Operational Water Prediction (OWP) Flood Inundation Maps (FIM)
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
L160. HPC is already defined. 
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
L167. MR is already defined.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
L178. USGS is not yet defined.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Line 347: NLD has already been introduced on page 11.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Line 524: MS should be first defined here not on line 537.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

\begin{point}
Throughout the manuscript, please be consistent with the terms you use.
Use either Manning's n or Manning's N everywhere.
\label{pt:bar}
\end{point}

\begin{reply}

\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Begin a new reviewer section
\reviewersection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
	This is the first point of Reviewer \thereviewer. With some more words foo
	bar foo bar ...
\end{point}

\begin{reply}
	Our reply to it with reference to one of our points above using the \LaTeX's 
	label/ref system (see also \ref{pt:foo}).
\end{reply}

\end{document}

### 2022-08-05 manuscripts/owp_fim4_2022/reviews/first_review/rebuttal.tex
* almost done with all of review #1s comments * just need to make method overview figure

@@
 \usepackage{fullpage}
+\usepackage{apacite}
+
+% bibstyle
+\bibliographystyle{apacite}
 
@@
 I have some major, minor, and editorial points that I hope authors find helpful in revising this manuscript.
-    \label{pt:foo}
+\label{pt:intro_1}
 \end{point}
@@
 This needs to be at least discussed in your discussion section.
-    \label{pt:bar}
+\label{pt:discussion}
 \end{point}
@@
 \begin{reply}
-We thank you very much for making this important point as one of the main purposes of the paper is to introduce how exactly drainage reduction techniques change HAND derived inundation performance.
+We thank you very much for making this important point as one of the main purposes of the paper was to introduce how exactly drainage reduction techniques change HAND derived inundation performance.
 After reading your comment and our initial discussion section submission, we did notice a missing portion that properly explains how drainage order reduction increases catchment boundary extents and how that drives increased inundation extents and Probability of Detection (POD).
-We noted that a fair portion of the discussion (second paragraph in Discussion) was dedicated to how reducing drainage order promotes stage reduction and an associated marginal reduction in False Alarm Rate (FAR).
 This, as you correctly noted, is a competing effect to the increased catchment size and associated increase in inundation area and POD.
@@
 
-Motivated by this comment, we have inserted a paragraph that better describes the increase POD as driven by catchment size enlargement especially around the critical areas described in the paper.
-Additionally due to your comment on the competing effects, we have inserted another paragraph that gives the audience a better conceptual idea as to how these effects both compete in inundation area but complement each other in terms of performance.
-
-Limited catchment boundaries drive pools of water that are artificially built up without room for the water to naturall spread. 
+Motivated by this comment, we have completely restructured the Discussion section to make sure your comment is properly highlighted in the paper.
+We have inserted some paragraphs that better describe the increase in inundation extents and POD as driven by catchment size enlargement especially around the critical areas described in the paper.
+These larger, overlapping catchments better handle riverine inundation from multiple sources as opposed to the traditional single source "nearest drainage" that HAND is limited to.
+Limited catchment boundaries drive pools of water that are artificially built up without room for the water to naturally spread. 
 This can be conceptually visualized as a column of water that extends beyond its containers top edge.
 This inherent limitation in HAND skews the stage-discharge relationship in addition to limiting the inunduation extents.
-Thanks to your comment we have contributed a significant addition that better highlights some of the points made in this reply.
 
-Thank you for this important question as one of the main discussion points proposed in the paper was introducing how this demonstrated drainage order reduction technique for HAND computation affects FIM skill.
-In this study, we noted that FIM skill was enhanced with the increased use of drainage order reduction via two interconnected avenues: catchment boundary modifications and river stage reduction.
-For inundation mapping purporses, HAND works on the principle of independent catchments that translate reach-level stream flow values to stages with the use of reach-level synthetic rating curves.
-This relationship between stage and discharge was derived using geometric values sampled from the HAND and catchment rasters that are inputted into the reach-averaged Manning' N equation (Equation 2).
-
-The critical argument here is conveying that a given's reach stage value can only be propogated as far as the given's reach catchment boundary. 
-The term "nearest" in HAND is a limiting factor as higher order stream networks have locations with multiple contributing reaches that are not being considered.
-When delineating HAND with unit-order streams, catchments are extended to their more natural extents thus making them more expansive.
-These catchments can overlap considerably allowing for multi-source flooding effects.
-
-The phenomena of larger catchments is illustrated in Figure 10 where catchments delineated at full stream resolution in Figure 10a are shown in white.
-When compared to the catchments also symbolized in white in Figure 10b for the unit-stream order network, one observes a sizeable increase in the catchment size for the horizontally oriented mainstem illustrated in the image.
-By expanding the catchment sizes by unconstraing catchment lines to only their nearest drainage line, catchments begin to include more expansive drainage areas that would potentially include those of upstream tributaries.
-This size increase enables the inundation depth to increase considerably.
-
-As an opposing effect, larger catchments influence the geometric properties that are used in the reach-averaged Manning's equation including bed area and volume.
-Figure 9 attempts to illustrate these changes where we show how the more dependency on lower resolution stream networks the lower the rating curve biases.
-These changes are influenced by the shape of catchments by making more volume and bed area for water to settle into.
-While volume and bed area denote competing forces on discharge, Figure 9d illustrates how the net affect increases discharge with respect to a given stage value.
-
-So in summary, yes stage does decrease with increased reliance of unit drainage order networks which is due to the larger, overlapping catchments that are created.
-This reduction in stage, itself, likely would create a reduction in total inundation area if it weren't for the fact that these new larger catchments 
+Overall, accounting for multiple sources of inundation by delineating catchments from independent level paths of unit stream order better accounts for multi-source inundation thus increasing inundation extent (and POD) while also lowering river stage values thus reducing FAR in other areas.
+We believe the updated Discussion section better reflects this critical finding of the paper.
 \end{reply}
@@
 \begin{reply}
-And our reply to it.
+We acknowledge your point that FEMA, or Base Level Engineering (BLE), flood extents are derived from HEC-RAS 1D models which are themselves subject to a set of assumptions and limitations based on the modeling techniques and data sources employed.
+It can be expressed confidently that the use of BLE maps for evaluation does not offer the sort of fidelity that a ground truthed, real-time survey would provide.
+However, the BLE does offer many advantages that are useful when evaluating a large-domain flood inundation map derived from HAND.
+
+Despite being an engineering scale model, the BLE is available over a very large domain (49 HUC8s and 9 states) at high spatial resolutions which allows for a broader evaluation of HAND which is being applied to continental scale flood forecasting for our purposes.
+Additionally, the BLE offers the availability of tightly spaced cross-sections that provide the stream flows used in making the flood extents.
+This high resolution dataset of stream flows and inundation extents provides the opportunity to remove the meterological, hydrological, and some of the hydraulic assumptions, errors, and limitations from the comparison.
+Stream gages and in-situ observations, while clearly higher fidelity in nature, don't offer the spatial coverage that the BLE offers both in terms of flood extents and their associated stream flows that can be used to detangle confounding factors that introduce errors into the evaluation.
+Nevertheless, FEMA maps have a storied use in the literature as an evaluation dataset due to some of the advantages and trade-offs listed above \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
+Additionally, it is important to note that almost any other source of FIM is subject to errors and limitations including crowd-sourced data, remote sensing observation, and modeled extents.
+Unfortunatley, flood inundation mapping is an imprecise science that depends on compromises to be made in the search for objective benchmarking.
+
+To address your concern within the paper, we have added some verbage in the evaluation section of the draft that better recognizes and discusses the limitations of the BLE HEC-RAS based FIMs.
+While we can't assure that the BLE maps are perfect, they are considered more sophisticated than our simplified continental technique and have been established well in the literature as a common benchmarking model.
+Hopefully this properly addresses your concern with the use of the BLE HEC-RAS 1D models for evaluation.
 \end{reply}
@@
 \begin{reply}
-And our reply to it.
+We respect your opinion that the methodology may seem to take away from the novel scientific contribution of adding multi-source riverine inundation to HAND.
+While the major and foremost scientific contribution of the paper is a proposed method to cure the nearest drainage limitation within HAND, we note that a significant amount of work was devised in developing the methods for computing HAND at any scale.
+These hydro-conditioning and geospatial techniques can have a significant impact on performance and on the quality of the final operational product.
+We found it important to make sure these methods were properly documented within scientific literature for the benefit of the academic and forecasting communities.
+Based on your recommendation, we did go through sections 2.1-4 and tried to remove any information that maybe superflous to the goals and scope of the paper.
+This includes much of the stream thalweg breaching technique in Section 2.3.4 that was more than adequately described in the citation it was developed in.
+
+Additionally, we initiated a substantial revamp of Section 2.5 to ensure the main contribution of the paper is properly explained and defended.
+The Section NWM MS (2.5.1) does represent methods conducted by us therefore we have elected to keep it in the paper to adequately describe a stream reduction technique that covers 4\% of the country that is of primary concern to river forecasters.
+Also, we added an extra Figure (Figure \#6 in the revised document) that expands upon the explaination in Figure 5. 
+This provides the reader with a more in-depth look at how drainage order reduction down to the level path scale expands catchment and inundation extents.
+This allows for multiple sources of fluvial inundation to be accounted for in regions where it is applicable.
+We believe that this better conveys the main contribution of the paper to the audience.
 \end{reply}
@@
 \begin{reply}
-And our reply to it.
+Thank you for pointing out an opportunity to better explain the limitation we are trying to resolve within HAND for FIM applications.
+To better explain this limitation, we included an additional Sub-section (1.4 HAND's Assumptions and Limitations) that better introduces all of the fundametal assumptions HAND employs and their resulting limitations.
+Much effort is expelled here to ensure the reader better understands HAND's limitation to only model fluvial inundation from one nearest drainage source thus ignoring other possible fluvial sources of inundation.
+We would also like to point out the reference to Figure 3 and the new Figure 6 which also illustrate this problem and proposed solution.
+Please reference the explanation in the Methodology Section 2.5.2 on more discussion on the issue and proposed solution.
+And lastly, yes, the citations for this known limitation were provided on L241-242 but also included in the new sub-section to ensure the reader is aware of prior literature acknowledging this issue.
+
+With respect to your question about always producing overestimation (False positives) or underestimation (False negatives), we note in the results and discussion sections that you can't say that this problem always leads to FPs or FNs.
+It can only lead to FN's if a given catchment boundary is restricting water inundating a neighboring catchment.
+This is commonly seen in areas where junctions are present where a reach's drainage area is restricted by the nearest drainage area of neighboring tributaries.
+However, as the wall of a catchment boundary begins to fill artificially with water, the lack of spread could cause FPs in other regions.
+By reducing the effect of the neigboring boundaries, we see a general reduction in the rating curve thus leading to lower stage values at given discharges.
+We believe that the new version of the Discussion section adequetely details these competing forces as per your prior recommention (\ref{pt:discussion}).
 \end{reply}
@@
 \begin{reply}
-And our reply to it.
+Based on your suggestion we have reviewed the citations to all datasets and software used to ensure they follow the journal's requirements.
+In addition, to citing the datasets and software properly in their respective sections, 2.1 and 2.2, we have added another column in the datasets table to ensure they are properly cited there as well.
+The URL's to all the software and datasets are included as part of the reference as indicated by the journal.
 \end{reply}
@@
 \begin{reply}
-And our reply to it.
+The values for buffer distance, smooth drop and sharp drop were set fixed to 70m, 1000m, and 10m, respectively.
+These values were selected due to prior research in the citations provided on original Lines 375-377.
+They were decided to be keep fixed to limit the scope of the article to its main objectives though we acknowledge that much work can be done to optimize these parameters in the future.
+Based on your comment, we have added the important information within this reply into the revised manuscript in the Stream Network Enforcement Section (2.3.1).
 \end{reply}
@@
 \begin{reply}
-And our reply to it.
+We use the native elevations from the NHD DEM in the areas not corresponding to the thalweg or the areas excavated with the AGREE DEM method.
+Doing so allows the HAND value within non-pit filled regions to be computed based on its original value and not its hydro-conditioned value.
+We inherited this technique from Djokic, 2019 which we now cite on that line in the paper.
 \end{reply}
@@
 \begin{reply}
-And our reply to it.
+The full explanation for this figure is provided by its in-text reference but based on your recommendation we have expanded this caption to properly describe the Figure in place.
 \end{reply}
@@
 \begin{reply}
-And our reply to it.
+The main scientific contribution of the paper was to allow for multiple sources of fluvial inundation to be considered. 
+At a given pixel with multiple fluvial inundation sources, we decided, intuitively, that the maximum contributing depth should be selected.
+We respect that flooding is governed by the shallow water equations and these methods are just a simplified proxy for this.
+We have updated the paper to include this and invite future researchers to investigate this.
 \end{reply}
@@
 \begin{reply}
-
+The term ``extents'' was added to these two references (key point and original L39) to clarify that we are using flood extents to evaluate for skill.
 \end{reply}
@@
 It is realistic because it is based on some terrain physics.
-You may want to revise this sentence and use "physically-based" rather than "realistic".
+You may want to revise this sentence and use ``physically-based'' rather than ``realistic''.
 \label{pt:bar}
@@
 \begin{reply}
-
+The term ``realistic'' was changed for ``physically-based'' per your suggestion.
 \end{reply}
@@
 \begin{reply}
-
+This suggestion was incorporated but the term ``hydrofabric'' was recycled in favor for the term ``geofabric''.
 \end{reply}
@@
 \begin{reply}
-
+Changed.
 \end{reply}
@@
 \begin{reply}
-
+Due to restructuring of the Results section, this reference was changed to refer to the Figure with violin plots instead in the following sentence.
 \end{reply}
@@
 \begin{reply}
-
+Specified to NHDPlusHR.
 \end{reply}
@@
 \begin{reply}
-
+Citations added.
 \end{reply}
@@
 \begin{reply}
-
+Removed.
 \end{reply}
@@
 \begin{reply}
-
+Section 2.3.1 is about enforcing the general location and density of the stream network used for HAND.
+Additionally, an excavation procedure is detailed that is used to counter flat bathymetric regions in the DEM that can severly distort synthetic rating curves (SRC) and catchment delineations.
+We use the term ``hydro-conditioning'', as stated in Section 2.3, as the series of steps required to generate monotonically decreasing elevations, flow directions that fully drain processing regions, and monotonically draining stream network thalweg lines.
+These are fundamental assumptions and inputs required for HAND, catchment, and SRC calculation.
+We believe that these steps should fall under the term hydro-conditioning as we defined it in the paper. 
 \end{reply}
@@
 \begin{reply}
-
+We exposed this parameter to the user as we acknowledge it could benefit from future calibration.
+We decided to set a generous value to avoid edge effects as well as possibly limiting catchment sizes in regions with lower slope where inundation could extend out some ways.
+This large buffer can add significant computational expense to smaller rivers as well as added data storage requirements.
+Overall, we recommend future researchers and developers to consider this parameter to better balance skill with computational expenses.
+We have added this discussion to the paper.
 \end{reply}
@@
 \begin{reply}
-
+Based on your request, we have changed the results for the MS and GMS models to represent the percentage change from the respective metric within the FR method based on the metric, Manning's n, and magnitude combination.
 \end{reply}
@@
 Is it something you had expected?
-Another question that arises here is why using a higher Manning's N shifts the KDE graphs up and whether this means results get better or worse.
+Another question that arises here is why using a higher Manning's n shifts the KDE graphs up and whether this means results get better or worse.
 You discuss this figure on Line 686-690, but I think it could be better to bring this part further up where you first mention Figure 8.
@@
 \begin{reply}
-
+Based on your suggestions, we have significantly improved to the Results section.
+The first paragraph briefly introduces the results while the second paragraph gives the reader and in-depth description of the previously numbered Figure 8.
+We have made sure all elements of the Figure are properly described and that a specific sub-figured is dived into more detail to ensure the reader has a proper understanding of the many elements at play here.
+In addition to this enhanced description, we have added a significant amount of information in the third paragraph of the results section that begins to dissect the results and the trends observed in the study.
+In a related point and in response to your critique of the Discussion section, we have made significant enhancements to the Discussion section that better elaborate on the material and trends presented in this Figure.
+We hope that these changes fully address your questions and support this Figure that is central to the paper's findings.
 \end{reply}
@@
 \begin{reply}
-
+Corrected.
 \end{reply}
@@
 It can help a reader quickly understand the flow numbers on the figure as you talk about in the content.
-\label{pt:bar}
 \end{point}
@@
 \begin{reply}
-
+The caption for Figure 3 has been significantly revised addressing your concerns.
+Flow direction arrows have been added to the figure as well.
+The term main stem was removed from this caption to avoid confusions.
+Mainstems (MS) refers to the subset of the NWM FR network that is at or downstream of existing forecast points.
+Here main stem was used to refer to the higher stream order segment in the middle of the image.
 \end{reply}
@@
 Please mention the HUC8 identifier.
-\label{pt:bar}
 \end{point}
@@
 \begin{reply}
-
+We have added the HUC8 identifier, 12090301, to the figure's caption.
+This particular HUC8 has 372 unique level path IDs which we have also indicated in the caption.
 \end{reply}
@@
 Please clarify and revise the figure and caption accordingly. 
-\label{pt:bar}
 \end{point}
@@
 \begin{reply}
-
+Reference to 0.1\% recurrence flow was removed in the caption.
+There isn't enough granularity to show both the 0.1\% and 0.2\% flows at this scale as they would both look the same.
+The main purpose of the figure is to illustrate the spatial extents and distributions of the BLE validation maps.
 \end{reply}
@@
 Is this the FR or MR NWM v2.1 stream or the one you created (GSM)?
-\label{pt:bar}
 \end{point}
@@
 \begin{reply}
-
+These names of the rivers are not present in the NWM V2.1 but instead I had to look at their equivalents in the NHDPlusHR stream line data.
+While the figure is quite busy, I agree that showing the NWM catchments does provide the reader some sense as to how many BLE cross sections might be associated with a given NWM forecast identifier.
+We have updated the caption to include this new layer and provide some more spatial context.
 \end{reply}
@@
 Please add this to the caption and mention it in the content where you explain results in Figure 10.
-\label{pt:bar}
 \end{point}
@@
 \begin{reply}
-
+We used the 500yr magnitude and added this to both the caption and in-text content as well.
 \end{reply}
@@
 \subsection*{EDITORIAL COMMENTS}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{reply}
-
+Done.
 \end{reply}
@@
 \begin{reply}
-
+Addressed.
 \end{reply}
@@
 \begin{reply}
-
+This is the first occurrence of Medium Resolution (MR) that we are seeing. 
+It is possible prior uses were removed during edits.
 \end{reply}
@@
 \begin{reply}
-
+We opted for the use of main segment here to avoid confusion with the Mainstems stream network used throughout the study.
 \end{reply}
@@
 \begin{reply}
-
+We have opted for the use of the lower case, Manning's n.
 \end{reply}
@@
 \begin{point}
-	This is the first point of Reviewer \thereviewer. With some more words foo
-	bar foo bar ...
+This study presents recent development made by federal agencies and their contractors on streamlining HAND-based flood inundation mapping.
+By introducing the idea of level paths, they have also tested the hypothesis that unary stream order networks enhance flood mapping performance skills with HAND.
+The study presents some solid work, and the results support their statements.
+It's important to expose this workflow, which has been adopted by government agencies in daily operation, to the academic community for broader discussion.
+I'm a little concerned about the relevance of this draft.
+To publish on Water Resources Research, you need first to highlight the problem you want to solve in water science.
+Therefore, presenting some research that is a byproduct of tool development won't interest potential readers of this journal.
+The authors may consider reorganizing the draft in a more science-driven way instead of technology.
+Based on the reasons above, a major revision is recommended for the editors' consideration.
 \end{point}
@@
 \begin{reply}
-	Our reply to it with reference to one of our points above using the \LaTeX's 
-	label/ref system (see also \ref{pt:foo}).
+We understand your concern and have made every effort to highlight the novel contributions of the manuscript which is an operational, continental scale version of HAND that applies the latest datasets and hydro-conditioning methods while considering multiple sources of fluvial inundation instead of just that of the nearest drainage line.
+We hope that the revisions completed based on your suggestions and the suggestions of the other two reviewers better highlight the studies' novel findings in a more science-driven format.
 \end{reply}
 
+\begin{point}
+Line 82-84 Introducing these two NWM networks in section 1.2 makes more sense.
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+\begin{point}
+Line 112 There isn't a figure 1.2.
+\end{point}
+
+\begin{reply}
+Fixed.
+\end{reply}
+
+\begin{point}
+Line 128-129 Here, a statement may be added to explain that running a continental-scale hydrodynamic model regularly is not feasible in operational forecasting, making the simplified alternative like HAND a better solution.
+(Like line 795-798)
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+\begin{point}
+Line 212-230 ``In addition to developed tooling, we introduce research'' I don't like how the authors present the study's purpose. This is not a draft submitted to Environmental Modelling \& Software.
+Presenting some research that is a byproduct of tool development won't interest the potential readers of this journal.
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+\begin{point}
+Although very comprehensive, the introduction section seems a little bit long.
+This is not a HAND-study review paper, so you don't have to go through all the studies.
+The authors may consider removing a few sentences to make it more focused.
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+\begin{point}
+Figure 2 How are the pruned headwaters derived? Why are they moved away from the original NHDHR endpoints?
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+\begin{point}
+Line 513 I don't think "work for many instances fine" is formal enough as a sentence in a scientific journal article.
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+\begin{point}
+Line 532-533 Four sections have been written in the background before the authors finally dive into the method related to the hypothesis of the current draft, which seems too much.
+I appreciate the comprehensiveness, but this is a journal article, not a technical document.
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+\begin{point}
+Line 544 "only find an MS stream network of stream order 1 (i.e., headwater)" From Figure 1, we can tell that most AHPS sites are located on big rivers, which means their stream orders are pretty low.
+Then why do you say they are headwater links?
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+\begin{point}
+Figure 4 This figure is confusing and needs some revision.
+Demonstrating one network is enough to explain the idea of level paths, but there are a few in this plot.
+Also, the annotations are too dense, and their alignment needs to be adjusted.
+For example, I'm confused when I see two numbers (6.13 and 10.95) on the left purple link.
+How come two numbers there?
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+\begin{point}
+Table 2 \& Figure 8 If only staring at these index numbers and plots, the improvement brought by the current study looks relatively marginal.
+The authors may look from other angles to see the benefits brought by analyses with unary stream order networks.
+Otherwise, the significance of this study does not look very convincing.
+\end{point}
+
+\begin{reply}
+\end{reply}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Begin a new reviewer section
+\reviewersection
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{point}
+This paper details the methods employed and findings encountered as the authors endeavored to produce a nationwide flood inundation map (FIM) from National Water Model (NWM) streamflow estimates over the Continental United States.
+This is an application of large-domain hydrological modeling.
+The relevance of this work to society is not to be understated; a forecast of flood inundation extent and depth from an operational flood forecast model is essential to reducing risks associated with major hydrologic events.
+The originality of this work lies in the scale of the problem being addressed and the relevance of operational forecasting to mitigating risk to life and property.
+
+This manuscript adds to the growing body of literature using HAND as a method for determining potential flood inundation in the absence of a full-physics (hydraulic) model.
+Ultimately, the process of producing inundation estimates for >2.7M stream segments at high spatial resolution (~10m) is one which required much iteration.
+In the end, the authors found that generalizing the stream networks, and using smaller basin areas (HUC8) combined with aggregating individual level-paths yielded better estimates of flood inundation when compared to results from HEC-RAS BLE maps.
+Novel methods and computational algorithms were employed along the way, resulting in potentially useful workflows and repositories that readers may follow to produce similar FIMs over the same regions (CONUS).
+
+Although there is a growing body of literature surrounding the use of relative elevation maps such as HAND for flood inundation mapping, this study scales the process to the continental scale and addresses many of the limitations encountered along the way.
+Some of the most novel findings in the paper come from the methodological choices employed, such as identifying the appropriate thalweg elevation in the presence of competing channels (DEM-based thalweg vs.
+hydrography-based thalweg).
+Another innovation is in the mosaicking of the FIMs in the presence of tributaries and lower-order streams.
+Another very useful innovation lies in 'burning' levee elevations onto the DEMs to enforce flood control measures on to the HAND surface.
+The approach detailed in this manuscript can be seen as a guide for mitigating the 'catchment boundary problem', the importance of which should be emphasized more extensively in this paper.
+
+As a reader, I find the manuscript title and some of the conclusions to be somewhat misdirected.
+The idea that Horton-Strahler stream order has a direct impact on the resulting HAND inundation product is misleading.
+The stream order of a given reach is determined by the drainage density, and as such is a product of 'upstream' decisions made during the derivation of the flow network.
+Stream order is not an input into the HAND model, nor is it very meaningful when comparing networks that are derived from different headwater locations (such as a pruned version of an existing network such as NHDPlusHR).
+Thus, simply reducing stream order should not have any impact on HAND.
+There are references in the paper to the impact of drainage density on maximum stream order, but the Conclusions indicate that Horton-Strahler reduction is the reason for improved FIM results, when that reduction is simply an associated result of reduced stream network density.
+For example, the authors indicate "We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed...".
+In reality, the reduction Horton-Strahler stream order is the result of, and not the cause for, a reduction in drainage density.
+It may be useful to emphasize that the discretization of the network into coherent regions based on level-path or the regionalization methods (such as using HUC8 boundaries), as well as the priority queue calculation (Equation 6) during the HAND raster mosaic process has a bigger impact on the result than the Horton-Strahler number of the network employed.
+
+Further revision could include a more thorough description of how using level-paths eliminates the catchment boundary problem, when each level path may have many tributaries (and thus multiple catchment boundaries).
+There is a bit of missing information in Section 2.5.2 that is necessary for understanding how the catchment boundary problem is mitigated.
+It seems that the way around the catchment boundary problem is to calculate HAND for a given level-path while leaving tributaries out of the analysis, and allowing the mosaic process to result in a HAND/FIM result with more extensive coverage over catchment boundary areas.
+
+Another limitation in the methodology lies in using synthetic rating curves (SRCs) that result from reach-averaged geometry parameter derived from the AGREE-DEM (a hydro-conditioning process).
+This results in an artificially impacted cross-section and bathymetry structure.
+The parameters used to constrain the AGREE method will have a profound impact on the SRCs derived from that DEM due to the excavation process in the AGREE method.
+
+The manuscript is, in places, difficult to understand due in part to the large number of acronyms and persistent jargon that pervades it.
+There are also many grammatical and syntactical errors that should be edited before re-submission.
+The attached annotated manuscript contains many suggestions for resolving these issues along with other comments and feedback, which the authors may find helpful.
+For this reason and those outlined above, I suggest to return the manuscript to the author for these (minor) revisions to increase the clarity of the methods and significance of the methodological choices on the results.
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+\begin{point}
+
+\end{point}
+
+\begin{reply}
+
+\end{reply}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% references and end document
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\clearpage % this clears figures before references
+\bibliography{rebuttal_bib}
+
 \end{document}

### 2022-08-05 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
* almost done with all of review #1s comments * just need to make method overview figure

@@
 \usepackage{soul}
+\usepackage{rotating}
 
@@
 \item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
-\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation.
+\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation extents.
 \end{keypoints}
@@
 \begin{abstract}
-The National Water Model (NWM) currently requires the post-processing of forecast discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
+The National Water Model (NWM) currently requires additional modeling from discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
 Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is worthy of producing high-resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
 However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
-To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that reduces the Horton-Strahler stream order of the stream network.
-GMS contains all segments within the NWM stream network but instead of deriving HAND by accounting for all river segments at once, it is derived independently at the level path (LP) scale.
+We propose here that HAND is limited to producing inundation only when sourced from its nearest drainage line thus lacks the ability to source inundation from multiple fluvial sources.
+To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
 LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
-These FIMs are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
-Improvement in mapping skill is observed when compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
+The FIMs associate from each independent LP are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
+Improvement in mapping skill is observed when the extents are compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
 A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
+This novel contribution is proposed and studied in a new open-source implementation that utilizes the latest combination of hydro-conditioning techniques to enforce drainage and counter limitations in the elevation data.
+This framework scales to the entire NWN modeling domain including Hawaii and Puerto Rico to help forecasters provide timely and accurate flood warnings.
 \end{abstract}
@@
 We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-We compared these HAND derived FIMs to maps from more realistic models and found improvement in mapping performance.
+We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance.
 %
@@
 With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
-Trends related to flood damages and fatalities have been steadily increasing over recent decades.
-\cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
+Trends related to flood damages and fatalities have been steadily increasing over recent decades \cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
 Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
@@
 \includegraphics[scale=1.0]{figures/forecast_points.jpg}
-\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
-Also show are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
+\caption{Forecast points with and without Flood Inundation Maps (FIM) in United States' Advanced Hydrologic Prediction System (AHPS).
+Note that only a small fraction of the AHPS forecast points have existing FIM.
+Also shown are the National Water Model (NWM) stream networks at the full resolution (FR) and Mainstems (MS) resolution.
+The FR network constitutes the entire NWM stream network while the MS resolution network is the FR network at or downstream of the AHPS forecast points shown.
+}
 \label{fig:forecast_points}
@@
 Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the NWM which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
 The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
@@
 The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
-Figure \ref{ssec:national_water_model} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
+Figure \ref{fig:forecast_points} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
 The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
@@
 The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
+The NWM provides stream forecasts at these hydrofabric segments using the Muskingham-Cunge method to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
 Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{HAND's Assumptions and Limitations}
+\label{sssec:hand_assumptions_limitations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+HAND operates on many underlying assumptions since it can only be used as an inundation proxy and not a true hydrodynmic inundation model.
+HAND, to our knowledge, has only been applied for riverine inundation applications thus it is also missing pluvial, coastal, ground water and dam break components among other possible sources of flooding.
+Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest stream line which is used for catchment allocation and relative elevation calculation.
+Since elevations don't naturally do this, they must undergo a long series of hydro-conditioning processess to enforce monotonically decreasing elevations across an entire processing unit along with hydrologically correct directions of flow.
+The level of DEM manipulation required to enforce this assumption can be substantial depending on the region and can be a significant source of error.
+The drainage enforcing assumption also interacts with an inabilty to properly account for fluvial inundation in regions of DEM depressions.
+
+True to its name, when HAND, where the D denotes drainage, is used for FIM applications it assumes that a given area's drainage area is the only reach it eligble to source inundation from.
+Catchments are independent of one another for FIM purposes meaning a reaches' stage value is only used to threshold the HAND values within its respective catchment.
+This assumption plays to the ``nearest drainage'' term in HAND and creates a significant limitation within HAND for FIM applications \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
+At the junction of high stream order and high flow rivers with lower flow tributaries, there can a lack of inundation extents exhibited which is known colloaquially in the forecasting community as the catchment boundary problem.
+The academic community has somewhat referenced this issue before but it has been characterized more as a problem to due with the stream delineation process that comes from thresholding the drainage accumulation maps.
+Later in this study, we will re-introduce this problem and demonstrate how we start with a starting stream network, that of the NWM's, and thus avoid having to threshold accumulations to some arbitrary value to define stream networks.
+We illustrate how computing HAND independently for stream lines of unit stream order can significantly enhance FIM performance by accounting for multiple sources of fluvial inundation that may exist in certain regions and flow scenerios.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{HAND Implementations}
@@
 Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
-As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
+As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a HPC \cite{maidment2017conceptual,liu2016cybergis}.
 NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
@@
 
-Related to these efforts, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
+Related to these efforts, the United States Geological Survey (USGS) has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
 Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{OWP FIM}
+\subsection{Office of Water Prediction Flood Inundation Mapping}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
-Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
-This capability will be referred to as Generalized Mainstems (GMS).
+Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which is closely related to stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
+Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
+This capability is introduced progressively as Mainstems (MS) and to a higher degree Generalized Mainstems (GMS) which will be explained later on.
 The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
@@
 \subsection{Software Dependencies and Architecture}
+\label{ssec:software}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}.
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
-For enforcing levee data into the DEMs, the USACE National Levee Database (NLD) is used to burn feature elevations \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{OpenStreetMap,GreatLakesHydrographyDataset}.
+We utilized a series of data products from the NHDPlusHR including the BurnLineEvents \cite{nhdplus2022vectors}, Value Added Attributes (VAA) \cite{nhdplus2022vectors}, Water Boundaries (WBD) or HUC Layers \cite{nhdplus2022wbd}, and the DEM elevation rasters \cite{nhdplus2022dems}.
+It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes \cite{nwm2022hydrofabric}.
+For enforcing levee data into the DEMs, the USACE NLD is used to burn feature elevations \cite{engineers2016national}.
+Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{osm2021landsea,GreatLakesHydrographyDataset}.
 Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
@@
 %
-\begin{table}
-\caption{Data sources, names, and descriptions.}
+\begin{sidewaystable}
+\caption{Data sources, names, descriptions, and citations.}
 \label{tab:data}
 \centering
-\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
+\begin{tabular}{|p{1.75cm}|p{4cm}|p{11cm}|p{5cm}|}
 %\begin{tabular}{l c c}
 \hline
-Source & Name & Description \\
-\hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
-\hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
-\hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters \\
-\hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. \\
-\hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. \\
-\hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. \\
-\hline
-USACE & NLD & Levee database of locations and elevations  \\
-\hline
-OSM & Land-Sea Border & Border of land and sea. \\
-\hline
-GLHD & Land-Great Lakes Border & Border of land and Great Lakes. \\
-\hline
-InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
-\hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
+Source & Name & Description & Citations\\
+\hline
+USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement & \cite{nhdplus2022vectors} \\
+\hline
+USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display & \cite{nhdplus2022vectors} \\
+\hline
+USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters & \cite{nhdplus2022dems} \\
+\hline
+USGS & NHDPlusHR WBD & Water Boundaries (WBD) or HUCs used for spatial processing units & \cite{nhdplus2022wbd} \\
+\hline
+NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. & \cite{nwm2022hydrofabric} \\
+\hline
+NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. & \cite{nwm2022hydrofabric} \\
+\hline
+NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. & \cite{nwm2022hydrofabric} \\
+\hline
+USACE & NLD & Levee database of locations and elevations & \cite{engineers2016national} \\
+\hline
+OSM & Land-Sea Border & Border of land and sea. & \cite{osm2021landsea} \\
+\hline
+GLHD & Land-Great Lakes Border & Border of land and Great Lakes. & \cite{GreatLakesHydrographyDataset} \\
+\hline
+InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
+\hline
+InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
 \hline
 \end{tabular}
-\end{table}
+%\end{table}
+\end{sidewaystable}
 %
@@
 %
-The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
+The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping with HAND. 
 These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
@@
 %
-The location of the stream network is enforced to ensure general agreement with established stream networks.
+The location of the stream network is enforced to ensure general agreement with established stream networks especially the placement of the NWM network which is used for forecasting the stream flow inputs.
 The NHDPlusHR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
-However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
+However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus MR, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
 For every NWM headwater segment a headwater point is derived and linearly interpolated to the nearest Burnline Event segment.
 Burnline Event headwater segments are split at the adjusted headwater point to match NWM flowlines.
-The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
+The resulting pruned NHDPlusHR stream network is what gets hydro-enforced in subsequent operations.
+This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHDPlusHR network corresponds to the full density NHDPlusHR network at NWM V2.1 headwater locations only.
 Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
+This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents. 
 %
@@
 %
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.
 
@@
 The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
+In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop which were set to fixed values of 70m, 1000m, and 10m, respectively.
+While the value of these parameters are critical to the inundation extents produced, especially for lower flow rates, finding their optimal value for OWP FIM was not done since it was out of the main scientific scope of this article.
 Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
@@
 %
-The DEM at 10 m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
+Coarse DEM's at 10m, 30m, and higher resolutions can lack sufficient representation of fine grain features such as embankments, floodwalls, and closure structures \cite{arundel2018assimilation,dobbs2010evaluation,wang2005comparison,sanders2007evaluation}.
 In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10 m DEM.
@@
 The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
-\citeA{garousi2019terrain} proposed an algorithm that traverses stream thalweg pixels in a depth first manner starting with adding all the headwater pixels to a queue. 
-The connectivity of the thalweg pixels is defined by the D-8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
-At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in Equation \ref{eq:thalweg_breach},
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:thalweg_breach}
-\textbf{D}[x] = \min_{y\ drains\ to\ x} {(\ \textbf{D}[x]\ ,\ \textbf{D}[y]\ )}
-\end{equation}
-\end{linenomath*}
-%
-, in which \textbf{D} represents the array of thalweg adjusted elevations indexed by x and y where by y is upstream of x. 
-When a pixel's upstream neighbors are all evaluated, the downstream pixel is added into the queue thus the depth first traversal of the drainage network.
-This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in Critical Success Index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
+\citeA{garousi2019terrain} proposed an algorithm that breaches stream thalweg pixel elevations in a depth first manner. 
+This procedure was found to increase the Critical Success Index of resulting FIMs from HAND and is employed in OWP FIM to enforce monotonically decreasing elevations with thalweg pixel networks.
 %
@@
 The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
-The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southeast direction as 8. 
 Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
@@
 The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
-Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
-By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
+OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
 %
@@
 Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
+
 Each stream link in this derived FIM stream network is split into equidistant reaches.
@@
 The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
-Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling. 
+Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling \cite{djokic2019arc}.
 %
@@
 %
-FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
-Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
+As previously discussed, HAND based FIMs are subject to many assumptions and limitations in order serve as a suitable inundation proxy for large scale, high resolution domains.
+HAND produced FIMs are limited in only providing inundation sourced from the nearest drainage which in many instances can only be of a several sources of fluvial surface water inundation.
+The forecasting community, in reviewing HAND, have noted significant negative effects at the confluence of lower flow tributaries with higher flow rivers for which the term catchment boundary issue has been termed.
+In previous studies, FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND which is closely related to the maximum Horton-Strahler stream order of the network \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+This sensitivity is partly in due to the limitation that catchment boundaries place on inundation extents.
+
+Figure \ref{fig:catchment_boundaries_issue} illustrates the exact situation our solution proposes to address where two tributaries converge with a higher order stream segment. 
 An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the MS of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
+The higher discharge along the main segment in Figure \ref{fig:catchment_boundaries_issue} of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS.
 This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the MS that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the MS catchments.
+As a parallel problem, there is excess water accumulated along the mainstem that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the mainstem catchments.
 %
@@
 \includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{False negatives associated with confluence of tributaries with MS. Integers represent flow values from BLE 100 year event for the associated areas. 
-No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
+\caption{The figure represents an agreement map between a HAND derived FIM and one produced from the Base Level Engineering (BLE) program for a 100yr magnitude event at HUC8 12090301.
+Agreement maps are symbolized by false negatives (FN), true negatives (TN), false positives (FP), and true positives (TP) where inundated represents the postive condition (see Section \ref{ssec:evaluation} for more details).
+The stream flows associated with each river segment are shown in CMS while the flow directions are symbolized as red arrows.
+The presence of FNs at the confluence of tributaries (circled in red) with the main segment is associated with lower flow rates in the tributaries that don't account for backwater effects.
+Additionally, the flow of 1900 CMS from the main segment cannot extend to the neighboring catchments belonging to its tributaries shown here.
+Water pools up vertically along the catchment boundaries of the higher order segment distorting rating curve behavior (Section \ref{sec:discussion}).
+Sourcing fluvial inundation from HAND is limited to only its nearest drainage line which is the main issue this study aims to address.
+}
 \label{fig:catchment_boundaries_issue}
@@
 %
-We seek to resolve this problem by deriving HAND for processing units with stream networks of reduced stream order. 
-We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
-The resulting FIMs from the overlapping HAND processing units are mosaiced together taking any inundated area to be inundated but more will be explained in Section \ref{ssec:inundation_mapping}.
+
+We seek to resolve this catchment boundary problem or nearest drainage limitation by discretizing the target stream network into stream networks of reduced, unit stream order to avoid the constraining of catchments by those belonging to lower order neighbors.
+By discretizing the network into stream networks of unit stream order (later defined as level paths), we remove the influence of neighboring catchments that constrain the inundation extent.
+This creates much larger and overlapping catchments that can source fluvial inundation from multiple reaches as required by the given river stage at current flow conditions.
+We present two successive methods, National Weather Service MS (Section \ref{sssec:nws_mainstems}) and GMS (Section \ref{sssec:generalized_mainstems}), implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND by expanding the nearest drainage definition to increase potential inundation areas.
 %
@@
 %
+The initial attempt at drainage order reduction to solve the catchment boundary issue was to use a stream network relevant to the NWS forecasting community. 
 The Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
@@
 It also spans 121,724 reaches across 1,608 HUC8s.
-HAND was originally derived for this stream network to enhance mapping skill along these critical MS segments \cite{djokic2019arc}. 
-The inundation derived from this stream network is mosaiced with the inundation from the FR network to form the MS FIMs. 
+In this technique, we derive HAND using the FR stream network as well as the MS network which was originally proposed by \citeA{djokic2019arc}.
+Inundation is derived independently from the resulting FR and MS HAND hydrofabric are mosaiced together using the technique proposed in Section \ref{ssec:inundation_mapping} to form the MS FIMs. 
 Within each HUC, you'll typically only find a MS stream network of stream order 1 (i.e.
 headwater) but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
+So while we may refer to the MS network as that of one with unit stream order, we acknowledge there are many cases where additional or converging forecast points create multiple branches within a given processing unit.
 %
@@
 %
-To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call GMS.
-Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
+Since MS only covers 4\% of the entire FR stream network, we sought to expand drainage order reduction techniques to all reaches within the NWM modeling domain.
+In order to do this, we discretized the NWM network into level paths which when considered individually have unit Horton-Strahler stream orders thus significantly reducing drainage density.
 Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
@@
 For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
+
 Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
 Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
+This buffer was selected to avoid edge contamination \cite{lindsay2013measuring} and to ensure adequate data availability for wide rivers with large catchments in regions with low slope.
+Further work could be dedicated to tune this user exposed parameter to better balance its effect on FIM extents and computational expense since larger buffers create additional floating points and calculations to process and store.
 At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
 The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
+Once the NWM FR stream network is discretized into level paths, we independently compute HAND using each level path as the target stream network to be used.
 
@@
 Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
-Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}c.
+Lastly, these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}d.
+
+For a more intimate look at the drainage order reduction procedure and its effects, GMS, we allude to Figure \ref{fig:gms_methods_2} which references the same area (in HUC8 12090301) and set of river junctions as in Figure \ref{fig:catchment_boundaries_issue}.
+The catchments and stream lines for HAND computed at the FR scale are illustrated in Figure \ref{fig:gms_methods_2}a where the respective inundation (for 100 yr magnitude) is heavily constrained by the limited catchment extents especially at junctions.
+In subsequent sub-figures, we show the same datasets for the HAND computation problem for this region but discretized into independent level paths for the main level path (b), the eastern tributary (c), and the western tributary (d).
+Notably, inspecting (b) one sees how removing the tributaries creates much larger catchments for the main level path. 
+These catchments include drainage areas that would traditionally be considered nearest to the tributaries thus ineligible to receive inundation sourced from the main level path.
+The inundation extents in (b) overlap those of (c) and (d) and are mosaiced together by methods explained in Section \ref{ssec:inundation_mapping}.
 %
@@
 \includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates how level paths for the NWM are derived.
-Level paths symbolized by lines of unique colors are propagated upstream following the direction of maximum arbolate sum at each junction.
+\caption{Illustrates how level paths (LP) for the NWM stream network are derived.
+They are derived by starting at an outlet point with a unique ID which are symbolized by unique colors.
+Once a junction is reached, the current level path identifier is propagated in the direction of highest arbolate sum.
+The remaining converging members at the given junction are each assigned a new unique LP identifier and the process is repeated recursively until all reaches have been assigned a level path.
+The values shown labeling the reaches above are arbolate sums which is defined as the cumulative distance of all upstream reaches.
+Thus, level paths serve as a proxy means of assigning membership to a given river when presented with a confluence.
+Each individual level path has a unit Horton-Strahler stream order thus serves as a great method for our proposed technique.
 }
@@
 \includegraphics[scale=1.0]{figures/gms_methods.jpg}
-\caption{Overall procedure for GMS HAND.
-In (a), we illustrate all NWM stream lines symbolized by their level path.
+\caption{Overall procedure for GMS HAND at HUC8 12090301.
+In (a), we illustrate all NWM stream lines symbolized by their level path with 372 unique level path IDs in this HUC.
 Meanwhile (b), demonstrates the DEM clipped to a 7 km buffer around two selected level paths.
@@
 %
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.70]{figures/gms_methods_2.jpg}
+\caption{This image, with the same spatial domain as Figure \ref{fig:catchment_boundaries_issue} (HUC 12090301), demonstrates how computing HAND on level path bases leads to larger, independent catchments and more expansive inundation extents (100 yr flows).
+In (a), the catchments and stream network are shown for HAND computed in Full-Resolution (FR) method which shows constrained inundation extents around the two junctions.
+(b) demonstrates the level path associated with this region's highest order river.
+By delineating catchments at this scale indepedent of the neighboring tributaries, the drainage areas are allowed to expand thus allowing inundation extents to cover previously restricted areas.
+In (c) and (d), we show the stream lines, catchments, and inundation extents of the two tributaries.
+Later in Section \ref{ssec:inundation_mapping}, we describe how the inundation in (b), (c), and (d) are mosaiced together to form one seamless inundation map.
+This process allows for multiple, possible contributing sources of fluvial inundation to be considered thus enhancing FIM skill.
+}
+\label{fig:gms_methods_2}
+\end{figure}
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
+Consolidating the depths using a maximum function was decided upon based on intuition which we believe to best represent the depth of water in an area with multiple contributing fluvial inundation sources.
+Other aggregation methods could lead to different results but were not investigated here.
 Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
@@
 %
+Possible benchmark FIM candidates for evaluation purposes include high-water marks, remote sensing observation, crowd-sourced information, and modeled extents.
+Unfortunately, all of these sources or using a combination of them are subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to lack of spatial coverage, signal interference, lack of stream flow data, inaccurate stream flow data, physics-based assumptions, and errors in input data.
+While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated stream flow data necessary to make FIMs to compare to.
+A suitable dataset needed to be selected to best balance advantages and limitations with future studies interested in expanding to additional datasets as able to.
+
 Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
+This dataset was selected due to its large spatial coverage, availability of cross-sections with stream flow information, more sophistication when compared to HAND based FIMs, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
 49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
@@
 We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
+
 By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
@@
 \includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
-\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states. This dataset for 1\% and 0.2\% recurrence flows were used as benchmarks.}
+\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states at 0.2\% recurrence magnitude for flow rates.}
 \label{fig:all_ble_maps}
@@
 \includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
-\caption{Illustrates Base Level Engineering (BLE) cross sections and stream lines.
-The BLE stream network, which is denser than the NWM V2.1 stream lines, is also shown.
-BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM reach.
+\caption{
+Illustrates Base Level Engineering (BLE) cross sections and stream lines at the HUC8 12100203 near the confluences of West Fork Plum Creek and Clear Fork Plum Creek with Plum Creek.
+BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM V2.1 Full Resolution (FR) stream lines.
+Additionally, we illustrate the NWM V2.1 catchments to provide a sense of how many cross-sections may intersect a given NWM flow line.
+The BLE stream network is also shown which is denser than the NWM V2.1 stream lines meaning there are several lower order streams represented in the BLE stream network that are not in the NWM V2.1 stream lines.
+This creates additional inundation areas in the validation data that are not modeled with our HAND based FIMs.
 }
@@
 %
+
 The metrics employed in this study to evaluate inundation extents include Critical Success Index (CSI), Probability of Detection (POD), and False Alarm Ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
@@
 Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
+Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
 Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
+In other words, POD is an indicator of model skill in inundated regions while FAR is an indicated of model skill in non-inundated regions.
+We use all three secondary metrics here to add value to the discussion while avoiding aggregating away the meaning of all four primary metrics.
+
 While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Mapping Performance}
-\label{ssec:mapping_performance}
+\subsection{Flood Mapping Performance}
+\label{ssec:flood_mapping_performance}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
-We computed the statistics (CSI, POD, and FAR) for both 100 and 500 year events for Mannings N set to 0.06 and 0.12. 
-The distribution of these statistics can be examined in Figure \ref{fig:violin_plot} as violin plots.
+We computed the statistics, CSI, POD, and FAR, for both 100 and 500 year events for Mannings N set to 0.06 and 0.12.
+These results are presented in Figure \ref{fig:violin_plot} as violin plots and in Table \ref{tab:aggregate_metrics} as aggregated metrics with the results for MS and GMS presented as percentage changes from their respective FR values.To be more specific, Table \ref{tab:aggregate_metrics} sums the primary metrics, TP, FP, FN, and TN, across all HUC8's then recomputes the secondary metrics which was done to better account for large variances in HUC8 size.
+The same trends discussed below are consistent across both reporting methods (Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}).
+
+The distribution of these flood extent metrics can be examined in Figure \ref{fig:violin_plot} as violin plots.
 Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
-We also denote trend lines for each metric and Manning's n setting as well as their respective slope estimate and one-tailed p-value denoting the level of significance of the trend.
-
-Aggregating the metrics in the method above treats each HUC8 as it's own unit and does little to consider the size differences of the HUCs. 
-In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics}  the CSI, POD, and FAR recomputed for the entire domain using the sum of all the TPs, FPs, and FNs. 
+For example, let's examine the violin plot for the row marked CSI and column for Manning's n = 0.06.
+This sub-figure shows the CSI distributions across all 49 HUC8's when Manning's n is set to 0.06.
+Each independent, whole violin represents the distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100yr and orange for 500yr).
+The horizontal dashed and dotted lines represent the 25th, 50th, and 75th percentiles from bottom to top, respectively.
+Additionally, we show trend lines symbolized in green that for each metric and Manning's n combination denotes the best fit line for the three methods (FR, MS, and GMS).
+To avoid having two trend lines per sub-figure, we elected to aggregate the two magnitudes togther as they tend to observe similar trends across the three models.
+The slope of each trend line is quantified in the Figure by its $\beta_1$ value and the p-value for that statistic which tests the significance of that trend (deviation from a zero sloped trend line).
+
+% Interpretation of metrics
+Both Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} contain a fair amount of information that is central to the objectives of this study.
+As previously stated in Section \ref{ssec:evaluation}, we consider CSI as a general proxy for the skill of the inundation extents with POD denoting skill on inundated areas and FAR indicating skill on non-inundated areas.
+Again, the main objective of the study is to introduce how computing HAND with disaggregated stream networks to those with unit stream order can enhance the fidelity of FIMs by capturing fluvial inundation from multiple sources as opposed to that of just the nearest drainage line.
+As can be seen in Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, CSI generally increases from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
+This increase is primarily driven by an increase in POD thus generally increasing the probability of correctly detecting inundation.
+Also, we note that FAR is somewhat, albeit marginally, decreased from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
+The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}.
+So overall and in other words, the entire distribution of HUC8's improves across the three methods.
+Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
+We move more details on the relationship betweeen stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).
+
+Additional noteworthy trends in Figure \ref{fig:violin_plot} center around the all-around better performance of FIM's for those of higher Manning's n values and recurrence flows.
+The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value used in the BLE model, 0.11 \cite{fema2021base,fema2021estimated}.
+Most of this improvement is driven by significant increases in POD but, unfortunately, leading to a significant amount of overprediction as observed by the increase in FAR.
+More work can be invested to better regionalize Manning's n values for FIM purposes with HAND.
+We also observe additional trends associated with the magnitude or recurrence interval of the flow rates used with the higher flow rates exhibting better overall CSI, POD, and FAR values than the lower, 100 yr magnitude.
+We introduce in the Discussion (Section \ref{sec:discussion}) that this skill premium exhibited by higher flow events could be due higher quality elevation data in regions that are not described as bathymetric areas.
 %
@@
 \begin{table}[h!]
-\caption{Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated for BLE domain.
+\caption{ Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated to the BLE domain.
+         The values for MS and GMS methods are expressed in percentage change (\%) from their respective values with the same Manning's N, magnitude, and metric combination in the Full Resolution (FR) method columns.
          The best value across models is highlighted in bold.}
@@
 \hline
-\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{MS} & \multicolumn{2}{|c|}{GMS} \\
+\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{\shortstack{MS\\(\% Change)}} & \multicolumn{2}{|c|}{\shortstack{GMS\\(\% Change)}} \\
 \cline{3-8}
@@
 \hline
-\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & \textbf{0.5796} & \textbf{0.6075} \\
+\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 2.53 & 2.59 & \textbf{3.95} & \textbf{4.04} \\
 \cline{2-8}
-  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & \textbf{0.6182} & \textbf{0.6435} \\
-\hline
-\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & \textbf{0.6633} & \textbf{0.6863} \\
+  & 0.12 & 0.5915 & 0.6149 & 2.35 & 2.26 & \textbf{4.51} & \textbf{4.65} \\
+\hline
+\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 2.68 & 2.74 & \textbf{4.39} & \textbf{4.38} \\
 \cline{2-8}
-  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & \textbf{0.7606} & \textbf{0.7810} \\
-\hline
-\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1787 & \textbf{0.1589} & \textbf{0.1778} & \textbf{0.1589} \\
+  & 0.12 & 0.7255 & 0.7446 & 2.83 & 2.71 & \textbf{4.84} & \textbf{4.89} \\
+\hline
+\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & -0.72 & \textbf{-1.24} & \textbf{-1.22} & \textbf{-1.24} \\
 \cline{2-8}
-  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & \textbf{0.2324} & \textbf{0.2148} \\
+  & 0.12 & 0.2379 & 0.2208 & -0.21 & -0.18 & \textbf{-2.31} & \textbf{-2.72} \\
 \hline
@@
 \end{table}
-%
-% Interpretation of metrics
-From Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, we denote several meaningful trends. 
-Using CSI as an overall proxy for skill of the FIM, we note that generally speaking the skill is correlated with a reduction of the stream orders of the processing units used for HAND.
-In other words, the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
-While this trend is evident for both sets of Manning's n values, the trend is slightly more significant for the higher value of 0.12.
-Other trends related to this Figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
-We also note how the higher Manning's n value enhances performance for both of these recurrence intervals across all models.
-
-Dissecting the improvements and trends presented in the previous paragraph comes down mostly to improvement in POD or a reduction in absolute amount of FNs.
-POD being the primary driver in skill enhancement is evident across models by comparing the slope of the POD lines with the slope of the FAR lines.
-Even though aggregating metrics by HUC8 yields a statistically zero trend, one does see a slight reduction in FAR across models that reduce HAND's maximum stream order.
-Additionally, we note that POD is a primary driver in enhancing performance across Manning's n values as well.
-This significant improvement comes at a cost of false alarms as the FAR increases significantly across Manning's n values.
 %
@@
 %
-Overall, we note a positive relationship between FIM skill and a reduction of the stream order of the stream network we use to derive the HAND datasets.
-Most of this change is accounted for by increasing POD thus reducing FNs especially along higher order rivers with higher flow magnitudes.
-We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
-This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
-
-In analyzing the data, we found a slight reduction in FAR was detected and more digging pointed to a bias in rating curves introduced by stream order reduction.
+Overall, the main observation of this study was how FIM performance can be improved by reducing the Horton-Strahler stream order of the target stream network used for HAND computation.
+Most of this change is accounted for by substantially increasing POD and inundation extents in some areas thus reducing FNs.
+We believe, as we later argue for, that the increase in POD is primarily driven by an increase in the catchment sizes that is inhertently enabled by dividing up the stream network into independent stream networks of unit stream order.
+Additionally, we note that reducing drainage order also has some minor influence on reducing inundation extents and the rate of false alarms.
+We believe that this effect is driven by a reduction in some inundation areas led by a change in the stage-discharge relationship where a given streamflow now leads to lower river stage values when HAND is computed with target stream networks of unit drainage order.
+We seek to explain that these two effects, catchment boundary enlargment and stage-discharge curve lowering, are highly interrelated and cannot be easily detangled.
+Lastly, we discuss the diminishing effects on performance that the MS and GMS techniques may have and also any additional effects including enhanced cross-walking abilities.
+
+As evident in the results of the study in Section \ref{sec:results}, a sizeable amount of the increase in CSI observed by reducing stream order for HAND computation can be attributed to increases in POD.
+This can be inferred by close inspection of Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} where changes in POD are significantly higher than that of FAR.
+This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
+
+Upon investigation of the performance of HAND derived FIM, we note a general increase of catchment sizes for the MS and GMS methods when compared to the FR method as they are now delineated independently of any tributaries that would constrain catchment sizes otherwise.
+Additionally, we note significantly less water build up along catchment boundaries especially at the junction of lower order tributaries with lower flow rates and higher order rivers with more flow.
+This allows for inundation extents to expand across regions previously encapsulated by catchments of joining reaches in lower flow tributaries.
+The water built up along the catchment boundaries can be thought of as a column of water in a cyclindrical container (catchment) that has exceeded the elevation of the containers rim.
+This, of course, does not represent accurate physics but it is what does occur in HAND when water accumulates along catchment boundaries.
+
+Figure \ref{fig:gms_enhancement} best illustrates this behaviour in real context.
+The figure is comprised of two agreement rasters for two different HAND based FIMs compared to the validation dataset for a given region with a high flow mainstem (500 yr recurrence flow) running horizontally along the region.
+Sub-figure \ref{fig:gms_enhancement}a demonstrates the agreement raster for the FR stream network as well as it's respective catchment boundary lines symbolized in white and stream network shown in green.
+Inspection of this sub-figure denotes a clear spatial pattern where TP's or areas correctly inundated are pooled alongside catchment boundary lines. 
+On the other side of the catchment boundary, one can witness large swaths of FN's that should be inundated. 
+The FN's also exhibit a spatial pattern as in they tend to collocate within catchments of the pictured mainstems tributaries.
+This sort of behaviour was introduced early in the paper and shown qualitatively in Figure \ref{fig:catchment_boundaries_issue}.
+
+As an enhancement, this paper proposes computing HAND for stream networks comprised of level-paths independently of one another.
+In sub-figure \ref{fig:gms_enhancement}b, the agreement raster for the GMS technique is illustrated as well as the stream network lines in green.
+While the entire mosaiced inundation map from GMS (as described in Section \ref{ssec:inundation_mapping} and Equation \ref{eq:hand_fim_depth}) is used to produce this agreement map, we only show the catchments associated with the mainstem of this region that is shown to follow a clear horizontal path.
+Showing all the catchments for the tributaries that were all derived independently would convulate the image.
+The main message illustrated here is that the catchments associated with the mainstem of this area significantly increase in size.
+Since they are not restricted by the catchments of tributaries that lie in the same drainage areas as those of the mainstem, they extend to include those as well.
+The consequence for inundation extents is a general increase in spatial coverage of the river's water which shows to be in much better agreement with the benchmark map.
+The TP's are no longer bounded by the catchment lines and allowed the expand to their natural extents.
+
+We note here as a contribution of this study that a major inherant, limitation of HAND is the ``nearest drainage'' constraint or the idea that a given river reach only drains or, in HAND's case, inundates its immediate, unique drainage area or catchment.
+In other words, HAND based FIM's are limited to producing fluvial inundation to only their nearest drainage area or catchment.
+However, we know that fluvial inundation can be sourced from several streams nearby that also serve as drainage outlets to the area in question.
+Generally speaking, drainage areas are known to be hiearchical in nature so a given drainage area for a given outlet point can be decomposed into nested drainage areas for outlet points that lie in the original drainage area.
+A perfect example of this are points that lie on tributary reaches closely neighboring a mainstem.
+These points lie in the drainage area of reaches in the mainstem but inundation from the mainstem cannot reach these tributary catchments because of the "nearest" assumption in HAND.
+Hence it's important to state that just like there are different sources of flooding such as fluvial, pluvial, groundwater, reservoir, and coastal, there can also be multiple sources of a riverine flood.
+HAND is only equipped to handle rivirine flooding from the nearest drainage line.
+Other relevant drainage lines that produce fluvial flood waters are not considered here especially if the routing model used doesn't consider backwater effects.
+
+The remaining portion of the improvement in CSI was found to come from a marginal yet noteable reduction in FAR.
+Upon investigation of the spatial results in the agreement maps, we found some areas of slight reductions in FPs especially where changes in catchment boundaries may have occurred due to the reduction in stream order.
+These observations pointed to changes in the rating curves introduced by stream order reduction and catchment definition adjustments.
 Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
@@
 This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
+
 Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
@@
 Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
+
 Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments that encompass neighboring tributaries.
-These tributaries have a significant amount of bathymetry that is low-lying thus easily including the SRC derivation. 
+These tributaries have a significant amount of bathymetry that is low-lying thus easily included in the geometry for the SRC derivation. 
 They also contribute volume and bed area that is technically not perpendicular to the flux of streamflow being accounted for in the stream in question. 
@@
 We consider this fact to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
+
+We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
+This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
+One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
 
@@
 Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
-The NWM stream network was derived using the NHD medium resolution dataset which was derived from coarser DEMs than those used here. 
+The NWM stream network was derived using the NHDPlus MR  dataset which was derived from coarser DEMs than those used here. 
 Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
@@
 Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
+We suspect that a significant amount of the difference in performance between 100yr and 500yr magnitude events can be attributed to poor synthetic rating curve performance due to poor bathymetric representation.
+Lower magnitude events are, logically, more susceptible to poor bathymetric data due a greater proportion of the inundation being attributable to areas that are more typically under normal flow conditions.
+Higher flow events tend to cover regions
 
@@
 \includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
-\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105.
+\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105 at the 500yr recurrence magnitude.
 Catchment boundaries and stream lines are shown in white and dotted green, respectively.
@@
 Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
+FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
 We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
 This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
-FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
+
+Dissecting the stream network into regions of level-paths with unit stream order is necessary because HAND has a ``nearest drainage'' limitation meaning it only accounts for riverine inundation sourced from the nearest drainage line.
+While this is all that is required in many, lower flow instances, typically larger flow scenerios can call for multi-source riverine flooding to occur especially at the junction of major rivers with their smaller tributaries.
+Computing HAND independently for each level-path removes the boundary of neighboring catchments thus allowing for 
 Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.
@@
 Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
-%The [type of data] data used for [brief context, description] in the study are available at [repository, source name] via [DOI, persistent identifier link] with  [license, access conditions] [optional in-text citation in References]
-%[Version number] of the [software name] used for [brief context, description of what the software was used for] is preserved at [DOI, persistent identifier link], available via [license type, access conditions] and developed openly at [software development platform link].* [optional in-text citation in References].
+All data availability and open research information is cited in the Software and Datasets sections, \ref{ssec:datasets} and \ref{ssec:software}, respectively.
+Please see those sections, Table \ref{tab:datasets}, and their References for more information.
 

### 2022-08-05 manuscripts/owp_fim4_2022/sections/conclusions.tex
spelling updates
* almost done with all of review #1s comments * just need to make method overview figure

@@
 Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
+FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
 We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
 This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
-FIM skill was evaluated over large spatial scales by comparsion to HEC-RAS 1D models.
+
+Dissecting the stream network into regions of level-paths with unit stream order is necessary because HAND has a ``nearest drainage'' limitation meaning it only accounts for riverine inundation sourced from the nearest drainage line.
+While this is all that is required in many, lower flow instances, typically larger flow scenerios can call for multi-source riverine flooding to occur especially at the junction of major rivers with their smaller tributaries.
+Computing HAND independently for each level-path removes the boundary of neighboring catchments thus allowing for 
 Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.

### 2022-08-05 manuscripts/owp_fim4_2022/sections/discussion.tex
minor changes
* almost done with all of review #1s comments * just need to make method overview figure

--- 2022-01-10 manuscripts/owp_fim4_2021/sections/discussion.tex
@@
 %
-Overall, we note a positive relationship between FIM skill and a reduction of the stream order of the stream network we use to derive the HAND datasets.
-Most of this change is accounted for by increasing POD thus reducing FNs especially along higher order rivers with higher flow magnitudes.
-We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
-This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
+Overall, the main observation of this study was how FIM performance can be improved by reducing the Horton-Strahler stream order of the target stream network used for HAND computation.
+Most of this change is accounted for by substantially increasing POD and inundation extents in some areas thus reducing FNs.
+We believe, as we later argue for, that the increase in POD is primarily driven by an increase in the catchment sizes that is inhertently enabled by dividing up the stream network into independent stream networks of unit stream order.
+Additionally, we note that reducing drainage order also has some minor influence on reducing inundation extents and the rate of false alarms.
+We believe that this effect is driven by a reduction in some inundation areas led by a change in the stage-discharge relationship where a given streamflow now leads to lower river stage values when HAND is computed with target stream networks of unit drainage order.
+We seek to explain that these two effects, catchment boundary enlargment and stage-discharge curve lowering, are highly interrelated and cannot be easily detangled.
+Lastly, we discuss the diminishing effects on performance that the MS and GMS techniques may have and also any additional effects including enhanced cross-walking abilities.
 
-In analyzing the data, we found a slight reduction in FAR was detected and more digging pointed to a bias in rating curves introduced by stream order reduction.
+As evident in the results of the study in Section \ref{sec:results}, a sizeable amount of the increase in CSI observed by reducing stream order for HAND computation can be attributed to increases in POD.
+This can be inferred by close inspection of Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} where changes in POD are significantly higher than that of FAR.
+This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
+
+Upon investigation of the performance of HAND derived FIM, we note a general increase of catchment sizes for the MS and GMS methods when compared to the FR method as they are now delineated independently of any tributaries that would constrain catchment sizes otherwise.
+Additionally, we note significantly less water build up along catchment boundaries especially at the junction of lower order tributaries with lower flow rates and higher order rivers with more flow.
+This allows for inundation extents to expand across regions previously encapsulated by catchments of joining reaches in lower flow tributaries.
+The water built up along the catchment boundaries can be thought of as a column of water in a cyclindrical container (catchment) that has exceeded the elevation of the containers rim.
+This, of course, does not represent accurate physics but it is what does occur in HAND when water accumulates along catchment boundaries.
+
+Figure \ref{fig:gms_enhancement} best illustrates this behaviour in real context.
+The figure is comprised of two agreement rasters for two different HAND based FIMs compared to the validation dataset for a given region with a high flow mainstem (500 yr recurrence flow) running horizontally along the region.
+Sub-figure \ref{fig:gms_enhancement}a demonstrates the agreement raster for the FR stream network as well as it's respective catchment boundary lines symbolized in white and stream network shown in green.
+Inspection of this sub-figure denotes a clear spatial pattern where TP's or areas correctly inundated are pooled alongside catchment boundary lines. 
+On the other side of the catchment boundary, one can witness large swaths of FN's that should be inundated. 
+The FN's also exhibit a spatial pattern as in they tend to collocate within catchments of the pictured mainstems tributaries.
+This sort of behaviour was introduced early in the paper and shown qualitatively in Figure \ref{fig:catchment_boundaries_issue}.
+
+As an enhancement, this paper proposes computing HAND for stream networks comprised of level-paths independently of one another.
+In sub-figure \ref{fig:gms_enhancement}b, the agreement raster for the GMS technique is illustrated as well as the stream network lines in green.
+While the entire mosaiced inundation map from GMS (as described in Section \ref{ssec:inundation_mapping} and Equation \ref{eq:hand_fim_depth}) is used to produce this agreement map, we only show the catchments associated with the mainstem of this region that is shown to follow a clear horizontal path.
+Showing all the catchments for the tributaries that were all derived independently would convulate the image.
+The main message illustrated here is that the catchments associated with the mainstem of this area significantly increase in size.
+Since they are not restricted by the catchments of tributaries that lie in the same drainage areas as those of the mainstem, they extend to include those as well.
+The consequence for inundation extents is a general increase in spatial coverage of the river's water which shows to be in much better agreement with the benchmark map.
+The TP's are no longer bounded by the catchment lines and allowed the expand to their natural extents.
+
+We note here as a contribution of this study that a major inherant, limitation of HAND is the ``nearest drainage'' constraint or the idea that a given river reach only drains or, in HAND's case, inundates its immediate, unique drainage area or catchment.
+In other words, HAND based FIM's are limited to producing fluvial inundation to only their nearest drainage area or catchment.
+However, we know that fluvial inundation can be sourced from several streams nearby that also serve as drainage outlets to the area in question.
+Generally speaking, drainage areas are known to be hiearchical in nature so a given drainage area for a given outlet point can be decomposed into nested drainage areas for outlet points that lie in the original drainage area.
+A perfect example of this are points that lie on tributary reaches closely neighboring a mainstem.
+These points lie in the drainage area of reaches in the mainstem but inundation from the mainstem cannot reach these tributary catchments because of the "nearest" assumption in HAND.
+Hence it's important to state that just like there are different sources of flooding such as fluvial, pluvial, groundwater, reservoir, and coastal, there can also be multiple sources of a riverine flood.
+HAND is only equipped to handle rivirine flooding from the nearest drainage line.
+Other relevant drainage lines that produce fluvial flood waters are not considered here especially if the routing model used doesn't consider backwater effects.
+
+The remaining portion of the improvement in CSI was found to come from a marginal yet noteable reduction in FAR.
+Upon investigation of the spatial results in the agreement maps, we found some areas of slight reductions in FPs especially where changes in catchment boundaries may have occurred due to the reduction in stream order.
+These observations pointed to changes in the rating curves introduced by stream order reduction and catchment definition adjustments.
 Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
@@
 \includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
-\caption{OWP FIM `Cahaba' inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105.
+\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105 at the 500yr recurrence magnitude.
 Catchment boundaries and stream lines are shown in white and dotted green, respectively.

### 2022-08-05 manuscripts/owp_fim4_2022/sections/introduction.tex
* almost done with all of review #1s comments * just need to make method overview figure

@@
 \includegraphics[scale=1.0]{figures/forecast_points.jpg}
-\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
-Also shown are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
+\caption{Forecast points with and without Flood Inundation Maps (FIM) in United States' Advanced Hydrologic Prediction System (AHPS).
+Note that only a small fraction of the AHPS forecast points have existing FIM.
+Also shown are the National Water Model (NWM) stream networks at the full resolution (FR) and Mainstems (MS) resolution.
+The FR network constitutes the entire NWM stream network while the MS resolution network is the FR network at or downstream of the AHPS forecast points shown.
+}
 \label{fig:forecast_points}

### 2022-08-05 manuscripts/owp_fim4_2022/sections/keypoints.tex
* almost done with all of review #1s comments * just need to make method overview figure

--- 2022-02-09 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Keypoints, final entry on title page.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%  List up to three key points (at least one is required)
+%  Key Points summarize the main points and conclusions of the article
+%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
+
+% Example:
+% \begin{keypoints}
+% \item	List up to three key points (at least one is required)
+% \item	Key Points summarize the main points and conclusions of the article
+% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
+% \end{keypoints}
 %
-OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
+\begin{keypoints}
+\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
+\item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
+\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation extents.
+\end{keypoints}
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Software Dependencies and Architecture}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
-Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
-A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub \cite{inundationMapping2022}.
-
-The pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
-Preprocessing acquires and prepares datasets for production of the FIM hydrofabric. 
-The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
-Functionality is included to turn FIM hydrofabric and streamflows into inundation maps represented in both binary and depths.
-Lastly, a test suite includes means of calculating evaluation metrics compared to a variety of preprocessed test case data.
-A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}
-\label{ssec:datasets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP) \cite{esipData2022}.
-The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
-For enforcing levee data into the DEMs, the USACE National Levee Database (NLD) is used to burn feature elevations \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{OpenStreetMap,GreatLakesHydrographyDataset}.
-Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
-These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
-The full input datasets presented by source are listed in Table \ref{tab:data}.
-%
-\begin{table}
-\caption{Data sources, names, and descriptions.}
-\label{tab:data}
-\centering
-\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
-%\begin{tabular}{l c c}
-\hline
-Source & Name & Description \\
-\hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
-\hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
-\hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters \\
-\hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. \\
-\hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. \\
-\hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. \\
-\hline
-USACE & NLD & Levee database of locations and elevations  \\
-\hline
-OSM & Land-Sea Border & Border of land and sea. \\
-\hline
-GLHD & Land-Great Lakes Border & Border of land and Great Lakes. \\
-\hline
-InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
-\hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
-\hline
-\end{tabular}
-\end{table}
-%
-Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and GMS networks and 1,604 HUC8s for the MS method. 
-These methods will be explained more later.
-An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hydro-conditioning}
-\label{ssec:hydro_conditioning}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
-These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
-HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
-In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligble for flooding.
-Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
-It can also be used to simulate more accurate bathymetry which is not accounted for in the 10 m DEM \cite{gesch2002national}.
-
-Specifically within the context of OWP FIM, the hydro-conditioning operations that take place in sequential order are presented. 
-Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing units of size HUC8. 
-The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reach code's membership in the processing unit.
-Hydro-conditioning raster operations take place on buffered boundary definitions to avoid edge contamination and effects \cite{lindsay2013measuring}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Network Enforcement} 
-\label{ssec:stream_network_enforcment}
-%
-The location of the stream network is enforced to ensure general agreement with established stream networks.
-The NHDPlusHR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
-However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-For every NWM headwater segment a headwater point is derived and linearly interpolated to the nearest Burnline Event segment.
-Burnline Event headwater segments are split at the adjusted headwater point to match NWM flowlines.
-The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection and linear interpolation.
-Resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR Burnline Events.}
-\label{fig:stream_density_pruning}
-\end{figure}
-%
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.
-
-The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
-The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
-Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
-Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
-Some of these drawbacks are later addressed by additional conditioning techniques later on.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Levee Enforcement}
-%
-The DEM at 10 m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
-In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10 m DEM.
-The elevations found in the NLD are burned into the DEM if those elevations were found to exceed those already in place.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Depression Filling}
-\label{sssec:depression_filling}
-%
-Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
-The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
-Priority-Flood developed by \citeA{barnes2014priority} is an algorithm for filling said depressions and shown to have improved performance over early works in the field by \citeA{jenson1988extracting} implemented in \citeA{tarboton2005terrain} as well as \citeA{planchon2002fast}.
-The depression filling algorithm used in our pipeline is a Priority-Flood variant developed by \cite{zhou2016efficient} with enhanced single-thread performance and a time complexity of O(n log n) for floating point grids.
-This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
-The depression technique employed here does leave the existence of flat regions where pits existed a prior thus later requiring the need for resolving these flats.
-The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Thalweg Elevation Conditioning}
-\label{sssec:stream_thalweg_elevation_conditioning}
-%
-Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
-In order to prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
-Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50 m.
-The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
-
-The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
-\citeA{garousi2019terrain} proposed an algorithm that traverses stream thalweg pixels in a depth first manner starting with adding all the headwater pixels to a queue. 
-The connectivity of the thalweg pixels is defined by the D-8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
-At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in Equation \ref{eq:thalweg_breach},
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:thalweg_breach}
-\textbf{D}[x] = \min_{y\ drains\ to\ x} {(\ \textbf{D}[x]\ ,\ \textbf{D}[y]\ )}
-\end{equation}
-\end{linenomath*}
-%
-, in which \textbf{D} represents the array of thalweg adjusted elevations indexed by x and y where by y is upstream of x. 
-When a pixel's upstream neighbors are all evaluated, the downstream pixel is added into the queue thus the depth first traversal of the drainage network.
-This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in Critical Success Index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Deriving FIM Hydrofabric}
-\label{ssec:deriving_fim_hydrofabric}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Flow Directions and Flats Resolution}
-\label{ssec:flow_direction_and_flat_resolution}
-%
-To facilitate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D-8 flow directions.
-D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
-The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southeast direction as 8. 
-Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
-
-Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
-Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
-The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
-Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
-By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Deriving FIM Stream Network}
-\label{sssec:deriving_fim_stream_network}
-%
-The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
-The FIM stream network is of similar drainage density as the NWM V2.1 network but fully converges at all junctions leaving no divergences in the network.
-This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
-Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
-Each stream link in this derived FIM stream network is split into equidistant reaches.
-Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments.
-Stream links are then further segmented at NWM lakes and HUC8 boundaries.
-Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretation of the stream network.
-Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherent the synthetic rating curve properties of the upstream or downstream segment, depending on the topology.
-Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
-Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
-This stream network is important since it drives the HAND calculation and derivation of catchments.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Catchments}
-\label{sssec:catchments}
-%
-Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
-Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
-The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
-Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
-The other catchments are derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Height Above Nearest Drainage}
-%
-Once the pixel level catchments are derived, the final relative elevations can be computed.
-Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment described in Section \ref{sssec:catchments}.
-The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
-Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Synthetic Rating Curves}
-\label{sssec:synthetic_rating_curve}
-%
- A method for converting forecast river discharges from the NWM to stages or river depths is necessary for producing FIMs with HAND. 
-For one-dimensional models such as the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a synthetic rating curve at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
-For this application, we utilized the reach averaged approach for developing synthetic rating curves (SRC) \cite{zheng2018river}.
-The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-The reach averaged Manning's formula is derived to be 
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:reach_averaged_mannings_equation}
-\textbf{Q} = \frac{1}{n} \frac{V^{5/3}S^{1/2}}{L B^{2/3}} 
-\end{equation}
-\end{linenomath*}
-%
-where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
-Q, V, and B are taken a specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
-All units are international given the 1 numerator above n.
-The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
-The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the channel length.
-
-Setting of the Manning's n roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
-These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cross-walking with NWM Stream Network}
-\label{sssec:cross_walking_networks}
-%
-The stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with a NWM reach identifier so that a discharge can be converted to stage and later inundation extent.
-For the methods already discussed, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent methods discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
-Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
-While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
-More discussion on this will follow in Section \ref{sec:discussion}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stream Order Reduction}
-\label{ssec:stream_order_reduction}
-%
-FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
-Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
-An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the MS of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
-This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the MS that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the MS catchments.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{False negatives associated with confluence of tributaries with MS. Integers represent flow values from BLE 100 year event for the associated areas. 
-No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
-\label{fig:catchment_boundaries_issue}
-\end{figure}
-%
-We seek to resolve this problem by deriving HAND for processing units with stream networks of reduced stream order. 
-We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
-The resulting FIMs from the overlapping HAND processing units are mosaiced together taking any inundated area to be inundated but more will be explained in Section \ref{ssec:inundation_mapping}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Main-stems}
-\label{sssec:nws_mainstems}
-%
-The Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
-The MS network comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
-It also spans 121,724 reaches across 1,608 HUC8s.
-HAND was originally derived for this stream network to enhance mapping skill along these critical MS segments \cite{djokic2019arc}. 
-The inundation derived from this stream network is mosaiced with the inundation from the FR network to form the MS FIMs. 
-Within each HUC, you'll typically only find a MS stream network of stream order 1 (i.e.
-headwater) but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Generalized Mainstems}
-\label{sssec:generalized_mainstems}
-%
-To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call GMS.
-Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
-Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-Starting at the outlet, a unique level path is propagated upstream. 
-At every confluence, the direction of maximum flow path length is sought to propagate the current level path identifier.
-For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
-Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
-At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
-
-To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
-In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
-For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each level path is buffered to a polygon with a user-available distance parameter of 7 km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
-We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
-Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}c.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates how level paths for the NWM are derived.
-Level paths symbolized by lines of unique colors are propagated upstream following the direction of maximum arbolate sum at each junction.
-}
-\label{fig:level_path_methods}
-\end{figure}
-%
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/gms_methods.jpg}
-\caption{Overall procedure for GMS HAND.
-In (a), we illustrate all NWM stream lines symbolized by their level path.
-Meanwhile (b), demonstrates the DEM clipped to a 7 km buffer around two selected level paths.
-In (c), we show how HAND can be computed just for each one of these two level paths independently. 
-We also show inundation maps created for these two level paths in (c). 
-In (d), we show all the inundation maps for all the level paths mosaiced together. }
-\label{fig:gms_methods}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Inundation Mapping}
-\label{ssec:inundation_mapping}
-%
-The FIM hydrofabric consisting of the relative elevations grid, catchments grid, catchment polygons, rating curve, and cross-walking data are all used to convert forecasts from the NWM into forecasts extents.
-For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
-From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 8 spatial processing units in a rapid, parallel operation.
-The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
-
-Utilizing the stage-discharge relationships in the synthetic rating curves, each forecast for each catchment identifier is assigned a stage value. 
-The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
-We use the basic logic already established in previous works to conduct this \cite{nobre2016hand,liu2016cybergis,maidment2017conceptual}.
-Mathematically, the HAND values, $H_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
-For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
-For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim_depth}
-    D_{ij} = S_i - H_{ij}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim}
-    F_{ij} = D_{ij} > 0
-\end{equation}
-\end{linenomath*}
-%
-For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast in the form of a single raster file.
-For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
-Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
-Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim_depths}
-    D_{ij} = \max_{k=[1,...,K]} D_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-For the MS and GMS methods, the contributing areas are defined differently.
-For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to 2 for that case.
-For GMS, all FIMs from all the level paths in a given area are mosaiced together then K is set to this number of level paths.
-Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim}
-    F_{ij} = \max_{k=[1,...,K]} F_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}
-\label{ssec:evaluation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
-49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM so we used those corresponding discharges and mapping extents for evaluation.
-We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
-Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
-We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
-To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
-This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
-While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
-\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states. This dataset for 1\% and 0.2\% recurrence flows were used as benchmarks.}
-\label{fig:all_ble_maps}
-\end{figure}
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
-\caption{Illustrates Base Level Engineering (BLE) cross sections and stream lines.
-The BLE stream network, which is denser than the NWM V2.1 stream lines, is also shown.
-BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM reach.
-}
-\label{fig:ble_evaluation_method}
-\end{figure}
-%
-The metrics employed in this study to evaluate inundation extents include Critical Success Index (CSI), Probability of Detection (POD), and False Alarm Ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
-To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.
-The two types of errors consist of false positives (FP), or type I errors, which is dry in benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in benchmark by predicted dry.
-Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
-Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
-While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
-Thus, frequency dependent statistics should be used with caution when comparing across sites with varying frequencies. 
-Lastly, approximately 6 HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:csi}
-CSI = \frac{TP}{TP + FN + FP}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:pod}
-POD = \frac{TP}{TP + FN}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:far}
-FAR = \frac{FP}{TP + FP}
-\end{equation}
-\end{linenomath*}
-%

### 2022-08-05 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
* almost done with all of review #1s comments * just need to make method overview figure

@@
 %
-\begin{table}
-\caption{Data sources, names, and descriptions.}
+\begin{sidewaystable}
+\caption{Data sources, names, descriptions, and citations.}
 \label{tab:data}
 \centering
-\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
+\begin{tabular}{|p{1.75cm}|p{4cm}|p{11cm}|p{5cm}|}
 %\begin{tabular}{l c c}
 \hline
-Source & Name & Description \\
+Source & Name & Description & Citations\\
 \hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
+USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement & \cite{nhdplus2022vectors} \\
 \hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
+USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display & \cite{nhdplus2022vectors} \\
 \hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters \\
+USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters & \cite{nhdplus2022dems} \\
 \hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. \\
+USGS & NHDPlusHR WBD & Water Boundaries (WBD) or HUCs used for spatial processing units & \cite{nhdplus2022wbd} \\
 \hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. \\
+NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. & \cite{nwm2022hydrofabric} \\
 \hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. \\
+NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. & \cite{nwm2022hydrofabric} \\
 \hline
-USACE & NLD & Levee database of locations and elevations  \\
+NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. & \cite{nwm2022hydrofabric} \\
 \hline
-OSM & Land-Sea Border & Border of land and sea. \\
+USACE & NLD & Levee database of locations and elevations & \cite{engineers2016national} \\
 \hline
-GLHD & Land-Great Lakes Border & Border of land and Great Lakes. \\
+OSM & Land-Sea Border & Border of land and sea. & \cite{osm2021landsea} \\
 \hline
-InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
+GLHD & Land-Great Lakes Border & Border of land and Great Lakes. & \cite{GreatLakesHydrographyDataset} \\
 \hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
+InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
+\hline
+InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
 \hline
 \end{tabular}
-\end{table}
+%\end{table}
+\end{sidewaystable}
 %
@@
 \label{fig:gms_methods}
+\end{figure}
+%
+%
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.70]{figures/gms_methods_2.jpg}
+\caption{This image, with the same spatial domain as Figure \ref{fig:catchment_boundaries_issue} (HUC 12090301), demonstrates how computing HAND on level path bases leads to larger, independent catchments and more expansive inundation extents (100 yr flows).
+In (a), the catchments and stream network are shown for HAND computed in Full-Resolution (FR) method which shows constrained inundation extents around the two junctions.
+(b) demonstrates the level path associated with this region's highest order river.
+By delineating catchments at this scale indepedent of the neighboring tributaries, the drainage areas are allowed to expand thus allowing inundation extents to cover previously restricted areas.
+In (c) and (d), we show the stream lines, catchments, and inundation extents of the two tributaries.
+Later in Section \ref{ssec:inundation_mapping}, we describe how the inundation in (b), (c), and (d) are mosaiced together to form one seamless inundation map.
+This process allows for multiple, possible contributing sources of fluvial inundation to be considered thus enhancing FIM skill.
+}
+\label{fig:gms_methods_2}
 \end{figure}

### 2022-08-05 manuscripts/owp_fim4_2022/sections/results.tex
minor changes
* almost done with all of review #1s comments * just need to make method overview figure

--- 2022-01-10 manuscripts/owp_fim4_2021/sections/results.tex
@@
 The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
-We computed the statistics (CSI, POD, and FAR) for both 100 and 500 year events for Mannings N set to 0.06 and 0.12. 
-The distribution of these statistics can be examined in Figure \ref{fig:violin_plot} as violin plots.
+We computed the statistics, CSI, POD, and FAR, for both 100 and 500 year events for Mannings N set to 0.06 and 0.12.
+These results are presented in Figure \ref{fig:violin_plot} as violin plots and in Table \ref{tab:aggregate_metrics} as aggregated metrics with the results for MS and GMS presented as percentage changes from their respective FR values.To be more specific, Table \ref{tab:aggregate_metrics} sums the primary metrics, TP, FP, FN, and TN, across all HUC8's then recomputes the secondary metrics which was done to better account for large variances in HUC8 size.
+The same trends discussed below are consistent across both reporting methods (Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}).
+
+The distribution of these flood extent metrics can be examined in Figure \ref{fig:violin_plot} as violin plots.
 Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
-We also denote trend lines for each metric and Manning's n setting as well as their respective slope estimate and one-tailed p-value denoting the level of significance of the trend.
+For example, let's examine the violin plot for the row marked CSI and column for Manning's n = 0.06.
+This sub-figure shows the CSI distributions across all 49 HUC8's when Manning's n is set to 0.06.
+Each independent, whole violin represents the distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100yr and orange for 500yr).
+The horizontal dashed and dotted lines represent the 25th, 50th, and 75th percentiles from bottom to top, respectively.
+Additionally, we show trend lines symbolized in green that for each metric and Manning's n combination denotes the best fit line for the three methods (FR, MS, and GMS).
+To avoid having two trend lines per sub-figure, we elected to aggregate the two magnitudes togther as they tend to observe similar trends across the three models.
+The slope of each trend line is quantified in the Figure by its $\beta_1$ value and the p-value for that statistic which tests the significance of that trend (deviation from a zero sloped trend line).
 
-Aggregating the metrics in the method above treats each HUC8 as it's own unit and does little to consider the size differences of the HUCs. 
-In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics}  the CSI, POD, and FAR recomputed for the entire domain using the sum of all the TPs, FPs, and FNs. 
+% Interpretation of metrics
+Both Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} contain a fair amount of information that is central to the objectives of this study.
+As previously stated in Section \ref{ssec:evaluation}, we consider CSI as a general proxy for the skill of the inundation extents with POD denoting skill on inundated areas and FAR indicating skill on non-inundated areas.
+Again, the main objective of the study is to introduce how computing HAND with disaggregated stream networks to those with unit stream order can enhance the fidelity of FIMs by capturing fluvial inundation from multiple sources as opposed to that of just the nearest drainage line.
+As can be seen in Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, CSI generally increases from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
+This increase is primarily driven by an increase in POD thus generally increasing the probability of correctly detecting inundation.
+Also, we note that FAR is somewhat, albeit marginally, decreased from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
+The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}.
+So overall and in other words, the entire distribution of HUC8's improves across the three methods.
+Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
+We move more details on the relationship betweeen stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).
+
+Additional noteworthy trends in Figure \ref{fig:violin_plot} center around the all-around better performance of FIM's for those of higher Manning's n values and recurrence flows.
+The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value used in the BLE model, 0.11 \cite{fema2021base,fema2021estimated}.
+Most of this improvement is driven by significant increases in POD but, unfortunately, leading to a significant amount of overprediction as observed by the increase in FAR.
+More work can be invested to better regionalize Manning's n values for FIM purposes with HAND.
+We also observe additional trends associated with the magnitude or recurrence interval of the flow rates used with the higher flow rates exhibting better overall CSI, POD, and FAR values than the lower, 100 yr magnitude.
+We introduce in the Discussion (Section \ref{sec:discussion}) that this skill premium exhibited by higher flow events could be due higher quality elevation data in regions that are not described as bathymetric areas.
 %
@@
 \begin{table}[h!]
-\caption{Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated for BLE domain.
-         Best value for across models is highlighted in bold.}
+\caption{ Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated to the BLE domain.
+         The values for MS and GMS methods are expressed in percentage change (\%) from their respective values with the same Manning's N, magnitude, and metric combination in the Full Resolution (FR) method columns.
+         The best value across models is highlighted in bold.}
 \label{tab:aggregate_metrics}
@@
 \hline
-\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & \textbf{0.5796} & \textbf{0.6075} \\
+\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 2.53 & 2.59 & \textbf{3.95} & \textbf{4.04} \\
 \cline{2-8}
-  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & \textbf{0.6182} & \textbf{0.6435} \\
+  & 0.12 & 0.5915 & 0.6149 & 2.35 & 2.26 & \textbf{4.51} & \textbf{4.65} \\
 \hline
-\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & \textbf{0.6633} & \textbf{0.6863} \\
+\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 2.68 & 2.74 & \textbf{4.39} & \textbf{4.38} \\
 \cline{2-8}
-  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & \textbf{0.7606} & \textbf{0.7810} \\
+  & 0.12 & 0.7255 & 0.7446 & 2.83 & 2.71 & \textbf{4.84} & \textbf{4.89} \\
 \hline
-\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1787 & \textbf{0.1589} & \textbf{0.1778} & \textbf{0.1589} \\
+\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & -0.72 & \textbf{-1.24} & \textbf{-1.22} & \textbf{-1.24} \\
 \cline{2-8}
-  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & \textbf{0.2324} & \textbf{0.2148} \\
+  & 0.12 & 0.2379 & 0.2208 & -0.21 & -0.18 & \textbf{-2.31} & \textbf{-2.72} \\
 \hline

### 2022-08-05 manuscripts/owp_fim4_2022/Makefile.sh

@@ -14,7 +14,7 @@ pdflatex "$fullFile"
 
 # expands all tex files into one file
 echo 'Lat-Expand to Create One Tex File ...'
-/usr/bin/latexpand/latexpand --keep-comments --biber bibliography/owp_fim4_2021.bib /foss_fim/manuscripts/owp_fim4_2022/owp_fim4.tex -o owp_fim4_expanded.tex
+/usr/bin/latexpand/latexpand --keep-comments --biber bibliography/owp_fim4_2021.bib owp_fim4.tex -o owp_fim4_expanded.tex
 
 
 

### 2022-08-05 manuscripts/owp_fim4_2022/reviews/first_review/makefile_rebuttal.sh

@@ -1,0 +1,13 @@
+#!/bin/bash -e
+
+#parent_path=$( cd "$(dirname "${BASH_SOURCE[0]}")" ; pwd -P )
+#cd $parent_path
+
+fullFile=rebuttal.tex
+baseName="${fullFile%%.*}"
+extension="${fullFile#*.}"
+
+pdflatex "$fullFile"
+bibtex "$baseName".aux
+pdflatex "$fullFile"
+pdflatex "$fullFile"

### 2022-08-21 manuscripts/owp_fim4_2022/reviews/first_review/rebuttal.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

@@
 \begin{reply}
-We thank the reviewers for carefully reading the manuscript and for providing valuable feedback.
+We thank you, the editor, as well as the reviewers for carefully reading the manuscript and for providing valuable feedback to improve its quality.
 We have gone through the reviews and made every effort to address the concerns while staying within the scope of the paper.
-We feel like we have addressed the concerns around not highlighting the primary contribution of the paper while strengthening the explaination of the contribution's methods in Section 2.5.
+Substantial revision has led to the removal of superfluous report like information while ensuring adequate background not published in other sources is provided in the manuscript.
+This has left some room for better highlighting and explaining the novel contribution of the paper which is countering the ``nearest drainage'' limitation within Height Above Nearest Drainage for flood inundation mapping applications.
+We have added several key clarifications in response to reviewer points and several new or updated figures that better explain our methods and results.
+There are few concerns that we had to leave out of this manuscript as this submission already covers a very substantial amount of work.
+The questions left open concern issues surrounding the synthetic rating curves including localizing Manning's n as well as bathymetry inclusion, estimation, and/or calibration.
+We hope this greatly enhanced submission meets the expectations set by WRR and the concerns of our three reviewers adequately.
 \end{reply}
@@
 % Let's start point-by-point with Reviewer 1
+\clearpage
 \reviewersection
@@
 After reading your comment and our initial discussion section submission, we did notice a missing portion that properly explains how drainage order reduction increases catchment boundary extents and how that drives increased inundation extents and Probability of Detection (POD).
-This, as you correctly noted, is a competing effect to the increased catchment size and associated increase in inundation area and POD.
+This, as you correctly noted, is a competing effect to the increased catchment size and associated decrease in inundation area due to a downward shift in the synthetic rating curve (SRC).
 
@@
 We have inserted some paragraphs that better describe the increase in inundation extents and POD as driven by catchment size enlargement especially around the critical areas described in the paper.
-These larger, overlapping catchments better handle riverine inundation from multiple sources as opposed to the traditional single source "nearest drainage" that HAND is limited to.
+These larger, overlapping catchments better handle riverine inundation from multiple sources as opposed to the traditional single source ``nearest drainage'' that HAND is limited to.
 Limited catchment boundaries drive pools of water that are artificially built up without room for the water to naturally spread.
-This can be conceptually visualized as a column of water that extends beyond its containers top edge.
-This inherent limitation in HAND skews the stage-discharge relationship in addition to limiting the inunduation extents.
+This can be conceptually visualized as a column of water that extends beyond its container's top edge.
+This inherent limitation in HAND skews the stage-discharge relationship in addition to limiting the inundation extents.
 
@@
 How would you say this reduction of false negatives is really in a right direction when you did not compare results with ground truth data?
-    \label{pt:bar}
 \end{point}
@@
 Despite being an engineering scale model, the BLE is available over a very large domain (49 HUC8s and 9 states) at high spatial resolutions which allows for a broader evaluation of HAND which is being applied to continental scale flood forecasting for our purposes.
-Additionally, the BLE offers the availability of tightly spaced cross-sections that provide the stream flows used in making the flood extents.
-This high resolution dataset of stream flows and inundation extents provides the opportunity to remove the meterological, hydrological, and some of the hydraulic assumptions, errors, and limitations from the comparison.
-Stream gages and in-situ observations, while clearly higher fidelity in nature, don't offer the spatial coverage that the BLE offers both in terms of flood extents and their associated stream flows that can be used to detangle confounding factors that introduce errors into the evaluation.
+Additionally, the BLE offers the availability of tightly spaced cross-sections that provide the streamflows used in making the flood extents.
+This high resolution dataset of streamflows and inundation extents provides the opportunity to remove the meteorological, hydrological, and some of the hydraulic assumptions, errors, and limitations from the comparison.
+Stream gages and in-situ observations, while clearly higher fidelity in nature, don't offer the spatial coverage that the BLE offers both in terms of flood extents and their associated streamflows that can be used to detangle confounding factors that introduce errors into the evaluation.
 Nevertheless, FEMA maps have a storied use in the literature as an evaluation dataset due to some of the advantages and trade-offs listed above \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
 Additionally, it is important to note that almost any other source of FIM is subject to errors and limitations including crowd-sourced data, remote sensing observation, and modeled extents.
-Unfortunatley, flood inundation mapping is an imprecise science that depends on compromises to be made in the search for objective benchmarking.
+Unfortunately, flood inundation mapping is an imprecise science that depends on compromises to be made in the search for objective benchmarking.
 
-To address your concern within the paper, we have added some verbage in the evaluation section of the draft that better recognizes and discusses the limitations of the BLE HEC-RAS based FIMs.
-While we can't assure that the BLE maps are perfect, they are considered more sophisticated than our simplified continental technique and have been established well in the literature as a common benchmarking model.
+To address your concern within the paper, we have added some wording in the evaluation section of the draft that better recognizes and discusses the limitations of the BLE HEC-RAS based FIMs.
+While we can't assure that the BLE maps are perfect, they are considered more sophisticated than our simplified, zero-physics, continental-scale technique and have been established well in the literature as a common benchmarking model.
 Hopefully this properly addresses your concern with the use of the BLE HEC-RAS 1D models for evaluation.
@@
 Again, this is the most important part of you work. 
-\label{pt:bar}
 \end{point}
@@
 We found it important to make sure these methods were properly documented within scientific literature for the benefit of the academic and forecasting communities.
-Based on your recommendation, we did go through sections 2.1-4 and tried to remove any information that maybe superflous to the goals and scope of the paper.
+Based on your recommendation, we did go through sections 2.1-4 and tried to remove any information that maybe superfluous to the goals and scope of the paper.
 This includes much of the stream thalweg breaching technique in Section 2.3.4 that was more than adequately described in the citation it was developed in.
@@
 Additionally, we initiated a substantial revamp of Section 2.5 to ensure the main contribution of the paper is properly explained and defended.
-The Section NWM MS (2.5.1) does represent methods conducted by us therefore we have elected to keep it in the paper to adequately describe a stream reduction technique that covers 4\% of the country that is of primary concern to river forecasters.
-Also, we added an extra Figure (Figure \#6 in the revised document) that expands upon the explaination in Figure 5. 
+The Section NWM MS (2.5.1) does represent methods conducted by us in that we computed HAND for that specific stream network.
+Therefore we have elected to keep it in the paper to adequately describe a stream reduction technique that covers 4\% of the country that is of primary concern to river forecasters.
+Also, we added an extra figure (Figure 7 in the revised document) that expands upon the explanation in Figure 5. 
 This provides the reader with a more in-depth look at how drainage order reduction down to the level path scale expands catchment and inundation extents.
@@
 Could you also clarify whether this limitation always results in more overestimation or underestimation? 
-\label{pt:bar}
 \end{point}
@@
 \begin{reply}
-Thank you for pointing out an opportunity to better explain the limitation we are trying to resolve within HAND for FIM applications.
-To better explain this limitation, we included an additional Sub-section (1.4 HAND's Assumptions and Limitations) that better introduces all of the fundametal assumptions HAND employs and their resulting limitations.
+To better explain this limitation, we included an additional Sub-section (1.4 HAND's Assumptions and Limitations) that better introduces all of the fundamental assumptions HAND employs and their resulting limitations.
 Much effort is expelled here to ensure the reader better understands HAND's limitation to only model fluvial inundation from one nearest drainage source thus ignoring other possible fluvial sources of inundation.
-We would also like to point out the reference to Figure 3 and the new Figure 6 which also illustrate this problem and proposed solution.
+We would also like to point out the reference to Figure 3 and the new Figure 7 which also illustrate this problem and proposed solution.
 Please reference the explanation in the Methodology Section 2.5.2 on more discussion on the issue and proposed solution.
@@
 With respect to your question about always producing overestimation (False positives) or underestimation (False negatives), we note in the results and discussion sections that you can't say that this problem always leads to FPs or FNs.
-It can only lead to FN's if a given catchment boundary is restricting water inundating a neighboring catchment.
+It can only lead to FNs if a given catchment boundary is restricting water inundating a neighboring catchment.
 This is commonly seen in areas where junctions are present where a reach's drainage area is restricted by the nearest drainage area of neighboring tributaries.
 However, as the wall of a catchment boundary begins to fill artificially with water, the lack of spread could cause FPs in other regions.
-By reducing the effect of the neigboring boundaries, we see a general reduction in the rating curve thus leading to lower stage values at given discharges.
-We believe that the new version of the Discussion section adequetely details these competing forces as per your prior recommention (\ref{pt:discussion}).
+By reducing the effect of the neighboring boundaries, we see a general reduction in the rating curve thus leading to lower stage values at given discharges.
+We believe that the new version of the Discussion section adequately details these competing forces as per your prior recommendation (\ref{pt:discussion}).
 \end{reply}
@@
 \begin{point}
-Links to all datasets introduced in section 2.2 should be included in your manuscript either in the suggested table or within the "Open Research" section at the end of your manuscript. 
-\label{pt:bar}
+Links to all datasets introduced in section 2.2 should be included in your manuscript either in the suggested table or within the ``Open Research'' section at the end of your manuscript. 
 \end{point}
@@
 Based on your suggestion we have reviewed the citations to all datasets and software used to ensure they follow the journal's requirements.
-In addition, to citing the datasets and software properly in their respective sections, 2.1 and 2.2, we have added another column in the datasets table to ensure they are properly cited there as well.
-The URL's to all the software and datasets are included as part of the reference as indicated by the journal.
+In addition to citing the datasets and software properly in their respective sections, 2.1 and 2.2, we have added another column in the datasets table to ensure they are properly cited there as well.
+The URLs to all the software and datasets are included as part of the references as indicated by the journal.
+We have also added all this information to an Open Research section at the end of the manuscript.
 \end{reply}
@@
 How did you choose these values? 
-\label{pt:bar}
 \end{point}
@@
 It is sometimes hard to follow along. 
-\label{pt:bar}
 \end{point}
@@
 \begin{reply}
-And our reply to it.
+A new Figure 2 was added to the beginning of the Materials and Methods section that gives the reader a high-level overview of the methods employed.
+A brief paragraph was added to help explain this new figure as well at the beginning of the methodology.
 \end{reply}
@@
 Any other studies that observed this that you could cite?
-\label{pt:bar}
 \end{point}
@@
 Doing so allows the HAND value within non-pit filled regions to be computed based on its original value and not its hydro-conditioned value.
+HAND requires a long series of hydro-conditioning operations that can greatly manipulate elevations to ensure drainage.
+We thought it appropriate to use native elevations for this one step to minimize the error associated with those assumptions.
 We inherited this technique from Djokic, 2019 which we now cite on that line in the paper.
+The explaination as to why we elected to use the native elevations was expanded upon in this section as well.
 \end{reply}
@@
 You mention that the level path method starts from an outlet, so it is probably better to show an outlet in Figure 4.
-Please clarify what you mean by "arbolate sum"? I cannot understand what this summation is.
-\label{pt:bar}
+Please clarify what you mean by ``arbolate sum''? I cannot understand what this summation is.
 \end{point}
@@
 \begin{reply}
-The full explanation for this figure is provided by its in-text reference but based on your recommendation we have expanded this caption to properly describe the Figure in place.
+The full explanation for this figure is provided by its in-text reference but based on your recommendation we have expanded this caption to properly describe the figure in place.
 \end{reply}
@@
 Please provide your thoughts on this and add it to the manuscript. 
-\label{pt:bar}
 \end{point}
@@
 Does that refer to flood depth or flood extent?
-\label{pt:bar}
 \end{point}
@@
 You may want to revise this sentence and use ``physically-based'' rather than ``realistic''.
-\label{pt:bar}
 \end{point}
@@
 You might want to start this sentence like this: The NWM provides stream forecasts at these geofabric segments using the Muskingham-Cunge method. 
-\label{pt:bar}
 \end{point}
@@
 \begin{point}
-Page 29: It could be better to change the title of subsection 3.1 to "Flood mapping performance". 
-\label{pt:bar}
+Page 29: It could be better to change the title of subsection 3.1 to ``Flood mapping performance''. 
 \end{point}
@@
 \begin{point}
-L671. Please add "(Table 2)" at the end of the sentence.
+L671. Please add ``(Table 2)'' at the end of the sentence.
 This guides the reader to know what you are referring to immediately. 
-\label{pt:bar}
 \end{point}
@@
 \begin{reply}
-Due to restructuring of the Results section, this reference was changed to refer to the Figure with violin plots instead in the following sentence.
+Due to restructuring of the Results section, this reference was changed to refer to the figure with violin plots instead in the following sentence.
 \end{reply}
@@
 Does that refer to the NHDPlusHR or NHDPlus medium resolution as used in the NWM v2.1?
-\label{pt:bar}
 \end{point}
@@
 Line 344-345: This statement requires citation(s).
-\label{pt:bar}
 \end{point}
@@
 I would suggest removing lines 534-535. 
-\label{pt:bar}
 \end{point}
@@
 The hydro conditioning of the DEM section should start with levee enforcement (section 2.3.2).
-\label{pt:bar}
 \end{point}
@@
 Section 2.3.1 is about enforcing the general location and density of the stream network used for HAND.
-Additionally, an excavation procedure is detailed that is used to counter flat bathymetric regions in the DEM that can severly distort synthetic rating curves (SRC) and catchment delineations.
+Additionally, an excavation procedure is detailed that is used to counter flat bathymetric regions in the DEM that can severely distort synthetic rating curves (SRC) and catchment delineations.
 We use the term ``hydro-conditioning'', as stated in Section 2.3, as the series of steps required to generate monotonically decreasing elevations, flow directions that fully drain processing regions, and monotonically draining stream network thalweg lines.
+Additionally, we enforce the general locations of hydrography-based stream lines as well as levee elvations which we also consider hydro-conditioning.
 These are fundamental assumptions and inputs required for HAND, catchment, and SRC calculation.
@@
 What is your recommendation or suggestion?
-\label{pt:bar}
 \end{point}
@@
 Maybe reporting the percentage of improvements in Table 2 would help better understanding the contribution of your work.
-\label{pt:bar}
 \end{point}
@@
 You discuss this figure on Line 686-690, but I think it could be better to bring this part further up where you first mention Figure 8.
-\label{pt:bar}
 \end{point}
@@
 The first paragraph briefly introduces the results while the second paragraph gives the reader and in-depth description of the previously numbered Figure 8.
-We have made sure all elements of the Figure are properly described and that a specific sub-figured is dived into more detail to ensure the reader has a proper understanding of the many elements at play here.
+We have made sure all elements of the figure are properly described and that a specific sub-figured is dived into more detail to ensure the reader has a proper understanding of the many elements at play here.
 In addition to this enhanced description, we have added a significant amount of information in the third paragraph of the results section that begins to dissect the results and the trends observed in the study.
-In a related point and in response to your critique of the Discussion section, we have made significant enhancements to the Discussion section that better elaborate on the material and trends presented in this Figure.
-We hope that these changes fully address your questions and support this Figure that is central to the paper's findings.
+In a related point and in response to your critique of the Discussion section, we have made significant enhancements to the Discussion section that better elaborate on the material and trends presented in this figure.
+We hope that these changes fully address your questions and support this figure that is central to the paper's findings.
 \end{reply}
@@
 Fig 1. The NWM FR streams line is hardly visible in the legend. Please revise. 
-\label{pt:bar}
 \end{point}
@@
 Line 575. This should be Figure 5d not 5c.
-\label{pt:bar}
 \end{point}
@@
 \begin{reply}
-These names of the rivers are not present in the NWM V2.1 but instead I had to look at their equivalents in the NHDPlusHR stream line data.
-While the figure is quite busy, I agree that showing the NWM catchments does provide the reader some sense as to how many BLE cross sections might be associated with a given NWM forecast identifier.
+These names of the rivers are not present in the NWM V2.1 but instead we had to look at their equivalents in the NHDPlusHR stream line data.
+While the figure is quite busy, we agree that showing the NWM catchments does provide the reader some sense as to how many BLE cross sections might be associated with a given NWM forecast identifier.
 We have updated the caption to include this new layer and provide some more spatial context.
@@
 \begin{reply}
-We used the 500yr magnitude and added this to both the caption and in-text content as well.
+We used the 500 yr magnitude and added this to both the caption and in-text content as well.
 \end{reply}
@@
 L66. Remove the period before the parenthesis, followed by the citations.
-\label{pt:bar}
 \end{point}
@@
 L101. NWM is already defined. Just use the abbreviation here. 
-\label{pt:bar}
 \end{point}
@@
 Title of section 1.5: It could be better to use the Operational Water Prediction (OWP) Flood Inundation Maps (FIM)
-\label{pt:bar}
 \end{point}
@@
 L160. HPC is already defined. 
-\label{pt:bar}
 \end{point}
@@
 L167. MR is already defined.
-\label{pt:bar}
 \end{point}
@@
 L178. USGS is not yet defined.
-\label{pt:bar}
 \end{point}
@@
 Line 347: NLD has already been introduced on page 11.
-\label{pt:bar}
 \end{point}
@@
 Line 524: MS should be first defined here not on line 537.
-\label{pt:bar}
 \end{point}
@@
 Use either Manning's n or Manning's N everywhere.
-\label{pt:bar}
 \end{point}
@@
 % Begin a new reviewer section
+\clearpage
 \reviewersection
@@
 \begin{reply}
+Thank you for taking the time to read our paper and providing ample feedback for improving its content.
 We understand your concern and have made every effort to highlight the novel contributions of the manuscript which is an operational, continental scale version of HAND that applies the latest datasets and hydro-conditioning methods while considering multiple sources of fluvial inundation instead of just that of the nearest drainage line.
@@
 \begin{reply}
+To your point, we have added an additional, high-level explaination of the two stream networks for the NWM in this section to provide the reader with a general idea of what these two illustrated networks mean.
+We did leave some additional detail to be discussed in the NWM section since it is more relevant to that section.
 \end{reply}
@@
 \begin{reply}
+Added a sentence specifically mentioning that a continental scale hydrodynamic model is computationally prohibitive.
 \end{reply}
@@
 Presenting some research that is a byproduct of tool development won't interest the potential readers of this journal.
+\label{pt:science_oriented_purpose}
 \end{point}
@@
 \begin{reply}
+Based on your suggestion, we have rewritten this section to better highlight OWP FIM as a research study and a scientific enhancement.
 \end{reply}
@@
 \begin{reply}
+Several sentences in the introduction of AHPS and NWM were removed to only include the most relevant information to this study.
+The section introducing OWP FIM was completely revised to address \ref{pt:science_oriented_purpose} and an entire paragraph on some of the superflous HAND-related studies that are mentioned in the Methods later on was also removed.
 \end{reply}
@@
 \begin{reply}
+The first paragraph in original section 2.3.1 was meant to describe the pruning of the NHDPlusHR to better match the resolution of the NWM stream network which was originally derived from the NHDPlus network.
+In response to your questions, we have revamped this paragraph to offer the reader a better explanation.
+Additionally, we have added more information to the caption of original Figure 2 to provide a better explaination in context.
+We hope this level of detail adequately answers your questions.
 \end{reply}
@@
 \begin{point}
-Line 513 I don't think "work for many instances fine" is formal enough as a sentence in a scientific journal article.
+Line 513 I don't think ``work for many instances fine'' is formal enough as a sentence in a scientific journal article.
 \end{point}
@@
 \begin{reply}
+This informal phrase was removed and the sentence was merged with the following sentence.
 \end{reply}
@@
 \begin{reply}
+Yes we do understand that the methods are quite long but these methods are very much central to the study.
+We believe that all methods involved in developing the conclusion of the study should be properly explained or cited.
+The methods in the four sub-sections you referenced involve a listing of the primary software dependencies used, description of the datasets used, explanation of the hydro-conditioning processes employed, and a detailing of the derivation of the FIM hydrofabric.
+All of these methods are central to the conclusions of the study and must be properly explained here since they have not been published previously.
+We argue that keeping these sub-sections intact are required to ensure reproducible, transparent, and open research.
 \end{reply}
@@
 \begin{point}
-Line 544 "only find an MS stream network of stream order 1 (i.e., headwater)" From Figure 1, we can tell that most AHPS sites are located on big rivers, which means their stream orders are pretty low.
+Line 544 ``only find an MS stream network of stream order 1 (i.e., headwater)'' From Figure 1, we can tell that most AHPS sites are located on big rivers, which means their stream orders are pretty low.
 Then why do you say they are headwater links?
@@
 \begin{reply}
+We agree this sentence is fairly ambigous so we have clarified this to be more clear.
+Yes, typically MS river segments are large rivers with higher Horton-Strahler stream order.
+Horton-Strahler stream order starts at order 1 for headwaters and compounds downstream until the outlet point.
+However, HAND is computed at the HUC8 spatial resolution for the MS version.
+When a mainstem river is clipped to a HUC8 segment, it typically will have an effective stream order of 1 within this particular area.
+This is relevant to the computation of HAND as in it removes tributaries that may limit catchment boundary sizes and resulting inundation extents.
+We have changed the phrasing to reflect ``uniform stream order'' instead which we believe better describes the circumstances.
 \end{reply}
@@
 \begin{reply}
+Based on your appropriate feedback, we have elected to show the same figure but in the context of a HUC12.
+Illustrating a HUC12 shows the dendritic structure relevant to the explaination of arbolate sums and level paths (LP).
+To be clear, the arbolate sums and LPs were derived at the HUC8 scale but for the figure we only show a HUC12 to show fine grain features more clearly.
+The numbers in the figure represent arbolate sums which are cumulative drainage distances of all inclusive upstream reaches.
+The arbolate sums are now much more clearly labeled at the whole km value for each reach defined as a junction to junction stream segment.
 \end{reply}
@@
 \begin{reply}
+We understand that a few metric points may seem like a marginal increase.
+In order to better highlight the value of the proposed work, we have changed Table 2 to report the percentage improvement for the proposed models, MS and GMS, when compared to the baseline model, FR.
+Each percentage improvement corresponds to a specific metric (CSI, POD, and FAR), magnitude (100 and 500 yr), and Manning's n values (0.06 and 0.12).
+These percentage improvements for CSI range from just over 2\% for MS to almost 5\% improvement for GMS which we deem convincing.
+Additionally, we argue and highlight that these metrics are for evaluations at the HUC8 scale.
+HUC8s are very large regions that aggregate away fine grain results and impacts.
+This up to 5\% improvement assessed at a much local scale can seem like a significant improvement.
+While the OWP is not currently using an impact based data source such as building footprints or land cover/use maps, viewing the enhancements qualitatively and computing the results at this scale could provide an argument that these results can be significant at the impact scale.
+We provided original Figure 10 to highlight the enhancement of this improvement at this scale.
+To respond to this point, we have inserted some verbage in the Discussion that acknowledges the limitations of large, HUC8, scale evaluations and hints towards future usage of impact datasets and finer grain evaluations to counter this limitation.
+For now, we note an absolute improvement in FIM skill by account for multiple fluvial inundation sources by reducing the scale of HAND computation to utilize stream networks of unit Horton-Strahler stream order.
 \end{reply}
@@
 This manuscript adds to the growing body of literature using HAND as a method for determining potential flood inundation in the absence of a full-physics (hydraulic) model.
-Ultimately, the process of producing inundation estimates for >2.7M stream segments at high spatial resolution (~10m) is one which required much iteration.
+Ultimately, the process of producing inundation estimates for $>$2.7M stream segments at high spatial resolution ($\sim$10m) is one which required much iteration.
 In the end, the authors found that generalizing the stream networks, and using smaller basin areas (HUC8) combined with aggregating individual level-paths yielded better estimates of flood inundation when compared to results from HEC-RAS BLE maps.
@@
 Another innovation is in the mosaicking of the FIMs in the presence of tributaries and lower-order streams.
-Another very useful innovation lies in 'burning' levee elevations onto the DEMs to enforce flood control measures on to the HAND surface.
+Another very useful innovation lies in ``burning'' levee elevations onto the DEMs to enforce flood control measures on to the HAND surface.
 The approach detailed in this manuscript can be seen as a guide for mitigating the 'catchment boundary problem', the importance of which should be emphasized more extensively in this paper.
+\label{pt:intro_pt_reviewer_3}
+\end{point}
 
+\begin{reply}
+We thank you for your overall assessment of the paper and for noting the important contributions this work represents.
+While the central scientific contribution of the paper was allowing for multiple fluvial source inundation within HAND based FIMs, we are pleased to hear that the many of the supporting contributions including continental scaling, levee enforcement, open-source software availability, forecasting applications, and thalweg delineation methods were properly conveyed and appreciated.
+We hope that the following points in your review are properly addressed to ensure that the main contribution of the paper is properly communicated in a clear and scientific manner to add to the body of literature in this sub-field of hydrology.
+\end{reply}
+
+\begin{point}
 As a reader, I find the manuscript title and some of the conclusions to be somewhat misdirected.
 The idea that Horton-Strahler stream order has a direct impact on the resulting HAND inundation product is misleading.
-The stream order of a given reach is determined by the drainage density, and as such is a product of 'upstream' decisions made during the derivation of the flow network.
+The stream order of a given reach is determined by the drainage density, and as such is a product of ``upstream'' decisions made during the derivation of the flow network.
 Stream order is not an input into the HAND model, nor is it very meaningful when comparing networks that are derived from different headwater locations (such as a pruned version of an existing network such as NHDPlusHR).
@@
 There are references in the paper to the impact of drainage density on maximum stream order, but the Conclusions indicate that Horton-Strahler reduction is the reason for improved FIM results, when that reduction is simply an associated result of reduced stream network density.
-For example, the authors indicate "We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed...".
+For example, the authors indicate ``We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed...''.
 In reality, the reduction Horton-Strahler stream order is the result of, and not the cause for, a reduction in drainage density.
 It may be useful to emphasize that the discretization of the network into coherent regions based on level-path or the regionalization methods (such as using HUC8 boundaries), as well as the priority queue calculation (Equation 6) during the HAND raster mosaic process has a bigger impact on the result than the Horton-Strahler number of the network employed.
+\end{point}
 
+\begin{reply}
+We acknowledge that the original messaging in the paper maybe confusing.
+We note in this reply that as described in the methodology surrounding datasets and stream network enforcement as well as your mention \ref{pt:intro_pt_reviewer_3}, stream network is indeed used as an input in our process to compute HAND.
+We differ from previous studies mentioned in the paper that utilize accumulated drainage pixels across the entire processing area then threshold those accumulations to produce a stream thalweg network.
+In our methods, we consider a forecast network, that of the NWM, and a network that better agrees with the DEM employed, NHDPlusHR Burnlines, to help enforce drainage in the processing unit which is one of the fundamental assumptions and requirements of HAND.
+We then use select headwater points to seed a new stream thalweg network that is heavily influenced by the burning and conditioning operations inspired by the input networks, NWM and NHDPlusHR.
+The thalweg datum used to detrend or normalize elevations away from mean sea level to that of nearest drainage is thus a product of input stream networks.
+
+Where we believe our explaination lacked, and with respect to your point, is we failed to properly communicate that discretizing the Full Resolution NWM stream network into either Mainstems (MS) network or independent level paths doesn't change the maximum Horton-Strahler stream order or the drainage density.
+It does, however, change the effective maximum stream order used in each independent computation of HAND.
+This distinction is not properly conveyed in the paper and could lead readers to be confused as to what is actually happening.
+In response to this ambigiouty, we have added language in the Stream Order Reduction sub-section of the Methods to ensure this distincition is more clear.
+
+Most notably, we propose a title change to the paper to ensure that the primary objective of considering multi-source fluvial inundation is properly highlighted.
+The new proposed title ``Extending Height Above Nearest Drainage to Model Multiple Fluvial Sources in Flood Inundation Mapping Applications for the U.S.
+National Water Model'' better accounts for the primary purpose of the paper while not allowing an ambigouty to settle in without proper explanation to occur first.
+We hope these changes better account for the fundamental contributions of the paper and convey them in a much more clear manner.
+\end{reply}
+
+\begin{point}
+Another limitation in the methodology lies in using synthetic rating curves (SRCs) that result from reach-averaged geometry parameter derived from the AGREE-DEM (a hydro-conditioning process).
+This results in an artificially impacted cross-section and bathymetry structure.
+The parameters used to constrain the AGREE method will have a profound impact on the SRCs derived from that DEM due to the excavation process in the AGREE method.
+\end{point}
+
+\begin{reply}
+We agree with you that the AGREE DEM proceduce does introduce a significant amount of bathymetric manipulation that affects SRC and FIM extent quality.
+To your point, we have noted this limitation in multiple sections throughout the paper including the methods, discussion, and conclusions to ensure that the reader is fully aware of this limitation and its effects.
+We do motivate future work in this realm as SRC quality is of primary importance to FIM skill.
+We invite future researchers to investigate factors that contribute to SRC quality including Manning's n localization as well as bathymetry assimiliation, estimation, and calibration.
+Unfortunately, the scope of this article is already too extensive to go much further.
+\end{reply}
+
+\begin{point}
 Further revision could include a more thorough description of how using level-paths eliminates the catchment boundary problem, when each level path may have many tributaries (and thus multiple catchment boundaries).
@@
 It seems that the way around the catchment boundary problem is to calculate HAND for a given level-path while leaving tributaries out of the analysis, and allowing the mosaic process to result in a HAND/FIM result with more extensive coverage over catchment boundary areas.
+\end{point}
 
-Another limitation in the methodology lies in using synthetic rating curves (SRCs) that result from reach-averaged geometry parameter derived from the AGREE-DEM (a hydro-conditioning process).
-This results in an artificially impacted cross-section and bathymetry structure.
-The parameters used to constrain the AGREE method will have a profound impact on the SRCs derived from that DEM due to the excavation process in the AGREE method.
+\begin{reply}
+We agree that the original submission lacked sufficient detail as to how the catchment boundary problem is mitigated by level path disaggregation for HAND computation.
+We have completely revamped this section to ensure the core contribution of this paper is properly explained.
+Additionally, we have introduced a new Figure 7 that better describes how the catchments are delineated at the level path scale as well as the associated inundation.
+We hope these considerable changes better communicate this critical section of the paper.
+\end{reply}
 
+\begin{point}
 The manuscript is, in places, difficult to understand due in part to the large number of acronyms and persistent jargon that pervades it.
@@
 \begin{reply}
-
+We have gone through your attached annotated manuscript and have made every effort to address every concern listed.
+For any annotations that require an additional response, we provide those below referencing the original line numbers those annotations occur. 
 \end{reply}
@@
 \begin{point}
-
+L98 - Does the AHPS have a coarser resolution than NWM?
 \end{point}
@@
 \begin{reply}
-
+AHPS only has two forecast horizons while NWM has three forecast horizons and one analysis product (lookback).  
+These are now accounted for in the new manuscript.
 \end{reply}
@@
 \begin{point}
-
+L368 - 375:
+Why does this need to be done?
+Is there often a conflict after running AGREE between the DEM thalweg and the AGREE-DEM thalweg?
+Depending on the smooth-drop parameter used in AGREE, I wouldn't think there would be any parallel channels in the resulting DEM.
 \end{point}
@@
 \begin{reply}
-
+The citations in the manuscript refer to the parallel streams problem that arises by using AGREE DEM.
+The AGREE DEM enforces a hydrography based streamline and associated notch while not considering some semblemce of the thalweg that might be occuring natively in the DEM.
+This creates the notion of a parallel stream when comparing two different DEMs, the AGREE one and the native, unconditioned one.
+Additionally, this step provides native elevations to be fed into the thalweg breaching algorithm as opposed to the vastly artificial values in the sharp drop portion of AGREE.
+To your point, we have attempted to better explain this in the manuscript.
 \end{reply}
@@
 \begin{point}
-
+L454: Couldn't you just use the value from the flow accumulation grid at each thalweg pixel rather than derive catchments at each thalweg pixel?
 \end{point}
@@
 \begin{reply}
-
+The term ``contributing area'' is ambigious here.
+We seek to clarify that we are not computing the contributing area in some square km value.
+We are just trying the derive catchments or unique drainage polygons for two separate circumstances.
+We have changed the language here to bring about more clarity.
 \end{reply}
@@
 \begin{point}
-
+L478: perimeter?
 \end{point}
@@
 \begin{reply}
-
+In the reach averaged version of the Manning's Equation, wetted perimeter in the traditional Manning's Equation is translated to bed area given the consideration to the reach length, L, in the formula as well.
 \end{reply}
@@
 \begin{point}
-
+L491: 0.05 does not appear to be used later. How did the authors decide to use 0.06 and 0.12?
 \end{point}
@@
 \begin{reply}
-
+In our evaluations, we have noted better general performance with higher Manning's n values thus selected the value of 0.06 from the Djokic work as our baseline.
 \end{reply}
@@
 \begin{point}
-
+L498: NWM channel parameters, for consistency with the source model? These parameters exhibit variability and have been evaluated during mdoel testing.
 \end{point}
@@
 \begin{reply}
-
+We tried those parameters and unfortunately they do not yield better results in terms of FIM.
+It's possible that those regionalized values are not optimal for FIM purposes as they are tuned more for 1D hydralic routing purposes.
+It's likely that lack of accurate bathymetric information interacts heavily with the optimal Manning's n value which is only supposed to account for roughness.
+We are keen on working more with the combined effects of bathymetry and roughness parameters to optimize the SRCs for better FIM performance.
+In the meantime, we focused on the core contributions of this paper already presented.
 \end{reply}
@@
 \begin{point}
-
+L684:  This is a key finding of this research and could be emphasized more perhaps in the abstract or conclusions.
 \end{point}
@@
 \begin{reply}
-
+This key finding has now been further elaborated on in the conclusions and abstract sections.
 \end{reply}
@@
 \begin{point}
-
+L765-767: I thought lower stream order produced better FIM maps?
 \end{point}
@@
 \begin{reply}
-
+This sentence is referring to an issue associated with conflation or cross-walking of the FR version of HAND.
+This has now been specified so the reader understands that it is likely improved with the new GMS version that ignores neighboring tributaries thus less susceptible to cross-walking errors.
 \end{reply}
-
-\begin{point}
-
-\end{point}
-
-\begin{reply}
-
-\end{reply}
-
-\begin{point}
-
-\end{point}
-
-\begin{reply}
-
-\end{reply}
-
-\begin{point}
-
-\end{point}
-
-\begin{reply}
-
-\end{reply}
-
-\begin{point}
-
-\end{point}
-
-\begin{reply}
-
-\end{reply}
-
-\begin{point}
-
-\end{point}
-
-\begin{reply}
-
-\end{reply}
-
-\begin{point}
-
-\end{point}
-
-\begin{reply}
-
-\end{reply}
-
-\begin{point}
-
-\end{point}
-
-\begin{reply}
-
-\end{reply}
-
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-08-21 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

@@
 \usepackage{amsmath}
+\usepackage{float}
 
@@
 %
-\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
+%\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
+\title{Extending Height Above Nearest Drainage to Model Multiple Fluvial Sources in Flood Inundation Mapping Applications for the U.S. National Water Model}
 %
@@
 \begin{keypoints}
-\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
+\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to a ``nearest drainage'' limitation most evident at river junctions.
 \item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
@@
 LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
-The FIMs associate from each independent LP are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
+The FIMs associated with each independent LP are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
 Improvement in mapping skill is observed when the extents are compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
 A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
+We observe that the improvement from extending drainage order reduction techniques from 4-5\% of the entire stream network to 100\% of the network is about the same magnitude improvement as going from no drainage order reduction to 4-5\% of the network.
+Thus, we conclude that drainage order reduction techniques circumvent the nearest drainage limitation while offering diminishing returns to skill improvement and penalizing with additional computational expenses.
 This novel contribution is proposed and studied in a new open-source implementation that utilizes the latest combination of hydro-conditioning techniques to enforce drainage and counter limitations in the elevation data.
-This framework scales to the entire NWN modeling domain including Hawaii and Puerto Rico to help forecasters provide timely and accurate flood warnings.
+This framework scales to the entire NWM modeling domain including Hawaii and Puerto Rico to help forecasters provide timely and accurate flood warnings.
 \end{abstract}
@@
 In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
+This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance because only fluvial inundation, or inundation sourced from rivers, is only considered from the nearest drainage line.
+As rivers flood, multiple neighboring rivers could contribute to a regions inundation and HAND has no means of modeling for that.
 We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance.
+This is done by splitting the stream network into continuous rivers known as level paths.
+These level paths have no tributaries thus are known to be stream lines with a unit stream order indicating no branching.
+HAND is computed independently for each level path and the resulting FIMs are mosaiced together to form one seamless map.
+We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance to counter one of HAND's fundamental limitations.
 %
@@
 More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
-With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
+With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}, with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf Coast.
 Trends related to flood damages and fatalities have been steadily increasing over recent decades \cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
-Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
-Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.
+Some are expecting that the hydrologic cycle will intensify due to climate change which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
+Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest and Great Lakes regions while decreasing in the coastal Southeast, Southwest, and California.
 %
@@
 %
-Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
+Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life threatening and property damaging events. 
 One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble stream flows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
-Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
-AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
-On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
-Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons to address the increasingly complex water challenges facing the US.
-%
-\begin{figure}[h!]
+AHPS provides forecasting services in the form of ensemble streamflows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
+Additionally, two forecasting networks, Full Resolution (FR) and Mainstems (MS) stream networks, relevant to the National Water Model (NWM) (see Section \ref{ssec:national_water_model}) are rendered in Figure \ref{fig:forecast_points}.
+The FR network refers to the entire NWM forecasting domain while MS refers to the subset of the FR network that is at or downstream of AHPS forecasting points (see Section \ref{ssec:national_water_model}).
+On an approximate basis, there is only one forecast point every 1,450 km of river (FR) and one forecast point with FIM every 29,000 km of river (FR).
+Despite the AHPS advances in operational flood forecasting, it lacks sufficient domain coverage, spatial resolution, and long-range forecast horizons to address the increasingly complex water challenges facing the US.
+%
+\begin{figure}[H]
 \centering
@@
 Note that only a small fraction of the AHPS forecast points have existing FIM.
-Also shown are the National Water Model (NWM) stream networks at the full resolution (FR) and Mainstems (MS) resolution.
+Also shown are the National Water Model (NWM) stream networks at the Full Resolution (FR) and Mainstems (MS) resolution.
 The FR network constitutes the entire NWM stream network while the MS resolution network is the FR network at or downstream of the AHPS forecast points shown.
@@
 %
-Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the NWM which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
-The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
+Additional work is required to address the gaps that the AHPS leaves in terms of spatial resolution, spatial coverage, and temporal forecast horizons.
+In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners at the National Center for Atmospheric Research (NCAR) have developed and implemented operationally the NWM which is a configuration of the Weather Research and Forecasting Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including lookback-range (3-28 hrs), short-range (18 hr), medium-range (10 day) and long-range (30 day) forecast horizons.
 The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
-As a complement to the operational NWM, OWP has also developed a configuration of the NWM that extends RFC forecasts downstream by assimilating and routing forecast streamflows to the next downstream AHPS forecast point.
-This configuration of the NWM is used to enhance forecasting skill by leveraging best available regional-scale data.
-The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
-Figure \ref{fig:forecast_points} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
+As a complement to the operational NWM, RFC forecasts from AHPS forecast points are assimilated in the NWM and routed downstream to the next downstream AHPS forecast point where the process iterates again.
+This assimilation into NWM is used to enhance forecasting skill by leveraging best available regional-scale forecasts.
+The river network upon which this special assimilation technique operates on is herein referred to as the Mainstem (MS) stream network.
+Figure \ref{fig:forecast_points} shows the NWM V2.1 FR stream network as well as the NWM V2.1 MS network.
 The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
 
-The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
-The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The NWM provides stream forecasts at these hydrofabric segments using the Muskingham-Cunge method to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
-Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
-As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
-The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
+The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the ``hydrofabric'' in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus,nhdplus2022vectors}.
+The term ``hydrofabric'' is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
+The NWM provides stream forecasts at these hydrofabric segments using the Muskingum-Cunge method to reduce computational requirements of a continental scale model but fails to consider backwater dynamics \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
+The need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
+The use of a 2-dimensional (2D) hydrodynamic model across a continental-scale and high spatial resolutions is very cost prohibitive especially in an operational setting.
+The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert 1-dimensional (1D) riverine discharges to stages, and finally to inundation extents and depths.
 %
@@
 %
-HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
+HAND normalizes topography along the nearest drainage path and it has been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
 \citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
-The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
-\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
+However, a core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
+The terrain index also provides additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
+\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as SRCs by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
 This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
@@
 %
-HAND operates on many underlying assumptions since it can only be used as an inundation proxy and not a true hydrodynmic inundation model.
-HAND, to our knowledge, has only been applied for riverine inundation applications thus it is also missing pluvial, coastal, ground water and dam break components among other possible sources of flooding.
-Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest stream line which is used for catchment allocation and relative elevation calculation.
-Since elevations don't naturally do this, they must undergo a long series of hydro-conditioning processess to enforce monotonically decreasing elevations across an entire processing unit along with hydrologically correct directions of flow.
+HAND operates on many underlying assumptions since it can only be used as an inundation proxy or no physics model and thus, not a true hydrodynamic inundation model \cite{nobre2016hand,liu2016cybergis,liu2020height}.
+HAND, to our knowledge, has only been applied to natural, inland, and riverine inundation applications thus it is also missing pluvial, coastal, ground water, and dam break components among other possible sources of flooding.
+Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest stream line which is used for catchment allocation and relative elevation calculation \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,maidment2017conceptual,garousi2019terrain,zheng2018river,zheng2018geoflood,johnson2019integrated,renno2008hand}.
+Stream thalweg networks must also collectively drain to a singular outlet point for a given processing region \cite{nobre2016hand,zheng2018geoflood,renno2008hand}.
+Since elevations don't naturally do this, they must undergo a long series of hydro-conditioning processes to enforce monotonically decreasing elevations across an entire processing unit along with hydrologically correct directions of flow \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,donchyts2016global,renno2008hand}.
 The level of DEM manipulation required to enforce this assumption can be substantial depending on the region and can be a significant source of error.
-The drainage enforcing assumption also interacts with an inabilty to properly account for fluvial inundation in regions of DEM depressions.
-
-True to its name, when HAND, where the D denotes drainage, is used for FIM applications it assumes that a given area's drainage area is the only reach it eligble to source inundation from.
-Catchments are independent of one another for FIM purposes meaning a reaches' stage value is only used to threshold the HAND values within its respective catchment.
-This assumption plays to the ``nearest drainage'' term in HAND and creates a significant limitation within HAND for FIM applications \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-At the junction of high stream order and high flow rivers with lower flow tributaries, there can a lack of inundation extents exhibited which is known colloaquially in the forecasting community as the catchment boundary problem.
-The academic community has somewhat referenced this issue before but it has been characterized more as a problem to due with the stream delineation process that comes from thresholding the drainage accumulation maps.
-Later in this study, we will re-introduce this problem and demonstrate how we start with a starting stream network, that of the NWM's, and thus avoid having to threshold accumulations to some arbitrary value to define stream networks.
-We illustrate how computing HAND independently for stream lines of unit stream order can significantly enhance FIM performance by accounting for multiple sources of fluvial inundation that may exist in certain regions and flow scenerios.
+The drainage enforcing assumption also interacts with an inability to properly account for fluvial inundation in regions of DEM depressions that lack natural drainage to riverine areas \cite{nobre2016hand,renno2008hand}.
+
+When used for FIM applications, HAND assumes only fluvial inundation sourced from its nearest drainage line is accounted for \cite{nobre2016hand,mcgehee2016modified}.
+Catchments are independent of one another for FIM purposes meaning a reaches' stage value is only used to threshold the HAND values within its respective catchment \cite{liu2016cybergis,zheng2018river,zheng2018geoflood}.
+This assumption plays to the ``Nearest Drainage'' term in HAND and creates a significant limitation within HAND for FIM applications \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+At the junction of high stream order and high flow rivers with lower flow tributaries, there can be a lack of inundation extents exhibited which is known colloquially in the forecasting community as the ``catchment boundary problem''.
+The academic community has somewhat referenced this issue before but it has been characterized more as a problem with the stream delineation process that comes from thresholding the drainage accumulation maps \cite{nobre2016hand,li2020evaluation}.
+Later in this study, we will re-introduce this problem and demonstrate how we initialize with a stream network (that of the NWM's) and thus avoid having to threshold accumulations to some arbitrary value to define stream networks.
+We illustrate how computing HAND independently for stream lines of unit stream order can significantly enhance FIM performance by accounting for multiple sources of fluvial inundation that may exist in certain regions and flow scenarios.
 %
@@
 %
-Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
+Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (approximately 10 m depending on latitude) scale, HAND has been implemented into software for large-scale, continental computation.
 As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a HPC \cite{maidment2017conceptual,liu2016cybergis}.
 NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
-By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
-Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC as an unofficial demonstrations to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
-Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
-More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
-
-Related to these efforts, the United States Geological Survey (USGS) has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
+By allocating 31 nodes at 20 cores per node for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
+Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (NHDPlusHR) which better agrees with the 10 m DEM from the NHDPlusHR program \cite{liu2020height}.
+The original NFIE dataset was employed by the NWC as an unofficial demonstration to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
+Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
+More notably the software added the ability to derive HAND on both the NWM FR and MS stream networks to consider multiple sources of fluvial inundation along high impact rivers of primary forecasting concern.
+
+Related to these efforts, the United States Geological Survey (USGS) has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses SRCs with cross-sections for stage-discharge relationships \cite{verdin2016software}.
 Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
@@
 %
-Many of those in the academic community assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
-\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
-The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
-Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
-Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3 m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
-The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
-Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
-Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
-Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
-These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.
-
-With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
+In order to mitigate the ever increasing threat of flooding to life and property, an operational capability is required to extend NWM streamflow forecasts to river stages, inundation extents, and inundation depths.
+OWP FIM is introduced here as a continental scale capability that generates these products at high spatial and temporal resolutions.
 Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
-This framework is open-source and modular enabling and accelerating the research to operations development pipeline.
-Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
-In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
-Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which is closely related to stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
-This capability is introduced progressively as Mainstems (MS) and to a higher degree Generalized Mainstems (GMS) which will be explained later on.
+Within the operational framework of OWP FIM, we introduce research demonstrating how FIM performance skill with HAND can be improved by discretizing stream networks into units of an effective unit Horton-Strahler stream order \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} for HAND computation contexts.
+Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to the stream accumulation threshold which is closely related to stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
+Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths (LP), can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
+This capability is introduced progressively as MS (whose network represents about 4\% of FR network) and to a higher degree Generalized Mainstems (GMS) (covers entire FR network) which will be explained later on.
 The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
@@
 OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
+Figure \ref{fig:methods_overview} gives a high-level overview of the methodology used in OWP FIM and in this study.
+Input data from multiple sources are preprocessed (not illustrated) and then subset to processing areas based on the model used, FR, MS, or GMS.
+The standard processing unit of OWP FIM is a HUC8 however for the MS and GMS versions only the NWM MS stream network and the NWM FR network discretized into LPs are used, respectively.
+A series of hydro-conditioning steps enforces the location of stream lines, monotonically decreasing elevations, excavated bathymetry, stream thalweg breaching, and levee enforcement.
+After a DEM suitable for HAND's assumptions is conditioned, the FIM hydrofabric is generated including stream network, catchments, HAND, SRCs, and cross-walk table.
+The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains SRC and cross-walk information).
+In operational circumstances, the NWM streamflows are used in conjunction with the FIM hydrofabric to derive forecast FIMs.
+However for evaluation purposes, we use the streamflows from the cross-sections of our benchmark model.
+As later discussed in Section \ref{ssec:inundation_mapping}, independent FIMs from multiple fluvial sources are mosaiced together.
+For the case of MS, two sources are mosaiced together (FR and MS) while for GMS the inundation from every LP is composited together.
+The evaluation FIM extents are compared to the extents of the benchmark model and metrics are computed.
+%
+\begin{figure}[H]
+\centering
+\includegraphics[scale=6.0]{figures/methods_overview.jpg}
+\caption{
+Methodology overview detailing high-level steps followed in the study.
+The flow chart begins with the input data organized by source.
+Subsetting the data into processing units depends on which model is being considered.
+FR utilizes the entire NWM stream network processed at HUC8 processing areas.
+MS only computes HAND using the NWM stream at or downstream of legacy forecasting points.
+The resulting inundation from the MS HAND is eventually layered with the FIM from FR HAND to account for high levels of inundation contributed by the mainstem.
+Generalized Mainstem (GMS) discretized NWM streams into level paths (LP) then computes HAND and the FIMs independently only to mosaic them later. 
+This better accounts for multiple possible sources of fluvial inundation.
+The dotted lines denote the use of NWM streamflow forecasts to produce operational FIM but not used in this study.
+All acronyms used in the figure are defined in the paper.
+}
+\label{fig:methods_overview}
+\end{figure}
+%
 %
@@
 Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
+To simplify setup and enhance portability across host operating systems, OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
 A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub \cite{inundationMapping2022}.
-
-The pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
-Preprocessing acquires and prepares datasets for production of the FIM hydrofabric. 
-The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
-Functionality is included to turn FIM hydrofabric and streamflows into inundation maps represented in both binary and depths.
-Lastly, a test suite includes means of calculating evaluation metrics compared to a variety of preprocessed test case data.
-A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
+Source code is made available for this project on GitHub where a user could consult the Readme.md page for more information on how to acquire the datasets and reproduce the pipeline \cite{inundationMapping2022}.
 %
@@
 Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP) \cite{esipData2022}.
+Instructions for accessing data processed for OWP FIM are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by the Earth Science Information Partners (ESIP) \cite{esipData2022}.
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}.
 We utilized a series of data products from the NHDPlusHR including the BurnLineEvents \cite{nhdplus2022vectors}, Value Added Attributes (VAA) \cite{nhdplus2022vectors}, Water Boundaries (WBD) or HUC Layers \cite{nhdplus2022wbd}, and the DEM elevation rasters \cite{nhdplus2022dems}.
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes \cite{nwm2022hydrofabric}.
-For enforcing levee data into the DEMs, the USACE NLD is used to burn feature elevations \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{osm2021landsea,GreatLakesHydrographyDataset}.
-Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
-These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
+These BurnLines used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk against for forecasting purposes \cite{nwm2022hydrofabric}.
+For enforcing levee data, the USACE NLD is used to burn feature elevations into DEMs \cite{engineers2016national}.
+Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) \cite{osm2021landsea} and the land-lake border from Great Lakes Hydrography Dataset (GLHD) \cite{GreatLakesHydrographyDataset} to exclude those areas from production of FIMs.
+Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of nine states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year or yr) and 0.2\% (500 year or yr), are used for validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
+These BLE datasets are provided at the watershed scale (HUC8) utilizing best available DEMs and simulations from the Hydrologic Engineering Center's River Analysis System (HEC-RAS) model \cite{us2022hydrologic}.
 The full input datasets presented by source are listed in Table \ref{tab:data}.
@@
 \hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement & \cite{nhdplus2022vectors} \\
-\hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display & \cite{nhdplus2022vectors} \\
-\hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters & \cite{nhdplus2022dems} \\
-\hline
-USGS & NHDPlusHR WBD & Water Boundaries (WBD) or HUCs used for spatial processing units & \cite{nhdplus2022wbd} \\
-\hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. & \cite{nwm2022hydrofabric} \\
-\hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. & \cite{nwm2022hydrofabric} \\
-\hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. & \cite{nwm2022hydrofabric} \\
-\hline
-USACE & NLD & Levee database of locations and elevations & \cite{engineers2016national} \\
+USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlusHR for hydro-enforcement. & \cite{nhdplus2022vectors} \\
+\hline
+USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display. & \cite{nhdplus2022vectors} \\
+\hline
+USGS & NHDPlusHR DEM & DEM used for NHDPlusHR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters. & \cite{nhdplus2022dems} \\
+\hline
+USGS & NHDPlusHR WBD & Water Boundaries (WBD) or HUCs used for spatial processing units. & \cite{nhdplus2022wbd} \\
+\hline
+NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting adapted from NHDPlus V2 NHDFlowline\_Network feature class. & \cite{nwm2022hydrofabric} \\
+\hline
+NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM adapted from NHDPlus V2 Catchment feature class. & \cite{nwm2022hydrofabric} \\
+\hline
+NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes adapted from NHDPlus V2 NHDWaterbody feature class. & \cite{nwm2022hydrofabric} \\
+\hline
+USACE & NLD & Levee database of locations and elevations. & \cite{engineers2016national} \\
 \hline
@@
 \hline
-InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
+InFRM & Cross-Sections & HEC-RAS 1D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
 \hline
@@
 Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and GMS networks and 1,604 HUC8s for the MS method. 
-These methods will be explained more later.
-An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
+These methods will be explained in more detail later.
+An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which are expansion domains in the NWM V2.0 and V2.1, respectively.
 %
@@
 HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
-In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligble for flooding.
+In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligible for flooding.
 Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
@@
 %
-The location of the stream network is enforced to ensure general agreement with established stream networks especially the placement of the NWM network which is used for forecasting the stream flow inputs.
-The NHDPlusHR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
-However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus MR, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-For every NWM headwater segment a headwater point is derived and linearly interpolated to the nearest Burnline Event segment.
-Burnline Event headwater segments are split at the adjusted headwater point to match NWM flowlines.
-The resulting pruned NHDPlusHR stream network is what gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHDPlusHR network corresponds to the full density NHDPlusHR network at NWM V2.1 headwater locations only.
+The location of the stream network is enforced to ensure general agreement with the NWM network which is used for forecasting the streamflow inputs.
+The NHDPlusHR Beta BurnLineEvent layer is used to enforce stream locations in the NHDPlusHR workflow and best agrees with thalweg locations in the DEM used so it is also used here for hydro-enforcement \cite{moore2019user}.
+
+However, to better match the drainage density of the NWM FR V2.1 stream network, which is based on the NHDPlus V2, the BurnlineEvents are pruned utilizing a nearest neighbor search around the NWM flowlines.
+Headwater points are first derived for the NWM FR V2.1.
+For every NWM headwater point, the nearest NHDPlusHR point is selected and placed into a set and those not belonging to this nearest neighbor set are discarded.
+Only the nearest point on the NHDPlusHR is used so any portion of the NHDPlusHR network upstream of this nearest point is discarded to avoid extending inundation too far above the modeling domain.
+The points in this nearest neighbor set are then traversed downstream.
+Any headwater portion in the NHDPlusHR or any other stream not traversed are pruned away to better match the resolution and spatial locations of the NWM stream network and its headwater points.
+The resulting pruned NHDPlusHR stream network gets hydro-enforced in subsequent operations.
+This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHDPlusHR network corresponds to the NHDPlusHR network at or downstream of NWM V2.1 headwater locations only.
 Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
@@
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
 \includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection and linear interpolation.
-Resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR Burnline Events.}
+\caption{This figure illustrates some of the datasets that result from the pruning of NHDPlusHR Beta BurnlineEvents (dotted black) to the stream density of the NWM FR V2.1 density (blue).
+The stream network used for forecasting, NWM FR V2.1, is of higher stream density than that of the NHDPlusHR which has better agreement with the thalweg locations in the DEM used.
+Thus we opt to prune the NHDPlusHR network to match the general location and density of the NWM network.
+The nearest neighbor segment in the NHDPlusHR of each NWM headwater locations and the nearest point on that segment is determined to match the closest point to that of the NWM headwater.
+These points are then traversed downstream and any segments not traversed are pruned away.
+The resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR BurnlineEvents.}
 \label{fig:stream_density_pruning}
@@
 The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop which were set to fixed values of 70m, 1000m, and 10m, respectively.
-While the value of these parameters are critical to the inundation extents produced, especially for lower flow rates, finding their optimal value for OWP FIM was not done since it was out of the main scientific scope of this article.
+The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop and buffer distance. 
+The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling an inverted triangular notch shape with a significantly lower elevation along the thalweg line only.
+In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop which were set to fixed values of 70 m, 1000 m, and 10 m, respectively, but available to the user via the parameter file.
+While the values for these parameters are critical to the inundation extents produced, especially for lower flow rates where bathymetric information has more influence, finding their optimal values for OWP FIM was not done since it was out of the main scientific scope of this article.
 Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
+\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement techniques especially when computational costs are considered. 
 Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
-Some of these drawbacks are later addressed by additional conditioning techniques later on.
+Some of these drawbacks are addressed by additional conditioning techniques applied later on.
 %
@@
 %
-Coarse DEM's at 10m, 30m, and higher resolutions can lack sufficient representation of fine grain features such as embankments, floodwalls, and closure structures \cite{arundel2018assimilation,dobbs2010evaluation,wang2005comparison,sanders2007evaluation}.
-In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10 m DEM.
-The elevations found in the NLD are burned into the DEM if those elevations were found to exceed those already in place.
+Coarse DEM's at 10 m, 30 m, and higher resolutions can lack sufficient representation of fine grain features such as embankments, flood walls, and closure structures \cite{arundel2018assimilation,dobbs2010evaluation,wang2005comparison,sanders2007evaluation}.
+In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 1/3 arc-second DEM.
+The elevations found in the NLD are burned onto the DEM if those elevations were found to exceed those already in place.
 %
@@
 %
-Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
-The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
+Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes with HAND.
+The partially conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
 Priority-Flood developed by \citeA{barnes2014priority} is an algorithm for filling said depressions and shown to have improved performance over early works in the field by \citeA{jenson1988extracting} implemented in \citeA{tarboton2005terrain} as well as \citeA{planchon2002fast}.
@@
 This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
-The depression technique employed here does leave the existence of flat regions where pits existed a prior thus later requiring the need for resolving these flats.
+The depression filling technique employed here does leave the existence of flat regions where pits previously existed thus later requiring the need for resolving these flats.
 The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
@@
 %
-Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
-In order to prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
+Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing, elevations are derived prior to the normalizing of elevations.
+Work on the AGREE DEM method from several authors have illustrated that the AGREE DEM method does not prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another \cite{hellweger1997agree,baker2006comparison,saunders1999preparation,saunders1996gis,quenzer1998gis,saunders1995grid}.
+These works observe that the artificial elevations enforced by the hydrographically based stream network and AGREE DEM disagree with those naturally occurring in the native DEM.
+In order to mitigate this documented issue, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum on the original, unconditioned DEM for each thalweg pixel.
 Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50 m.
 The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
+This step essentially enforces an estimate of the native DEM thalweg elevations onto the sharp drop enforced thalweg elevations from the AGREE procedure.
 
@@
 \citeA{garousi2019terrain} proposed an algorithm that breaches stream thalweg pixel elevations in a depth first manner. 
-This procedure was found to increase the Critical Success Index of resulting FIMs from HAND and is employed in OWP FIM to enforce monotonically decreasing elevations with thalweg pixel networks.
+This procedure was found to increase the Critical Success Index (CSI) of resulting FIMs from HAND and is employed in OWP FIM to enforce monotonically decreasing elevations with thalweg pixel networks.
 %
@@
 The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
+These datasets include the HAND or relative elevation model (REM) raster, reach-level catchments raster/polygons, DEM-derived streamlines, SRCs, and cross-walk table.
+The following sub-sections describe how the subset and hydro-enforced geospatial datasets are converted into the FIM hydrofabric.
 %
@@
 D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
-Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
-
-Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
+The horizontal component of slope is defined as one for the four neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$ by means of the Pythagorean theorem.
+Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression-filled cells must be resolved.
+
+Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment} where flats are resolved by incrementing elevations iteratively.
 Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
 The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
+The established literature in this niche field of hydrology discusses how prevalent flats can be in given study areas and how difficult the problem is from both computational and hydrologic stand-points \cite{garbrecht1997assignment,tarboton2009generalized,tarboton2005terrain,survila2016scalable,barnes2014efficient,tesfa2011extraction,wallis2009parallel}.
 OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
@@
 %
-The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
-The FIM stream network is of similar drainage density as the NWM V2.1 network but fully converges at all junctions leaving no divergences in the network.
+The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new, DEM based, FIM stream network. 
+The FIM stream network is of similar drainage density as the NWM V2.1 network and fully converges at all junctions leaving no divergences in the network.
 This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
+These seeds points are pruned headwater locations of the NHDPlusHR Beta BurnlineEvents layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
 Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
 
-Each stream link in this derived FIM stream network is split into equidistant reaches.
+Each stream link in this derived FIM stream network is split into equidistant reaches of 1.5 km in length which is a user exposed parameter.
 Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments.
@@
 Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretation of the stream network.
-Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherent the synthetic rating curve properties of the upstream or downstream segment, depending on the topology.
-Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
-Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
+Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve SRC and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
+Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretization of the stream network.
+Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherit the SRC properties of the upstream or downstream segment, depending on the topology.
+Section \ref{sssec:synthetic_rating_curve} details the derivation of the SRC and the dependence on channel length. 
+Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC8 membership.
 This stream network is important since it drives the HAND calculation and derivation of catchments.
@@
 Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
-The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
-Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
-The other catchments are derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
+The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing, nearest drainage region for each outlet point.
+Two sets of catchments are derived, one set of catchments denotes the unique drainage region for each thalweg pixel which is used for relative elevation calculation.
+The other catchments are derived for the drainage region for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
 %
@@
 Once the pixel level catchments are derived, the final relative elevations can be computed.
-Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment described in Section \ref{sssec:catchments}.
+The elevation of every thalweg pixel is subtracted from the elevations of the non-thalweg pixels within the same, corresponding pixel-level catchment described in Section \ref{sssec:catchments}.
 The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
 Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling \cite{djokic2019arc}.
+Any negative values resulting from this subtraction with native elevations are replaced by zero.
+Again, HAND assumes and requires processing areas to drain thus have monotonically decreasing elevations with hydrologically correct flow directions all leading to a singular outlet point.
+While this is required for the generation of DEM-derived catchments and stream lines, it is not necessarily required for the computation of the relative elevations.
+Since the use of hydro-conditioning processes to fit the drainage requirement for HAND can be extensive, we found it more fitting to use the native elevations this final HAND computation thus avoid the use of manipulated values that fit modeling assumptions.
 %
@@
 %
- A method for converting forecast river discharges from the NWM to stages or river depths is necessary for producing FIMs with HAND. 
-For one-dimensional models such as the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a synthetic rating curve at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
-For this application, we utilized the reach averaged approach for developing synthetic rating curves (SRC) \cite{zheng2018river}.
+A method for converting forecast river discharges from the NWM to stages or river depths at the reach scale is necessary for producing FIMs with HAND. 
+For 1D hydrodynamic models such as the routing methods in the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a SRC at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
+For this application, we utilized the reach averaged approach for developing SRCs \cite{zheng2018river}.
 The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-The reach averaged Manning's formula is derived to be 
+Previously not reported in literature to our knowledge in this form, the reach averaged Manning's formula is derived to be 
 %
@@
 \label{eq:reach_averaged_mannings_equation}
-\textbf{Q} = \frac{1}{n} \frac{V^{5/3}S^{1/2}}{L B^{2/3}} 
+Q(y) = \frac{1}{n} \frac{V(y)^{5/3}S^{1/2}}{L B(y)^{2/3}} 
 \end{equation}
@@
 %
-where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
-Q, V, and B are taken a specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
-All units are international given the 1 numerator above n.
-The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
+where Q is discharge at stage y, n is the Manning's n roughness coefficient, V is volume at y, S is channel slope, L is the along-flow reach length, and B is wetted bed area at y.
+Q, V, and B are taken at specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
+All units are international (SI) given the one in the numerator above n.
+The reach averaged method has been compared to rating curves from HEC-RAS and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
 The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the channel length.
-
-Setting of the Manning's n roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
+Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the reach length.
+While the AGREE DEM is subject to hydro-conditioning processes, it does introduce some notion of bathymetry estimation that the native DEMs lack while being sensitive to additional parameters that could yield further errors in the FIM.
+We leave this issue open in this study and elaborate on needs with respect to bathymetry and Manning's n values in the Discussion section (Section \ref{sec:discussion}).
+
+Setting of the Manning's n roughness coefficient has precedent in previous continental-scale FIM (CFIM) studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
 These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
+We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but not limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
+For now and for the purpose of this study, we examine the SRCs with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
+After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values (y in Eq.
+\ref{eq:reach_averaged_mannings_equation}) from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
 %
@@
 %
-The stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with a NWM reach identifier so that a discharge can be converted to stage and later inundation extent.
-For the methods already discussed, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent methods discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
-Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
-While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
-More discussion on this will follow in Section \ref{sec:discussion}.
+The DEM based stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with NWM reach identifiers so that a discharge can be converted to stage and later inundation extents and depths.
+For the FR version of HAND, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
+For two subsequent HAND methods, MS and GMS, discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, respectively, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
+Additionally, only relevant catchments from the NWM for the given LP are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
+While these conflation methods are approximate, they can lead to some substantial errors which will be discussed more in Section \ref{sec:discussion}.
 %
@@
 As previously discussed, HAND based FIMs are subject to many assumptions and limitations in order serve as a suitable inundation proxy for large scale, high resolution domains.
-HAND produced FIMs are limited in only providing inundation sourced from the nearest drainage which in many instances can only be of a several sources of fluvial surface water inundation.
-The forecasting community, in reviewing HAND, have noted significant negative effects at the confluence of lower flow tributaries with higher flow rivers for which the term catchment boundary issue has been termed.
+HAND produced FIMs are limited in only providing inundation sourced from the nearest drainage line, however, depending on flow conditions and topography, a given area may have multiple contributing fluvial sources of surface water inundation.
+The forecasting community, in reviewing HAND, have noted significant negative effects at the confluence of lower flow tributaries with higher flow rivers for which the phrase ``catchment boundary issue'' has been termed.
 In previous studies, FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND which is closely related to the maximum Horton-Strahler stream order of the network \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is partly in due to the limitation that catchment boundaries place on inundation extents.
+This sensitivity is partly in due to the limitation that catchment boundaries place on inundation extents where only the nearest drainage line can source inundation for any particular area.
 
 Figure \ref{fig:catchment_boundaries_issue} illustrates the exact situation our solution proposes to address where two tributaries converge with a higher order stream segment. 
-An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
+An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 yr extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
 The higher discharge along the main segment in Figure \ref{fig:catchment_boundaries_issue} of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS.
@@
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
 \includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{The figure represents an agreement map between a HAND derived FIM and one produced from the Base Level Engineering (BLE) program for a 100yr magnitude event at HUC8 12090301.
-Agreement maps are symbolized by false negatives (FN), true negatives (TN), false positives (FP), and true positives (TP) where inundated represents the postive condition (see Section \ref{ssec:evaluation} for more details).
-The stream flows associated with each river segment are shown in CMS while the flow directions are symbolized as red arrows.
+\caption{The figure represents an agreement map between a HAND derived FIM and one produced from the Base Level Engineering (BLE) program for a 100 yr magnitude event at HUC8 12090301.
+Agreement maps are symbolized by false negatives (FN), true negatives (TN), false positives (FP), and true positives (TP) where inundated represents the positive condition (see Section \ref{ssec:evaluation} for more details).
+The streamflows associated with each river segment are shown in CMS while the flow directions are symbolized as red arrows.
 The presence of FNs at the confluence of tributaries (circled in red) with the main segment is associated with lower flow rates in the tributaries that don't account for backwater effects.
@@
 We seek to resolve this catchment boundary problem or nearest drainage limitation by discretizing the target stream network into stream networks of reduced, unit stream order to avoid the constraining of catchments by those belonging to lower order neighbors.
-By discretizing the network into stream networks of unit stream order (later defined as level paths), we remove the influence of neighboring catchments that constrain the inundation extent.
+By discretizing the network into stream networks of unit stream order (later defined as LP), we remove the influence of neighboring catchments that constrain the inundation extent.
 This creates much larger and overlapping catchments that can source fluvial inundation from multiple reaches as required by the given river stage at current flow conditions.
-We present two successive methods, National Weather Service MS (Section \ref{sssec:nws_mainstems}) and GMS (Section \ref{sssec:generalized_mainstems}), implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND by expanding the nearest drainage definition to increase potential inundation areas.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Main-stems}
+We present two successive methods, National Weather Service MS (Section \ref{sssec:nws_mainstems}) and GMS (Section \ref{sssec:generalized_mainstems}), implemented that reduce the effective Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND by expanding the nearest drainage definition to increase potential inundation areas.
+
+To clarify the phrase ``reducing Horton-Strahler stream order'' used extensively in this paper, every FIM used in evaluation contains a flood extent sourced from every NWM forecast point in the given evaluation domain.
+What we do to reduce stream order is discretize the NWM FR network into different units of size, MS network (\ref{sssec:nws_mainstems}) and GMS LPs (\ref{sssec:generalized_mainstems}), that effectively reduce the HAND computation to independent networks of unit stream order.
+These independent HAND datasets are later used to produce FIM independently and mosaiced together (see Section \ref{ssec:inundation_mapping}).
+The inundation from the MS HAND is mosaiced with the inundation from FR HAND, while the inundation of each individual LP from GMS is mosaiced together.
+The Horton-Strahler stream order is only reduced for HAND computation purposes to reduce the negative effects of the nearest drainage limitation inherent to HAND.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{NWS Mainstems}
 \label{sssec:nws_mainstems}
@@
 The initial attempt at drainage order reduction to solve the catchment boundary issue was to use a stream network relevant to the NWS forecasting community. 
-The Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
+The Mainstems (MS) network is a subset of the NWM FR network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
 The MS network comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
@@
 Inundation is derived independently from the resulting FR and MS HAND hydrofabric are mosaiced together using the technique proposed in Section \ref{ssec:inundation_mapping} to form the MS FIMs. 
-Within each HUC, you'll typically only find a MS stream network of stream order 1 (i.e.
-headwater) but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
+Within each HUC, one might typically only find a MS stream network of uniform stream order but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
 So while we may refer to the MS network as that of one with unit stream order, we acknowledge there are many cases where additional or converging forecast points create multiple branches within a given processing unit.
@@
 Since MS only covers 4\% of the entire FR stream network, we sought to expand drainage order reduction techniques to all reaches within the NWM modeling domain.
-In order to do this, we discretized the NWM network into level paths which when considered individually have unit Horton-Strahler stream orders thus significantly reducing drainage density.
-Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-Starting at the outlet, a unique level path is propagated upstream. 
-At every confluence, the direction of maximum flow path length is sought to propagate the current level path identifier.
-For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
-
-Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
+In order to do this, we discretized the NWM network into LPs which when considered individually have unit Horton-Strahler stream orders.
+LPs group flowlines by maximizing the length of each flow path and minimizing the number of LP identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
+In order to derive LPs for the NWM FR stream network at the HUC8 processing area, we first compute arbolate sums which are defined as the cumulative drainage distance of all upstream drainage lines.
+Arbolate sum is also inclusive of the current drainage reach as well.
+Arbolate sums are computed by starting at the headwater points and summing up drainage distances as you traverse downstream.
+
+Arbolate sum is critical to discretizing the NWM network into LP identifiers.
+Starting at a HUC8's outlet, a unique LP is propagated upstream. 
+At every confluence, the direction of maximum arbolate sum is sought to propagate the current LP identifier.
+For the remaining parent reaches of the given junction, a new LP identifier is assigned and the process recursively continues with them.
+Figure \ref{fig:level_path_methods} illustrates how LPs (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
+The figure shows computed arbolate sums and unique LP identifiers for the HUC12 120903010404 for clarity but were computed at the corresponding HUC8.
+The mainstem of the figure runs from the red ellipses to the black one which is the outlet.
+From the figure, we can see how unique colors are propagated in the direction of the maximum arbolate sum.
+
+Each HUC8 is discretized into LPs independently and the relevant inputs as described in Table \ref{tab:data} are assigned to each LP processing unit given a buffer of seven km.
 This buffer was selected to avoid edge contamination \cite{lindsay2013measuring} and to ensure adequate data availability for wide rivers with large catchments in regions with low slope.
-Further work could be dedicated to tune this user exposed parameter to better balance its effect on FIM extents and computational expense since larger buffers create additional floating points and calculations to process and store.
-At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
-Once the NWM FR stream network is discretized into level paths, we independently compute HAND using each level path as the target stream network to be used.
-
+Further work could be dedicated to tune this user exposed parameter to better balance its effect on FIM extents and computational expense since larger buffers create additional floating point calculations and storage requirements.
+For the time being, we designate this issue to be out of scope.
+
+At the LP scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the LP in question at the time.
+The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement by burning these lines and seeding from its headwater points directly instead of going through the NHDPlusHR network as described in Section \ref{ssec:stream_network_enforcment} .
+This decision was motivated by the difficulty in deriving LPs in the NWM stream network with high agreement with the LPs derived for the NHDPlusHR stream lines.
+We found that the same algorithm to compute arbolate sums and LPs could yield enough disagreements associated with disordered branches or slight differences in arbolate sums that could significantly affect the agreement of the LP identifiers in the NWM and NHDPlusHR networks.
+This yielded enough error to justify the use of the NWM directly for hydro-enforcement operations.
+
+Once the NWM FR stream network is discretized into LPs, we independently compute HAND using each LP as the target stream network to be used.
 To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
-In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
+In Figure \ref{fig:gms_methods}a, we uniquely color code the LPs derived for the NWM stream network. 
 For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each level path is buffered to a polygon with a user-available distance parameter of 7 km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
-We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
+Each LP is buffered to a polygon with a user-exposed, distance parameter of seven km that is used to subset the original DEM for two selected LPs in Figure \ref{fig:gms_methods}b.
+We illustrate two HAND grids for two of the LPs in this HUC8 in Figure \ref{fig:gms_methods}c.
+Once the FIM hydrofabrics for each LP are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
 Lastly, these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}d.
@@
 For a more intimate look at the drainage order reduction procedure and its effects, GMS, we allude to Figure \ref{fig:gms_methods_2} which references the same area (in HUC8 12090301) and set of river junctions as in Figure \ref{fig:catchment_boundaries_issue}.
-The catchments and stream lines for HAND computed at the FR scale are illustrated in Figure \ref{fig:gms_methods_2}a where the respective inundation (for 100 yr magnitude) is heavily constrained by the limited catchment extents especially at junctions.
-In subsequent sub-figures, we show the same datasets for the HAND computation problem for this region but discretized into independent level paths for the main level path (b), the eastern tributary (c), and the western tributary (d).
-Notably, inspecting (b) one sees how removing the tributaries creates much larger catchments for the main level path. 
-These catchments include drainage areas that would traditionally be considered nearest to the tributaries thus ineligible to receive inundation sourced from the main level path.
+The catchments and stream lines for HAND computed at the FR scale are illustrated in Figure \ref{fig:gms_methods_2}a where the respective inundation at the 100 yr magnitude is heavily constrained by the limited catchment extents especially at junctions.
+In subsequent sub-figures, we show the same datasets for the HAND computation problem for this region but discretized into independent LPs for the main LP (b), the eastern tributary (c), and the western tributary (d).
+Notably, inspecting (b) one sees how removing the tributaries creates much larger catchments for the main LP. 
+These catchments include drainage areas that would traditionally be considered nearest to the tributaries thus ineligible to receive inundation sourced from the main LP.
 The inundation extents in (b) overlap those of (c) and (d) and are mosaiced together by methods explained in Section \ref{ssec:inundation_mapping}.
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
 \includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates how level paths (LP) for the NWM stream network are derived.
-They are derived by starting at an outlet point with a unique ID which are symbolized by unique colors.
-Once a junction is reached, the current level path identifier is propagated in the direction of highest arbolate sum.
-The remaining converging members at the given junction are each assigned a new unique LP identifier and the process is repeated recursively until all reaches have been assigned a level path.
-The values shown labeling the reaches above are arbolate sums which is defined as the cumulative distance of all upstream reaches.
-Thus, level paths serve as a proxy means of assigning membership to a given river when presented with a confluence.
-Each individual level path has a unit Horton-Strahler stream order thus serves as a great method for our proposed technique.
+\caption{Illustrates the NWM Full Resolution V2.1 stream network discretized into LPs (LP), symbolized by unique colors, as well as the values of the arbolate sums to the nearest whole km distance.
+The LPs were derived on a HUC8 level (12090301) but only illustrated for a HUC12 (120903010404) for clarity.
+The mainstem of this HUC12 runs from the red ellipses to the outlet denoted by the black ellipses.
+Arbolate sums are defined as the cumulative drainage distances of all upstream stream lines.
+Arbolate sums are computed for the NWM network by starting at the headwater points then traversing downstream and adding the distances cumulatively. 
+LPs are derived by starting at an outlet point with a unique identifier (ID).
+The unique LP ID is propagated upstream until a junction is reached where the current LP ID is propagated in the direction of maximum arbolate sum.
+The remaining converging segments at the given junction are each assigned a new unique LP ID and the process is repeated recursively until all reaches have been assigned a LP.
+Thus, LP serve as a proxy means of assigning membership to a given river when presented with a confluence.
+Each individual LP has a unit Horton-Strahler stream order thus serves as a great method for our proposed technique.
 }
@@
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
@@
 \caption{Overall procedure for GMS HAND at HUC8 12090301.
-In (a), we illustrate all NWM stream lines symbolized by their level path with 372 unique level path IDs in this HUC.
-Meanwhile (b), demonstrates the DEM clipped to a 7 km buffer around two selected level paths.
-In (c), we show how HAND can be computed just for each one of these two level paths independently. 
-We also show inundation maps created for these two level paths in (c). 
-In (d), we show all the inundation maps for all the level paths mosaiced together. }
+In (a), we illustrate all NWM stream lines symbolized by their LP with 372 unique LP IDs in this HUC.
+Meanwhile (b), demonstrates the DEM clipped to a seven km buffer around two selected LPs.
+In (c), we show how HAND can be computed just for each one of these two LPs independently. 
+We also show inundation maps created for these two LPs in (c). 
+In (d), we show all the inundation maps for all the LPs mosaiced together. }
 \label{fig:gms_methods}
@@
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
 \includegraphics[scale=0.70]{figures/gms_methods_2.jpg}
-\caption{This image, with the same spatial domain as Figure \ref{fig:catchment_boundaries_issue} (HUC 12090301), demonstrates how computing HAND on level path bases leads to larger, independent catchments and more expansive inundation extents (100 yr flows).
+\caption{This image, with the same spatial domain as Figure \ref{fig:catchment_boundaries_issue} (HUC 12090301), demonstrates how computing HAND on level path (LP) bases leads to larger, independent catchments and more expansive inundation extents (100 yr flows).
 In (a), the catchments and stream network are shown for HAND computed in Full-Resolution (FR) method which shows constrained inundation extents around the two junctions.
-(b) demonstrates the level path associated with this region's highest order river.
-By delineating catchments at this scale indepedent of the neighboring tributaries, the drainage areas are allowed to expand thus allowing inundation extents to cover previously restricted areas.
+(b) demonstrates the LP associated with this region's highest order river.
+By delineating catchments at this scale independent of the neighboring tributaries, the drainage areas are allowed to expand thus allowing inundation extents to cover previously restricted areas.
 In (c) and (d), we show the stream lines, catchments, and inundation extents of the two tributaries.
@@
 For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
-From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 8 spatial processing units in a rapid, parallel operation.
+From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC8 spatial processing units in a rapid, parallel operation.
 The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
 
-Utilizing the stage-discharge relationships in the synthetic rating curves, each forecast for each catchment identifier is assigned a stage value. 
+Utilizing the stage-discharge relationships in the SRCs, each forecast for each catchment identifier is assigned a stage value. 
 The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
@@
 Other aggregation methods could lead to different results but were not investigated here.
-Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
-Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps.
+Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i,
 %
@@
 %
-For the MS and GMS methods, the contributing areas are defined differently.
-For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to 2 for that case.
-For GMS, all FIMs from all the level paths in a given area are mosaiced together then K is set to this number of level paths.
-Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
+. Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps,
 %
@@
 \end{linenomath*}
+. For the MS and GMS methods, the contributing areas are defined differently.
+For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to two for that case.
+For GMS, all FIMs from all the LPs in a given area are mosaiced together then K is set to this number of LPs.
+Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
 %
@@
 Possible benchmark FIM candidates for evaluation purposes include high-water marks, remote sensing observation, crowd-sourced information, and modeled extents.
-Unfortunately, all of these sources or using a combination of them are subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to lack of spatial coverage, signal interference, lack of stream flow data, inaccurate stream flow data, physics-based assumptions, and errors in input data.
-While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated stream flow data necessary to make FIMs to compare to.
+Unfortunately, all of these sources or using a combination of them are subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
+While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to isolating out hydrological factors.
 A suitable dataset needed to be selected to best balance advantages and limitations with future studies interested in expanding to additional datasets as able to.
 
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
-This dataset was selected due to its large spatial coverage, availability of cross-sections with stream flow information, more sophistication when compared to HAND based FIMs, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
-49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM so we used those corresponding discharges and mapping extents for evaluation.
-We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
-
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
+Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
+This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, more sophistication when compared to HAND based FIMs, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
+49 HUC8s spanning about 185 thousand $km^2$ were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
+The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM and thus corresponding discharges and mapping extents for evaluation.
+We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation sourced from OWP FIM.
+
+By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, and others that we would have potentially affected our conclusions if we used NWM forecasted discharges.
 Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
@@
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
 \includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
-\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states at 0.2\% recurrence magnitude for flow rates.}
+\caption{
+Shows 185 thousand $km^2$ of modeled areas for the Base Level Engineering (BLE) domain of 49 HUC8s across nine states at 0.2\% recurrence magnitude for flow rates.
+BLE maps are produced for two recurrence flows, 1\% (100 yr) and 0.2\% (500 yr), using 1D HEC-RAS models.
+The maps are used as benchmarks for validation purposes of OWP FIM.
+}
 \label{fig:all_ble_maps}
@@
 BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM V2.1 Full Resolution (FR) stream lines.
-Additionally, we illustrate the NWM V2.1 catchments to provide a sense of how many cross-sections may intersect a given NWM flow line.
+Additionally, we illustrate the NWM V2.1 catchments to provide a sense of how many cross-sections may intersect a given NWM flowline.
 The BLE stream network is also shown which is denser than the NWM V2.1 stream lines meaning there are several lower order streams represented in the BLE stream network that are not in the NWM V2.1 stream lines.
@@
 
-The metrics employed in this study to evaluate inundation extents include Critical Success Index (CSI), Probability of Detection (POD), and False Alarm Ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
-To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.
-The two types of errors consist of false positives (FP), or type I errors, which is dry in benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in benchmark by predicted dry.
+The metrics employed in this study to evaluate inundation extents include CSI, Probability of Detection (POD), and False Alarm Ratio (FAR) and are presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
+To calculate these secondary metrics, one must define three primary metrics starting with true positives (TP) which is predicted wet and wet in the BLE benchmark dataset.
+The two types of errors consist of false positives (FP), or type I errors, which is dry in the benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in the benchmark but predicted dry.
 Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
-Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
 Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-In other words, POD is an indicator of model skill in inundated regions while FAR is an indicated of model skill in non-inundated regions.
+In other words, POD is an indicator of model skill in inundated regions while FAR is an indicator of model skill in non-inundated regions.
+Some work by \citeA{gerapetritis2004behavior} denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
 We use all three secondary metrics here to add value to the discussion while avoiding aggregating away the meaning of all four primary metrics.
@@
 Thus, frequency dependent statistics should be used with caution when comparing across sites with varying frequencies. 
-Lastly, approximately 6 HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability.
+Lastly, approximately six HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability to compare GMS metrics to.
 %
@@
 We computed the statistics, CSI, POD, and FAR, for both 100 and 500 year events for Mannings N set to 0.06 and 0.12.
-These results are presented in Figure \ref{fig:violin_plot} as violin plots and in Table \ref{tab:aggregate_metrics} as aggregated metrics with the results for MS and GMS presented as percentage changes from their respective FR values.To be more specific, Table \ref{tab:aggregate_metrics} sums the primary metrics, TP, FP, FN, and TN, across all HUC8's then recomputes the secondary metrics which was done to better account for large variances in HUC8 size.
+These results are presented in Figure \ref{fig:violin_plot} as violin plots and in Table \ref{tab:aggregate_metrics} as aggregated metrics with the results for MS and GMS presented as percentage changes from their respective FR values.
+To be more specific, Table \ref{tab:aggregate_metrics} sums the primary metrics, TP, FP, FN, and TN, across all HUC8s then recomputes the secondary metrics which was done to better account for large variances in HUC8 size.
 The same trends discussed below are consistent across both reporting methods (Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}).
@@
 For example, let's examine the violin plot for the row marked CSI and column for Manning's n = 0.06.
-This sub-figure shows the CSI distributions across all 49 HUC8's when Manning's n is set to 0.06.
-Each independent, whole violin represents the distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100yr and orange for 500yr).
+This sub-figure shows the CSI distributions across all 49 HUC8s when Manning's n is set to 0.06.
+Each independent, whole violin represents the distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100 yr and orange for 500 yr).
 The horizontal dashed and dotted lines represent the 25th, 50th, and 75th percentiles from bottom to top, respectively.
 Additionally, we show trend lines symbolized in green that for each metric and Manning's n combination denotes the best fit line for the three methods (FR, MS, and GMS).
-To avoid having two trend lines per sub-figure, we elected to aggregate the two magnitudes togther as they tend to observe similar trends across the three models.
-The slope of each trend line is quantified in the Figure by its $\beta_1$ value and the p-value for that statistic which tests the significance of that trend (deviation from a zero sloped trend line).
+To avoid having two trend lines per sub-figure, we elected to aggregate the two magnitudes together as they tend to observe similar trends across the three models.
+The slope of each trend line is quantified in the figure by its $\beta_1$ value and the p-value for that statistic which tests the significance of that trend (deviation from a zero sloped trend line).
 
@@
 Also, we note that FAR is somewhat, albeit marginally, decreased from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
-The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}.
-So overall and in other words, the entire distribution of HUC8's improves across the three methods.
-Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
-We move more details on the relationship betweeen stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).
-
-Additional noteworthy trends in Figure \ref{fig:violin_plot} center around the all-around better performance of FIM's for those of higher Manning's n values and recurrence flows.
+The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}).
+So overall and in other words, the entire distribution of HUC8s improves across the three methods.
+Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit stream order and mosaic the resulting FIMs, the better those FIMs perform.
+We move more details on the relationship between stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).
+
+Additional noteworthy trends in Figure \ref{fig:violin_plot} center around the all-around better performance of FIMs for those of higher Manning's n values and recurrence flows.
 The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value used in the BLE model, 0.11 \cite{fema2021base,fema2021estimated}.
-Most of this improvement is driven by significant increases in POD but, unfortunately, leading to a significant amount of overprediction as observed by the increase in FAR.
+Most of this improvement is driven by significant increases in POD but, unfortunately, leading to a significant amount of over-prediction as observed by the increase in FAR.
 More work can be invested to better regionalize Manning's n values for FIM purposes with HAND.
-We also observe additional trends associated with the magnitude or recurrence interval of the flow rates used with the higher flow rates exhibting better overall CSI, POD, and FAR values than the lower, 100 yr magnitude.
+We also observe additional trends associated with the magnitude or recurrence interval of the flow rates used with the higher flow rates exhibiting better overall CSI, POD, and FAR values than the lower, 100 yr magnitude.
 We introduce in the Discussion (Section \ref{sec:discussion}) that this skill premium exhibited by higher flow events could be due higher quality elevation data in regions that are not described as bathymetric areas.
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
@@
 %
-\begin{table}[h!]
+\begin{table}[H]
 \caption{ Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated to the BLE domain.
-         The values for MS and GMS methods are expressed in percentage change (\%) from their respective values with the same Manning's N, magnitude, and metric combination in the Full Resolution (FR) method columns.
+         The values for MS and GMS methods are expressed in percentage change (\%) from their respective values with the same Manning's n, magnitude, and metric combination in the Full Resolution (FR) method columns.
          The best value across models is highlighted in bold.}
@@
 \cline{3-8}
-  &  & 100yr & 500yr & 100yr & 500yr & 100yr & 500yr \\
+  &  & 100 yr & 500 yr & 100 yr & 500 yr & 100 yr & 500 yr \\
 \hline
@@
 For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
-GMS which generalizes HAND production to level path scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
+GMS which generalizes HAND production to the LP scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
 %
@@
 Most of this change is accounted for by substantially increasing POD and inundation extents in some areas thus reducing FNs.
-We believe, as we later argue for, that the increase in POD is primarily driven by an increase in the catchment sizes that is inhertently enabled by dividing up the stream network into independent stream networks of unit stream order.
-Additionally, we note that reducing drainage order also has some minor influence on reducing inundation extents and the rate of false alarms.
-We believe that this effect is driven by a reduction in some inundation areas led by a change in the stage-discharge relationship where a given streamflow now leads to lower river stage values when HAND is computed with target stream networks of unit drainage order.
-We seek to explain that these two effects, catchment boundary enlargment and stage-discharge curve lowering, are highly interrelated and cannot be easily detangled.
+We believe, as we later argue for, that the increase in POD is primarily driven by an increase in the catchment sizes that is inherently enabled by dividing up the stream network into independent stream networks of unit stream order.
+Additionally, we note that reducing drainage order also has some minor influence on reducing inundation extents in other areas and the rate of false alarms.
+We believe that this effect is driven by a change in the stage-discharge relationship where a given streamflow now leads to lower river stage values when HAND is computed with target stream networks of unit drainage order.
+We seek to explain that these two effects, catchment boundary enlargement and stage-discharge curve lowering, are highly interrelated and cannot be easily detangled.
 Lastly, we discuss the diminishing effects on performance that the MS and GMS techniques may have and also any additional effects including enhanced cross-walking abilities.
 
-As evident in the results of the study in Section \ref{sec:results}, a sizeable amount of the increase in CSI observed by reducing stream order for HAND computation can be attributed to increases in POD.
+As evident in the results of the study in Section \ref{sec:results}, a sizable amount of the increase in CSI observed by reducing stream order for HAND computation can be attributed to increases in POD.
 This can be inferred by close inspection of Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} where changes in POD are significantly higher than that of FAR.
-This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
-
+%This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
 Upon investigation of the performance of HAND derived FIM, we note a general increase of catchment sizes for the MS and GMS methods when compared to the FR method as they are now delineated independently of any tributaries that would constrain catchment sizes otherwise.
@@
 This allows for inundation extents to expand across regions previously encapsulated by catchments of joining reaches in lower flow tributaries.
-The water built up along the catchment boundaries can be thought of as a column of water in a cyclindrical container (catchment) that has exceeded the elevation of the containers rim.
+The water built up along the catchment boundaries can be thought of as a column of water in a cylindrical container (catchment) that has exceeded the elevation of the containers rim.
 This, of course, does not represent accurate physics but it is what does occur in HAND when water accumulates along catchment boundaries.
 
-Figure \ref{fig:gms_enhancement} best illustrates this behaviour in real context.
+Large scale HUC8 level evaluations can fail to demonstrate fine grain enhancements as they aggregate away many changes that are only clear at more local scales.
+Future assessments of OWP FIM should consider finer grain evaluation units as well possible impact assessments using asset information such as building footprints to better illustrate fine grain changes in a more relevant manner to stakeholders.
+For now, we offer up Figure \ref{fig:gms_enhancement} which best illustrates the improvement offered by multi-source fluvial flooding capabilities in a more local context.
 The figure is comprised of two agreement rasters for two different HAND based FIMs compared to the validation dataset for a given region with a high flow mainstem (500 yr recurrence flow) running horizontally along the region.
-Sub-figure \ref{fig:gms_enhancement}a demonstrates the agreement raster for the FR stream network as well as it's respective catchment boundary lines symbolized in white and stream network shown in green.
-Inspection of this sub-figure denotes a clear spatial pattern where TP's or areas correctly inundated are pooled alongside catchment boundary lines. 
-On the other side of the catchment boundary, one can witness large swaths of FN's that should be inundated. 
-The FN's also exhibit a spatial pattern as in they tend to collocate within catchments of the pictured mainstems tributaries.
-This sort of behaviour was introduced early in the paper and shown qualitatively in Figure \ref{fig:catchment_boundaries_issue}.
+Sub-figure \ref{fig:gms_enhancement}a demonstrates the agreement raster for the FR stream network as well as its respective catchment boundary lines symbolized in white and stream network shown in green.
+Inspection of this sub-figure denotes a clear spatial pattern where TPs or areas correctly inundated are pooled alongside catchment boundary lines. 
+On the other side of the catchment boundary, one can witness large swaths of FNs that should be inundated. 
+The FNs also exhibit a spatial pattern as in they tend to collocate within catchments of the pictured mainstems tributaries.
+This sort of behavior was introduced early in the paper and shown qualitatively in Figure \ref{fig:catchment_boundaries_issue}.
 
@@
 While the entire mosaiced inundation map from GMS (as described in Section \ref{ssec:inundation_mapping} and Equation \ref{eq:hand_fim_depth}) is used to produce this agreement map, we only show the catchments associated with the mainstem of this region that is shown to follow a clear horizontal path.
-Showing all the catchments for the tributaries that were all derived independently would convulate the image.
+Showing all the catchments for the tributaries that were all derived independently would convolute the image.
 The main message illustrated here is that the catchments associated with the mainstem of this area significantly increase in size.
@@
 The consequence for inundation extents is a general increase in spatial coverage of the river's water which shows to be in much better agreement with the benchmark map.
-The TP's are no longer bounded by the catchment lines and allowed the expand to their natural extents.
-
-We note here as a contribution of this study that a major inherant, limitation of HAND is the ``nearest drainage'' constraint or the idea that a given river reach only drains or, in HAND's case, inundates its immediate, unique drainage area or catchment.
-In other words, HAND based FIM's are limited to producing fluvial inundation to only their nearest drainage area or catchment.
+The TPs are no longer bounded by the catchment lines and allowed the expand to their natural extents.
+
+We note here as a contribution of this study that a major inherent, limitation of HAND is the ``nearest drainage'' constraint or the idea that a given river reach only drains or, in HAND's case, inundates its immediate, unique drainage area or catchment.
+In other words, HAND based FIMs are limited to producing fluvial inundation to only their nearest drainage area or catchment.
 However, we know that fluvial inundation can be sourced from several streams nearby that also serve as drainage outlets to the area in question.
-Generally speaking, drainage areas are known to be hiearchical in nature so a given drainage area for a given outlet point can be decomposed into nested drainage areas for outlet points that lie in the original drainage area.
+Generally speaking, drainage areas are known to be hierarchical in nature so a given drainage area for a given outlet point can be decomposed into nested drainage areas for outlet points that lie in the original drainage area.
 A perfect example of this are points that lie on tributary reaches closely neighboring a mainstem.
-These points lie in the drainage area of reaches in the mainstem but inundation from the mainstem cannot reach these tributary catchments because of the "nearest" assumption in HAND.
-Hence it's important to state that just like there are different sources of flooding such as fluvial, pluvial, groundwater, reservoir, and coastal, there can also be multiple sources of a riverine flood.
-HAND is only equipped to handle rivirine flooding from the nearest drainage line.
+These points lie in the drainage area of reaches in the mainstem but inundation from the mainstem cannot reach these tributary catchments because of the ``nearest'' assumption in HAND.
+Hence it's important to state that just like there are different sources of flooding such as fluvial, pluvial, groundwater, reservoir, barrier failure (dam/levee/embankment), and coastal, there can also be multiple sources of a riverine flood.
+HAND is only equipped to handle riverine flooding from the nearest drainage line.
 Other relevant drainage lines that produce fluvial flood waters are not considered here especially if the routing model used doesn't consider backwater effects.
 
-The remaining portion of the improvement in CSI was found to come from a marginal yet noteable reduction in FAR.
-Upon investigation of the spatial results in the agreement maps, we found some areas of slight reductions in FPs especially where changes in catchment boundaries may have occurred due to the reduction in stream order.
-These observations pointed to changes in the rating curves introduced by stream order reduction and catchment definition adjustments.
-Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
-Sub-figure \ref{fig:rating_curve_comparison}a shows how the average rating curves for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to bias down (and to the right) with ever increasing stream order reduction (FR to MS to GMS).
+The remaining portion of the improvement in CSI was found to come from a marginal yet notable reduction in FAR.
+Upon investigation of the spatial results in the agreement maps, we found some areas of slight reductions in FPs especially where changes in catchment boundaries may have occurred due to the reduction in effective stream order in computing HAND.
+These observations pointed to changes in the SRCs introduced by stream order reduction and catchment definition adjustments.
+Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on SRCs.
+Sub-figure \ref{fig:rating_curve_comparison}a shows how the average SRCs for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to shift the curve down and to the right with ever increasing stream order reduction (FR to MS to GMS).
 This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
 
-Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
+Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged SRC relationship between stage and discharge.
 Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
 While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
-Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius.
+Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius but these associations only hold when reach length, L, is considered.
 Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively (see Equation \ref{eq:reach_averaged_mannings_equation}).
-Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
+Figures \ref{fig:rating_curve_comparison}b and \ref{fig:rating_curve_comparison}c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
 Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
 Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
-We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
+We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the SRC Q up and changes the SRC accordingly.
 In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
@@
 Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
-While the cause of these errors can be varied, we assert here that conflating 4 networks for use in evaluations leads to significant error.
+While the cause of these errors can be varied, we assert here that conflating four networks for use in evaluations leads to significant error.
 As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
-One of the issues is when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
-The utilization of MS and GMS only conflates to NWM catchments directly associated with the level path in question which is inherently easy to do with those methods. 
+One of the issues with the FR version of HAND occurs when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
+The utilization of MS and GMS only conflates to NWM catchments directly associated with the LP in question which is inherently easy to do with those methods. 
 Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
-The NWM stream network was derived using the NHDPlus MR  dataset which was derived from coarser DEMs than those used here. 
+The NWM stream network was derived using the NHDPlus V2 dataset which was derived from coarser DEMs than those used here. 
 Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
@@
 
-Our qualitative analysis suggests that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM for future development.
+Our qualitative analysis suggests that the SRCs offer a significant opportunity for improvement in HAND based FIM for future development.
 The bathymetry of the 10 m DEM from 3DEP is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
@@
 Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
-We suspect that a significant amount of the difference in performance between 100yr and 500yr magnitude events can be attributed to poor synthetic rating curve performance due to poor bathymetric representation.
+We suspect that a significant amount of the difference in performance between 100 yr and 500 yr magnitude events can be attributed to poor SRC performance due to poor bathymetric representation.
 Lower magnitude events are, logically, more susceptible to poor bathymetric data due a greater proportion of the inundation being attributable to areas that are more typically under normal flow conditions.
-Higher flow events tend to cover regions
+Higher flow events tend to cover regions with more floodplain inundation thus less sensitive to errors from bathymetric data quality.
+On a related note, the use of the AGREE DEM method discussed in Section \ref{ssec:stream_network_enforcment} also interacts with the bathymetry issue introducing several artificial geometry parameters that affect SRC shape and quality.
+Due to focus on the nearest drainage problem, we leave future work related to SRC representation including roughness estimation, bathymetric data assimilation, and bathymetry adjustments as opportunities for major enhancements in HAND based FIM.
 
@@
 These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
-Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
-%
-\begin{figure}[h!]
+Ultimately, the use of more advanced 2D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
+%
+\begin{figure}[H]
 \centering
@@
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
 \includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
-\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105 at the 500yr recurrence magnitude.
+\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105 at the 500 yr recurrence magnitude.
 Catchment boundaries and stream lines are shown in white and dotted green, respectively.
@@
 Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
-The National Water Model (NWM) developed by the National Oceanic and Atmospheric Administration's Office of Water Prediction, along with partners, provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
-Additional processing is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
+The National Water Model (NWM), developed by the National Oceanic and Atmospheric Administration's (NOAA) Office of Water Prediction (OWP) and the National Center for Atmospheric Research (NCAR), provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
+Additional modeling is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
 Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
 HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
-Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
+Scalable, open-source software, known as OWP FIM, was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
 HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
-Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
+Finally, we used this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
 FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
+
+The main contribution and conclusion of this work centers around a fundamental limitation in HAND based FIM which is a failure to account for multiple possible sources of fluvial inundation since HAND only considers inundation from the nearest drainage line.
 We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
-This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
-
-Dissecting the stream network into regions of level-paths with unit stream order is necessary because HAND has a ``nearest drainage'' limitation meaning it only accounts for riverine inundation sourced from the nearest drainage line.
-While this is all that is required in many, lower flow instances, typically larger flow scenerios can call for multi-source riverine flooding to occur especially at the junction of major rivers with their smaller tributaries.
-Computing HAND independently for each level-path removes the boundary of neighboring catchments thus allowing for 
-Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.
+In order to reduce stream order of the NWM stream network for HAND computation, we dissected the NWM network into two simpler units of singular Horton-Strahler stream order and mosaiced the resulting FIMs derived from each.
+The NWM Mainstems (MS) stream network, which covers roughly 4-5\% of the NWM Full Resolution (FR) network, spans all established forecasting points in the Advanced Hydrologic Prediction System (AHPS) and downstream reaches.
+The inundation from MS derived HAND is mosaiced together with the inundation of FR derived HAND.
+Extending order reduction to the entire network, the Generalized Mainstem (GMS) technique discretizes the NWM FR network into level paths (LP) of unit stream order for HAND computation.
+All LP based FIM derived from LP based HAND datasets are mosaiced together to form one seamless FIM.
+Dissecting the stream network into regions of LPs with unit stream order is necessary because HAND has a ``nearest drainage'' limitation meaning it only accounts for riverine inundation sourced from the nearest drainage line.
+In our evaluation of this technique, we observe that HAND based FIM improves in skill as we extend from nearest drainage inundation in FR to multiple drainage support in MS for only 4-5\% of the FR network.
+Extending multiple drainage support to the entire FR network with GMS based HAND improves upon skill at around the same magnitude that MS improved upon FR.
+Thus we conclude that deriving HAND with independent stream networks of unit Horton-Strahler stream order enhances the skill of FIM but offers diminishing returns as we extend from 4-5\% of the network with MS to 100\% of the network with GMS since deriving HAND and FIMs at these localized scale does add computational expense.
+
+This primary contribution also affects the parameters used to compute stage-discharge relationships shifting discharge higher at given stages which reduced the rate of false positives.
+This shift in SRC behavior is driven by larger catchments that influence reach averaged geometric parameters in the Manning's equation.
+Related to SRCs, we noted better results and more sensitivity to unit stream order networks with the higher Manning's n value of 0.12 when compared to 0.06 for high magnitude events at 1\% (100 year or yr) and 0.2\% (500 year or yr) recurrence intervals.
+Additionally, we noted better performance for more intense 500 yr events which we attribute to a stronger influence of poor quality bathymetric data in 100 yr magnitude inundation extents.
+While the AGREE DEM procedure is meant to add some bathymetry primarily motivated to enhance catchment and stream line delineation, it does introduce three parameters that have major implications in the quality of SRCs and the resulting FIMs.
 Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
+We leave questions related to Manning's n localization as well as bathymetry integration, estimation, and/or calibration open for future research to answer.
+Two other issues left open for improvement include the integration of higher resolution Lidar-based digital elevation maps (DEM) as well as the use of physics-based models for continental scale, high resolution forecasting applications.
 Due to inherent limitations with HAND, scalable, physics-based methods are necessary to consider to provide a better representation of flood extent dynamics in steady and unsteady conditions.
  %
+% Acknowledgments
+\clearpage % this clears figures before references
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Acknowledgments
+\acknowledgments
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+This work was funded by the Office of Water Prediction (OWP) which is part of the National Oceanic and Atmospheric Administration's (NOAA) National Weather Service (NWS).
+Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
+We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
+Additionally, David Blodgett from the United States Geolgical Survey (USGS) Water Mission Area was instrumental in helping define level paths and other hydrographic work.
+More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
+Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
+ %
+% Open Research
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\openresearch
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
+OWP Flood Inundation Mapping (FIM) capabilities rely extensively on the use of the National Hydrography Plus High Resolution (NHDPlusHR) datasets including BurnLineEvents \cite{nhdplus2022vectors}, value-added attributes (VAA) \cite{nhdplus2022vectors}, water boundaries (WBD) or hydrologic unit code (HUC) geometries \cite{nhdplus2022wbd}, and digital elevation maps (DEM) \cite{nhdplus2022dems}.
+Some additional datasets for processing include the National Levee Database (NLD) furnished by the United States Army Core of Engineers (USACE) \cite{engineers2016national}, Land-sea border from the Great Lakes Hydrography Dataset (GLHD) furnished by the Great Lakes Aquatic Habitat Framework (GLAHF) \cite{GreatLakesHydrographyDataset}, and a Land-sea border provided by OpenStreetMap (OSM) \cite{osm2021landsea}.
+Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
+Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
+
+Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022}.
+ %
+% Bibliography
+\clearpage % this clears figures before references
+\bibliography{bibliography/owp_fim4_2022}
 %
@@
 % AGU does not normally allow dedications.
-
-\acknowledgments
-This work was funded by the OWP which is part of NOAA's National Weather Service.
-Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
-We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
-Additionally, David Blodgett from the USGS Water Mission Area was instrumental in helping define level paths and other hydrographic work.
-More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
-Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
-All data availability and open research information is cited in the Software and Datasets sections, \ref{ssec:datasets} and \ref{ssec:software}, respectively.
-Please see those sections, Table \ref{tab:datasets}, and their References for more information.
-
 %% ------------------------------------------------------------------------ %%
@@
 %
-\clearpage % this clears figures before references
-\bibliography{bibliography/owp_fim4_2022}
 %

### 2022-08-21 manuscripts/owp_fim4_2022/sections/acknowledgments.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

--- 2022-02-14 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-%
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements, including customized APA reference formatting.
-%
-% You may copy this file and give it your
-% article name, and enter your text.
-%
-%
-% Step 1: Set the \documentclass
-%
-%
-
-%% To submit your paper:
-\documentclass[draft]{dependencies/agujournal2019}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage[inline]{dependencies/trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-
-%% added
-\usepackage{multirow}
-\usepackage{amsmath}
-
-\linenumbers
-
-%%%%%%%
-% As of 2018 we recommend use of the TrackChanges package to mark revisions.
-% Add track changes editors
-\addeditor{Fernando Aristizabal}
-
-% The trackchanges package adds five new LaTeX commands:
-%
-%  \note[editor]{The note}
-%  \annote[editor]{Text to annotate}{The note}
-%  \add[editor]{Text to add}
-%  \remove[editor]{Text to remove}
-%  \change[editor]{Text to remove}{Text to add}
-%
-% complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%
-
-\drafttrue
-
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-% ie, \journalname{Water Resources Research}
-
-\journalname{Water Resources Research}
-
-% starts document
-\begin{document}
-
-%% Title and Authors
-%% ------------------------------------------------------------------------ %%
-%  Title
-%
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
-
-% Example: \title{This is a test title}
-%
-\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
-
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-%
-%
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
-%
-%
-\affiliation{1}{Lynker, Leesburg, VA, USA}
-\affiliation{2}{Center for Remote Sensing, Agricultural and Biological Engineering Department, University of Florida, Gainesville, FL, USA}
-\affiliation{3}{National Water Center, Office of Water Prediction, National Weather Service, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
-\affiliation{4}{New York Water Science Center, Hydrologic Applied Innovations Lab, United States Geological Survey, Troy, NY, USA}
-\affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Maryland, College Park, MD, USA}
-%
-%(repeat as many times as is necessary)
-%
-%% Corresponding Author:
-% Corresponding author mailing address and e-mail address:
-
-% (include name and email addresses of the corresponding author.  More
-% than one corresponding author is allowed in this LaTeX file and for
-% publication; but only one corresponding author is allowed in our
-% editorial system.)
-%
-% Example: \correspondingauthor{First and Last Name}{email@address.edu}
-%
-\correspondingauthor{Fernando Aristizabal}{fernando.aristizabal@noaa.gov}
-
-%% Keypoints
-%% Keypoints, final entry on title page.
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-
-% Example:
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-% \end{keypoints}
-%
-\begin{keypoints}
-\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
-\item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
-\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation.
-\end{keypoints}
-%
-
-%% Abstract
-% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT and PLAIN LANGUAGE SUMMARY
-%
-% A good Abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
-
-% The Plain Language Summary should be written for a broad audience,
-% including journalists and the science-interested public, that will not have 
-% a background in your field.
-%
-% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
-% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
-% see http://sharingscience.agu.org/creating-plain-language-summary/)
-%
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
-
-\begin{abstract}
-The National Water Model (NWM) currently requires the post-processing of forecast discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
-Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is worthy of producing high-resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
-However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
-To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that reduces the Horton-Strahler stream order of the stream network.
-GMS contains all segments within the NWM stream network but instead of deriving HAND by accounting for all river segments at once, it is derived independently at the level path (LP) scale.
-LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
-These FIMs are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
-Improvement in mapping skill is observed when compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
-A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
-\end{abstract}
-%
-\section*{Plain Language Summary}
-Flooding is one of the most impactful natural disasters on life and property.
-The United States National Water Model (NWM) provides flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
-In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
-We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-We compared these HAND derived FIMs to maps from more realistic models and found improvement in mapping performance.
-%
-
-%%% Suggested section heads:
-% \section{Introduction}
-%
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
-
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-%
-% \section{Data} (Or section title might be a descriptive heading about data)
-%
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-%
-% \section{Conclusions}
-
-% Introduction
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
-In 2017 and 2019, river and flash flooding combined represented the leading cause of death and the second leading cause in 2018 among all natural disasters in the US \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
-More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
-With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
-Trends related to flood damages and fatalities have been steadily increasing over recent decades.
-\cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
-Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
-Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Operational Forecasting}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
-One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble stream flows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
-Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
-AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
-On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
-Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons to address the increasingly complex water challenges facing the US.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/forecast_points.jpg}
-\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
-Also show are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
-\label{fig:forecast_points}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{National Water Model}
-\label{ssec:national_water_model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
-The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
-The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
-As a complement to the operational NWM, OWP has also developed a configuration of the NWM that extends RFC forecasts downstream by assimilating and routing forecast streamflows to the next downstream AHPS forecast point.
-This configuration of the NWM is used to enhance forecasting skill by leveraging best available regional-scale data.
-The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
-Figure \ref{ssec:national_water_model} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
-The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
-
-The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
-The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
-Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
-As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
-The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Height Above Nearest Drainage}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
-\citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
-The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
-\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
-This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{HAND Implementations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
-As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
-NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
-By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
-Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC as an unofficial demonstrations to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
-Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
-More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
-
-Related to these efforts, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
-Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
-In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30 m resolutions \cite{yamazaki2019merit,donchyts2016global}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{OWP FIM}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Many of those in the academic community assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
-\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
-The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
-Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
-Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3 m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
-The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
-Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
-Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
-Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
-These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.
-
-With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
-Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
-This framework is open-source and modular enabling and accelerating the research to operations development pipeline.
-Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
-In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
-Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
-This capability will be referred to as Generalized Mainstems (GMS).
-The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
-%
- %
-% Materials and Methods
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
+% Acknowledgments
+\acknowledgments
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Software Dependencies and Architecture}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
-Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
-A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub \cite{inundationMapping2022}.
-
-The pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
-Preprocessing acquires and prepares datasets for production of the FIM hydrofabric. 
-The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
-Functionality is included to turn FIM hydrofabric and streamflows into inundation maps represented in both binary and depths.
-Lastly, a test suite includes means of calculating evaluation metrics compared to a variety of preprocessed test case data.
-A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}
-\label{ssec:datasets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP) \cite{esipData2022}.
-The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
-For enforcing levee data into the DEMs, the USACE National Levee Database (NLD) is used to burn feature elevations \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{OpenStreetMap,GreatLakesHydrographyDataset}.
-Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
-These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
-The full input datasets presented by source are listed in Table \ref{tab:data}.
-%
-\begin{table}
-\caption{Data sources, names, and descriptions.}
-\label{tab:data}
-\centering
-\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
-%\begin{tabular}{l c c}
-\hline
-Source & Name & Description \\
-\hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
-\hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
-\hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters \\
-\hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. \\
-\hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. \\
-\hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. \\
-\hline
-USACE & NLD & Levee database of locations and elevations  \\
-\hline
-OSM & Land-Sea Border & Border of land and sea. \\
-\hline
-GLHD & Land-Great Lakes Border & Border of land and Great Lakes. \\
-\hline
-InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
-\hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
-\hline
-\end{tabular}
-\end{table}
-%
-Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and GMS networks and 1,604 HUC8s for the MS method. 
-These methods will be explained more later.
-An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hydro-conditioning}
-\label{ssec:hydro_conditioning}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
-These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
-HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
-In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligble for flooding.
-Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
-It can also be used to simulate more accurate bathymetry which is not accounted for in the 10 m DEM \cite{gesch2002national}.
-
-Specifically within the context of OWP FIM, the hydro-conditioning operations that take place in sequential order are presented. 
-Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing units of size HUC8. 
-The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reach code's membership in the processing unit.
-Hydro-conditioning raster operations take place on buffered boundary definitions to avoid edge contamination and effects \cite{lindsay2013measuring}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Network Enforcement} 
-\label{ssec:stream_network_enforcment}
-%
-The location of the stream network is enforced to ensure general agreement with established stream networks.
-The NHDPlusHR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
-However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-For every NWM headwater segment a headwater point is derived and linearly interpolated to the nearest Burnline Event segment.
-Burnline Event headwater segments are split at the adjusted headwater point to match NWM flowlines.
-The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection and linear interpolation.
-Resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR Burnline Events.}
-\label{fig:stream_density_pruning}
-\end{figure}
-%
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.
-
-The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
-The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
-Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
-Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
-Some of these drawbacks are later addressed by additional conditioning techniques later on.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Levee Enforcement}
-%
-The DEM at 10 m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
-In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10 m DEM.
-The elevations found in the NLD are burned into the DEM if those elevations were found to exceed those already in place.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Depression Filling}
-\label{sssec:depression_filling}
-%
-Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
-The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
-Priority-Flood developed by \citeA{barnes2014priority} is an algorithm for filling said depressions and shown to have improved performance over early works in the field by \citeA{jenson1988extracting} implemented in \citeA{tarboton2005terrain} as well as \citeA{planchon2002fast}.
-The depression filling algorithm used in our pipeline is a Priority-Flood variant developed by \cite{zhou2016efficient} with enhanced single-thread performance and a time complexity of O(n log n) for floating point grids.
-This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
-The depression technique employed here does leave the existence of flat regions where pits existed a prior thus later requiring the need for resolving these flats.
-The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Thalweg Elevation Conditioning}
-\label{sssec:stream_thalweg_elevation_conditioning}
-%
-Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
-In order to prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
-Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50 m.
-The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
-
-The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
-\citeA{garousi2019terrain} proposed an algorithm that traverses stream thalweg pixels in a depth first manner starting with adding all the headwater pixels to a queue. 
-The connectivity of the thalweg pixels is defined by the D-8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
-At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in Equation \ref{eq:thalweg_breach},
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:thalweg_breach}
-\textbf{D}[x] = \min_{y\ drains\ to\ x} {(\ \textbf{D}[x]\ ,\ \textbf{D}[y]\ )}
-\end{equation}
-\end{linenomath*}
-%
-, in which \textbf{D} represents the array of thalweg adjusted elevations indexed by x and y where by y is upstream of x. 
-When a pixel's upstream neighbors are all evaluated, the downstream pixel is added into the queue thus the depth first traversal of the drainage network.
-This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in Critical Success Index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Deriving FIM Hydrofabric}
-\label{ssec:deriving_fim_hydrofabric}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Flow Directions and Flats Resolution}
-\label{ssec:flow_direction_and_flat_resolution}
-%
-To facilitate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D-8 flow directions.
-D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
-The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southeast direction as 8. 
-Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
-
-Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
-Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
-The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
-Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
-By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Deriving FIM Stream Network}
-\label{sssec:deriving_fim_stream_network}
-%
-The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
-The FIM stream network is of similar drainage density as the NWM V2.1 network but fully converges at all junctions leaving no divergences in the network.
-This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
-Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
-Each stream link in this derived FIM stream network is split into equidistant reaches.
-Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments.
-Stream links are then further segmented at NWM lakes and HUC8 boundaries.
-Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretation of the stream network.
-Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherent the synthetic rating curve properties of the upstream or downstream segment, depending on the topology.
-Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
-Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
-This stream network is important since it drives the HAND calculation and derivation of catchments.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Catchments}
-\label{sssec:catchments}
-%
-Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
-Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
-The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
-Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
-The other catchments are derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Height Above Nearest Drainage}
-%
-Once the pixel level catchments are derived, the final relative elevations can be computed.
-Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment described in Section \ref{sssec:catchments}.
-The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
-Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Synthetic Rating Curves}
-\label{sssec:synthetic_rating_curve}
-%
- A method for converting forecast river discharges from the NWM to stages or river depths is necessary for producing FIMs with HAND. 
-For one-dimensional models such as the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a synthetic rating curve at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
-For this application, we utilized the reach averaged approach for developing synthetic rating curves (SRC) \cite{zheng2018river}.
-The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-The reach averaged Manning's formula is derived to be 
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:reach_averaged_mannings_equation}
-\textbf{Q} = \frac{1}{n} \frac{V^{5/3}S^{1/2}}{L B^{2/3}} 
-\end{equation}
-\end{linenomath*}
-%
-where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
-Q, V, and B are taken a specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
-All units are international given the 1 numerator above n.
-The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
-The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the channel length.
-
-Setting of the Manning's n roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
-These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cross-walking with NWM Stream Network}
-\label{sssec:cross_walking_networks}
-%
-The stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with a NWM reach identifier so that a discharge can be converted to stage and later inundation extent.
-For the methods already discussed, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent methods discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
-Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
-While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
-More discussion on this will follow in Section \ref{sec:discussion}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stream Order Reduction}
-\label{ssec:stream_order_reduction}
-%
-FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
-Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
-An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the MS of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
-This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the MS that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the MS catchments.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{False negatives associated with confluence of tributaries with MS. Integers represent flow values from BLE 100 year event for the associated areas. 
-No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
-\label{fig:catchment_boundaries_issue}
-\end{figure}
-%
-We seek to resolve this problem by deriving HAND for processing units with stream networks of reduced stream order. 
-We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
-The resulting FIMs from the overlapping HAND processing units are mosaiced together taking any inundated area to be inundated but more will be explained in Section \ref{ssec:inundation_mapping}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Main-stems}
-\label{sssec:nws_mainstems}
-%
-The Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
-The MS network comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
-It also spans 121,724 reaches across 1,608 HUC8s.
-HAND was originally derived for this stream network to enhance mapping skill along these critical MS segments \cite{djokic2019arc}. 
-The inundation derived from this stream network is mosaiced with the inundation from the FR network to form the MS FIMs. 
-Within each HUC, you'll typically only find a MS stream network of stream order 1 (i.e.
-headwater) but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Generalized Mainstems}
-\label{sssec:generalized_mainstems}
-%
-To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call GMS.
-Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
-Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-Starting at the outlet, a unique level path is propagated upstream. 
-At every confluence, the direction of maximum flow path length is sought to propagate the current level path identifier.
-For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
-Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
-At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
-
-To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
-In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
-For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each level path is buffered to a polygon with a user-available distance parameter of 7 km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
-We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
-Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}c.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates how level paths for the NWM are derived.
-Level paths symbolized by lines of unique colors are propagated upstream following the direction of maximum arbolate sum at each junction.
-}
-\label{fig:level_path_methods}
-\end{figure}
-%
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/gms_methods.jpg}
-\caption{Overall procedure for GMS HAND.
-In (a), we illustrate all NWM stream lines symbolized by their level path.
-Meanwhile (b), demonstrates the DEM clipped to a 7 km buffer around two selected level paths.
-In (c), we show how HAND can be computed just for each one of these two level paths independently. 
-We also show inundation maps created for these two level paths in (c). 
-In (d), we show all the inundation maps for all the level paths mosaiced together. }
-\label{fig:gms_methods}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Inundation Mapping}
-\label{ssec:inundation_mapping}
-%
-The FIM hydrofabric consisting of the relative elevations grid, catchments grid, catchment polygons, rating curve, and cross-walking data are all used to convert forecasts from the NWM into forecasts extents.
-For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
-From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 8 spatial processing units in a rapid, parallel operation.
-The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
-
-Utilizing the stage-discharge relationships in the synthetic rating curves, each forecast for each catchment identifier is assigned a stage value. 
-The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
-We use the basic logic already established in previous works to conduct this \cite{nobre2016hand,liu2016cybergis,maidment2017conceptual}.
-Mathematically, the HAND values, $H_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
-For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
-For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim_depth}
-    D_{ij} = S_i - H_{ij}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim}
-    F_{ij} = D_{ij} > 0
-\end{equation}
-\end{linenomath*}
-%
-For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast in the form of a single raster file.
-For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
-Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
-Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim_depths}
-    D_{ij} = \max_{k=[1,...,K]} D_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-For the MS and GMS methods, the contributing areas are defined differently.
-For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to 2 for that case.
-For GMS, all FIMs from all the level paths in a given area are mosaiced together then K is set to this number of level paths.
-Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim}
-    F_{ij} = \max_{k=[1,...,K]} F_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}
-\label{ssec:evaluation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
-49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM so we used those corresponding discharges and mapping extents for evaluation.
-We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
-Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
-We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
-To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
-This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
-While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
-\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states. This dataset for 1\% and 0.2\% recurrence flows were used as benchmarks.}
-\label{fig:all_ble_maps}
-\end{figure}
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
-\caption{Illustrates Base Level Engineering (BLE) cross sections and stream lines.
-The BLE stream network, which is denser than the NWM V2.1 stream lines, is also shown.
-BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM reach.
-}
-\label{fig:ble_evaluation_method}
-\end{figure}
-%
-The metrics employed in this study to evaluate inundation extents include Critical Success Index (CSI), Probability of Detection (POD), and False Alarm Ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
-To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.
-The two types of errors consist of false positives (FP), or type I errors, which is dry in benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in benchmark by predicted dry.
-Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
-Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
-While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
-Thus, frequency dependent statistics should be used with caution when comparing across sites with varying frequencies. 
-Lastly, approximately 6 HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:csi}
-CSI = \frac{TP}{TP + FN + FP}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:pod}
-POD = \frac{TP}{TP + FN}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:far}
-FAR = \frac{FP}{TP + FP}
-\end{equation}
-\end{linenomath*}
-%
- %
-% Results
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{sec:results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Mapping Performance}
-\label{ssec:mapping_performance}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-We produced FIMs for the entire BLE domain within the 49 HUC8 areas across several states in the south central US. 
-The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
-We computed the statistics (CSI, POD, and FAR) for both 100 and 500 year events for Mannings N set to 0.06 and 0.12. 
-The distribution of these statistics can be examined in Figure \ref{fig:violin_plot} as violin plots.
-Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
-We also denote trend lines for each metric and Manning's n setting as well as their respective slope estimate and one-tailed p-value denoting the level of significance of the trend.
-
-Aggregating the metrics in the method above treats each HUC8 as it's own unit and does little to consider the size differences of the HUCs. 
-In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics}  the CSI, POD, and FAR recomputed for the entire domain using the sum of all the TPs, FPs, and FNs. 
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.9]{figures/violin_plots.jpg}
-\caption{Shows kernel density estimation of the distributions (sample size = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
-The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's n (0.06 and 0.12).
-Trend lines for each metric - Mannings combination are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}
-\label{fig:violin_plot}
-\end{figure}
-%
-\begin{table}[h!]
-\caption{Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated for BLE domain.
-         The best value across models is highlighted in bold.}
-\label{tab:aggregate_metrics}
-\centering
-%\begin{tabular}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
-\begin{tabular}{|c|c||c|c|c|c|c|c|}
-\hline
-\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{MS} & \multicolumn{2}{|c|}{GMS} \\
-\cline{3-8}
-  &  & 100yr & 500yr & 100yr & 500yr & 100yr & 500yr \\
-\hline
-\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & \textbf{0.5796} & \textbf{0.6075} \\
-\cline{2-8}
-  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & \textbf{0.6182} & \textbf{0.6435} \\
-\hline
-\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & \textbf{0.6633} & \textbf{0.6863} \\
-\cline{2-8}
-  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & \textbf{0.7606} & \textbf{0.7810} \\
-\hline
-\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1787 & \textbf{0.1589} & \textbf{0.1778} & \textbf{0.1589} \\
-\cline{2-8}
-  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & \textbf{0.2324} & \textbf{0.2148} \\
-\hline
-\end{tabular}
-\end{table}
-%
-% Interpretation of metrics
-From Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, we denote several meaningful trends. 
-Using CSI as an overall proxy for skill of the FIM, we note that generally speaking the skill is correlated with a reduction of the stream orders of the processing units used for HAND.
-In other words, the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
-While this trend is evident for both sets of Manning's n values, the trend is slightly more significant for the higher value of 0.12.
-Other trends related to this Figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
-We also note how the higher Manning's n value enhances performance for both of these recurrence intervals across all models.
-
-Dissecting the improvements and trends presented in the previous paragraph comes down mostly to improvement in POD or a reduction in absolute amount of FNs.
-POD being the primary driver in skill enhancement is evident across models by comparing the slope of the POD lines with the slope of the FAR lines.
-Even though aggregating metrics by HUC8 yields a statistically zero trend, one does see a slight reduction in FAR across models that reduce HAND's maximum stream order.
-Additionally, we note that POD is a primary driver in enhancing performance across Manning's n values as well.
-This significant improvement comes at a cost of false alarms as the FAR increases significantly across Manning's n values.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Computational Performance}
-\label{ssec:compuational_performance}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The NFIE experiments were able to produce HAND for 331 HUC6's in 1.34 CPU years \cite{liu2016cybergis} and estimates using work from \citeA{djokic2019arc} put producing HAND at the FR NWM at 0.55 CPU years.
-For our work, we were able to produce HAND at the full NWM resolution in 0.13 CPU years which represents a substantial speed-up compared to previous works.
-For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
-GMS which generalizes HAND production to level path scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
-%
- %
-% Discussion
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-\label{sec:discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Overall, we note a positive relationship between FIM skill and a reduction of the stream order of the stream network we use to derive the HAND datasets.
-Most of this change is accounted for by increasing POD thus reducing FNs especially along higher order rivers with higher flow magnitudes.
-We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
-This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
-
-In analyzing the data, we found a slight reduction in FAR was detected and more digging pointed to a bias in rating curves introduced by stream order reduction.
-Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
-Sub-figure \ref{fig:rating_curve_comparison}a shows how the average rating curves for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to bias down (and to the right) with ever increasing stream order reduction (FR to MS to GMS).
-This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
-Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
-Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
-While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
-Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius.
-Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively (see Equation \ref{eq:reach_averaged_mannings_equation}).
-Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
-Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
-Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
-We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
-In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
-Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
-Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments that encompass neighboring tributaries.
-These tributaries have a significant amount of bathymetry that is low-lying thus easily including the SRC derivation. 
-They also contribute volume and bed area that is technically not perpendicular to the flux of streamflow being accounted for in the stream in question. 
-Careful examination of Figure \ref{fig:gms_enhancement}b shows how much larger catchments include neighboring tributaries and the geometry associated with those tributaries. 
-This geometry is not perpendicular to the flow that is associated with the main reach thus leading to biases in the SRC.
-We consider this fact to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
-
-Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
-While the cause of these errors can be varied, we assert here that conflating 4 networks for use in evaluations leads to significant error.
-As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
-One of the issues is when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
-The utilization of MS and GMS only conflates to NWM catchments directly associated with the level path in question which is inherently easy to do with those methods. 
-Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
-The NWM stream network was derived using the NHD medium resolution dataset which was derived from coarser DEMs than those used here. 
-Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
-Until a singular stream network is used for the NWM, BLE benchmark, and for HAND based FIM, conflation will continue being a source of error.
-
-Our qualitative analysis suggests that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM for future development.
-The bathymetry of the 10 m DEM from 3DEP is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
-Manning's n which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
-Some adjusting parameter must be introduced to enhance the estimation of the bathymetric representation.
-Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
-
-Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
-For HAND to be derived and thus create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
-In other words, an entire area eligible for flooding must monotonically decrease in elevation. 
-DEM's naturally don't do this and the dynamics of true flood events don't follow drainage patterns.
-Enforcing this assumption for HAND leads to significant amount of DEM manipulations that introduce basic errors.
-These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
-Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/rating_curve_comparison.jpg}
-\caption{Illustrates average quantities for the three methods, FR, MS, and GMS, for each stage value (m). 
-The values are (a) Discharge $m^3s^{-1}$, (b) Volume $m^3$, (c) Bed Area $m^2$, (d) a function of Volume and Bed Area, and (e) number of pixels.
-}
-\label{fig:rating_curve_comparison}
-\end{figure}
-%
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
-\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105.
-Catchment boundaries and stream lines are shown in white and dotted green, respectively.
-Sub-figure (a) shows agreement of FR HAND denoting significant areas of under-prediction due to junctions and catchment boundaries.
-Meanwhile, (b) shows the agreement for GMS and much larger catchments leading to much better inundation agreement for this given reach. 
-Overall, this illustrates the benefits of stream order reduction for deriving HAND datasets.
-}
-\label{fig:gms_enhancement}
-\end{figure}
-%
- %
-% Conclusions
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-\label{sec:conclusions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Floods present a significant, under-served, and expanding risk to life, property, and resources.
-Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
-The National Water Model (NWM) developed by the National Oceanic and Atmospheric Administration's Office of Water Prediction, along with partners, provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
-Additional processing is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
-Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
-HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
-Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
-HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
-Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
-We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
-This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
-FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
-Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.
-Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
-Due to inherent limitations with HAND, scalable, physics-based methods are necessary to consider to provide a better representation of flood extent dynamics in steady and unsteady conditions. 
-%
- %
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-%\section{= enter section title =}
-%Text here ===>>>
-%\section{= enter section title =}
-%Text here ===>>>
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables;  other caption information
-%  should be placed in last line of the table, using
-% \multicolumn2l{$^a$ This is a table note.}
-%
-%----------------
-% EXAMPLE FIGURES
-%
-% \begin{figure}
-% \includegraphics{example.png}
-% \caption{caption}
-% \end{figure}
-%
-% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
-% \begin{figure}
-% \noindent\includegraphics[width=\textwidth]{anothersample.png}
-%\caption{caption}
-%\label{pngfiguresample}
-%\end{figure}
-%
-%
-% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
-This is common when trying to add a PDF figure without pdflatex.
-% \begin{figure}
-% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
-%\caption{caption}
-%\label{pdffiguresample}
-%\end{figure}
-%
-%
-% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
-%
-
-%
-% ---------------
-% EXAMPLE TABLE
-%
-% \begin{table}
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \centering
-% \begin{tabular}{l c}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Footnote text here.}
-% \end{tabular}
-% \end{table}
-
-%% SIDEWAYS FIGURE and TABLE
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{figsamp}
-% \caption{caption here}
-% \label{newfig}
-% \end{sidewaysfigure}
-%
-%  \begin{sidewaystable}
-%  \caption{Caption here}
-% \label{tab:signif_gap_clos}
-%  \begin{tabular}{ccc}
-% one&two&three\\
-% four&five&six
-%  \end{tabular}
-%  \end{sidewaystable}
-
-%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
-%\begin{linenomath*}
-%\begin{equation}
-%y|{f} \sim g(m, \sigma),
-%\end{equation}
-%\end{linenomath*}
-
-%%% End of body of article
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary, Notation or Acronym section goes here:
-%
-%%%%%%%%%%%%%%
-% Glossary is only allowed in Reviews of Geophysics
-%  \begin{glossary}
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \end{glossary}
-
-%
-%%%%%%%%%%%%%%
-% Acronyms
-%   \begin{acronyms}
-%   \acro{Acronym}
-%   Definition here
-%   \acro{EMOS}
-%   Ensemble model output statistics
-%   \acro{ECMWF}
-%   Centre for Medium-Range Weather Forecasts
-%   \end{acronyms}
-
-%
-%%%%%%%%%%%%%%
-% Notation
-%   \begin{notation}
-%   \notation{$a+b$} Notation Definition here
-%   \notation{$e=mc^2$}
-%   Equation in German-born physicist Albert Einstein's theory of special
-%  relativity that showed that the increased relativistic mass ($m$) of a
-%  body comes from the energy of motion of the body—that is, its kinetic
-%  energy ($E$)—divided by the speed of light squared ($c^2$).
-%   \end{notation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-%
-% The acknowledgments must list:
-%
-% >>>>	A statement that indicates to the reader where the data
-% 	supporting the conclusions can be obtained (for example, in the
-% 	references, tables, supporting information, and other databases).
-%
-% 	All funding sources related to this work from all authors
-%
-% 	Any real or perceived financial conflicts of interests for any
-%	author
-%
-% 	Other affiliations for any author that may be perceived as
-% 	having a conflict of interest with respect to the results of this
-% 	paper.
-%
-%
-% It is also the appropriate place to thank colleagues and other contributors.
-% AGU does not normally allow dedications.
-
-\acknowledgments
-This work was funded by the OWP which is part of NOAA's National Weather Service.
+This work was funded by the Office of Water Prediction (OWP) which is part of the National Oceanic and Atmospheric Administration's (NOAA) National Weather Service (NWS).
 Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
 We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
-Additionally, David Blodgett from the USGS Water Mission Area was instrumental in helping define level paths and other hydrographic work.
+Additionally, David Blodgett from the United States Geolgical Survey (USGS) Water Mission Area was instrumental in helping define level paths and other hydrographic work.
 More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
 Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
-%The [type of data] data used for [brief context, description] in the study are available at [repository, source name] via [DOI, persistent identifier link] with  [license, access conditions] [optional in-text citation in References]
-%[Version number] of the [software name] used for [brief context, description of what the software was used for] is preserved at [DOI, persistent identifier link], available via [license type, access conditions] and developed openly at [software development platform link].* [optional in-text citation in References].
-
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% \bibliography{<name of your .bib file>} don't specify the file extension
-%
-% don't specify bibliography style
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\clearpage % this clears figures before references
-\bibliography{bibliography/owp_fim4_2022}
-%
-%Reference citation instructions and examples:
-%
-% Please use ONLY \cite and \citeA for reference citations.
-% \cite for parenthetical references
-% ...as shown in recent studies (Simpson et al., 2019)
-% \citeA for in-text citations
-% ...Simpson et al. (2019) have shown...
-%
-%
-%...as shown by \citeA{jskilby}.
-%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
-%...has been shown \cite{jskilbye}.
-%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
-%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
-%...has been shown by \cite <e.g.,>[and others]{lewin76}.
-%
-% apacite uses < > for prenotes and [ ] for postnotes
-% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
-%
-
-\end{document}
-
-%%%% More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-% If you are using line numbers, remember to surround
-% equations with \begin{linenomath*}...\end{linenomath*}
-
-%  To add line numbers to lines in equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}

### 2022-08-21 manuscripts/owp_fim4_2022/sections/conclusions.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

--- 2022-02-09 manuscripts/owp_fim4_2022/sections/conclusions.tex
@@
 Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
-The National Water Model (NWM) developed by the National Oceanic and Atmospheric Administration's Office of Water Prediction, along with partners, provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
-Additional processing is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
+The National Water Model (NWM), developed by the National Oceanic and Atmospheric Administration's (NOAA) Office of Water Prediction (OWP) and the National Center for Atmospheric Research (NCAR), provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
+Additional modeling is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
 Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
 HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
-Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
+Scalable, open-source software, known as OWP FIM, was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
 HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
-Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
+Finally, we used this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
+FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
+
+The main contribution and conclusion of this work centers around a fundamental limitation in HAND based FIM which is a failure to account for multiple possible sources of fluvial inundation since HAND only considers inundation from the nearest drainage line.
 We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
-This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
-FIM skill was evaluated over large spatial scales by comparsion to HEC-RAS 1D models.
-Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.
+In order to reduce stream order of the NWM stream network for HAND computation, we dissected the NWM network into two simpler units of singular Horton-Strahler stream order and mosaiced the resulting FIMs derived from each.
+The NWM Mainstems (MS) stream network, which covers roughly 4-5\% of the NWM Full Resolution (FR) network, spans all established forecasting points in the Advanced Hydrologic Prediction System (AHPS) and downstream reaches.
+The inundation from MS derived HAND is mosaiced together with the inundation of FR derived HAND.
+Extending order reduction to the entire network, the Generalized Mainstem (GMS) technique discretizes the NWM FR network into level paths (LP) of unit stream order for HAND computation.
+All LP based FIM derived from LP based HAND datasets are mosaiced together to form one seamless FIM.
+Dissecting the stream network into regions of LPs with unit stream order is necessary because HAND has a ``nearest drainage'' limitation meaning it only accounts for riverine inundation sourced from the nearest drainage line.
+In our evaluation of this technique, we observe that HAND based FIM improves in skill as we extend from nearest drainage inundation in FR to multiple drainage support in MS for only 4-5\% of the FR network.
+Extending multiple drainage support to the entire FR network with GMS based HAND improves upon skill at around the same magnitude that MS improved upon FR.
+Thus we conclude that deriving HAND with independent stream networks of unit Horton-Strahler stream order enhances the skill of FIM but offers diminishing returns as we extend from 4-5\% of the network with MS to 100\% of the network with GMS since deriving HAND and FIMs at these localized scale does add computational expense.
+
+This primary contribution also affects the parameters used to compute stage-discharge relationships shifting discharge higher at given stages which reduced the rate of false positives.
+This shift in SRC behavior is driven by larger catchments that influence reach averaged geometric parameters in the Manning's equation.
+Related to SRCs, we noted better results and more sensitivity to unit stream order networks with the higher Manning's n value of 0.12 when compared to 0.06 for high magnitude events at 1\% (100 year or yr) and 0.2\% (500 year or yr) recurrence intervals.
+Additionally, we noted better performance for more intense 500 yr events which we attribute to a stronger influence of poor quality bathymetric data in 100 yr magnitude inundation extents.
+While the AGREE DEM procedure is meant to add some bathymetry primarily motivated to enhance catchment and stream line delineation, it does introduce three parameters that have major implications in the quality of SRCs and the resulting FIMs.
 Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
+We leave questions related to Manning's n localization as well as bathymetry integration, estimation, and/or calibration open for future research to answer.
+Two other issues left open for improvement include the integration of higher resolution Lidar-based digital elevation maps (DEM) as well as the use of physics-based models for continental scale, high resolution forecasting applications.
 Due to inherent limitations with HAND, scalable, physics-based methods are necessary to consider to provide a better representation of flood extent dynamics in steady and unsteady conditions.
-%

### 2022-08-21 manuscripts/owp_fim4_2022/sections/discussion.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

--- 2022-01-10 manuscripts/owp_fim4_2021/sections/discussion.tex
@@
 %
-Overall, we note a positive relationship between FIM skill and a reduction of the stream order of the stream network we use to derive the HAND datasets.
-Most of this change is accounted for by increasing POD thus reducing FNs especially along higher order rivers with higher flow magnitudes.
-We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
-This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
+Overall, the main observation of this study was how FIM performance can be improved by reducing the Horton-Strahler stream order of the target stream network used for HAND computation.
+Most of this change is accounted for by substantially increasing POD and inundation extents in some areas thus reducing FNs.
+We believe, as we later argue for, that the increase in POD is primarily driven by an increase in the catchment sizes that is inherently enabled by dividing up the stream network into independent stream networks of unit stream order.
+Additionally, we note that reducing drainage order also has some minor influence on reducing inundation extents in other areas and the rate of false alarms.
+We believe that this effect is driven by a change in the stage-discharge relationship where a given streamflow now leads to lower river stage values when HAND is computed with target stream networks of unit drainage order.
+We seek to explain that these two effects, catchment boundary enlargement and stage-discharge curve lowering, are highly interrelated and cannot be easily detangled.
+Lastly, we discuss the diminishing effects on performance that the MS and GMS techniques may have and also any additional effects including enhanced cross-walking abilities.
 
-In analyzing the data, we found a slight reduction in FAR was detected and more digging pointed to a bias in rating curves introduced by stream order reduction.
-Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
-Sub-figure \ref{fig:rating_curve_comparison}a shows how the average rating curves for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to bias down (and to the right) with ever increasing stream order reduction (FR to MS to GMS).
+As evident in the results of the study in Section \ref{sec:results}, a sizable amount of the increase in CSI observed by reducing stream order for HAND computation can be attributed to increases in POD.
+This can be inferred by close inspection of Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} where changes in POD are significantly higher than that of FAR.
+%This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
+Upon investigation of the performance of HAND derived FIM, we note a general increase of catchment sizes for the MS and GMS methods when compared to the FR method as they are now delineated independently of any tributaries that would constrain catchment sizes otherwise.
+Additionally, we note significantly less water build up along catchment boundaries especially at the junction of lower order tributaries with lower flow rates and higher order rivers with more flow.
+This allows for inundation extents to expand across regions previously encapsulated by catchments of joining reaches in lower flow tributaries.
+The water built up along the catchment boundaries can be thought of as a column of water in a cylindrical container (catchment) that has exceeded the elevation of the containers rim.
+This, of course, does not represent accurate physics but it is what does occur in HAND when water accumulates along catchment boundaries.
+
+Large scale HUC8 level evaluations can fail to demonstrate fine grain enhancements as they aggregate away many changes that are only clear at more local scales.
+Future assessments of OWP FIM should consider finer grain evaluation units as well possible impact assessments using asset information such as building footprints to better illustrate fine grain changes in a more relevant manner to stakeholders.
+For now, we offer up Figure \ref{fig:gms_enhancement} which best illustrates the improvement offered by multi-source fluvial flooding capabilities in a more local context.
+The figure is comprised of two agreement rasters for two different HAND based FIMs compared to the validation dataset for a given region with a high flow mainstem (500 yr recurrence flow) running horizontally along the region.
+Sub-figure \ref{fig:gms_enhancement}a demonstrates the agreement raster for the FR stream network as well as its respective catchment boundary lines symbolized in white and stream network shown in green.
+Inspection of this sub-figure denotes a clear spatial pattern where TPs or areas correctly inundated are pooled alongside catchment boundary lines. 
+On the other side of the catchment boundary, one can witness large swaths of FNs that should be inundated. 
+The FNs also exhibit a spatial pattern as in they tend to collocate within catchments of the pictured mainstems tributaries.
+This sort of behavior was introduced early in the paper and shown qualitatively in Figure \ref{fig:catchment_boundaries_issue}.
+
+As an enhancement, this paper proposes computing HAND for stream networks comprised of level-paths independently of one another.
+In sub-figure \ref{fig:gms_enhancement}b, the agreement raster for the GMS technique is illustrated as well as the stream network lines in green.
+While the entire mosaiced inundation map from GMS (as described in Section \ref{ssec:inundation_mapping} and Equation \ref{eq:hand_fim_depth}) is used to produce this agreement map, we only show the catchments associated with the mainstem of this region that is shown to follow a clear horizontal path.
+Showing all the catchments for the tributaries that were all derived independently would convolute the image.
+The main message illustrated here is that the catchments associated with the mainstem of this area significantly increase in size.
+Since they are not restricted by the catchments of tributaries that lie in the same drainage areas as those of the mainstem, they extend to include those as well.
+The consequence for inundation extents is a general increase in spatial coverage of the river's water which shows to be in much better agreement with the benchmark map.
+The TPs are no longer bounded by the catchment lines and allowed the expand to their natural extents.
+
+We note here as a contribution of this study that a major inherent, limitation of HAND is the ``nearest drainage'' constraint or the idea that a given river reach only drains or, in HAND's case, inundates its immediate, unique drainage area or catchment.
+In other words, HAND based FIMs are limited to producing fluvial inundation to only their nearest drainage area or catchment.
+However, we know that fluvial inundation can be sourced from several streams nearby that also serve as drainage outlets to the area in question.
+Generally speaking, drainage areas are known to be hierarchical in nature so a given drainage area for a given outlet point can be decomposed into nested drainage areas for outlet points that lie in the original drainage area.
+A perfect example of this are points that lie on tributary reaches closely neighboring a mainstem.
+These points lie in the drainage area of reaches in the mainstem but inundation from the mainstem cannot reach these tributary catchments because of the ``nearest'' assumption in HAND.
+Hence it's important to state that just like there are different sources of flooding such as fluvial, pluvial, groundwater, reservoir, barrier failure (dam/levee/embankment), and coastal, there can also be multiple sources of a riverine flood.
+HAND is only equipped to handle riverine flooding from the nearest drainage line.
+Other relevant drainage lines that produce fluvial flood waters are not considered here especially if the routing model used doesn't consider backwater effects.
+
+The remaining portion of the improvement in CSI was found to come from a marginal yet notable reduction in FAR.
+Upon investigation of the spatial results in the agreement maps, we found some areas of slight reductions in FPs especially where changes in catchment boundaries may have occurred due to the reduction in effective stream order in computing HAND.
+These observations pointed to changes in the SRCs introduced by stream order reduction and catchment definition adjustments.
+Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on SRCs.
+Sub-figure \ref{fig:rating_curve_comparison}a shows how the average SRCs for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to shift the curve down and to the right with ever increasing stream order reduction (FR to MS to GMS).
 This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
-Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
+
+Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged SRC relationship between stage and discharge.
 Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
 While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
-Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius.
+Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius but these associations only hold when reach length, L, is considered.
 Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively (see Equation \ref{eq:reach_averaged_mannings_equation}).
-Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
+Figures \ref{fig:rating_curve_comparison}b and \ref{fig:rating_curve_comparison}c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
 Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
 Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
-We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
+We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the SRC Q up and changes the SRC accordingly.
 In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
 Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
+
 Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments that encompass neighboring tributaries.
-These tributaries have a significant amount of bathymetry that is low-lying thus easily including the SRC derivation. 
+These tributaries have a significant amount of bathymetry that is low-lying thus easily included in the geometry for the SRC derivation. 
 They also contribute volume and bed area that is technically not perpendicular to the flux of streamflow being accounted for in the stream in question. 
@@
 
+We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
+This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
+One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
+
 Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
-While the cause of these errors can be varied, we assert here that conflating 4 networks for use in evaluations leads to significant error.
+While the cause of these errors can be varied, we assert here that conflating four networks for use in evaluations leads to significant error.
 As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
-One of the issues is when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
-The utilization of MS and GMS only conflates to NWM catchments directly associated with the level path in question which is inherently easy to do with those methods. 
+One of the issues with the FR version of HAND occurs when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
+The utilization of MS and GMS only conflates to NWM catchments directly associated with the LP in question which is inherently easy to do with those methods. 
 Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
-The NWM stream network was derived using the NHD medium resolution dataset which was derived from coarser DEMs than those used here. 
+The NWM stream network was derived using the NHDPlus V2 dataset which was derived from coarser DEMs than those used here. 
 Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
@@
 Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
+We suspect that a significant amount of the difference in performance between 100 yr and 500 yr magnitude events can be attributed to poor SRC performance due to poor bathymetric representation.
+Lower magnitude events are, logically, more susceptible to poor bathymetric data due a greater proportion of the inundation being attributable to areas that are more typically under normal flow conditions.
+Higher flow events tend to cover regions with more floodplain inundation thus less sensitive to errors from bathymetric data quality.
+On a related note, the use of the AGREE DEM method discussed in Section \ref{ssec:stream_network_enforcment} also interacts with the bathymetry issue introducing several artificial geometry parameters that affect SRC shape and quality.
+Due to focus on the nearest drainage problem, we leave future work related to SRC representation including roughness estimation, bathymetric data assimilation, and bathymetry adjustments as opportunities for major enhancements in HAND based FIM.
 
@@
 These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
-Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
+Ultimately, the use of more advanced 2D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
@@
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
 \includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
-\caption{OWP FIM `Cahaba' inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105.
+\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105 at the 500 yr recurrence magnitude.
 Catchment boundaries and stream lines are shown in white and dotted green, respectively.

### 2022-08-21 manuscripts/owp_fim4_2022/sections/introduction.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

--- 2022-02-24 manuscripts/owp_fim4_2022/sections/introduction.tex
@@
 More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
-With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
-Trends related to flood damages and fatalities have been steadily increasing over recent decades.
-\cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
-Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
-Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.
+With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}, with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf Coast.
+Trends related to flood damages and fatalities have been steadily increasing over recent decades \cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
+Some are expecting that the hydrologic cycle will intensify due to climate change which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
+Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest and Great Lakes regions while decreasing in the coastal Southeast, Southwest, and California.
 %
@@
 %
-Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
+Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life threatening and property damaging events. 
 One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble stream flows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
-Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
-AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
-On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
-Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons to address the increasingly complex water challenges facing the US.
+AHPS provides forecasting services in the form of ensemble streamflows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
+Additionally, two forecasting networks, Full Resolution (FR) and Mainstems (MS) stream networks, relevant to the National Water Model (NWM) (see Section \ref{ssec:national_water_model}) are rendered in Figure \ref{fig:forecast_points}.
+The FR network refers to the entire NWM forecasting domain while MS refers to the subset of the FR network that is at or downstream of AHPS forecasting points (see Section \ref{ssec:national_water_model}).
+On an approximate basis, there is only one forecast point every 1,450 km of river (FR) and one forecast point with FIM every 29,000 km of river (FR).
+Despite the AHPS advances in operational flood forecasting, it lacks sufficient domain coverage, spatial resolution, and long-range forecast horizons to address the increasingly complex water challenges facing the US.
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
 \includegraphics[scale=1.0]{figures/forecast_points.jpg}
-\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
-Also shown are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
+\caption{Forecast points with and without Flood Inundation Maps (FIM) in United States' Advanced Hydrologic Prediction System (AHPS).
+Note that only a small fraction of the AHPS forecast points have existing FIM.
+Also shown are the National Water Model (NWM) stream networks at the Full Resolution (FR) and Mainstems (MS) resolution.
+The FR network constitutes the entire NWM stream network while the MS resolution network is the FR network at or downstream of the AHPS forecast points shown.
+}
 \label{fig:forecast_points}
@@
 %
-Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
-The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
+Additional work is required to address the gaps that the AHPS leaves in terms of spatial resolution, spatial coverage, and temporal forecast horizons.
+In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners at the National Center for Atmospheric Research (NCAR) have developed and implemented operationally the NWM which is a configuration of the Weather Research and Forecasting Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including lookback-range (3-28 hrs), short-range (18 hr), medium-range (10 day) and long-range (30 day) forecast horizons.
 The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
-As a complement to the operational NWM, OWP has also developed a configuration of the NWM that extends RFC forecasts downstream by assimilating and routing forecast streamflows to the next downstream AHPS forecast point.
-This configuration of the NWM is used to enhance forecasting skill by leveraging best available regional-scale data.
-The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
-Figure \ref{ssec:national_water_model} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
+As a complement to the operational NWM, RFC forecasts from AHPS forecast points are assimilated in the NWM and routed downstream to the next downstream AHPS forecast point where the process iterates again.
+This assimilation into NWM is used to enhance forecasting skill by leveraging best available regional-scale forecasts.
+The river network upon which this special assimilation technique operates on is herein referred to as the Mainstem (MS) stream network.
+Figure \ref{fig:forecast_points} shows the NWM V2.1 FR stream network as well as the NWM V2.1 MS network.
 The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
 
-The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
-The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
-Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
-As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
-The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
+The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the ``hydrofabric'' in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus,nhdplus2022vectors}.
+The term ``hydrofabric'' is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
+The NWM provides stream forecasts at these hydrofabric segments using the Muskingum-Cunge method to reduce computational requirements of a continental scale model but fails to consider backwater dynamics \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
+The need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
+The use of a 2-dimensional (2D) hydrodynamic model across a continental-scale and high spatial resolutions is very cost prohibitive especially in an operational setting.
+The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert 1-dimensional (1D) riverine discharges to stages, and finally to inundation extents and depths.
 %
@@
 %
-HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
+HAND normalizes topography along the nearest drainage path and it has been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
 \citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
-The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
-\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
+However, a core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
+The terrain index also provides additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
+\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as SRCs by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
 This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{HAND's Assumptions and Limitations}
+\label{sssec:hand_assumptions_limitations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+HAND operates on many underlying assumptions since it can only be used as an inundation proxy or no physics model and thus, not a true hydrodynamic inundation model \cite{nobre2016hand,liu2016cybergis,liu2020height}.
+HAND, to our knowledge, has only been applied to natural, inland, and riverine inundation applications thus it is also missing pluvial, coastal, ground water, and dam break components among other possible sources of flooding.
+Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest stream line which is used for catchment allocation and relative elevation calculation \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,maidment2017conceptual,garousi2019terrain,zheng2018river,zheng2018geoflood,johnson2019integrated,renno2008hand}.
+Stream thalweg networks must also collectively drain to a singular outlet point for a given processing region \cite{nobre2016hand,zheng2018geoflood,renno2008hand}.
+Since elevations don't naturally do this, they must undergo a long series of hydro-conditioning processes to enforce monotonically decreasing elevations across an entire processing unit along with hydrologically correct directions of flow \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,donchyts2016global,renno2008hand}.
+The level of DEM manipulation required to enforce this assumption can be substantial depending on the region and can be a significant source of error.
+The drainage enforcing assumption also interacts with an inability to properly account for fluvial inundation in regions of DEM depressions that lack natural drainage to riverine areas \cite{nobre2016hand,renno2008hand}.
+
+When used for FIM applications, HAND assumes only fluvial inundation sourced from its nearest drainage line is accounted for \cite{nobre2016hand,mcgehee2016modified}.
+Catchments are independent of one another for FIM purposes meaning a reaches' stage value is only used to threshold the HAND values within its respective catchment \cite{liu2016cybergis,zheng2018river,zheng2018geoflood}.
+This assumption plays to the ``Nearest Drainage'' term in HAND and creates a significant limitation within HAND for FIM applications \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+At the junction of high stream order and high flow rivers with lower flow tributaries, there can be a lack of inundation extents exhibited which is known colloquially in the forecasting community as the ``catchment boundary problem''.
+The academic community has somewhat referenced this issue before but it has been characterized more as a problem with the stream delineation process that comes from thresholding the drainage accumulation maps \cite{nobre2016hand,li2020evaluation}.
+Later in this study, we will re-introduce this problem and demonstrate how we initialize with a stream network (that of the NWM's) and thus avoid having to threshold accumulations to some arbitrary value to define stream networks.
+We illustrate how computing HAND independently for stream lines of unit stream order can significantly enhance FIM performance by accounting for multiple sources of fluvial inundation that may exist in certain regions and flow scenarios.
 %
@@
 %
-Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
-As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
+Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (approximately 10 m depending on latitude) scale, HAND has been implemented into software for large-scale, continental computation.
+As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a HPC \cite{maidment2017conceptual,liu2016cybergis}.
 NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
-By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
-Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC as an unofficial demonstrations to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
-Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
-More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
+By allocating 31 nodes at 20 cores per node for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
+Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (NHDPlusHR) which better agrees with the 10 m DEM from the NHDPlusHR program \cite{liu2020height}.
+The original NFIE dataset was employed by the NWC as an unofficial demonstration to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
+Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
+More notably the software added the ability to derive HAND on both the NWM FR and MS stream networks to consider multiple sources of fluvial inundation along high impact rivers of primary forecasting concern.
 
-Related to these efforts, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
+Related to these efforts, the United States Geological Survey (USGS) has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses SRCs with cross-sections for stage-discharge relationships \cite{verdin2016software}.
 Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{OWP FIM}
+\subsection{Office of Water Prediction Flood Inundation Mapping}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-Many of those in the academic community assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
-\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
-The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
-Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
-Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3 m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
-The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
-Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
-Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
-Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
-These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.
-
-With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
+In order to mitigate the ever increasing threat of flooding to life and property, an operational capability is required to extend NWM streamflow forecasts to river stages, inundation extents, and inundation depths.
+OWP FIM is introduced here as a continental scale capability that generates these products at high spatial and temporal resolutions.
 Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
-This framework is open-source and modular enabling and accelerating the research to operations development pipeline.
-Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
-In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
-Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
-This capability will be referred to as Generalized Mainstems (GMS).
+Within the operational framework of OWP FIM, we introduce research demonstrating how FIM performance skill with HAND can be improved by discretizing stream networks into units of an effective unit Horton-Strahler stream order \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} for HAND computation contexts.
+Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to the stream accumulation threshold which is closely related to stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
+Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths (LP), can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
+This capability is introduced progressively as MS (whose network represents about 4\% of FR network) and to a higher degree Generalized Mainstems (GMS) (covers entire FR network) which will be explained later on.
 The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.

### 2022-08-21 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

@@
 \hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement & \cite{nhdplus2022vectors} \\
+USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlusHR for hydro-enforcement. & \cite{nhdplus2022vectors} \\
 \hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display & \cite{nhdplus2022vectors} \\
+USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display. & \cite{nhdplus2022vectors} \\
 \hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters & \cite{nhdplus2022dems} \\
+USGS & NHDPlusHR DEM & DEM used for NHDPlusHR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters. & \cite{nhdplus2022dems} \\
 \hline
-USGS & NHDPlusHR WBD & Water Boundaries (WBD) or HUCs used for spatial processing units & \cite{nhdplus2022wbd} \\
+USGS & NHDPlusHR WBD & Water Boundaries (WBD) or HUCs used for spatial processing units. & \cite{nhdplus2022wbd} \\
 \hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. & \cite{nwm2022hydrofabric} \\
+NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting adapted from NHDPlus V2 NHDFlowline\_Network feature class. & \cite{nwm2022hydrofabric} \\
 \hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. & \cite{nwm2022hydrofabric} \\
+NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM adapted from NHDPlus V2 Catchment feature class. & \cite{nwm2022hydrofabric} \\
 \hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. & \cite{nwm2022hydrofabric} \\
+NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes adapted from NHDPlus V2 NHDWaterbody feature class. & \cite{nwm2022hydrofabric} \\
 \hline
-USACE & NLD & Levee database of locations and elevations & \cite{engineers2016national} \\
+USACE & NLD & Levee database of locations and elevations. & \cite{engineers2016national} \\
 \hline
@@
 D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
-Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
+The horizontal component of slope is defined as one for the four neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$ by means of the Pythagorean theorem.
+Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression-filled cells must be resolved.
 
-Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
+Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment} where flats are resolved by incrementing elevations iteratively.
 Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
 The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
+The established literature in this niche field of hydrology discusses how prevalent flats can be in given study areas and how difficult the problem is from both computational and hydrologic stand-points \cite{garbrecht1997assignment,tarboton2009generalized,tarboton2005terrain,survila2016scalable,barnes2014efficient,tesfa2011extraction,wallis2009parallel}.
 OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
@@
 %
-where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
-Q, V, and B are taken a specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
-All units are international given the 1 numerator above n.
-The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
+where Q is discharge at stage y, n is the Manning's n roughness coefficient, V is volume at y, S is channel slope, L is the along-flow reach length, and B is wetted bed area at y.
+Q, V, and B are taken at specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
+All units are international (SI) given the one in the numerator above n.
+The reach averaged method has been compared to rating curves from HEC-RAS and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
 The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the channel length.
+Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the reach length.
+While the AGREE DEM is subject to hydro-conditioning processes, it does introduce some notion of bathymetry estimation that the native DEMs lack while being sensitive to additional parameters that could yield further errors in the FIM.
+We leave this issue open in this study and elaborate on needs with respect to bathymetry and Manning's n values in the Discussion section (Section \ref{sec:discussion}).
 
-Setting of the Manning's n roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
+Setting of the Manning's n roughness coefficient has precedent in previous continental-scale FIM (CFIM) studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
 These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
+We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but not limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
+For now and for the purpose of this study, we examine the SRCs with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
+After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values (y in Eq.
+\ref{eq:reach_averaged_mannings_equation}) from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
 %
@@
 We seek to resolve this catchment boundary problem or nearest drainage limitation by discretizing the target stream network into stream networks of reduced, unit stream order to avoid the constraining of catchments by those belonging to lower order neighbors.
-By discretizing the network into stream networks of unit stream order (later defined as level paths), we remove the influence of neighboring catchments that constrain the inundation extent.
+By discretizing the network into stream networks of unit stream order (later defined as LP), we remove the influence of neighboring catchments that constrain the inundation extent.
 This creates much larger and overlapping catchments that can source fluvial inundation from multiple reaches as required by the given river stage at current flow conditions.
-We present two successive methods, National Weather Service MS (Section \ref{sssec:nws_mainstems}) and GMS (Section \ref{sssec:generalized_mainstems}), implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND by expanding the nearest drainage definition to increase potential inundation areas.
+We present two successive methods, National Weather Service MS (Section \ref{sssec:nws_mainstems}) and GMS (Section \ref{sssec:generalized_mainstems}), implemented that reduce the effective Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND by expanding the nearest drainage definition to increase potential inundation areas.
+
+To clarify the phrase ``reducing Horton-Strahler stream order'' used extensively in this paper, every FIM used in evaluation contains a flood extent sourced from every NWM forecast point in the given evaluation domain.
+What we do to reduce stream order is discretize the NWM FR network into different units of size, MS network (\ref{sssec:nws_mainstems}) and GMS LPs (\ref{sssec:generalized_mainstems}), that effectively reduce the HAND computation to independent networks of unit stream order.
+These independent HAND datasets are later used to produce FIM independently and mosaiced together (see Section \ref{ssec:inundation_mapping}).
+The inundation from the MS HAND is mosaiced with the inundation from FR HAND, while the inundation of each individual LP from GMS is mosaiced together.
+The Horton-Strahler stream order is only reduced for HAND computation purposes to reduce the negative effects of the nearest drainage limitation inherent to HAND.
 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Main-stems}
+\subsubsection{NWS Mainstems}
 \label{sssec:nws_mainstems}
@@
 Since MS only covers 4\% of the entire FR stream network, we sought to expand drainage order reduction techniques to all reaches within the NWM modeling domain.
-In order to do this, we discretized the NWM network into level paths which when considered individually have unit Horton-Strahler stream orders thus significantly reducing drainage density.
-Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-Starting at the outlet, a unique level path is propagated upstream. 
-At every confluence, the direction of maximum flow path length is sought to propagate the current level path identifier.
-For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
+In order to do this, we discretized the NWM network into LPs which when considered individually have unit Horton-Strahler stream orders.
+LPs group flowlines by maximizing the length of each flow path and minimizing the number of LP identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
+In order to derive LPs for the NWM FR stream network at the HUC8 processing area, we first compute arbolate sums which are defined as the cumulative drainage distance of all upstream drainage lines.
+Arbolate sum is also inclusive of the current drainage reach as well.
+Arbolate sums are computed by starting at the headwater points and summing up drainage distances as you traverse downstream.
 
-Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
+Arbolate sum is critical to discretizing the NWM network into LP identifiers.
+Starting at a HUC8's outlet, a unique LP is propagated upstream. 
+At every confluence, the direction of maximum arbolate sum is sought to propagate the current LP identifier.
+For the remaining parent reaches of the given junction, a new LP identifier is assigned and the process recursively continues with them.
+Figure \ref{fig:level_path_methods} illustrates how LPs (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
+The figure shows computed arbolate sums and unique LP identifiers for the HUC12 120903010404 for clarity but were computed at the corresponding HUC8.
+The mainstem of the figure runs from the red ellipses to the black one which is the outlet.
+From the figure, we can see how unique colors are propagated in the direction of the maximum arbolate sum.
+
+Each HUC8 is discretized into LPs independently and the relevant inputs as described in Table \ref{tab:data} are assigned to each LP processing unit given a buffer of seven km.
 This buffer was selected to avoid edge contamination \cite{lindsay2013measuring} and to ensure adequate data availability for wide rivers with large catchments in regions with low slope.
-Further work could be dedicated to tune this user exposed parameter to better balance its effect on FIM extents and computational expense since larger buffers create additional floating points and calculations to process and store.
-At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
-Once the NWM FR stream network is discretized into level paths, we independently compute HAND using each level path as the target stream network to be used.
+Further work could be dedicated to tune this user exposed parameter to better balance its effect on FIM extents and computational expense since larger buffers create additional floating point calculations and storage requirements.
+For the time being, we designate this issue to be out of scope.
 
+At the LP scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the LP in question at the time.
+The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement by burning these lines and seeding from its headwater points directly instead of going through the NHDPlusHR network as described in Section \ref{ssec:stream_network_enforcment} .
+This decision was motivated by the difficulty in deriving LPs in the NWM stream network with high agreement with the LPs derived for the NHDPlusHR stream lines.
+We found that the same algorithm to compute arbolate sums and LPs could yield enough disagreements associated with disordered branches or slight differences in arbolate sums that could significantly affect the agreement of the LP identifiers in the NWM and NHDPlusHR networks.
+This yielded enough error to justify the use of the NWM directly for hydro-enforcement operations.
+
+Once the NWM FR stream network is discretized into LPs, we independently compute HAND using each LP as the target stream network to be used.
 To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
-In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
+In Figure \ref{fig:gms_methods}a, we uniquely color code the LPs derived for the NWM stream network. 
 For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each level path is buffered to a polygon with a user-available distance parameter of 7 km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
-We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
+Each LP is buffered to a polygon with a user-exposed, distance parameter of seven km that is used to subset the original DEM for two selected LPs in Figure \ref{fig:gms_methods}b.
+We illustrate two HAND grids for two of the LPs in this HUC8 in Figure \ref{fig:gms_methods}c.
+Once the FIM hydrofabrics for each LP are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
 Lastly, these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}d.
@@
 Possible benchmark FIM candidates for evaluation purposes include high-water marks, remote sensing observation, crowd-sourced information, and modeled extents.
-Unfortunately, all of these sources or using a combination of them are subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to lack of spatial coverage, signal interference, lack of stream flow data, inaccurate stream flow data, physics-based assumptions, and errors in input data.
-While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated stream flow data necessary to make FIMs to compare to.
+Unfortunately, all of these sources or using a combination of them are subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
+While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to isolating out hydrological factors.
 A suitable dataset needed to be selected to best balance advantages and limitations with future studies interested in expanding to additional datasets as able to.
 
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
-This dataset was selected due to its large spatial coverage, availability of cross-sections with stream flow information, more sophistication when compared to HAND based FIMs, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
-49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM so we used those corresponding discharges and mapping extents for evaluation.
-We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
+Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
+This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, more sophistication when compared to HAND based FIMs, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
+49 HUC8s spanning about 185 thousand $km^2$ were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
+The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM and thus corresponding discharges and mapping extents for evaluation.
+We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation sourced from OWP FIM.
 
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
+By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, and others that we would have potentially affected our conclusions if we used NWM forecasted discharges.
 Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.

### 2022-08-21 manuscripts/owp_fim4_2022/sections/openresearch.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

--- 2022-02-14 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-%
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements, including customized APA reference formatting.
-%
-% You may copy this file and give it your
-% article name, and enter your text.
-%
-%
-% Step 1: Set the \documentclass
-%
-%
-
-%% To submit your paper:
-\documentclass[draft]{dependencies/agujournal2019}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage[inline]{dependencies/trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-
-%% added
-\usepackage{multirow}
-\usepackage{amsmath}
-
-\linenumbers
-
-%%%%%%%
-% As of 2018 we recommend use of the TrackChanges package to mark revisions.
-% Add track changes editors
-\addeditor{Fernando Aristizabal}
-
-% The trackchanges package adds five new LaTeX commands:
-%
-%  \note[editor]{The note}
-%  \annote[editor]{Text to annotate}{The note}
-%  \add[editor]{Text to add}
-%  \remove[editor]{Text to remove}
-%  \change[editor]{Text to remove}{Text to add}
-%
-% complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%
-
-\drafttrue
-
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-% ie, \journalname{Water Resources Research}
-
-\journalname{Water Resources Research}
-
-% starts document
-\begin{document}
-
-%% Title and Authors
-%% ------------------------------------------------------------------------ %%
-%  Title
-%
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
-
-% Example: \title{This is a test title}
-%
-\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
-
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-%
-%
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
-%
-%
-\affiliation{1}{Lynker, Leesburg, VA, USA}
-\affiliation{2}{Center for Remote Sensing, Agricultural and Biological Engineering Department, University of Florida, Gainesville, FL, USA}
-\affiliation{3}{National Water Center, Office of Water Prediction, National Weather Service, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
-\affiliation{4}{New York Water Science Center, Hydrologic Applied Innovations Lab, United States Geological Survey, Troy, NY, USA}
-\affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Maryland, College Park, MD, USA}
-%
-%(repeat as many times as is necessary)
-%
-%% Corresponding Author:
-% Corresponding author mailing address and e-mail address:
-
-% (include name and email addresses of the corresponding author.  More
-% than one corresponding author is allowed in this LaTeX file and for
-% publication; but only one corresponding author is allowed in our
-% editorial system.)
-%
-% Example: \correspondingauthor{First and Last Name}{email@address.edu}
-%
-\correspondingauthor{Fernando Aristizabal}{fernando.aristizabal@noaa.gov}
-
-%% Keypoints
-%% Keypoints, final entry on title page.
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-
-% Example:
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-% \end{keypoints}
-%
-\begin{keypoints}
-\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
-\item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
-\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation.
-\end{keypoints}
-%
-
-%% Abstract
-% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT and PLAIN LANGUAGE SUMMARY
-%
-% A good Abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
-
-% The Plain Language Summary should be written for a broad audience,
-% including journalists and the science-interested public, that will not have 
-% a background in your field.
-%
-% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
-% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
-% see http://sharingscience.agu.org/creating-plain-language-summary/)
-%
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
-
-\begin{abstract}
-The National Water Model (NWM) currently requires the post-processing of forecast discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
-Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is worthy of producing high-resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
-However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
-To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that reduces the Horton-Strahler stream order of the stream network.
-GMS contains all segments within the NWM stream network but instead of deriving HAND by accounting for all river segments at once, it is derived independently at the level path (LP) scale.
-LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
-These FIMs are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
-Improvement in mapping skill is observed when compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
-A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
-\end{abstract}
-%
-\section*{Plain Language Summary}
-Flooding is one of the most impactful natural disasters on life and property.
-The United States National Water Model (NWM) provides flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
-In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
-We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-We compared these HAND derived FIMs to maps from more realistic models and found improvement in mapping performance.
-%
-
-%%% Suggested section heads:
-% \section{Introduction}
-%
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
-
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-%
-% \section{Data} (Or section title might be a descriptive heading about data)
-%
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-%
-% \section{Conclusions}
-
-% Introduction
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
-In 2017 and 2019, river and flash flooding combined represented the leading cause of death and the second leading cause in 2018 among all natural disasters in the US \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
-More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
-With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
-Trends related to flood damages and fatalities have been steadily increasing over recent decades.
-\cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
-Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
-Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Operational Forecasting}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
-One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble stream flows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
-Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
-AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
-On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
-Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons to address the increasingly complex water challenges facing the US.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/forecast_points.jpg}
-\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
-Also show are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
-\label{fig:forecast_points}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{National Water Model}
-\label{ssec:national_water_model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
-The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
-The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
-As a complement to the operational NWM, OWP has also developed a configuration of the NWM that extends RFC forecasts downstream by assimilating and routing forecast streamflows to the next downstream AHPS forecast point.
-This configuration of the NWM is used to enhance forecasting skill by leveraging best available regional-scale data.
-The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
-Figure \ref{ssec:national_water_model} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
-The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
-
-The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
-The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
-Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
-As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
-The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Height Above Nearest Drainage}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
-\citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
-The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
-\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
-This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{HAND Implementations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
-As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
-NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
-By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
-Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC as an unofficial demonstrations to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
-Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
-More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
-
-Related to these efforts, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
-Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
-In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30 m resolutions \cite{yamazaki2019merit,donchyts2016global}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{OWP FIM}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Many of those in the academic community assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
-\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
-The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
-Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
-Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3 m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
-The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
-Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
-Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
-Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
-These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.
-
-With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
-Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
-This framework is open-source and modular enabling and accelerating the research to operations development pipeline.
-Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
-In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
-Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
-This capability will be referred to as Generalized Mainstems (GMS).
-The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
-%
- %
-% Materials and Methods
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
+\openresearch
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Software Dependencies and Architecture}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
-Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
-A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub \cite{inundationMapping2022}.
+National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
+OWP Flood Inundation Mapping (FIM) capabilities rely extensively on the use of the National Hydrography Plus High Resolution (NHDPlusHR) datasets including BurnLineEvents \cite{nhdplus2022vectors}, value-added attributes (VAA) \cite{nhdplus2022vectors}, water boundaries (WBD) or hydrologic unit code (HUC) geometries \cite{nhdplus2022wbd}, and digital elevation maps (DEM) \cite{nhdplus2022dems}.
+Some additional datasets for processing include the National Levee Database (NLD) furnished by the United States Army Core of Engineers (USACE) \cite{engineers2016national}, Land-sea border from the Great Lakes Hydrography Dataset (GLHD) furnished by the Great Lakes Aquatic Habitat Framework (GLAHF) \cite{GreatLakesHydrographyDataset}, and a Land-sea border provided by OpenStreetMap (OSM) \cite{osm2021landsea}.
+Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
+Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
 
-The pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
-Preprocessing acquires and prepares datasets for production of the FIM hydrofabric. 
-The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
-Functionality is included to turn FIM hydrofabric and streamflows into inundation maps represented in both binary and depths.
-Lastly, a test suite includes means of calculating evaluation metrics compared to a variety of preprocessed test case data.
-A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}
-\label{ssec:datasets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP) \cite{esipData2022}.
-The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
-For enforcing levee data into the DEMs, the USACE National Levee Database (NLD) is used to burn feature elevations \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{OpenStreetMap,GreatLakesHydrographyDataset}.
-Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
-These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
-The full input datasets presented by source are listed in Table \ref{tab:data}.
-%
-\begin{table}
-\caption{Data sources, names, and descriptions.}
-\label{tab:data}
-\centering
-\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
-%\begin{tabular}{l c c}
-\hline
-Source & Name & Description \\
-\hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
-\hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
-\hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters \\
-\hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. \\
-\hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. \\
-\hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. \\
-\hline
-USACE & NLD & Levee database of locations and elevations  \\
-\hline
-OSM & Land-Sea Border & Border of land and sea. \\
-\hline
-GLHD & Land-Great Lakes Border & Border of land and Great Lakes. \\
-\hline
-InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
-\hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
-\hline
-\end{tabular}
-\end{table}
-%
-Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and GMS networks and 1,604 HUC8s for the MS method. 
-These methods will be explained more later.
-An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hydro-conditioning}
-\label{ssec:hydro_conditioning}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
-These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
-HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
-In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligble for flooding.
-Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
-It can also be used to simulate more accurate bathymetry which is not accounted for in the 10 m DEM \cite{gesch2002national}.
-
-Specifically within the context of OWP FIM, the hydro-conditioning operations that take place in sequential order are presented. 
-Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing units of size HUC8. 
-The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reach code's membership in the processing unit.
-Hydro-conditioning raster operations take place on buffered boundary definitions to avoid edge contamination and effects \cite{lindsay2013measuring}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Network Enforcement} 
-\label{ssec:stream_network_enforcment}
-%
-The location of the stream network is enforced to ensure general agreement with established stream networks.
-The NHDPlusHR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
-However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-For every NWM headwater segment a headwater point is derived and linearly interpolated to the nearest Burnline Event segment.
-Burnline Event headwater segments are split at the adjusted headwater point to match NWM flowlines.
-The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection and linear interpolation.
-Resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR Burnline Events.}
-\label{fig:stream_density_pruning}
-\end{figure}
-%
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.
-
-The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
-The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
-Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
-Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
-Some of these drawbacks are later addressed by additional conditioning techniques later on.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Levee Enforcement}
-%
-The DEM at 10 m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
-In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10 m DEM.
-The elevations found in the NLD are burned into the DEM if those elevations were found to exceed those already in place.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Depression Filling}
-\label{sssec:depression_filling}
-%
-Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
-The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
-Priority-Flood developed by \citeA{barnes2014priority} is an algorithm for filling said depressions and shown to have improved performance over early works in the field by \citeA{jenson1988extracting} implemented in \citeA{tarboton2005terrain} as well as \citeA{planchon2002fast}.
-The depression filling algorithm used in our pipeline is a Priority-Flood variant developed by \cite{zhou2016efficient} with enhanced single-thread performance and a time complexity of O(n log n) for floating point grids.
-This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
-The depression technique employed here does leave the existence of flat regions where pits existed a prior thus later requiring the need for resolving these flats.
-The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Thalweg Elevation Conditioning}
-\label{sssec:stream_thalweg_elevation_conditioning}
-%
-Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
-In order to prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
-Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50 m.
-The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
-
-The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
-\citeA{garousi2019terrain} proposed an algorithm that traverses stream thalweg pixels in a depth first manner starting with adding all the headwater pixels to a queue. 
-The connectivity of the thalweg pixels is defined by the D-8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
-At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in Equation \ref{eq:thalweg_breach},
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:thalweg_breach}
-\textbf{D}[x] = \min_{y\ drains\ to\ x} {(\ \textbf{D}[x]\ ,\ \textbf{D}[y]\ )}
-\end{equation}
-\end{linenomath*}
-%
-, in which \textbf{D} represents the array of thalweg adjusted elevations indexed by x and y where by y is upstream of x. 
-When a pixel's upstream neighbors are all evaluated, the downstream pixel is added into the queue thus the depth first traversal of the drainage network.
-This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in Critical Success Index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Deriving FIM Hydrofabric}
-\label{ssec:deriving_fim_hydrofabric}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Flow Directions and Flats Resolution}
-\label{ssec:flow_direction_and_flat_resolution}
-%
-To facilitate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D-8 flow directions.
-D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
-The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southeast direction as 8. 
-Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
-
-Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
-Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
-The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
-Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
-By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Deriving FIM Stream Network}
-\label{sssec:deriving_fim_stream_network}
-%
-The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
-The FIM stream network is of similar drainage density as the NWM V2.1 network but fully converges at all junctions leaving no divergences in the network.
-This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
-Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
-Each stream link in this derived FIM stream network is split into equidistant reaches.
-Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments.
-Stream links are then further segmented at NWM lakes and HUC8 boundaries.
-Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretation of the stream network.
-Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherent the synthetic rating curve properties of the upstream or downstream segment, depending on the topology.
-Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
-Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
-This stream network is important since it drives the HAND calculation and derivation of catchments.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Catchments}
-\label{sssec:catchments}
-%
-Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
-Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
-The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
-Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
-The other catchments are derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Height Above Nearest Drainage}
-%
-Once the pixel level catchments are derived, the final relative elevations can be computed.
-Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment described in Section \ref{sssec:catchments}.
-The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
-Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Synthetic Rating Curves}
-\label{sssec:synthetic_rating_curve}
-%
- A method for converting forecast river discharges from the NWM to stages or river depths is necessary for producing FIMs with HAND. 
-For one-dimensional models such as the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a synthetic rating curve at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
-For this application, we utilized the reach averaged approach for developing synthetic rating curves (SRC) \cite{zheng2018river}.
-The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-The reach averaged Manning's formula is derived to be 
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:reach_averaged_mannings_equation}
-\textbf{Q} = \frac{1}{n} \frac{V^{5/3}S^{1/2}}{L B^{2/3}} 
-\end{equation}
-\end{linenomath*}
-%
-where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
-Q, V, and B are taken a specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
-All units are international given the 1 numerator above n.
-The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
-The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the channel length.
-
-Setting of the Manning's n roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
-These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cross-walking with NWM Stream Network}
-\label{sssec:cross_walking_networks}
-%
-The stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with a NWM reach identifier so that a discharge can be converted to stage and later inundation extent.
-For the methods already discussed, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent methods discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
-Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
-While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
-More discussion on this will follow in Section \ref{sec:discussion}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stream Order Reduction}
-\label{ssec:stream_order_reduction}
-%
-FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
-Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
-An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the MS of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
-This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the MS that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the MS catchments.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{False negatives associated with confluence of tributaries with MS. Integers represent flow values from BLE 100 year event for the associated areas. 
-No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
-\label{fig:catchment_boundaries_issue}
-\end{figure}
-%
-We seek to resolve this problem by deriving HAND for processing units with stream networks of reduced stream order. 
-We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
-The resulting FIMs from the overlapping HAND processing units are mosaiced together taking any inundated area to be inundated but more will be explained in Section \ref{ssec:inundation_mapping}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Main-stems}
-\label{sssec:nws_mainstems}
-%
-The Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
-The MS network comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
-It also spans 121,724 reaches across 1,608 HUC8s.
-HAND was originally derived for this stream network to enhance mapping skill along these critical MS segments \cite{djokic2019arc}. 
-The inundation derived from this stream network is mosaiced with the inundation from the FR network to form the MS FIMs. 
-Within each HUC, you'll typically only find a MS stream network of stream order 1 (i.e.
-headwater) but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Generalized Mainstems}
-\label{sssec:generalized_mainstems}
-%
-To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call GMS.
-Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
-Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-Starting at the outlet, a unique level path is propagated upstream. 
-At every confluence, the direction of maximum flow path length is sought to propagate the current level path identifier.
-For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
-Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
-At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
-
-To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
-In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
-For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each level path is buffered to a polygon with a user-available distance parameter of 7 km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
-We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
-Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}c.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates how level paths for the NWM are derived.
-Level paths symbolized by lines of unique colors are propagated upstream following the direction of maximum arbolate sum at each junction.
-}
-\label{fig:level_path_methods}
-\end{figure}
-%
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/gms_methods.jpg}
-\caption{Overall procedure for GMS HAND.
-In (a), we illustrate all NWM stream lines symbolized by their level path.
-Meanwhile (b), demonstrates the DEM clipped to a 7 km buffer around two selected level paths.
-In (c), we show how HAND can be computed just for each one of these two level paths independently. 
-We also show inundation maps created for these two level paths in (c). 
-In (d), we show all the inundation maps for all the level paths mosaiced together. }
-\label{fig:gms_methods}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Inundation Mapping}
-\label{ssec:inundation_mapping}
-%
-The FIM hydrofabric consisting of the relative elevations grid, catchments grid, catchment polygons, rating curve, and cross-walking data are all used to convert forecasts from the NWM into forecasts extents.
-For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
-From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 8 spatial processing units in a rapid, parallel operation.
-The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
-
-Utilizing the stage-discharge relationships in the synthetic rating curves, each forecast for each catchment identifier is assigned a stage value. 
-The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
-We use the basic logic already established in previous works to conduct this \cite{nobre2016hand,liu2016cybergis,maidment2017conceptual}.
-Mathematically, the HAND values, $H_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
-For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
-For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim_depth}
-    D_{ij} = S_i - H_{ij}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim}
-    F_{ij} = D_{ij} > 0
-\end{equation}
-\end{linenomath*}
-%
-For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast in the form of a single raster file.
-For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
-Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
-Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim_depths}
-    D_{ij} = \max_{k=[1,...,K]} D_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-For the MS and GMS methods, the contributing areas are defined differently.
-For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to 2 for that case.
-For GMS, all FIMs from all the level paths in a given area are mosaiced together then K is set to this number of level paths.
-Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim}
-    F_{ij} = \max_{k=[1,...,K]} F_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}
-\label{ssec:evaluation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
-49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM so we used those corresponding discharges and mapping extents for evaluation.
-We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
-Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
-We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
-To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
-This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
-While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
-\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states. This dataset for 1\% and 0.2\% recurrence flows were used as benchmarks.}
-\label{fig:all_ble_maps}
-\end{figure}
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
-\caption{Illustrates Base Level Engineering (BLE) cross sections and stream lines.
-The BLE stream network, which is denser than the NWM V2.1 stream lines, is also shown.
-BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM reach.
-}
-\label{fig:ble_evaluation_method}
-\end{figure}
-%
-The metrics employed in this study to evaluate inundation extents include Critical Success Index (CSI), Probability of Detection (POD), and False Alarm Ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
-To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.
-The two types of errors consist of false positives (FP), or type I errors, which is dry in benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in benchmark by predicted dry.
-Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
-Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
-While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
-Thus, frequency dependent statistics should be used with caution when comparing across sites with varying frequencies. 
-Lastly, approximately 6 HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:csi}
-CSI = \frac{TP}{TP + FN + FP}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:pod}
-POD = \frac{TP}{TP + FN}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:far}
-FAR = \frac{FP}{TP + FP}
-\end{equation}
-\end{linenomath*}
-%
- %
-% Results
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{sec:results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Mapping Performance}
-\label{ssec:mapping_performance}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-We produced FIMs for the entire BLE domain within the 49 HUC8 areas across several states in the south central US. 
-The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
-We computed the statistics (CSI, POD, and FAR) for both 100 and 500 year events for Mannings N set to 0.06 and 0.12. 
-The distribution of these statistics can be examined in Figure \ref{fig:violin_plot} as violin plots.
-Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
-We also denote trend lines for each metric and Manning's n setting as well as their respective slope estimate and one-tailed p-value denoting the level of significance of the trend.
-
-Aggregating the metrics in the method above treats each HUC8 as it's own unit and does little to consider the size differences of the HUCs. 
-In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics}  the CSI, POD, and FAR recomputed for the entire domain using the sum of all the TPs, FPs, and FNs. 
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.9]{figures/violin_plots.jpg}
-\caption{Shows kernel density estimation of the distributions (sample size = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
-The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's n (0.06 and 0.12).
-Trend lines for each metric - Mannings combination are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}
-\label{fig:violin_plot}
-\end{figure}
-%
-\begin{table}[h!]
-\caption{Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated for BLE domain.
-         The best value across models is highlighted in bold.}
-\label{tab:aggregate_metrics}
-\centering
-%\begin{tabular}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
-\begin{tabular}{|c|c||c|c|c|c|c|c|}
-\hline
-\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{MS} & \multicolumn{2}{|c|}{GMS} \\
-\cline{3-8}
-  &  & 100yr & 500yr & 100yr & 500yr & 100yr & 500yr \\
-\hline
-\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & \textbf{0.5796} & \textbf{0.6075} \\
-\cline{2-8}
-  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & \textbf{0.6182} & \textbf{0.6435} \\
-\hline
-\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & \textbf{0.6633} & \textbf{0.6863} \\
-\cline{2-8}
-  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & \textbf{0.7606} & \textbf{0.7810} \\
-\hline
-\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1787 & \textbf{0.1589} & \textbf{0.1778} & \textbf{0.1589} \\
-\cline{2-8}
-  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & \textbf{0.2324} & \textbf{0.2148} \\
-\hline
-\end{tabular}
-\end{table}
-%
-% Interpretation of metrics
-From Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, we denote several meaningful trends. 
-Using CSI as an overall proxy for skill of the FIM, we note that generally speaking the skill is correlated with a reduction of the stream orders of the processing units used for HAND.
-In other words, the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
-While this trend is evident for both sets of Manning's n values, the trend is slightly more significant for the higher value of 0.12.
-Other trends related to this Figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
-We also note how the higher Manning's n value enhances performance for both of these recurrence intervals across all models.
-
-Dissecting the improvements and trends presented in the previous paragraph comes down mostly to improvement in POD or a reduction in absolute amount of FNs.
-POD being the primary driver in skill enhancement is evident across models by comparing the slope of the POD lines with the slope of the FAR lines.
-Even though aggregating metrics by HUC8 yields a statistically zero trend, one does see a slight reduction in FAR across models that reduce HAND's maximum stream order.
-Additionally, we note that POD is a primary driver in enhancing performance across Manning's n values as well.
-This significant improvement comes at a cost of false alarms as the FAR increases significantly across Manning's n values.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Computational Performance}
-\label{ssec:compuational_performance}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The NFIE experiments were able to produce HAND for 331 HUC6's in 1.34 CPU years \cite{liu2016cybergis} and estimates using work from \citeA{djokic2019arc} put producing HAND at the FR NWM at 0.55 CPU years.
-For our work, we were able to produce HAND at the full NWM resolution in 0.13 CPU years which represents a substantial speed-up compared to previous works.
-For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
-GMS which generalizes HAND production to level path scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
-%
- %
-% Discussion
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-\label{sec:discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Overall, we note a positive relationship between FIM skill and a reduction of the stream order of the stream network we use to derive the HAND datasets.
-Most of this change is accounted for by increasing POD thus reducing FNs especially along higher order rivers with higher flow magnitudes.
-We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
-This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
-
-In analyzing the data, we found a slight reduction in FAR was detected and more digging pointed to a bias in rating curves introduced by stream order reduction.
-Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
-Sub-figure \ref{fig:rating_curve_comparison}a shows how the average rating curves for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to bias down (and to the right) with ever increasing stream order reduction (FR to MS to GMS).
-This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
-Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
-Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
-While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
-Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius.
-Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively (see Equation \ref{eq:reach_averaged_mannings_equation}).
-Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
-Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
-Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
-We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
-In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
-Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
-Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments that encompass neighboring tributaries.
-These tributaries have a significant amount of bathymetry that is low-lying thus easily including the SRC derivation. 
-They also contribute volume and bed area that is technically not perpendicular to the flux of streamflow being accounted for in the stream in question. 
-Careful examination of Figure \ref{fig:gms_enhancement}b shows how much larger catchments include neighboring tributaries and the geometry associated with those tributaries. 
-This geometry is not perpendicular to the flow that is associated with the main reach thus leading to biases in the SRC.
-We consider this fact to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
-
-Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
-While the cause of these errors can be varied, we assert here that conflating 4 networks for use in evaluations leads to significant error.
-As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
-One of the issues is when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
-The utilization of MS and GMS only conflates to NWM catchments directly associated with the level path in question which is inherently easy to do with those methods. 
-Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
-The NWM stream network was derived using the NHD medium resolution dataset which was derived from coarser DEMs than those used here. 
-Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
-Until a singular stream network is used for the NWM, BLE benchmark, and for HAND based FIM, conflation will continue being a source of error.
-
-Our qualitative analysis suggests that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM for future development.
-The bathymetry of the 10 m DEM from 3DEP is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
-Manning's n which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
-Some adjusting parameter must be introduced to enhance the estimation of the bathymetric representation.
-Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
-
-Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
-For HAND to be derived and thus create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
-In other words, an entire area eligible for flooding must monotonically decrease in elevation. 
-DEM's naturally don't do this and the dynamics of true flood events don't follow drainage patterns.
-Enforcing this assumption for HAND leads to significant amount of DEM manipulations that introduce basic errors.
-These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
-Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/rating_curve_comparison.jpg}
-\caption{Illustrates average quantities for the three methods, FR, MS, and GMS, for each stage value (m). 
-The values are (a) Discharge $m^3s^{-1}$, (b) Volume $m^3$, (c) Bed Area $m^2$, (d) a function of Volume and Bed Area, and (e) number of pixels.
-}
-\label{fig:rating_curve_comparison}
-\end{figure}
-%
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
-\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105.
-Catchment boundaries and stream lines are shown in white and dotted green, respectively.
-Sub-figure (a) shows agreement of FR HAND denoting significant areas of under-prediction due to junctions and catchment boundaries.
-Meanwhile, (b) shows the agreement for GMS and much larger catchments leading to much better inundation agreement for this given reach. 
-Overall, this illustrates the benefits of stream order reduction for deriving HAND datasets.
-}
-\label{fig:gms_enhancement}
-\end{figure}
-%
- %
-% Conclusions
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-\label{sec:conclusions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Floods present a significant, under-served, and expanding risk to life, property, and resources.
-Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
-The National Water Model (NWM) developed by the National Oceanic and Atmospheric Administration's Office of Water Prediction, along with partners, provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
-Additional processing is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
-Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
-HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
-Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
-HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
-Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
-We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
-This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
-FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
-Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.
-Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
-Due to inherent limitations with HAND, scalable, physics-based methods are necessary to consider to provide a better representation of flood extent dynamics in steady and unsteady conditions. 
-%
- %
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-%\section{= enter section title =}
-%Text here ===>>>
-%\section{= enter section title =}
-%Text here ===>>>
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables;  other caption information
-%  should be placed in last line of the table, using
-% \multicolumn2l{$^a$ This is a table note.}
-%
-%----------------
-% EXAMPLE FIGURES
-%
-% \begin{figure}
-% \includegraphics{example.png}
-% \caption{caption}
-% \end{figure}
-%
-% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
-% \begin{figure}
-% \noindent\includegraphics[width=\textwidth]{anothersample.png}
-%\caption{caption}
-%\label{pngfiguresample}
-%\end{figure}
-%
-%
-% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
-This is common when trying to add a PDF figure without pdflatex.
-% \begin{figure}
-% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
-%\caption{caption}
-%\label{pdffiguresample}
-%\end{figure}
-%
-%
-% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
-%
-
-%
-% ---------------
-% EXAMPLE TABLE
-%
-% \begin{table}
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \centering
-% \begin{tabular}{l c}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Footnote text here.}
-% \end{tabular}
-% \end{table}
-
-%% SIDEWAYS FIGURE and TABLE
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{figsamp}
-% \caption{caption here}
-% \label{newfig}
-% \end{sidewaysfigure}
-%
-%  \begin{sidewaystable}
-%  \caption{Caption here}
-% \label{tab:signif_gap_clos}
-%  \begin{tabular}{ccc}
-% one&two&three\\
-% four&five&six
-%  \end{tabular}
-%  \end{sidewaystable}
-
-%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
-%\begin{linenomath*}
-%\begin{equation}
-%y|{f} \sim g(m, \sigma),
-%\end{equation}
-%\end{linenomath*}
-
-%%% End of body of article
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary, Notation or Acronym section goes here:
-%
-%%%%%%%%%%%%%%
-% Glossary is only allowed in Reviews of Geophysics
-%  \begin{glossary}
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \end{glossary}
-
-%
-%%%%%%%%%%%%%%
-% Acronyms
-%   \begin{acronyms}
-%   \acro{Acronym}
-%   Definition here
-%   \acro{EMOS}
-%   Ensemble model output statistics
-%   \acro{ECMWF}
-%   Centre for Medium-Range Weather Forecasts
-%   \end{acronyms}
-
-%
-%%%%%%%%%%%%%%
-% Notation
-%   \begin{notation}
-%   \notation{$a+b$} Notation Definition here
-%   \notation{$e=mc^2$}
-%   Equation in German-born physicist Albert Einstein's theory of special
-%  relativity that showed that the increased relativistic mass ($m$) of a
-%  body comes from the energy of motion of the body—that is, its kinetic
-%  energy ($E$)—divided by the speed of light squared ($c^2$).
-%   \end{notation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-%
-% The acknowledgments must list:
-%
-% >>>>	A statement that indicates to the reader where the data
-% 	supporting the conclusions can be obtained (for example, in the
-% 	references, tables, supporting information, and other databases).
-%
-% 	All funding sources related to this work from all authors
-%
-% 	Any real or perceived financial conflicts of interests for any
-%	author
-%
-% 	Other affiliations for any author that may be perceived as
-% 	having a conflict of interest with respect to the results of this
-% 	paper.
-%
-%
-% It is also the appropriate place to thank colleagues and other contributors.
-% AGU does not normally allow dedications.
-
-\acknowledgments
-This work was funded by the OWP which is part of NOAA's National Weather Service.
-Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
-We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
-Additionally, David Blodgett from the USGS Water Mission Area was instrumental in helping define level paths and other hydrographic work.
-More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
-Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
-%The [type of data] data used for [brief context, description] in the study are available at [repository, source name] via [DOI, persistent identifier link] with  [license, access conditions] [optional in-text citation in References]
-%[Version number] of the [software name] used for [brief context, description of what the software was used for] is preserved at [DOI, persistent identifier link], available via [license type, access conditions] and developed openly at [software development platform link].* [optional in-text citation in References].
-
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% \bibliography{<name of your .bib file>} don't specify the file extension
-%
-% don't specify bibliography style
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\clearpage % this clears figures before references
-\bibliography{bibliography/owp_fim4_2022}
-%
-%Reference citation instructions and examples:
-%
-% Please use ONLY \cite and \citeA for reference citations.
-% \cite for parenthetical references
-% ...as shown in recent studies (Simpson et al., 2019)
-% \citeA for in-text citations
-% ...Simpson et al. (2019) have shown...
-%
-%
-%...as shown by \citeA{jskilby}.
-%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
-%...has been shown \cite{jskilbye}.
-%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
-%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
-%...has been shown by \cite <e.g.,>[and others]{lewin76}.
-%
-% apacite uses < > for prenotes and [ ] for postnotes
-% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
-%
-
-\end{document}
-
-%%%% More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-% If you are using line numbers, remember to surround
-% equations with \begin{linenomath*}...\end{linenomath*}
-
-%  To add line numbers to lines in equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
+Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022}.

### 2022-08-21 manuscripts/owp_fim4_2022/sections/results.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

@@
 Also, we note that FAR is somewhat, albeit marginally, decreased from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
-The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}.
-So overall and in other words, the entire distribution of HUC8's improves across the three methods.
-Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
-We move more details on the relationship betweeen stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).
+The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}).
+So overall and in other words, the entire distribution of HUC8s improves across the three methods.
+Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit stream order and mosaic the resulting FIMs, the better those FIMs perform.
+We move more details on the relationship between stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).
 
-Additional noteworthy trends in Figure \ref{fig:violin_plot} center around the all-around better performance of FIM's for those of higher Manning's n values and recurrence flows.
+Additional noteworthy trends in Figure \ref{fig:violin_plot} center around the all-around better performance of FIMs for those of higher Manning's n values and recurrence flows.
 The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value used in the BLE model, 0.11 \cite{fema2021base,fema2021estimated}.
-Most of this improvement is driven by significant increases in POD but, unfortunately, leading to a significant amount of overprediction as observed by the increase in FAR.
+Most of this improvement is driven by significant increases in POD but, unfortunately, leading to a significant amount of over-prediction as observed by the increase in FAR.
 More work can be invested to better regionalize Manning's n values for FIM purposes with HAND.
-We also observe additional trends associated with the magnitude or recurrence interval of the flow rates used with the higher flow rates exhibting better overall CSI, POD, and FAR values than the lower, 100 yr magnitude.
+We also observe additional trends associated with the magnitude or recurrence interval of the flow rates used with the higher flow rates exhibiting better overall CSI, POD, and FAR values than the lower, 100 yr magnitude.
 We introduce in the Discussion (Section \ref{sec:discussion}) that this skill premium exhibited by higher flow events could be due higher quality elevation data in regions that are not described as bathymetric areas.
 %
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering

### 2022-08-21 manuscripts/owp_fim4_2022/sections/title_authors.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments

--- 2022-02-14 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-%
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements, including customized APA reference formatting.
-%
-% You may copy this file and give it your
-% article name, and enter your text.
-%
-%
-% Step 1: Set the \documentclass
-%
-%
-
-%% To submit your paper:
-\documentclass[draft]{dependencies/agujournal2019}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage[inline]{dependencies/trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-
-%% added
-\usepackage{multirow}
-\usepackage{amsmath}
-
-\linenumbers
-
-%%%%%%%
-% As of 2018 we recommend use of the TrackChanges package to mark revisions.
-% Add track changes editors
-\addeditor{Fernando Aristizabal}
-
-% The trackchanges package adds five new LaTeX commands:
-%
-%  \note[editor]{The note}
-%  \annote[editor]{Text to annotate}{The note}
-%  \add[editor]{Text to add}
-%  \remove[editor]{Text to remove}
-%  \change[editor]{Text to remove}{Text to add}
-%
-% complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%
-
-\drafttrue
-
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-% ie, \journalname{Water Resources Research}
-
-\journalname{Water Resources Research}
-
-% starts document
-\begin{document}
-
-%% Title and Authors
 %% ------------------------------------------------------------------------ %%
@@
 \correspondingauthor{Fernando Aristizabal}{fernando.aristizabal@noaa.gov}
-
-%% Keypoints
-%% Keypoints, final entry on title page.
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-
-% Example:
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-% \end{keypoints}
-%
-\begin{keypoints}
-\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
-\item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
-\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation.
-\end{keypoints}
-%
-
-%% Abstract
-% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT and PLAIN LANGUAGE SUMMARY
-%
-% A good Abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
-
-% The Plain Language Summary should be written for a broad audience,
-% including journalists and the science-interested public, that will not have 
-% a background in your field.
-%
-% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
-% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
-% see http://sharingscience.agu.org/creating-plain-language-summary/)
-%
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
-
-\begin{abstract}
-The National Water Model (NWM) currently requires the post-processing of forecast discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
-Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is worthy of producing high-resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
-However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
-To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that reduces the Horton-Strahler stream order of the stream network.
-GMS contains all segments within the NWM stream network but instead of deriving HAND by accounting for all river segments at once, it is derived independently at the level path (LP) scale.
-LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
-These FIMs are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
-Improvement in mapping skill is observed when compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
-A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
-\end{abstract}
-%
-\section*{Plain Language Summary}
-Flooding is one of the most impactful natural disasters on life and property.
-The United States National Water Model (NWM) provides flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
-In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
-We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-We compared these HAND derived FIMs to maps from more realistic models and found improvement in mapping performance.
-%
-
-%%% Suggested section heads:
-% \section{Introduction}
-%
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
-
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-%
-% \section{Data} (Or section title might be a descriptive heading about data)
-%
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-%
-% \section{Conclusions}
-
-% Introduction
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
-In 2017 and 2019, river and flash flooding combined represented the leading cause of death and the second leading cause in 2018 among all natural disasters in the US \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
-More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
-With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
-Trends related to flood damages and fatalities have been steadily increasing over recent decades.
-\cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
-Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
-Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Operational Forecasting}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
-One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble stream flows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
-Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
-AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
-On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
-Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons to address the increasingly complex water challenges facing the US.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/forecast_points.jpg}
-\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
-Also show are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
-\label{fig:forecast_points}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{National Water Model}
-\label{ssec:national_water_model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
-The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
-The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
-As a complement to the operational NWM, OWP has also developed a configuration of the NWM that extends RFC forecasts downstream by assimilating and routing forecast streamflows to the next downstream AHPS forecast point.
-This configuration of the NWM is used to enhance forecasting skill by leveraging best available regional-scale data.
-The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
-Figure \ref{ssec:national_water_model} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
-The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
-
-The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
-The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
-Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
-As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
-The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Height Above Nearest Drainage}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
-\citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
-The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
-\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
-This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{HAND Implementations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
-As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
-NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
-By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
-Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC as an unofficial demonstrations to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
-Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
-More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
-
-Related to these efforts, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
-Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
-In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30 m resolutions \cite{yamazaki2019merit,donchyts2016global}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{OWP FIM}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Many of those in the academic community assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
-\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
-The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
-Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
-Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3 m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
-The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
-Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
-Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
-Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
-These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.
-
-With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
-Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
-This framework is open-source and modular enabling and accelerating the research to operations development pipeline.
-Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
-In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
-Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
-This capability will be referred to as Generalized Mainstems (GMS).
-The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
-%
- %
-% Materials and Methods
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Software Dependencies and Architecture}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
-Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
-A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub \cite{inundationMapping2022}.
-
-The pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
-Preprocessing acquires and prepares datasets for production of the FIM hydrofabric. 
-The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
-Functionality is included to turn FIM hydrofabric and streamflows into inundation maps represented in both binary and depths.
-Lastly, a test suite includes means of calculating evaluation metrics compared to a variety of preprocessed test case data.
-A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}
-\label{ssec:datasets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP) \cite{esipData2022}.
-The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
-It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
-For enforcing levee data into the DEMs, the USACE National Levee Database (NLD) is used to burn feature elevations \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{OpenStreetMap,GreatLakesHydrographyDataset}.
-Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
-These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
-The full input datasets presented by source are listed in Table \ref{tab:data}.
-%
-\begin{table}
-\caption{Data sources, names, and descriptions.}
-\label{tab:data}
-\centering
-\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
-%\begin{tabular}{l c c}
-\hline
-Source & Name & Description \\
-\hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
-\hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
-\hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters \\
-\hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. \\
-\hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. \\
-\hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. \\
-\hline
-USACE & NLD & Levee database of locations and elevations  \\
-\hline
-OSM & Land-Sea Border & Border of land and sea. \\
-\hline
-GLHD & Land-Great Lakes Border & Border of land and Great Lakes. \\
-\hline
-InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
-\hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
-\hline
-\end{tabular}
-\end{table}
-%
-Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and GMS networks and 1,604 HUC8s for the MS method. 
-These methods will be explained more later.
-An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hydro-conditioning}
-\label{ssec:hydro_conditioning}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
-These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
-HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
-In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligble for flooding.
-Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
-It can also be used to simulate more accurate bathymetry which is not accounted for in the 10 m DEM \cite{gesch2002national}.
-
-Specifically within the context of OWP FIM, the hydro-conditioning operations that take place in sequential order are presented. 
-Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing units of size HUC8. 
-The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reach code's membership in the processing unit.
-Hydro-conditioning raster operations take place on buffered boundary definitions to avoid edge contamination and effects \cite{lindsay2013measuring}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Network Enforcement} 
-\label{ssec:stream_network_enforcment}
-%
-The location of the stream network is enforced to ensure general agreement with established stream networks.
-The NHDPlusHR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
-However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
-For every NWM headwater segment a headwater point is derived and linearly interpolated to the nearest Burnline Event segment.
-Burnline Event headwater segments are split at the adjusted headwater point to match NWM flowlines.
-The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection and linear interpolation.
-Resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR Burnline Events.}
-\label{fig:stream_density_pruning}
-\end{figure}
-%
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.
-
-The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
-The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
-Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
-Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
-Some of these drawbacks are later addressed by additional conditioning techniques later on.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Levee Enforcement}
-%
-The DEM at 10 m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
-In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10 m DEM.
-The elevations found in the NLD are burned into the DEM if those elevations were found to exceed those already in place.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Depression Filling}
-\label{sssec:depression_filling}
-%
-Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
-The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
-Priority-Flood developed by \citeA{barnes2014priority} is an algorithm for filling said depressions and shown to have improved performance over early works in the field by \citeA{jenson1988extracting} implemented in \citeA{tarboton2005terrain} as well as \citeA{planchon2002fast}.
-The depression filling algorithm used in our pipeline is a Priority-Flood variant developed by \cite{zhou2016efficient} with enhanced single-thread performance and a time complexity of O(n log n) for floating point grids.
-This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
-The depression technique employed here does leave the existence of flat regions where pits existed a prior thus later requiring the need for resolving these flats.
-The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Thalweg Elevation Conditioning}
-\label{sssec:stream_thalweg_elevation_conditioning}
-%
-Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
-In order to prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
-Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50 m.
-The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
-
-The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
-\citeA{garousi2019terrain} proposed an algorithm that traverses stream thalweg pixels in a depth first manner starting with adding all the headwater pixels to a queue. 
-The connectivity of the thalweg pixels is defined by the D-8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
-At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in Equation \ref{eq:thalweg_breach},
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:thalweg_breach}
-\textbf{D}[x] = \min_{y\ drains\ to\ x} {(\ \textbf{D}[x]\ ,\ \textbf{D}[y]\ )}
-\end{equation}
-\end{linenomath*}
-%
-, in which \textbf{D} represents the array of thalweg adjusted elevations indexed by x and y where by y is upstream of x. 
-When a pixel's upstream neighbors are all evaluated, the downstream pixel is added into the queue thus the depth first traversal of the drainage network.
-This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in Critical Success Index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Deriving FIM Hydrofabric}
-\label{ssec:deriving_fim_hydrofabric}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Flow Directions and Flats Resolution}
-\label{ssec:flow_direction_and_flat_resolution}
-%
-To facilitate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D-8 flow directions.
-D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
-The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southeast direction as 8. 
-Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
-
-Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
-Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
-The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
-Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
-By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Deriving FIM Stream Network}
-\label{sssec:deriving_fim_stream_network}
-%
-The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
-The FIM stream network is of similar drainage density as the NWM V2.1 network but fully converges at all junctions leaving no divergences in the network.
-This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
-Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
-Each stream link in this derived FIM stream network is split into equidistant reaches.
-Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments.
-Stream links are then further segmented at NWM lakes and HUC8 boundaries.
-Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretation of the stream network.
-Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherent the synthetic rating curve properties of the upstream or downstream segment, depending on the topology.
-Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
-Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
-This stream network is important since it drives the HAND calculation and derivation of catchments.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Catchments}
-\label{sssec:catchments}
-%
-Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
-Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
-The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
-Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
-The other catchments are derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Height Above Nearest Drainage}
-%
-Once the pixel level catchments are derived, the final relative elevations can be computed.
-Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment described in Section \ref{sssec:catchments}.
-The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
-Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Synthetic Rating Curves}
-\label{sssec:synthetic_rating_curve}
-%
- A method for converting forecast river discharges from the NWM to stages or river depths is necessary for producing FIMs with HAND. 
-For one-dimensional models such as the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a synthetic rating curve at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
-For this application, we utilized the reach averaged approach for developing synthetic rating curves (SRC) \cite{zheng2018river}.
-The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-The reach averaged Manning's formula is derived to be 
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:reach_averaged_mannings_equation}
-\textbf{Q} = \frac{1}{n} \frac{V^{5/3}S^{1/2}}{L B^{2/3}} 
-\end{equation}
-\end{linenomath*}
-%
-where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
-Q, V, and B are taken a specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
-All units are international given the 1 numerator above n.
-The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
-The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the channel length.
-
-Setting of the Manning's n roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
-These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cross-walking with NWM Stream Network}
-\label{sssec:cross_walking_networks}
-%
-The stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with a NWM reach identifier so that a discharge can be converted to stage and later inundation extent.
-For the methods already discussed, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent methods discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
-Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
-While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
-More discussion on this will follow in Section \ref{sec:discussion}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stream Order Reduction}
-\label{ssec:stream_order_reduction}
-%
-FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
-Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
-An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the MS of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
-This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the MS that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the MS catchments.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{False negatives associated with confluence of tributaries with MS. Integers represent flow values from BLE 100 year event for the associated areas. 
-No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
-\label{fig:catchment_boundaries_issue}
-\end{figure}
-%
-We seek to resolve this problem by deriving HAND for processing units with stream networks of reduced stream order. 
-We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
-The resulting FIMs from the overlapping HAND processing units are mosaiced together taking any inundated area to be inundated but more will be explained in Section \ref{ssec:inundation_mapping}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Main-stems}
-\label{sssec:nws_mainstems}
-%
-The Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
-The MS network comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
-It also spans 121,724 reaches across 1,608 HUC8s.
-HAND was originally derived for this stream network to enhance mapping skill along these critical MS segments \cite{djokic2019arc}. 
-The inundation derived from this stream network is mosaiced with the inundation from the FR network to form the MS FIMs. 
-Within each HUC, you'll typically only find a MS stream network of stream order 1 (i.e.
-headwater) but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Generalized Mainstems}
-\label{sssec:generalized_mainstems}
-%
-To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call GMS.
-Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
-Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-Starting at the outlet, a unique level path is propagated upstream. 
-At every confluence, the direction of maximum flow path length is sought to propagate the current level path identifier.
-For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
-Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
-At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
-
-To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
-In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
-For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each level path is buffered to a polygon with a user-available distance parameter of 7 km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
-We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
-Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}c.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates how level paths for the NWM are derived.
-Level paths symbolized by lines of unique colors are propagated upstream following the direction of maximum arbolate sum at each junction.
-}
-\label{fig:level_path_methods}
-\end{figure}
-%
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/gms_methods.jpg}
-\caption{Overall procedure for GMS HAND.
-In (a), we illustrate all NWM stream lines symbolized by their level path.
-Meanwhile (b), demonstrates the DEM clipped to a 7 km buffer around two selected level paths.
-In (c), we show how HAND can be computed just for each one of these two level paths independently. 
-We also show inundation maps created for these two level paths in (c). 
-In (d), we show all the inundation maps for all the level paths mosaiced together. }
-\label{fig:gms_methods}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Inundation Mapping}
-\label{ssec:inundation_mapping}
-%
-The FIM hydrofabric consisting of the relative elevations grid, catchments grid, catchment polygons, rating curve, and cross-walking data are all used to convert forecasts from the NWM into forecasts extents.
-For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
-From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 8 spatial processing units in a rapid, parallel operation.
-The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
-
-Utilizing the stage-discharge relationships in the synthetic rating curves, each forecast for each catchment identifier is assigned a stage value. 
-The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
-We use the basic logic already established in previous works to conduct this \cite{nobre2016hand,liu2016cybergis,maidment2017conceptual}.
-Mathematically, the HAND values, $H_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
-For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
-For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim_depth}
-    D_{ij} = S_i - H_{ij}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim}
-    F_{ij} = D_{ij} > 0
-\end{equation}
-\end{linenomath*}
-%
-For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast in the form of a single raster file.
-For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
-Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
-Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim_depths}
-    D_{ij} = \max_{k=[1,...,K]} D_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-For the MS and GMS methods, the contributing areas are defined differently.
-For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to 2 for that case.
-For GMS, all FIMs from all the level paths in a given area are mosaiced together then K is set to this number of level paths.
-Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim}
-    F_{ij} = \max_{k=[1,...,K]} F_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}
-\label{ssec:evaluation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
-49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM so we used those corresponding discharges and mapping extents for evaluation.
-We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
-Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
-We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
-To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
-This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
-While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
-\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states. This dataset for 1\% and 0.2\% recurrence flows were used as benchmarks.}
-\label{fig:all_ble_maps}
-\end{figure}
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
-\caption{Illustrates Base Level Engineering (BLE) cross sections and stream lines.
-The BLE stream network, which is denser than the NWM V2.1 stream lines, is also shown.
-BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM reach.
-}
-\label{fig:ble_evaluation_method}
-\end{figure}
-%
-The metrics employed in this study to evaluate inundation extents include Critical Success Index (CSI), Probability of Detection (POD), and False Alarm Ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
-To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.
-The two types of errors consist of false positives (FP), or type I errors, which is dry in benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in benchmark by predicted dry.
-Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
-Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
-While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
-Thus, frequency dependent statistics should be used with caution when comparing across sites with varying frequencies. 
-Lastly, approximately 6 HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:csi}
-CSI = \frac{TP}{TP + FN + FP}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:pod}
-POD = \frac{TP}{TP + FN}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:far}
-FAR = \frac{FP}{TP + FP}
-\end{equation}
-\end{linenomath*}
-%
- %
-% Results
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{sec:results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Mapping Performance}
-\label{ssec:mapping_performance}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-We produced FIMs for the entire BLE domain within the 49 HUC8 areas across several states in the south central US. 
-The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
-We computed the statistics (CSI, POD, and FAR) for both 100 and 500 year events for Mannings N set to 0.06 and 0.12. 
-The distribution of these statistics can be examined in Figure \ref{fig:violin_plot} as violin plots.
-Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
-We also denote trend lines for each metric and Manning's n setting as well as their respective slope estimate and one-tailed p-value denoting the level of significance of the trend.
-
-Aggregating the metrics in the method above treats each HUC8 as it's own unit and does little to consider the size differences of the HUCs. 
-In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics}  the CSI, POD, and FAR recomputed for the entire domain using the sum of all the TPs, FPs, and FNs. 
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.9]{figures/violin_plots.jpg}
-\caption{Shows kernel density estimation of the distributions (sample size = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
-The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's n (0.06 and 0.12).
-Trend lines for each metric - Mannings combination are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}
-\label{fig:violin_plot}
-\end{figure}
-%
-\begin{table}[h!]
-\caption{Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated for BLE domain.
-         The best value across models is highlighted in bold.}
-\label{tab:aggregate_metrics}
-\centering
-%\begin{tabular}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
-\begin{tabular}{|c|c||c|c|c|c|c|c|}
-\hline
-\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{MS} & \multicolumn{2}{|c|}{GMS} \\
-\cline{3-8}
-  &  & 100yr & 500yr & 100yr & 500yr & 100yr & 500yr \\
-\hline
-\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & \textbf{0.5796} & \textbf{0.6075} \\
-\cline{2-8}
-  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & \textbf{0.6182} & \textbf{0.6435} \\
-\hline
-\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & \textbf{0.6633} & \textbf{0.6863} \\
-\cline{2-8}
-  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & \textbf{0.7606} & \textbf{0.7810} \\
-\hline
-\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1787 & \textbf{0.1589} & \textbf{0.1778} & \textbf{0.1589} \\
-\cline{2-8}
-  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & \textbf{0.2324} & \textbf{0.2148} \\
-\hline
-\end{tabular}
-\end{table}
-%
-% Interpretation of metrics
-From Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, we denote several meaningful trends. 
-Using CSI as an overall proxy for skill of the FIM, we note that generally speaking the skill is correlated with a reduction of the stream orders of the processing units used for HAND.
-In other words, the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
-While this trend is evident for both sets of Manning's n values, the trend is slightly more significant for the higher value of 0.12.
-Other trends related to this Figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
-We also note how the higher Manning's n value enhances performance for both of these recurrence intervals across all models.
-
-Dissecting the improvements and trends presented in the previous paragraph comes down mostly to improvement in POD or a reduction in absolute amount of FNs.
-POD being the primary driver in skill enhancement is evident across models by comparing the slope of the POD lines with the slope of the FAR lines.
-Even though aggregating metrics by HUC8 yields a statistically zero trend, one does see a slight reduction in FAR across models that reduce HAND's maximum stream order.
-Additionally, we note that POD is a primary driver in enhancing performance across Manning's n values as well.
-This significant improvement comes at a cost of false alarms as the FAR increases significantly across Manning's n values.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Computational Performance}
-\label{ssec:compuational_performance}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The NFIE experiments were able to produce HAND for 331 HUC6's in 1.34 CPU years \cite{liu2016cybergis} and estimates using work from \citeA{djokic2019arc} put producing HAND at the FR NWM at 0.55 CPU years.
-For our work, we were able to produce HAND at the full NWM resolution in 0.13 CPU years which represents a substantial speed-up compared to previous works.
-For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
-GMS which generalizes HAND production to level path scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
-%
- %
-% Discussion
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-\label{sec:discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Overall, we note a positive relationship between FIM skill and a reduction of the stream order of the stream network we use to derive the HAND datasets.
-Most of this change is accounted for by increasing POD thus reducing FNs especially along higher order rivers with higher flow magnitudes.
-We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
-This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
-
-In analyzing the data, we found a slight reduction in FAR was detected and more digging pointed to a bias in rating curves introduced by stream order reduction.
-Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
-Sub-figure \ref{fig:rating_curve_comparison}a shows how the average rating curves for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to bias down (and to the right) with ever increasing stream order reduction (FR to MS to GMS).
-This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
-Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
-Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
-While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
-Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius.
-Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively (see Equation \ref{eq:reach_averaged_mannings_equation}).
-Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
-Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
-Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
-We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
-In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
-Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
-Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments that encompass neighboring tributaries.
-These tributaries have a significant amount of bathymetry that is low-lying thus easily including the SRC derivation. 
-They also contribute volume and bed area that is technically not perpendicular to the flux of streamflow being accounted for in the stream in question. 
-Careful examination of Figure \ref{fig:gms_enhancement}b shows how much larger catchments include neighboring tributaries and the geometry associated with those tributaries. 
-This geometry is not perpendicular to the flow that is associated with the main reach thus leading to biases in the SRC.
-We consider this fact to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
-
-Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
-While the cause of these errors can be varied, we assert here that conflating 4 networks for use in evaluations leads to significant error.
-As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
-One of the issues is when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
-The utilization of MS and GMS only conflates to NWM catchments directly associated with the level path in question which is inherently easy to do with those methods. 
-Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
-The NWM stream network was derived using the NHD medium resolution dataset which was derived from coarser DEMs than those used here. 
-Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
-Until a singular stream network is used for the NWM, BLE benchmark, and for HAND based FIM, conflation will continue being a source of error.
-
-Our qualitative analysis suggests that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM for future development.
-The bathymetry of the 10 m DEM from 3DEP is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
-Manning's n which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
-Some adjusting parameter must be introduced to enhance the estimation of the bathymetric representation.
-Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
-
-Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
-For HAND to be derived and thus create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
-In other words, an entire area eligible for flooding must monotonically decrease in elevation. 
-DEM's naturally don't do this and the dynamics of true flood events don't follow drainage patterns.
-Enforcing this assumption for HAND leads to significant amount of DEM manipulations that introduce basic errors.
-These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
-Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/rating_curve_comparison.jpg}
-\caption{Illustrates average quantities for the three methods, FR, MS, and GMS, for each stage value (m). 
-The values are (a) Discharge $m^3s^{-1}$, (b) Volume $m^3$, (c) Bed Area $m^2$, (d) a function of Volume and Bed Area, and (e) number of pixels.
-}
-\label{fig:rating_curve_comparison}
-\end{figure}
-%
-%
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
-\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105.
-Catchment boundaries and stream lines are shown in white and dotted green, respectively.
-Sub-figure (a) shows agreement of FR HAND denoting significant areas of under-prediction due to junctions and catchment boundaries.
-Meanwhile, (b) shows the agreement for GMS and much larger catchments leading to much better inundation agreement for this given reach. 
-Overall, this illustrates the benefits of stream order reduction for deriving HAND datasets.
-}
-\label{fig:gms_enhancement}
-\end{figure}
-%
- %
-% Conclusions
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-\label{sec:conclusions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Floods present a significant, under-served, and expanding risk to life, property, and resources.
-Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
-The National Water Model (NWM) developed by the National Oceanic and Atmospheric Administration's Office of Water Prediction, along with partners, provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
-Additional processing is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
-Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
-HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
-Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
-HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
-Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
-We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
-This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
-FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
-Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.
-Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
-Due to inherent limitations with HAND, scalable, physics-based methods are necessary to consider to provide a better representation of flood extent dynamics in steady and unsteady conditions. 
-%
- %
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-%\section{= enter section title =}
-%Text here ===>>>
-%\section{= enter section title =}
-%Text here ===>>>
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables;  other caption information
-%  should be placed in last line of the table, using
-% \multicolumn2l{$^a$ This is a table note.}
-%
-%----------------
-% EXAMPLE FIGURES
-%
-% \begin{figure}
-% \includegraphics{example.png}
-% \caption{caption}
-% \end{figure}
-%
-% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
-% \begin{figure}
-% \noindent\includegraphics[width=\textwidth]{anothersample.png}
-%\caption{caption}
-%\label{pngfiguresample}
-%\end{figure}
-%
-%
-% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
-This is common when trying to add a PDF figure without pdflatex.
-% \begin{figure}
-% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
-%\caption{caption}
-%\label{pdffiguresample}
-%\end{figure}
-%
-%
-% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
-%
-
-%
-% ---------------
-% EXAMPLE TABLE
-%
-% \begin{table}
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \centering
-% \begin{tabular}{l c}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Footnote text here.}
-% \end{tabular}
-% \end{table}
-
-%% SIDEWAYS FIGURE and TABLE
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{figsamp}
-% \caption{caption here}
-% \label{newfig}
-% \end{sidewaysfigure}
-%
-%  \begin{sidewaystable}
-%  \caption{Caption here}
-% \label{tab:signif_gap_clos}
-%  \begin{tabular}{ccc}
-% one&two&three\\
-% four&five&six
-%  \end{tabular}
-%  \end{sidewaystable}
-
-%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
-%\begin{linenomath*}
-%\begin{equation}
-%y|{f} \sim g(m, \sigma),
-%\end{equation}
-%\end{linenomath*}
-
-%%% End of body of article
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary, Notation or Acronym section goes here:
-%
-%%%%%%%%%%%%%%
-% Glossary is only allowed in Reviews of Geophysics
-%  \begin{glossary}
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \end{glossary}
-
-%
-%%%%%%%%%%%%%%
-% Acronyms
-%   \begin{acronyms}
-%   \acro{Acronym}
-%   Definition here
-%   \acro{EMOS}
-%   Ensemble model output statistics
-%   \acro{ECMWF}
-%   Centre for Medium-Range Weather Forecasts
-%   \end{acronyms}
-
-%
-%%%%%%%%%%%%%%
-% Notation
-%   \begin{notation}
-%   \notation{$a+b$} Notation Definition here
-%   \notation{$e=mc^2$}
-%   Equation in German-born physicist Albert Einstein's theory of special
-%  relativity that showed that the increased relativistic mass ($m$) of a
-%  body comes from the energy of motion of the body—that is, its kinetic
-%  energy ($E$)—divided by the speed of light squared ($c^2$).
-%   \end{notation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-%
-% The acknowledgments must list:
-%
-% >>>>	A statement that indicates to the reader where the data
-% 	supporting the conclusions can be obtained (for example, in the
-% 	references, tables, supporting information, and other databases).
-%
-% 	All funding sources related to this work from all authors
-%
-% 	Any real or perceived financial conflicts of interests for any
-%	author
-%
-% 	Other affiliations for any author that may be perceived as
-% 	having a conflict of interest with respect to the results of this
-% 	paper.
-%
-%
-% It is also the appropriate place to thank colleagues and other contributors.
-% AGU does not normally allow dedications.
-
-\acknowledgments
-This work was funded by the OWP which is part of NOAA's National Weather Service.
-Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
-We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
-Additionally, David Blodgett from the USGS Water Mission Area was instrumental in helping define level paths and other hydrographic work.
-More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
-Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
-%The [type of data] data used for [brief context, description] in the study are available at [repository, source name] via [DOI, persistent identifier link] with  [license, access conditions] [optional in-text citation in References]
-%[Version number] of the [software name] used for [brief context, description of what the software was used for] is preserved at [DOI, persistent identifier link], available via [license type, access conditions] and developed openly at [software development platform link].* [optional in-text citation in References].
-
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% \bibliography{<name of your .bib file>} don't specify the file extension
-%
-% don't specify bibliography style
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\clearpage % this clears figures before references
-\bibliography{bibliography/owp_fim4_2022}
-%
-%Reference citation instructions and examples:
-%
-% Please use ONLY \cite and \citeA for reference citations.
-% \cite for parenthetical references
-% ...as shown in recent studies (Simpson et al., 2019)
-% \citeA for in-text citations
-% ...Simpson et al. (2019) have shown...
-%
-%
-%...as shown by \citeA{jskilby}.
-%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
-%...has been shown \cite{jskilbye}.
-%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
-%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
-%...has been shown by \cite <e.g.,>[and others]{lewin76}.
-%
-% apacite uses < > for prenotes and [ ] for postnotes
-% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
-%
-
-\end{document}
-
-%%%% More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-% If you are using line numbers, remember to surround
-% equations with \begin{linenomath*}...\end{linenomath*}
-
-%  To add line numbers to lines in equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}

### 2022-08-29 manuscripts/owp_3dep_2022/owp_3dep.tex
* started lidar manuscript * advances lidar experiment script

%% 
%% Copyright 2007-2020 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references

\documentclass[preprint,12pt]{dependencies/elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Additional packages
\usepackage{acronym}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\journal{Advances in Water Resources}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%% \fntext[label3]{}

\title{Varying the Spatial Resolution of LiDAR Derived Elevations to Balance Skill and Computational Costs for Flood Inundation Mapping Applications}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author[lynk,nwc,uf]{Fernando Aristizabal}
\author[nwc]{Fernando Salas}
\author[uh]{Taher Chegini}
\author[usgs]{Gregory Petrochenkov}
\author[uf]{Jasmeet Judge}

\affiliation[lynk]{
             organization={Lynker}, %Department and Organization
             addressline={338 E Market St}, 
             city={Leesburg},
             postcode={20176}, 
             state={VA},
             country={USA}
            }

\affiliation[nwc]{
             organization={National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration}, %Department and Organization
             addressline={205 Hackberry Ln}, 
             city={Tuscaloosa},
             postcode={35401}, 
             state={VA},
             country={USA}
            }

\affiliation[uf]{
             organization={Center for Remote Sensing, Agricultural and Biological Engineering, University of Florida}, %Department and Organization
             addressline={1741 Museum Rd}, 
             city={Gainesville},
             postcode={32603}, 
             state={FL},
             country={USA}
            }

\affiliation[uh]{
             organization={Civil and Environmental Engineering, University of Houston},
             addressline={4226 Martin Luther King Boulevard}, 
             city={Houston},
             postcode={77204}, 
             state={TX},
             country={USA}
            }

\affiliation[usgs]{
             organization={Hydrologic Applied Innovations Lab, New York Water Science Center, United States Geological Survey},
             addressline={425 Jordan Rd}, 
             city={Troy},
             postcode={12180}, 
             state={NY},
             country={USA}
            }

\begin{abstract}
%% Text of abstract

\end{abstract}

%%Graphical abstract
%\begin{graphicalabstract}
%\includegraphics{grabs}
%\end{graphicalabstract}

%%Research highlights
\begin{highlights}
\item Continental-scale flood inundation mapping requires simplifying assumptions
\item Utilizing LiDAR derived data from the USGS 3DEP program enhances the skill on flood inundation maps produced from HAND.
\item Varying the spatial resolution provides no skill enhancement when evaluated at large scales across three study sites and benchmark datasets.
\end{highlights}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Acronyms
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acronym}

\acro{OWP}[OWP]{Office of Water Prediction}

\end{acronym}

\linenumbers

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% main text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}
%
Floods are among the most frequent, damaging, and deadly of natural disasters \citet{doocy2013human,stromberg2007natural,kahn2005death}. 
What's worse, trends in flood events have been increasing in frequency and intensity driven by climate change, anthropogenic development, \citet{berz2000flood,mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect} with more hydrologic extremes expected in the future \citet{kahn2005death,tabari2020climate,milly2002increasing,wing2018estimates}.
Floods impact mortality and morbidity through drowning, trauma, or increased risk of infectious disease \citet{jonkman2005global,beinin2012medical,alajo2006cholera,french1983mortality}.
Flooding impacts population inequitably \citet{kahn2005death,smiley2021social,stromberg2007natural,jonkman2005global,tellman2020using,tellman2021satellite}.
Early-warning systems can help \citet{stromberg2007natural}
The \ac{OWP}.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% For citations use: 
%%       \citet{<label>} ==> Jones et al. [21]
%%       \citep{<label>} ==> [21]
%%

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
\bibliographystyle{elsarticle-num-names} 
\bibliography{bibliography/owp_3dep_2022}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

%\begin{thebibliography}{00}

%% \bibitem[Author(year)]{label}
%% Text of bibliographic item

%\bibitem[ ()]{}

%\end{thebibliography}

\end{document}
\endinput
%%
%% End of file `elsarticle-template-num-names.tex'.

### 2022-08-29 dem_3dep_comparison.py

@@ -1,0 +1,254 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import pandas as pd
+import argparse
+from foss_fim.src.acquire_and_preprocess_inputs import manage_preprocessing
+import subprocess
+import shutil
+import os
+from ast import literal_eval
+
+def Dem_3dep_comparison( huc4=1202,
+                         resolutions=[20,15,10,5,3,1], 
+                         tile_sizes={20:10000,15:10000,10:5000,5:2000,3:1500,1:1000},
+                         huc_sizes ={20:8, 15:8, 10:8, 5:8, 3:12, 1:12},
+                         stream_resolutions=['FR','MS', 'GMS'],
+                         acquire_data=True,
+                         resolutions_to_acquire_data_for = {20,15,10,5,3,1},
+                         skip_hydrofabric=False,
+                         skip_evaluate=False,
+                         manning = 12,
+                         production = True,
+                         overwrite = True,
+                         overwrite_eval = True,
+                         remove_inputs = True,
+                         remove_outputs = True,
+                         hand_jobs_dict = { 
+                                            'FR' : 
+                                              { 1 : 3, 3 : 3, 5 : 7, 10 : 7, 15 : 7 , 20 : 7 },
+                                            'MS' :
+                                              { 1 : 5, 3 : 3, 5 : 7, 10 : 7, 15 : 7 , 20 : 7 },
+                                            'GMS' :
+                                              { 1 : [10,5], 3 : [15,15], 5 : [7,15], 10 : [7,15], 15 : [7,15] , 20 : [7,15] }
+                                           },
+                         eval_jobs_dict = { 
+                                            'FR' :
+                                              { 1 : 12, 3 : 12, 5 : 4, 10 : 7, 15 : 7 , 20 : 7 },
+                                            'MS' :
+                                              { 1 : 12, 3 : 12, 5 : 4, 10 : 7, 15 : 7 , 20 : 7 },
+                                            'GMS' :
+                                              { 1 : [3,4], 3 : [3,4], 5 : [2,4], 10 : [3,4], 15 : [3,4] , 20 : [3,4]}
+                                              }
+        ):
+
+    resolutions_to_acquire_data_for = set(resolutions_to_acquire_data_for)
+
+    # huc4, tile_size, resolution
+    #for resolution, tile_size, huc_size in zip(resolutions,tile_sizes,huc_sizes):
+    for resolution in resolutions:
+
[…]

### 2022-08-29 gms_run_branch.sh

@@ -134,4 +134,4 @@ fi
 
 # -------------------
 ## GET NON ZERO EXIT CODES ##
-grep -ER 'Exit status: [1-9]' $outputRunDataDir/logs/branch/*_branch_*.log > $outputRunDataDir/logs/branch/non_zero_exit_codes.log
+# grep -ER 'Exit status: [1-9]' $outputRunDataDir/logs/branch/*_branch_*.log > $outputRunDataDir/logs/branch/non_zero_exit_codes.log

### 2022-08-29 gms_run_unit.sh

@@ -118,7 +118,7 @@ export input_GL_boundaries=$inputDataDir/landsea/gl_water_polygons.gpkg
 export deny_gms_unit_list=$deny_gms_unit_list
 
 ## Input handling ##
-$srcDir/check_huc_inputs.py -u "$hucList"
+#$srcDir/check_huc_inputs.py -u "$hucList"
 
 ## Make output and data directories ##
 if [ "$retry" = "--retry-failed" ]; then
@@ -158,4 +158,4 @@ python3 $srcDir/gms/aggregate_branch_lists.py -l $hucList
 
 ## GET NON ZERO EXIT CODES ##
 # get positive non-zero exit codes 
-grep -ER 'Exit status: [1-9]' $outputRunDataDir/logs/unit/*_unit.log > $outputRunDataDir/logs/unit/non_zero_exit_codes.log
+# grep -ER 'Exit status: [1-9]' $outputRunDataDir/logs/unit/*_unit.log > $outputRunDataDir/logs/unit/non_zero_exit_codes.log

### 2022-08-29 manuscripts/owp_3dep_2022/Makefile.sh

@@ -1,0 +1,20 @@
+#!/bin/bash -e
+
+parent_path=$( cd "$(dirname "${BASH_SOURCE[0]}")" ; pwd -P )
+cd $parent_path
+
+fullFile=owp_3dep.tex
+baseName="${fullFile%%.*}"
+extension="${fullFile#*.}"
+
+pdflatex "$fullFile"
+bibtex "$baseName".aux
+pdflatex "$fullFile"
+pdflatex "$fullFile"
+
+# expands all tex files into one file
+#echo 'Lat-Expand to Create One Tex File ...'
+#/usr/bin/latexpand/latexpand --keep-comments --biber bibliography/owp_3dep_2022.bib owp_3dep.tex -o owp_3dep_expanded.tex
+
+
+

### 2022-08-29 manuscripts/owp_3dep_2022/docker_run_and_make.sh

@@ -1,0 +1,5 @@
+#!/bin/bash -e
+
+codeDir=$1
+
+docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex:20220201 /foss_fim/manuscripts/owp_fim4_2022/Makefile.sh

### 2022-08-29 src/acquire_and_preprocess_inputs.py

@@ -318,7 +318,6 @@ def retrieve_and_reproject_3dep_for_huc( wbd,
                                          wbd_buffer=2000
                                        ):
 
-    
     def __fishnet_geometry(geometry,cell_size_meters=1000,cell_buffer=10):
         
         xmin, ymin, xmax, ymax = geometry.bounds
@@ -455,7 +454,8 @@ def retrieve_and_reproject_3dep_for_huc( wbd,
                                     ) 
                 break
             
-            except (ReadTimeout,HTTPError,RasterioIOError) as e:
+            #except (ReadTimeout,HTTPError,RasterioIOError,pygeoogc.exceptions.ServiceUnavailable) as e:
+            except Exception as e:
                 
                 print(f'{e} - idx: {idx} | retries: {retries}')
                 retries += 1
@@ -467,28 +467,10 @@ def retrieve_and_reproject_3dep_for_huc( wbd,
                     dem = __get_tile_from_nhd(geometry,huc)
                     break
             
-            """
-            except RasterioIOError:
-
-                print(f'RasterIOError - idx: {idx} | retries: {retries}')
-                retries += 1
-                
-                if retries < max_retries:
-                    continue
-                else:
-                    print(f'RasterIOError - idx: {idx} | retries: {retries} - Using NHD')
-                    dem = __get_tile_from_nhd(geometry,huc)
-                    break
-        dem = py3dep.get_map( 'DEM',
-                              geometry=geometry,
-                              resolution=dem_resolution,
-                              geo_crs=wbd_crs
-                            ) 
-        """
         if dem is None:
             return
+        
         # reproject and resample
-        #print(f'{huc} reprojecting ')
         #dem = dask.delayed(dem.rio.reproject)( wbd_crs,
         dem = dem.rio.reproject( wbd_crs,
                                  resolution=dem_resolution,
@@ -526,6 +508,8 @@ def retrieve_and_reproject_3dep_for_huc( wbd,
     
     for huc4_dir in tqdm(huc4_directories,desc=' Merging tiles'):
         
+        huc4_dir = os.path.join(output_dir,huc4_dir)
+        
         if not os.path.isdir(huc4_dir):
             continue
         
@@ -543,7 +527,7 @@ def retrieve_and_reproject_3dep_for_huc( wbd,
         
[…]

### 2022-08-29 src/agreedem.py

@@ -54,6 +54,7 @@ def agreedem(rivers_raster, dem, output_raster, workspace, grass_workspace, buff
     smo_profile = dem_profile.copy()
     smo_profile.update(nodata = 0)
     smo_profile.update(dtype = 'float32')
+    smo_profile.update(BIGTIFF='yes')
     smo_output = os.path.join(workspace, 'agree_smogrid.tif')
 
     # Windowed reading/calculating/writing
@@ -95,7 +96,15 @@ def agreedem(rivers_raster, dem, output_raster, workspace, grass_workspace, buff
     # Compute allocation and proximity grid using GRASS gis
     # r.grow.distance tool. Output distance grid in meters. Set datatype
     # for output allocation and proximity grids to float32.
+    ####################
+    #### OPEN ISSUE ####
+    ## Both of these grids below are written out as geotiff. Documentation on Grass GIS doesn't  seem
+    ## to accomodate other raster drivers or even the bigtiff option.
+    ## Not having this ability creates a reading error lower.
+    ## Please run HUC12 120200020502 for more information.
+    ####################
     vectdist_grid, vectallo_grid = r_grow_distance(smo_output, grass_workspace, 'Float32', 'Float32')
+
 
     #------------------------------------------------------------------
     # 4. From Hellweger documentation: Compute the buffer grid
@@ -118,7 +127,7 @@ def agreedem(rivers_raster, dem, output_raster, workspace, grass_workspace, buff
         with rasterio.open(buf_output, 'w', **buf_profile) as raster:
             for ji, window in elev.block_windows(1):
                 # read distance, allocation, and elevation datasets
-                vectdist_data_window = vectdist.read(1, window = window)
+                vectdist_data_window = vectdist.read(1, window = window) # READ ERROR FOR LARGE TIFFs HERE
                 vectallo_data_window = vectallo.read(1, window = window)
                 elev_data_window = elev.read(1, window = window)
 
@@ -162,6 +171,7 @@ def agreedem(rivers_raster, dem, output_raster, workspace, grass_workspace, buff
     agree_output = output_raster
     agree_profile = dem_profile.copy()
     agree_profile.update(dtype = 'float32')
+    agree_profile.update(BIGTIFF='YES')
 
     # Windowed reading/calculating/writing
     with rasterio.Env():
@@ -221,6 +231,7 @@ def agreedem(rivers_raster, dem, output_raster, workspace, grass_workspace, buff
     vectallo.close()
[…]

### 2022-08-29 src/entrypoint.sh

@@ -4,3 +4,4 @@ umask 002
 echo "Group ID: $GID | Group Name: $GN"
 newgrp $GN
 exec "$@"
+

### 2022-08-29 src/gms/derive_level_paths.py

@@ -25,6 +25,7 @@ def Derive_level_paths(in_stream_network, out_stream_network,branch_id_attribute
     # load file
     if verbose:
         print("Loading stream network ...")
+
     stream_network = StreamNetwork.from_file(in_stream_network)
 
     inlets_attribute = 'inlet_id'
@@ -32,6 +33,7 @@ def Derive_level_paths(in_stream_network, out_stream_network,branch_id_attribute
     outlet_linestring_index = -1
 
     # converts multi-linestrings to linestrings
+    #print(stream_network.crs);exit(0)
     stream_network = stream_network.multilinestrings_to_linestrings()
 
     # derive nodes

### 2022-08-29 src/gms/stream_branches.py

@@ -24,7 +24,7 @@ class StreamNetwork(gpd.GeoDataFrame):
     values_excluded = None
     attribute_excluded = None
 
-    def __init__(self,*args,**kwargs):
+    def __init__(self, *args, **kwargs):
         
         if kwargs:
             branch_id_attribute = kwargs.pop("branch_id_attribute",None)
@@ -32,7 +32,7 @@ class StreamNetwork(gpd.GeoDataFrame):
             attribute_excluded = kwargs.pop("attribute_excluded",None)
         
         super().__init__(*args,**kwargs)
-        
+
         self.branch_id_attribute = branch_id_attribute
         self.values_excluded = values_excluded
         self.attribute_excluded = attribute_excluded
@@ -57,7 +57,12 @@ class StreamNetwork(gpd.GeoDataFrame):
         if verbose: 
             print('Loading file')
         
-        return(cls(gpd.read_file(filename,*args,**kwargs),**inputs))
+        gdf = gpd.read_file(filename)
+        inputs['crs'] = gdf.crs
+
+        f = cls(gdf,**inputs)
+
+        return(f)
 
 
     def write(self,fileName,layer=None,index=True,verbose=False):

### 2022-08-29 tools/compare_dems.py

@@ -5,8 +5,183 @@ import numpy as np
 import argparse
 import os
 from glob import glob
+import geopandas as gpd
+import pandas as pd
+from foss_fim.src.utils.shared_variables import PREP_PROJECTION
+import csv
+from pygeohydro import WBD
+from tqdm.dask import TqdmCallback
+from utils.shared_variables import DEM_3DEP_RASTER_DIR_NAME, PREP_PROJECTION
+from inundation import __append_huc_code_to_file_name
+import dask
 
 
-def get_dem_filenames(resolution,dem_source):
+def Compare_dems(hucs, desired_crs=PREP_PROJECTION, output_raster_file=None, output_stats_file=None):
 
-    glob(f'/data/inputs/{dem_source}/dem_3dep_*_{resolution}.vrt'
+    # Parse HUCs from hucs.
+    if isinstance(hucs,list):
+        if len(hucs) == 1:
+            try:
+                with open(hucs[0]) as csv_file:  # Does not have to be CSV format.
+                    hucs = [i[0] for i in csv.reader(csv_file)]
+            except FileNotFoundError:
+                hucs = hucs
+        else:
+                hucs = hucs
+    elif isinstance(hucs,str):
+        try:
+            with open(hucs) as csv_file:  # Does not have to be CSV format.
+                hucs = [i[0] for i in csv.reader(csv_file)]
+        except FileNotFoundError:
+            hucs = list(hucs)
+        
+    huc_length = [ len(h) for h in hucs ]
+    huc_length = set(huc_length)
+
+    if len(huc_length) > 1:
+        raise ValueError("Pass equivalent length HUCs")
+
+    huc_length = list(huc_length)[0]
+
+    wbd = WBD(f'huc{huc_length}')
+    wbd_df = wbd.byids(f'huc{huc_length}',hucs)
+    wbd_df = wbd_df.to_crs(PREP_PROJECTION)
+
+    dask.compute(__process_comparison(wbd_df,huc_length,output_raster_file,output_stats_file))
+
+
+def __process_comparison(wbd_df,huc_length,output_raster_file=None,output_stats_file=None):
+
+    for _,e in wbd_df.iterrows():
+    
+        huc = e[f'huc{huc_length}']
+        huc_geom = e['geometry']
+        
+        # get NHD dems
+        nhd_dem = __get_nhd_dem(huc,huc_geom)
+        #nhd_dem_arr, nhd_dem_trans, nhd_dem_profile = nhd_dem[0],nhd_dem[1],nhd_dem[2]
+
[…]

### 2022-08-29 tools/consolidate_metrics.py

@@ -4,6 +4,7 @@ import pandas as pd
 import os
 import re
 import json
+import csv
 import numpy as np
 from glob import glob
 from tqdm import tqdm
@@ -22,8 +23,10 @@ pd.set_option('display.max_columns', None)
 
 def Consolidate_metrics( benchmarks=['all'],versions=['all'],
                          zones=['total_area'],matching_hucs_only=True,
+                         hucs_of_interest=None,
                          metrics_output_csv=None,
-                         impute_missing_ms=False
+                         impute_missing_ms=False,
+                         quiet=False
                        ):
 
     """ Consolidates metrics into single dataframe """
@@ -49,6 +52,24 @@ def Consolidate_metrics( benchmarks=['all'],versions=['all'],
     else:
         zones = list(zones)
     
+    # Parse HUCs from hucs_of_interest.
+    if isinstance(hucs_of_interest,list):
+        if len(hucs_of_interest) == 1:
+            try:
+                with open(hucs_of_interest[0]) as csv_file:  # Does not have to be CSV format.
+                    huc_list = [i[0] for i in csv.reader(csv_file)]
+            except FileNotFoundError:
+                huc_list = hucs_of_interest
+        else:
+                huc_list = hucs_of_interest
+    elif isinstance(hucs_of_interest,str):
+        try:
+            with open(hucs_of_interest) as csv_file:  # Does not have to be CSV format.
+                huc_list = [i[0] for i in csv.reader(csv_file)]
+        except FileNotFoundError:
+            huc_list = list(hucs_of_interest)
+    elif hucs_of_interest is None:
+        pass
 
     # loop through benchmarks
     consolidated_metrics_df = [ f for f in return_dataframe_for_benchmark_source(benchmarks,zones) ]
@@ -58,6 +79,9 @@ def Consolidate_metrics( benchmarks=['all'],versions=['all'],
 
     # find matching rows
     consolidated_metrics_df = find_matching_rows_by_attribute_value(consolidated_metrics_df,'version',versions)
+
+    if hucs_of_interest is not None:
+        consolidated_metrics_df = find_matching_rows_by_attribute_value(consolidated_metrics_df,'huc',huc_list)
 
     if impute_missing_ms:
         consolidated_metrics_df = impute_missing_MS_with_FR(consolidated_metrics_df)
@@ -75,9 +99,13 @@ def Consolidate_metrics( benchmarks=['all'],versions=['all'],
                                                 aggfunc=np.sum
                                                )
[…]

### 2022-08-29 tools/get_sources.py

@@ -5,9 +5,13 @@ import py3dep
 import argparse
 import csv
 import pandas as pd
+import dask
+from tqdm.dask import TqdmCallback
+from foss_fim.src.utils.shared_variables import PREP_PROJECTION
+import geopandas as gpd
 
 
-def get_sources_by_hucs(hucs,output_file=None):
+def get_sources_by_hucs(hucs,desired_crs=PREP_PROJECTION,output_file=None):
 
     # Parse HUCs from hucs.
     if isinstance(hucs,list):
@@ -33,20 +37,45 @@ def get_sources_by_hucs(hucs,output_file=None):
         raise ValueError("Pass equivalent length HUCs")
 
     huc_length = list(huc_length)[0]
+    huc_column_name = f'huc{huc_length}'
 
     wbd = WBD(f'huc{huc_length}')
     wbd_df = wbd.byids(f'huc{huc_length}',hucs)
 
-    bounds = [g.bounds for g in wbd_df.geometry]
+    wbd_df = wbd_df.to_crs(desired_crs)
 
-    source_polygons = [py3dep.query_3dep_sources(b,crs=wbd_df.crs) for b in bounds]
+    bounds_hucs = [(r['geometry'].bounds,r[huc_column_name]) for _,r in wbd_df.iterrows()]
 
+    def __get_sources_for_dask(b,h):
+        
+        try:
+            sources = py3dep.query_3dep_sources(b,crs=wbd_df.crs)
+        except Exception as e:
+            print(f'HUC {h}: {e}')
+        else:
+            return(sources)
+
+    #source_polygons = [py3dep.query_3dep_sources(b,crs=wbd_df.crs) for b in bounds]
+    compute_list = [dask.delayed(__get_sources_for_dask)(b,h) for b,h in bounds_hucs]
+
+    with TqdmCallback(desc='Get data sources'):
+        source_polygons = dask.compute(*compute_list)
+
+    print('Creating dataframe with results')
     source_polygons = pd.concat(source_polygons).reset_index(drop=True)
     source_polygons = source_polygons.explode()
     source_polygons = source_polygons.reset_index(drop=True)
+    source_polygons = source_polygons.loc[source_polygons.is_valid,:]
+    source_polygons = source_polygons.to_crs(desired_crs)
 
-    print(source_polygons)
-    breakpoint()
+    source_polygons = source_polygons.loc[:,['dem_res','geometry']]
+
+    # clipping
+    print('Clipping to WBDs')
[…]

### 2022-08-29 tools/gms_tools/inundate_gms.py

@@ -21,7 +21,7 @@ def Inundate_gms(
                   log_file=None,
                   output_fileNames=None
                  ):
-
+    
     # input handling
     if hucs is not None:
         try:
@@ -39,7 +39,7 @@ def Inundate_gms(
         if os.path.exists(log_file):
             os.remove(log_file)
         
-        print('HUC8,BranchID,Exception',file=open(log_file,'w'))
+        print('HUC,BranchID,Exception',file=open(log_file,'w'))
     #if log_file:
         #logging.basicConfig(filename=log_file, level=logging.INFO)
         #logging.info('HUC8,BranchID,Exception')
@@ -80,7 +80,7 @@ def Inundate_gms(
     hucCodes = [None] * number_of_branches
     branch_ids = [None] * number_of_branches
        
-
+    
     executor_generator = { 
                 executor.submit(inundate,**inp) : ids for inp,ids in inundate_input_generator 
                 }
@@ -168,7 +168,7 @@ def __inundate_gms_generator(
                               depths_raster,
                               forecast,verbose=False
                             ):
-
+    
     # iterate over branches
     for idx,row in hucs_branches.iterrows():
         

### 2022-08-29 tools/gms_tools/mosaic_inundation.py

@@ -135,7 +135,7 @@ if __name__ == '__main__':
     parser.add_argument('-s','--subset', help='Subset units', required=False,default=None,type=str,nargs='+')
     parser.add_argument('-n','--nodata', help='Inundation Maps', required=False,default=elev_raster_ndv)
     parser.add_argument('-w','--workers', help='Number of Workers', required=False,default=4,type=int)
-    parser.add_argument('-t','--mosaic-attribute', help='Mosaiced inundation Maps', required=False,default=None)
+    parser.add_argument('-t','--mosaic-attribute', help='Mosaiced inundation Maps', required=False,default='inundation_rasters')
     parser.add_argument('-m','--mosaic-output', help='Mosaiced inundation Maps', required=False,default=None)
     parser.add_argument('-r','--remove-inputs', help='Remove original input inundation Maps', required=False,default=False,action='store_true')
     parser.add_argument('-v','--verbose', help='Remove original input inundation Maps', required=False,default=False,action='store_true')

### 2022-08-29 tools/make_dem_skill_results_plot.py

@@ -1,0 +1,377 @@
+#!/usr/bin/env python3
+
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import numpy as np
+import argparse
+from foss_fim.tools.consolidate_metrics import Consolidate_metrics
+from matplotlib.ticker import FixedLocator, FixedFormatter, NullFormatter, FormatStrFormatter
+from statsmodels.robust.robust_linear_model import RLM
+#from foss_fim.tools.gms_tools.plots import preparing_data_for_plotting
+
+
+def Make_results_plot( benchmarks=['all'],versions=['all'],
+                       zones=['total_area'],matching_hucs_only=True,
+                       hucs_of_interest=None,
+                       metrics_output_csv=None,
+                       impute_missing_ms=False,
+                       output_fig=None,
+                       model=None,
+                       nhd_v_3dep=False,
+                       quiet=False
+                     ):
+
+    consolidated_metrics_df,consolidated_secondary_metrics = Consolidate_metrics( 
+                         benchmarks,
+                         versions, 
+                         zones,
+                         matching_hucs_only,
+                         hucs_of_interest,
+                         metrics_output_csv,
+                         impute_missing_ms,
+                         quiet=True
+                        )
+
+    consolidated_metrics_df = __preparing_data_for_plotting(consolidated_metrics_df,nhd_v_3dep)
+    violin_plots(consolidated_metrics_df,nhd_v_3dep,model=model,output_fig=output_fig)
+
+
+def __preparing_data_for_plotting(metrics_table,nhd_v_3dep):
+
+    metrics_table.rename(columns= {'extent_config':'Model',
+                                   'magnitude': 'Magnitude',
+                                   'TPR' : 'POD',
+                                   'huc' : 'HUC'
+                                  },
+                         inplace=True)
+
+    set_mannings = lambda df : 0.06 if 'n_6' in df['version'] else 0.12
+
+    metrics_table['Mannings N'] = metrics_table.apply(set_mannings,axis=1)
+
+    
+    def __set_resolution(metrics_table):
+        
+        if '10m' in metrics_table['version']:
+            resolution = 10
+        elif '15m' in metrics_table['version']:
+            resolution = 15
[…]

### 2022-08-29 tools/run_test_case.py

@@ -29,6 +29,7 @@ def run_alpha_test( fim_run_dir, version, test_id, magnitude,
                 inclusion_area_buffer=0, light_run=False, 
                 overwrite=True, fr_run_dir=None, 
                 gms_workers=1,verbose=False,
+                keep_gms=False,
                 gms_verbose=False
               ):
 
@@ -191,6 +192,12 @@ def run_alpha_test( fim_run_dir, version, test_id, magnitude,
                                             os.path.split(inundation_raster)[1].replace('.tif', '_'+current_huc+'.tif')
                                                     )  
             
+            # assign current huc variable for inundation mapping
+            if huc12:
+                ch = current_huc12
+            else:
+                ch = current_huc
+                    
             try:
                 if model == 'GMS':
                     
@@ -198,7 +205,7 @@ def run_alpha_test( fim_run_dir, version, test_id, magnitude,
                                              hydrofabric_dir=hydrofabric_dir, 
                                              forecast=forecast, 
                                              num_workers=gms_workers,
-                                             hucs=current_huc,
+                                             hucs=ch,
                                              inundation_raster=inundation_raster,
                                              inundation_polygon=None, depths_raster=None,
                                              verbose=gms_verbose,
@@ -213,7 +220,7 @@ def run_alpha_test( fim_run_dir, version, test_id, magnitude,
                                         mosaic_output=inundation_raster,
                                         mask=mask_path_gms,unit_attribute_name='huc8',
                                         nodata=elev_raster_ndv,workers=1,
-                                        remove_inputs=True,
+                                        remove_inputs=(not keep_gms),
                                         subset=None,verbose=verbose
                                       )
                 
@@ -283,7 +290,8 @@ def run_alpha_test( fim_run_dir, version, test_id, magnitude,
                     all_computed_bool = [True if h12 in computed_huc12s_set else False for h12 in all_huc12s_in_current_huc ]
                     all_computed = all(all_computed_bool)
 
-                    print(current_huc,current_huc12,last_huc12,all_computed,magnitude)
+                    if all_computed:
+                        print(current_huc,current_huc12,last_huc12,all_computed,magnitude)
 
                     if all_computed & (last_huc12 == current_huc12):
                         
@@ -367,9 +375,10 @@ if __name__ == '__main__':
[…]

### 2022-08-29 tools/synthesize_test_cases.py

@@ -419,15 +419,18 @@ if __name__ == '__main__':
     # delete version dirs with HUC12
     if overwrite:
         for ch,ata in procs_dict.items():
-            if ata['all_huc12s_in_current_huc'] is not None:
-                path_to_remove = os.path.join( TEST_CASES_DIR,
-                                               ata['test_id'].split('_')[1]+'_test_cases',
-                                               ata['test_id'],
-                                               'testing_versions',ata['version']
-                                              )
-                shutil.rmtree( path_to_remove,
-                              ignore_errors=True
-                             )
+            try:
+                if ata['all_huc12s_in_current_huc'] is not None:
+                    path_to_remove = os.path.join( TEST_CASES_DIR,
+                                                   ata['test_id'].split('_')[1]+'_test_cases',
+                                                   ata['test_id'],
+                                                   'testing_versions',ata['version']
+                                                  )
+                    shutil.rmtree( path_to_remove,
+                                  ignore_errors=True
+                                 )
+            except KeyError:
+                pass
     
     if job_number_huc == 1:
         
@@ -445,7 +448,7 @@ if __name__ == '__main__':
     # Multiprocess alpha test runs.
     if job_number_huc > 1:
         
-        #print(procs_dict)
+        #print(procs_dict);exit()
         executor = ProcessPoolExecutor(max_workers=job_number_huc)
         
         executor_generator = { 

### 2022-08-31 manuscripts/owp_fim4_2022/sections/introduction.tex
* updated references * updated affiliations

@@
 Additional work is required to address the gaps that the AHPS leaves in terms of spatial resolution, spatial coverage, and temporal forecast horizons.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners at the National Center for Atmospheric Research (NCAR) have developed and implemented operationally the NWM which is a configuration of the Weather Research and Forecasting Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners at the National Center for Atmospheric Research (NCAR) have developed and implemented operationally the NWM which is a configuration of the Weather Research and Forecasting Hydrologic Model (WRF-Hydro) \cite{salas2018towards,gochis2018wrf,cosgrove2019evolution}.
 The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including lookback-range (3-28 hrs), short-range (18 hr), medium-range (10 day) and long-range (30 day) forecast horizons.

### 2022-08-31 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
* updated references * updated affiliations
* update WRF reference

@@
 \affiliation{3}{National Water Center, Office of Water Prediction, National Weather Service, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
-\affiliation{4}{New York Water Science Center, Hydrologic Applied Innovations Lab, United States Geological Survey, Troy, NY, USA}
+\affiliation{4}{Hydrologic Applied Innovations Lab, New York Water Science Center, United States Geological Survey, Troy, NY, USA}
 \affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Maryland, College Park, MD, USA}
@@
 Additional work is required to address the gaps that the AHPS leaves in terms of spatial resolution, spatial coverage, and temporal forecast horizons.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners at the National Center for Atmospheric Research (NCAR) have developed and implemented operationally the NWM which is a configuration of the Weather Research and Forecasting Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners at the National Center for Atmospheric Research (NCAR) have developed and implemented operationally the NWM which is a configuration of the Weather Research and Forecasting Hydrologic Model (WRF-Hydro) \cite{salas2018towards,gochis2021wrf,cosgrove2019evolution}.
 The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including lookback-range (3-28 hrs), short-range (18 hr), medium-range (10 day) and long-range (30 day) forecast horizons.
@@
 The term ``hydrofabric'' is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The NWM provides stream forecasts at these hydrofabric segments using the Muskingum-Cunge method to reduce computational requirements of a continental scale model but fails to consider backwater dynamics \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
+The NWM provides stream forecasts at these hydrofabric segments using the Muskingum-Cunge method to reduce computational requirements of a continental scale model but fails to consider backwater dynamics \cite{bedient2008hydrology,ponce1994variable,gochis2021wrf}.
 The need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.

### 2022-09-14 manuscripts/owp_fim4_2022/reviews/first_review/responses.tex
changed rebuttal to responses

% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from 
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
% 
% Licensed under cc by-sa 3.0 with attribution required.
% See https://creativecommons.org/licenses/by-sa/3.0/
% and https://stackoverflow.blog/2009/06/25/attribution-required/

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{apacite}

% bibstyle
\bibliographystyle{apacite}

% import Eq and Section references from the main manuscript where needed
% \usepackage{xr}
% \externaldocument{manuscript}

% package needed for optional arguments
\usepackage{xifthen}
% define counters for reviewers and their points
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% This refines the format of how the reviewer/point reference will appear.
\renewcommand{\thepoint}{P\,\thereviewer.\arabic{point}}

% command declarations for reviewer points and our responses
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
                  \section*{Reviewer \thereviewer}}

\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Reviewer~Point~\thepoint} } ---\ }
   {\par }

\newcommand{\shortpoint}[1]{\refstepcounter{point}  \bigskip \noindent 
	{\textbf{Reviewer~Point~\thepoint} } ---~#1\par }

\newenvironment{reply}
   {\medskip \noindent \begin{sf}\textbf{Reply}:\  }
   {\medskip \end{sf}}

\newcommand{\shortreply}[2][]{\medskip \noindent \begin{sf}\textbf{Reply}:\  #2
	\ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
	\medskip \end{sf}}

\begin{document}

\section*{Response to the reviewers}

\subsection*{Editor Response}

Associate Editor Evaluations: \\
Recommendation (Required): Return to author for major revisions \\
Accurate Key Points: Yes \\
\noindent
Associate Editor (Remarks to Author):

Dear authors,

We have received three very detailed reviews for your manuscript, which suggest that major revisions are needed before your manuscript can be reconsidered for publication in WRR.
All reviewers agree that the manuscript's structure has to be improved and that the method description needs clarification.
Reviewers 1 and 2 highlight the need to better work out the novel contribution of this study, which is hidden behind the report-like structure of this manuscript.
I agree with the reviewers that the manuscript needs a clearer problem statement, more focus on the major contribution, results and discussion.
Some clarification regarding methodology seems to be needed in Section 2.5.
In addition, Reviewer 3 points out that substantial editing is needed before re-submission.
I am looking forward to reading a substantially revised and edited version of this manuscript.\\

Best regards

\begin{reply}
We thank you, the editor, as well as the reviewers for carefully reading the manuscript and for providing valuable feedback to improve its quality.
We have gone through the reviews and made every effort to address the concerns while staying within the scope of the paper.
Substantial revision has led to the removal of superfluous report like information while ensuring adequate background not published in other sources is provided in the manuscript.
This has left some room for better highlighting and explaining the novel contribution of the paper which is countering the ``nearest drainage'' limitation within Height Above Nearest Drainage for flood inundation mapping applications.
We have added several key clarifications in response to reviewer points and several new or updated figures that better explain our methods and results.
There are few concerns that we had to leave out of this manuscript as this submission already covers a very substantial amount of work.
The questions left open concern issues surrounding the synthetic rating curves including localizing Manning's n as well as bathymetry inclusion, estimation, and/or calibration.
We hope this greatly enhanced submission meets the expectations set by WRR and the concerns of our three reviewers adequately.
\end{reply}

% Let's start point-by-point with Reviewer 1
\clearpage
\reviewersection

\subsection*{Introductory Comment}

% Point one description 
\begin{point}
This work improved the accuracy of the height above nearest drainage (HAND) based flood inundation mapping.
This study first compared the original HAND-based flood inundation mapping approach with HEC-RAS 1D results and showed that HAND suffers from a limitation caused by independent neighboring catchments that cannot cross catchment boundaries.
Then, the study proposed a series of terrain analysis steps to resolve this issue.
The most crucial new method compared to other steps that had already been used in previous studies was to reduce the Horton-Strahler stream order of a HAND processing unit down to one.
Overall, I support this work and find it a valuable research area.
The paper was well written, and the work is original in its effort to resolve some issues associated with the HAND-based flood mapping method.
However, I sometimes found it more like a technical report than a research paper.
I appreciate the effort authors took to explain every step of the methodology in detail; however, I hoped to see more results and discussion than learning every detail mentioned in the methodology section.
Also, the main part of the methodology section (section 2.5) requires modification as it does not clearly present the proposed method (please see my major comments below).
I have some major, minor, and editorial points that I hope authors find helpful in revising this manuscript.
\label{pt:intro_1}
\end{point}

% Our reply
\begin{reply}
Thank you for providing an overall assessment of the paper.
We have carefully reviewed your comments and made every effort to clarify or correct any issues you've brought up.
Any references to line numbers, paragraphs, sections, figures, or tables in the proceeding responses to your comments will refer to those in the original PDF document unless otherwise noted.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Major Comments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
My first comment is regarding comparing results with the HEC-RAS model results.
In this comparison, you evaluated the flood inundation extent and not depth.
My question is, how the proposed method affects the estimated flood stage (depth)?
Figure 9a shows that the proposed method resulted in a smaller estimated stage value.
This should then result in getting a smaller HAND-based inundation extent.
But you showed a larger inundated extent in Figure 10.
How do you separate these two effects?
This needs to be at least discussed in your discussion section.
\label{pt:discussion}
\end{point}

\begin{reply}
We thank you very much for making this important point as one of the main purposes of the paper was to introduce how exactly drainage reduction techniques change HAND derived inundation performance.
After reading your comment and our initial discussion section submission, we did notice a missing portion that properly explains how drainage order reduction increases catchment boundary extents and how that drives increased inundation extents and Probability of Detection (POD).
This, as you correctly noted, is a competing effect to the increased catchment size and associated decrease in inundation area due to a downward shift in the synthetic rating curve (SRC).

These competing effects are very difficult to detangle since catchment definitions are used to derive synthetic rating curves (SRCs).
When viewed from the perspective of change in the total inundation area, these two effects can correctly be described as competing since one tends to increase inundation area while the other generally decreases it.
However, when viewed in the paradigm of overall FIM performance, we believe they are complimentary in nature as hinted by the concurrent increase in POD and slight decrease in FAR.

Motivated by this comment, we have completely restructured the Discussion section to make sure your comment is properly highlighted in the paper.
We have inserted some paragraphs that better describe the increase in inundation extents and POD as driven by catchment size enlargement especially around the critical areas described in the paper.
These larger, overlapping catchments better handle riverine inundation from multiple sources as opposed to the traditional single source ``nearest drainage'' that HAND is limited to.
Limited catchment boundaries drive pools of water that are artificially built up without room for the water to naturally spread.
This can be conceptually visualized as a column of water that extends beyond its container's top edge.
This inherent limitation in HAND skews the stage-discharge relationship in addition to limiting the inundation extents.

Overall, accounting for multiple sources of inundation by delineating catchments from independent level paths of unit stream order better accounts for multi-source inundation thus increasing inundation extent (and POD) while also lowering river stage values thus reducing FAR in other areas.
We believe the updated Discussion section better reflects this critical finding of the paper.
\end{reply}

\begin{point}
My second question is that it seems the proposed method reduces the false alarm rate.
This means the underestimation of the original HAND-based flood mapping compared to the FEMA maps is resolved.
However, we know there are uncertainties associated with the FEMA flood extent.
How would you say this reduction of false negatives is really in a right direction when you did not compare results with ground truth data?
\end{point}

\begin{reply}
We acknowledge your point that FEMA, or Base Level Engineering (BLE), flood extents are derived from HEC-RAS 1D models which are themselves subject to a set of assumptions and limitations based on the modeling techniques and data sources employed.
It can be expressed confidently that the use of BLE maps for evaluation does not offer the sort of fidelity that a ground truthed, real-time survey would provide.
However, the BLE does offer many advantages that are useful when evaluating a large-domain flood inundation map derived from HAND.

Despite being an engineering scale model, the BLE is available over a very large domain (49 HUC8s and 9 states) at high spatial resolutions which allows for a broader evaluation of HAND which is being applied to continental scale flood forecasting for our purposes.
Additionally, the BLE offers the availability of tightly spaced cross-sections that provide the streamflows used in making the flood extents.
This high resolution dataset of streamflows and inundation extents provides the opportunity to remove the meteorological, hydrological, and some of the hydraulic assumptions, errors, and limitations from the comparison.
Stream gages and in-situ observations, while clearly higher fidelity in nature, don't offer the spatial coverage that the BLE offers both in terms of flood extents and their associated streamflows that can be used to detangle confounding factors that introduce errors into the evaluation.
Nevertheless, FEMA maps have a storied use in the literature as an evaluation dataset due to some of the advantages and trade-offs listed above \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
Additionally, it is important to note that almost any other source of FIM is subject to errors and limitations including crowd-sourced data, remote sensing observation, and modeled extents.
Unfortunately, flood inundation mapping is an imprecise science that depends on compromises to be made in the search for objective benchmarking.

To address your concern within the paper, we have added some wording in the evaluation section of the draft that better recognizes and discusses the limitations of the BLE HEC-RAS based FIMs.
While we can't assure that the BLE maps are perfect, they are considered more sophisticated than our simplified, zero-physics, continental-scale technique and have been established well in the literature as a common benchmarking model.
Hopefully this properly addresses your concern with the use of the BLE HEC-RAS 1D models for evaluation.
\end{reply}

\begin{point}
My second major comment is that the methodology section tends to be very long, particularly sections 2.1-2.4.
The main contribution of this paper is section 2.5, but it took me as a reader a long time to get to this point and figure out what the paper presents.
Also, section 2.5 is somehow confusing and could be revised and simplified.
For example, in L531, you mentioned you present two successive methods.
Is the first one sub-setting MS?
The NWM main-stem part (2.5.1) seems (at least to me) like an introduction, not something that you did.
Please clarify and revise this part such that a reader can follow the story that connects 2.5.1 and 2.5.2 with the main paragraph in Section 2.5.
Again, this is the most important part of you work. 
\end{point}

\begin{reply}
We respect your opinion that the methodology may seem to take away from the novel scientific contribution of adding multi-source riverine inundation to HAND.
While the major and foremost scientific contribution of the paper is a proposed method to cure the nearest drainage limitation within HAND, we note that a significant amount of work was devised in developing the methods for computing HAND at any scale.
These hydro-conditioning and geospatial techniques can have a significant impact on performance and on the quality of the final operational product.
We found it important to make sure these methods were properly documented within scientific literature for the benefit of the academic and forecasting communities.
Based on your recommendation, we did go through sections 2.1-4 and tried to remove any information that maybe superfluous to the goals and scope of the paper.
This includes much of the stream thalweg breaching technique in Section 2.3.4 that was more than adequately described in the citation it was developed in.

Additionally, we initiated a substantial revamp of Section 2.5 to ensure the main contribution of the paper is properly explained and defended.
The Section NWM MS (2.5.1) does represent methods conducted by us in that we computed HAND for that specific stream network.
Therefore we have elected to keep it in the paper to adequately describe a stream reduction technique that covers 4\% of the country that is of primary concern to river forecasters.
Also, we added an extra figure (Figure 7 in the revised document) that expands upon the explanation in Figure 5. 
This provides the reader with a more in-depth look at how drainage order reduction down to the level path scale expands catchment and inundation extents.
This allows for multiple sources of fluvial inundation to be accounted for in regions where it is applicable.
We believe that this better conveys the main contribution of the paper to the audience.
\end{reply}

\begin{point}
The limitation that the authors attempted to resolve is introduced as a known limitation in the abstract.
Is that really a known limitation?
If yes, did you cite it properly in your literature review?
I believe this was not introduced well as a limitation.
You only pointed out to this in your results (Figure 10).
This is somehow explained in L203-210.
But not very well explained.
This is an important part and the gap your research is trying to resolve so be generous and explain more.
Could you also clarify whether this limitation always results in more overestimation or underestimation? 
\end{point}

\begin{reply}
To better explain this limitation, we included an additional Sub-section (1.4 HAND's Assumptions and Limitations) that better introduces all of the fundamental assumptions HAND employs and their resulting limitations.
Much effort is expelled here to ensure the reader better understands HAND's limitation to only model fluvial inundation from one nearest drainage source thus ignoring other possible fluvial sources of inundation.
We would also like to point out the reference to Figure 3 and the new Figure 7 which also illustrate this problem and proposed solution.
Please reference the explanation in the Methodology Section 2.5.2 on more discussion on the issue and proposed solution.
And lastly, yes, the citations for this known limitation were provided on L241-242 but also included in the new sub-section to ensure the reader is aware of prior literature acknowledging this issue.

With respect to your question about always producing overestimation (False positives) or underestimation (False negatives), we note in the results and discussion sections that you can't say that this problem always leads to FPs or FNs.
It can only lead to FNs if a given catchment boundary is restricting water inundating a neighboring catchment.
This is commonly seen in areas where junctions are present where a reach's drainage area is restricted by the nearest drainage area of neighboring tributaries.
However, as the wall of a catchment boundary begins to fill artificially with water, the lack of spread could cause FPs in other regions.
By reducing the effect of the neighboring boundaries, we see a general reduction in the rating curve thus leading to lower stage values at given discharges.
We believe that the new version of the Discussion section adequately details these competing forces as per your prior recommendation (\ref{pt:discussion}).
\end{reply}

\begin{point}
Links to all datasets introduced in section 2.2 should be included in your manuscript either in the suggested table or within the ``Open Research'' section at the end of your manuscript. 
\end{point}

\begin{reply}
Based on your suggestion we have reviewed the citations to all datasets and software used to ensure they follow the journal's requirements.
In addition to citing the datasets and software properly in their respective sections, 2.1 and 2.2, we have added another column in the datasets table to ensure they are properly cited there as well.
The URLs to all the software and datasets are included as part of the references as indicated by the journal.
We have also added all this information to an Open Research section at the end of the manuscript.
\end{reply}

\begin{point}
Line 331-332: what were the values for buffer distance, smooth drop, and sharp drop you used?
Were they fixed numbers or differ?
How did you choose these values? 
\end{point}

\begin{reply}
The values for buffer distance, smooth drop and sharp drop were set fixed to 70m, 1000m, and 10m, respectively.
These values were selected due to prior research in the citations provided on original Lines 375-377.
They were decided to be keep fixed to limit the scope of the article to its main objectives though we acknowledge that much work can be done to optimize these parameters in the future.
Based on your comment, we have added the important information within this reply into the revised manuscript in the Stream Network Enforcement Section (2.3.1).
\end{reply}

\begin{point}
I believe a diagram showing all the steps and methods you took to prepare your data and method could be helpful.
Your material and data section has about 14 sub-sections collectively.
It is sometimes hard to follow along. 
\end{point}

\begin{reply}
A new Figure 2 was added to the beginning of the Materials and Methods section that gives the reader a high-level overview of the methods employed.
A brief paragraph was added to help explain this new figure as well at the beginning of the methodology.
\end{reply}

\begin{point}
LINE 462-464. Could you explain more about why this is required?
Is it to improve flat areas?
Can't we see any results associated with this error improvement?
Any other studies that observed this that you could cite?
\end{point}

\begin{reply}
We use the native elevations from the NHD DEM in the areas not corresponding to the thalweg or the areas excavated with the AGREE DEM method.
Doing so allows the HAND value within non-pit filled regions to be computed based on its original value and not its hydro-conditioned value.
HAND requires a long series of hydro-conditioning operations that can greatly manipulate elevations to ensure drainage.
We thought it appropriate to use native elevations for this one step to minimize the error associated with those assumptions.
We inherited this technique from Djokic, 2019 which we now cite on that line in the paper.
The explaination as to why we elected to use the native elevations was expanded upon in this section as well.
\end{reply}

\begin{point}
Fig 4. You need to explain this figure in the content more thoroughly.
What are the numbers?
How does one dominate the other?
You mention that the level path method starts from an outlet, so it is probably better to show an outlet in Figure 4.
Please clarify what you mean by ``arbolate sum''? I cannot understand what this summation is.
\end{point}

\begin{reply}
The full explanation for this figure is provided by its in-text reference but based on your recommendation we have expanded this caption to properly describe the figure in place.
\end{reply}

\begin{point}
In equations 5 and 6, When you apply the mosaic method, what is the reason to choose the max?
Is that because you always see the original HAND approach, as shown in Figure 3, underestimating the flood extent?
This may not be true everywhere, or it might be.
Please provide your thoughts on this and add it to the manuscript. 
\end{point}

\begin{reply}
The main scientific contribution of the paper was to allow for multiple sources of fluvial inundation to be considered. 
At a given pixel with multiple fluvial inundation sources, we decided, intuitively, that the maximum contributing depth should be selected.
We respect that flooding is governed by the shallow water equations and these methods are just a simplified proxy for this.
We have updated the paper to include this and invite future researchers to investigate this.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Minor Comments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the short-hand macros for one-liners.
\begin{point}
In the third key point, what do you mean by higher skill inundation?
Please be more specific.
Do you mean more accurate inundation depth or inundation extent?
Same for L39.
Mapping skill itself is not clear.
Does that refer to flood depth or flood extent?
\end{point}

\begin{reply}
The term ``extents'' was added to these two references (key point and original L39) to clarify that we are using flood extents to evaluate for skill.
\end{reply}

\begin{point}
The last line of your plain language summary: Philosophically speaking, if you compared the HAND to a more realistic model and found improvements, what is the point you offer using HAND instead of a more realistic model?
If HAND is not a realistic model, the logic is that we should not use an unrealistic model.
Please revise this.
I agree HAND is a simplified model, but that does not make it unrealistic.
It is realistic because it is based on some terrain physics.
You may want to revise this sentence and use ``physically-based'' rather than ``realistic''.
\end{point}

\begin{reply}
The term ``realistic'' was changed for ``physically-based'' per your suggestion.
\end{reply}

\begin{point}
L121. There is a gap between the geofabric concept you introduced and the Muskingam-Cunge routing method.
You might want to start this sentence like this: The NWM provides stream forecasts at these geofabric segments using the Muskingham-Cunge method. 
\end{point}

\begin{reply}
This suggestion was incorporated but the term ``hydrofabric'' was recycled in favor for the term ``geofabric''.
\end{reply}

\begin{point}
Page 29: It could be better to change the title of subsection 3.1 to ``Flood mapping performance''. 
\end{point}

\begin{reply}
Changed.
\end{reply}

\begin{point}
L671. Please add ``(Table 2)'' at the end of the sentence.
This guides the reader to know what you are referring to immediately. 
\end{point}

\begin{reply}
Due to restructuring of the Results section, this reference was changed to refer to the figure with violin plots instead in the following sentence.
\end{reply}

\begin{point}
Line 315-318: Please clarify what NHD is?
Does that refer to the NHDPlusHR or NHDPlus medium resolution as used in the NWM v2.1?
\end{point}

\begin{reply}
Specified to NHDPlusHR.
\end{reply}

\begin{point}
Line 344-345: This statement requires citation(s).
\end{point}

\begin{reply}
Citations added.
\end{reply}

\begin{point}
I would suggest removing lines 534-535. 
\end{point}

\begin{reply}
Removed.
\end{reply}

\begin{point}
I believe section 2.3.1 is not a DEM hydro conditioning step.
It is more like creating an input (as seed points) that you need when delineating the stream network.
The hydro conditioning of the DEM section should start with levee enforcement (section 2.3.2).
\end{point}

\begin{reply}
Section 2.3.1 is about enforcing the general location and density of the stream network used for HAND.
Additionally, an excavation procedure is detailed that is used to counter flat bathymetric regions in the DEM that can severely distort synthetic rating curves (SRC) and catchment delineations.
We use the term ``hydro-conditioning'', as stated in Section 2.3, as the series of steps required to generate monotonically decreasing elevations, flow directions that fully drain processing regions, and monotonically draining stream network thalweg lines.
Additionally, we enforce the general locations of hydrography-based stream lines as well as levee elvations which we also consider hydro-conditioning.
These are fundamental assumptions and inputs required for HAND, catchment, and SRC calculation.
We believe that these steps should fall under the term hydro-conditioning as we defined it in the paper. 
\end{reply}

\begin{point}
Line 570. What is the logic behind using a 7 km buffer?
Did you first come with this number or after some test?
What is your recommendation or suggestion?
\end{point}

\begin{reply}
We exposed this parameter to the user as we acknowledge it could benefit from future calibration.
We decided to set a generous value to avoid edge effects as well as possibly limiting catchment sizes in regions with lower slope where inundation could extend out some ways.
This large buffer can add significant computational expense to smaller rivers as well as added data storage requirements.
Overall, we recommend future researchers and developers to consider this parameter to better balance skill with computational expenses.
We have added this discussion to the paper.
\end{reply}

\begin{point}
Maybe reporting the percentage of improvements in Table 2 would help better understanding the contribution of your work.
\end{point}

\begin{reply}
Based on your request, we have changed the results for the MS and GMS models to represent the percentage change from the respective metric within the FR method based on the metric, Manning's n, and magnitude combination.
\end{reply}

\begin{point}
Figure 8 contains a lot of information, and readers may require a substantial explanation to understand what these graphs and numbers all shown in a figure mean.
You introduced the figure in the first paragraph of the result section (L672-677).
I suggest you take one example from this figure and walk the reader through the numbers.
For example, what does it mean when the KDE of CSI for the GMS model has the most left-skewed graph compared to the other models?
Is it considered good or bad?
Is it something you had expected?
Another question that arises here is why using a higher Manning's n shifts the KDE graphs up and whether this means results get better or worse.
You discuss this figure on Line 686-690, but I think it could be better to bring this part further up where you first mention Figure 8.
\end{point}

\begin{reply}
Based on your suggestions, we have significantly improved to the Results section.
The first paragraph briefly introduces the results while the second paragraph gives the reader and in-depth description of the previously numbered Figure 8.
We have made sure all elements of the figure are properly described and that a specific sub-figured is dived into more detail to ensure the reader has a proper understanding of the many elements at play here.
In addition to this enhanced description, we have added a significant amount of information in the third paragraph of the results section that begins to dissect the results and the trends observed in the study.
In a related point and in response to your critique of the Discussion section, we have made significant enhancements to the Discussion section that better elaborate on the material and trends presented in this figure.
We hope that these changes fully address your questions and support this figure that is central to the paper's findings.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{FIGURES}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
Fig 1. The NWM FR streams line is hardly visible in the legend. Please revise. 
\end{point}

\begin{reply}
Corrected.
\end{reply}

\begin{point}
Fig 3. Please define MS in the figure caption (Is it mainstem?).
Also, TN, FN, FP, and TP should be described here.
You introduced these terms later in the evaluation section.
A reader has no idea what these mean unless they first read your evaluation section (section 2.7).
Please add flow direction to this map using arrows.
It can help a reader quickly understand the flow numbers on the figure as you talk about in the content.
\end{point}

\begin{reply}
The caption for Figure 3 has been significantly revised addressing your concerns.
Flow direction arrows have been added to the figure as well.
The term main stem was removed from this caption to avoid confusions.
Mainstems (MS) refers to the subset of the NWM FR network that is at or downstream of existing forecast points.
Here main stem was used to refer to the higher stream order segment in the middle of the image.
\end{reply}

\begin{point}
Line 575. This should be Figure 5d not 5c.
\end{point}

\begin{reply}
Changed.
\end{reply}

\begin{point}
Fig 5. How many different color codes (level paths) are there?
Is fig 5a one HUC 8?
Please mention the HUC8 identifier.
\end{point}

\begin{reply}
We have added the HUC8 identifier, 12090301, to the figure's caption.
This particular HUC8 has 372 unique level path IDs which we have also indicated in the caption.
\end{reply}

\begin{point}
Fig 6. It seems that the blue areas are inundated areas for the 0.2\% recurrence flow.
How about the inundated areas for the 0.1\% recurrence flow?
I am asking this because the caption mentions 0.2\% and 0.1\%.
Please clarify and revise the figure and caption accordingly. 
\end{point}

\begin{reply}
Reference to 0.1\% recurrence flow was removed in the caption.
There isn't enough granularity to show both the 0.1\% and 0.2\% flows at this scale as they would both look the same.
The main purpose of the figure is to illustrate the spatial extents and distributions of the BLE validation maps.
\end{reply}

\begin{point}
Fig 7. It could be good to add some labels on the figure indicating where this location is within your study area.
What is the mainstem's name?
Would it be possible to add NWM v2.1 catchments to the map?
I understand this might make the figure busy, but I would like to know how many cross sections falls within a NWM catchment?
Is this the FR or MR NWM v2.1 stream or the one you created (GSM)?
\end{point}

\begin{reply}
These names of the rivers are not present in the NWM V2.1 but instead we had to look at their equivalents in the NHDPlusHR stream line data.
While the figure is quite busy, we agree that showing the NWM catchments does provide the reader some sense as to how many BLE cross sections might be associated with a given NWM forecast identifier.
We have updated the caption to include this new layer and provide some more spatial context.
\end{reply}

\begin{point}
Fig 10. Which recurrence event was used to create this inundation map?
Please add this to the caption and mention it in the content where you explain results in Figure 10.
\end{point}

\begin{reply}
We used the 500 yr magnitude and added this to both the caption and in-text content as well.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{EDITORIAL COMMENTS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
L66. Remove the period before the parenthesis, followed by the citations.
\end{point}

\begin{reply}
Done.
\end{reply}

\begin{point}
L101. NWM is already defined. Just use the abbreviation here. 
\end{point}

\begin{reply}
Done.
\end{reply}

\begin{point}
Title of section 1.5: It could be better to use the Operational Water Prediction (OWP) Flood Inundation Maps (FIM)
\end{point}

\begin{reply}
Done.
\end{reply}

\begin{point}
L160. HPC is already defined. 
\end{point}

\begin{reply}
Addressed.
\end{reply}

\begin{point}
L167. MR is already defined.
\end{point}

\begin{reply}
This is the first occurrence of Medium Resolution (MR) that we are seeing. 
It is possible prior uses were removed during edits.
\end{reply}

\begin{point}
L178. USGS is not yet defined.
\end{point}

\begin{reply}
Done.
\end{reply}

\begin{point}
Line 347: NLD has already been introduced on page 11.
\end{point}

\begin{reply}
Done.
\end{reply}

\begin{point}
Line 524: MS should be first defined here not on line 537.
\end{point}

\begin{reply}
We opted for the use of main segment here to avoid confusion with the Mainstems stream network used throughout the study.
\end{reply}

\begin{point}
Throughout the manuscript, please be consistent with the terms you use.
Use either Manning's n or Manning's N everywhere.
\end{point}

\begin{reply}
We have opted for the use of the lower case, Manning's n.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Begin a new reviewer section
\clearpage
\reviewersection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
This study presents recent development made by federal agencies and their contractors on streamlining HAND-based flood inundation mapping.
By introducing the idea of level paths, they have also tested the hypothesis that unary stream order networks enhance flood mapping performance skills with HAND.
The study presents some solid work, and the results support their statements.
It's important to expose this workflow, which has been adopted by government agencies in daily operation, to the academic community for broader discussion.
I'm a little concerned about the relevance of this draft.
To publish on Water Resources Research, you need first to highlight the problem you want to solve in water science.
Therefore, presenting some research that is a byproduct of tool development won't interest potential readers of this journal.
The authors may consider reorganizing the draft in a more science-driven way instead of technology.
Based on the reasons above, a major revision is recommended for the editors' consideration.
\end{point}

\begin{reply}
Thank you for taking the time to read our paper and providing ample feedback for improving its content.
We understand your concern and have made every effort to highlight the novel contributions of the manuscript which is an operational, continental scale version of HAND that applies the latest datasets and hydro-conditioning methods while considering multiple sources of fluvial inundation instead of just that of the nearest drainage line.
We hope that the revisions completed based on your suggestions and the suggestions of the other two reviewers better highlight the studies' novel findings in a more science-driven format.
\end{reply}

\begin{point}
Line 82-84 Introducing these two NWM networks in section 1.2 makes more sense.
\end{point}

\begin{reply}
To your point, we have added an additional, high-level explaination of the two stream networks for the NWM in this section to provide the reader with a general idea of what these two illustrated networks mean. 
We did leave some additional detail to be discussed in the NWM section since it is more relevant to that section.
\end{reply}

\begin{point}
Line 112 There isn't a figure 1.2.
\end{point}

\begin{reply}
Fixed.
\end{reply}

\begin{point}
Line 128-129 Here, a statement may be added to explain that running a continental-scale hydrodynamic model regularly is not feasible in operational forecasting, making the simplified alternative like HAND a better solution. (Like line 795-798)
\end{point}

\begin{reply}
Added a sentence specifically mentioning that a continental scale hydrodynamic model is computationally prohibitive.
\end{reply}

\begin{point}
Line 212-230 ``In addition to developed tooling, we introduce research'' I don't like how the authors present the study's purpose. This is not a draft submitted to Environmental Modelling \& Software.
Presenting some research that is a byproduct of tool development won't interest the potential readers of this journal.
\label{pt:science_oriented_purpose}
\end{point}

\begin{reply}
Based on your suggestion, we have rewritten this section to better highlight OWP FIM as a research study and a scientific enhancement.
\end{reply}

\begin{point}
Although very comprehensive, the introduction section seems a little bit long.
This is not a HAND-study review paper, so you don't have to go through all the studies.
The authors may consider removing a few sentences to make it more focused.
\end{point}

\begin{reply}
Several sentences in the introduction of AHPS and NWM were removed to only include the most relevant information to this study.
The section introducing OWP FIM was completely revised to address \ref{pt:science_oriented_purpose} and an entire paragraph on some of the superflous HAND-related studies that are mentioned in the Methods later on was also removed.
\end{reply}

\begin{point}
Figure 2 How are the pruned headwaters derived? Why are they moved away from the original NHDHR endpoints?
\end{point}

\begin{reply}
The first paragraph in original section 2.3.1 was meant to describe the pruning of the NHDPlusHR to better match the resolution of the NWM stream network which was originally derived from the NHDPlus network.
In response to your questions, we have revamped this paragraph to offer the reader a better explanation.
Additionally, we have added more information to the caption of original Figure 2 to provide a better explaination in context.
We hope this level of detail adequately answers your questions.
\end{reply}

\begin{point}
Line 513 I don't think ``work for many instances fine'' is formal enough as a sentence in a scientific journal article.
\end{point}

\begin{reply}
This informal phrase was removed and the sentence was merged with the following sentence.
\end{reply}

\begin{point}
Line 532-533 Four sections have been written in the background before the authors finally dive into the method related to the hypothesis of the current draft, which seems too much.
I appreciate the comprehensiveness, but this is a journal article, not a technical document.
\end{point}

\begin{reply}
Yes we do understand that the methods are quite long but these methods are very much central to the study.
We believe that all methods involved in developing the conclusion of the study should be properly explained or cited.
The methods in the four sub-sections you referenced involve a listing of the primary software dependencies used, description of the datasets used, explanation of the hydro-conditioning processes employed, and a detailing of the derivation of the FIM hydrofabric.
All of these methods are central to the conclusions of the study and must be properly explained here since they have not been published previously.
We argue that keeping these sub-sections intact are required to ensure reproducible, transparent, and open research.
\end{reply}

\begin{point}
Line 544 ``only find an MS stream network of stream order 1 (i.e., headwater)'' From Figure 1, we can tell that most AHPS sites are located on big rivers, which means their stream orders are pretty low.
Then why do you say they are headwater links?
\end{point}

\begin{reply}
We agree this sentence is fairly ambigous so we have clarified this to be more clear.
Yes, typically MS river segments are large rivers with higher Horton-Strahler stream order.
Horton-Strahler stream order starts at order 1 for headwaters and compounds downstream until the outlet point.
However, HAND is computed at the HUC8 spatial resolution for the MS version.
When a mainstem river is clipped to a HUC8 segment, it typically will have an effective stream order of 1 within this particular area.
This is relevant to the computation of HAND as in it removes tributaries that may limit catchment boundary sizes and resulting inundation extents.
We have changed the phrasing to reflect ``uniform stream order'' instead which we believe better describes the circumstances.
\end{reply}

\begin{point}
Figure 4 This figure is confusing and needs some revision.
Demonstrating one network is enough to explain the idea of level paths, but there are a few in this plot.
Also, the annotations are too dense, and their alignment needs to be adjusted.
For example, I'm confused when I see two numbers (6.13 and 10.95) on the left purple link.
How come two numbers there?
\end{point}

\begin{reply}
Based on your appropriate feedback, we have elected to show the same figure but in the context of a HUC12.
Illustrating a HUC12 shows the dendritic structure relevant to the explaination of arbolate sums and level paths (LP).
To be clear, the arbolate sums and LPs were derived at the HUC8 scale but for the figure we only show a HUC12 to show fine grain features more clearly.
The numbers in the figure represent arbolate sums which are cumulative drainage distances of all inclusive upstream reaches.
The arbolate sums are now much more clearly labeled at the whole km value for each reach defined as a junction to junction stream segment.
\end{reply}

\begin{point}
Table 2 \& Figure 8 If only staring at these index numbers and plots, the improvement brought by the current study looks relatively marginal.
The authors may look from other angles to see the benefits brought by analyses with unary stream order networks.
Otherwise, the significance of this study does not look very convincing.
\end{point}

\begin{reply}
We understand that a few metric points may seem like a marginal increase.
In order to better highlight the value of the proposed work, we have changed Table 2 to report the percentage improvement for the proposed models, MS and GMS, when compared to the baseline model, FR.
Each percentage improvement corresponds to a specific metric (CSI, POD, and FAR), magnitude (100 and 500 yr), and Manning's n values (0.06 and 0.12).
These percentage improvements for CSI range from just over 2\% for MS to almost 5\% improvement for GMS which we deem convincing.
Additionally, we argue and highlight that these metrics are for evaluations at the HUC8 scale.
HUC8s are very large regions that aggregate away fine grain results and impacts.
This up to 5\% improvement assessed at a much local scale can seem like a significant improvement.
While the OWP is not currently using an impact based data source such as building footprints or land cover/use maps, viewing the enhancements qualitatively and computing the results at this scale could provide an argument that these results can be significant at the impact scale.
We provided original Figure 10 to highlight the enhancement of this improvement at this scale.
To respond to this point, we have inserted some verbage in the Discussion that acknowledges the limitations of large, HUC8, scale evaluations and hints towards future usage of impact datasets and finer grain evaluations to counter this limitation.
For now, we note an absolute improvement in FIM skill by account for multiple fluvial inundation sources by reducing the scale of HAND computation to utilize stream networks of unit Horton-Strahler stream order.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Begin a new reviewer section
\clearpage
\reviewersection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
This paper details the methods employed and findings encountered as the authors endeavored to produce a nationwide flood inundation map (FIM) from National Water Model (NWM) streamflow estimates over the Continental United States.
This is an application of large-domain hydrological modeling.
The relevance of this work to society is not to be understated; a forecast of flood inundation extent and depth from an operational flood forecast model is essential to reducing risks associated with major hydrologic events.
The originality of this work lies in the scale of the problem being addressed and the relevance of operational forecasting to mitigating risk to life and property.

This manuscript adds to the growing body of literature using HAND as a method for determining potential flood inundation in the absence of a full-physics (hydraulic) model.
Ultimately, the process of producing inundation estimates for $>$2.7M stream segments at high spatial resolution ($\sim$10m) is one which required much iteration.
In the end, the authors found that generalizing the stream networks, and using smaller basin areas (HUC8) combined with aggregating individual level-paths yielded better estimates of flood inundation when compared to results from HEC-RAS BLE maps.
Novel methods and computational algorithms were employed along the way, resulting in potentially useful workflows and repositories that readers may follow to produce similar FIMs over the same regions (CONUS).

Although there is a growing body of literature surrounding the use of relative elevation maps such as HAND for flood inundation mapping, this study scales the process to the continental scale and addresses many of the limitations encountered along the way.
Some of the most novel findings in the paper come from the methodological choices employed, such as identifying the appropriate thalweg elevation in the presence of competing channels (DEM-based thalweg vs. hydrography-based thalweg).
Another innovation is in the mosaicking of the FIMs in the presence of tributaries and lower-order streams.
Another very useful innovation lies in ``burning'' levee elevations onto the DEMs to enforce flood control measures on to the HAND surface.
The approach detailed in this manuscript can be seen as a guide for mitigating the 'catchment boundary problem', the importance of which should be emphasized more extensively in this paper.
\label{pt:intro_pt_reviewer_3}
\end{point}

\begin{reply}
We thank you for your overall assessment of the paper and for noting the important contributions this work represents.
While the central scientific contribution of the paper was allowing for multiple fluvial source inundation within HAND based FIMs, we are pleased to hear that the many of the supporting contributions including continental scaling, levee enforcement, open-source software availability, forecasting applications, and thalweg delineation methods were properly conveyed and appreciated.
We hope that the following points in your review are properly addressed to ensure that the main contribution of the paper is properly communicated in a clear and scientific manner to add to the body of literature in this sub-field of hydrology.
\end{reply}

\begin{point}
As a reader, I find the manuscript title and some of the conclusions to be somewhat misdirected.
The idea that Horton-Strahler stream order has a direct impact on the resulting HAND inundation product is misleading.
The stream order of a given reach is determined by the drainage density, and as such is a product of ``upstream'' decisions made during the derivation of the flow network.
Stream order is not an input into the HAND model, nor is it very meaningful when comparing networks that are derived from different headwater locations (such as a pruned version of an existing network such as NHDPlusHR).
Thus, simply reducing stream order should not have any impact on HAND.
There are references in the paper to the impact of drainage density on maximum stream order, but the Conclusions indicate that Horton-Strahler reduction is the reason for improved FIM results, when that reduction is simply an associated result of reduced stream network density.
For example, the authors indicate ``We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed...''.
In reality, the reduction Horton-Strahler stream order is the result of, and not the cause for, a reduction in drainage density.
It may be useful to emphasize that the discretization of the network into coherent regions based on level-path or the regionalization methods (such as using HUC8 boundaries), as well as the priority queue calculation (Equation 6) during the HAND raster mosaic process has a bigger impact on the result than the Horton-Strahler number of the network employed.
\end{point}

\begin{reply}
We acknowledge that the original messaging in the paper maybe confusing.
We note in this reply that as described in the methodology surrounding datasets and stream network enforcement as well as your mention \ref{pt:intro_pt_reviewer_3}, stream network is indeed used as an input in our process to compute HAND.
We differ from previous studies mentioned in the paper that utilize accumulated drainage pixels across the entire processing area then threshold those accumulations to produce a stream thalweg network.
In our methods, we consider a forecast network, that of the NWM, and a network that better agrees with the DEM employed, NHDPlusHR Burnlines, to help enforce drainage in the processing unit which is one of the fundamental assumptions and requirements of HAND.
We then use select headwater points to seed a new stream thalweg network that is heavily influenced by the burning and conditioning operations inspired by the input networks, NWM and NHDPlusHR.
The thalweg datum used to detrend or normalize elevations away from mean sea level to that of nearest drainage is thus a product of input stream networks.

Where we believe our explaination lacked, and with respect to your point, is we failed to properly communicate that discretizing the Full Resolution NWM stream network into either Mainstems (MS) network or independent level paths doesn't change the maximum Horton-Strahler stream order or the drainage density.
It does, however, change the effective maximum stream order used in each independent computation of HAND.
This distinction is not properly conveyed in the paper and could lead readers to be confused as to what is actually happening.
In response to this ambigiouty, we have added language in the Stream Order Reduction sub-section of the Methods to ensure this distincition is more clear.

Most notably, we propose a title change to the paper to ensure that the primary objective of considering multi-source fluvial inundation is properly highlighted.
The new proposed title ``Extending Height Above Nearest Drainage to Model Multiple Fluvial Sources in Flood Inundation Mapping Applications for the U.S. National Water Model'' better accounts for the primary purpose of the paper while not allowing an ambigouty to settle in without proper explanation to occur first.
We hope these changes better account for the fundamental contributions of the paper and convey them in a much more clear manner.
\end{reply}

\begin{point}
Another limitation in the methodology lies in using synthetic rating curves (SRCs) that result from reach-averaged geometry parameter derived from the AGREE-DEM (a hydro-conditioning process).
This results in an artificially impacted cross-section and bathymetry structure.
The parameters used to constrain the AGREE method will have a profound impact on the SRCs derived from that DEM due to the excavation process in the AGREE method.
\end{point}

\begin{reply}
We agree with you that the AGREE DEM proceduce does introduce a significant amount of bathymetric manipulation that affects SRC and FIM extent quality.
To your point, we have noted this limitation in multiple sections throughout the paper including the methods, discussion, and conclusions to ensure that the reader is fully aware of this limitation and its effects.
We do motivate future work in this realm as SRC quality is of primary importance to FIM skill.
We invite future researchers to investigate factors that contribute to SRC quality including Manning's n localization as well as bathymetry assimiliation, estimation, and calibration.
Unfortunately, the scope of this article is already too extensive to go much further.
\end{reply}

\begin{point}
Further revision could include a more thorough description of how using level-paths eliminates the catchment boundary problem, when each level path may have many tributaries (and thus multiple catchment boundaries).
There is a bit of missing information in Section 2.5.2 that is necessary for understanding how the catchment boundary problem is mitigated.
It seems that the way around the catchment boundary problem is to calculate HAND for a given level-path while leaving tributaries out of the analysis, and allowing the mosaic process to result in a HAND/FIM result with more extensive coverage over catchment boundary areas.
\end{point}

\begin{reply}
We agree that the original submission lacked sufficient detail as to how the catchment boundary problem is mitigated by level path disaggregation for HAND computation.
We have completely revamped this section to ensure the core contribution of this paper is properly explained.
Additionally, we have introduced a new Figure 7 that better describes how the catchments are delineated at the level path scale as well as the associated inundation.
We hope these considerable changes better communicate this critical section of the paper.
\end{reply}

\begin{point}
The manuscript is, in places, difficult to understand due in part to the large number of acronyms and persistent jargon that pervades it.
There are also many grammatical and syntactical errors that should be edited before re-submission.
The attached annotated manuscript contains many suggestions for resolving these issues along with other comments and feedback, which the authors may find helpful.
For this reason and those outlined above, I suggest to return the manuscript to the author for these (minor) revisions to increase the clarity of the methods and significance of the methodological choices on the results.
\end{point}

\begin{reply}
We have gone through your attached annotated manuscript and have made every effort to address every concern listed.
For any annotations that require an additional response, we provide those below referencing the original line numbers those annotations occur. 
\end{reply}

\begin{point}
L98 - Does the AHPS have a coarser resolution than NWM?
\end{point}

\begin{reply}
AHPS only has two forecast horizons while NWM has three forecast horizons and one analysis product (lookback).  
These are now accounted for in the new manuscript.
\end{reply}

\begin{point}
L368 - 375:
Why does this need to be done?
Is there often a conflict after running AGREE between the DEM thalweg and the AGREE-DEM thalweg?
Depending on the smooth-drop parameter used in AGREE, I wouldn't think there would be any parallel channels in the resulting DEM.
\end{point}

\begin{reply}
The citations in the manuscript refer to the parallel streams problem that arises by using AGREE DEM.
The AGREE DEM enforces a hydrography based streamline and associated notch while not considering some semblemce of the thalweg that might be occuring natively in the DEM.
This creates the notion of a parallel stream when comparing two different DEMs, the AGREE one and the native, unconditioned one.
Additionally, this step provides native elevations to be fed into the thalweg breaching algorithm as opposed to the vastly artificial values in the sharp drop portion of AGREE.
To your point, we have attempted to better explain this in the manuscript.
\end{reply}

\begin{point}
L454: Couldn't you just use the value from the flow accumulation grid at each thalweg pixel rather than derive catchments at each thalweg pixel?
\end{point}

\begin{reply}
The term ``contributing area'' is ambigious here.
We seek to clarify that we are not computing the contributing area in some square km value.
We are just trying the derive catchments or unique drainage polygons for two separate circumstances.
We have changed the language here to bring about more clarity.
\end{reply}

\begin{point}
L478: perimeter?
\end{point}

\begin{reply}
In the reach averaged version of the Manning's Equation, wetted perimeter in the traditional Manning's Equation is translated to bed area given the consideration to the reach length, L, in the formula as well.
\end{reply}

\begin{point}
L491: 0.05 does not appear to be used later. How did the authors decide to use 0.06 and 0.12?
\end{point}

\begin{reply}
In our evaluations, we have noted better general performance with higher Manning's n values thus selected the value of 0.06 from the Djokic work as our baseline.
\end{reply}

\begin{point}
L498: NWM channel parameters, for consistency with the source model? These parameters exhibit variability and have been evaluated during mdoel testing.
\end{point}

\begin{reply}
We tried those parameters and unfortunately they do not yield better results in terms of FIM.
It's possible that those regionalized values are not optimal for FIM purposes as they are tuned more for 1D hydralic routing purposes.
It's likely that lack of accurate bathymetric information interacts heavily with the optimal Manning's n value which is only supposed to account for roughness.
We are keen on working more with the combined effects of bathymetry and roughness parameters to optimize the SRCs for better FIM performance.
In the meantime, we focused on the core contributions of this paper already presented.
\end{reply}

\begin{point}
L684:  This is a key finding of this research and could be emphasized more perhaps in the abstract or conclusions.
\end{point}

\begin{reply}
This key finding has now been further elaborated on in the conclusions and abstract sections.
\end{reply}

\begin{point}
L765-767: I thought lower stream order produced better FIM maps?
\end{point}

\begin{reply}
This sentence is referring to an issue associated with conflation or cross-walking of the FR version of HAND.
This has now been specified so the reader understands that it is likely improved with the new GMS version that ignores neighboring tributaries thus less susceptible to cross-walking errors.
\end{reply}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% references and end document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage % this clears figures before references
\bibliography{responses}

\end{document}

### 2022-09-14 manuscripts/owp_fim4_2022/reviews/first_review/makefile_responses.sh

@@ -3,7 +3,7 @@
 #parent_path=$( cd "$(dirname "${BASH_SOURCE[0]}")" ; pwd -P )
 #cd $parent_path
 
-fullFile=rebuttal.tex
+fullFile=responses.tex
 baseName="${fullFile%%.*}"
 extension="${fullFile#*.}"
 

### 2022-09-15 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
* latest updates * possible resubmission candidate

@@
 The National Water Model (NWM) currently requires additional modeling from discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
-Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is worthy of producing high-resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
+Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is a means worthy of producing high resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
 However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
-We propose here that HAND is limited to producing inundation only when sourced from its nearest drainage line thus lacks the ability to source inundation from multiple fluvial sources.
-To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
+We propose here that HAND is limited to producing inundation only when sourced from its nearest drainage line, thus lacks the ability to source inundation from multiple fluvial sources.
+To counter this constraint, a version of HAND, known as Generalized Mainstems (GMS), is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
 LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
 The FIMs associated with each independent LP are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
-Improvement in mapping skill is observed when the extents are compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
+Improvement in mapping skill is observed by significantly reducing false negatives at river junctions when the inundation extents are compared to FIMs from the Hydrologic Engineering Center's River Analysis System (HEC-RAS).
 A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
 We observe that the improvement from extending drainage order reduction techniques from 4-5\% of the entire stream network to 100\% of the network is about the same magnitude improvement as going from no drainage order reduction to 4-5\% of the network.
-Thus, we conclude that drainage order reduction techniques circumvent the nearest drainage limitation while offering diminishing returns to skill improvement and penalizing with additional computational expenses.
+Thus, we conclude that drainage order reduction techniques circumvent the nearest drainage limitation while offering diminishing returns to skill improvement.
+Drainage order reduction techniques also penalize with additional computational expenses in terms of processing and storage.
 This novel contribution is proposed and studied in a new open-source implementation that utilizes the latest combination of hydro-conditioning techniques to enforce drainage and counter limitations in the elevation data.
@@
 In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance because only fluvial inundation, or inundation sourced from rivers, is only considered from the nearest drainage line.
-As rivers flood, multiple neighboring rivers could contribute to a regions inundation and HAND has no means of modeling for that.
+This model, known as Height Above Nearest Drainage (HAND), suffers from issues in mapping performance because only fluvial inundation, or inundation sourced from rivers, is considered from the nearest drainage line.
+As rivers flood, multiple neighboring rivers could contribute to a region's inundation for which HAND has no means of modeling for.
 We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-This is done by splitting the stream network into continuous rivers known as level paths.
-These level paths have no tributaries thus are known to be stream lines with a unit stream order indicating no branching.
-HAND is computed independently for each level path and the resulting FIMs are mosaiced together to form one seamless map.
+This is done by splitting the stream network into continuous river segments known as level paths (LPs).
+These LPs have no tributaries, thus are known to be stream lines with a unit stream order indicating no branching.
+HAND is computed independently for each LP and the resulting FIMs are mosaiced together to form one seamless map.
 We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance to counter one of HAND's fundamental limitations.
@@
 As a complement to the operational NWM, RFC forecasts from AHPS forecast points are assimilated in the NWM and routed downstream to the next downstream AHPS forecast point where the process iterates again.
-This assimilation into NWM is used to enhance forecasting skill by leveraging best available regional-scale forecasts.
+This assimilation into the NWM is used to enhance forecasting skill by leveraging best available regional-scale forecasts.
 The river network upon which this special assimilation technique operates on is herein referred to as the Mainstem (MS) stream network.
@@
 The NWM provides stream forecasts at these hydrofabric segments using the Muskingum-Cunge method to reduce computational requirements of a continental scale model but fails to consider backwater dynamics \cite{bedient2008hydrology,ponce1994variable,gochis2021wrf}.
-The need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
+The need for high resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
 The use of a 2-dimensional (2D) hydrodynamic model across a continental-scale and high spatial resolutions is very cost prohibitive especially in an operational setting.
@@
 \citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-However, a core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
-The terrain index also provides additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
+The terrain index also provides additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel,aristizabal2021mapping}.
 \citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as SRCs by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
@@
 Since elevations don't naturally do this, they must undergo a long series of hydro-conditioning processes to enforce monotonically decreasing elevations across an entire processing unit along with hydrologically correct directions of flow \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,donchyts2016global,renno2008hand}.
-The level of DEM manipulation required to enforce this assumption can be substantial depending on the region and can be a significant source of error.
+The level of digital elevation map (DEM) manipulation required to enforce this assumption can be substantial depending on the region and can be a significant source of error.
 The drainage enforcing assumption also interacts with an inability to properly account for fluvial inundation in regions of DEM depressions that lack natural drainage to riverine areas \cite{nobre2016hand,renno2008hand}.
@@
 %
-Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (approximately 10 m depending on latitude) scale, HAND has been implemented into software for large-scale, continental computation.
+Due to significant advances in high performance computing (HPC) and large scale high resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (approximately 10 m depending on latitude) scale, HAND has been implemented into software for large-scale, continental computation.
 As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a HPC \cite{maidment2017conceptual,liu2016cybergis}.
-NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
-By allocating 31 nodes at 20 cores per node for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
+NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 central processing unit (CPU) years.
+By allocating 31 nodes at 20 cores per node for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2 terrabyte (TB) of peak memory and 5 TB of total disk space.
 Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (NHDPlusHR) which better agrees with the 10 m DEM from the NHDPlusHR program \cite{liu2020height}.
@@
 Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to the stream accumulation threshold which is closely related to stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths (LP), can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
+Here we demonstrate how reducing a HAND processing unit's stream network into discrete level paths (LPs) of singular, effective stream order, can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
 This capability is introduced progressively as MS (whose network represents about 4\% of FR network) and to a higher degree Generalized Mainstems (GMS) (covers entire FR network) which will be explained later on.
-The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
+The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Materials and Methods}
+\label{sec:materials_and_methods}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
-Figure \ref{fig:methods_overview} gives a high-level overview of the methodology used in OWP FIM and in this study.
+Figure \ref{fig:methods_overview} gives a high level overview of the methodology used in OWP FIM and in this study.
 Input data from multiple sources are preprocessed (not illustrated) and then subset to processing areas based on the model used, FR, MS, or GMS.
-The standard processing unit of OWP FIM is a HUC8 however for the MS and GMS versions only the NWM MS stream network and the NWM FR network discretized into LPs are used, respectively.
+The standard processing unit of OWP FIM is a HUC8 and the entire NWM FR stream network is used for enforcment.
+Later, we explain how only the NWM MS stream network is used for the MS version of HAND, while for GMS, the FR stream network is discretized into LPs before computing HAND.
 A series of hydro-conditioning steps enforces the location of stream lines, monotonically decreasing elevations, excavated bathymetry, stream thalweg breaching, and levee enforcement.
@@
 \caption{
-Methodology overview detailing high-level steps followed in the study.
+Methodology overview detailing high level steps followed in the study.
 The flow chart begins with the input data organized by source.
@@
 Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems, OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
+To simplify setup and enhance portability across host operating systems, OWP FIM packages all dependencies up in a Docker image \cite{merkel2014docker}. 
 A user only needs to install Docker on their host machine and build the image from the provided recipe. 
@@
 These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
-HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
-In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligible for flooding.
+HAND inherently requires all areas eligible for inundation to drain to the designated drainage network.
+So to satisfy this requirment, DEMs must undergo significant manipulation.
+In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations to enable eligiblity for flooding.
 Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
@@
 Headwater points are first derived for the NWM FR V2.1.
-For every NWM headwater point, the nearest NHDPlusHR point is selected and placed into a set and those not belonging to this nearest neighbor set are discarded.
+For every NWM headwater point, the nearest NHDPlusHR point is selected and placed into a set while those excluded are discarded.
 Only the nearest point on the NHDPlusHR is used so any portion of the NHDPlusHR network upstream of this nearest point is discarded to avoid extending inundation too far above the modeling domain.
@@
 \caption{This figure illustrates some of the datasets that result from the pruning of NHDPlusHR Beta BurnlineEvents (dotted black) to the stream density of the NWM FR V2.1 density (blue).
-The stream network used for forecasting, NWM FR V2.1, is of higher stream density than that of the NHDPlusHR which has better agreement with the thalweg locations in the DEM used.
-Thus we opt to prune the NHDPlusHR network to match the general location and density of the NWM network.
+The stream network used for forecasting, NWM FR V2.1, is of lower stream density than that of the NHDPlusHR which has better agreement with the thalweg locations in the DEM used.
+Thus, we opt to prune the NHDPlusHR network to match the general location and density of the NWM network.
 The nearest neighbor segment in the NHDPlusHR of each NWM headwater locations and the nearest point on that segment is determined to match the closest point to that of the NWM headwater.
@@
 These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but not limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
+We note significant opportunity to enhance CFIM skill by better localizing Manning's n according to available data including but not limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
 For now and for the purpose of this study, we examine the SRCs with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
 After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values (y in Eq.
-\ref{eq:reach_averaged_mannings_equation}) from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
+\ref{eq:reach_averaged_mannings_equation}) from 0 to 25 meters in depth at a third of a meter increments to calculate the discharge values for each stage value.
 %
@@
 In this technique, we derive HAND using the FR stream network as well as the MS network which was originally proposed by \citeA{djokic2019arc}.
-Inundation is derived independently from the resulting FR and MS HAND hydrofabric are mosaiced together using the technique proposed in Section \ref{ssec:inundation_mapping} to form the MS FIMs. 
+Inundation is derived independently from the resulting FR and MS HAND hydrofabrics and are mosaiced together using the technique proposed in Section \ref{ssec:inundation_mapping} to form the MS FIMs. 
 Within each HUC, one might typically only find a MS stream network of uniform stream order but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
@@
 Figure \ref{fig:level_path_methods} illustrates how LPs (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-The figure shows computed arbolate sums and unique LP identifiers for the HUC12 120903010404 for clarity but were computed at the corresponding HUC8.
+The figure shows computed arbolate sums and unique LP identifiers on a HUC12 (120903010404) for clarity but were computed at the corresponding HUC8.
 The mainstem of the figure runs from the red ellipses to the black one which is the outlet.
@@
 
-For a more intimate look at the drainage order reduction procedure and its effects, GMS, we allude to Figure \ref{fig:gms_methods_2} which references the same area (in HUC8 12090301) and set of river junctions as in Figure \ref{fig:catchment_boundaries_issue}.
+For a more intimate look at the drainage order reduction procedure GMS, and its effects, we allude to Figure \ref{fig:gms_methods_2} which references the same area (in HUC8 12090301) and set of river junctions as in Figure \ref{fig:catchment_boundaries_issue}.
 The catchments and stream lines for HAND computed at the FR scale are illustrated in Figure \ref{fig:gms_methods_2}a where the respective inundation at the 100 yr magnitude is heavily constrained by the limited catchment extents especially at junctions.
 In subsequent sub-figures, we show the same datasets for the HAND computation problem for this region but discretized into independent LPs for the main LP (b), the eastern tributary (c), and the western tributary (d).
-Notably, inspecting (b) one sees how removing the tributaries creates much larger catchments for the main LP. 
+Notably, inspecting (b), one sees how removing the tributaries creates much larger catchments for the main LP. 
 These catchments include drainage areas that would traditionally be considered nearest to the tributaries thus ineligible to receive inundation sourced from the main LP.
@@
 \includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates the NWM Full Resolution V2.1 stream network discretized into LPs (LP), symbolized by unique colors, as well as the values of the arbolate sums to the nearest whole km distance.
+\caption{Illustrates the NWM Full Resolution V2.1 stream network discretized into level paths (LP), symbolized by unique colors, as well as the values of the arbolate sums to the nearest whole km distance.
 The LPs were derived on a HUC8 level (12090301) but only illustrated for a HUC12 (120903010404) for clarity.
@@
 %
-Possible benchmark FIM candidates for evaluation purposes include high-water marks, remote sensing observation, crowd-sourced information, and modeled extents.
-Unfortunately, all of these sources or using a combination of them are subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
-While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to isolating out hydrological factors.
-A suitable dataset needed to be selected to best balance advantages and limitations with future studies interested in expanding to additional datasets as able to.
-
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
-This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, more sophistication when compared to HAND based FIMs, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
-49 HUC8s spanning about 185 thousand $km^2$ were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM and thus corresponding discharges and mapping extents for evaluation.
+Possible benchmark FIM candidates for evaluation purposes include high water marks, remote sensing observation, crowd-sourced information, and modeled extents.
+These sources are all subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to a lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
+While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to as to isolate out other hydrological factors.
+
+Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
+This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, higher level of sophistication when compared to HAND, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
+We selected 49 available HUC8s, shown in Figure \ref{fig:all_ble_maps}, which span about 185 thousand $km^2$ across nine states.
+The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM as well as the corresponding discharges and mapping extents for evaluation.
 We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation sourced from OWP FIM.
 
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, and others that we would have potentially affected our conclusions if we used NWM forecasted discharges.
+By using the same HEC-RAS derived discharges and FIM extents for creating maps with OWP FIM, we are able to separate out errors introduced by NWM inputs and processes including land surface interactions, groundwater fluxes, atmospheric forcings, hydraulic routing, and others that would have potentially affected our conclusions if we had used NWM forecasted discharges.
 Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
@@
 This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
+Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with unit Horton-Strahler order due to the lack of representation of these additional headwater streams in the NWM network.
 While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
@@
 The distribution of these flood extent metrics can be examined in Figure \ref{fig:violin_plot} as violin plots.
-Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
+Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), Manning's n value (0.06, 0.12), recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
 For example, let's examine the violin plot for the row marked CSI and column for Manning's n = 0.06.
 This sub-figure shows the CSI distributions across all 49 HUC8s when Manning's n is set to 0.06.
-Each independent, whole violin represents the distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100 yr and orange for 500 yr).
+Each independent, whole violin represents the HUC8 metric value distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100 yr and orange for 500 yr).
 The horizontal dashed and dotted lines represent the 25th, 50th, and 75th percentiles from bottom to top, respectively.
@@
 The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}).
-So overall and in other words, the entire distribution of HUC8s improves across the three methods.
+So overall and in other words, the broader distribution of HUC8s improves across the three methods.
 Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit stream order and mosaic the resulting FIMs, the better those FIMs perform.
@@
 Additional noteworthy trends in Figure \ref{fig:violin_plot} center around the all-around better performance of FIMs for those of higher Manning's n values and recurrence flows.
-The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value used in the BLE model, 0.11 \cite{fema2021base,fema2021estimated}.
-Most of this improvement is driven by significant increases in POD but, unfortunately, leading to a significant amount of over-prediction as observed by the increase in FAR.
+The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value of 0.11 used in the BLE model \cite{fema2021base,fema2021estimated}.
+Most of this improvement is driven by significant increases in POD, but unfortunately, it also leads to a significant amount of over-prediction as observed by the increase in FAR.
 More work can be invested to better regionalize Manning's n values for FIM purposes with HAND.
@@
 \centering
-\includegraphics[scale=0.9]{figures/violin_plots.jpg}
+\includegraphics[scale=0.8]{figures/violin_plots.jpg}
 \caption{Shows kernel density estimation of the distributions (sample size = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
 The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's n (0.06 and 0.12).
-Trend lines for each metric - Mannings combination are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}
+Trend lines for each combination of metric and Manning's n are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}
 \label{fig:violin_plot}
@@
 \multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{\shortstack{MS\\(\% Change)}} & \multicolumn{2}{|c|}{\shortstack{GMS\\(\% Change)}} \\
+%\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{\shortstack{MS\\(\% Change)}} & \multicolumn{2}{|c|}{\shortstack{GMS\\(\% Change)}} \\
 \cline{3-8}
@@
 For our work, we were able to produce HAND at the full NWM resolution in 0.13 CPU years which represents a substantial speed-up compared to previous works.
-For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
+For the MS resolution, an additional 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
 GMS which generalizes HAND production to the LP scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
@@
 Most of this change is accounted for by substantially increasing POD and inundation extents in some areas thus reducing FNs.
-We believe, as we later argue for, that the increase in POD is primarily driven by an increase in the catchment sizes that is inherently enabled by dividing up the stream network into independent stream networks of unit stream order.
+We believe, as we later argue, that the increase in POD is primarily driven by an increase in the catchment sizes that is inherently enabled by dividing up the stream network into independent stream networks of unit stream order.
 Additionally, we note that reducing drainage order also has some minor influence on reducing inundation extents in other areas and the rate of false alarms.
-We believe that this effect is driven by a change in the stage-discharge relationship where a given streamflow now leads to lower river stage values when HAND is computed with target stream networks of unit drainage order.
+We believe that this effect is driven by a change in the stage-discharge relationship where a given streamflow leads to lower river stage values when HAND is computed with target stream networks of unit drainage order.
 We seek to explain that these two effects, catchment boundary enlargement and stage-discharge curve lowering, are highly interrelated and cannot be easily detangled.
@@
 This allows for inundation extents to expand across regions previously encapsulated by catchments of joining reaches in lower flow tributaries.
-The water built up along the catchment boundaries can be thought of as a column of water in a cylindrical container (catchment) that has exceeded the elevation of the containers rim.
-This, of course, does not represent accurate physics but it is what does occur in HAND when water accumulates along catchment boundaries.
+The water built up along the catchment boundaries can be thought of as a column of water in a cylindrical container (catchment) that has exceeded the elevation of the container's rim which does not represent accurate physics.
 
@@
 Future assessments of OWP FIM should consider finer grain evaluation units as well possible impact assessments using asset information such as building footprints to better illustrate fine grain changes in a more relevant manner to stakeholders.
-For now, we offer up Figure \ref{fig:gms_enhancement} which best illustrates the improvement offered by multi-source fluvial flooding capabilities in a more local context.
+For now, we provide Figure \ref{fig:gms_enhancement} which best illustrates the improvement offered by multi-source fluvial flooding capabilities in a more local context.
 The figure is comprised of two agreement rasters for two different HAND based FIMs compared to the validation dataset for a given region with a high flow mainstem (500 yr recurrence flow) running horizontally along the region.
@@
 The consequence for inundation extents is a general increase in spatial coverage of the river's water which shows to be in much better agreement with the benchmark map.
-The TPs are no longer bounded by the catchment lines and allowed the expand to their natural extents.
+The TPs are no longer bounded by the catchment lines and allowed to expand to their natural extents.
 
@@
 Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
-We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the SRC Q up and changes the SRC accordingly.
+We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the Q up and changes the SRC accordingly.
 In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
@@
 This geometry is not perpendicular to the flow that is associated with the main reach thus leading to biases in the SRC.
-We consider this fact to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
-
-We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
+We consider this to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
+
+We note that reducing stream order does suffer from diminishing returns where the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
 This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
-
-Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
+One option could be removing lower stream orders (e.g. 1 and 2) from stream order reduction and simply using the inundation from FR from these areas.
+
+Additional analysis of Figure \ref{fig:gms_enhancement}a, reveals that some catchments don't have inundation or significant inundation.
 While the cause of these errors can be varied, we assert here that conflating four networks for use in evaluations leads to significant error.
-As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
+Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
 One of the issues with the FR version of HAND occurs when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
@@
 In our evaluation of this technique, we observe that HAND based FIM improves in skill as we extend from nearest drainage inundation in FR to multiple drainage support in MS for only 4-5\% of the FR network.
-Extending multiple drainage support to the entire FR network with GMS based HAND improves upon skill at around the same magnitude that MS improved upon FR.
+Extending multiple drainage support to the entire FR network with GMS based HAND improves skill at around the same magnitude that MS improved upon FR.
 Thus we conclude that deriving HAND with independent stream networks of unit Horton-Strahler stream order enhances the skill of FIM but offers diminishing returns as we extend from 4-5\% of the network with MS to 100\% of the network with GMS since deriving HAND and FIMs at these localized scale does add computational expense.
@@
 %
-National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
+National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs \cite{nwm2022hydrofabric}.
+These are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
 OWP Flood Inundation Mapping (FIM) capabilities rely extensively on the use of the National Hydrography Plus High Resolution (NHDPlusHR) datasets including BurnLineEvents \cite{nhdplus2022vectors}, value-added attributes (VAA) \cite{nhdplus2022vectors}, water boundaries (WBD) or hydrologic unit code (HUC) geometries \cite{nhdplus2022wbd}, and digital elevation maps (DEM) \cite{nhdplus2022dems}.

### 2022-09-15 manuscripts/owp_fim4_2022/sections/discussion.tex
* latest updates * possible resubmission candidate

@@
 This geometry is not perpendicular to the flow that is associated with the main reach thus leading to biases in the SRC.
-We consider this fact to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
+We consider this to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
 
-We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
+We note that reducing stream order does suffer from diminishing returns where the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
 This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
+One option could be removing lower stream orders (e.g. 1 and 2) from stream order reduction and simply using the inundation from FR from these areas.
 
-Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
+Additional analysis of Figure \ref{fig:gms_enhancement}a, reveals that some catchments don't have inundation or significant inundation.
 While the cause of these errors can be varied, we assert here that conflating four networks for use in evaluations leads to significant error.
-As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
+Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
 One of the issues with the FR version of HAND occurs when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.

### 2022-09-15 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
* latest updates * possible resubmission candidate

@@
 %
-Possible benchmark FIM candidates for evaluation purposes include high-water marks, remote sensing observation, crowd-sourced information, and modeled extents.
-Unfortunately, all of these sources or using a combination of them are subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
-While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to isolating out hydrological factors.
-A suitable dataset needed to be selected to best balance advantages and limitations with future studies interested in expanding to additional datasets as able to.
+Possible benchmark FIM candidates for evaluation purposes include high water marks, remote sensing observation, crowd-sourced information, and modeled extents.
+These sources are all subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to a lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
+While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to as to isolate out other hydrological factors.
 
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
-This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, more sophistication when compared to HAND based FIMs, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
-49 HUC8s spanning about 185 thousand $km^2$ were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM and thus corresponding discharges and mapping extents for evaluation.
+Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
+This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, higher level of sophistication when compared to HAND, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
+We selected 49 available HUC8s, shown in Figure \ref{fig:all_ble_maps}, which span about 185 thousand $km^2$ across nine states.
+The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM as well as the corresponding discharges and mapping extents for evaluation.
 We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation sourced from OWP FIM.
 
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, and others that we would have potentially affected our conclusions if we used NWM forecasted discharges.
+By using the same HEC-RAS derived discharges and FIM extents for creating maps with OWP FIM, we are able to separate out errors introduced by NWM inputs and processes including land surface interactions, groundwater fluxes, atmospheric forcings, hydraulic routing, and others that would have potentially affected our conclusions if we had used NWM forecasted discharges.
 Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.

### 2022-09-15 manuscripts/owp_fim4_2022/docker_run_and_make.sh

@@ -2,4 +2,4 @@
 
 codeDir=$1
 
-docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex:20220201 /foss_fim/manuscripts/owp_fim4_2022/Makefile.sh
+docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex:20220821 /foss_fim/manuscripts/owp_fim4_2022/Makefile.sh

### 2022-09-16 manuscripts/owp_fim4_2022/Makefile.sh
Version sent to JJ

@@ -14,7 +14,7 @@ pdflatex "$fullFile"
 
 # expands all tex files into one file
 echo 'Lat-Expand to Create One Tex File ...'
-/usr/bin/latexpand/latexpand --keep-comments --biber bibliography/owp_fim4_2021.bib owp_fim4.tex -o owp_fim4_expanded.tex
+latexpand --keep-comments --biber bibliography/owp_fim4_2021.bib owp_fim4.tex -o owp_fim4_expanded.tex
 
 
 

### 2022-09-19 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
Resubmitted to WRR

@@
 \begin{keypoints}
-\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to a ``nearest drainage'' limitation most evident at river junctions.
-\item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
-\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation extents.
+\item Flood maps derived from Height Above Nearest Drainage (HAND) are subject to nearest drainage line limitations that affect inundation skill.
+\item A means of resolving this limitation is provided by reducing HAND processing units to level paths with effective unit stream order.
+\item Discretizing the stream network for HAND computation affects the stage-discharge relationship and leads to higher skill inundation.
 \end{keypoints}
@@
 \begin{abstract}
-The National Water Model (NWM) currently requires additional modeling from discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
-Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is a means worthy of producing high resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
-However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
-We propose here that HAND is limited to producing inundation only when sourced from its nearest drainage line, thus lacks the ability to source inundation from multiple fluvial sources.
-To counter this constraint, a version of HAND, known as Generalized Mainstems (GMS), is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
-LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
+Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is a means able of producing high resolution flood inundation maps (FIMs) from the National Water Model (NWM) at large scales and high resolutions using reach-averaged synthetic rating curves.
+We highlight here that HAND is limited to producing inundation only when sourced from its nearest drainage line, thus lacks the ability to source inundation from multiple fluvial sources.
+A version of HAND, known as Generalized Mainstems (GMS), is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
 The FIMs associated with each independent LP are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
-Improvement in mapping skill is observed by significantly reducing false negatives at river junctions when the inundation extents are compared to FIMs from the Hydrologic Engineering Center's River Analysis System (HEC-RAS).
-A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
-We observe that the improvement from extending drainage order reduction techniques from 4-5\% of the entire stream network to 100\% of the network is about the same magnitude improvement as going from no drainage order reduction to 4-5\% of the network.
-Thus, we conclude that drainage order reduction techniques circumvent the nearest drainage limitation while offering diminishing returns to skill improvement.
-Drainage order reduction techniques also penalize with additional computational expenses in terms of processing and storage.
-This novel contribution is proposed and studied in a new open-source implementation that utilizes the latest combination of hydro-conditioning techniques to enforce drainage and counter limitations in the elevation data.
-This framework scales to the entire NWM modeling domain including Hawaii and Puerto Rico to help forecasters provide timely and accurate flood warnings.
+Improvement in mapping skill is observed by significantly reducing false negatives at river junctions when the inundation extents are compared to FIMs from that of benchmarks.
+A more marginal reduction in the false alarm rate is also observed due to a shift introduced in the stage-discharge relationship by increasing the size of the catchments.
+We observe that the improvement of this method applied at 4-5\% of the entire stream network to 100\% of the network is about the same magnitude improvement as going from no drainage order reduction to 4-5\% of the network.
+This novel contribution is framed in a new open-source implementation that utilizes the latest combination of hydro-conditioning techniques to enforce drainage and counter limitations in the input data.
 \end{abstract}
@@
 The United States National Water Model (NWM) provides flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
-In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model, known as Height Above Nearest Drainage (HAND), suffers from issues in mapping performance because only fluvial inundation, or inundation sourced from rivers, is considered from the nearest drainage line.
-As rivers flood, multiple neighboring rivers could contribute to a region's inundation for which HAND has no means of modeling for.
+In order to convert flow rates from the NWM to flood inundation maps (FIM), a model, known as Height Above Nearest Drainage (HAND), is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
+This model suffers from issues in mapping performance because inundation sourced from rivers is only considered from the nearest river line.
 We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
@@
 HAND is computed independently for each LP and the resulting FIMs are mosaiced together to form one seamless map.
-We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance to counter one of HAND's fundamental limitations.
+We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance.
 %
@@
  %
+% Open Research
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Open Research}
+%\openresearch
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs \cite{nwm2022hydrofabric}.
+These are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
+OWP Flood Inundation Mapping (FIM) capabilities rely extensively on the use of the National Hydrography Plus High Resolution (NHDPlusHR) datasets including BurnLineEvents \cite{nhdplus2022vectors}, value-added attributes (VAA) \cite{nhdplus2022vectors}, water boundaries (WBD) or hydrologic unit code (HUC) geometries \cite{nhdplus2022wbd}, and digital elevation maps (DEM) \cite{nhdplus2022dems}.
+Some additional datasets for processing include the National Levee Database (NLD) furnished by the United States Army Core of Engineers (USACE) \cite{engineers2016national}, Land-sea border from the Great Lakes Hydrography Dataset (GLHD) furnished by the Great Lakes Aquatic Habitat Framework (GLAHF) \cite{GreatLakesHydrographyDataset}, and a Land-sea border provided by OpenStreetMap (OSM) \cite{osm2021landsea}.
+Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
+Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
+
+Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022}.
+ %
 % Acknowledgments
@@
 Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
- %
-% Open Research
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\openresearch
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs \cite{nwm2022hydrofabric}.
-These are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
-OWP Flood Inundation Mapping (FIM) capabilities rely extensively on the use of the National Hydrography Plus High Resolution (NHDPlusHR) datasets including BurnLineEvents \cite{nhdplus2022vectors}, value-added attributes (VAA) \cite{nhdplus2022vectors}, water boundaries (WBD) or hydrologic unit code (HUC) geometries \cite{nhdplus2022wbd}, and digital elevation maps (DEM) \cite{nhdplus2022dems}.
-Some additional datasets for processing include the National Levee Database (NLD) furnished by the United States Army Core of Engineers (USACE) \cite{engineers2016national}, Land-sea border from the Great Lakes Hydrography Dataset (GLHD) furnished by the Great Lakes Aquatic Habitat Framework (GLAHF) \cite{GreatLakesHydrographyDataset}, and a Land-sea border provided by OpenStreetMap (OSM) \cite{osm2021landsea}.
-Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
-Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
-
-Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022}.
  %

### 2022-09-19 manuscripts/owp_fim4_2022/owp_fim4_expanded_original.tex
Resubmitted to WRR

--- 2022-08-31 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
@@
 \usepackage{soul}
-\usepackage{rotating}
 
@@
 \usepackage{amsmath}
-\usepackage{float}
 
@@
 %
-%\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
-\title{Extending Height Above Nearest Drainage to Model Multiple Fluvial Sources in Flood Inundation Mapping Applications for the U.S. National Water Model}
+\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
 %
@@
 \affiliation{3}{National Water Center, Office of Water Prediction, National Weather Service, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
-\affiliation{4}{Hydrologic Applied Innovations Lab, New York Water Science Center, United States Geological Survey, Troy, NY, USA}
+\affiliation{4}{New York Water Science Center, Hydrologic Applied Innovations Lab, United States Geological Survey, Troy, NY, USA}
 \affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Maryland, College Park, MD, USA}
@@
 \begin{keypoints}
-\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to a ``nearest drainage'' limitation most evident at river junctions.
+\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
 \item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
-\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation extents.
+\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation.
 \end{keypoints}
@@
 \begin{abstract}
-The National Water Model (NWM) currently requires additional modeling from discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
+The National Water Model (NWM) currently requires the post-processing of forecast discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
 Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is worthy of producing high-resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
 However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
-We propose here that HAND is limited to producing inundation only when sourced from its nearest drainage line thus lacks the ability to source inundation from multiple fluvial sources.
-To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
+To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that reduces the Horton-Strahler stream order of the stream network.
+GMS contains all segments within the NWM stream network but instead of deriving HAND by accounting for all river segments at once, it is derived independently at the level path (LP) scale.
 LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
-The FIMs associated with each independent LP are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
-Improvement in mapping skill is observed when the extents are compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
+These FIMs are then mosaiced together, effectively turning the stream network into discrete groups of homogenous unit stream order by removing the influence of neighboring tributaries.
+Improvement in mapping skill is observed when compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
 A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
-We observe that the improvement from extending drainage order reduction techniques from 4-5\% of the entire stream network to 100\% of the network is about the same magnitude improvement as going from no drainage order reduction to 4-5\% of the network.
-Thus, we conclude that drainage order reduction techniques circumvent the nearest drainage limitation while offering diminishing returns to skill improvement and penalizing with additional computational expenses.
-This novel contribution is proposed and studied in a new open-source implementation that utilizes the latest combination of hydro-conditioning techniques to enforce drainage and counter limitations in the elevation data.
-This framework scales to the entire NWM modeling domain including Hawaii and Puerto Rico to help forecasters provide timely and accurate flood warnings.
 \end{abstract}
@@
 In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance because only fluvial inundation, or inundation sourced from rivers, is only considered from the nearest drainage line.
-As rivers flood, multiple neighboring rivers could contribute to a regions inundation and HAND has no means of modeling for that.
+This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
 We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-This is done by splitting the stream network into continuous rivers known as level paths.
-These level paths have no tributaries thus are known to be stream lines with a unit stream order indicating no branching.
-HAND is computed independently for each level path and the resulting FIMs are mosaiced together to form one seamless map.
-We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance to counter one of HAND's fundamental limitations.
+We compared these HAND derived FIMs to maps from more realistic models and found improvement in mapping performance.
 %
@@
 More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
-With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}, with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf Coast.
-Trends related to flood damages and fatalities have been steadily increasing over recent decades \cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
-Some are expecting that the hydrologic cycle will intensify due to climate change which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
-Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest and Great Lakes regions while decreasing in the coastal Southeast, Southwest, and California.
+With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018} with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf coast.
+Trends related to flood damages and fatalities have been steadily increasing over recent decades.
+\cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
+Some are expecting that the hydrologic cycle will intensify which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
+Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest/Great Lakes region while decreasing in coastal Southeast, Southwest and California.
 %
@@
 %
-Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life threatening and property damaging events. 
+Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life or property damaging events occur. 
 One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble streamflows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
-Additionally, two forecasting networks, Full Resolution (FR) and Mainstems (MS) stream networks, relevant to the National Water Model (NWM) (see Section \ref{ssec:national_water_model}) are rendered in Figure \ref{fig:forecast_points}.
-The FR network refers to the entire NWM forecasting domain while MS refers to the subset of the FR network that is at or downstream of AHPS forecasting points (see Section \ref{ssec:national_water_model}).
-On an approximate basis, there is only one forecast point every 1,450 km of river (FR) and one forecast point with FIM every 29,000 km of river (FR).
-Despite the AHPS advances in operational flood forecasting, it lacks sufficient domain coverage, spatial resolution, and long-range forecast horizons to address the increasingly complex water challenges facing the US.
-%
-\begin{figure}[H]
+AHPS provides forecasting services in the form of ensemble stream flows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
+Additionally, two forecasting networks relevant to the National Water Model (NWM) which will be introduced in Section \ref{ssec:national_water_model} are rendered in Figure \ref{fig:forecast_points}.
+AHPS implements a series of advances including model calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}, distributed modeling approaches \cite{reed2004overall,koren2004hydrology,duan2002results}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, enhanced data analysis procedures \cite{mcenery2005noaa}, flood-forecasting inundation maps \cite{cajina2002fldview}, hydraulic routing models \cite{fread1973technique,cajina2002fldview}, and multi-sensor precipitation techniques \cite{breidenbach1999accounting,kondragunta2001outlier,seo2002real,bonnin1996noaa}.
+On an approximate basis, there is only one forecast point every 1,450 km of river and one forecast point with FIM every 29,000 km.
+Despite the AHPS advances in operational flood forecasting, it lacks sufficient spatial coverage and long-range forecast horizons to address the increasingly complex water challenges facing the US.
+%
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=1.0]{figures/forecast_points.jpg}
-\caption{Forecast points with and without Flood Inundation Maps (FIM) in United States' Advanced Hydrologic Prediction System (AHPS).
-Note that only a small fraction of the AHPS forecast points have existing FIM.
-Also shown are the National Water Model (NWM) stream networks at the Full Resolution (FR) and Mainstems (MS) resolution.
-The FR network constitutes the entire NWM stream network while the MS resolution network is the FR network at or downstream of the AHPS forecast points shown.
-}
+\caption{Forecast points with and without FIM in United States' Advanced Hydrologic Prediction System.
+Also show are the National Water Model stream networks at the full resolution (FR) and Mainstems (MS) resolution.}
 \label{fig:forecast_points}
@@
 %
-Additional work is required to address the gaps that the AHPS leaves in terms of spatial resolution, spatial coverage, and temporal forecast horizons.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners at the National Center for Atmospheric Research (NCAR) have developed and implemented operationally the NWM which is a configuration of the Weather Research and Forecasting Hydrologic Model (WRF-Hydro) \cite{salas2018towards,gochis2021wrf,cosgrove2019evolution}.
-The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including lookback-range (3-28 hrs), short-range (18 hr), medium-range (10 day) and long-range (30 day) forecast horizons.
+Additional work is required to fill-in the gaps that the AHPS leaves in terms of spatial and temporal coverage.
+In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners has developed and implemented operationally the National Water Model (NWM) which is a version of the Weather Research and Forecast Hydrologic Model (WRF-Hydro) \cite{gochis2018wrf,cosgrove2019evolution}.
+The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including some medium (10 day) and long (30 day) range forecast horizons.
 The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
-As a complement to the operational NWM, RFC forecasts from AHPS forecast points are assimilated in the NWM and routed downstream to the next downstream AHPS forecast point where the process iterates again.
-This assimilation into NWM is used to enhance forecasting skill by leveraging best available regional-scale forecasts.
-The river network upon which this special assimilation technique operates on is herein referred to as the Mainstem (MS) stream network.
-Figure \ref{fig:forecast_points} shows the NWM V2.1 FR stream network as well as the NWM V2.1 MS network.
+As a complement to the operational NWM, OWP has also developed a configuration of the NWM that extends RFC forecasts downstream by assimilating and routing forecast streamflows to the next downstream AHPS forecast point.
+This configuration of the NWM is used to enhance forecasting skill by leveraging best available regional-scale data.
+The river network upon which this special configuration operates is herein referred to as the Main-Stem (MS) modeling stream network.
+Figure \ref{ssec:national_water_model} shows the Full-Resolution (FR) modeling stream network as well as the MS network.
 The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
 
-The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the ``hydrofabric'' in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus,nhdplus2022vectors}.
-The term ``hydrofabric'' is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The NWM provides stream forecasts at these hydrofabric segments using the Muskingum-Cunge method to reduce computational requirements of a continental scale model but fails to consider backwater dynamics \cite{bedient2008hydrology,ponce1994variable,gochis2021wrf}.
-The need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
-The use of a 2-dimensional (2D) hydrodynamic model across a continental-scale and high spatial resolutions is very cost prohibitive especially in an operational setting.
-The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert 1-dimensional (1D) riverine discharges to stages, and finally to inundation extents and depths.
+The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
+The term hydrofabric is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
+The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
+Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
+As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
+The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages then finally to inundation extents.
 %
@@
 %
-HAND normalizes topography along the nearest drainage path and it has been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
+HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
 \citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-However, a core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
-The terrain index also provides additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
-\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as SRCs by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
+A core assumption made for HAND based FIM is enforcing drainage across the entire area of interest which requires significant digital elevation maps (DEM) manipulations to make a reality.
+The terrain index has even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
+\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
 This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{HAND's Assumptions and Limitations}
-\label{sssec:hand_assumptions_limitations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-HAND operates on many underlying assumptions since it can only be used as an inundation proxy or no physics model and thus, not a true hydrodynamic inundation model \cite{nobre2016hand,liu2016cybergis,liu2020height}.
-HAND, to our knowledge, has only been applied to natural, inland, and riverine inundation applications thus it is also missing pluvial, coastal, ground water, and dam break components among other possible sources of flooding.
-Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest stream line which is used for catchment allocation and relative elevation calculation \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,maidment2017conceptual,garousi2019terrain,zheng2018river,zheng2018geoflood,johnson2019integrated,renno2008hand}.
-Stream thalweg networks must also collectively drain to a singular outlet point for a given processing region \cite{nobre2016hand,zheng2018geoflood,renno2008hand}.
-Since elevations don't naturally do this, they must undergo a long series of hydro-conditioning processes to enforce monotonically decreasing elevations across an entire processing unit along with hydrologically correct directions of flow \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,donchyts2016global,renno2008hand}.
-The level of DEM manipulation required to enforce this assumption can be substantial depending on the region and can be a significant source of error.
-The drainage enforcing assumption also interacts with an inability to properly account for fluvial inundation in regions of DEM depressions that lack natural drainage to riverine areas \cite{nobre2016hand,renno2008hand}.
-
-When used for FIM applications, HAND assumes only fluvial inundation sourced from its nearest drainage line is accounted for \cite{nobre2016hand,mcgehee2016modified}.
-Catchments are independent of one another for FIM purposes meaning a reaches' stage value is only used to threshold the HAND values within its respective catchment \cite{liu2016cybergis,zheng2018river,zheng2018geoflood}.
-This assumption plays to the ``Nearest Drainage'' term in HAND and creates a significant limitation within HAND for FIM applications \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-At the junction of high stream order and high flow rivers with lower flow tributaries, there can be a lack of inundation extents exhibited which is known colloquially in the forecasting community as the ``catchment boundary problem''.
-The academic community has somewhat referenced this issue before but it has been characterized more as a problem with the stream delineation process that comes from thresholding the drainage accumulation maps \cite{nobre2016hand,li2020evaluation}.
-Later in this study, we will re-introduce this problem and demonstrate how we initialize with a stream network (that of the NWM's) and thus avoid having to threshold accumulations to some arbitrary value to define stream networks.
-We illustrate how computing HAND independently for stream lines of unit stream order can significantly enhance FIM performance by accounting for multiple sources of fluvial inundation that may exist in certain regions and flow scenarios.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{HAND Implementations}
@@
 %
-Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (approximately 10 m depending on latitude) scale, HAND has been implemented into software for large-scale, continental computation.
-As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a HPC \cite{maidment2017conceptual,liu2016cybergis}.
+Due to significant advances in high-performance computing (HPC) and large scale high-resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (10 m) scale, HAND has been implemented into software for large-scale, continental computation.
+As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a high-performance computer (HPC) \cite{maidment2017conceptual,liu2016cybergis}.
 NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 CPU years.
-By allocating 31 nodes at 20 cores per node for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
-Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (NHDPlusHR) which better agrees with the 10 m DEM from the NHDPlusHR program \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC as an unofficial demonstration to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
-Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
-More notably the software added the ability to derive HAND on both the NWM FR and MS stream networks to consider multiple sources of fluvial inundation along high impact rivers of primary forecasting concern.
-
-Related to these efforts, the United States Geological Survey (USGS) has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses SRCs with cross-sections for stage-discharge relationships \cite{verdin2016software}.
+By allocating 31 nodes at 20 cores per for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2TB of peak memory and 5TB of total disk space.
+Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (HR) \cite{liu2020height}.
+The original NFIE dataset was employed by the NWC as an unofficial demonstrations to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
+Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE-DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
+More notably the software added the ability to derive drainage potentials on both the NWM FR and MS stream networks which leverages the lower drainage density and Horton-Strahler stream order of the MS network.
+
+Related to these efforts, the USGS has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses synthetic rating curves with cross-sections for stage-discharge relationships \cite{verdin2016software}.
 Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Office of Water Prediction Flood Inundation Mapping}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-In order to mitigate the ever increasing threat of flooding to life and property, an operational capability is required to extend NWM streamflow forecasts to river stages, inundation extents, and inundation depths.
-OWP FIM is introduced here as a continental scale capability that generates these products at high spatial and temporal resolutions.
+\subsection{OWP FIM}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+Many of those in the academic community assessing HAND's efficacy for producing FIM have noted opportunities for improvement. 
+\citeA{godbout2019error} found how reach length and slope are important parameters for maximizing mapping skill with the moderate values performing best. 
+The co-linearity of reach length and slope led \citeA{godbout2019error} to propose that reaches of extreme lengths performed worse because of the extreme slope values, a parameter directly represented in Manning's equation.
+Issues with the reach-average approaches have been documented in \citeA{tuozzolo2019impact} where large within reach variance of the roughness Manning's n coefficient have been observed.
+Furthermore, \citeA{garousi2019terrain} noted improvements to mapping efficacy by conditioning monotonically decreasing thalweg elevations, adjusting the Manning's n roughness coefficient, and using higher resolution (3 m) Digital Elevation Model's (DEM) derived from light detection and ranging (Lidar).
+The use of higher resolution DEMs in that study was motivated by previous work with Lidar DEMs and least-cost thalweg derivations \cite{zheng2018geoflood}.
+Further work by \citeA{johnson2019integrated} noted the general under-prediction of HAND and suggested tuning the Manning's n parameter to better align SRC's with observations. 
+Additionally, the sensitivity to low topographic relief and channel slope have been observed \cite{johnson2019integrated,godbout2019error}. 
+Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
+These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM.
+
+With all of the latest developments in the realm of continental FIM (CFIM) noted in the previous paragraph, a fast, open-source framework for accelerating the research to operations pipeline is required.
 Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
-Within the operational framework of OWP FIM, we introduce research demonstrating how FIM performance skill with HAND can be improved by discretizing stream networks into units of an effective unit Horton-Strahler stream order \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} for HAND computation contexts.
-Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to the stream accumulation threshold which is closely related to stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths (LP), can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
-This capability is introduced progressively as MS (whose network represents about 4\% of FR network) and to a higher degree Generalized Mainstems (GMS) (covers entire FR network) which will be explained later on.
+This framework is open-source and modular enabling and accelerating the research to operations development pipeline.
+Automated evaluation tools and processed test case data enable the rapid testing and evaluation of new techniques in consistent contexts.
+In addition to developed tooling, we introduce research demonstrating how FIM performance skill with HAND can be improved by reducing Horton-Strahler stream orders \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} of the stream networks.
+Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to stream threshold which serves as a proxy for stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
+Here we demonstrate how reducing a HAND processing unit's stream network to a singular stream order discretized by level paths, can enhance FIM skill.
+This capability will be referred to as Generalized Mainstems (GMS).
 The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM with applications.
@@
 OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
-Figure \ref{fig:methods_overview} gives a high-level overview of the methodology used in OWP FIM and in this study.
-Input data from multiple sources are preprocessed (not illustrated) and then subset to processing areas based on the model used, FR, MS, or GMS.
-The standard processing unit of OWP FIM is a HUC8 however for the MS and GMS versions only the NWM MS stream network and the NWM FR network discretized into LPs are used, respectively.
-A series of hydro-conditioning steps enforces the location of stream lines, monotonically decreasing elevations, excavated bathymetry, stream thalweg breaching, and levee enforcement.
-After a DEM suitable for HAND's assumptions is conditioned, the FIM hydrofabric is generated including stream network, catchments, HAND, SRCs, and cross-walk table.
-The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains SRC and cross-walk information).
-In operational circumstances, the NWM streamflows are used in conjunction with the FIM hydrofabric to derive forecast FIMs.
-However for evaluation purposes, we use the streamflows from the cross-sections of our benchmark model.
-As later discussed in Section \ref{ssec:inundation_mapping}, independent FIMs from multiple fluvial sources are mosaiced together.
-For the case of MS, two sources are mosaiced together (FR and MS) while for GMS the inundation from every LP is composited together.
-The evaluation FIM extents are compared to the extents of the benchmark model and metrics are computed.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=6.0]{figures/methods_overview.jpg}
-\caption{
-Methodology overview detailing high-level steps followed in the study.
-The flow chart begins with the input data organized by source.
-Subsetting the data into processing units depends on which model is being considered.
-FR utilizes the entire NWM stream network processed at HUC8 processing areas.
-MS only computes HAND using the NWM stream at or downstream of legacy forecasting points.
-The resulting inundation from the MS HAND is eventually layered with the FIM from FR HAND to account for high levels of inundation contributed by the mainstem.
-Generalized Mainstem (GMS) discretized NWM streams into level paths (LP) then computes HAND and the FIMs independently only to mosaic them later. 
-This better accounts for multiple possible sources of fluvial inundation.
-The dotted lines denote the use of NWM streamflow forecasts to produce operational FIM but not used in this study.
-All acronyms used in the figure are defined in the paper.
-}
-\label{fig:methods_overview}
-\end{figure}
-%
 %
@@
 \subsection{Software Dependencies and Architecture}
-\label{ssec:software}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems, OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
+To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in a Docker image (\url{https://docs.docker.com/engine/install/}). 
 A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub where a user could consult the Readme.md page for more information on how to acquire the datasets and reproduce the pipeline \cite{inundationMapping2022}.
+Source code is made available for this project on GitHub \cite{inundationMapping2022}.
+
+The pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. 
+Preprocessing acquires and prepares datasets for production of the FIM hydrofabric. 
+The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
+Functionality is included to turn FIM hydrofabric and streamflows into inundation maps represented in both binary and depths.
+Lastly, a test suite includes means of calculating evaluation metrics compared to a variety of preprocessed test case data.
+A user should visit the Readme.md page on GitHub for more information on how to acquire the datasets and reproduce the pipeline.
 %
@@
 Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data processed for OWP FIM are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by the Earth Science Information Partners (ESIP) \cite{esipData2022}.
+Instructions for accessing data are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by Earth Science Information Partners (ESIP) \cite{esipData2022}.
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
-We utilized a series of data products from the NHDPlusHR including the BurnLineEvents \cite{nhdplus2022vectors}, Value Added Attributes (VAA) \cite{nhdplus2022vectors}, Water Boundaries (WBD) or HUC Layers \cite{nhdplus2022wbd}, and the DEM elevation rasters \cite{nhdplus2022dems}.
-These BurnLines used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk against for forecasting purposes \cite{nwm2022hydrofabric}.
-For enforcing levee data, the USACE NLD is used to burn feature elevations into DEMs \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) \cite{osm2021landsea} and the land-lake border from Great Lakes Hydrography Dataset (GLHD) \cite{GreatLakesHydrographyDataset} to exclude those areas from production of FIMs.
-Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of nine states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year or yr) and 0.2\% (500 year or yr), are used for validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
-These BLE datasets are provided at the watershed scale (HUC8) utilizing best available DEMs and simulations from the Hydrologic Engineering Center's River Analysis System (HEC-RAS) model \cite{us2022hydrologic}.
+It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
+For enforcing levee data into the DEMs, the USACE National Levee Database (NLD) is used to burn feature elevations \cite{engineers2016national}.
+Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) and the land-lake border from Great Lakes Hydrography Dataset (GLHD) to exclude those areas from production of FIMs \cite{OpenStreetMap,GreatLakesHydrographyDataset}.
+Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of 9 states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year) and 0.2\% (500 year), are used as validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
+These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
 The full input datasets presented by source are listed in Table \ref{tab:data}.
 %
-\begin{sidewaystable}
-\caption{Data sources, names, descriptions, and citations.}
+\begin{table}
+\caption{Data sources, names, and descriptions.}
 \label{tab:data}
 \centering
-\begin{tabular}{|p{1.75cm}|p{4cm}|p{11cm}|p{5cm}|}
+\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
 %\begin{tabular}{l c c}
 \hline
-Source & Name & Description & Citations\\
-\hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlusHR for hydro-enforcement. & \cite{nhdplus2022vectors} \\
-\hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display. & \cite{nhdplus2022vectors} \\
-\hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlusHR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters. & \cite{nhdplus2022dems} \\
-\hline
-USGS & NHDPlusHR WBD & Water Boundaries (WBD) or HUCs used for spatial processing units. & \cite{nhdplus2022wbd} \\
-\hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting adapted from NHDPlus V2 NHDFlowline\_Network feature class. & \cite{nwm2022hydrofabric} \\
-\hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM adapted from NHDPlus V2 Catchment feature class. & \cite{nwm2022hydrofabric} \\
-\hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes adapted from NHDPlus V2 NHDWaterbody feature class. & \cite{nwm2022hydrofabric} \\
-\hline
-USACE & NLD & Levee database of locations and elevations. & \cite{engineers2016national} \\
-\hline
-OSM & Land-Sea Border & Border of land and sea. & \cite{osm2021landsea} \\
-\hline
-GLHD & Land-Great Lakes Border & Border of land and Great Lakes. & \cite{GreatLakesHydrographyDataset} \\
-\hline
-InFRM & Cross-Sections & HEC-RAS 1D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
-\hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
+Source & Name & Description \\
+\hline
+USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
+\hline
+USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
+\hline
+USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters \\
+\hline
+NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting. \\
+\hline
+NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM. \\
+\hline
+NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes. \\
+\hline
+USACE & NLD & Levee database of locations and elevations  \\
+\hline
+OSM & Land-Sea Border & Border of land and sea. \\
+\hline
+GLHD & Land-Great Lakes Border & Border of land and Great Lakes. \\
+\hline
+InFRM & Cross-Sections & HEC-RAS 1-D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. \\
+\hline
+InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. \\
 \hline
 \end{tabular}
-%\end{table}
-\end{sidewaystable}
+\end{table}
 %
 Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and GMS networks and 1,604 HUC8s for the MS method. 
-These methods will be explained in more detail later.
-An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which are expansion domains in the NWM V2.0 and V2.1, respectively.
+These methods will be explained more later.
+An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which the NWM V2.1 will cover.
 %
@@
 %
-The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping with HAND. 
+The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping. 
 These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
 HAND inherently requires all areas eligible for inundation to drain to the designated drainage network so DEMs must undergo significant manipulation to make this the case.
-In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligible for flooding.
+In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations if we wish to have them be eligble for flooding.
 Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
@@
 %
-The location of the stream network is enforced to ensure general agreement with the NWM network which is used for forecasting the streamflow inputs.
-The NHDPlusHR Beta BurnLineEvent layer is used to enforce stream locations in the NHDPlusHR workflow and best agrees with thalweg locations in the DEM used so it is also used here for hydro-enforcement \cite{moore2019user}.
-
-However, to better match the drainage density of the NWM FR V2.1 stream network, which is based on the NHDPlus V2, the BurnlineEvents are pruned utilizing a nearest neighbor search around the NWM flowlines.
-Headwater points are first derived for the NWM FR V2.1.
-For every NWM headwater point, the nearest NHDPlusHR point is selected and placed into a set and those not belonging to this nearest neighbor set are discarded.
-Only the nearest point on the NHDPlusHR is used so any portion of the NHDPlusHR network upstream of this nearest point is discarded to avoid extending inundation too far above the modeling domain.
-The points in this nearest neighbor set are then traversed downstream.
-Any headwater portion in the NHDPlusHR or any other stream not traversed are pruned away to better match the resolution and spatial locations of the NWM stream network and its headwater points.
-The resulting pruned NHDPlusHR stream network gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHDPlusHR network corresponds to the NHDPlusHR network at or downstream of NWM V2.1 headwater locations only.
+The location of the stream network is enforced to ensure general agreement with established stream networks.
+The NHDPlusHR Beta Burnline Event layer is used to enforce stream locations in the NHDPlusHR workflow so it is also used here for hydro-enforcement \cite{moore2019user}. 
+However, to better match the drainage density of the NWM V2.1 stream network which is based on the NHDPlus Medium Resolution, the Burnline Events are pruned utilizing a nearest neighbor search around the NWM flowlines.
+For every NWM headwater segment a headwater point is derived and linearly interpolated to the nearest Burnline Event segment.
+Burnline Event headwater segments are split at the adjusted headwater point to match NWM flowlines.
+The resulting pruned NHD stream network is what gets hydro-enforced in subsequent operations.
+This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHD network corresponds to the full density NHD network at NWM V2.1 headwater locations only. 
 Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents. 
-%
-\begin{figure}[H]
+%
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{This figure illustrates some of the datasets that result from the pruning of NHDPlusHR Beta BurnlineEvents (dotted black) to the stream density of the NWM FR V2.1 density (blue).
-The stream network used for forecasting, NWM FR V2.1, is of higher stream density than that of the NHDPlusHR which has better agreement with the thalweg locations in the DEM used.
-Thus we opt to prune the NHDPlusHR network to match the general location and density of the NWM network.
-The nearest neighbor segment in the NHDPlusHR of each NWM headwater locations and the nearest point on that segment is determined to match the closest point to that of the NWM headwater.
-These points are then traversed downstream and any segments not traversed are pruned away.
-The resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR BurnlineEvents.}
+\caption{Pruning of NHDPlus HR Beta Burnline Events (dotted black) to NWM V2.1 stream density (blue) using nearest neighbor selection and linear interpolation.
+Resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR Burnline Events.}
 \label{fig:stream_density_pruning}
@@
 %
+This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents.
 
@@
 The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop and buffer distance. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling an inverted triangular notch shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop which were set to fixed values of 70 m, 1000 m, and 10 m, respectively, but available to the user via the parameter file.
-While the values for these parameters are critical to the inundation extents produced, especially for lower flow rates where bathymetric information has more influence, finding their optimal values for OWP FIM was not done since it was out of the main scientific scope of this article.
+The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop. 
+The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling a trapezoidal shape with a significantly lower elevation along the thalweg line only.
+In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop. 
 Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement techniques especially when computational costs are considered. 
+\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement especially when computational costs are considered. 
 Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
-Some of these drawbacks are addressed by additional conditioning techniques applied later on.
+Some of these drawbacks are later addressed by additional conditioning techniques later on.
 %
@@
 %
-Coarse DEM's at 10 m, 30 m, and higher resolutions can lack sufficient representation of fine grain features such as embankments, flood walls, and closure structures \cite{arundel2018assimilation,dobbs2010evaluation,wang2005comparison,sanders2007evaluation}.
-In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 1/3 arc-second DEM.
-The elevations found in the NLD are burned onto the DEM if those elevations were found to exceed those already in place.
+The DEM at 10 m resolution lacks sufficient representation of fine grain features such as embankments, floodwalls, and closure structures.
+In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 10 m DEM.
+The elevations found in the NLD are burned into the DEM if those elevations were found to exceed those already in place.
 %
@@
 %
-Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes with HAND.
-The partially conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
+Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes.
+The conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
 Priority-Flood developed by \citeA{barnes2014priority} is an algorithm for filling said depressions and shown to have improved performance over early works in the field by \citeA{jenson1988extracting} implemented in \citeA{tarboton2005terrain} as well as \citeA{planchon2002fast}.
@@
 This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
-The depression filling technique employed here does leave the existence of flat regions where pits previously existed thus later requiring the need for resolving these flats.
+The depression technique employed here does leave the existence of flat regions where pits existed a prior thus later requiring the need for resolving these flats.
 The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
@@
 %
-Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing, elevations are derived prior to the normalizing of elevations.
-Work on the AGREE DEM method from several authors have illustrated that the AGREE DEM method does not prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another \cite{hellweger1997agree,baker2006comparison,saunders1999preparation,saunders1996gis,quenzer1998gis,saunders1995grid}.
-These works observe that the artificial elevations enforced by the hydrographically based stream network and AGREE DEM disagree with those naturally occurring in the native DEM.
-In order to mitigate this documented issue, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum on the original, unconditioned DEM for each thalweg pixel.
+Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing elevations are derived prior to normalizing of elevations.
+In order to prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum for each thalweg pixel.
 Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50 m.
 The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
-This step essentially enforces an estimate of the native DEM thalweg elevations onto the sharp drop enforced thalweg elevations from the AGREE procedure.
 
 The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
-\citeA{garousi2019terrain} proposed an algorithm that breaches stream thalweg pixel elevations in a depth first manner. 
-This procedure was found to increase the Critical Success Index (CSI) of resulting FIMs from HAND and is employed in OWP FIM to enforce monotonically decreasing elevations with thalweg pixel networks.
+\citeA{garousi2019terrain} proposed an algorithm that traverses stream thalweg pixels in a depth first manner starting with adding all the headwater pixels to a queue. 
+The connectivity of the thalweg pixels is defined by the D-8 flow directions further discussed in Section \ref{ssec:flow_direction_and_flat_resolution}.
+At every thalweg pixel, the minimum elevation among itself and its upstream contributing thalweg pixels is taken as shown in Equation \ref{eq:thalweg_breach},
+%
+\begin{linenomath*}
+\begin{equation}
+\label{eq:thalweg_breach}
+\textbf{D}[x] = \min_{y\ drains\ to\ x} {(\ \textbf{D}[x]\ ,\ \textbf{D}[y]\ )}
+\end{equation}
+\end{linenomath*}
+%
+, in which \textbf{D} represents the array of thalweg adjusted elevations indexed by x and y where by y is upstream of x. 
+When a pixel's upstream neighbors are all evaluated, the downstream pixel is added into the queue thus the depth first traversal of the drainage network.
+This procedure enforces the location of streams and ensures that thalweg elevations are hydrologically correct which yielded a 7\% improvement in Critical Success Index (CSI) per an evaluation for an event in 2017 on the Malad river \cite{garousi2019terrain}.
 %
@@
 The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
-These datasets include the HAND or relative elevation model (REM) raster, reach-level catchments raster/polygons, DEM-derived streamlines, SRCs, and cross-walk table.
-The following sub-sections describe how the subset and hydro-enforced geospatial datasets are converted into the FIM hydrofabric.
 %
@@
 D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the four neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$ by means of the Pythagorean theorem.
-Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression-filled cells must be resolved.
-
-Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment} where flats are resolved by incrementing elevations iteratively.
+The horizontal component of slope is defined as one for the 4 neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$. 
+The flow direction is encoded as integers 1 through 8 corresponding with the cardinal direction East as 1 and continuing counter-clockwise to the Southeast direction as 8. 
+Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression filled cells must be resolved.
+
+Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment}.  
 Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
 The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
-The established literature in this niche field of hydrology discusses how prevalent flats can be in given study areas and how difficult the problem is from both computational and hydrologic stand-points \cite{garbrecht1997assignment,tarboton2009generalized,tarboton2005terrain,survila2016scalable,barnes2014efficient,tesfa2011extraction,wallis2009parallel}.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
+Specifically the work by \citeA{survila2016scalable} enables the use of parallel processing and made smoother catchments from our informal experience than those from \citeA{barnes2014efficient}.
+By processing flats local to each partition separately from flats shared with other partitions, the accelerated flat resolution algorithm demonstrated an average speed up of 468x when compared to prior implementations \cite{survila2016scalable}.
+OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats \cite{survila2016scalable,cybergis2016}.
 %
@@
 %
-The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new, DEM based, FIM stream network. 
-The FIM stream network is of similar drainage density as the NWM V2.1 network and fully converges at all junctions leaving no divergences in the network.
+The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new FIM stream network. 
+The FIM stream network is of similar drainage density as the NWM V2.1 network but fully converges at all junctions leaving no divergences in the network.
 This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are pruned headwater locations of the NHDPlusHR Beta BurnlineEvents layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
+These seeds points are headwater locations of the NHDPlusHR Beta Burnline Events layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
 Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
-
-Each stream link in this derived FIM stream network is split into equidistant reaches of 1.5 km in length which is a user exposed parameter.
+Each stream link in this derived FIM stream network is split into equidistant reaches.
 Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments.
@@
 Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve SRC and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretization of the stream network.
-Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherit the SRC properties of the upstream or downstream segment, depending on the topology.
-Section \ref{sssec:synthetic_rating_curve} details the derivation of the SRC and the dependence on channel length. 
-Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC8 membership.
+Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve synthetic rating curve and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
+Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretation of the stream network.
+Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherent the synthetic rating curve properties of the upstream or downstream segment, depending on the topology.
+Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 
+Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC 8 membership.
 This stream network is important since it drives the HAND calculation and derivation of catchments.
@@
 Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
-The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing, nearest drainage region for each outlet point.
-Two sets of catchments are derived, one set of catchments denotes the unique drainage region for each thalweg pixel which is used for relative elevation calculation.
-The other catchments are derived for the drainage region for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
+The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing area for each gage.
+Two sets of catchments are derived, one of which assigns the contributing area for each thalweg pixel which is used for relative elevation calculation.
+The other catchments are derived for the contributing area for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
 %
@@
 Once the pixel level catchments are derived, the final relative elevations can be computed.
-The elevation of every thalweg pixel is subtracted from the elevations of the non-thalweg pixels within the same, corresponding pixel-level catchment described in Section \ref{sssec:catchments}.
+Every non-thalweg elevation is subtracted from the thalweg elevation within the same pixel-level catchment described in Section \ref{sssec:catchments}.
 The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
-Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling \cite{djokic2019arc}.
-Any negative values resulting from this subtraction with native elevations are replaced by zero.
-Again, HAND assumes and requires processing areas to drain thus have monotonically decreasing elevations with hydrologically correct flow directions all leading to a singular outlet point.
-While this is required for the generation of DEM-derived catchments and stream lines, it is not necessarily required for the computation of the relative elevations.
-Since the use of hydro-conditioning processes to fit the drainage requirement for HAND can be extensive, we found it more fitting to use the native elevations this final HAND computation thus avoid the use of manipulated values that fit modeling assumptions.
+Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling. 
 %
@@
 %
-A method for converting forecast river discharges from the NWM to stages or river depths at the reach scale is necessary for producing FIMs with HAND. 
-For 1D hydrodynamic models such as the routing methods in the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a SRC at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
-For this application, we utilized the reach averaged approach for developing SRCs \cite{zheng2018river}.
+ A method for converting forecast river discharges from the NWM to stages or river depths is necessary for producing FIMs with HAND. 
+For one-dimensional models such as the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a synthetic rating curve at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
+For this application, we utilized the reach averaged approach for developing synthetic rating curves (SRC) \cite{zheng2018river}.
 The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-Previously not reported in literature to our knowledge in this form, the reach averaged Manning's formula is derived to be 
+The reach averaged Manning's formula is derived to be 
 %
@@
 \label{eq:reach_averaged_mannings_equation}
-Q(y) = \frac{1}{n} \frac{V(y)^{5/3}S^{1/2}}{L B(y)^{2/3}} 
+\textbf{Q} = \frac{1}{n} \frac{V^{5/3}S^{1/2}}{L B^{2/3}} 
 \end{equation}
@@
 %
-where Q is discharge at stage y, n is the Manning's n roughness coefficient, V is volume at y, S is channel slope, L is the along-flow reach length, and B is wetted bed area at y.
-Q, V, and B are taken at specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
-All units are international (SI) given the one in the numerator above n.
-The reach averaged method has been compared to rating curves from HEC-RAS and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
+where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
+Q, V, and B are taken a specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
+All units are international given the 1 numerator above n.
+The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
 The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the reach length.
-While the AGREE DEM is subject to hydro-conditioning processes, it does introduce some notion of bathymetry estimation that the native DEMs lack while being sensitive to additional parameters that could yield further errors in the FIM.
-We leave this issue open in this study and elaborate on needs with respect to bathymetry and Manning's n values in the Discussion section (Section \ref{sec:discussion}).
-
-Setting of the Manning's n roughness coefficient has precedent in previous continental-scale FIM (CFIM) studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
+Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the channel length.
+
+Setting of the Manning's n roughness coefficient has precedent in previous CFIM studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
 These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but not limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the SRCs with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values (y in Eq.
-\ref{eq:reach_averaged_mannings_equation}) from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
+We note significant opportunity to enhance CFIM skill by better parameterizing Manning's n according to available data including but limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
+For now and for the purpose of this study, we examine the developed ecosystem of tools with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
+After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values from 0 to 25 meters in depth at one third of a meter increments to calculate the discharge values for each stage value.
 %
@@
 %
-The DEM based stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with NWM reach identifiers so that a discharge can be converted to stage and later inundation extents and depths.
-For the FR version of HAND, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent HAND methods, MS and GMS, discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, respectively, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
-Additionally, only relevant catchments from the NWM for the given LP are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
-While these conflation methods are approximate, they can lead to some substantial errors which will be discussed more in Section \ref{sec:discussion}.
+The stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with a NWM reach identifier so that a discharge can be converted to stage and later inundation extent.
+For the methods already discussed, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
+For two subsequent methods discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
+Additionally, only relevant catchments from the NWM for the given level path are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
+While these conflation methods are approximate, they work for many instances just fine but do lead to areas with substantial error. 
+More discussion on this will follow in Section \ref{sec:discussion}.
 %
@@
 %
-As previously discussed, HAND based FIMs are subject to many assumptions and limitations in order serve as a suitable inundation proxy for large scale, high resolution domains.
-HAND produced FIMs are limited in only providing inundation sourced from the nearest drainage line, however, depending on flow conditions and topography, a given area may have multiple contributing fluvial sources of surface water inundation.
-The forecasting community, in reviewing HAND, have noted significant negative effects at the confluence of lower flow tributaries with higher flow rivers for which the phrase ``catchment boundary issue'' has been termed.
-In previous studies, FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND which is closely related to the maximum Horton-Strahler stream order of the network \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is partly in due to the limitation that catchment boundaries place on inundation extents where only the nearest drainage line can source inundation for any particular area.
-
-Figure \ref{fig:catchment_boundaries_issue} illustrates the exact situation our solution proposes to address where two tributaries converge with a higher order stream segment. 
-An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 yr extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the main segment in Figure \ref{fig:catchment_boundaries_issue} of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS.
+FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+In our evaluations, we note negative effects at the confluence of lower flow tributaries with higher flow rivers partly due to the independent nature of the catchments within HAND methods.
+Figure \ref{fig:catchment_boundaries_issue} illustrates this exact situation where two tributaries converge with a higher order stream segment. 
+An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 year extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
+The higher discharge along the MS of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS. 
 This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the mainstem that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the mainstem catchments.
-%
-\begin{figure}[H]
+As a parallel problem, there is excess water accumulated along the MS that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the MS catchments.
+%
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{The figure represents an agreement map between a HAND derived FIM and one produced from the Base Level Engineering (BLE) program for a 100 yr magnitude event at HUC8 12090301.
-Agreement maps are symbolized by false negatives (FN), true negatives (TN), false positives (FP), and true positives (TP) where inundated represents the positive condition (see Section \ref{ssec:evaluation} for more details).
-The streamflows associated with each river segment are shown in CMS while the flow directions are symbolized as red arrows.
-The presence of FNs at the confluence of tributaries (circled in red) with the main segment is associated with lower flow rates in the tributaries that don't account for backwater effects.
-Additionally, the flow of 1900 CMS from the main segment cannot extend to the neighboring catchments belonging to its tributaries shown here.
-Water pools up vertically along the catchment boundaries of the higher order segment distorting rating curve behavior (Section \ref{sec:discussion}).
-Sourcing fluvial inundation from HAND is limited to only its nearest drainage line which is the main issue this study aims to address.
-}
+\caption{False negatives associated with confluence of tributaries with MS. Integers represent flow values from BLE 100 year event for the associated areas. 
+No backwater consideration is implemented and the independent nature of the HAND catchments prohibits pluvial inundation from taking place.}
 \label{fig:catchment_boundaries_issue}
@@
 %
-
-We seek to resolve this catchment boundary problem or nearest drainage limitation by discretizing the target stream network into stream networks of reduced, unit stream order to avoid the constraining of catchments by those belonging to lower order neighbors.
-By discretizing the network into stream networks of unit stream order (later defined as LP), we remove the influence of neighboring catchments that constrain the inundation extent.
-This creates much larger and overlapping catchments that can source fluvial inundation from multiple reaches as required by the given river stage at current flow conditions.
-We present two successive methods, National Weather Service MS (Section \ref{sssec:nws_mainstems}) and GMS (Section \ref{sssec:generalized_mainstems}), implemented that reduce the effective Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND by expanding the nearest drainage definition to increase potential inundation areas.
-
-To clarify the phrase ``reducing Horton-Strahler stream order'' used extensively in this paper, every FIM used in evaluation contains a flood extent sourced from every NWM forecast point in the given evaluation domain.
-What we do to reduce stream order is discretize the NWM FR network into different units of size, MS network (\ref{sssec:nws_mainstems}) and GMS LPs (\ref{sssec:generalized_mainstems}), that effectively reduce the HAND computation to independent networks of unit stream order.
-These independent HAND datasets are later used to produce FIM independently and mosaiced together (see Section \ref{ssec:inundation_mapping}).
-The inundation from the MS HAND is mosaiced with the inundation from FR HAND, while the inundation of each individual LP from GMS is mosaiced together.
-The Horton-Strahler stream order is only reduced for HAND computation purposes to reduce the negative effects of the nearest drainage limitation inherent to HAND.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Mainstems}
+We seek to resolve this problem by deriving HAND for processing units with stream networks of reduced stream order. 
+We present two successive methods implemented that reduce drainage densities by reducing Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND.
+The resulting FIMs from the overlapping HAND processing units are mosaiced together taking any inundated area to be inundated but more will be explained in Section \ref{ssec:inundation_mapping}.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{NWS Main-stems}
 \label{sssec:nws_mainstems}
 %
-The initial attempt at drainage order reduction to solve the catchment boundary issue was to use a stream network relevant to the NWS forecasting community. 
-The Mainstems (MS) network is a subset of the NWM FR network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
+The Mainstems (MS) network is a subset of the NWM full-resolution (FR) network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
 The MS network comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
 It also spans 121,724 reaches across 1,608 HUC8s.
-In this technique, we derive HAND using the FR stream network as well as the MS network which was originally proposed by \citeA{djokic2019arc}.
-Inundation is derived independently from the resulting FR and MS HAND hydrofabric are mosaiced together using the technique proposed in Section \ref{ssec:inundation_mapping} to form the MS FIMs. 
-Within each HUC, one might typically only find a MS stream network of uniform stream order but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
-So while we may refer to the MS network as that of one with unit stream order, we acknowledge there are many cases where additional or converging forecast points create multiple branches within a given processing unit.
+HAND was originally derived for this stream network to enhance mapping skill along these critical MS segments \cite{djokic2019arc}. 
+The inundation derived from this stream network is mosaiced with the inundation from the FR network to form the MS FIMs. 
+Within each HUC, you'll typically only find a MS stream network of stream order 1 (i.e.
+headwater) but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
 %
@@
 %
-Since MS only covers 4\% of the entire FR stream network, we sought to expand drainage order reduction techniques to all reaches within the NWM modeling domain.
-In order to do this, we discretized the NWM network into LPs which when considered individually have unit Horton-Strahler stream orders.
-LPs group flowlines by maximizing the length of each flow path and minimizing the number of LP identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-In order to derive LPs for the NWM FR stream network at the HUC8 processing area, we first compute arbolate sums which are defined as the cumulative drainage distance of all upstream drainage lines.
-Arbolate sum is also inclusive of the current drainage reach as well.
-Arbolate sums are computed by starting at the headwater points and summing up drainage distances as you traverse downstream.
-
-Arbolate sum is critical to discretizing the NWM network into LP identifiers.
-Starting at a HUC8's outlet, a unique LP is propagated upstream. 
-At every confluence, the direction of maximum arbolate sum is sought to propagate the current LP identifier.
-For the remaining parent reaches of the given junction, a new LP identifier is assigned and the process recursively continues with them.
-Figure \ref{fig:level_path_methods} illustrates how LPs (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-The figure shows computed arbolate sums and unique LP identifiers for the HUC12 120903010404 for clarity but were computed at the corresponding HUC8.
-The mainstem of the figure runs from the red ellipses to the black one which is the outlet.
-From the figure, we can see how unique colors are propagated in the direction of the maximum arbolate sum.
-
-Each HUC8 is discretized into LPs independently and the relevant inputs as described in Table \ref{tab:data} are assigned to each LP processing unit given a buffer of seven km.
-This buffer was selected to avoid edge contamination \cite{lindsay2013measuring} and to ensure adequate data availability for wide rivers with large catchments in regions with low slope.
-Further work could be dedicated to tune this user exposed parameter to better balance its effect on FIM extents and computational expense since larger buffers create additional floating point calculations and storage requirements.
-For the time being, we designate this issue to be out of scope.
-
-At the LP scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the LP in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement by burning these lines and seeding from its headwater points directly instead of going through the NHDPlusHR network as described in Section \ref{ssec:stream_network_enforcment} .
-This decision was motivated by the difficulty in deriving LPs in the NWM stream network with high agreement with the LPs derived for the NHDPlusHR stream lines.
-We found that the same algorithm to compute arbolate sums and LPs could yield enough disagreements associated with disordered branches or slight differences in arbolate sums that could significantly affect the agreement of the LP identifiers in the NWM and NHDPlusHR networks.
-This yielded enough error to justify the use of the NWM directly for hydro-enforcement operations.
-
-Once the NWM FR stream network is discretized into LPs, we independently compute HAND using each LP as the target stream network to be used.
+To further the efforts implemented by MS, we sought to derive HAND at a level path scale which we call GMS.
+Since the MS network only covers a small percentage of the NWM forecasting domain, we sought to expand the benefits of stream order reduction within HAND processing units to the entire FR domain.
+Level paths group flowlines by maximizing the length of each flow path and minimizing the number of level path identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
+Starting at the outlet, a unique level path is propagated upstream. 
+At every confluence, the direction of maximum flow path length is sought to propagate the current level path identifier.
+For the remaining parent reaches of the given junction, a new level path identifier is assigned and the process recursively continues with them.
+Figure \ref{fig:level_path_methods} illustrate how level paths (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
+Each HUC8 is discretized into level paths independently and relevant inputs are assigned to each level path processing unit given a buffer of 7 km.
+At the level path scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the level path in question at the time.
+The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement which was motivated by the difficulty in deriving level paths in the NWM stream network with high agreement with the NHDPlusHR stream lines.
+
 To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
-In Figure \ref{fig:gms_methods}a, we uniquely color code the LPs derived for the NWM stream network. 
+In Figure \ref{fig:gms_methods}a, we uniquely color code the level paths derived for the NWM stream network. 
 For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each LP is buffered to a polygon with a user-exposed, distance parameter of seven km that is used to subset the original DEM for two selected LPs in Figure \ref{fig:gms_methods}b.
-We illustrate two HAND grids for two of the LPs in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each LP are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
-Lastly, these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}d.
-
-For a more intimate look at the drainage order reduction procedure and its effects, GMS, we allude to Figure \ref{fig:gms_methods_2} which references the same area (in HUC8 12090301) and set of river junctions as in Figure \ref{fig:catchment_boundaries_issue}.
-The catchments and stream lines for HAND computed at the FR scale are illustrated in Figure \ref{fig:gms_methods_2}a where the respective inundation at the 100 yr magnitude is heavily constrained by the limited catchment extents especially at junctions.
-In subsequent sub-figures, we show the same datasets for the HAND computation problem for this region but discretized into independent LPs for the main LP (b), the eastern tributary (c), and the western tributary (d).
-Notably, inspecting (b) one sees how removing the tributaries creates much larger catchments for the main LP. 
-These catchments include drainage areas that would traditionally be considered nearest to the tributaries thus ineligible to receive inundation sourced from the main LP.
-The inundation extents in (b) overlap those of (c) and (d) and are mosaiced together by methods explained in Section \ref{ssec:inundation_mapping}.
-%
-\begin{figure}[H]
+Each level path is buffered to a polygon with a user-available distance parameter of 7 km and this polygon is used to subset the original DEM for two selected level paths in Figure \ref{fig:gms_methods}b.
+We illustrate two HAND grids for two of the level paths in this HUC8 in Figure \ref{fig:gms_methods}c.
+Once the FIM hydrofabrics for each level path are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
+Lastly these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}c.
+%
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates the NWM Full Resolution V2.1 stream network discretized into LPs (LP), symbolized by unique colors, as well as the values of the arbolate sums to the nearest whole km distance.
-The LPs were derived on a HUC8 level (12090301) but only illustrated for a HUC12 (120903010404) for clarity.
-The mainstem of this HUC12 runs from the red ellipses to the outlet denoted by the black ellipses.
-Arbolate sums are defined as the cumulative drainage distances of all upstream stream lines.
-Arbolate sums are computed for the NWM network by starting at the headwater points then traversing downstream and adding the distances cumulatively. 
-LPs are derived by starting at an outlet point with a unique identifier (ID).
-The unique LP ID is propagated upstream until a junction is reached where the current LP ID is propagated in the direction of maximum arbolate sum.
-The remaining converging segments at the given junction are each assigned a new unique LP ID and the process is repeated recursively until all reaches have been assigned a LP.
-Thus, LP serve as a proxy means of assigning membership to a given river when presented with a confluence.
-Each individual LP has a unit Horton-Strahler stream order thus serves as a great method for our proposed technique.
+\caption{Illustrates how level paths for the NWM are derived.
+Level paths symbolized by lines of unique colors are propagated upstream following the direction of maximum arbolate sum at each junction.
 }
@@
 %
-\begin{figure}[H]
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=1.0]{figures/gms_methods.jpg}
-\caption{Overall procedure for GMS HAND at HUC8 12090301.
-In (a), we illustrate all NWM stream lines symbolized by their LP with 372 unique LP IDs in this HUC.
-Meanwhile (b), demonstrates the DEM clipped to a seven km buffer around two selected LPs.
-In (c), we show how HAND can be computed just for each one of these two LPs independently. 
-We also show inundation maps created for these two LPs in (c). 
-In (d), we show all the inundation maps for all the LPs mosaiced together. }
+\caption{Overall procedure for GMS HAND.
+In (a), we illustrate all NWM stream lines symbolized by their level path.
+Meanwhile (b), demonstrates the DEM clipped to a 7 km buffer around two selected level paths.
+In (c), we show how HAND can be computed just for each one of these two level paths independently. 
+We also show inundation maps created for these two level paths in (c). 
+In (d), we show all the inundation maps for all the level paths mosaiced together. }
 \label{fig:gms_methods}
@@
 %
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.70]{figures/gms_methods_2.jpg}
-\caption{This image, with the same spatial domain as Figure \ref{fig:catchment_boundaries_issue} (HUC 12090301), demonstrates how computing HAND on level path (LP) bases leads to larger, independent catchments and more expansive inundation extents (100 yr flows).
-In (a), the catchments and stream network are shown for HAND computed in Full-Resolution (FR) method which shows constrained inundation extents around the two junctions.
-(b) demonstrates the LP associated with this region's highest order river.
-By delineating catchments at this scale independent of the neighboring tributaries, the drainage areas are allowed to expand thus allowing inundation extents to cover previously restricted areas.
-In (c) and (d), we show the stream lines, catchments, and inundation extents of the two tributaries.
-Later in Section \ref{ssec:inundation_mapping}, we describe how the inundation in (b), (c), and (d) are mosaiced together to form one seamless inundation map.
-This process allows for multiple, possible contributing sources of fluvial inundation to be considered thus enhancing FIM skill.
-}
-\label{fig:gms_methods_2}
-\end{figure}
-%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
-From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC8 spatial processing units in a rapid, parallel operation.
+From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 8 spatial processing units in a rapid, parallel operation.
 The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
 
-Utilizing the stage-discharge relationships in the SRCs, each forecast for each catchment identifier is assigned a stage value. 
+Utilizing the stage-discharge relationships in the synthetic rating curves, each forecast for each catchment identifier is assigned a stage value. 
 The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
@@
 For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
-Consolidating the depths using a maximum function was decided upon based on intuition which we believe to best represent the depth of water in an area with multiple contributing fluvial inundation sources.
-Other aggregation methods could lead to different results but were not investigated here.
-Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i,
+Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i.
+Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps.
 %
@@
 %
-. Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps,
+For the MS and GMS methods, the contributing areas are defined differently.
+For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to 2 for that case.
+For GMS, all FIMs from all the level paths in a given area are mosaiced together then K is set to this number of level paths.
+Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
 %
@@
 \end{linenomath*}
-. For the MS and GMS methods, the contributing areas are defined differently.
-For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to two for that case.
-For GMS, all FIMs from all the LPs in a given area are mosaiced together then K is set to this number of LPs.
-Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
 %
@@
 %
-Possible benchmark FIM candidates for evaluation purposes include high-water marks, remote sensing observation, crowd-sourced information, and modeled extents.
-Unfortunately, all of these sources or using a combination of them are subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
-While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to isolating out hydrological factors.
-A suitable dataset needed to be selected to best balance advantages and limitations with future studies interested in expanding to additional datasets as able to.
-
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
-This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, more sophistication when compared to HAND based FIMs, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
-49 HUC8s spanning about 185 thousand $km^2$ were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
-The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM and thus corresponding discharges and mapping extents for evaluation.
-We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation sourced from OWP FIM.
-
-By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, and others that we would have potentially affected our conclusions if we used NWM forecasted discharges.
+Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated}.
+49 HUC 8's spanning about 185 thousand square km were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
+The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM so we used those corresponding discharges and mapping extents for evaluation.
+We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation.
+By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
 Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
@@
 %
-\begin{figure}[H]
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
-\caption{
-Shows 185 thousand $km^2$ of modeled areas for the Base Level Engineering (BLE) domain of 49 HUC8s across nine states at 0.2\% recurrence magnitude for flow rates.
-BLE maps are produced for two recurrence flows, 1\% (100 yr) and 0.2\% (500 yr), using 1D HEC-RAS models.
-The maps are used as benchmarks for validation purposes of OWP FIM.
-}
+\caption{Shows 185 thousand square km of modeled areas for BLE domain of 49 HUC8s across 9 states. This dataset for 1\% and 0.2\% recurrence flows were used as benchmarks.}
 \label{fig:all_ble_maps}
@@
 %
-\begin{figure}[H]
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
-\caption{
-Illustrates Base Level Engineering (BLE) cross sections and stream lines at the HUC8 12100203 near the confluences of West Fork Plum Creek and Clear Fork Plum Creek with Plum Creek.
-BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM V2.1 Full Resolution (FR) stream lines.
-Additionally, we illustrate the NWM V2.1 catchments to provide a sense of how many cross-sections may intersect a given NWM flowline.
-The BLE stream network is also shown which is denser than the NWM V2.1 stream lines meaning there are several lower order streams represented in the BLE stream network that are not in the NWM V2.1 stream lines.
-This creates additional inundation areas in the validation data that are not modeled with our HAND based FIMs.
+\caption{Illustrates Base Level Engineering (BLE) cross sections and stream lines.
+The BLE stream network, which is denser than the NWM V2.1 stream lines, is also shown.
+BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM reach.
 }
@@
 %
-
-The metrics employed in this study to evaluate inundation extents include CSI, Probability of Detection (POD), and False Alarm Ratio (FAR) and are presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
-To calculate these secondary metrics, one must define three primary metrics starting with true positives (TP) which is predicted wet and wet in the BLE benchmark dataset.
-The two types of errors consist of false positives (FP), or type I errors, which is dry in the benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in the benchmark but predicted dry.
+The metrics employed in this study to evaluate inundation extents include Critical Success Index (CSI), Probability of Detection (POD), and False Alarm Ratio (FAR) and presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
+To calculate these secondary metrics, one must define three primary metrics including true positives (TP) which is predicted wet and wet in benchmark dataset.
+The two types of errors consist of false positives (FP), or type I errors, which is dry in benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in benchmark by predicted dry.
 Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
 Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-In other words, POD is an indicator of model skill in inundated regions while FAR is an indicator of model skill in non-inundated regions.
-Some work by \citeA{gerapetritis2004behavior} denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
-We use all three secondary metrics here to add value to the discussion while avoiding aggregating away the meaning of all four primary metrics.
-
+Some work by \citeA{gerapetritis2004behavior} while at the NWM denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
 While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
 Thus, frequency dependent statistics should be used with caution when comparing across sites with varying frequencies. 
-Lastly, approximately six HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability to compare GMS metrics to.
+Lastly, approximately 6 HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability.
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Flood Mapping Performance}
-\label{ssec:flood_mapping_performance}
+\subsection{Mapping Performance}
+\label{ssec:mapping_performance}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
-We computed the statistics, CSI, POD, and FAR, for both 100 and 500 year events for Mannings N set to 0.06 and 0.12.
-These results are presented in Figure \ref{fig:violin_plot} as violin plots and in Table \ref{tab:aggregate_metrics} as aggregated metrics with the results for MS and GMS presented as percentage changes from their respective FR values.
-To be more specific, Table \ref{tab:aggregate_metrics} sums the primary metrics, TP, FP, FN, and TN, across all HUC8s then recomputes the secondary metrics which was done to better account for large variances in HUC8 size.
-The same trends discussed below are consistent across both reporting methods (Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}).
-
-The distribution of these flood extent metrics can be examined in Figure \ref{fig:violin_plot} as violin plots.
+We computed the statistics (CSI, POD, and FAR) for both 100 and 500 year events for Mannings N set to 0.06 and 0.12. 
+The distribution of these statistics can be examined in Figure \ref{fig:violin_plot} as violin plots.
 Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
-For example, let's examine the violin plot for the row marked CSI and column for Manning's n = 0.06.
-This sub-figure shows the CSI distributions across all 49 HUC8s when Manning's n is set to 0.06.
-Each independent, whole violin represents the distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100 yr and orange for 500 yr).
-The horizontal dashed and dotted lines represent the 25th, 50th, and 75th percentiles from bottom to top, respectively.
-Additionally, we show trend lines symbolized in green that for each metric and Manning's n combination denotes the best fit line for the three methods (FR, MS, and GMS).
-To avoid having two trend lines per sub-figure, we elected to aggregate the two magnitudes together as they tend to observe similar trends across the three models.
-The slope of each trend line is quantified in the figure by its $\beta_1$ value and the p-value for that statistic which tests the significance of that trend (deviation from a zero sloped trend line).
-
-% Interpretation of metrics
-Both Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} contain a fair amount of information that is central to the objectives of this study.
-As previously stated in Section \ref{ssec:evaluation}, we consider CSI as a general proxy for the skill of the inundation extents with POD denoting skill on inundated areas and FAR indicating skill on non-inundated areas.
-Again, the main objective of the study is to introduce how computing HAND with disaggregated stream networks to those with unit stream order can enhance the fidelity of FIMs by capturing fluvial inundation from multiple sources as opposed to that of just the nearest drainage line.
-As can be seen in Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, CSI generally increases from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
-This increase is primarily driven by an increase in POD thus generally increasing the probability of correctly detecting inundation.
-Also, we note that FAR is somewhat, albeit marginally, decreased from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
-The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}).
-So overall and in other words, the entire distribution of HUC8s improves across the three methods.
-Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit stream order and mosaic the resulting FIMs, the better those FIMs perform.
-We move more details on the relationship between stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).
-
-Additional noteworthy trends in Figure \ref{fig:violin_plot} center around the all-around better performance of FIMs for those of higher Manning's n values and recurrence flows.
-The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value used in the BLE model, 0.11 \cite{fema2021base,fema2021estimated}.
-Most of this improvement is driven by significant increases in POD but, unfortunately, leading to a significant amount of over-prediction as observed by the increase in FAR.
-More work can be invested to better regionalize Manning's n values for FIM purposes with HAND.
-We also observe additional trends associated with the magnitude or recurrence interval of the flow rates used with the higher flow rates exhibiting better overall CSI, POD, and FAR values than the lower, 100 yr magnitude.
-We introduce in the Discussion (Section \ref{sec:discussion}) that this skill premium exhibited by higher flow events could be due higher quality elevation data in regions that are not described as bathymetric areas.
-%
-\begin{figure}[H]
+We also denote trend lines for each metric and Manning's n setting as well as their respective slope estimate and one-tailed p-value denoting the level of significance of the trend.
+
+Aggregating the metrics in the method above treats each HUC8 as it's own unit and does little to consider the size differences of the HUCs. 
+In an opposing aggregation method, we illustrate in Table \ref{tab:aggregate_metrics}  the CSI, POD, and FAR recomputed for the entire domain using the sum of all the TPs, FPs, and FNs. 
+%
+\begin{figure}[h!]
 \centering
@@
 %
-\begin{table}[H]
-\caption{ Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated to the BLE domain.
-         The values for MS and GMS methods are expressed in percentage change (\%) from their respective values with the same Manning's n, magnitude, and metric combination in the Full Resolution (FR) method columns.
+\begin{table}[h!]
+\caption{Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated for BLE domain.
          The best value across models is highlighted in bold.}
@@
 \hline
-\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{\shortstack{MS\\(\% Change)}} & \multicolumn{2}{|c|}{\shortstack{GMS\\(\% Change)}} \\
+\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{MS} & \multicolumn{2}{|c|}{GMS} \\
 \cline{3-8}
-  &  & 100 yr & 500 yr & 100 yr & 500 yr & 100 yr & 500 yr \\
-\hline
-\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 2.53 & 2.59 & \textbf{3.95} & \textbf{4.04} \\
+  &  & 100yr & 500yr & 100yr & 500yr & 100yr & 500yr \\
+\hline
+\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 0.5717 & 0.5990 & \textbf{0.5796} & \textbf{0.6075} \\
 \cline{2-8}
-  & 0.12 & 0.5915 & 0.6149 & 2.35 & 2.26 & \textbf{4.51} & \textbf{4.65} \\
-\hline
-\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 2.68 & 2.74 & \textbf{4.39} & \textbf{4.38} \\
+  & 0.12 & 0.5915 & 0.6149 & 0.6054 & 0.6288 & \textbf{0.6182} & \textbf{0.6435} \\
+\hline
+\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 0.6524 & 0.6755 & \textbf{0.6633} & \textbf{0.6863} \\
 \cline{2-8}
-  & 0.12 & 0.7255 & 0.7446 & 2.83 & 2.71 & \textbf{4.84} & \textbf{4.89} \\
-\hline
-\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & -0.72 & \textbf{-1.24} & \textbf{-1.22} & \textbf{-1.24} \\
+  & 0.12 & 0.7255 & 0.7446 & 0.7460 & 0.7648 & \textbf{0.7606} & \textbf{0.7810} \\
+\hline
+\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & 0.1787 & \textbf{0.1589} & \textbf{0.1778} & \textbf{0.1589} \\
 \cline{2-8}
-  & 0.12 & 0.2379 & 0.2208 & -0.21 & -0.18 & \textbf{-2.31} & \textbf{-2.72} \\
+  & 0.12 & 0.2379 & 0.2208 & 0.2374 & 0.2204 & \textbf{0.2324} & \textbf{0.2148} \\
 \hline
@@
 \end{table}
+%
+% Interpretation of metrics
+From Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, we denote several meaningful trends. 
+Using CSI as an overall proxy for skill of the FIM, we note that generally speaking the skill is correlated with a reduction of the stream orders of the processing units used for HAND.
+In other words, the more we derive HAND on networks of unit drainage density and mosaic the resulting FIMs, the better those FIMs perform.
+While this trend is evident for both sets of Manning's n values, the trend is slightly more significant for the higher value of 0.12.
+Other trends related to this Figure include the general performance premium for 0.2\% events as opposed to lower 1\% events.
+We also note how the higher Manning's n value enhances performance for both of these recurrence intervals across all models.
+
+Dissecting the improvements and trends presented in the previous paragraph comes down mostly to improvement in POD or a reduction in absolute amount of FNs.
+POD being the primary driver in skill enhancement is evident across models by comparing the slope of the POD lines with the slope of the FAR lines.
+Even though aggregating metrics by HUC8 yields a statistically zero trend, one does see a slight reduction in FAR across models that reduce HAND's maximum stream order.
+Additionally, we note that POD is a primary driver in enhancing performance across Manning's n values as well.
+This significant improvement comes at a cost of false alarms as the FAR increases significantly across Manning's n values.
 %
@@
 For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
-GMS which generalizes HAND production to the LP scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
+GMS which generalizes HAND production to level path scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
 %
@@
 %
-Overall, the main observation of this study was how FIM performance can be improved by reducing the Horton-Strahler stream order of the target stream network used for HAND computation.
-Most of this change is accounted for by substantially increasing POD and inundation extents in some areas thus reducing FNs.
-We believe, as we later argue for, that the increase in POD is primarily driven by an increase in the catchment sizes that is inherently enabled by dividing up the stream network into independent stream networks of unit stream order.
-Additionally, we note that reducing drainage order also has some minor influence on reducing inundation extents in other areas and the rate of false alarms.
-We believe that this effect is driven by a change in the stage-discharge relationship where a given streamflow now leads to lower river stage values when HAND is computed with target stream networks of unit drainage order.
-We seek to explain that these two effects, catchment boundary enlargement and stage-discharge curve lowering, are highly interrelated and cannot be easily detangled.
-Lastly, we discuss the diminishing effects on performance that the MS and GMS techniques may have and also any additional effects including enhanced cross-walking abilities.
-
-As evident in the results of the study in Section \ref{sec:results}, a sizable amount of the increase in CSI observed by reducing stream order for HAND computation can be attributed to increases in POD.
-This can be inferred by close inspection of Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} where changes in POD are significantly higher than that of FAR.
-%This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
-Upon investigation of the performance of HAND derived FIM, we note a general increase of catchment sizes for the MS and GMS methods when compared to the FR method as they are now delineated independently of any tributaries that would constrain catchment sizes otherwise.
-Additionally, we note significantly less water build up along catchment boundaries especially at the junction of lower order tributaries with lower flow rates and higher order rivers with more flow.
-This allows for inundation extents to expand across regions previously encapsulated by catchments of joining reaches in lower flow tributaries.
-The water built up along the catchment boundaries can be thought of as a column of water in a cylindrical container (catchment) that has exceeded the elevation of the containers rim.
-This, of course, does not represent accurate physics but it is what does occur in HAND when water accumulates along catchment boundaries.
-
-Large scale HUC8 level evaluations can fail to demonstrate fine grain enhancements as they aggregate away many changes that are only clear at more local scales.
-Future assessments of OWP FIM should consider finer grain evaluation units as well possible impact assessments using asset information such as building footprints to better illustrate fine grain changes in a more relevant manner to stakeholders.
-For now, we offer up Figure \ref{fig:gms_enhancement} which best illustrates the improvement offered by multi-source fluvial flooding capabilities in a more local context.
-The figure is comprised of two agreement rasters for two different HAND based FIMs compared to the validation dataset for a given region with a high flow mainstem (500 yr recurrence flow) running horizontally along the region.
-Sub-figure \ref{fig:gms_enhancement}a demonstrates the agreement raster for the FR stream network as well as its respective catchment boundary lines symbolized in white and stream network shown in green.
-Inspection of this sub-figure denotes a clear spatial pattern where TPs or areas correctly inundated are pooled alongside catchment boundary lines. 
-On the other side of the catchment boundary, one can witness large swaths of FNs that should be inundated. 
-The FNs also exhibit a spatial pattern as in they tend to collocate within catchments of the pictured mainstems tributaries.
-This sort of behavior was introduced early in the paper and shown qualitatively in Figure \ref{fig:catchment_boundaries_issue}.
-
-As an enhancement, this paper proposes computing HAND for stream networks comprised of level-paths independently of one another.
-In sub-figure \ref{fig:gms_enhancement}b, the agreement raster for the GMS technique is illustrated as well as the stream network lines in green.
-While the entire mosaiced inundation map from GMS (as described in Section \ref{ssec:inundation_mapping} and Equation \ref{eq:hand_fim_depth}) is used to produce this agreement map, we only show the catchments associated with the mainstem of this region that is shown to follow a clear horizontal path.
-Showing all the catchments for the tributaries that were all derived independently would convolute the image.
-The main message illustrated here is that the catchments associated with the mainstem of this area significantly increase in size.
-Since they are not restricted by the catchments of tributaries that lie in the same drainage areas as those of the mainstem, they extend to include those as well.
-The consequence for inundation extents is a general increase in spatial coverage of the river's water which shows to be in much better agreement with the benchmark map.
-The TPs are no longer bounded by the catchment lines and allowed the expand to their natural extents.
-
-We note here as a contribution of this study that a major inherent, limitation of HAND is the ``nearest drainage'' constraint or the idea that a given river reach only drains or, in HAND's case, inundates its immediate, unique drainage area or catchment.
-In other words, HAND based FIMs are limited to producing fluvial inundation to only their nearest drainage area or catchment.
-However, we know that fluvial inundation can be sourced from several streams nearby that also serve as drainage outlets to the area in question.
-Generally speaking, drainage areas are known to be hierarchical in nature so a given drainage area for a given outlet point can be decomposed into nested drainage areas for outlet points that lie in the original drainage area.
-A perfect example of this are points that lie on tributary reaches closely neighboring a mainstem.
-These points lie in the drainage area of reaches in the mainstem but inundation from the mainstem cannot reach these tributary catchments because of the ``nearest'' assumption in HAND.
-Hence it's important to state that just like there are different sources of flooding such as fluvial, pluvial, groundwater, reservoir, barrier failure (dam/levee/embankment), and coastal, there can also be multiple sources of a riverine flood.
-HAND is only equipped to handle riverine flooding from the nearest drainage line.
-Other relevant drainage lines that produce fluvial flood waters are not considered here especially if the routing model used doesn't consider backwater effects.
-
-The remaining portion of the improvement in CSI was found to come from a marginal yet notable reduction in FAR.
-Upon investigation of the spatial results in the agreement maps, we found some areas of slight reductions in FPs especially where changes in catchment boundaries may have occurred due to the reduction in effective stream order in computing HAND.
-These observations pointed to changes in the SRCs introduced by stream order reduction and catchment definition adjustments.
-Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on SRCs.
-Sub-figure \ref{fig:rating_curve_comparison}a shows how the average SRCs for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to shift the curve down and to the right with ever increasing stream order reduction (FR to MS to GMS).
+Overall, we note a positive relationship between FIM skill and a reduction of the stream order of the stream network we use to derive the HAND datasets.
+Most of this change is accounted for by increasing POD thus reducing FNs especially along higher order rivers with higher flow magnitudes.
+We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
+This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
+One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
+
+In analyzing the data, we found a slight reduction in FAR was detected and more digging pointed to a bias in rating curves introduced by stream order reduction.
+Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on synthetic rating curves.
+Sub-figure \ref{fig:rating_curve_comparison}a shows how the average rating curves for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to bias down (and to the right) with ever increasing stream order reduction (FR to MS to GMS).
 This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
-
-Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged SRC relationship between stage and discharge.
+Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged synthetic rating curve relationship between stage and discharge.
 Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
 While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
-Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius but these associations only hold when reach length, L, is considered.
+Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius.
 Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively (see Equation \ref{eq:reach_averaged_mannings_equation}).
-Figures \ref{fig:rating_curve_comparison}b and \ref{fig:rating_curve_comparison}c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
+Figures \ref{fig:rating_curve_comparison} b and c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
 Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
 Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
-We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the SRC Q up and changes the SRC accordingly.
+We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the rating curve Q up and biases the rating curve down.
 In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
 Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
-
 Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments that encompass neighboring tributaries.
-These tributaries have a significant amount of bathymetry that is low-lying thus easily included in the geometry for the SRC derivation. 
+These tributaries have a significant amount of bathymetry that is low-lying thus easily including the SRC derivation. 
 They also contribute volume and bed area that is technically not perpendicular to the flux of streamflow being accounted for in the stream in question. 
@@
 
-We note that reducing stream order does in turn suffer from diminishing returns in which the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
-This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing stream orders ones and possibly twos and threes from stream order reduction and simply using the inundation from FR from these areas.
-
 Additional careful analysis of Figure \ref{fig:gms_enhancement}a, leads one to note many catchments that don't have inundation or significant inundation.
-While the cause of these errors can be varied, we assert here that conflating four networks for use in evaluations leads to significant error.
+While the cause of these errors can be varied, we assert here that conflating 4 networks for use in evaluations leads to significant error.
 As one may remember, Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
-One of the issues with the FR version of HAND occurs when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
-The utilization of MS and GMS only conflates to NWM catchments directly associated with the LP in question which is inherently easy to do with those methods. 
+One of the issues is when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
+The utilization of MS and GMS only conflates to NWM catchments directly associated with the level path in question which is inherently easy to do with those methods. 
 Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
-The NWM stream network was derived using the NHDPlus V2 dataset which was derived from coarser DEMs than those used here. 
+The NWM stream network was derived using the NHD medium resolution dataset which was derived from coarser DEMs than those used here. 
 Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
@@
 
-Our qualitative analysis suggests that the SRCs offer a significant opportunity for improvement in HAND based FIM for future development.
+Our qualitative analysis suggests that the synthetic rating curves offer a significant opportunity for improvement in HAND based FIM for future development.
 The bathymetry of the 10 m DEM from 3DEP is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
@@
 Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
-We suspect that a significant amount of the difference in performance between 100 yr and 500 yr magnitude events can be attributed to poor SRC performance due to poor bathymetric representation.
-Lower magnitude events are, logically, more susceptible to poor bathymetric data due a greater proportion of the inundation being attributable to areas that are more typically under normal flow conditions.
-Higher flow events tend to cover regions with more floodplain inundation thus less sensitive to errors from bathymetric data quality.
-On a related note, the use of the AGREE DEM method discussed in Section \ref{ssec:stream_network_enforcment} also interacts with the bathymetry issue introducing several artificial geometry parameters that affect SRC shape and quality.
-Due to focus on the nearest drainage problem, we leave future work related to SRC representation including roughness estimation, bathymetric data assimilation, and bathymetry adjustments as opportunities for major enhancements in HAND based FIM.
 
@@
 These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
-Ultimately, the use of more advanced 2D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
-%
-\begin{figure}[H]
+Ultimately, the use of more advanced 2-D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
+%
+\begin{figure}[h!]
 \centering
@@
 %
-\begin{figure}[H]
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
-\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105 at the 500 yr recurrence magnitude.
+\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105.
 Catchment boundaries and stream lines are shown in white and dotted green, respectively.
@@
 Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
-The National Water Model (NWM), developed by the National Oceanic and Atmospheric Administration's (NOAA) Office of Water Prediction (OWP) and the National Center for Atmospheric Research (NCAR), provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
-Additional modeling is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
+The National Water Model (NWM) developed by the National Oceanic and Atmospheric Administration's Office of Water Prediction, along with partners, provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
+Additional processing is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
 Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
 HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
-Scalable, open-source software, known as OWP FIM, was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
+Scalable, open-source software was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
 HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
-Finally, we used this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
-FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
-
-The main contribution and conclusion of this work centers around a fundamental limitation in HAND based FIM which is a failure to account for multiple possible sources of fluvial inundation since HAND only considers inundation from the nearest drainage line.
+Finally, we use this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
 We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
-In order to reduce stream order of the NWM stream network for HAND computation, we dissected the NWM network into two simpler units of singular Horton-Strahler stream order and mosaiced the resulting FIMs derived from each.
-The NWM Mainstems (MS) stream network, which covers roughly 4-5\% of the NWM Full Resolution (FR) network, spans all established forecasting points in the Advanced Hydrologic Prediction System (AHPS) and downstream reaches.
-The inundation from MS derived HAND is mosaiced together with the inundation of FR derived HAND.
-Extending order reduction to the entire network, the Generalized Mainstem (GMS) technique discretizes the NWM FR network into level paths (LP) of unit stream order for HAND computation.
-All LP based FIM derived from LP based HAND datasets are mosaiced together to form one seamless FIM.
-Dissecting the stream network into regions of LPs with unit stream order is necessary because HAND has a ``nearest drainage'' limitation meaning it only accounts for riverine inundation sourced from the nearest drainage line.
-In our evaluation of this technique, we observe that HAND based FIM improves in skill as we extend from nearest drainage inundation in FR to multiple drainage support in MS for only 4-5\% of the FR network.
-Extending multiple drainage support to the entire FR network with GMS based HAND improves upon skill at around the same magnitude that MS improved upon FR.
-Thus we conclude that deriving HAND with independent stream networks of unit Horton-Strahler stream order enhances the skill of FIM but offers diminishing returns as we extend from 4-5\% of the network with MS to 100\% of the network with GMS since deriving HAND and FIMs at these localized scale does add computational expense.
-
-This primary contribution also affects the parameters used to compute stage-discharge relationships shifting discharge higher at given stages which reduced the rate of false positives.
-This shift in SRC behavior is driven by larger catchments that influence reach averaged geometric parameters in the Manning's equation.
-Related to SRCs, we noted better results and more sensitivity to unit stream order networks with the higher Manning's n value of 0.12 when compared to 0.06 for high magnitude events at 1\% (100 year or yr) and 0.2\% (500 year or yr) recurrence intervals.
-Additionally, we noted better performance for more intense 500 yr events which we attribute to a stronger influence of poor quality bathymetric data in 100 yr magnitude inundation extents.
-While the AGREE DEM procedure is meant to add some bathymetry primarily motivated to enhance catchment and stream line delineation, it does introduce three parameters that have major implications in the quality of SRCs and the resulting FIMs.
+This also affects the parameters used to compute stage-discharge relationships biasing discharge higher at given stages which reduced the rate of false positives.
+FIM skill was evaluated over large spatial scales by comparsion to HEC-RAS 1D models.
+Further investment in the SRC's is warranted by accounting for bathymetric errors inherited by the DEM and better accounting for localized friction values at varying flow magnitudes.
 Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
-We leave questions related to Manning's n localization as well as bathymetry integration, estimation, and/or calibration open for future research to answer.
-Two other issues left open for improvement include the integration of higher resolution Lidar-based digital elevation maps (DEM) as well as the use of physics-based models for continental scale, high resolution forecasting applications.
 Due to inherent limitations with HAND, scalable, physics-based methods are necessary to consider to provide a better representation of flood extent dynamics in steady and unsteady conditions. 
 %
-% Acknowledgments
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Acknowledgments
-\acknowledgments
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-This work was funded by the Office of Water Prediction (OWP) which is part of the National Oceanic and Atmospheric Administration's (NOAA) National Weather Service (NWS).
-Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
-We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
-Additionally, David Blodgett from the United States Geolgical Survey (USGS) Water Mission Area was instrumental in helping define level paths and other hydrographic work.
-More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
-Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
- %
-% Open Research
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\openresearch
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
-OWP Flood Inundation Mapping (FIM) capabilities rely extensively on the use of the National Hydrography Plus High Resolution (NHDPlusHR) datasets including BurnLineEvents \cite{nhdplus2022vectors}, value-added attributes (VAA) \cite{nhdplus2022vectors}, water boundaries (WBD) or hydrologic unit code (HUC) geometries \cite{nhdplus2022wbd}, and digital elevation maps (DEM) \cite{nhdplus2022dems}.
-Some additional datasets for processing include the National Levee Database (NLD) furnished by the United States Army Core of Engineers (USACE) \cite{engineers2016national}, Land-sea border from the Great Lakes Hydrography Dataset (GLHD) furnished by the Great Lakes Aquatic Habitat Framework (GLAHF) \cite{GreatLakesHydrographyDataset}, and a Land-sea border provided by OpenStreetMap (OSM) \cite{osm2021landsea}.
-Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
-Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
-
-Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022}.
- %
-% Bibliography
-\clearpage % this clears figures before references
-\bibliography{bibliography/owp_fim4_2022}
  %
@@
 % AGU does not normally allow dedications.
+
+\acknowledgments
+This work was funded by the OWP which is part of NOAA's National Weather Service.
+Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
+We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
+Additionally, David Blodgett from the USGS Water Mission Area was instrumental in helping define level paths and other hydrographic work.
+More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
+Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
+%The [type of data] data used for [brief context, description] in the study are available at [repository, source name] via [DOI, persistent identifier link] with  [license, access conditions] [optional in-text citation in References]
+%[Version number] of the [software name] used for [brief context, description of what the software was used for] is preserved at [DOI, persistent identifier link], available via [license type, access conditions] and developed openly at [software development platform link].* [optional in-text citation in References].
+
 %% ------------------------------------------------------------------------ %%
@@
 %
+\clearpage % this clears figures before references
+\bibliography{bibliography/owp_fim4_2022}
 %

### 2022-09-19 manuscripts/owp_fim4_2022/reviews/first_review/responses.tex
Resubmitted to WRR

@@
 We have added several key clarifications in response to reviewer points and several new or updated figures that better explain our methods and results.
-There are few concerns that we had to leave out of this manuscript as this submission already covers a very substantial amount of work.
+However, we felt that we had to opt out of addressing a few of the concerns presented by the reviewers as this submission already covers a very substantial amount of work.
 The questions left open concern issues surrounding the synthetic rating curves including localizing Manning's n as well as bathymetry inclusion, estimation, and/or calibration.
+We believe these issues to be out of scope and ensured the resubmitted manuscript motivated these areas for future research. 
 We hope this greatly enhanced submission meets the expectations set by WRR and the concerns of our three reviewers adequately.

### 2022-09-19 manuscripts/owp_fim4_2022/sections/abstract_plainTextSummary.tex
Resubmitted to WRR

--- 2022-08-05 manuscripts/owp_fim4_2022/sections/abstract_plainTextSummary.tex
@@
 \begin{abstract}
-The National Water Model (NWM) currently requires additional modeling from discharges to produce forecast flood inundation maps (FIM) for protecting life and property. 
-Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is worthy of producing high-resolution FIMs at large spatial scales and frequent time steps using reach-averaged synthetic rating curves.
-However, HAND based FIMs suffer from a known limitation caused by independent catchments that lack the ability to cross catchment boundaries and ridgelines.
-We propose here that HAND is limited to producing inundation only when sourced from its nearest drainage line thus lacks the ability to source inundation from multiple fluvial sources.
-To counter this constraint, a version of HAND known as Generalized Mainstems (GMS) is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
-LPs are unique identifiers propagated upstream from a sub-basin’s outlet along the direction of maximum flow distance and repeated recursively until all segments are assigned LP identifiers.
-The FIMs associate from each independent LP are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
-Improvement in mapping skill is observed when the extents are compared to HEC-RAS 1D models by significantly reducing false negatives at river junctions.
-A more marginal reduction in the false alarm rate is also observed due to a bias introduced in the stage-discharge relationship by increasing the size of the catchments.
-This novel contribution is proposed and studied in a new open-source implementation that utilizes the latest combination of hydro-conditioning techniques to enforce drainage and counter limitations in the elevation data.
-This framework scales to the entire NWN modeling domain including Hawaii and Puerto Rico to help forecasters provide timely and accurate flood warnings.
+Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is a means able of producing high resolution flood inundation maps (FIMs) from the National Water Model (NWM) at large scales and high resolutions using reach-averaged synthetic rating curves.
+We highlight here that HAND is limited to producing inundation only when sourced from its nearest drainage line, thus lacks the ability to source inundation from multiple fluvial sources.
+A version of HAND, known as Generalized Mainstems (GMS), is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
+The FIMs associated with each independent LP are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
+Improvement in mapping skill is observed by significantly reducing false negatives at river junctions when the inundation extents are compared to FIMs from that of benchmarks.
+A more marginal reduction in the false alarm rate is also observed due to a shift introduced in the stage-discharge relationship by increasing the size of the catchments.
+We observe that the improvement of this method applied at 4-5\% of the entire stream network to 100\% of the network is about the same magnitude improvement as going from no drainage order reduction to 4-5\% of the network.
+This novel contribution is framed in a new open-source implementation that utilizes the latest combination of hydro-conditioning techniques to enforce drainage and counter limitations in the input data.
 \end{abstract}
@@
 The United States National Water Model (NWM) provides flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
-In order to convert forecasted flow rates from the NWM to flood inundation maps (FIM), a model is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model known as Height Above Nearest Drainage (HAND) suffers from issues in mapping performance where rivers converge.
+In order to convert flow rates from the NWM to flood inundation maps (FIM), a model, known as Height Above Nearest Drainage (HAND), is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
+This model suffers from issues in mapping performance because inundation sourced from rivers is only considered from the nearest river line.
 We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
+This is done by splitting the stream network into continuous river segments known as level paths (LPs).
+These LPs have no tributaries, thus are known to be stream lines with a unit stream order indicating no branching.
+HAND is computed independently for each LP and the resulting FIMs are mosaiced together to form one seamless map.
 We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance.

### 2022-09-19 manuscripts/owp_fim4_2022/sections/keypoints.tex
* Draft containing all revisions after first round of reviews * This version was submitted to co-authors for comments
Resubmitted to WRR

@@
 \begin{keypoints}
-\item Flood inundation maps derived from Height Above Nearest Drainage (HAND) are subject to limitations at river junctions.
-\item A means of resolving this limitation is provided by reducing HAND processing units to level paths of unit stream order.
-\item Changing the scale of the stream network for HAND processing affects the stage-discharge relationship and leads to higher skill inundation extents.
+\item Flood maps derived from Height Above Nearest Drainage (HAND) are subject to nearest drainage line limitations that affect inundation skill.
+\item A means of resolving this limitation is provided by reducing HAND processing units to level paths with effective unit stream order.
+\item Discretizing the stream network for HAND computation affects the stage-discharge relationship and leads to higher skill inundation.
 \end{keypoints}

### 2022-09-19 manuscripts/owp_fim4_2022/Makefile.sh

@@ -1,20 +1,33 @@
 #!/bin/bash -e
 
+# declare files
+mainFile=owp_fim4.tex
+bibFile=bibliography/owp_fim4_2022.bib
+expandedFile=owp_fim4_expanded.tex
+differenceFile=owp_fim4_expanded_diff.tex
+originalExandedFile=owp_fim4_expanded_original.tex
+
+# record parent path
 parent_path=$( cd "$(dirname "${BASH_SOURCE[0]}")" ; pwd -P )
 cd $parent_path
 
-fullFile=owp_fim4.tex
-baseName="${fullFile%%.*}"
-extension="${fullFile#*.}"
-
-pdflatex "$fullFile"
-bibtex "$baseName".aux
-pdflatex "$fullFile"
-pdflatex "$fullFile"
+# compile document
+pdflatex "$mainFile"
+bibtex "${mainFile%%.*}".aux
+pdflatex "$mainFile"
+pdflatex "$mainFile"
 
 # expands all tex files into one file
-echo 'Lat-Expand to Create One Tex File ...'
-latexpand --keep-comments --biber bibliography/owp_fim4_2021.bib owp_fim4.tex -o owp_fim4_expanded.tex
+echo 'Latexpand to Create One Tex File ...'
+latexpand --keep-comments --biber "$bibFile" "$mainFile" -o "$expandedFile"
 
-
+# makes difference file
+if test -f "$originalExandedFile"; then
+    echo 'Latexdiff to make a difference file'
+    latexdiff "$originalExandedFile" "$expandedFile" > "$differenceFile"
+    pdflatex "$differenceFile"
+    bibtex "${differenceFile%%.*}".aux
+    pdflatex "$differenceFile"
+    pdflatex "$differenceFile"
+fi
 

### 2022-09-21 manuscripts/owp_3dep_2022/owp_3dep.tex
- potentially working 3dep file - latex document upgrades to intro - dockerfile manuscript removal - temp file cleanup - build dir added

@@
 
-\documentclass[preprint,12pt]{dependencies/elsarticle}
+\documentclass[preprint,review,12pt]{dependencies/elsarticle}
 
@@
 \usepackage{lineno}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 %% Additional packages
-\usepackage{acronym}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\usepackage[printonlyused]{acronym}
+%\usepackage{hyperref}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \journal{Advances in Water Resources}
-
+%
 \begin{document}
-
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% FRONT MATTER
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{frontmatter}
-
+%
 %% Title, authors and addresses
-
+%
 %% use the tnoteref command within \title for footnotes;
@@
 
-\title{Varying the Spatial Resolution of LiDAR Derived Elevations to Balance Skill and Computational Costs for Flood Inundation Mapping Applications}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% TITLE 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
+\title{Investigating the Effect of Varying the Spatial Resolutions of LiDAR Derived Elevations on Flood Inundation Maps Produced from Height Above Nearest Drainage}
 
-\author[lynk,nwc,uf]{Fernando Aristizabal}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% AUTHORS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\author[lynk,nwc,uf]{Fernando Aristizabal\corref{fa_corr}}
+\ead{fernando.aristizabal@noaa.gov}
+\ead[url]{https://www.linkedin.com/in/fernando-aristizabal}
+%\fntext[fa_corr]{Please direct correspondence to Fernando Aristizabal.}
+\cortext[fa_corr]{Please direct correspondence to Fernando Aristizabal.}
+
 \author[nwc]{Fernando Salas}
@@
 \author[uf]{Jasmeet Judge}
-
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% AFFILIATIONS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \affiliation[lynk]{
-             organization={Lynker}, %Department and Organization
+             organization={Lynker},
              addressline={338 E Market St}, 
@@
 \affiliation[nwc]{
-             organization={National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration}, %Department and Organization
+             organization={National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration},
              addressline={205 Hackberry Ln}, 
@@
             }
-
+%
 \affiliation[uf]{
-             organization={Center for Remote Sensing, Agricultural and Biological Engineering, University of Florida}, %Department and Organization
+             organization={Center for Remote Sensing, Agricultural and Biological Engineering, University of Florida},
              addressline={1741 Museum Rd}, 
@@
             }
-
+%
 \affiliation[uh]{
@@
             }
-
+%
 \affiliation[usgs]{
@@
             }
-
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ABSTRACT
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \begin{abstract}
-%% Text of abstract
 
 \end{abstract}
-
-%%Graphical abstract
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Graphical abstract
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 %\begin{graphicalabstract}
@@
 %\end{graphicalabstract}
-
-%%Research highlights
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Research highlights
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \begin{highlights}
 \item Continental-scale flood inundation mapping requires simplifying assumptions
-\item Utilizing LiDAR derived data from the USGS 3DEP program enhances the skill on flood inundation maps produced from HAND.
+\item Utilizing LiDAR derived data from the USGS 3DEP program enhances the skill on flood inundation maps produced from \ac{HAND}.
 \item Varying the spatial resolution provides no skill enhancement when evaluated at large scales across three study sites and benchmark datasets.
 \end{highlights}
-
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% KEYWORDS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% MAX SIX WORDS %%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 \begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-
+Hydrology \sep
+Height Above Nearest Drainage \sep
+Flood Inundation Mapping \sep
+Digital Elevation Maps \sep
+Forecasting
+%
 %% PACS codes here, in the form: \PACS code \sep code
-
+\PACS 92.40.-t \sep 92.40.Qk \sep *92.40.Q- \sep *92.40.qp \sep 89.60.Gg \sep 92.60.Wc \sep 91.10.Jf \sep 84.40.Xb
+%
 %% MSC codes here, in the form: \MSC code \sep code
 %% or \MSC[2008] code \sep code (2000 is the default)
-
+%
 \end{keyword}
-
+%
 \end{frontmatter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Acronyms
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acronym}
-
-\acro{OWP}[OWP]{Office of Water Prediction}
-
-\end{acronym}
-
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% MAIN TEXT
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% prints line numbers
 \linenumbers
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% main text
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% INTRODUCTION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 %
-Floods are among the most frequent, damaging, and deadly of natural disasters \citet{doocy2013human,stromberg2007natural,kahn2005death}. 
-What's worse, trends in flood events have been increasing in frequency and intensity driven by climate change, anthropogenic development, \citet{berz2000flood,mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect} with more hydrologic extremes expected in the future \citet{kahn2005death,tabari2020climate,milly2002increasing,wing2018estimates}.
-Floods impact mortality and morbidity through drowning, trauma, or increased risk of infectious disease \citet{jonkman2005global,beinin2012medical,alajo2006cholera,french1983mortality}.
-Flooding impacts population inequitably \citet{kahn2005death,smiley2021social,stromberg2007natural,jonkman2005global,tellman2020using,tellman2021satellite}.
-Early-warning systems can help \citet{stromberg2007natural}
-The \ac{OWP}.
+% Flush acronym usage
+\acresetall 
+%
+%%%% Motivates and introduces flooding %%%%
+Floods are among the most frequent, damaging, and deadly of natural disasters \citep{doocy2013human,stromberg2007natural,kahn2005death}. 
+The frequency and intensity of flood events as well as the exposure of people and property to them have been increasing in recent times driven by secular changes in climate, infrastructure, and demographics \citep{berz2000flood,mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
+Unfortunately, these trends are expected to continue placing additional pressure on hydrological extremes \citep{kahn2005death,tabari2020climate,milly2002increasing,wing2018estimates}.
+Floods impact mortality and morbidity through drowning or physical trauma at the individual health scale, while increasing the risk of infectious disease at the public health level \citep{jonkman2005global,beinin2012medical,alajo2006cholera,french1983mortality}.
+Flooding disrupts systems providing human needs such as transportation routes, supply chains, water delivery, waste management, communications, and energy grids \citep{wijkman2021natural}.
+These impacts disproportionately affect certain demographics such as the socioeconomically-disadvantaged, youth, and elderly who are more likely to live in vulnerable areas with less access to educational resources, \acp{EWS}, and the capacity or resources to evacuate impacted areas \citep{kahn2005death,smiley2022social,stromberg2007natural,jonkman2005global,tellman2020using,tellman2021satellite}.
+These inequitable impacts further entrench poverty and inequalities \citep{stallings1988conflict,birkmann2010extreme}.
+In political terms, severe disasters, including floods, can reduce social order, strain governance systems, collapse social safety nets, increase the risk of social conflict \citep{drury1998disasters,xu2016natural,zahran2009natural}.
+These dire consequences motivate adaption and mitigation efforts such as \acp{EWS}, protective infrastructure (e.g.
+storage, defenses, drainage, infiltration), public awareness and education, and zoning regulations \citep{tumbare2000mitigating,tauhid2018mitigating,charlesworth201115}.
 
+%%%% Motivates and introduces NWM %%%%
+Due to the growing consequences and risks presented by increasing flood impacts, \acp{EWS}, or forecasting systems, can help understand future conditions and provide intelligence to furnish adequate warnings to protect life, prevent damages, and enhance resilience \citep{stromberg2007natural,cools2016lessons,unisdr2015making,baudoin2014early,golnaraghi2012overview,unep2012early,liu2018review}.
+The early warning of flood disasters at national scales requires the use of continental-scale, forecast hydrology models and modeling frameworks that span intranational political boundaries.
+The applications of these models extend beyond \acp{EWS} to provide historical trends for applications in infrastructure planning, public planning, insurance underwriting, and more.
+The \ac{OWP}, an office of the \ac{NOAA} along with partners at the \ac{NCAR}, developed such a continental-scale model known as the \ac{US} \ac{NWM} \citep{salas2018towards,gochis2021wrf,cosgrove2019evolution,cohen2018featured,noaa2016national,water2022nwm}.
+The \ac{NWM} is based on a configuration of the \ac{WRF-Hydro} model that accounts for land surface processes as well as overland and channel routing \citep{gochis2021wrf,salas2018towards,cosgrove2019evolution}.
+Operationally, the \ac{NWM} produces streamflow analysis and forecasts at multiple time horizons depending on location which include the \ac{CONUS}, \ac{PR}, and \ac{HI} \citep{cosgrove2019evolution,noaa2016national,water2022nwm}.
+The \ac{NWM} routes streamflow across the \ac{NWM} \ac{V2.1} stream network, based on the \ac{NHDPlusV2} network, is comprised of more than 5.5 million \acp{km} of lines discretized into more than 2.8 million forecast points \citep{aristizabal2022reducing}.
+The \ac{NWM} \ac{V2.1} stream network belongs to the \ac{NWM} ``hydrofabric'' defined as a catalog of geospatial layers relevant to hydrology modeling including stream network lines, catchments, reservoirs, and more \citep{water2022nwm,cosgrove2019evolution}.
+While streamflow is an important variable for engineering and scientific applications of fluvial flooding, flood inundation stages, extents and depths are much more tangible variables to the stakeholders flood events directly impact.
+
+%%%% Introduces and motivates HAND %%%%
+The \acp{SWE}, a system of two hyperbolic partial differential equations, formally govern the flow of fluvial surface water by conserving both mass (first equation) and momentum (second equation) and can be expressed in both the \ac{1D} (Saint Venant Equations) and \ac{2D} forms.
+Solving this system in full \ac{2D} form requires numerical methods that can be very cost prohibitive and numerically unstable in an operational setting across continental-scales at high spatial discretizations (10 \ac{m} or higher).
+This use case motivates the implementation of an inundation proxy, also known as a zero-physics or a simplified conceptual model, that is agnostic to the \acp{SWE} while still computing accurate fluvial inundation extents and depths for this problem \citep{teng2015rapid}.
+\ac{HAND} detrends elevations within \acp{DEM} to compute drainage potentials by normalizing elevations to the nearest, relevant drainage line instead of datums that represent mean sea level \citep{renno2008hand,nobre2011height,nobre2016hand}.
+\ac{HAND} as a terrain index has been used extensively for \ac{FIM} purposes from both modeled or observed stream flows and stages \citep{nobre2016hand,afshari2018comparison,garousi2019terrain,johnson2019integrated,zheng2018geoflood,zheng2018river,zhang2018comparative,teng2015rapid,li2022accounting,li2020evaluation}, as well as for assisting the remote sensing detection of fluvial inundation \citep{aristizabal2020high,shastry2019using,aristizabal2021mapping,huang2017comparison,twele2016sentinel}.
+\ac{HAND} operates as an inundation proxy by thresholding the relative elevation (or \ac{HAND}) values with a singular river stage value for each catchment corresponding to the drainage area of a given river reach \citep{nobre2016hand,garousi2019terrain,johnson2019integrated,zheng2018geoflood,teng2015rapid,li2020evaluation,liu2016cybergis,maidment2017conceptual,liu2018cybergis,liu2020height,liu2018review}.
+When used to generate inundation extents and depths from streamflow, reach-averaged \ac{SRC} sample geometric variables along an entire reach and normalize using the length of the reach to create stage-discharge relationships \citep{zheng2018river,aristizabal2022reducing,godbout2019error}.
+These relationships depend on a friction parameter, Manning's n, and used to convert streamflows to stages for eventual \ac{2D} mapping with \ac{HAND}.
+Numerous investigations have validated the use of \ac{HAND} for flood mapping applications as a suitable alternative to more sophisticated physics-based techniques for large scale and high resolution use cases \citep{johnson2019integrated,li2020evaluation,li2022comparative,aristizabal2022reducing,nobre2016hand,godbout2019error,afshari2018comparison,zhang2018comparative,teng2015rapid,teng2017flood,diehl2021improving}.
+
+%%%% Motivates and introduces OWP FIM %%%%
+Several prior and active large-scale \ac{HAND} implementations catered to operational \ac{EWS} applications including the \ac{NFIE} \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis}, GeoFlood \citep{zheng2018geoflood}, and \ac{PyGFT}, i.e.
+\ac{GIS} \citep{petrochenkov2020pygft,verdin2016software}.
+The \ac{NFIE} was a broad, inter-institutional effort to apply HAND to the initial versions of the \ac{NWM} which leveraged 10 \ac{m} (1/3 arc-second) elevation data available at the time \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis} from the \ac{USGS}'s \ac{NED} \citep{gesch2009national,gesch2014accuracy} which now has been subsumed \citep{arundel2018assimilation} to The National Map program \citep{gesch2009national,archuleta2017national}.
+\citet{zheng2018geoflood} applies HAND for operational applications to \ac{LiDAR} derived 1 \ac{m} elevation data with a novel least cost, geodesic based stream delineation method \citep{passalacqua2010geometric,passalacqua2012automatic,zheng2018geoflood}.
+For applications with the \ac{NWM}, an advanced version of \ac{HAND} coupled with the use of \acp{SRC}, known as \ac{OWP} \ac{FIM}, converts \ac{NWM} analysis, reanalysis, and forecast streamflows to river stages and fluvial inundation depths and extents on an operational basis to \ac{CONUS} while extending the modeling domain to \ac{PR} and \ac{HI} \citep{aristizabal2022reducing,inundationMapping2022}.
+\ac{OWP} \ac{FIM} utilizes some of the latest datasets including the \ac{NHDPlusHR}, \ac{NLD}, and \ac{NWM} \ac{V2.1} hydrofabric to enforce drainage, hydrologically relevant features such as levees, and conflation with the forecast stream network \citep{aristizabal2022reducing,inundationMapping2022}.
+Additionally, \ac{OWP} \ac{FIM} advanced a fundamental limitation of \ac{HAND} that limits sourcing fluvial inundation only from the nearest, relevant drainage line \citep{mcgehee2016modified,aristizabal2022reducing,zhang2018comparative,li2022comparative,zheng2018geoflood,zheng2018river,nobre2016hand}.
+Stream lines of higher Horton-Strahler stream order that could contribute inundation to a given area have no way of extending beyond catchment lines which creates artificial bottlenecks in inundation extents especially along junctions of high order rivers with their lower flow tributaries \citep{aristizabal2022reducing,mcgehee2016modified}.
+To resolve this limitation, \ac{OWP} \ac{FIM} disaggregates the \ac{NWM} \ac{V2.1} stream network into segments of effective unit stream order called \acp{LP} \citep{aristizabal2022reducing}.
+In terms of terrain data, \ac{OWP} \ac{FIM} utilizes the 10 \ac{m} \ac{DEM} from the \ac{NHDPlusHR} dataset which is based on the \ac{NED} due to its high fidelity and continental availability \citep{aristizabal2022reducing,moore2019user}.
+The previous advances in \ac{OWP} \ac{FIM} stopped short of accounting for new \ac{LiDAR} datasets that are now approaching continental scale availability from the \ac{USGS} \citep{aristizabal2022reducing}.
+
+%%%% 3DEP program introduction with LiDAR intro %%%%
+The \ac{USGS}'s \ac{3DEP} extends The National Map Program from the already existing 1/3, 1, and 2 arc-second spatial resolution seamless \acp{DEM} products to include a 1 \ac{m}, \ac{LiDAR} derived \ac{DEM} product as well.
+While other sources of observation are available for \ac{DEM} derivation such as \ac{IfSAR} in Alaska, we focus here on 1 \ac{m} data derived from \ac{LiDAR} and use that synonymously with the term \ac{3DEP}.
+
+3DEP divided up into hydrography, topographic, 
+Base specifications from USGS are \citep{arundel20151} for 1m and the rest are \citep{archuleta2017national}.
+
+Lidar technical details \citep{stoker2015usgs}.
+
+NED evaluation \citep{gesch2014accuracy,dobbs2010evaluation}
+
+Compares Lidar data to NED for flooding using HEC-RAS \citep{wang2005comparison}.
+
+The 1 \ac{m} \ac{3DEP} product is a hydro-flattened, topographic bare-earth raster \ac{DEM} gridded to 1 \ac{km} square shaped tiles with 6 pixels of overlap \citep{arundel20151}.
+The 1 \ac{m} \ac{DEM} is derived almost entirely from \ac{LiDAR} except in a few places where photogrammetry falling under the specifications of the \ac{LiDAR} data are used \citep{arundel20151}.
+According to specifications, the horizontal accuracy of 1 \ac{m} \ac{3DEP} is expected to be within 1 \ac{m} while the vertical accuracies are within 19.6 \ac{cm} and 30 \ac{cm} at the 95\% confidence interval for non-vegetative and vegetative regions, respectively \citep{arundel20151}.
+Non-vegetative vertical accuracies fall within 10 \ac{cm} \ac{RSME} \ac{arundel20151}.
+
+Elevation-derived hydrography specifications \citep{terziotti2020elevation}.
+
+% Specifications
+- Lidar base specifications \citep{heidemann2012lidar}
+- national map seamless products \citep{archuleta2017national}
+- 1m specifications \citep{arundel20151}
+
+%%%% assessments on 3DEP data %%%%
+Assessments of 3DEP \citep{stoker2022accuracy,kim2020positional,chirico2020evaluating,callahan2022vertical,scott2022statewide,dobbs2010evaluation}
+Assessment of airborne Lidar on bathymetry data quality \citep{hilldale2008assessing}.
+Assessment of Lidar \citep{tarolli2014high}
+
+%%%% FIM studies on importance of Lidar and elevation data %%%%
+Numerous reports and academic studies discuss the principal effect that topographic data quality has on the quality of \ac{FIM}.
+The National Enhanced Elevation Assessment conducted by \citet{dewberry2011final} quantified the estimated annual benefits of high resolution, enhanced elevation data by 27 business sectors and concluded flood risk management to have the highest conservative estimate among the sectors in 2015 \ac{USD}.
+Several technical reports document that the quality of the elevation map used is of primal significance to the quality of the flood maps produced \citep{national2007elevation,national2009mapping}.
+
+Make sure to mention geoflood \citep{zheng2018geoflood,zheng2019automatic}.
+Make sure to mention the three limitations of elevation data include bathymetry, embankment delineations, and anthropogenic features such as bridges.
+
+DEMs and flooding \citep{casas2006topographic,thomas2016quantifying,savage2016does,passalacqua2010geometric,passalacqua2012automatic,munoth2019effects}
+
+Effects of resolution on hilly areas \citep{dai2019effects}
+
+Changed the horizontal resolution of DEM for HAND \citep{li2022accounting}, showed that DEM resolution was the most important factor influencing FIM extent skill when varied alongside water depth and drainage threshold.
+
+Flood specific assessments for 3DEP \citep{carswell20183d,bales2009sources,gesch2018best,podhoranyi2015inaccuracy,lamichhane2018effect,tsubaki2013uncertainty,dobbs2010evaluation,arrighi2019effects,zazo2015analysis,bhuyian2018accounting,gesch2012elevation}.
+
+\citep{garousi2019terrain} using 3m vs 10m with HAND.
+
+Effects of flat areas on FIM quality and overestimation \citep{garousi2019terrain,godbout2019error,jafarzadegan2017based,papaioannou2017probabilistic}
+
+\citep{lopez2018influence} how DEM resolution affects hydro connectivity and inundation extents.
+
+%%%% Motivation paragraph for this study %%%%
+With these latest developments in terrain data availability, we propose investigating the integration of \ac{3DEP} data into \ac{OWP} \ac{FIM} for continental-scale inundation forecasting abilities.
+We will investigate utilizing \ac{3DEP} data for \ac{HAND} computation to generate the \ac{FIM} hydrofabric.
+Additionally, we investigate the utility of varying spatial resolutions from 1, 3, 5, 10, 15, and 20 \ac{m}.
+\ac{HAND} depends on the drainage assumptions which requires \acp{DEM} to undergo a long series of enforcement processes to ensure monotonically decreasing elevations with hydrologically correct flow directions \citep{garousi2019terrain,nobre2011height,nobre2016hand,aristizabal2022reducing}.
+The resampling of \acp{DEM} into varying spatial resolutions could interact with these hydro-conditioning operations thus influencing the \ac{FIM} hydrofabric and the resulting quality of the \acp{FIM} produced.
+We seek to quantify this relationship and assess how these interact with one another to optimize for the quality of continental-scale \acp{FIM} from the \ac{NWM}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% MATERIAL AND METHODS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Material and Methods}
+\label{sec:material_and_matheds}
+%
+\ac{HAND} assumes monotonically decreasing elevations with hydrologically relevant flow paths ensuring all cells in a region drain \citep{renno2008hand,nobre2011height,nobre2016hand}.
+This requires the use of extensive hydro-conditioning techniques on \acp{DEM} to enforce drainage and agreement with existing hydrography \citep{aristizabal2022reducing,maidment2017conceptual,liu2016cybergis,liu2020height}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{\ac{3DEP}}
+\label{ssec:3dep}
+%
+%% make a table of different 3DEP data sources
+\begin{table}[h!]
+ \begin{center}
+  \caption{\acf{3DEP} product specifications.}
+  \label{tab:table1}
+  \begin{tabular}{c|c|c|c|c|c} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
+   Program & Product & Arc-Second Resolution & \ac{m} Resolution & Horizontal Accuracy (\ac{m}) & Vertical Accuracy (\ac{m})\\
+   \hline
+   1 & 1110.1 & a &  &  &  \\
+   \hline
+   2 & 10.1 & b &  &  &  \\
+   \hline
+   3 & 23.113231 & c &  &  &  \\
+   \hline
+  \end{tabular}
+ \end{center}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% RESULTS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+\label{sec:results}
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% DISCUSSION
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+\label{sec:discussion}
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% CONCLUSIONS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+\label{sec:conclusions}
+%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Flush acronym usage
+\acresetall 
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ACKNOWLEDGMENTS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Acknowledgments}
+\label{sec:acknowledgments}
+%
+Collate acknowledgements in a separate section at the end of the article before the references and do not, therefore, include them on the title page, as a footnote to the title or otherwise.
+List here those individuals who provided help during the research (e.g., providing language help, writing assistance or proof reading the article, etc.).
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% FUNDING SOURCES
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Funding Sources}
+\label{sec:funding_sources}
+%
+List funding sources in this standard way to facilitate compliance to funder's requirements:
+
+Funding: This work was supported by the National Institutes of Health [grant numbers xxxx, yyyy]; the Bill \& Melinda Gates Foundation, Seattle, WA [grant number zzzz]; and the United States Institutes of Peace [grant number aaaa].
+
+It is not necessary to include detailed descriptions on the program or type of grants and awards.
+When funding is from a block grant or other resources available to a university, college, or other research institution, submit the name of the institute or organization that provided the funding.
+If no funding has been provided for the research, it is recommended to include the following sentence:
+This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% APPENDIX
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 %% The Appendices part is started with the command \appendix;
 %% appendix sections are then done as normal sections
-%% \appendix
+\appendix
 
@@
 %% \label{}
-
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ACRONYMS
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Acronyms}
+\label{sec:acronyms}
+%
+\begin{acronym}
+\acro{OWP}{Office of Water Prediction}
+\acro{EWS}{early warning system}
+\acro{NWM}{National Water Model}
+\acro{NOAA}{National Oceanic and Atmospheric Administration}
+\acro{NWC}{National Water Center}
+\acro{NFIE}{National Flood Interoperability Experiment}
+\acro{NWS}{National Water Service}
+\acro{NCAR}{National Center for Atmospheric Research}
+\acro{WRF-Hydro}{Weather Research and Forecasting Hydro}
+\acro{US}{United States}
+\acro{USGS}[US Geological Survey]{United States Geological Survey}
+\acro{USD}{US Dollar}
+\acro{BLE}{Base Level Engineering}
+\acro{FEMA}{Federal Emergency Management Agency}
+\acro{HAND}{Height Above Nearest Drainage}
+\acro{TP}{true positive}
+\acro{FP}{false positive}
+\acro{TN}{true negative}
+\acro{FN}{false negative}
+\acro{CSI}{critical success index}
+\acro{POD}{probability of detection}
+\acro{FAR}{false alarm rate}
+\acro{3D}{3-Dimensional}
+\acro{2D}{2-Dimensional}
+\acro{1D}{1-Dimensional}
+\acro{3DEP}{3-Dimensional Elevation Program}
+\acro{DEM}{digital elevation model}
+\acro{NLD}{National Levee Database}
+\acro{NHD}{National Hydrography Dataset}
+\acro{NED}{National Elevation Dataset}
+\acro{NHDPlus}{National Hydrography Dataset Plus}
+\acro{NHDPlusV2}{National Hydrography Dataset Plus Version 2}
+\acro{NHDPlusHR}{National Hydrography Dataset Plus High Resolution}
+\acro{LP}{level path}
+\acro{SRC}{synthetic rating curve}
+\acro{SWE}{Shallow Water Equations}
+\acro{CONUS}{Continental United States}
+\acro{PR}{Puerto Rico}
+\acro{HI}{Hawaii}
+\acro{V2.1}{Version 2.1}
+\acro{V4}{Version 4}
+\acro{km}{kilometer}
+\acro{cm}{centimeter}
+\acro{RMSE}{root mean squared error}
+\acro{$km^2$}{square kilometer}
+\acro{$m^2$}{square meter}
+\acro{m}{meter}
+\acro{FIM}{flood inundation map}
+\acro{LiDAR}{Light Detection and Ranging}
+\acro{PyGFT}{Python GIS Flood Tool}
+\acro{GIS}{Geographic Information Systems}
+\end{acronym}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BIBLIOGRAPHY
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
 %% For citations use: 
@@
 
-%% else use the following coding to input the bibitems directly in the
-%% TeX file.
-
-%\begin{thebibliography}{00}
-
-%% \bibitem[Author(year)]{label}
-%% Text of bibliographic item
-
-%\bibitem[ ()]{}
-
-%\end{thebibliography}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% END
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}
 \endinput
-%%
-%% End of file `elsarticle-template-num-names.tex'.

### 2022-09-21 dem_3dep_comparison.py

@@ -8,39 +8,64 @@ import subprocess
 import shutil
 import os
 from ast import literal_eval
+import boto3
+import tqdm
 
-def Dem_3dep_comparison( huc4=1202,
-                         resolutions=[20,15,10,5,3,1], 
-                         tile_sizes={20:10000,15:10000,10:5000,5:2000,3:1500,1:1000},
-                         huc_sizes ={20:8, 15:8, 10:8, 5:8, 3:12, 1:12},
-                         stream_resolutions=['FR','MS', 'GMS'],
-                         acquire_data=True,
-                         resolutions_to_acquire_data_for = {20,15,10,5,3,1},
-                         skip_hydrofabric=False,
-                         skip_evaluate=False,
-                         manning = 12,
-                         production = True,
-                         overwrite = True,
-                         overwrite_eval = True,
-                         remove_inputs = True,
-                         remove_outputs = True,
-                         hand_jobs_dict = { 
-                                            'FR' : 
-                                              { 1 : 3, 3 : 3, 5 : 7, 10 : 7, 15 : 7 , 20 : 7 },
-                                            'MS' :
-                                              { 1 : 5, 3 : 3, 5 : 7, 10 : 7, 15 : 7 , 20 : 7 },
-                                            'GMS' :
-                                              { 1 : [10,5], 3 : [15,15], 5 : [7,15], 10 : [7,15], 15 : [7,15] , 20 : [7,15] }
-                                           },
-                         eval_jobs_dict = { 
-                                            'FR' :
-                                              { 1 : 12, 3 : 12, 5 : 4, 10 : 7, 15 : 7 , 20 : 7 },
-                                            'MS' :
-                                              { 1 : 12, 3 : 12, 5 : 4, 10 : 7, 15 : 7 , 20 : 7 },
-                                            'GMS' :
-                                              { 1 : [3,4], 3 : [3,4], 5 : [2,4], 10 : [3,4], 15 : [3,4] , 20 : [3,4]}
-                                              }
-        ):
+huc4 = 1202
+resolutions=[20,15,10,5,3,1]
+tile_sizes = {100:100000,20:10000,15:10000,10:5000,5:2000,3:1500,1:1000}
+huc_sizes = {100:8,20:8, 15:8, 10:8, 5:8, 3:12, 1:12}
+stream_resolutions=['FR','MS', 'GMS']
+acquire_data=True
[…]

### 2022-09-21 manuscripts/owp_3dep_2022/Makefile.sh

@@ -1,20 +1,42 @@
 #!/bin/bash -e
 
+# declare files
+mainFile=owp_3dep.tex
+bibFile=bibliography/owp_3dep_2022.bib
+#expandedFile=owp_3dep_expanded.tex
+#differenceFile=owp_3dep_expanded_diff.tex
+#originalExandedFile=owp_3dep_expanded_original.tex
+
+# record build directory
+buildDir=build
+
+# record parent path
 parent_path=$( cd "$(dirname "${BASH_SOURCE[0]}")" ; pwd -P )
 cd $parent_path
 
-fullFile=owp_3dep.tex
-baseName="${fullFile%%.*}"
-extension="${fullFile#*.}"
+# get main file basename
+mainfile_base="${mainFile%%.*}"
 
-pdflatex "$fullFile"
-bibtex "$baseName".aux
-pdflatex "$fullFile"
-pdflatex "$fullFile"
+# compile document
+pdflatex -output-directory "$buildDir" "$mainFile"
+bibtex $buildDir/"$mainfile_base".aux
+pdflatex -output-directory "$buildDir" "$mainFile"
+pdflatex -output-directory "$buildDir" "$mainFile"
+
+# move pdf output
+mv $buildDir/"$mainfile_base".pdf "$mainfile_base".pdf
 
 # expands all tex files into one file
-#echo 'Lat-Expand to Create One Tex File ...'
-#/usr/bin/latexpand/latexpand --keep-comments --biber bibliography/owp_3dep_2022.bib owp_3dep.tex -o owp_3dep_expanded.tex
+#echo 'Latexpand to create one tex file ...'
+#latexpand --keep-comments --biber "$bibFile" "$mainFile" -o "$expandedFile"
 
-
+# makes difference file
+#if test -f "$originalExandedFile"; then
+#    echo 'Latexdiff to make a difference file'
+#    latexdiff "$originalExandedFile" "$expandedFile" > "$differenceFile"
+#    pdflatex "$differenceFile"
+#    bibtex "${differenceFile%%.*}".aux
+#    pdflatex "$differenceFile"
+#    pdflatex "$differenceFile"
+#fi
 

### 2022-09-21 manuscripts/owp_3dep_2022/docker_run_and_make.sh

@@ -1,5 +1,0 @@
-#!/bin/bash -e
-
-codeDir=$1
-
-docker run --rm -v "$1":/foss_fim fernandoaristizabal/latex:20220201 /foss_fim/manuscripts/owp_fim4_2022/Makefile.sh

### 2022-09-21 src/acquire_and_preprocess_inputs.py

@@ -8,6 +8,7 @@ sys.path.append('/foss_fim/src')
 import shutil
 from multiprocessing import Pool
 import geopandas as gpd
+import pandas as pd
 from urllib.error import HTTPError
 from tqdm import tqdm
 import xarray as xr
@@ -15,6 +16,7 @@ import py3dep
 import fiona
 from fiona.crs import to_string
 from shapely.geometry import shape,box,Polygon, MultiPolygon
+import shapely
 from rasterio.enums import Resampling
 from rasterio.errors import RasterioIOError
 import dask
@@ -315,7 +317,8 @@ def retrieve_and_reproject_3dep_for_huc( wbd,
                                          output_dir=None,
                                          ndv=elev_raster_ndv,
                                          num_workers=1,
-                                         wbd_buffer=2000
+                                         wbd_buffer=2000,
+                                         retry_3dep=False
                                        ):
 
     def __fishnet_geometry(geometry,cell_size_meters=1000,cell_buffer=10):
@@ -384,11 +387,12 @@ def retrieve_and_reproject_3dep_for_huc( wbd,
         # huc4 output dir
         huc4_output_dir = os.path.join(output_dir,f'{current_huc4}_{int(dem_resolution)}m')
         
-        # remove
-        shutil.rmtree(huc4_output_dir,ignore_errors=True)
+        if not retry_3dep:
+            # remove
+            shutil.rmtree(huc4_output_dir,ignore_errors=True)
         
-        # re-create
-        os.makedirs(huc4_output_dir,exist_ok=True)
+            # re-create
+            os.makedirs(huc4_output_dir,exist_ok=True)
 
         # buffer
         geometry = geometry.buffer(wbd_buffer)
@@ -418,7 +422,7 @@ def retrieve_and_reproject_3dep_for_huc( wbd,
                    [ndv for _ in range(number_of_boxes)],
                    huc4_output_dirs,
                    hucs,
-                   [idx for idx in range(number_of_boxes)]
+                   [int(idx) for idx in range(number_of_boxes)]
                  ]
 
 
@@ -443,15 +447,19 @@ def retrieve_and_reproject_3dep_for_huc( wbd,
 
     def __retrieve_and_process_single_3dep_dem(geometry,dem_resolution,wbd_crs,ndv,output_dir,huc,idx):
 
-        max_retries = 5
-        retries = 0
+        max_retries = 5; retries = 0
+        nhd_failed, failed_3dep = False, False
         while True:
             try:
+                
[…]

### 2022-09-21 tools/synthesize_test_cases.py

@@ -226,7 +226,7 @@ if __name__ == '__main__':
     parser.add_argument('-c','--config',help='Save outputs to development_versions or previous_versions? Options: "DEV" or "PREV"',required=False,default='DEV')
     parser.add_argument('-l','--calibrated',help='Denotes use of calibrated n values. This should be taken from meta-data from hydrofabric dir',required=False, default=False,action='store_true')
     parser.add_argument('-e','--model',help='Denotes model used. FR, MS, or GMS allowed. This should be taken from meta-data in hydrofabric dir.',required=True)
-    parser.add_argument('-v','--fim-version',help='Name of fim version to cache.',required=False, default="all")
+    parser.add_argument('-v','--fim-version',help='Name of fim version to cache.',required=False, default="all",nargs='+')
     parser.add_argument('-jh','--job-number-huc',help='Number of processes to use for HUC scale operations. HUC and Batch job numbers should multiply to no more than one less than the CPU count of the machine.',required=False, default=1,type=int)
     parser.add_argument('-jb','--job-number-branch',help='Number of processes to use for Branch scale operations. HUC and Batch job numbers should multiply to no more than one less than the CPU count of the machine.',required=False, default=1,type=int)
     parser.add_argument('-s','--special-string',help='Add a special name to the end of the branch.',required=False, default="")
@@ -266,7 +266,7 @@ if __name__ == '__main__':
                         )
 
     # Default to processing all possible versions in PREVIOUS_FIM_DIR. Otherwise, process only the user-supplied version.
-    if fim_version != "all":
+    if (fim_version != "all") & (not isinstance(fim_version,list)):
         previous_fim_list = [fim_version]
     else:
         if config == 'PREV':
@@ -315,6 +315,7 @@ if __name__ == '__main__':
                     # Loop through versions.
                     for version in previous_fim_list:
                         
+                        #print(version,type(version),previous_fim_list,type(previous_fim_list));exit()
                         version = os.path.basename(version)
                         
                         if config == 'DEV':

### 2022-09-21 manuscripts/owp_3dep_2022/owp_3dep.tex
added build dir mv pdf to makefile

@@
 The \ac{USGS}'s \ac{3DEP} extends The National Map Program from the already existing 1/3, 1, and 2 arc-second spatial resolution seamless \acp{DEM} products to include a 1 \ac{m}, \ac{LiDAR} derived \ac{DEM} product as well.
-While other sources of observation are available for \ac{DEM} derivation such as \ac{IfSAR} in Alaska, we focus here on 1 \ac{m} data derived from \ac{LiDAR} and use that synonymously with the term \ac{3DEP}.
+While other sources of observation are available for \ac{DEM} derivation such as \ac{InSAR} in Alaska, we focus here on 1 \ac{m} data derived from \ac{LiDAR} and use that synonymously with the term \ac{3DEP}.
 
@@
 \acro{GIS}{Geographic Information Systems}
+\acro{InSAR}{interferometric synthetic aperture radar}
 \end{acronym}

### 2022-09-21 clean_up_gms_files_script.sh

@@ -1,0 +1,10 @@
+#!/bin/bash -e
+
+for d in $(ls -d /data/outputs/3dep_test_*_1m_GMS_*of2/**/branches/*);do 
+    huc8=${d:24:8}
+    huc12=${d:45:12}
+    branch=$(basename $d)
+    version=${d:40:1}
+    
+    /foss_fim/src/gms/outputs_cleanup.py -d /data/outputs/3dep_test_"$huc8"_1m_GMS_"$version"of2/"$huc12"/branches/"$branch" -l /foss_fim/config/deny_gms_branches_default.lst -v -b
+done

### 2022-09-21 dem_3dep_comparison.py

@@ -157,7 +157,7 @@ def Dem_3dep_comparison( huc4 = huc4,
                                               '-n', f'3dep_test_{huc4}_{resolution}m_GMS',
                                               '-c', f'/foss_fim/config/params_{resolution}m.env',
                                               '-j', f'{bjobs}',
-                                              '-d', f'/foss_fim/config/deny_gms_branches_default.lst'
+                                              '-d', f'/foss_fim/config/deny_gms_branches_default.lst',
                                             ]
                     
                     if production:

### 2022-09-21 manuscripts/owp_3dep_2022/Makefile.sh

@@ -24,6 +24,7 @@ pdflatex -output-directory "$buildDir" "$mainFile"
 pdflatex -output-directory "$buildDir" "$mainFile"
 
 # move pdf output
+echo Moving $buildDir/$mainfile_base.pdf to main directory.
 mv $buildDir/"$mainfile_base".pdf "$mainfile_base".pdf
 
 # expands all tex files into one file

### 2022-09-21 manuscripts/owp_3dep_2022/Makefile.sh
Minor makefile changes

@@ -35,9 +35,9 @@ mv $buildDir/"$mainfile_base".pdf "$mainfile_base".pdf
 #if test -f "$originalExandedFile"; then
 #    echo 'Latexdiff to make a difference file'
 #    latexdiff "$originalExandedFile" "$expandedFile" > "$differenceFile"
-#    pdflatex "$differenceFile"
-#    bibtex "${differenceFile%%.*}".aux
-#    pdflatex "$differenceFile"
-#    pdflatex "$differenceFile"
+#    pdflatex -output-directory "$buildDir" "$differenceFile"
+#    bibtex -output-directory "$buildDir" "$buildDir"/"${differenceFile%%.*}".aux
+#    pdflatex -output-directory "$buildDir" "$differenceFile"
+#    pdflatex -output-directory "$buildDir" "$differenceFile"
 #fi
 

### 2022-09-22 manuscripts/owp_3dep_2022/owp_3dep.tex
manuscript updates

@@
 \usepackage[printonlyused]{acronym}
+\usepackage{multirow}
+\usepackage{comment}
 %\usepackage{hyperref}
@@
 
-\title{Investigating the Effect of Varying the Spatial Resolutions of LiDAR Derived Elevations on Flood Inundation Maps Produced from Height Above Nearest Drainage}
+\title{On the Effects of Spatial Resolution of LiDAR Derived Elevations on Flood Inundation Maps Produced from Height Above Nearest Drainage}
 
@@
 These relationships depend on a friction parameter, Manning's n, and used to convert streamflows to stages for eventual \ac{2D} mapping with \ac{HAND}.
-Numerous investigations have validated the use of \ac{HAND} for flood mapping applications as a suitable alternative to more sophisticated physics-based techniques for large scale and high resolution use cases \citep{johnson2019integrated,li2020evaluation,li2022comparative,aristizabal2022reducing,nobre2016hand,godbout2019error,afshari2018comparison,zhang2018comparative,teng2015rapid,teng2017flood,diehl2021improving}.
+Numerous investigations have validated the use of \ac{HAND} for flood mapping applications as a suitable alternative to more sophisticated physics-based techniques for large scale and high resolution use cases \citep{johnson2019integrated,li2020evaluation,li2022comparative,aristizabal2022reducing,nobre2016hand,godbout2019error,afshari2018comparison,zhang2018comparative,teng2015rapid,teng2017flood,diehl2021improving,hocini2021performance}.
 
 %%%% Motivates and introduces OWP FIM %%%%
-Several prior and active large-scale \ac{HAND} implementations catered to operational \ac{EWS} applications including the \ac{NFIE} \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis}, GeoFlood \citep{zheng2018geoflood}, and \ac{PyGFT}, i.e.
+Several prior and active large-scale \ac{HAND} implementations catered to operational \ac{EWS} applications including the \ac{NFIE} \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis}, GeoFlood \citep{zheng2018geoflood,hocini2021performance,d2022identification,carruthers2021assessment,zheng2022application}, and \ac{PyGFT}, i.e.
 \ac{GIS} \citep{petrochenkov2020pygft,verdin2016software}.
-The \ac{NFIE} was a broad, inter-institutional effort to apply HAND to the initial versions of the \ac{NWM} which leveraged 10 \ac{m} (1/3 arc-second) elevation data available at the time \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis} from the \ac{USGS}'s \ac{NED} \citep{gesch2009national,gesch2014accuracy} which now has been subsumed \citep{arundel2018assimilation} to The National Map program \citep{gesch2009national,archuleta2017national}.
-\citet{zheng2018geoflood} applies HAND for operational applications to \ac{LiDAR} derived 1 \ac{m} elevation data with a novel least cost, geodesic based stream delineation method \citep{passalacqua2010geometric,passalacqua2012automatic,zheng2018geoflood}.
+The \ac{NFIE} was a broad, inter-institutional, and pioneering effort to apply HAND to the initial versions of the \ac{NWM} which leveraged 10 \ac{m} (1/3 arc-second) elevation data available at the time \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis} from the \ac{USGS}'s \ac{NED} \citep{gesch2002national,gesch2007digital}.
+\citet{zheng2018geoflood} applies HAND for operational applications to \ac{LiDAR} derived 1 \ac{m} elevation data with a novel least cost, geodesic based stream delineation method \citep{passalacqua2010geometric,passalacqua2012automatic,zheng2018geoflood,zheng2019automatic,carruthers2021assessment,d2022identification,zheng2022application}.
 For applications with the \ac{NWM}, an advanced version of \ac{HAND} coupled with the use of \acp{SRC}, known as \ac{OWP} \ac{FIM}, converts \ac{NWM} analysis, reanalysis, and forecast streamflows to river stages and fluvial inundation depths and extents on an operational basis to \ac{CONUS} while extending the modeling domain to \ac{PR} and \ac{HI} \citep{aristizabal2022reducing,inundationMapping2022}.
@@
 %%%% 3DEP program introduction with LiDAR intro %%%%
-The \ac{USGS}'s \ac{3DEP} extends The National Map Program from the already existing 1/3, 1, and 2 arc-second spatial resolution seamless \acp{DEM} products to include a 1 \ac{m}, \ac{LiDAR} derived \ac{DEM} product as well.
-While other sources of observation are available for \ac{DEM} derivation such as \ac{InSAR} in Alaska, we focus here on 1 \ac{m} data derived from \ac{LiDAR} and use that synonymously with the term \ac{3DEP}.
+The \ac{NGP}, under the \ac{USGS}, is the primary authority on collecting, processing, and maintaining elevation data within the \ac{US} in collaboration with Federal partners \citep{omb2016circularA16}.
+Withinthe \ac{NGP}, the \ac{NED} \citep{gesch2002national,gesch2007digital}, originally released in 1997, forms the elevation layers of the \ac{TNM} \citep{gesch2009national,archuleta2017national,arundel2015preparing,arundel2018assimilation}.
 
-3DEP divided up into hydrography, topographic, 
-Base specifications from USGS are \citep{arundel20151} for 1m and the rest are \citep{archuleta2017national}.
-
-Lidar technical details \citep{stoker2015usgs}.
-
-NED evaluation \citep{gesch2014accuracy,dobbs2010evaluation}
-
-Compares Lidar data to NED for flooding using HEC-RAS \citep{wang2005comparison}.
+These seamless \acp{DEM} were generated via mosaicing and resampling of other non-seamless \acp{DEM} from a variety of elevation data sources including cartographic maps, \ac{LiDAR}, \ac{InSAR} \citep{gesch2002national}.
+The \ac{3DEP}, from the \ac{USGS}'s \ac{NGP}, extends \ac{TNM} - Elevation Program from the already existing 1/3, 1, and 2 arc-second spatial resolution seamless \acp{DEM} layers to include a 1 \ac{m}, \ac{LiDAR} derived \ac{DEM} product as well.
+While other sources of observation are available for \ac{DEM} derivation such as \ac{InSAR} in \ac{AK}, we focus here on 1 \ac{m} data derived from \ac{LiDAR} and use that synonymously with the term \ac{3DEP}.
+Previous continental scale \ac{FIM} efforts avoided use of \ac{3DEP} \ac{LiDAR} derived 1 \ac{m} \ac{DEM} products due to lack of spatial coverage and no seamless data availability.
+With the \ac{3DEP} target for full 1 \ac{m} coverage outside of \ac{AK}
 
 The 1 \ac{m} \ac{3DEP} product is a hydro-flattened, topographic bare-earth raster \ac{DEM} gridded to 1 \ac{km} square shaped tiles with 6 pixels of overlap \citep{arundel20151}.
+Hydro-flattening refers to a process in which hydrologic features such as lakes, reservoirs, streams, rivers, and more are flattened in elevation for bathymetric regions from bank to bank that are represented by breaklines \citep{archuleta2017national}.
+This flattening excludes along gradient directions, parallel to the direction of the breaklines, for hydrologic features that naturally exhibit water conveyance such as streams, rivers, and long reservoirs \citep{arundel20151}.
 The 1 \ac{m} \ac{DEM} is derived almost entirely from \ac{LiDAR} except in a few places where photogrammetry falling under the specifications of the \ac{LiDAR} data are used \citep{arundel20151}.
@@
 
-Elevation-derived hydrography specifications \citep{terziotti2020elevation}.
+%% make a table of different 3DEP data sources
+\begin{comment}
+\begin{table}[h!]
+ \begin{center}
+  \caption{\acf{3DEP} product specifications.}
+  \label{tab:table1}
+  \begin{tabular}{c|c|c|c|c|c} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
+   Program & Sources & Arc-Second Resolution & \ac{m} Resolution & Horizontal Accuracy (\ac{m}) & Vertical Accuracy (\ac{m})\\
+   \hline
+   \multirow{3}*{\ac{NED}} & 1/3 & 10 & varies by source &  \\
+   \hline
+   \ac{The National Map} & 10.1 & b &  &  \\
+   \hline
+   3 & 23.113231 & c &  &  \\
+   \hline
+  \end{tabular}
+ \end{center}
+\end{table}
 
-% Specifications
-- Lidar base specifications \citep{heidemann2012lidar}
-- national map seamless products \citep{archuleta2017national}
-- 1m specifications \citep{arundel20151}
+%%%% Specification Documents %%%%
+% - Lidar base specifications \citep{heidemann2018lidar}
+% - Lidar technical details \citep{stoker2015usgs}.
+% - national map seamless products \citep{archuleta2017national}
+% - 1m specifications \citep{arundel20151}
+% - NHDPlusHR Beta \citep{moore2019user}
+% - National Requirements for Enhanced Elevation Data \citep{snyder2013national}
+% - The 3D Elevation Program Initiative— A Call for Action \citep{sugarbaker20143d}
+% - Elevation-derived hydrography specifications \citep{terziotti2020elevation}.
 
-%%%% assessments on 3DEP data %%%%
-Assessments of 3DEP \citep{stoker2022accuracy,kim2020positional,chirico2020evaluating,callahan2022vertical,scott2022statewide,dobbs2010evaluation}
-Assessment of airborne Lidar on bathymetry data quality \citep{hilldale2008assessing}.
-Assessment of Lidar \citep{tarolli2014high}
+%%%% assessments on Elevation data %%%%
+% - Assessments of 3DEP \citep{stoker2022accuracy,kim2020positional,chirico2020evaluating,callahan2022vertical,scott2022statewide,dobbs2010evaluation}
+% - Assessment of airborne Lidar on bathymetry data quality \citep{hilldale2008assessing}.
+% - Assessment of Lidar \citep{tarolli2014high}
+% - NED evaluation \citep{gesch2014accuracy,dobbs2010evaluation}
+% - Compares Lidar data to NED for flooding using HEC-RAS \citep{wang2005comparison}.
 
@@
 
-Make sure to mention geoflood \citep{zheng2018geoflood,zheng2019automatic}.
 Make sure to mention the three limitations of elevation data include bathymetry, embankment delineations, and anthropogenic features such as bridges.
@@
 
+\end{comment}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-%% make a table of different 3DEP data sources
-\begin{table}[h!]
- \begin{center}
-  \caption{\acf{3DEP} product specifications.}
-  \label{tab:table1}
-  \begin{tabular}{c|c|c|c|c|c} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
-   Program & Product & Arc-Second Resolution & \ac{m} Resolution & Horizontal Accuracy (\ac{m}) & Vertical Accuracy (\ac{m})\\
-   \hline
-   1 & 1110.1 & a &  &  &  \\
-   \hline
-   2 & 10.1 & b &  &  &  \\
-   \hline
-   3 & 23.113231 & c &  &  &  \\
-   \hline
-  \end{tabular}
- \end{center}
-\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{\acs{Py3DEP}}
+\label{ssec:py3dep}
+
+Since the highest resolution \ac{3DEP} \ac{DEM} is at the 1/3 arc-second, we leverage a tool known as \ac{Py3DEP} to handle mosaicing and resampling of best available \acp{DEM} into a singular seamless one.
+This measure
 
@@
 \acro{SWE}{Shallow Water Equations}
-\acro{CONUS}{Continental United States}
+\acro{CONUS}{conterminous United States}
 \acro{PR}{Puerto Rico}
@@
 \acro{InSAR}{interferometric synthetic aperture radar}
+\acro{AK}{Alaska}
+\acro{NGP}{National Geospatial Program}
+\acro{TNM}{The National Map}
+\acro{Py3DEP}{Python 3DEP}{Python 3-Dimensional Elevation Program}
 \end{acronym}

### 2022-09-27 manuscripts/owp_fim4_2022/sections/abstract_plainTextSummary.tex
change to abstract

@@
 \begin{abstract}
-Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is a means able of producing high resolution flood inundation maps (FIMs) from the National Water Model (NWM) at large scales and high resolutions using reach-averaged synthetic rating curves.
+Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is a means able of producing flood inundation maps (FIMs) from the National Water Model (NWM) at large scales and high resolutions using reach-averaged synthetic rating curves.
 We highlight here that HAND is limited to producing inundation only when sourced from its nearest drainage line, thus lacks the ability to source inundation from multiple fluvial sources.

### 2022-10-04 manuscripts/owp_3dep_2022/owp_3dep.tex
added to intro

--- 2022-09-21 manuscripts/owp_3dep_2022/owp_3dep.tex
@@
 \usepackage[printonlyused]{acronym}
-%\usepackage{hyperref}
+\usepackage{multirow}
+\usepackage{comment}
+\usepackage{url}
 %
@@
 
-\title{Investigating the Effect of Varying the Spatial Resolutions of LiDAR Derived Elevations on Flood Inundation Maps Produced from Height Above Nearest Drainage}
+\title{On the Effects of Varying Spatial Resolutions of 3DEP Data on Flood Inundation Extents Produced from Height Above Nearest Drainage}
 
@@
 \begin{abstract}
-
 \end{abstract}
@@
 \begin{highlights}
-\item Continental-scale flood inundation mapping requires simplifying assumptions
-\item Utilizing LiDAR derived data from the USGS 3DEP program enhances the skill on flood inundation maps produced from \ac{HAND}.
+\item HAND approximates flood inundation for large scale, high resolution, and operational applications.
+\item Utilizing LiDAR derived data from the USGS 3DEP program enhances the skill on flood inundation maps produced from HAND.
 \item Varying the spatial resolution provides no skill enhancement when evaluated at large scales across three study sites and benchmark datasets.
@@
 \begin{keyword}
-Hydrology \sep
-Height Above Nearest Drainage \sep
-Flood Inundation Mapping \sep
-Digital Elevation Maps \sep
-Forecasting
-%
 %% PACS codes here, in the form: \PACS code \sep code
-\PACS 92.40.-t \sep 92.40.Qk \sep *92.40.Q- \sep *92.40.qp \sep 89.60.Gg \sep 92.60.Wc \sep 91.10.Jf \sep 84.40.Xb
+Hydrology, surface water, water resources, water supply, floods, lakes, rivers, streamflow, natural and man-made disasters, weather analysis and prediction, topography
+
+\PACS
+92.40.-t \sep % hydrology, cryosphere
+92.40.Qk \sep % surface water, water resources, water supply, floods, lakes, rivers, streamflow,
+*92.40.Q- \sep % surface water
+*92.40.qp \sep % streamflow, floods
+89.60.Gg \sep % disasters (natural and man-made),
+92.60.Wc \sep % weather analysis and prediction
+91.10.Jf      % topography (earth)
 %
@@
 \end{keyword}
-%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{frontmatter}
@@
 Operationally, the \ac{NWM} produces streamflow analysis and forecasts at multiple time horizons depending on location which include the \ac{CONUS}, \ac{PR}, and \ac{HI} \citep{cosgrove2019evolution,noaa2016national,water2022nwm}.
-The \ac{NWM} routes streamflow across the \ac{NWM} \ac{V2.1} stream network, based on the \ac{NHDPlusV2} network, is comprised of more than 5.5 million \acp{km} of lines discretized into more than 2.8 million forecast points \citep{aristizabal2022reducing}.
+The \ac{NWM} routes streamflow across the \ac{NWM} \ac{V2.1} stream network, based on the \ac{NHDPlusV2} network, is comprised of more than 5.5 million \acp{km} of lines discretized into more than 2.8 million forecast points \citep{aristizabal2022extending}.
 The \ac{NWM} \ac{V2.1} stream network belongs to the \ac{NWM} ``hydrofabric'' defined as a catalog of geospatial layers relevant to hydrology modeling including stream network lines, catchments, reservoirs, and more \citep{water2022nwm,cosgrove2019evolution}.
@@
 %%%% Introduces and motivates HAND %%%%
-The \acp{SWE}, a system of two hyperbolic partial differential equations, formally govern the flow of fluvial surface water by conserving both mass (first equation) and momentum (second equation) and can be expressed in both the \ac{1D} (Saint Venant Equations) and \ac{2D} forms.
+The \ac{SWE}, a system of two hyperbolic partial differential equations, formally govern the flow of fluvial surface water by conserving both mass (first equation) and momentum (second equation) and can be expressed in both the \ac{1D} (Saint Venant Equations) and \ac{2D} forms.
 Solving this system in full \ac{2D} form requires numerical methods that can be very cost prohibitive and numerically unstable in an operational setting across continental-scales at high spatial discretizations (10 \ac{m} or higher).
-This use case motivates the implementation of an inundation proxy, also known as a zero-physics or a simplified conceptual model, that is agnostic to the \acp{SWE} while still computing accurate fluvial inundation extents and depths for this problem \citep{teng2015rapid}.
+This use case motivates the implementation of an inundation proxy, also known as a zero-physics or a simplified conceptual model, that is agnostic to the \acp{SWE} while still computing accurate fluvial inundation extents and depths for this problem \citep{teng2015rapid,bates2000simple}.
 \ac{HAND} detrends elevations within \acp{DEM} to compute drainage potentials by normalizing elevations to the nearest, relevant drainage line instead of datums that represent mean sea level \citep{renno2008hand,nobre2011height,nobre2016hand}.
@@
 \ac{HAND} operates as an inundation proxy by thresholding the relative elevation (or \ac{HAND}) values with a singular river stage value for each catchment corresponding to the drainage area of a given river reach \citep{nobre2016hand,garousi2019terrain,johnson2019integrated,zheng2018geoflood,teng2015rapid,li2020evaluation,liu2016cybergis,maidment2017conceptual,liu2018cybergis,liu2020height,liu2018review}.
-When used to generate inundation extents and depths from streamflow, reach-averaged \ac{SRC} sample geometric variables along an entire reach and normalize using the length of the reach to create stage-discharge relationships \citep{zheng2018river,aristizabal2022reducing,godbout2019error}.
+When used to generate inundation extents and depths from streamflow, reach-averaged \ac{SRC} sample geometric variables along an entire reach and normalize using the length of the reach to create stage-discharge relationships \citep{zheng2018river,aristizabal2022extending,godbout2019error}.
 These relationships depend on a friction parameter, Manning's n, and used to convert streamflows to stages for eventual \ac{2D} mapping with \ac{HAND}.
-Numerous investigations have validated the use of \ac{HAND} for flood mapping applications as a suitable alternative to more sophisticated physics-based techniques for large scale and high resolution use cases \citep{johnson2019integrated,li2020evaluation,li2022comparative,aristizabal2022reducing,nobre2016hand,godbout2019error,afshari2018comparison,zhang2018comparative,teng2015rapid,teng2017flood,diehl2021improving}.
+Numerous investigations have validated the use of \ac{HAND} for flood mapping applications as a suitable alternative to more sophisticated physics-based techniques for large scale and high resolution use cases \citep{johnson2019integrated,li2020evaluation,li2022comparative,aristizabal2022extending,nobre2016hand,godbout2019error,afshari2018comparison,zhang2018comparative,teng2015rapid,teng2017flood,diehl2021improving,hocini2021performance,bates2003optimal}.
 
 %%%% Motivates and introduces OWP FIM %%%%
-Several prior and active large-scale \ac{HAND} implementations catered to operational \ac{EWS} applications including the \ac{NFIE} \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis}, GeoFlood \citep{zheng2018geoflood}, and \ac{PyGFT}, i.e.
+Several prior and active large-scale \ac{HAND} implementations catered to operational \ac{EWS} applications including the \ac{NFIE} \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis}, GeoFlood \citep{zheng2018geoflood,hocini2021performance,d2022identification,carruthers2021assessment,zheng2022application}, and \ac{PyGFT}, i.e.
 \ac{GIS} \citep{petrochenkov2020pygft,verdin2016software}.
-The \ac{NFIE} was a broad, inter-institutional effort to apply HAND to the initial versions of the \ac{NWM} which leveraged 10 \ac{m} (1/3 arc-second) elevation data available at the time \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis} from the \ac{USGS}'s \ac{NED} \citep{gesch2009national,gesch2014accuracy} which now has been subsumed \citep{arundel2018assimilation} to The National Map program \citep{gesch2009national,archuleta2017national}.
-\citet{zheng2018geoflood} applies HAND for operational applications to \ac{LiDAR} derived 1 \ac{m} elevation data with a novel least cost, geodesic based stream delineation method \citep{passalacqua2010geometric,passalacqua2012automatic,zheng2018geoflood}.
-For applications with the \ac{NWM}, an advanced version of \ac{HAND} coupled with the use of \acp{SRC}, known as \ac{OWP} \ac{FIM}, converts \ac{NWM} analysis, reanalysis, and forecast streamflows to river stages and fluvial inundation depths and extents on an operational basis to \ac{CONUS} while extending the modeling domain to \ac{PR} and \ac{HI} \citep{aristizabal2022reducing,inundationMapping2022}.
-\ac{OWP} \ac{FIM} utilizes some of the latest datasets including the \ac{NHDPlusHR}, \ac{NLD}, and \ac{NWM} \ac{V2.1} hydrofabric to enforce drainage, hydrologically relevant features such as levees, and conflation with the forecast stream network \citep{aristizabal2022reducing,inundationMapping2022}.
-Additionally, \ac{OWP} \ac{FIM} advanced a fundamental limitation of \ac{HAND} that limits sourcing fluvial inundation only from the nearest, relevant drainage line \citep{mcgehee2016modified,aristizabal2022reducing,zhang2018comparative,li2022comparative,zheng2018geoflood,zheng2018river,nobre2016hand}.
-Stream lines of higher Horton-Strahler stream order that could contribute inundation to a given area have no way of extending beyond catchment lines which creates artificial bottlenecks in inundation extents especially along junctions of high order rivers with their lower flow tributaries \citep{aristizabal2022reducing,mcgehee2016modified}.
-To resolve this limitation, \ac{OWP} \ac{FIM} disaggregates the \ac{NWM} \ac{V2.1} stream network into segments of effective unit stream order called \acp{LP} \citep{aristizabal2022reducing}.
-In terms of terrain data, \ac{OWP} \ac{FIM} utilizes the 10 \ac{m} \ac{DEM} from the \ac{NHDPlusHR} dataset which is based on the \ac{NED} due to its high fidelity and continental availability \citep{aristizabal2022reducing,moore2019user}.
-The previous advances in \ac{OWP} \ac{FIM} stopped short of accounting for new \ac{LiDAR} datasets that are now approaching continental scale availability from the \ac{USGS} \citep{aristizabal2022reducing}.
+The \ac{NFIE} was a broad, inter-institutional, and pioneering effort to apply HAND to the initial versions of the \ac{NWM} which leveraged 1/3 arc-second (10 \ac{m}) seamless elevation data available at the time \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis} from the \ac{USGS}'s \ac{NED} \citep{gesch2002national,gesch2007digital}.
+\citet{zheng2018geoflood} applied HAND for operational applications with 1/27 arc-second (1 \ac{m}) elevation data with a novel least cost, geodesic based stream delineation method \citep{passalacqua2010geometric,passalacqua2012automatic,zheng2018geoflood,zheng2019automatic,carruthers2021assessment,d2022identification,zheng2022application}.
+For applications with the \ac{NWM}, an advanced version of \ac{HAND} coupled with the use of \acp{SRC}, known as \ac{OWP} \ac{FIM}, converts \ac{NWM} analysis, reanalysis, and forecast streamflows to river stages and fluvial inundation depths and extents on an operational basis to \ac{CONUS} while extending the modeling domain to \ac{PR} and \ac{HI} \citep{aristizabal2022extending,inundationMapping2022}.
+\ac{OWP} \ac{FIM} utilizes some of the latest datasets including the \ac{NHDPlusHR} \citep{moore2019user}, \ac{NLD} \citep{engineers2016national}, and the \ac{NWM} \ac{V2.1} hydrofabric \citep{water2022nwm,noaa2016national,nwm2022hydrofabric,gochis2021wrf}.
+These datasets enforce hydrologically relevant features such as levees and the general location of stream lines to facilitate conflation with the forecast stream network \citep{aristizabal2022extending,inundationMapping2022}.
+Additionally, \ac{OWP} \ac{FIM} advanced a fundamental limitation of \ac{HAND} that limits sourcing fluvial inundation only from the nearest, relevant drainage line \citep{mcgehee2016modified,aristizabal2022extending,zhang2018comparative,li2022comparative,zheng2018geoflood,zheng2018river,nobre2016hand}.
+Stream lines of higher Horton-Strahler stream order that could contribute inundation to a given area have no way of extending beyond catchment lines which creates artificial bottlenecks in inundation extents especially along junctions of high order rivers with their lower flow tributaries \citep{aristizabal2022extending,mcgehee2016modified}.
+To resolve this limitation, \ac{OWP} \ac{FIM} disaggregates the \ac{NWM} \ac{V2.1} stream network into segments of effective unit stream order called \acp{LP} \citep{aristizabal2022extending}.
+In terms of terrain data, \ac{OWP} \ac{FIM} utilizes the 10 \ac{m} \ac{DEM} from the \ac{NHDPlusHR} elevation dataset which is the elevation basis, derived in batches from \ac{3DEP}, for additional hydrography products within the \ac{NHDPlusHR} \citep{aristizabal2022extending,moore2019user}.
+The previous advances in \ac{OWP} \ac{FIM} stopped short of accounting for \ac{LiDAR} point elevation observations \citep{aristizabal2022extending} that are now nearing their first collection cycle to form a novel seamless, continental scale \ac{DEM} from \ac{3DEP} \citep{usgs2022status,usgs2022partnerships}.
 
-%%%% 3DEP program introduction with LiDAR intro %%%%
-The \ac{USGS}'s \ac{3DEP} extends The National Map Program from the already existing 1/3, 1, and 2 arc-second spatial resolution seamless \acp{DEM} products to include a 1 \ac{m}, \ac{LiDAR} derived \ac{DEM} product as well.
-While other sources of observation are available for \ac{DEM} derivation such as \ac{InSAR} in Alaska, we focus here on 1 \ac{m} data derived from \ac{LiDAR} and use that synonymously with the term \ac{3DEP}.
+%%%% 3DEP program introduction %%%%
+Broad scale terrain information in the form of \acp{DEM} is fundamental to all \ac{FIM} models and a significant influence on skill \citep{bales2009sources,dobbs2010evaluation,wang2005comparison,merwade2008uncertainty,witt2015evaluation}.
+The \ac{NGP}, under the \ac{USGS}, is the primary authority on collecting, processing, and maintaining terrestrial elevation data within the \ac{US} in collaboration with Federal partners within the \ac{NDEP} \citep{omb2016circularA16,dewberry2011final,national2007elevation,national2009mapping,sugarbaker20143d}.
+The \ac{NED} \citep{gesch2002national,gesch2007digital}, forms the seamless elevation layers of the \ac{TNM} \citep{gesch2009national,archuleta2017national,arundel2015preparing,arundel2018assimilation,kelmelis2003national}.
+Prior to the introduction of \ac{3DEP}, \ac{TNM} was originally composed of three seamless \acp{DEM} at 1/3 (10 \ac{m}), 1 (30 \ac{m}), and 2 (90 \ac{m}) arc-second resolutions produced from a variety legacy sources including digital photogrammetry, cartographic contours, mapped hydrography, and elevations from \ac{SRTM} .
+\citep{gesch2002national,gesch2007digital,arundel2015preparing}.
+High quality elevations derived from \ac{LiDAR} and \ac{InSAR} have been integrated into \ac{TNM} seamless elevation products as made available prior to and after the introduction of \ac{3DEP} \citep{snyder2013national,gesch2002national,arundel2015preparing}.
+Work by \citet{gesch2014accuracy},\citep{gesch2007digital}, and \citet{dobbs2010evaluation} illustrated that the inclusion of higher quality elevation data sources had significant improvement in the accuracy of \ac{NED} data when compared to the \ac{NGS} \citep{roman2010geodesy}.
+\citet{gesch2014accuracy} identified that the \ac{NED} 1/3 arc-second \ac{DEM}, as of April 2013, had a mean error of -0.29 \ac{m} with an \ac{RMSE} of 1.55 \ac{m} when compared to over 25 thousand reference points.
+At the time of evaluation, the \ac{NED} was subject to legacy, lower quality data sources dating almost a century in the past \citep{sugarbaker20143d,gesch2014accuracy,gesch2007digital}.
+This reduction in error and its impact on people and commerce \citep{dewberry2011final} motivated action on collection of elevation data from higher quality data sources \citep{sugarbaker20143d}.
 
-3DEP divided up into hydrography, topographic, 
-Base specifications from USGS are \citep{arundel20151} for 1m and the rest are \citep{archuleta2017national}.
+%%% 3DEP AND LIDAR %%%%%
+%%% NEED LIDAR REFERENCES %%%%
+\ac{3DEP} is a national, multi-organizational effort by the \ac{NDEP} to survey elevations with high quality sensors in response to growing stakeholder needs on a recurring collection cycle of no greater than 8 \acp{year} \citep{dewberry2011final,snyder2013national,sugarbaker20143d}.
+\ac{3DEP} leverages two main collection technologies including \ac{LiDAR} for the \ac{CONUS}, \ac{HI}, and \ac{US} territories as well as \ac{InSAR} for \ac{AK}.
+\ac{LiDAR}, the collection source of focus in this study, is a light reflection technology that beams concentrated powerful light and receives the return while recording the travel time and intensity of return.
+\ac{LiDAR} sensors are mounted on top of a variety of mobile or static platforms whose positions are geo-tracked as they collect \ac{LiDAR} returns.
+The travel time of the returns, along with knowledge of the speed of light, serve as a relative positioning of the target(s) referenced to a common vertical datum while the intensities serve as indicators of what the target(s) represent.
+Modes within the relationship of return intensities with respect to travel time/distance from the \ac{LiDAR} wave forms can be indicative of vegetation or other \acp{LULC} that reflect signals at varying distances and magnitudes and influence elevation errors \citep{gesch2014accuracy}.
+These modes can be discretized into varying \ac{DEM} products representing bare earth, structures, or canopy elevations.
+The horizontal and vertical accuracies and the horizontal resolutions of terrain observations derived from \ac{LiDAR}, and even the consequential economic benefits \citep{dewberry2011final,dewberry2022nation}, are dependent on a variety of sensor, platform, target, and collection specifications and practices such as nominal pulse spacing, nominal pulse density, and \ac{LULC} of the target \citep{heidemann2018lidar,passalacqua2015analyzing,smith2019determining,salach2018accuracy,gesch2014accuracy}.
+\ac{LiDAR} produces point cloud datasets which are scattered, geo-referenced points representing full wave forms or discretized return intensities.
+Various assessments of the vertical accuracries of \ac{LiDAR} point clouds have yielded satisfactory results in agreement with \ac{3DEP} requirements \citep{stoker2022accuracy,kim2022absolute,callahan2022vertical,kim2022absolute,salach2018accuracy,passalacqua2015analyzing}.
+Point clouds must undergo a series of operations to produce analysis ready, seamless \acp{DEM} \citep{passalacqua2015analyzing}.
 
-Lidar technical details \citep{stoker2015usgs}.
-
-NED evaluation \citep{gesch2014accuracy,dobbs2010evaluation}
-
-Compares Lidar data to NED for flooding using HEC-RAS \citep{wang2005comparison}.
-
-The 1 \ac{m} \ac{3DEP} product is a hydro-flattened, topographic bare-earth raster \ac{DEM} gridded to 1 \ac{km} square shaped tiles with 6 pixels of overlap \citep{arundel20151}.
-The 1 \ac{m} \ac{DEM} is derived almost entirely from \ac{LiDAR} except in a few places where photogrammetry falling under the specifications of the \ac{LiDAR} data are used \citep{arundel20151}.
-According to specifications, the horizontal accuracy of 1 \ac{m} \ac{3DEP} is expected to be within 1 \ac{m} while the vertical accuracies are within 19.6 \ac{cm} and 30 \ac{cm} at the 95\% confidence interval for non-vegetative and vegetative regions, respectively \citep{arundel20151}.
-Non-vegetative vertical accuracies fall within 10 \ac{cm} \ac{RSME} \ac{arundel20151}.
-
-Elevation-derived hydrography specifications \citep{terziotti2020elevation}.
-
-% Specifications
-- Lidar base specifications \citep{heidemann2012lidar}
-- national map seamless products \citep{archuleta2017national}
-- 1m specifications \citep{arundel20151}
-
-%%%% assessments on 3DEP data %%%%
-Assessments of 3DEP \citep{stoker2022accuracy,kim2020positional,chirico2020evaluating,callahan2022vertical,scott2022statewide,dobbs2010evaluation}
-Assessment of airborne Lidar on bathymetry data quality \citep{hilldale2008assessing}.
-Assessment of Lidar \citep{tarolli2014high}
+%%%% 3DEP DEMs and Accuracies %%%%
+\ac{3DEP} extends \ac{TNM} to include a 1/27 arc-second (1 \ac{m}), \ac{LiDAR} derived \ac{DEM} product for \ac{CONUS}, \ac{HI}, and \ac{US} territories as well as a 1/2 arc-second (5 \ac{m}) \ac{DEM} for \ac{AK} derived from \ac{InSAR} \citep{sugarbaker20143d,stoker2015usgs}.
+To create bare earth \acp{DEM}, \ac{LiDAR} observations must undergo a series of processes that filter out returns from vegetation, anthropogenic, and other features then grid the observations with resampling methods \citep{passalacqua2015analyzing}.
+The 1 \ac{m} \ac{3DEP} product is a \ac{hydro-flattened}, topographic, and bare-earth raster \ac{DEM} gridded to 1 \ac{km} square shaped tiles with 6 pixels of overlap \citep{arundel20151}.
+Hydro-flattening refers to a process in which hydrologic features such as lakes, reservoirs, streams, rivers, and more are flattened in elevation for bathymetric regions from lower bank to lower bank represented by breaklines \citep{archuleta2017national,maune2018digital}.
+This flattening excludes along gradient directions, parallel to the direction of the breaklines, for hydrologic features that naturally exhibit water conveyance such as streams, rivers, and long reservoirs \citep{arundel20151}.
+This process includes elevations underneath bridges that are not accurately observed from topographic \ac{LiDAR}.
+According to specifications, the horizontal accuracy of 1 \ac{m} \ac{3DEP} is expected to be within 1 \ac{m} while the vertical accuracies are within 19.6 \ac{cm} and 30 \ac{cm} at the 95\% confidence interval for non-vegetative and vegetative regions, respectively \citep{arundel20151,heidemann2018lidar}.
+Non-vegetative vertical accuracies fall within 10 \ac{cm} \ac{RMSE} \citep{arundel20151,heidemann2018lidar}.
+ Work by \citet{stoker2022accuracy,callahan2022vertical,kim2022absolute} have verified the vertical accuracies and general quality of the \acp{DEM} for \ac{3DEP} specifications.
 
 %%%% FIM studies on importance of Lidar and elevation data %%%%
-Numerous reports and academic studies discuss the principal effect that topographic data quality has on the quality of \ac{FIM}.
-The National Enhanced Elevation Assessment conducted by \citet{dewberry2011final} quantified the estimated annual benefits of high resolution, enhanced elevation data by 27 business sectors and concluded flood risk management to have the highest conservative estimate among the sectors in 2015 \ac{USD}.
-Several technical reports document that the quality of the elevation map used is of primal significance to the quality of the flood maps produced \citep{national2007elevation,national2009mapping}.
+As previously stated, the features of topographic information including its source, accuracies, resolutions, and processing methods are of primary importance to the quality of \acp{FIM} \citep{national2007elevation,national2009mapping,carswell20183d,bales2009sources}.
+Uncertainties in elevations and the processes used to determine them are propogated into \ac{FIM} extents.
+Uncertainty in FIM \citep{merwade2008uncertainty} including data sources, gridding, resampling, resolution, mosaicing, hydrologic conditioning, and more.
 
-Make sure to mention geoflood \citep{zheng2018geoflood,zheng2019automatic}.
+\citep{tarolli2014high} good review of channel extraction and other issues surrounding elevations and resolutions.
+
+Flood specific assessments for 3DEP \citep{carswell20183d,bales2009sources,gesch2018best,podhoranyi2015inaccuracy,lamichhane2018effect,tsubaki2013uncertainty,dobbs2010evaluation,arrighi2019effects,zazo2015analysis,bhuyian2018accounting,gesch2012elevation,witt2015evaluation}.
+
 Make sure to mention the three limitations of elevation data include bathymetry, embankment delineations, and anthropogenic features such as bridges.
@@
 DEMs and flooding \citep{casas2006topographic,thomas2016quantifying,savage2016does,passalacqua2010geometric,passalacqua2012automatic,munoth2019effects}
+ - Compares Lidar data to NED for flooding using HEC-RAS \citep{wang2005comparison}.
 
@@
 
+Concluded 10m DEM was good enough for hydro applications \citep{zhang1994digital}
+
+more FIM and elevations \citep{werner2001impact,omer2003impact,bates2003optimal,tate2002creating,colby2000modeling}.
+
 Changed the horizontal resolution of DEM for HAND \citep{li2022accounting}, showed that DEM resolution was the most important factor influencing FIM extent skill when varied alongside water depth and drainage threshold.
-
-Flood specific assessments for 3DEP \citep{carswell20183d,bales2009sources,gesch2018best,podhoranyi2015inaccuracy,lamichhane2018effect,tsubaki2013uncertainty,dobbs2010evaluation,arrighi2019effects,zazo2015analysis,bhuyian2018accounting,gesch2012elevation}.
-
 \citep{garousi2019terrain} using 3m vs 10m with HAND.
@@
 %%%% Motivation paragraph for this study %%%%
-With these latest developments in terrain data availability, we propose investigating the integration of \ac{3DEP} data into \ac{OWP} \ac{FIM} for continental-scale inundation forecasting abilities.
+Previous efforts with \ac{OWP} \ac{FIM} avoided use of 1 \ac{m} \ac{3DEP} \ac{DEM} products due to lack of spatial coverage and no seamless data availability.
+As the 1 \ac{m} \ac{3DEP} product rapidly approaches full \ac{CONUS} coverage in 2023, we propose investigating the integration of \ac{3DEP} data into \ac{OWP} \ac{FIM} for continental-scale inundation forecasting abilities \citep{usgs2022status,usgs2022partnerships}.
 We will investigate utilizing \ac{3DEP} data for \ac{HAND} computation to generate the \ac{FIM} hydrofabric.
 Additionally, we investigate the utility of varying spatial resolutions from 1, 3, 5, 10, 15, and 20 \ac{m}.
-\ac{HAND} depends on the drainage assumptions which requires \acp{DEM} to undergo a long series of enforcement processes to ensure monotonically decreasing elevations with hydrologically correct flow directions \citep{garousi2019terrain,nobre2011height,nobre2016hand,aristizabal2022reducing}.
-The resampling of \acp{DEM} into varying spatial resolutions could interact with these hydro-conditioning operations thus influencing the \ac{FIM} hydrofabric and the resulting quality of the \acp{FIM} produced.
-We seek to quantify this relationship and assess how these interact with one another to optimize for the quality of continental-scale \acp{FIM} from the \ac{NWM}.
+\ac{HAND} depends on the drainage assumptions which requires \acp{DEM} to undergo a long series of enforcement processes to ensure monotonically decreasing elevations with hydrologically correct flow directions \citep{garousi2019terrain,nobre2011height,nobre2016hand,aristizabal2022extending}.
+The resampling of \acp{DEM} into varying spatial resolutions could interact with these \ac{hydro-conditioning} operations thus influencing the \ac{FIM} hydrofabric and the resulting quality of the \acp{FIM} produced.
+As validation, we utilize \ac{1D} \ac{HEC-RAS} modeled flood inundation extents from both the \ac{BLE} published by \ac{InFRM} and from a novel package that automates the generation of \ac{FIM} from existing \ac{1D} \ac{HEC-RAS} models.
+As a third source of validation, we utilize proprietary \ac{FIM} derived from remote sensing earth observation \ac{SAR} furnished by ICEYE.
+By varying the spatial resolution of \ac{3DEP} \acp{DEM}, we seek to understand the relationship between spatial resolution and \ac{FIM} skill produced from \ac{HAND} that requires significant \ac{DEM} manipulations to satisfy inherent assumptions.
 
@@
 \ac{HAND} assumes monotonically decreasing elevations with hydrologically relevant flow paths ensuring all cells in a region drain \citep{renno2008hand,nobre2011height,nobre2016hand}.
-This requires the use of extensive hydro-conditioning techniques on \acp{DEM} to enforce drainage and agreement with existing hydrography \citep{aristizabal2022reducing,maidment2017conceptual,liu2016cybergis,liu2020height}.
+This requires the use of extensive \ac{hydro-conditioning} techniques on \acp{DEM} to enforce drainage and agreement with existing hydrography \citep{aristizabal2022extending,maidment2017conceptual,liu2016cybergis,liu2020height}.
 
@@
 %
+
+\begin{comment}
 %% make a table of different 3DEP data sources
+Table \ref{tab:tnm_layer_specifications} illustrates the various elevation layers of \ac{TNM} along with their respective specifications and sources.
 \begin{table}[h!]
  \begin{center}
-  \caption{\acf{3DEP} product specifications.}
-  \label{tab:table1}
+  \caption{\acf{TNM} - Elevation layers specifications. These layers are also known as the \acf{NED}}
+  \label{tab:tnm_layer_specifications}
   \begin{tabular}{c|c|c|c|c|c} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
-   Program & Product & Arc-Second Resolution & \ac{m} Resolution & Horizontal Accuracy (\ac{m}) & Vertical Accuracy (\ac{m})\\
+   \multirow{2}*{Layer} & \multicolumn{2}{|c|}{Spatial Resolution} & \multicolumn{2}{|c|}{Accuracy (\ac{m})} & \multirow{2}*{Sources} \\
+    & arc-second & (\ac{m}) & Horizontal & Vertical &  \\
    \hline
-   1 & 1110.1 & a &  &  &  \\
-   \hline
-   2 & 10.1 & b &  &  &  \\
-   \hline
-   3 & 23.113231 & c &  &  &  \\
+   1/3 arc-second seamless & 1/3 & 10 & varies & varies & Many  \\
    \hline
@@
 \end{table}
+\end{comment}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{\ac{3DEP} Data Acquisition and Preparation}
+\label{ssec:py3dep}
+%
+Since the highest resolution \ac{3DEP} \ac{DEM} is at the 1/3 arc-second, we leverage a tool known as \ac{Py3DEP} to handle mosaicing and resampling of best available \acp{DEM} into a singular seamless one.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evaluation}
+\label{ssec:evaluation}
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{\ac{BLE}}
+\label{sssec:ble}
+%
 
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{\ac{RAS2FIM}}
+\label{sssec:ras2fim}
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{\ac{InSAR}}
+\label{sssec:insar}
+%
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
+Future work could employ the use of STAC catalogs to acquire and process data from the cloud from Open Topography (\url{https://stacindex.org/catalogs/opentopography#/})
+Tangentially, we assert the gridding of source elevations from \ac{InSAR} and \ac{LiDAR} point clouds to be out of scope for this analysis elected instead to deal with already gridded elevations.
+Future work can further understand how \ac{HAND} is affect by gridding techniques most specifically that of point cloud conversion.
 
@@
 \acro{BLE}{Base Level Engineering}
+\acro{InSAR}{Interferometric Synthetic Aperture Radar}
+\acro{RAS2FIM}{River Analysis System-2-Flood Inundation Map}
 \acro{FEMA}{Federal Emergency Management Agency}
+\acro{InFRM}{Interagency Flood Risk Management}
 \acro{HAND}{Height Above Nearest Drainage}
@@
 \acro{InSAR}{interferometric synthetic aperture radar}
+\acro{AK}{Alaska}
+\acro{NGP}{National Geospatial Program}
+\acro{TNM}{The National Map}
+\acro{Py3DEP}[Python 3DEP]{Python 3-Dimensional Elevation Program}
+\acro{RMSE}{root mean squared error}
+\acro{SRTM}{Shuttle Radar Topography Mission}
+\acro{HEC-RAS}{Hydrologic Engineering Center River Analysis Center}
+\acro{NGS}{National Geodetic Survey}
+\acro{NDEP}{National Digital Elevation Program}
+\acro{year}{yr}
+\acro{hydro-conditioning}{hydrological conditioning}
+\acro{hydro-conditioned}{hydrologically conditioned}
+\acro{hydro-flattening}{hydrological flattening}
+\acro{hydro-flattened}{hydrologically conditioned}
+\acro{LULC}{land use/land cover}
 \end{acronym}

### 2022-10-04 manuscripts/owp_3dep_2022/dependencies/Definitions/journalnames.tex
added mdpi template

%%Updated on 28 July 2022
%,% Website Name,%,%,% Website Name,%%,% Full Name,%%,% Short Name,%%,% DOI abbr,%%,% ISSN,% \continuouspages
%\DeclareOption{}{\gdef\@journal{}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{acoustics}{\gdef\@journal{acoustics}\gdef\@journalfull{Acoustics}\gdef\@journalshort{Acoustics}\gdef\@doiabbr{acoustics}\gdef\@ISSN{2624-599X}\gdef\@continuouspages{yes}}
\DeclareOption{actuators}{\gdef\@journal{actuators}\gdef\@journalfull{Actuators}\gdef\@journalshort{Actuators}\gdef\@doiabbr{act}\gdef\@ISSN{2076-0825}}
\DeclareOption{addictions}{\gdef\@journal{addictions}\gdef\@journalfull{Addictions}\gdef\@journalshort{Addictions}\gdef\@doiabbr{}\gdef\@ISSN{0006-0006}}
\DeclareOption{admsci}{\gdef\@journal{admsci}\gdef\@journalfull{Administrative Sciences}\gdef\@journalshort{Adm. Sci.}\gdef\@doiabbr{admsci}\gdef\@ISSN{2076-3387}}
\DeclareOption{adolescents}{\gdef\@journal{adolescents}\gdef\@journalfull{Adolescents}\gdef\@journalshort{Adolescents}\gdef\@doiabbr{adolescents}\gdef\@ISSN{2673-7051}\gdef\@continuouspages{yes}}
\DeclareOption{aerospace}{\gdef\@journal{aerospace}\gdef\@journalfull{Aerospace}\gdef\@journalshort{Aerospace}\gdef\@doiabbr{aerospace}\gdef\@ISSN{2226-4310}}
\DeclareOption{agriculture}{\gdef\@journal{agriculture}\gdef\@journalfull{Agriculture}\gdef\@journalshort{Agriculture}\gdef\@doiabbr{agriculture}\gdef\@ISSN{2077-0472}}
\DeclareOption{agriengineering}{\gdef\@journal{agriengineering}\gdef\@journalfull{AgriEngineering}\gdef\@journalshort{AgriEngineering}\gdef\@doiabbr{agriengineering}\gdef\@ISSN{2624-7402}\gdef\@continuouspages{yes}}
\DeclareOption{agrochemicals}{\gdef\@journal{agrochemicals}\gdef\@journalfull{Agrochemicals}\gdef\@journalshort{Agrochemicals}\gdef\@doiabbr{agrochemicals}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{agronomy}{\gdef\@journal{agronomy}\gdef\@journalfull{Agronomy}\gdef\@journalshort{Agronomy}\gdef\@doiabbr{agronomy}\gdef\@ISSN{2073-4395}}
\DeclareOption{ai}{\gdef\@journal{ai}\gdef\@journalfull{AI}\gdef\@journalshort{AI}\gdef\@doiabbr{ai}\gdef\@ISSN{2673-2688}\gdef\@continuouspages{yes}}
\DeclareOption{air}{\gdef\@journal{air}\gdef\@journalfull{Air}\gdef\@journalshort{Air}\gdef\@doiabbr{air}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{algorithms}{\gdef\@journal{algorithms}\gdef\@journalfull{Algorithms}\gdef\@journalshort{Algorithms}\gdef\@doiabbr{a}\gdef\@ISSN{1999-4893}}
\DeclareOption{allergies}{\gdef\@journal{allergies}\gdef\@journalfull{Allergies}\gdef\@journalshort{Allergies}\gdef\@doiabbr{allergies}\gdef\@ISSN{2313-5786}}
\DeclareOption{alloys}{\gdef\@journal{alloys}\gdef\@journalfull{Alloys}\gdef\@journalshort{Alloys}\gdef\@doiabbr{alloys}\gdef\@ISSN{2674-063X}\gdef\@continuouspages{yes}}
\DeclareOption{analytica}{\gdef\@journal{analytica}\gdef\@journalfull{Analytica}\gdef\@journalshort{Analytica}\gdef\@doiabbr{analytica}\gdef\@ISSN{2673-4532}\gdef\@continuouspages{yes}}
\DeclareOption{analytics}{\gdef\@journal{analytics}\gdef\@journalfull{Analytics}\gdef\@journalshort{Analytics}\gdef\@doiabbr{analytics}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{anatomia}{\gdef\@journal{anatomia}\gdef\@journalfull{Anatomia}\gdef\@journalshort{Anatomia}\gdef\@doiabbr{anatomia}\gdef\@ISSN{2813-0545}\gdef\@continuouspages{yes}}
\DeclareOption{animals}{\gdef\@journal{animals}\gdef\@journalfull{Animals}\gdef\@journalshort{Animals}\gdef\@doiabbr{ani}\gdef\@ISSN{2076-2615}}
\DeclareOption{antibiotics}{\gdef\@journal{antibiotics}\gdef\@journalfull{Antibiotics}\gdef\@journalshort{Antibiotics}\gdef\@doiabbr{antibiotics}\gdef\@ISSN{2079-6382}}
\DeclareOption{antibodies}{\gdef\@journal{antibodies}\gdef\@journalfull{Antibodies}\gdef\@journalshort{Antibodies}\gdef\@doiabbr{antib}\gdef\@ISSN{2073-4468}}
\DeclareOption{antioxidants}{\gdef\@journal{antioxidants}\gdef\@journalfull{Antioxidants}\gdef\@journalshort{Antioxidants}\gdef\@doiabbr{antiox}\gdef\@ISSN{2076-3921}}
\DeclareOption{applbiosci}{\gdef\@journal{applbiosci}\gdef\@journalfull{Applied Biosciences}\gdef\@journalshort{Appl. Biosci.}\gdef\@doiabbr{applbiosci}\gdef\@ISSN{2813-0464}\gdef\@continuouspages{yes}}
\DeclareOption{appliedchem}{\gdef\@journal{appliedchem}\gdef\@journalfull{AppliedChem}\gdef\@journalshort{AppliedChem}\gdef\@doiabbr{appliedchem}\gdef\@ISSN{2673-9623}\gdef\@continuouspages{yes}}
\DeclareOption{appliedmath}{\gdef\@journal{appliedmath}\gdef\@journalfull{AppliedMath}\gdef\@journalshort{AppliedMath}\gdef\@doiabbr{appliedmath}\gdef\@ISSN{2673-9909}\gdef\@continuouspages{yes}}
\DeclareOption{applmech}{\gdef\@journal{applmech}\gdef\@journalfull{Applied Mechanics}\gdef\@journalshort{Appl. Mech.}\gdef\@doiabbr{applmech}\gdef\@ISSN{2673-3161}\gdef\@continuouspages{yes}}
\DeclareOption{applmicrobiol}{\gdef\@journal{applmicrobiol}\gdef\@journalfull{Applied Microbiology}\gdef\@journalshort{Appl. Microbiol.}\gdef\@doiabbr{applmicrobiol}\gdef\@ISSN{2673-8007}\gdef\@continuouspages{yes}}
\DeclareOption{applnano}{\gdef\@journal{applnano}\gdef\@journalfull{Applied Nano}\gdef\@journalshort{Appl. Nano}\gdef\@doiabbr{applnano}\gdef\@ISSN{2673-3501}\gdef\@continuouspages{yes}}
\DeclareOption{applsci}{\gdef\@journal{applsci}\gdef\@journalfull{Applied Sciences}\gdef\@journalshort{Appl. Sci.}\gdef\@doiabbr{app}\gdef\@ISSN{2076-3417}}
\DeclareOption{aquacj}{\gdef\@journal{aquacj}\gdef\@journalfull{Aquaculture Journal}\gdef\@journalshort{Aquac. J.}\gdef\@doiabbr{aj}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{architecture}{\gdef\@journal{architecture}\gdef\@journalfull{Architecture}\gdef\@journalshort{Architecture}\gdef\@doiabbr{architecture}\gdef\@ISSN{2673-8945}\gdef\@continuouspages{yes}}
\DeclareOption{arm}{\gdef\@journal{arm}\gdef\@journalfull{Advances in Respiratory Medicine}\gdef\@journalshort{Adv. Respir. Med.}\gdef\@doiabbr{arm}\gdef\@ISSN{2543-6031}\gdef\@continuouspages{yes}}
\DeclareOption{arthropoda}{\gdef\@journal{arthropoda}\gdef\@journalfull{Arthropoda}\gdef\@journalshort{Arthropoda}\gdef\@doiabbr{arthropoda}\gdef\@ISSN{2813-3323}\gdef\@continuouspages{yes}}
\DeclareOption{arts}{\gdef\@journal{arts}\gdef\@journalfull{Arts}\gdef\@journalshort{Arts}\gdef\@doiabbr{arts}\gdef\@ISSN{2076-0752}}
\DeclareOption{asc}{\gdef\@journal{asc}\gdef\@journalfull{Automatic Systems and Control}\gdef\@journalshort{Autom. Syst. Control}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{asi}{\gdef\@journal{asi}\gdef\@journalfull{Applied System Innovation}\gdef\@journalshort{Appl. Syst. Innov.}\gdef\@doiabbr{asi}\gdef\@ISSN{2571-5577}}
\DeclareOption{astronomy}{\gdef\@journal{astronomy}\gdef\@journalfull{Astronomy}\gdef\@journalshort{Astronomy}\gdef\@doiabbr{astronomy}\gdef\@ISSN{2674-0346}\gdef\@continuouspages{yes}}
\DeclareOption{atmosphere}{\gdef\@journal{atmosphere}\gdef\@journalfull{Atmosphere}\gdef\@journalshort{Atmosphere}\gdef\@doiabbr{atmos}\gdef\@ISSN{2073-4433}}
\DeclareOption{atoms}{\gdef\@journal{atoms}\gdef\@journalfull{Atoms}\gdef\@journalshort{Atoms}\gdef\@doiabbr{atoms}\gdef\@ISSN{2218-2004}}
\DeclareOption{audiolres}{\gdef\@journal{audiolres}\gdef\@journalfull{Audiology Research}\gdef\@journalshort{Audiol. Res.}\gdef\@doiabbr{audiolres}\gdef\@ISSN{2039-4349}\gdef\@continuouspages{yes}}
\DeclareOption{automation}{\gdef\@journal{automation}\gdef\@journalfull{Automation}\gdef\@journalshort{Automation}\gdef\@doiabbr{automation}\gdef\@ISSN{2673-4052}\gdef\@continuouspages{yes}}
\DeclareOption{axioms}{\gdef\@journal{axioms}\gdef\@journalfull{Axioms}\gdef\@journalshort{Axioms}\gdef\@doiabbr{axioms}\gdef\@ISSN{2075-1680}}
\DeclareOption{bacteria}{\gdef\@journal{bacteria}\gdef\@journalfull{Bacteria}\gdef\@journalshort{Bacteria}\gdef\@doiabbr{bacteria}\gdef\@ISSN{2674-1334}\gdef\@continuouspages{yes}}
\DeclareOption{batteries}{\gdef\@journal{batteries}\gdef\@journalfull{Batteries}\gdef\@journalshort{Batteries}\gdef\@doiabbr{batteries}\gdef\@ISSN{2313-0105}}
\DeclareOption{bdcc}{\gdef\@journal{bdcc}\gdef\@journalfull{Big Data and Cognitive Computing}\gdef\@journalshort{Big Data Cogn. Comput.}\gdef\@doiabbr{bdcc}\gdef\@ISSN{2504-2289}}
\DeclareOption{behavsci}{\gdef\@journal{behavsci}\gdef\@journalfull{Behavioral Sciences}\gdef\@journalshort{Behav. Sci.}\gdef\@doiabbr{bs}\gdef\@ISSN{2076-328X}}
\DeclareOption{beverages}{\gdef\@journal{beverages}\gdef\@journalfull{Beverages}\gdef\@journalshort{Beverages}\gdef\@doiabbr{beverages}\gdef\@ISSN{2306-5710}}
\DeclareOption{biochem}{\gdef\@journal{biochem}\gdef\@journalfull{BioChem}\gdef\@journalshort{BioChem}\gdef\@doiabbr{biochem}\gdef\@ISSN{2673-6411}\gdef\@continuouspages{yes}}
\DeclareOption{bioengineering}{\gdef\@journal{bioengineering}\gdef\@journalfull{Bioengineering}\gdef\@journalshort{Bioengineering}\gdef\@doiabbr{bioengineering}\gdef\@ISSN{2306-5354}}
\DeclareOption{biologics}{\gdef\@journal{biologics}\gdef\@journalfull{Biologics}\gdef\@journalshort{Biologics}\gdef\@doiabbr{biologics}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{biology}{\gdef\@journal{biology}\gdef\@journalfull{Biology}\gdef\@journalshort{Biology}\gdef\@doiabbr{biology}\gdef\@ISSN{2079-7737}}
\DeclareOption{biomass}{\gdef\@journal{biomass}\gdef\@journalfull{Biomass}\gdef\@journalshort{Biomass}\gdef\@doiabbr{biomass}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{biomechanics}{\gdef\@journal{biomechanics}\gdef\@journalfull{Biomechanics}\gdef\@journalshort{Biomechanics}\gdef\@doiabbr{biomechanics}\gdef\@ISSN{2673-7078}\gdef\@continuouspages{yes}}
\DeclareOption{biomed}{\gdef\@journal{biomed}\gdef\@journalfull{BioMed}\gdef\@journalshort{BioMed}\gdef\@doiabbr{biomed}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{biomedicines}{\gdef\@journal{biomedicines}\gdef\@journalfull{Biomedicines}\gdef\@journalshort{Biomedicines}\gdef\@doiabbr{biomedicines}\gdef\@ISSN{2227-9059}}
\DeclareOption{biomedinformatics}{\gdef\@journal{biomedinformatics}\gdef\@journalfull{Biomedinformatics}\gdef\@journalshort{Biomedinformatics}\gdef\@doiabbr{biomedinformatics}\gdef\@ISSN{2673-7426}\gdef\@continuouspages{yes}}
\DeclareOption{biomimetics}{\gdef\@journal{biomimetics}\gdef\@journalfull{Biomimetics}\gdef\@journalshort{Biomimetics}\gdef\@doiabbr{biomimetics}\gdef\@ISSN{2313-7673}}
\DeclareOption{biomolecules}{\gdef\@journal{biomolecules}\gdef\@journalfull{Biomolecules}\gdef\@journalshort{Biomolecules}\gdef\@doiabbr{biom}\gdef\@ISSN{2218-273X}}
\DeclareOption{biophysica}{\gdef\@journal{biophysica}\gdef\@journalfull{Biophysica}\gdef\@journalshort{Biophysica}\gdef\@doiabbr{biophysica}\gdef\@ISSN{2673-4125}\gdef\@continuouspages{yes}}
\DeclareOption{biosensors}{\gdef\@journal{biosensors}\gdef\@journalfull{Biosensors}\gdef\@journalshort{Biosensors}\gdef\@doiabbr{bios}\gdef\@ISSN{2079-6374}}
\DeclareOption{biotech}{\gdef\@journal{biotech}\gdef\@journalfull{BioTech}\gdef\@journalshort{BioTech}\gdef\@doiabbr{biotech}\gdef\@ISSN{2673-6284}}
\DeclareOption{birds}{\gdef\@journal{birds}\gdef\@journalfull{Birds}\gdef\@journalshort{Birds}\gdef\@doiabbr{birds}\gdef\@ISSN{2673-6004}\gdef\@continuouspages{yes}}
\DeclareOption{bloods}{\gdef\@journal{bloods}\gdef\@journalfull{Bloods}\gdef\@journalshort{Bloods}\gdef\@doiabbr{bloods}\gdef\@ISSN{2673-415X}}
\DeclareOption{blsf}{\gdef\@journal{blsf}\gdef\@journalfull{Biology and Life Sciences Forum}\gdef\@journalshort{Biol. Life Sci. Forum}\gdef\@doiabbr{blsf	}\gdef\@ISSN{2673-9976}}
\DeclareOption{brainsci}{\gdef\@journal{brainsci}\gdef\@journalfull{Brain Sciences}\gdef\@journalshort{Brain Sci.}\gdef\@doiabbr{brainsci}\gdef\@ISSN{2076-3425}}
\DeclareOption{breath}{\gdef\@journal{breath}\gdef\@journalfull{Breath}\gdef\@journalshort{Breath}\gdef\@doiabbr{breath}\gdef\@ISSN{}}
\DeclareOption{buildings}{\gdef\@journal{buildings}\gdef\@journalfull{Buildings}\gdef\@journalshort{Buildings}\gdef\@doiabbr{buildings}\gdef\@ISSN{2075-5309}}
\DeclareOption{businesses}{\gdef\@journal{businesses}\gdef\@journalfull{Businesses}\gdef\@journalshort{Businesses}\gdef\@doiabbr{businesses}\gdef\@ISSN{2673-7116}\gdef\@continuouspages{yes}}
\DeclareOption{cancers}{\gdef\@journal{cancers}\gdef\@journalfull{Cancers}\gdef\@journalshort{Cancers}\gdef\@doiabbr{cancers}\gdef\@ISSN{2072-6694}}
\DeclareOption{carbon}{\gdef\@journal{carbon}\gdef\@journalfull{C}\gdef\@journalshort{C}\gdef\@doiabbr{c}\gdef\@ISSN{2311-5629}}
\DeclareOption{cardiogenetics}{\gdef\@journal{cardiogenetics}\gdef\@journalfull{Cardiogenetics}\gdef\@journalshort{Cardiogenetics}\gdef\@doiabbr{cardiogenetics}\gdef\@ISSN{2035-8148}\gdef\@continuouspages{yes}}
\DeclareOption{catalysts}{\gdef\@journal{catalysts}\gdef\@journalfull{Catalysts}\gdef\@journalshort{Catalysts}\gdef\@doiabbr{catal}\gdef\@ISSN{2073-4344}}
\DeclareOption{cells}{\gdef\@journal{cells}\gdef\@journalfull{Cells}\gdef\@journalshort{Cells}\gdef\@doiabbr{cells}\gdef\@ISSN{2073-4409}}
\DeclareOption{ceramics}{\gdef\@journal{ceramics}\gdef\@journalfull{Ceramics}\gdef\@journalshort{Ceramics}\gdef\@doiabbr{ceramics}\gdef\@ISSN{2571-6131}\gdef\@continuouspages{yes}}
\DeclareOption{challenges}{\gdef\@journal{challenges}\gdef\@journalfull{Challenges}\gdef\@journalshort{Challenges}\gdef\@doiabbr{challe}\gdef\@ISSN{2078-1547}}
\DeclareOption{chemengineering}{\gdef\@journal{chemengineering}\gdef\@journalfull{ChemEngineering}\gdef\@journalshort{ChemEngineering}\gdef\@doiabbr{chemengineering}\gdef\@ISSN{2305-7084}}
\DeclareOption{chemistry}{\gdef\@journal{chemistry}\gdef\@journalfull{Chemistry}\gdef\@journalshort{Chemistry}\gdef\@doiabbr{chemistry}\gdef\@ISSN{2624-8549}\gdef\@continuouspages{yes}}
\DeclareOption{chemosensors}{\gdef\@journal{chemosensors}\gdef\@journalfull{Chemosensors}\gdef\@journalshort{Chemosensors}\gdef\@doiabbr{chemosensors}\gdef\@ISSN{2227-9040}}
\DeclareOption{chemproc}{\gdef\@journal{chemproc}\gdef\@journalfull{Chemistry Proceedings}\gdef\@journalshort{Chem. Proc.}\gdef\@doiabbr{chemproc}\gdef\@ISSN{2673-4583}}
\DeclareOption{children}{\gdef\@journal{children}\gdef\@journalfull{Children}\gdef\@journalshort{Children}\gdef\@doiabbr{children}\gdef\@ISSN{2227-9067}}
\DeclareOption{chips}{\gdef\@journal{chips}\gdef\@journalfull{Chips}\gdef\@journalshort{Chips}\gdef\@doiabbr{chips}\gdef\@ISSN{2674-0710}\gdef\@continuouspages{yes}}
\DeclareOption{cimb}{\gdef\@journal{cimb}\gdef\@journalfull{Current Issues in Molecular Biology}\gdef\@journalshort{Curr. Issues Mol. Biol.}\gdef\@doiabbr{cimb}\gdef\@ISSN{1467-3045}\gdef\@continuouspages{yes}}
\DeclareOption{civileng}{\gdef\@journal{civileng}\gdef\@journalfull{CivilEng}\gdef\@journalshort{CivilEng}\gdef\@doiabbr{civileng}\gdef\@ISSN{2673-4109}\gdef\@continuouspages{yes}}
\DeclareOption{cleantechnol}{\gdef\@journal{cleantechnol}\gdef\@journalfull{Clean Technologies}\gdef\@journalshort{Clean Technol.}\gdef\@doiabbr{cleantechnol}\gdef\@ISSN{2571-8797}\gdef\@continuouspages{yes}}
\DeclareOption{climate}{\gdef\@journal{climate}\gdef\@journalfull{Climate}\gdef\@journalshort{Climate}\gdef\@doiabbr{cli}\gdef\@ISSN{2225-1154}}
\DeclareOption{clinpract}{\gdef\@journal{clinpract}\gdef\@journalfull{Clinics and Practice}\gdef\@journalshort{Clin. Pract.}\gdef\@doiabbr{clinpract}\gdef\@ISSN{2039-7283}\gdef\@continuouspages{yes}}
\DeclareOption{clockssleep}{\gdef\@journal{clockssleep}\gdef\@journalfull{Clocks \& Sleep}\gdef\@journalshort{Clocks\&Sleep}\gdef\@doiabbr{clockssleep}\gdef\@ISSN{2624-5175}\gdef\@continuouspages{yes}}
\DeclareOption{cmd}{\gdef\@journal{cmd}\gdef\@journalfull{Corrosion and Materials Degradation}\gdef\@journalshort{Corros. Mater. Degrad.}\gdef\@doiabbr{cmd}\gdef\@ISSN{2624-5558}\gdef\@continuouspages{yes}}
\DeclareOption{coasts}{\gdef\@journal{coasts}\gdef\@journalfull{Coasts}\gdef\@journalshort{Coasts}\gdef\@doiabbr{coasts}\gdef\@ISSN{2673-964X}\gdef\@continuouspages{yes}}
\DeclareOption{coatings}{\gdef\@journal{coatings}\gdef\@journalfull{Coatings}\gdef\@journalshort{Coatings}\gdef\@doiabbr{coatings}\gdef\@ISSN{2079-6412}}
\DeclareOption{colloids}{\gdef\@journal{colloids}\gdef\@journalfull{Colloids Interfaces}\gdef\@journalshort{Colloids Interfaces}\gdef\@doiabbr{colloids}\gdef\@ISSN{2504-5377}}
\DeclareOption{colorants}{\gdef\@journal{colorants}\gdef\@journalfull{Colorants}\gdef\@journalshort{Colorants}\gdef\@doiabbr{colorants}\gdef\@ISSN{2079-6447}\gdef\@continuouspages{yes}}
\DeclareOption{commodities}{\gdef\@journal{commodities}\gdef\@journalfull{Commodities}\gdef\@journalshort{Commodities}\gdef\@doiabbr{commodities}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{compounds}{\gdef\@journal{compounds}\gdef\@journalfull{Compounds}\gdef\@journalshort{Compounds}\gdef\@doiabbr{compounds}\gdef\@ISSN{2673-6918}\gdef\@continuouspages{yes}}
\DeclareOption{computation}{\gdef\@journal{computation}\gdef\@journalfull{Computation}\gdef\@journalshort{Computation}\gdef\@doiabbr{computation}\gdef\@ISSN{2079-3197}}
\DeclareOption{computers}{\gdef\@journal{computers}\gdef\@journalfull{Computers}\gdef\@journalshort{Computers}\gdef\@doiabbr{computers}\gdef\@ISSN{2073-431X}}
\DeclareOption{condensedmatter}{\gdef\@journal{condensedmatter}\gdef\@journalfull{Condensed Matter}\gdef\@journalshort{Condens. Matter}\gdef\@doiabbr{condmat}\gdef\@ISSN{2410-3896}}
\DeclareOption{conservation}{\gdef\@journal{conservation}\gdef\@journalfull{Conservation}\gdef\@journalshort{Conservation}\gdef\@doiabbr{conservation}\gdef\@ISSN{2673-7159}\gdef\@continuouspages{yes}}
\DeclareOption{constrmater}{\gdef\@journal{constrmater}\gdef\@journalfull{Construction Materials}\gdef\@journalshort{Constr. Mater.}\gdef\@doiabbr{constrmater}\gdef\@ISSN{2673-7108}\gdef\@continuouspages{yes}}
\DeclareOption{cosmetics}{\gdef\@journal{cosmetics}\gdef\@journalfull{Cosmetics}\gdef\@journalshort{Cosmetics}\gdef\@doiabbr{cosmetics}\gdef\@ISSN{2079-9284}}
\DeclareOption{covid}{\gdef\@journal{covid}\gdef\@journalfull{COVID}\gdef\@journalshort{COVID}\gdef\@doiabbr{covid}\gdef\@ISSN{2673-8112}\gdef\@continuouspages{yes}}
\DeclareOption{crops}{\gdef\@journal{crops}\gdef\@journalfull{Crops}\gdef\@journalshort{Crops}\gdef\@doiabbr{crops}\gdef\@ISSN{2673-7655}\gdef\@continuouspages{yes}}
\DeclareOption{cryptography}{\gdef\@journal{cryptography}\gdef\@journalfull{Cryptography}\gdef\@journalshort{Cryptography}\gdef\@doiabbr{cryptography}\gdef\@ISSN{2410-387X}}
\DeclareOption{crystals}{\gdef\@journal{crystals}\gdef\@journalfull{Crystals}\gdef\@journalshort{Crystals}\gdef\@doiabbr{cryst}\gdef\@ISSN{2073-4352}}
\DeclareOption{csmf}{\gdef\@journal{csmf}\gdef\@journalfull{Computer Sciences \& Mathematics Forum}\gdef\@journalshort{Comput. Sci. Math. Forum}\gdef\@doiabbr{csmf}\gdef\@ISSN{2813-0324}}
\DeclareOption{ctn}{\gdef\@journal{ctn}\gdef\@journalfull{Clinical and Translational Neuroscience}\gdef\@journalshort{CTN}\gdef\@doiabbr{Clin. Transl. Neurosci.}\gdef\@ISSN{2514-183X}}
\DeclareOption{curroncol}{\gdef\@journal{curroncol}\gdef\@journalfull{Current Oncology}\gdef\@journalshort{Curr. Oncol.}\gdef\@doiabbr{curroncol}\gdef\@ISSN{1718-7729}\gdef\@continuouspages{yes}}
\DeclareOption{cyber}{\gdef\@journal{cyber}\gdef\@journalfull{Cyber}\gdef\@journalshort{Cyber}\gdef\@doiabbr{cyber}\gdef\@ISSN{2624-800X}}
\DeclareOption{dairy}{\gdef\@journal{dairy}\gdef\@journalfull{Dairy}\gdef\@journalshort{Dairy}\gdef\@doiabbr{dairy}\gdef\@ISSN{2624-862X}\gdef\@continuouspages{yes}}
\DeclareOption{data}{\gdef\@journal{data}\gdef\@journalfull{Data}\gdef\@journalshort{Data}\gdef\@doiabbr{data}\gdef\@ISSN{2306-5729}}
\DeclareOption{ddc}{\gdef\@journal{ddc}\gdef\@journalfull{Drugs and Drug Candidates}\gdef\@journalshort{Drugs Drug Candidates}\gdef\@doiabbr{ddc}\gdef\@ISSN{2813-2998}\gdef\@continuouspages{yes}}
\DeclareOption{dentistry}{\gdef\@journal{dentistry}\gdef\@journalfull{Dentistry Journal}\gdef\@journalshort{Dent. J.}\gdef\@doiabbr{dj}\gdef\@ISSN{2304-6767}}
\DeclareOption{dermato}{\gdef\@journal{dermato}\gdef\@journalfull{Dermato}\gdef\@journalshort{Dermato}\gdef\@doiabbr{dermato}\gdef\@ISSN{2673-6179}\gdef\@continuouspages{yes}}
\DeclareOption{dermatopathology}{\gdef\@journal{dermatopathology}\gdef\@journalfull{Dermatopathology}\gdef\@journalshort{Dermatopathology}\gdef\@doiabbr{dermatopathology}\gdef\@ISSN{2296-3529}\gdef\@continuouspages{yes}}
\DeclareOption{designs}{\gdef\@journal{designs}\gdef\@journalfull{Designs}\gdef\@journalshort{Designs}\gdef\@doiabbr{designs}\gdef\@ISSN{2411-9660}}
\DeclareOption{devices}{\gdef\@journal{devices}\gdef\@journalfull{Devices}\gdef\@journalshort{Devices}\gdef\@doiabbr{devices}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{diabetology}{\gdef\@journal{diabetology}\gdef\@journalfull{Diabetology}\gdef\@journalshort{Diabetology}\gdef\@doiabbr{diabetology}\gdef\@ISSN{2673-4540}\gdef\@continuouspages{yes}}
\DeclareOption{diagnostics}{\gdef\@journal{diagnostics}\gdef\@journalfull{Diagnostics}\gdef\@journalshort{Diagnostics}\gdef\@doiabbr{diagnostics}\gdef\@ISSN{2075-4418}}
\DeclareOption{dietetics}{\gdef\@journal{dietetics}\gdef\@journalfull{Dietetics}\gdef\@journalshort{Dietetics}\gdef\@doiabbr{dietetics}\gdef\@ISSN{2674-0311}\gdef\@continuouspages{yes}}
\DeclareOption{digital}{\gdef\@journal{digital}\gdef\@journalfull{Digital}\gdef\@journalshort{Digital}\gdef\@doiabbr{digital}\gdef\@ISSN{2673-6470}\gdef\@continuouspages{yes}}
\DeclareOption{disabilities}{\gdef\@journal{disabilities}\gdef\@journalfull{Disabilities}\gdef\@journalshort{Disabilities}\gdef\@doiabbr{disabilities}\gdef\@ISSN{2673-7272}\gdef\@continuouspages{yes}}
\DeclareOption{diseases}{\gdef\@journal{diseases}\gdef\@journalfull{Diseases}\gdef\@journalshort{Diseases}\gdef\@doiabbr{diseases}\gdef\@ISSN{2079-9721}}
\DeclareOption{diversity}{\gdef\@journal{diversity}\gdef\@journalfull{Diversity}\gdef\@journalshort{Diversity}\gdef\@doiabbr{d}\gdef\@ISSN{1424-2818}}
\DeclareOption{dna}{\gdef\@journal{dna}\gdef\@journalfull{DNA}\gdef\@journalshort{DNA}\gdef\@doiabbr{dna}\gdef\@ISSN{2673-8856}\gdef\@continuouspages{yes}}
\DeclareOption{drones}{\gdef\@journal{drones}\gdef\@journalfull{Drones}\gdef\@journalshort{Drones}\gdef\@doiabbr{drones}\gdef\@ISSN{2504-446X}}
\DeclareOption{dynamics}{\gdef\@journal{dynamics}\gdef\@journalfull{Dynamics}\gdef\@journalshort{Dynamics}\gdef\@doiabbr{dynamics}\gdef\@ISSN{2673-8716}\gdef\@continuouspages{yes}}
\DeclareOption{earth}{\gdef\@journal{earth}\gdef\@journalfull{Earth}\gdef\@journalshort{Earth}\gdef\@doiabbr{earth}\gdef\@ISSN{2673-4834}\gdef\@continuouspages{yes}}
\DeclareOption{ebj}{\gdef\@journal{ebj}\gdef\@journalfull{European Burn Journal}\gdef\@journalshort{Eur. Burn J.}\gdef\@doiabbr{ebj}\gdef\@ISSN{2673-1991}\gdef\@continuouspages{yes}}
\DeclareOption{ecologies}{\gdef\@journal{ecologies}\gdef\@journalfull{Ecologies}\gdef\@journalshort{Ecologies}\gdef\@doiabbr{ecologies}\gdef\@ISSN{2673-4133}\gdef\@continuouspages{yes}}
\DeclareOption{econometrics}{\gdef\@journal{econometrics}\gdef\@journalfull{Econometrics}\gdef\@journalshort{Econometrics}\gdef\@doiabbr{econometrics}\gdef\@ISSN{2225-1146}}
\DeclareOption{economies}{\gdef\@journal{economies}\gdef\@journalfull{Economies}\gdef\@journalshort{Economies}\gdef\@doiabbr{economies}\gdef\@ISSN{2227-7099}}
\DeclareOption{education}{\gdef\@journal{education}\gdef\@journalfull{Education Sciences}\gdef\@journalshort{Educ. Sci.}\gdef\@doiabbr{educsci}\gdef\@ISSN{2227-7102}}
\DeclareOption{ejihpe}{\gdef\@journal{ejihpe}\gdef\@journalfull{European Journal of Investigation in Health Psychology and Education}\gdef\@journalshort{Eur. J. Investig. Health Psychol. Educ.}\gdef\@doiabbr{ejihpe}\gdef\@ISSN{2254-9625}\gdef\@continuouspages{yes}}
\DeclareOption{electricity}{\gdef\@journal{electricity}\gdef\@journalfull{Electricity}\gdef\@journalshort{Electricity}\gdef\@doiabbr{electricity}\gdef\@ISSN{2673-4826}\gdef\@continuouspages{yes}}
\DeclareOption{electrochem}{\gdef\@journal{electrochem}\gdef\@journalfull{Electrochem}\gdef\@journalshort{Electrochem}\gdef\@doiabbr{electrochem}\gdef\@ISSN{2673-3293}\gdef\@continuouspages{yes}}
\DeclareOption{electronicmat}{\gdef\@journal{electronicmat}\gdef\@journalfull{Electronic Materials}\gdef\@journalshort{Electron. Mater.}\gdef\@doiabbr{electronicmat}\gdef\@ISSN{2673-3978}\gdef\@continuouspages{yes}}
\DeclareOption{electronics}{\gdef\@journal{electronics}\gdef\@journalfull{Electronics}\gdef\@journalshort{Electronics}\gdef\@doiabbr{electronics}\gdef\@ISSN{2079-9292}}
\DeclareOption{encyclopedia}{\gdef\@journal{encyclopedia}\gdef\@journalfull{Encyclopedia}\gdef\@journalshort{Encyclopedia}\gdef\@doiabbr{encyclopedia}\gdef\@ISSN{2673-8392}\gdef\@continuouspages{yes}}
\DeclareOption{endocrines}{\gdef\@journal{endocrines}\gdef\@journalfull{Endocrines}\gdef\@journalshort{Endocrines}\gdef\@doiabbr{endocrines}\gdef\@ISSN{2673-396X}\gdef\@continuouspages{yes}}
\DeclareOption{energies}{\gdef\@journal{energies}\gdef\@journalfull{Energies}\gdef\@journalshort{Energies}\gdef\@doiabbr{en}\gdef\@ISSN{1996-1073}}
\DeclareOption{eng}{\gdef\@journal{eng}\gdef\@journalfull{Eng-Advances in Engineering}\gdef\@journalshort{Eng}\gdef\@doiabbr{eng}\gdef\@ISSN{2673-4117}\gdef\@continuouspages{yes}}
\DeclareOption{engproc}{\gdef\@journal{engproc}\gdef\@journalfull{Engineering Proceedings}\gdef\@journalshort{Eng. Proc.}\gdef\@doiabbr{engproc}\gdef\@ISSN{2673-4591}}
%\DeclareOption{ent}{\gdef\@journal{ent}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{entomology}{\gdef\@journal{entomology}\gdef\@journalfull{Entomology}\gdef\@journalshort{Entomology}\gdef\@doiabbr{entomology}\gdef\@ISSN{2813-0502}\gdef\@continuouspages{yes}}	
\DeclareOption{entropy}{\gdef\@journal{entropy}\gdef\@journalfull{Entropy}\gdef\@journalshort{Entropy}\gdef\@doiabbr{e}\gdef\@ISSN{1099-4300}}
\DeclareOption{environments}{\gdef\@journal{environments}\gdef\@journalfull{Environments}\gdef\@journalshort{Environments}\gdef\@doiabbr{environments}\gdef\@ISSN{2076-3298}}
\DeclareOption{environsciproc}{\gdef\@journal{environsciproc}\gdef\@journalfull{Environmental Sciences Proceedings}\gdef\@journalshort{Environ. Sci. Proc.}\gdef\@doiabbr{environsciproc}\gdef\@ISSN{2673-4931}}
\DeclareOption{epidemiologia}{\gdef\@journal{epidemiologia}\gdef\@journalfull{Epidemiologia}\gdef\@journalshort{Epidemiologia}\gdef\@doiabbr{epidemiologia}\gdef\@ISSN{2673-3986}\gdef\@continuouspages{yes}}
\DeclareOption{epigenomes}{\gdef\@journal{epigenomes}\gdef\@journalfull{Epigenomes}\gdef\@journalshort{Epigenomes}\gdef\@doiabbr{epigenomes}\gdef\@ISSN{2075-4655}}
\DeclareOption{est}{\gdef\@journal{est}\gdef\@journalfull{Electrochemical Science and Technology}\gdef\@journalshort{Electrochem. Sci. Technol.}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{fermentation}{\gdef\@journal{fermentation}\gdef\@journalfull{Fermentation}\gdef\@journalshort{Fermentation}\gdef\@doiabbr{fermentation}\gdef\@ISSN{2311-5637}}
\DeclareOption{fibers}{\gdef\@journal{fibers}\gdef\@journalfull{Fibers}\gdef\@journalshort{Fibers}\gdef\@doiabbr{fib}\gdef\@ISSN{2079-6439}}
\DeclareOption{fintech}{\gdef\@journal{fintech}\gdef\@journalfull{FinTech}\gdef\@journalshort{FinTech}\gdef\@doiabbr{fintech}\gdef\@ISSN{2674-1032}\gdef\@continuouspages{yes}}
\DeclareOption{fire}{\gdef\@journal{fire}\gdef\@journalfull{Fire}\gdef\@journalshort{Fire}\gdef\@doiabbr{fire}\gdef\@ISSN{2571-6255}}
\DeclareOption{fishes}{\gdef\@journal{fishes}\gdef\@journalfull{Fishes}\gdef\@journalshort{Fishes}\gdef\@doiabbr{fishes}\gdef\@ISSN{2410-3888}}
\DeclareOption{fluids}{\gdef\@journal{fluids}\gdef\@journalfull{Fluids}\gdef\@journalshort{Fluids}\gdef\@doiabbr{fluids}\gdef\@ISSN{2311-5521}}
\DeclareOption{foods}{\gdef\@journal{foods}\gdef\@journalfull{Foods}\gdef\@journalshort{Foods}\gdef\@doiabbr{foods}\gdef\@ISSN{2304-8158}}
\DeclareOption{forecasting}{\gdef\@journal{forecasting}\gdef\@journalfull{Forecasting}\gdef\@journalshort{Forecasting}\gdef\@doiabbr{forecast}\gdef\@ISSN{2571-9394}\gdef\@continuouspages{yes}}
\DeclareOption{forensicsci}{\gdef\@journal{forensicsci}\gdef\@journalfull{Forensic Sciences}\gdef\@journalshort{Forensic Sci.}\gdef\@doiabbr{forensicsci}\gdef\@ISSN{2673-6756}\gdef\@continuouspages{yes}}
\DeclareOption{forests}{\gdef\@journal{forests}\gdef\@journalfull{Forests}\gdef\@journalshort{Forests}\gdef\@doiabbr{f}\gdef\@ISSN{1999-4907}}
\DeclareOption{foundations}{\gdef\@journal{foundations}\gdef\@journalfull{Foundations}\gdef\@journalshort{Foundations}\gdef\@doiabbr{foundations}\gdef\@ISSN{2673-9321}\gdef\@continuouspages{yes}}
\DeclareOption{fractalfract}{\gdef\@journal{fractalfract}\gdef\@journalfull{Fractal and Fractional}\gdef\@journalshort{Fractal Fract.}\gdef\@doiabbr{fractalfract}\gdef\@ISSN{2504-3110}}
\DeclareOption{fuels}{\gdef\@journal{fuels}\gdef\@journalfull{Fuels}\gdef\@journalshort{Fuels}\gdef\@doiabbr{fuels}\gdef\@ISSN{2673-3994}\gdef\@continuouspages{yes}}
\DeclareOption{future}{\gdef\@journal{future}\gdef\@journalfull{Future}\gdef\@journalshort{Future}\gdef\@doiabbr{future}\gdef\@ISSN{2813-2882}\gdef\@continuouspages{yes}}
\DeclareOption{futureinternet}{\gdef\@journal{futureinternet}\gdef\@journalfull{Future Internet}\gdef\@journalshort{Future Internet}\gdef\@doiabbr{fi}\gdef\@ISSN{1999-5903}}
%\DeclareOption{futureparasites}{\gdef\@journal{futureparasites}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{futurepharmacol}{\gdef\@journal{futurepharmacol}\gdef\@journalfull{Future Pharmacology}\gdef\@journalshort{Future Pharm.}\gdef\@doiabbr{futurepharmacol}\gdef\@ISSN{2673-9879}\gdef\@continuouspages{yes}}
\DeclareOption{futurephys}{\gdef\@journal{futurephys}\gdef\@journalfull{Future Physics}\gdef\@journalshort{Future Phys.}\gdef\@doiabbr{futurephys}\gdef\@ISSN{2624-6503}}
\DeclareOption{futuretransp}{\gdef\@journal{futuretransp}\gdef\@journalfull{Future Transportation}\gdef\@journalshort{Future Transp.}\gdef\@doiabbr{futuretransp}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{galaxies}{\gdef\@journal{galaxies}\gdef\@journalfull{Galaxies}\gdef\@journalshort{Galaxies}\gdef\@doiabbr{galaxies}\gdef\@ISSN{2075-4434}}
\DeclareOption{games}{\gdef\@journal{games}\gdef\@journalfull{Games}\gdef\@journalshort{Games}\gdef\@doiabbr{g}\gdef\@ISSN{2073-4336}}
\DeclareOption{gases}{\gdef\@journal{gases}\gdef\@journalfull{Gases}\gdef\@journalshort{Gases}\gdef\@doiabbr{gases}\gdef\@ISSN{2673-5628}\gdef\@continuouspages{yes}}
\DeclareOption{gastroent}{\gdef\@journal{gastroent}\gdef\@journalfull{Gastroenterology Insights}\gdef\@journalshort{Gastroenterol. Insights}\gdef\@doiabbr{gastroent}\gdef\@ISSN{2036-7422}\gdef\@continuouspages{yes}}
\DeclareOption{gastrointestdisord}{\gdef\@journal{gastrointestdisord}\gdef\@journalfull{Gastrointestinal Disorders}\gdef\@journalshort{Gastrointest. Disord.}\gdef\@doiabbr{gidisord}\gdef\@ISSN{2624-5647}\gdef\@continuouspages{yes}}
\DeclareOption{gels}{\gdef\@journal{gels}\gdef\@journalfull{Gels}\gdef\@journalshort{Gels}\gdef\@doiabbr{gels}\gdef\@ISSN{2310-2861}}
\DeclareOption{genealogy}{\gdef\@journal{genealogy}\gdef\@journalfull{Genealogy}\gdef\@journalshort{Genealogy}\gdef\@doiabbr{genealogy}\gdef\@ISSN{2313-5778}}
\DeclareOption{genes}{\gdef\@journal{genes}\gdef\@journalfull{Genes}\gdef\@journalshort{Genes}\gdef\@doiabbr{genes}\gdef\@ISSN{2073-4425}}
\DeclareOption{geographies}{\gdef\@journal{geographies}\gdef\@journalfull{Geographies}\gdef\@journalshort{Geographies}\gdef\@doiabbr{geographies}\gdef\@ISSN{2673-7086}\gdef\@continuouspages{yes}}
\DeclareOption{geohazards}{\gdef\@journal{geohazards}\gdef\@journalfull{GeoHazards}\gdef\@journalshort{GeoHazards}\gdef\@doiabbr{geohazards}\gdef\@ISSN{2624-795X}\gdef\@continuouspages{yes}}
\DeclareOption{geomatics}{\gdef\@journal{geomatics}\gdef\@journalfull{Geomatics}\gdef\@journalshort{Geomatics}\gdef\@doiabbr{geomatics}\gdef\@ISSN{2673-7418}\gdef\@continuouspages{yes}}
\DeclareOption{geosciences}{\gdef\@journal{geosciences}\gdef\@journalfull{Geosciences}\gdef\@journalshort{Geosciences}\gdef\@doiabbr{geosciences}\gdef\@ISSN{2076-3263}}
\DeclareOption{geotechnics}{\gdef\@journal{geotechnics}\gdef\@journalfull{Geotechnics}\gdef\@journalshort{Geotechnics}\gdef\@doiabbr{geotechnics}\gdef\@ISSN{2673-7094}\gdef\@continuouspages{yes}}
\DeclareOption{geriatrics}{\gdef\@journal{geriatrics}\gdef\@journalfull{Geriatrics}\gdef\@journalshort{Geriatrics}\gdef\@doiabbr{geriatrics}\gdef\@ISSN{2308-3417}}
\DeclareOption{grasses}{\gdef\@journal{grasses}\gdef\@journalfull{Grasses}\gdef\@journalshort{Grasses}\gdef\@doiabbr{grasses}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{hazardousmatters}{\gdef\@journal{hazardousmatters}\gdef\@journalfull{Hazardous Matters}\gdef\@journalshort{Hazard. Matters}\gdef\@doiabbr{}\gdef\@ISSN{0014-0014}}
\DeclareOption{healthcare}{\gdef\@journal{healthcare}\gdef\@journalfull{Healthcare}\gdef\@journalshort{Healthcare}\gdef\@doiabbr{healthcare}\gdef\@ISSN{2227-9032}}
\DeclareOption{hearts}{\gdef\@journal{hearts}\gdef\@journalfull{Hearts}\gdef\@journalshort{Hearts}\gdef\@doiabbr{hearts}\gdef\@ISSN{2673-3846}\gdef\@continuouspages{yes}}
\DeclareOption{hemato}{\gdef\@journal{hemato}\gdef\@journalfull{Hemato}\gdef\@journalshort{Hemato}\gdef\@doiabbr{hemato}\gdef\@ISSN{2673-6357}\gdef\@continuouspages{yes}}
\DeclareOption{hematolrep}{\gdef\@journal{hematolrep}\gdef\@journalfull{Hematology Reports}\gdef\@journalshort{Hematol. Rep.}\gdef\@doiabbr{hematolrep}\gdef\@ISSN{2038-8330}\gdef\@continuouspages{yes}}
\DeclareOption{heritage}{\gdef\@journal{heritage}\gdef\@journalfull{Heritage}\gdef\@journalshort{Heritage}\gdef\@doiabbr{heritage}\gdef\@ISSN{2571-9408}\gdef\@continuouspages{yes}}
\DeclareOption{higheredu}{\gdef\@journal{higheredu}\gdef\@journalfull{Trends in Higher Education}\gdef\@journalshort{Trends High. Educ.}\gdef\@doiabbr{higheredu}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{highthroughput}{\gdef\@journal{highthroughput}\gdef\@journalfull{High-Throughput}\gdef\@journalshort{High-Throughput}\gdef\@doiabbr{ht}\gdef\@ISSN{2571-5135}}
\DeclareOption{histories}{\gdef\@journal{histories}\gdef\@journalfull{Histories}\gdef\@journalshort{Histories}\gdef\@doiabbr{histories}\gdef\@ISSN{2409-9252}\gdef\@continuouspages{yes}}
\DeclareOption{horticulturae}{\gdef\@journal{horticulturae}\gdef\@journalfull{Horticulturae}\gdef\@journalshort{Horticulturae}\gdef\@doiabbr{horticulturae}\gdef\@ISSN{2311-7524}}
\DeclareOption{humanities}{\gdef\@journal{humanities}\gdef\@journalfull{Humanities}\gdef\@journalshort{Humanities}\gdef\@doiabbr{h}\gdef\@ISSN{2076-0787}}
\DeclareOption{humans}{\gdef\@journal{humans}\gdef\@journalfull{Humans}\gdef\@journalshort{Humans}\gdef\@doiabbr{humans}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{hydrobiology}{\gdef\@journal{hydrobiology}\gdef\@journalfull{Hydrobiology}\gdef\@journalshort{Hydrobiology	}\gdef\@doiabbr{hydrobiology}\gdef\@ISSN{2673-9917}\gdef\@continuouspages{yes}}
\DeclareOption{hydrogen}{\gdef\@journal{hydrogen}\gdef\@journalfull{Hydrogen}\gdef\@journalshort{Hydrogen}\gdef\@doiabbr{hydrogen}\gdef\@ISSN{2673-4141}\gdef\@continuouspages{yes}}
\DeclareOption{hydrology}{\gdef\@journal{hydrology}\gdef\@journalfull{Hydrology}\gdef\@journalshort{Hydrology}\gdef\@doiabbr{hydrology}\gdef\@ISSN{2306-5338}}
\DeclareOption{hygiene}{\gdef\@journal{hygiene}\gdef\@journalfull{Hygiene}\gdef\@journalshort{Hygiene}\gdef\@doiabbr{hygiene}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{idr}{\gdef\@journal{idr}\gdef\@journalfull{Infectious Disease Reports}\gdef\@journalshort{Infect. Dis. Rep.}\gdef\@doiabbr{idr}\gdef\@ISSN{2036-7449}\gdef\@continuouspages{yes}}
\DeclareOption{ijerph}{\gdef\@journal{ijerph}\gdef\@journalfull{International Journal of Environmental Research and Public Health}\gdef\@journalshort{Int. J. Environ. Res. Public Health}\gdef\@doiabbr{ijerph}\gdef\@ISSN{1660-4601}}
\DeclareOption{ijfs}{\gdef\@journal{ijfs}\gdef\@journalfull{International Journal of Financial Studies}\gdef\@journalshort{Int. J. Financial Stud.}\gdef\@doiabbr{ijfs}\gdef\@ISSN{2227-7072}}
\DeclareOption{ijgi}{\gdef\@journal{ijgi}\gdef\@journalfull{ISPRS International Journal of Geo-Information}\gdef\@journalshort{ISPRS Int. J. Geo-Inf.}\gdef\@doiabbr{ijgi}\gdef\@ISSN{2220-9964}}
\DeclareOption{ijms}{\gdef\@journal{ijms}\gdef\@journalfull{International Journal of Molecular Sciences}\gdef\@journalshort{Int. J. Mol. Sci.}\gdef\@doiabbr{ijms}\gdef\@ISSN{1422-0067}}
\DeclareOption{ijns}{\gdef\@journal{ijns}\gdef\@journalfull{International Journal of Neonatal Screening}\gdef\@journalshort{Int. J. Neonatal Screen.}\gdef\@doiabbr{ijns}\gdef\@ISSN{2409-515X}}
\DeclareOption{ijpb}{\gdef\@journal{ijpb}\gdef\@journalfull{International Journal of Plant Biology}\gdef\@journalshort{Int. J. Plant Biol.}\gdef\@doiabbr{ijpb}\gdef\@ISSN{2037-0164}\gdef\@continuouspages{yes}}
\DeclareOption{ijtm}{\gdef\@journal{ijtm}\gdef\@journalfull{International Journal of Translational Medicine}\gdef\@journalshort{Int. J. Transl. Med.}\gdef\@doiabbr{ijtm}\gdef\@ISSN{2673-8937}\gdef\@continuouspages{yes}}
\DeclareOption{ijtpp}{\gdef\@journal{ijtpp}\gdef\@journalfull{International Journal of Turbomachinery Propulsion and Power}\gdef\@journalshort{Int. J. Turbomach. Propuls. Power}\gdef\@doiabbr{ijtpp}\gdef\@ISSN{2504-186X}}
\DeclareOption{ime}{\gdef\@journal{ime}\gdef\@journalfull{International Medical Education}\gdef\@journalshort{Int. Med. Educ.}\gdef\@doiabbr{ime}\gdef\@ISSN{2813-141X}\gdef\@continuouspages{yes}}
\DeclareOption{immuno}{\gdef\@journal{immuno}\gdef\@journalfull{Immuno}\gdef\@journalshort{Immuno}\gdef\@doiabbr{immuno}\gdef\@ISSN{2673-5601}\gdef\@continuouspages{yes}}
\DeclareOption{informatics}{\gdef\@journal{informatics}\gdef\@journalfull{Informatics}\gdef\@journalshort{Informatics}\gdef\@doiabbr{informatics}\gdef\@ISSN{2227-9709}}
\DeclareOption{information}{\gdef\@journal{information}\gdef\@journalfull{Information}\gdef\@journalshort{Information}\gdef\@doiabbr{info}\gdef\@ISSN{2078-2489}}
\DeclareOption{infrastructures}{\gdef\@journal{infrastructures}\gdef\@journalfull{Infrastructures}\gdef\@journalshort{Infrastructures}\gdef\@doiabbr{infrastructures}\gdef\@ISSN{2412-3811}}
\DeclareOption{inorganics}{\gdef\@journal{inorganics}\gdef\@journalfull{Inorganics}\gdef\@journalshort{Inorganics}\gdef\@doiabbr{inorganics}\gdef\@ISSN{2304-6740}}
\DeclareOption{insects}{\gdef\@journal{insects}\gdef\@journalfull{Insects}\gdef\@journalshort{Insects}\gdef\@doiabbr{insects}\gdef\@ISSN{2075-4450}}
\DeclareOption{instruments}{\gdef\@journal{instruments}\gdef\@journalfull{Instruments}\gdef\@journalshort{Instruments}\gdef\@doiabbr{instruments}\gdef\@ISSN{2410-390X}}
\DeclareOption{inventions}{\gdef\@journal{inventions}\gdef\@journalfull{Inventions}\gdef\@journalshort{Inventions}\gdef\@doiabbr{inventions}\gdef\@ISSN{2411-5134}}
\DeclareOption{iot}{\gdef\@journal{iot}\gdef\@journalfull{IoT}\gdef\@journalshort{IoT}\gdef\@doiabbr{iot}\gdef\@ISSN{2624-831X}\gdef\@continuouspages{yes}}
\DeclareOption{j}{\gdef\@journal{j}\gdef\@journalfull{J}\gdef\@journalshort{J}\gdef\@doiabbr{j}\gdef\@ISSN{2571-8800}\gdef\@continuouspages{yes}}
\DeclareOption{jal}{\gdef\@journal{jal}\gdef\@journalfull{Journal of Ageing and Longevity}\gdef\@journalshort{J. Ageing Longev.}\gdef\@doiabbr{jal}\gdef\@ISSN{2673-9259}\gdef\@continuouspages{yes}}
\DeclareOption{jcdd}{\gdef\@journal{jcdd}\gdef\@journalfull{Journal of Cardiovascular Development and Disease}\gdef\@journalshort{J. Cardiovasc. Dev. Dis.}\gdef\@doiabbr{jcdd}\gdef\@ISSN{2308-3425}}
\DeclareOption{jcm}{\gdef\@journal{jcm}\gdef\@journalfull{Journal of Clinical Medicine}\gdef\@journalshort{J. Clin. Med.}\gdef\@doiabbr{jcm}\gdef\@ISSN{2077-0383}}
\DeclareOption{jcp}{\gdef\@journal{jcp}\gdef\@journalfull{Journal of Cybersecurity and Privacy}\gdef\@journalshort{J. Cybersecur. Priv.}\gdef\@doiabbr{jcp}\gdef\@ISSN{2624-800X}\gdef\@continuouspages{yes}}
\DeclareOption{jcs}{\gdef\@journal{jcs}\gdef\@journalfull{Journal of Composites Science}\gdef\@journalshort{J. Compos. Sci.}\gdef\@doiabbr{jcs}\gdef\@ISSN{2504-477X}}
\DeclareOption{jcto}{\gdef\@journal{jcto}\gdef\@journalfull{Journal of Clinical \& Translational Ophthalmology}\gdef\@journalshort{J. Clin. Transl. Ophthalmol.}\gdef\@doiabbr{jcto}\gdef\@ISSN{2813-1053}\gdef\@continuouspages{yes}}
\DeclareOption{jdb}{\gdef\@journal{jdb}\gdef\@journalfull{Journal of Developmental Biology}\gdef\@journalshort{J. Dev. Biol.}\gdef\@doiabbr{jdb}\gdef\@ISSN{2221-3759}}
\DeclareOption{jeta}{\gdef\@journal{jeta}\gdef\@journalfull{Journal of Experimental and Theoretical Analyses}\gdef\@journalshort{J. Exp. Theor. Anal.}\gdef\@doiabbr{jeta}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{jfb}{\gdef\@journal{jfb}\gdef\@journalfull{Journal of Functional Biomaterials}\gdef\@journalshort{J. Funct. Biomater.}\gdef\@doiabbr{jfb}\gdef\@ISSN{2079-4983}}
\DeclareOption{jfmk}{\gdef\@journal{jfmk}\gdef\@journalfull{Journal of Functional Morphology and Kinesiology}\gdef\@journalshort{J. Funct. Morphol. Kinesiol.}\gdef\@doiabbr{jfmk}\gdef\@ISSN{2411-5142}}
\DeclareOption{jimaging}{\gdef\@journal{jimaging}\gdef\@journalfull{Journal of Imaging}\gdef\@journalshort{J. Imaging}\gdef\@doiabbr{jimaging}\gdef\@ISSN{2313-433X}}
\DeclareOption{jintelligence}{\gdef\@journal{jintelligence}\gdef\@journalfull{Journal of Intelligence}\gdef\@journalshort{J. Intell.}\gdef\@doiabbr{jintelligence}\gdef\@ISSN{2079-3200}}
\DeclareOption{jlpea}{\gdef\@journal{jlpea}\gdef\@journalfull{Journal of Low Power Electronics and Applications}\gdef\@journalshort{J. Low Power Electron. Appl.}\gdef\@doiabbr{jlpea}\gdef\@ISSN{2079-9268}}
\DeclareOption{jmmp}{\gdef\@journal{jmmp}\gdef\@journalfull{Journal of Manufacturing and Materials Processing}\gdef\@journalshort{J. Manuf. Mater. Process.}\gdef\@doiabbr{jmmp}\gdef\@ISSN{2504-4494}}
\DeclareOption{jmp}{\gdef\@journal{jmp}\gdef\@journalfull{Journal of Molecular Pathology}\gdef\@journalshort{J. Mol. Pathol.}\gdef\@doiabbr{jmp}\gdef\@ISSN{2673-5261}\gdef\@continuouspages{yes}}
\DeclareOption{jmse}{\gdef\@journal{jmse}\gdef\@journalfull{Journal of Marine Science and Engineering}\gdef\@journalshort{J. Mar. Sci. Eng.}\gdef\@doiabbr{jmse}\gdef\@ISSN{2077-1312}}
\DeclareOption{jne}{\gdef\@journal{jne}\gdef\@journalfull{Journal of Nuclear Engineering}\gdef\@journalshort{J. Nucl. Eng.}\gdef\@doiabbr{jne}\gdef\@ISSN{2673-4362}\gdef\@continuouspages{yes}}
\DeclareOption{jnt}{\gdef\@journal{jnt}\gdef\@journalfull{Journal of Nanotheranostics}\gdef\@journalshort{J. Nanotheranostics}\gdef\@doiabbr{jnt}\gdef\@ISSN{2624-845X}\gdef\@continuouspages{yes}}
\DeclareOption{jof}{\gdef\@journal{jof}\gdef\@journalfull{Journal of Fungi}\gdef\@journalshort{J. Fungi}\gdef\@doiabbr{jof}\gdef\@ISSN{2309-608X}}
\DeclareOption{joitmc}{\gdef\@journal{joitmc}\gdef\@journalfull{Journal of Open Innovation: Technology Market and Complexity}\gdef\@journalshort{J. Open Innov. Technol. Mark. Complex.}\gdef\@doiabbr{joitmc}\gdef\@ISSN{2199-8531}}
\DeclareOption{jor}{\gdef\@journal{jor}\gdef\@journalfull{Journal of Respiration}\gdef\@journalshort{J. Respir.}\gdef\@doiabbr{jor}\gdef\@ISSN{2673-527X}\gdef\@continuouspages{yes}}
\DeclareOption{journalmedia}{\gdef\@journal{journalmedia}\gdef\@journalfull{Journalism and Media}\gdef\@journalshort{Journal. Media}\gdef\@doiabbr{journalmedia}\gdef\@ISSN{2673-5172}\gdef\@continuouspages{yes}}
\DeclareOption{jox}{\gdef\@journal{jox}\gdef\@journalfull{Journal of Xenobiotics}\gdef\@journalshort{J. Xenobiot.}\gdef\@doiabbr{jox}\gdef\@ISSN{2039-4713}\gdef\@continuouspages{yes}}
\DeclareOption{jpm}{\gdef\@journal{jpm}\gdef\@journalfull{Journal of Personalized Medicine}\gdef\@journalshort{J. Pers. Med.}\gdef\@doiabbr{jpm}\gdef\@ISSN{2075-4426}}
\DeclareOption{jrfm}{\gdef\@journal{jrfm}\gdef\@journalfull{Journal of Risk and Financial Management}\gdef\@journalshort{J. Risk Financial Manag.}\gdef\@doiabbr{jrfm}\gdef\@ISSN{1911-8074}}
\DeclareOption{jsan}{\gdef\@journal{jsan}\gdef\@journalfull{Journal of Sensor and Actuator Networks}\gdef\@journalshort{J. Sens. Actuator Netw.}\gdef\@doiabbr{jsan}\gdef\@ISSN{2224-2708}}
\DeclareOption{jtaer}{\gdef\@journal{jtaer}\gdef\@journalfull{Journal of Theoretical and Applied Electronic Commerce Research}\gdef\@journalshort{J. Theor. Appl. Electron. Commer. Res.}\gdef\@doiabbr{jtaer}\gdef\@ISSN{0718-1876}\gdef\@continuouspages{yes}}
\DeclareOption{jvd}{\gdef\@journal{jvd}\gdef\@journalfull{Journal of Vascular Diseases}\gdef\@journalshort{JVD}\gdef\@doiabbr{jvd}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{jzbg}{\gdef\@journal{jzbg}\gdef\@journalfull{Journal of Zoological and Botanical Gardens}\gdef\@journalshort{J. Zool. Bot. Gard.}\gdef\@doiabbr{jzbg}\gdef\@ISSN{2673-5636}\gdef\@continuouspages{yes}}
%\DeclareOption{kidney}{\gdef\@journal{kidney}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{kidneydial}{\gdef\@journal{kidneydial}\gdef\@journalfull{Kidney and Dialysis}\gdef\@journalshort{Kidney Dial.}\gdef\@doiabbr{kinasesphosphatases}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{kinasesphosphatases}{\gdef\@journal{kinasesphosphatases}\gdef\@journalfull{Kinases and Phosphatases}\gdef\@journalshort{Kinases Phosphatases}\gdef\@doiabbr{kinasesphosphatases}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{knowledge}{\gdef\@journal{knowledge}\gdef\@journalfull{Knowledge}\gdef\@journalshort{Knowledge}\gdef\@doiabbr{knowledge}\gdef\@ISSN{2673-9585}\gdef\@continuouspages{yes}}
\DeclareOption{land}{\gdef\@journal{land}\gdef\@journalfull{Land}\gdef\@journalshort{Land}\gdef\@doiabbr{land}\gdef\@ISSN{2073-445X}}
\DeclareOption{languages}{\gdef\@journal{languages}\gdef\@journalfull{Languages}\gdef\@journalshort{Languages}\gdef\@doiabbr{languages}\gdef\@ISSN{2226-471X}}
\DeclareOption{laws}{\gdef\@journal{laws}\gdef\@journalfull{Laws}\gdef\@journalshort{Laws}\gdef\@doiabbr{laws}\gdef\@ISSN{2075-471X}}
\DeclareOption{life}{\gdef\@journal{life}\gdef\@journalfull{Life}\gdef\@journalshort{Life}\gdef\@doiabbr{life}\gdef\@ISSN{2075-1729}}
\DeclareOption{liquids}{\gdef\@journal{liquids}\gdef\@journalfull{Liquids}\gdef\@journalshort{Liquids}\gdef\@doiabbr{liquids}\gdef\@ISSN{2673-8015}\gdef\@continuouspages{yes}}
\DeclareOption{literature}{\gdef\@journal{literature}\gdef\@journalfull{Literature}\gdef\@journalshort{Literature}\gdef\@doiabbr{literature}\gdef\@ISSN{2410-9789}\gdef\@continuouspages{yes}}
\DeclareOption{livers}{\gdef\@journal{livers}\gdef\@journalfull{Livers}\gdef\@journalshort{Livers}\gdef\@doiabbr{livers}\gdef\@ISSN{2673-4389}\gdef\@continuouspages{yes}}
\DeclareOption{logics}{\gdef\@journal{logics}\gdef\@journalfull{Logics}\gdef\@journalshort{Logics}\gdef\@doiabbr{logics}\gdef\@ISSN{2813-0405}\gdef\@continuouspages{yes}}
\DeclareOption{logistics}{\gdef\@journal{logistics}\gdef\@journalfull{Logistics}\gdef\@journalshort{Logistics}\gdef\@doiabbr{logistics}\gdef\@ISSN{2305-6290}}
\DeclareOption{lubricants}{\gdef\@journal{lubricants}\gdef\@journalfull{Lubricants}\gdef\@journalshort{Lubricants}\gdef\@doiabbr{lubricants}\gdef\@ISSN{2075-4442}}
\DeclareOption{lymphatics}{\gdef\@journal{lymphatics}\gdef\@journalfull{Lymphatics}\gdef\@journalshort{Lymphatics}\gdef\@doiabbr{lymphatics}\gdef\@ISSN{}\gdef\@continuouspages{yes}}	
\DeclareOption{machines}{\gdef\@journal{machines}\gdef\@journalfull{Machines}\gdef\@journalshort{Machines}\gdef\@doiabbr{machines}\gdef\@ISSN{2075-1702}}
\DeclareOption{macromol}{\gdef\@journal{macromol}\gdef\@journalfull{Macromol}\gdef\@journalshort{Macromol}\gdef\@doiabbr{macromol}\gdef\@ISSN{2673-6209}\gdef\@continuouspages{yes}}
\DeclareOption{magnetism}{\gdef\@journal{magnetism}\gdef\@journalfull{Magnetism}\gdef\@journalshort{Magnetism}\gdef\@doiabbr{magnetism}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{magnetochemistry}{\gdef\@journal{magnetochemistry}\gdef\@journalfull{Magnetochemistry}\gdef\@journalshort{Magnetochemistry}\gdef\@doiabbr{magnetochemistry}\gdef\@ISSN{2312-7481}}
\DeclareOption{make}{\gdef\@journal{make}\gdef\@journalfull{Machine Learning and Knowledge Extraction}\gdef\@journalshort{Mach. Learn. Knowl. Extr.}\gdef\@doiabbr{make}\gdef\@ISSN{2504-4990}\gdef\@continuouspages{yes}}
\DeclareOption{marinedrugs}{\gdef\@journal{marinedrugs}\gdef\@journalfull{Marine Drugs}\gdef\@journalshort{Mar. Drugs}\gdef\@doiabbr{md}\gdef\@ISSN{1660-3397}}
\DeclareOption{materials}{\gdef\@journal{materials}\gdef\@journalfull{Materials}\gdef\@journalshort{Materials}\gdef\@doiabbr{ma}\gdef\@ISSN{1996-1944}}
\DeclareOption{materproc}{\gdef\@journal{materproc}\gdef\@journalfull{Materials Proceedings}\gdef\@journalshort{Mater. Proc.}\gdef\@doiabbr{materproc}\gdef\@ISSN{2673-4605}}
\DeclareOption{mathematics}{\gdef\@journal{mathematics}\gdef\@journalfull{Mathematics}\gdef\@journalshort{Mathematics}\gdef\@doiabbr{math}\gdef\@ISSN{2227-7390}}
\DeclareOption{mca}{\gdef\@journal{mca}\gdef\@journalfull{Mathematical and Computational Applications}\gdef\@journalshort{Math. Comput. Appl.}\gdef\@doiabbr{mca}\gdef\@ISSN{2297-8747}}
\DeclareOption{measurements}{\gdef\@journal{measurements}\gdef\@journalfull{Measurements}\gdef\@journalshort{Measurements}\gdef\@doiabbr{measurements}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{medicina}{\gdef\@journal{medicina}\gdef\@journalfull{Medicina}\gdef\@journalshort{Medicina}\gdef\@doiabbr{medicina}\gdef\@ISSN{1648-9144}}
\DeclareOption{medicines}{\gdef\@journal{medicines}\gdef\@journalfull{Medicines}\gdef\@journalshort{Medicines}\gdef\@doiabbr{medicines}\gdef\@ISSN{2305-6320}}
\DeclareOption{medsci}{\gdef\@journal{medsci}\gdef\@journalfull{Medical Sciences}\gdef\@journalshort{Med. Sci.}\gdef\@doiabbr{medsci}\gdef\@ISSN{2076-3271}}
\DeclareOption{membranes}{\gdef\@journal{membranes}\gdef\@journalfull{Membranes}\gdef\@journalshort{Membranes}\gdef\@doiabbr{membranes}\gdef\@ISSN{2077-0375}}
\DeclareOption{merits}{\gdef\@journal{merits}\gdef\@journalfull{Merits}\gdef\@journalshort{Merits}\gdef\@doiabbr{merits}\gdef\@ISSN{2673-8104}\gdef\@continuouspages{yes}}
\DeclareOption{metabolites}{\gdef\@journal{metabolites}\gdef\@journalfull{Metabolites}\gdef\@journalshort{Metabolites}\gdef\@doiabbr{metabo}\gdef\@ISSN{2218-1989}}
\DeclareOption{metals}{\gdef\@journal{metals}\gdef\@journalfull{Metals}\gdef\@journalshort{Metals}\gdef\@doiabbr{met}\gdef\@ISSN{2075-4701}}
\DeclareOption{meteorology}{\gdef\@journal{meteorology}\gdef\@journalfull{Meteorology}\gdef\@journalshort{Meteorology}\gdef\@doiabbr{meteorology}\gdef\@ISSN{2674-0494}\gdef\@continuouspages{yes}}
\DeclareOption{methane}{\gdef\@journal{methane}\gdef\@journalfull{Methane}\gdef\@journalshort{Methane}\gdef\@doiabbr{methane}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{metrology}{\gdef\@journal{metrology}\gdef\@journalfull{Metrology}\gdef\@journalshort{Metrology}\gdef\@doiabbr{metrology}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{micro}{\gdef\@journal{micro}\gdef\@journalfull{Micro}\gdef\@journalshort{Micro}\gdef\@doiabbr{micro}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{microarrays}{\gdef\@journal{microarrays}\gdef\@journalfull{Microarrays}\gdef\@journalshort{Microarrays}\gdef\@doiabbr{microarrays}\gdef\@ISSN{2076-3905}}
\DeclareOption{microbiolres}{\gdef\@journal{microbiolres}\gdef\@journalfull{Microbiology Research}\gdef\@journalshort{Microbiol. Res.}\gdef\@doiabbr{microbiolres}\gdef\@ISSN{2036-7481}\gdef\@continuouspages{yes}}
\DeclareOption{micromachines}{\gdef\@journal{micromachines}\gdef\@journalfull{Micromachines}\gdef\@journalshort{Micromachines}\gdef\@doiabbr{mi}\gdef\@ISSN{2072-666X}}
\DeclareOption{microorganisms}{\gdef\@journal{microorganisms}\gdef\@journalfull{Microorganisms}\gdef\@journalshort{Microorganisms}\gdef\@doiabbr{microorganisms}\gdef\@ISSN{2076-2607}}
\DeclareOption{microplastics}{\gdef\@journal{microplastics}\gdef\@journalfull{Microplastics}\gdef\@journalshort{Microplastics}\gdef\@doiabbr{microplastics}\gdef\@ISSN{2673-8929}\gdef\@continuouspages{yes}}
\DeclareOption{minerals}{\gdef\@journal{minerals}\gdef\@journalfull{Minerals}\gdef\@journalshort{Minerals}\gdef\@doiabbr{min}\gdef\@ISSN{2075-163X}}
\DeclareOption{mining}{\gdef\@journal{mining}\gdef\@journalfull{Mining}\gdef\@journalshort{Mining}\gdef\@doiabbr{mining}\gdef\@ISSN{2673-6489}\gdef\@continuouspages{yes}}
\DeclareOption{modelling}{\gdef\@journal{modelling}\gdef\@journalfull{Modelling}\gdef\@journalshort{Modelling}\gdef\@doiabbr{modelling}\gdef\@ISSN{2673-3951}\gdef\@continuouspages{yes}}
\DeclareOption{molbank}{\gdef\@journal{molbank}\gdef\@journalfull{Molbank}\gdef\@journalshort{Molbank}\gdef\@doiabbr{M}\gdef\@ISSN{1422-8599}}
\DeclareOption{molecules}{\gdef\@journal{molecules}\gdef\@journalfull{Molecules}\gdef\@journalshort{Molecules}\gdef\@doiabbr{molecules}\gdef\@ISSN{1420-3049}}
\DeclareOption{mps}{\gdef\@journal{mps}\gdef\@journalfull{Methods and Protocols}\gdef\@journalshort{Methods Protoc.}\gdef\@doiabbr{mps}\gdef\@ISSN{2409-9279}}
\DeclareOption{msf}{\gdef\@journal{msf}\gdef\@journalfull{Medical Sciences Forum}\gdef\@journalshort{Med. Sci. Forum}\gdef\@doiabbr{msf}\gdef\@ISSN{2673-9992}}
\DeclareOption{mti}{\gdef\@journal{mti}\gdef\@journalfull{Multimodal Technologies and Interaction}\gdef\@journalshort{Multimodal Technol. Interact.}\gdef\@doiabbr{mti}\gdef\@ISSN{2414-4088}}
\DeclareOption{muscles}{\gdef\@journal{muscles}\gdef\@journalfull{Muscles}\gdef\@journalshort{Muscles}\gdef\@doiabbr{muscles}\gdef\@ISSN{2813-0413}\gdef\@continuouspages{yes}}
\DeclareOption{nanoenergyadv}{\gdef\@journal{nanoenergyadv}\gdef\@journalfull{Nanoenergy Advances}\gdef\@journalshort{Nanoenergy Adv.}\gdef\@doiabbr{nanoenergyadv}\gdef\@ISSN{2673-706X}\gdef\@continuouspages{yes}}
\DeclareOption{nanomanufacturing}{\gdef\@journal{nanomanufacturing}\gdef\@journalfull{Nanomanufacturing}\gdef\@journalshort{Nanomanufacturing}\gdef\@doiabbr{nanomanufacturing}\gdef\@ISSN{2673-687X}\gdef\@continuouspages{yes}}
\DeclareOption{nanomaterials}{\gdef\@journal{nanomaterials}\gdef\@journalfull{Nanomaterials}\gdef\@journalshort{Nanomaterials}\gdef\@doiabbr{nano}\gdef\@ISSN{2079-4991}}
\DeclareOption{ncrna}{\gdef\@journal{ncrna}\gdef\@journalfull{Non-Coding RNA}\gdef\@journalshort{Non-Coding RNA}\gdef\@doiabbr{ncrna}\gdef\@ISSN{2311-553X}}
\DeclareOption{network}{\gdef\@journal{network}\gdef\@journalfull{Network}\gdef\@journalshort{Network}\gdef\@doiabbr{network}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{neuroglia}{\gdef\@journal{neuroglia}\gdef\@journalfull{Neuroglia}\gdef\@journalshort{Neuroglia}\gdef\@doiabbr{neuroglia}\gdef\@ISSN{2571-6980}\gdef\@continuouspages{yes}}
\DeclareOption{neurolint}{\gdef\@journal{neurolint}\gdef\@journalfull{Neurology International}\gdef\@journalshort{Neurol. Int.}\gdef\@doiabbr{neurolint}\gdef\@ISSN{2035-8377}\gdef\@continuouspages{yes}}
\DeclareOption{neurosci}{\gdef\@journal{neurosci}\gdef\@journalfull{NeuroSci}\gdef\@journalshort{NeuroSci}\gdef\@doiabbr{neurosci}\gdef\@ISSN{2673-4087}\gdef\@continuouspages{yes}}
\DeclareOption{nitrogen}{\gdef\@journal{nitrogen}\gdef\@journalfull{Nitrogen}\gdef\@journalshort{Nitrogen}\gdef\@doiabbr{nitrogen}\gdef\@ISSN{2504-3129}\gdef\@continuouspages{yes}}
\DeclareOption{notspecified}{\gdef\@journal{notspecified}\gdef\@journalfull{Journal Not Specified}\gdef\@journalshort{Journal Not Specified}\gdef\@doiabbr{}\gdef\@ISSN{}}
%%\DeclareOption{nri}{\gdef\@journal{nri}\gdef\@journalfull{Natural Resources Informatics}\gdef\@journalshort{Nat. Resour. Inform.}\gdef\@doiabbr{nri}\gdef\@ISSN{2673-981X}\gdef\@continuouspages{yes}}
\DeclareOption{nursrep}{\gdef\@journal{nursrep}\gdef\@journalfull{Nursing Reports}\gdef\@journalshort{Nurs. Rep.}\gdef\@doiabbr{nursrep}\gdef\@ISSN{2039-4403}\gdef\@continuouspages{yes}}
\DeclareOption{nutraceuticals}{\gdef\@journal{nutraceuticals}\gdef\@journalfull{Nutraceuticals}\gdef\@journalshort{Nutraceuticals}\gdef\@doiabbr{nutraceuticals}\gdef\@ISSN{1661-3821}\gdef\@continuouspages{yes}}
\DeclareOption{nutrients}{\gdef\@journal{nutrients}\gdef\@journalfull{Nutrients}\gdef\@journalshort{Nutrients}\gdef\@doiabbr{nu}\gdef\@ISSN{2072-6643}}
\DeclareOption{obesities}{\gdef\@journal{obesities}\gdef\@journalfull{Obesities}\gdef\@journalshort{Obesities}\gdef\@doiabbr{obesities}\gdef\@ISSN{2673-4168}\gdef\@continuouspages{yes}}
\DeclareOption{oceans}{\gdef\@journal{oceans}\gdef\@journalfull{Oceans}\gdef\@journalshort{Oceans}\gdef\@doiabbr{oceans}\gdef\@ISSN{2673-1924}\gdef\@continuouspages{yes}}
\DeclareOption{ohbm}{\gdef\@journal{ohbm}\gdef\@journalfull{Journal of Otorhinolaryngology Hearing and Balance Medicine}\gdef\@journalshort{J. Otorhinolaryngol. Hear. Balance Med.}\gdef\@doiabbr{ohbm}\gdef\@ISSN{2504-463X}}
\DeclareOption{onco}{\gdef\@journal{onco}\gdef\@journalfull{Onco}\gdef\@journalshort{Onco}\gdef\@doiabbr{onco}\gdef\@ISSN{2673-7523}\gdef\@continuouspages{yes}}
%\DeclareOption{oncopathology}{\gdef\@journal{oncopathology}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{optics}{\gdef\@journal{optics}\gdef\@journalfull{Optics}\gdef\@journalshort{Optics}\gdef\@doiabbr{opt}\gdef\@ISSN{2673-3269}\gdef\@continuouspages{yes}}
\DeclareOption{oral}{\gdef\@journal{oral}\gdef\@journalfull{Oral}\gdef\@journalshort{Oral}\gdef\@doiabbr{oral}\gdef\@ISSN{2673-6373}\gdef\@continuouspages{yes}}
\DeclareOption{organics}{\gdef\@journal{organics}\gdef\@journalfull{Organics}\gdef\@journalshort{Organics}\gdef\@doiabbr{org}\gdef\@ISSN{2673-401X}\gdef\@continuouspages{yes}}
\DeclareOption{organoids}{\gdef\@journal{organoids}\gdef\@journalfull{Organoids}\gdef\@journalshort{Organoids}\gdef\@doiabbr{organoids}\gdef\@ISSN{2674-1172}\gdef\@continuouspages{yes}}
\DeclareOption{osteology}{\gdef\@journal{osteology}\gdef\@journalfull{Osteology}\gdef\@journalshort{Osteology}\gdef\@doiabbr{osteology}\gdef\@ISSN{2673-4036}\gdef\@continuouspages{yes}}
\DeclareOption{oxygen}{\gdef\@journal{oxygen}\gdef\@journalfull{Oxygen}\gdef\@journalshort{Oxygen}\gdef\@doiabbr{oxygen}\gdef\@ISSN{2673-9801}\gdef\@continuouspages{yes}}
\DeclareOption{parasites}{\gdef\@journal{parasites}\gdef\@journalfull{Parasites}\gdef\@journalshort{Parasites}\gdef\@doiabbr{parasites}\gdef\@ISSN{2673-5946}}
\DeclareOption{parasitologia}{\gdef\@journal{parasitologia}\gdef\@journalfull{Parasitologia}\gdef\@journalshort{Parasitologia}\gdef\@doiabbr{parasitologia}\gdef\@ISSN{2673-6772}\gdef\@continuouspages{yes}}
\DeclareOption{particles}{\gdef\@journal{particles}\gdef\@journalfull{Particles}\gdef\@journalshort{Particles}\gdef\@doiabbr{particles}\gdef\@ISSN{2571-712X}\gdef\@continuouspages{yes}}
\DeclareOption{pathogens}{\gdef\@journal{pathogens}\gdef\@journalfull{Pathogens}\gdef\@journalshort{Pathogens}\gdef\@doiabbr{pathogens}\gdef\@ISSN{2076-0817}}
\DeclareOption{pathophysiology}{\gdef\@journal{pathophysiology}\gdef\@journalfull{Pathophysiology}\gdef\@journalshort{Pathophysiology}\gdef\@doiabbr{pathophysiology}\gdef\@ISSN{0928-4680}\gdef\@continuouspages{yes}}
\DeclareOption{pediatrrep}{\gdef\@journal{pediatrrep}\gdef\@journalfull{Pediatric Reports}\gdef\@journalshort{Pediatr. Rep.}\gdef\@doiabbr{pediatric}\gdef\@ISSN{2036-7503}\gdef\@continuouspages{yes}}
\DeclareOption{pharmaceuticals}{\gdef\@journal{pharmaceuticals}\gdef\@journalfull{Pharmaceuticals}\gdef\@journalshort{Pharmaceuticals}\gdef\@doiabbr{ph}\gdef\@ISSN{1424-8247}}
\DeclareOption{pharmaceutics}{\gdef\@journal{pharmaceutics}\gdef\@journalfull{Pharmaceutics}\gdef\@journalshort{Pharmaceutics}\gdef\@doiabbr{pharmaceutics}\gdef\@ISSN{1999-4923}}
\DeclareOption{pharmacoepidemiology}{\gdef\@journal{pharmacoepidemiology}\gdef\@journalfull{Pharmacoepidemiology}\gdef\@journalshort{Pharmacoepidemiology}\gdef\@doiabbr{pharma}\gdef\@ISSN{2813-0618}\gdef\@continuouspages{yes}}
\DeclareOption{pharmacy}{\gdef\@journal{pharmacy}\gdef\@journalfull{Pharmacy}\gdef\@journalshort{Pharmacy}\gdef\@doiabbr{pharmacy}\gdef\@ISSN{2226-4787}}
\DeclareOption{philosophies}{\gdef\@journal{philosophies}\gdef\@journalfull{Philosophies}\gdef\@journalshort{Philosophies}\gdef\@doiabbr{philosophies}\gdef\@ISSN{2409-9287}}
\DeclareOption{photochem}{\gdef\@journal{photochem}\gdef\@journalfull{Photochem}\gdef\@journalshort{Photochem}\gdef\@doiabbr{photochem}\gdef\@ISSN{2673-7256}\gdef\@continuouspages{yes}}
\DeclareOption{photonics}{\gdef\@journal{photonics}\gdef\@journalfull{Photonics}\gdef\@journalshort{Photonics}\gdef\@doiabbr{photonics}\gdef\@ISSN{2304-6732}}
\DeclareOption{phycology}{\gdef\@journal{phycology}\gdef\@journalfull{Phycology}\gdef\@journalshort{Phycology}\gdef\@doiabbr{phycology}\gdef\@ISSN{2673-9410}\gdef\@continuouspages{yes}}
\DeclareOption{physchem}{\gdef\@journal{physchem}\gdef\@journalfull{Physchem}\gdef\@journalshort{Physchem}\gdef\@doiabbr{physchem}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{physics}{\gdef\@journal{physics}\gdef\@journalfull{Physics}\gdef\@journalshort{Physics}\gdef\@doiabbr{physics}\gdef\@ISSN{2624-8174}\gdef\@continuouspages{yes}}
\DeclareOption{physiologia}{\gdef\@journal{physiologia}\gdef\@journalfull{Physiologia}\gdef\@journalshort{Physiologia}\gdef\@doiabbr{physiologia}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{plants}{\gdef\@journal{plants}\gdef\@journalfull{Plants}\gdef\@journalshort{Plants}\gdef\@doiabbr{plants}\gdef\@ISSN{2223-7747}}
\DeclareOption{plasma}{\gdef\@journal{plasma}\gdef\@journalfull{Plasma}\gdef\@journalshort{Plasma}\gdef\@doiabbr{plasma}\gdef\@ISSN{2571-6182}\gdef\@continuouspages{yes}}
\DeclareOption{pollutants}{\gdef\@journal{pollutants}\gdef\@journalfull{Pollutants}\gdef\@journalshort{Pollutants}\gdef\@doiabbr{pollutants}\gdef\@ISSN{2673-4672}\gdef\@continuouspages{yes}}
\DeclareOption{polymers}{\gdef\@journal{polymers}\gdef\@journalfull{Polymers}\gdef\@journalshort{Polymers}\gdef\@doiabbr{polym}\gdef\@ISSN{2073-4360}}
\DeclareOption{polysaccharides}{\gdef\@journal{polysaccharides}\gdef\@journalfull{Polysaccharides}\gdef\@journalshort{Polysaccharides}\gdef\@doiabbr{polysaccharides}\gdef\@ISSN{2673-4176}\gdef\@continuouspages{yes}}
\DeclareOption{poultry}{\gdef\@journal{poultry}\gdef\@journalfull{Poultry}\gdef\@journalshort{Poultry}\gdef\@doiabbr{poultry}\gdef\@ISSN{2674-1164}\gdef\@continuouspages{yes}}
\DeclareOption{powders}{\gdef\@journal{powders}\gdef\@journalfull{Powders}\gdef\@journalshort{Powders}\gdef\@doiabbr{powders}\gdef\@ISSN{2674-0516}\gdef\@continuouspages{yes}}
\DeclareOption{preprints}{\gdef\@journal{preprints}\gdef\@journalfull{Preprints}\gdef\@journalshort{Preprints}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{proceedings}{\gdef\@journal{proceedings}\gdef\@journalfull{Proceedings}\gdef\@journalshort{Proceedings}\gdef\@doiabbr{proceedings}\gdef\@ISSN{2504-3900}}
\DeclareOption{processes}{\gdef\@journal{processes}\gdef\@journalfull{Processes}\gdef\@journalshort{Processes}\gdef\@doiabbr{pr}\gdef\@ISSN{2227-9717}}
\DeclareOption{prosthesis}{\gdef\@journal{prosthesis}\gdef\@journalfull{Prosthesis}\gdef\@journalshort{Prosthesis}\gdef\@doiabbr{prosthesis}\gdef\@ISSN{2673-1592}\gdef\@continuouspages{yes}}
\DeclareOption{proteomes}{\gdef\@journal{proteomes}\gdef\@journalfull{Proteomes}\gdef\@journalshort{Proteomes}\gdef\@doiabbr{proteomes}\gdef\@ISSN{2227-7382}}
\DeclareOption{psf}{\gdef\@journal{psf}\gdef\@journalfull{Physical Sciences Forum}\gdef\@journalshort{Phys. Sci. Forum}\gdef\@doiabbr{psf}\gdef\@ISSN{2673-9984}}
\DeclareOption{psych}{\gdef\@journal{psych}\gdef\@journalfull{Psych}\gdef\@journalshort{Psych}\gdef\@doiabbr{psych}\gdef\@ISSN{2624-8611}\gdef\@continuouspages{yes}}
\DeclareOption{psychiatryint}{\gdef\@journal{psychiatryint}\gdef\@journalfull{Psychiatry International}\gdef\@journalshort{Psychiatry Int.}\gdef\@doiabbr{psychiatryint}\gdef\@ISSN{2673-5318}\gdef\@continuouspages{yes}}
\DeclareOption{psychoactives}{\gdef\@journal{psychoactives}\gdef\@journalfull{Psychoactives}\gdef\@journalshort{Psychoactives}\gdef\@doiabbr{psychoactives}\gdef\@ISSN{2813-1851}\gdef\@continuouspages{yes}}
\DeclareOption{publications}{\gdef\@journal{publications}\gdef\@journalfull{Publications}\gdef\@journalshort{Publications}\gdef\@doiabbr{publications}\gdef\@ISSN{2304-6775}}
\DeclareOption{quantumrep}{\gdef\@journal{quantumrep}\gdef\@journalfull{Quantum Reports}\gdef\@journalshort{Quantum Rep.}\gdef\@doiabbr{quantum}\gdef\@ISSN{2624-960X}\gdef\@continuouspages{yes}}
\DeclareOption{quaternary}{\gdef\@journal{quaternary}\gdef\@journalfull{Quaternary}\gdef\@journalshort{Quaternary}\gdef\@doiabbr{quat}\gdef\@ISSN{2571-550X}}
\DeclareOption{qubs}{\gdef\@journal{qubs}\gdef\@journalfull{Quantum Beam Science}\gdef\@journalshort{Quantum Beam Sci.}\gdef\@doiabbr{qubs}\gdef\@ISSN{2412-382X}}
\DeclareOption{radiation}{\gdef\@journal{radiation}\gdef\@journalfull{Radiation}\gdef\@journalshort{Radiation}\gdef\@doiabbr{radiation}\gdef\@ISSN{2673-592X}\gdef\@continuouspages{yes}}
\DeclareOption{reactions}{\gdef\@journal{reactions}\gdef\@journalfull{Reactions}\gdef\@journalshort{Reactions}\gdef\@doiabbr{reactions}\gdef\@ISSN{2624-781X}\gdef\@continuouspages{yes}}
\DeclareOption{receptors}{\gdef\@journal{receptors}\gdef\@journalfull{Receptors}\gdef\@journalshort{Receptors}\gdef\@doiabbr{receptors}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{recycling}{\gdef\@journal{recycling}\gdef\@journalfull{Recycling}\gdef\@journalshort{Recycling}\gdef\@doiabbr{recycling}\gdef\@ISSN{2313-4321}}
\DeclareOption{regeneration}{\gdef\@journal{regeneration}\gdef\@journalfull{Regeneration}\gdef\@journalshort{Regeneration}\gdef\@doiabbr{regeneration}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{religions}{\gdef\@journal{religions}\gdef\@journalfull{Religions}\gdef\@journalshort{Religions}\gdef\@doiabbr{rel}\gdef\@ISSN{2077-1444}}
\DeclareOption{remotesensing}{\gdef\@journal{remotesensing}\gdef\@journalfull{Remote Sensing}\gdef\@journalshort{Remote Sens.}\gdef\@doiabbr{rs}\gdef\@ISSN{2072-4292}}
\DeclareOption{reports}{\gdef\@journal{reports}\gdef\@journalfull{Reports}\gdef\@journalshort{Reports}\gdef\@doiabbr{reports}\gdef\@ISSN{2571-841X}}
\DeclareOption{reprodmed}{\gdef\@journal{reprodmed}\gdef\@journalfull{Reproductive Medicine}\gdef\@journalshort{Reprod. Med.}\gdef\@doiabbr{reprodmed}\gdef\@ISSN{2673-3897}\gdef\@continuouspages{yes}}
\DeclareOption{resources}{\gdef\@journal{resources}\gdef\@journalfull{Resources}\gdef\@journalshort{Resources}\gdef\@doiabbr{resources}\gdef\@ISSN{2079-9276}}
\DeclareOption{rheumato}{\gdef\@journal{rheumato}\gdef\@journalfull{Rheumato}\gdef\@journalshort{Rheumato}\gdef\@doiabbr{rheumato}\gdef\@ISSN{2674-0621}\gdef\@continuouspages{yes}}
\DeclareOption{risks}{\gdef\@journal{risks}\gdef\@journalfull{Risks}\gdef\@journalshort{Risks}\gdef\@doiabbr{risks}\gdef\@ISSN{2227-9091}}
\DeclareOption{robotics}{\gdef\@journal{robotics}\gdef\@journalfull{Robotics}\gdef\@journalshort{Robotics}\gdef\@doiabbr{robotics}\gdef\@ISSN{2218-6581}}
\DeclareOption{ruminants}{\gdef\@journal{ruminants}\gdef\@journalfull{Ruminants}\gdef\@journalshort{Ruminants}\gdef\@doiabbr{ruminants}\gdef\@ISSN{2673-933X}\gdef\@continuouspages{yes}}
\DeclareOption{safety}{\gdef\@journal{safety}\gdef\@journalfull{Safety}\gdef\@journalshort{Safety}\gdef\@doiabbr{safety}\gdef\@ISSN{2313-576X}}
\DeclareOption{sci}{\gdef\@journal{sci}\gdef\@journalfull{Sci}\gdef\@journalshort{Sci}\gdef\@doiabbr{sci}\gdef\@ISSN{2413-4155}}
\DeclareOption{scipharm}{\gdef\@journal{scipharm}\gdef\@journalfull{Scientia Pharmaceutica}\gdef\@journalshort{Sci. Pharm.}\gdef\@doiabbr{scipharm}\gdef\@ISSN{2218-0532}}
\DeclareOption{seeds}{\gdef\@journal{seeds}\gdef\@journalfull{Seeds}\gdef\@journalshort{Seeds}\gdef\@doiabbr{seeds}\gdef\@ISSN{2674-1024}\gdef\@continuouspages{yes}}
\DeclareOption{sensors}{\gdef\@journal{sensors}\gdef\@journalfull{Sensors}\gdef\@journalshort{Sensors}\gdef\@doiabbr{s}\gdef\@ISSN{1424-8220}}
\DeclareOption{separations}{\gdef\@journal{separations}\gdef\@journalfull{Separations}\gdef\@journalshort{Separations}\gdef\@doiabbr{separations}\gdef\@ISSN{2297-8739}}
\DeclareOption{sexes}{\gdef\@journal{sexes}\gdef\@journalfull{Sexes}\gdef\@journalshort{Sexes}\gdef\@doiabbr{sexes}\gdef\@ISSN{2411-5118}\gdef\@continuouspages{yes}}
\DeclareOption{signals}{\gdef\@journal{signals}\gdef\@journalfull{Signals}\gdef\@journalshort{Signals}\gdef\@doiabbr{signals}\gdef\@ISSN{2624-6120}\gdef\@continuouspages{yes}}
\DeclareOption{sinusitis}{\gdef\@journal{sinusitis}\gdef\@journalfull{Sinusitis}\gdef\@journalshort{Sinusitis}\gdef\@doiabbr{sinusitis}\gdef\@ISSN{2309-107X}\gdef\@continuouspages{yes}}
\DeclareOption{skins}{\gdef\@journal{skins}\gdef\@journalfull{Skins}\gdef\@journalshort{Skins}\gdef\@doiabbr{skins}\gdef\@ISSN{}}
\DeclareOption{smartcities}{\gdef\@journal{smartcities}\gdef\@journalfull{Smart Cities}\gdef\@journalshort{Smart Cities}\gdef\@doiabbr{smartcities}\gdef\@ISSN{2624-6511}\gdef\@continuouspages{yes}}
\DeclareOption{sna}{\gdef\@journal{sna}\gdef\@journalfull{Sinusitis and Asthma}\gdef\@journalshort{Sinusitis Asthma}\gdef\@doiabbr{sna}\gdef\@ISSN{2624-7003}}
\DeclareOption{societies}{\gdef\@journal{societies}\gdef\@journalfull{Societies}\gdef\@journalshort{Societies}\gdef\@doiabbr{soc}\gdef\@ISSN{2075-4698}}
\DeclareOption{socsci}{\gdef\@journal{socsci}\gdef\@journalfull{Social Sciences}\gdef\@journalshort{Soc. Sci.}\gdef\@doiabbr{socsci}\gdef\@ISSN{2076-0760}}
\DeclareOption{software}{\gdef\@journal{software}\gdef\@journalfull{Software}\gdef\@journalshort{Software}\gdef\@doiabbr{software}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{soilsystems}{\gdef\@journal{soilsystems}\gdef\@journalfull{Soil Systems}\gdef\@journalshort{Soil Syst.}\gdef\@doiabbr{soilsystems}\gdef\@ISSN{2571-8789}}
\DeclareOption{solar}{\gdef\@journal{solar}\gdef\@journalfull{Solar}\gdef\@journalshort{Solar}\gdef\@doiabbr{solar}\gdef\@ISSN{2673-9941}\gdef\@continuouspages{yes}}
\DeclareOption{solids}{\gdef\@journal{solids}\gdef\@journalfull{Solids}\gdef\@journalshort{Solids}\gdef\@doiabbr{solids}\gdef\@ISSN{2673-6497}\gdef\@continuouspages{yes}}
\DeclareOption{spectroscj}{\gdef\@journal{spectroscj}\gdef\@journalfull{Spectroscopy Journal}\gdef\@journalshort{Spectrosc. J.}\gdef\@doiabbr{spectroscj}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{sports}{\gdef\@journal{sports}\gdef\@journalfull{Sports}\gdef\@journalshort{Sports}\gdef\@doiabbr{sports}\gdef\@ISSN{2075-4663}}
\DeclareOption{standards}{\gdef\@journal{standards}\gdef\@journalfull{Standards}\gdef\@journalshort{Standards}\gdef\@doiabbr{standards}\gdef\@ISSN{2305-6703}\gdef\@continuouspages{yes}}
\DeclareOption{stats}{\gdef\@journal{stats}\gdef\@journalfull{Stats}\gdef\@journalshort{Stats}\gdef\@doiabbr{stats}\gdef\@ISSN{2571-905X}\gdef\@continuouspages{yes}}
\DeclareOption{std}{\gdef\@journal{std}\gdef\@journalfull{Surgical Techniques Development}\gdef\@journalshort{Surg. Tech. Dev.}\gdef\@doiabbr{std}\gdef\@ISSN{2038-9582}\gdef\@continuouspages{yes}}
\DeclareOption{stresses}{\gdef\@journal{stresses}\gdef\@journalfull{Stresses}\gdef\@journalshort{Stresses}\gdef\@doiabbr{stresses}\gdef\@ISSN{2673-7140}\gdef\@continuouspages{yes}}
\DeclareOption{surfaces}{\gdef\@journal{surfaces}\gdef\@journalfull{Surfaces}\gdef\@journalshort{Surfaces}\gdef\@doiabbr{surfaces}\gdef\@ISSN{2571-9637}\gdef\@continuouspages{yes}}
\DeclareOption{surgeries}{\gdef\@journal{surgeries}\gdef\@journalfull{Surgeries}\gdef\@journalshort{Surgeries}\gdef\@doiabbr{surgeries}\gdef\@ISSN{2017-2017}\gdef\@continuouspages{yes}}
\DeclareOption{suschem}{\gdef\@journal{suschem}\gdef\@journalfull{Sustainable Chemistry}\gdef\@journalshort{Sustain. Chem.}\gdef\@doiabbr{suschem}\gdef\@ISSN{2673-4079}\gdef\@continuouspages{yes}}
\DeclareOption{sustainability}{\gdef\@journal{sustainability}\gdef\@journalfull{Sustainability}\gdef\@journalshort{Sustainability}\gdef\@doiabbr{su}\gdef\@ISSN{2071-1050}}
\DeclareOption{symmetry}{\gdef\@journal{symmetry}\gdef\@journalfull{Symmetry}\gdef\@journalshort{Symmetry}\gdef\@doiabbr{sym}\gdef\@ISSN{2073-8994}}
\DeclareOption{synbio}{\gdef\@journal{synbio}\gdef\@journalfull{Synbio}\gdef\@journalshort{Synbio}\gdef\@doiabbr{synbio}\gdef\@ISSN{2674-0583}\gdef\@continuouspages{yes}}
\DeclareOption{systems}{\gdef\@journal{systems}\gdef\@journalfull{Systems}\gdef\@journalshort{Systems}\gdef\@doiabbr{systems}\gdef\@ISSN{2079-8954}}
\DeclareOption{targets}{\gdef\@journal{targets}\gdef\@journalfull{Targets}\gdef\@journalshort{Targets}\gdef\@doiabbr{targets}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{taxonomy}{\gdef\@journal{taxonomy}\gdef\@journalfull{Taxonomy}\gdef\@journalshort{Taxonomy}\gdef\@doiabbr{taxonomy}\gdef\@ISSN{2673-6500}\gdef\@continuouspages{yes}}
\DeclareOption{technologies}{\gdef\@journal{technologies}\gdef\@journalfull{Technologies}\gdef\@journalshort{Technologies}\gdef\@doiabbr{technologies}\gdef\@ISSN{2227-7080}}
\DeclareOption{telecom}{\gdef\@journal{telecom}\gdef\@journalfull{Telecom}\gdef\@journalshort{Telecom}\gdef\@doiabbr{telecom}\gdef\@ISSN{2673-4001}\gdef\@continuouspages{yes}}
\DeclareOption{test}{\gdef\@journal{test}\gdef\@journalfull{Test}\gdef\@journalshort{Test}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{textiles}{\gdef\@journal{textiles}\gdef\@journalfull{Textiles}\gdef\@journalshort{Textiles}\gdef\@doiabbr{textiles}\gdef\@ISSN{2673-7248}\gdef\@continuouspages{yes}}
\DeclareOption{thalassrep}{\gdef\@journal{thalassrep}\gdef\@journalfull{Thalassemia Reports}\gdef\@journalshort{Thalass. Rep.}\gdef\@doiabbr{thalassrep}\gdef\@ISSN{2039-4365}\gdef\@continuouspages{yes}}
\DeclareOption{thermo}{\gdef\@journal{thermo}\gdef\@journalfull{Thermo}\gdef\@journalshort{Thermo}\gdef\@doiabbr{thermo}\gdef\@ISSN{2673-7264}\gdef\@continuouspages{yes}}
\DeclareOption{tomography}{\gdef\@journal{tomography}\gdef\@journalfull{Tomography}\gdef\@journalshort{Tomography}\gdef\@doiabbr{tomography}\gdef\@ISSN{2379-139X}\gdef\@continuouspages{yes}}
\DeclareOption{tourismhosp}{\gdef\@journal{tourismhosp}\gdef\@journalfull{Tourism and Hospitality}\gdef\@journalshort{Tour. Hosp.}\gdef\@doiabbr{tourhosp}\gdef\@ISSN{2673-5768}\gdef\@continuouspages{yes}}
\DeclareOption{toxics}{\gdef\@journal{toxics}\gdef\@journalfull{Toxics}\gdef\@journalshort{Toxics}\gdef\@doiabbr{toxics}\gdef\@ISSN{2305-6304}}
\DeclareOption{toxins}{\gdef\@journal{toxins}\gdef\@journalfull{Toxins}\gdef\@journalshort{Toxins}\gdef\@doiabbr{toxins}\gdef\@ISSN{2072-6651}}
\DeclareOption{transplantology}{\gdef\@journal{transplantology}\gdef\@journalfull{Transplantology}\gdef\@journalshort{Transplantology}\gdef\@doiabbr{transplantology}\gdef\@ISSN{2673-3943}\gdef\@continuouspages{yes}}
\DeclareOption{transportation}{\gdef\@journal{transportation}\gdef\@journalfull{Transportation}\gdef\@journalshort{Transportation}\gdef\@doiabbr{transportation}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{traumacare}{\gdef\@journal{traumacare}\gdef\@journalfull{Trauma Care}\gdef\@journalshort{Trauma Care}\gdef\@doiabbr{traumacare}\gdef\@ISSN{2673-866X}\gdef\@continuouspages{yes}}
\DeclareOption{traumas}{\gdef\@journal{traumas}\gdef\@journalfull{Traumas}\gdef\@journalshort{Traumas}\gdef\@doiabbr{traumas}\gdef\@ISSN{2673-7477}\gdef\@continuouspages{yes}}
\DeclareOption{tropicalmed}{\gdef\@journal{tropicalmed}\gdef\@journalfull{Tropical Medicine and Infectious Disease}\gdef\@journalshort{Trop. Med. Infect. Dis.}\gdef\@doiabbr{tropicalmed}\gdef\@ISSN{2414-6366}}
\DeclareOption{universe}{\gdef\@journal{universe}\gdef\@journalfull{Universe}\gdef\@journalshort{Universe}\gdef\@doiabbr{universe}\gdef\@ISSN{2218-1997}}
\DeclareOption{urbansci}{\gdef\@journal{urbansci}\gdef\@journalfull{Urban Science}\gdef\@journalshort{Urban Sci.}\gdef\@doiabbr{urbansci}\gdef\@ISSN{2413-8851}}
\DeclareOption{uro}{\gdef\@journal{uro}\gdef\@journalfull{Uro}\gdef\@journalshort{Uro}\gdef\@doiabbr{uro}\gdef\@ISSN{2673-4397}\gdef\@continuouspages{yes}}
\DeclareOption{vaccines}{\gdef\@journal{vaccines}\gdef\@journalfull{Vaccines}\gdef\@journalshort{Vaccines}\gdef\@doiabbr{vaccines}\gdef\@ISSN{2076-393X}}
\DeclareOption{vehicles}{\gdef\@journal{vehicles}\gdef\@journalfull{Vehicles}\gdef\@journalshort{Vehicles}\gdef\@doiabbr{vehicles}\gdef\@ISSN{2624-8921}\gdef\@continuouspages{yes}}
\DeclareOption{venereology}{\gdef\@journal{venereology}\gdef\@journalfull{Venereology}\gdef\@journalshort{Venereology}\gdef\@doiabbr{venereology}\gdef\@ISSN{2674-0710}\gdef\@continuouspages{yes}}
\DeclareOption{vetsci}{\gdef\@journal{vetsci}\gdef\@journalfull{Veterinary Sciences}\gdef\@journalshort{Vet. Sci.}\gdef\@doiabbr{vetsci}\gdef\@ISSN{2306-7381}}
\DeclareOption{vibration}{\gdef\@journal{vibration}\gdef\@journalfull{Vibration}\gdef\@journalshort{Vibration}\gdef\@doiabbr{vibration}\gdef\@ISSN{2571-631X}\gdef\@continuouspages{yes}}
\DeclareOption{virtualworlds}{\gdef\@journal{virtualworlds}\gdef\@journalfull{Virtual Worlds}\gdef\@journalshort{Virtual Worlds}\gdef\@doiabbr{virtualworlds}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{viruses}{\gdef\@journal{viruses}\gdef\@journalfull{Viruses}\gdef\@journalshort{Viruses}\gdef\@doiabbr{v}\gdef\@ISSN{1999-4915}}
\DeclareOption{vision}{\gdef\@journal{vision}\gdef\@journalfull{Vision}\gdef\@journalshort{Vision}\gdef\@doiabbr{vision}\gdef\@ISSN{2411-5150}}
\DeclareOption{waste}{\gdef\@journal{waste}\gdef\@journalfull{Waste}\gdef\@journalshort{Waste}\gdef\@doiabbr{waste}\gdef\@ISSN{2813-0391}\gdef\@continuouspages{yes}}
\DeclareOption{water}{\gdef\@journal{water}\gdef\@journalfull{Water}\gdef\@journalshort{Water}\gdef\@doiabbr{w}\gdef\@ISSN{2073-4441}}
\DeclareOption{wem}{\gdef\@journal{wem}\gdef\@journalfull{Wildlife Ecology and Management}\gdef\@journalshort{Wildl. Ecol. Manag.}\gdef\@doiabbr{}\gdef\@ISSN{1234-4321}}
\DeclareOption{wevj}{\gdef\@journal{wevj}\gdef\@journalfull{World Electric Vehicle Journal}\gdef\@journalshort{World Electr. Veh. J.}\gdef\@doiabbr{wevj}\gdef\@ISSN{2032-6653}}
\DeclareOption{wind}{\gdef\@journal{wind}\gdef\@journalfull{Wind}\gdef\@journalshort{Wind}\gdef\@doiabbr{wind}\gdef\@ISSN{2674-032X}\gdef\@continuouspages{yes}}
\DeclareOption{women}{\gdef\@journal{women}\gdef\@journalfull{Women}\gdef\@journalshort{Women}\gdef\@doiabbr{women}\gdef\@ISSN{2673-4184}\gdef\@continuouspages{yes}}
\DeclareOption{world}{\gdef\@journal{world}\gdef\@journalfull{World}\gdef\@journalshort{World}\gdef\@doiabbr{world}\gdef\@ISSN{2673-4060}\gdef\@continuouspages{yes}}
\DeclareOption{youth}{\gdef\@journal{youth}\gdef\@journalfull{Youth}\gdef\@journalshort{Youth}\gdef\@doiabbr{youth}\gdef\@ISSN{2673-995X}\gdef\@continuouspages{yes}}
\DeclareOption{zoonoticdis}{\gdef\@journal{zoonoticdis}\gdef\@journalfull{Zoonotic Diseases}\gdef\@journalshort{Zoonotic Dis.}\gdef\@doiabbr{zoonoticdis}\gdef\@ISSN{2813-0227}\gdef\@continuouspages{yes}}

### 2022-10-04 manuscripts/owp_3dep_2022/template.tex
added mdpi template

%  LaTeX support: latex@mdpi.com 
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[journal,article,submit,pdftex,moreauthors]{Definitions/mdpi} 
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept". The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, alloys, analytica, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, currophthalmol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, ent, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, futureinternet, futureparasites, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, kidneydial, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by: 
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
% MDPI internal commands
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2022}
\copyrightyear{2022}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{} 
%\daterevised{} % Only for the journal Acoustics
\dateaccepted{} 
\datepublished{} 
%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{2,\ddagger}$ and Firstname Lastname $^{2,}$*}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3.} 
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data
%\dataset{DOI number or link to the deposited data set if the data set is published separately. If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors. In this case, please submit the data set as a supplement.}
%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}

The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions on the authors' page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX-related questions please contact latex@mdpi.com.%\endnote{This is an endnote.} % To use endnotes, please un-comment \printendnotes below (before References). Only journal Laws uses \footnote.

% The order of the section titles is different for some journals. Please refer to the "Instructions for Authors” on the journal homepage.

\section{Introduction}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
\begin{quote}
This is an example of a quote.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\subsection{Subsection}
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}
\item	First bullet;
\item	Second bullet;
\item	Third bullet.
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}
\item	First item; 
\item	Second item;
\item	Third item.
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
\end{figure}   
\unskip

\begin{table}[H] 
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCC}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabularx}
\end{table}
\unskip

\begin{table}[H]
\caption{This is a wide table.\label{tab2}}
	\begin{adjustwidth}{-\extralength}{0cm}
		\newcolumntype{C}{>{\centering\arraybackslash}X}
		\begin{tabularx}{\fulllength}{CCCC}
			\toprule
			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
			\midrule
			Entry 1		& Data			& Data			& Data\\
			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
			\bottomrule
		\end{tabularx}
	\end{adjustwidth}
	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
\end{table}

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\columnwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\columnwidth}{1pt}
%\end{listing}

Text.

Text.

\subsection{Formatting of Mathematical Components}

This is the example 1 of equation:
\begin{linenomath}
\begin{equation}
a = 1,
\end{equation}
\end{linenomath}
the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

This is the example 2 of equation:
\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

% Example of a page in landscape format (with table and table footnote).
%\startlandscape
%\begin{table}[H] %% Table in wide page
%\caption{This is a very wide table.\label{tab3}}
%	\begin{tabularx}{\textwidth}{CCCC}
%		\toprule
%		\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}	& \textbf{Title 4}\\
%		\midrule
%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
%		\bottomrule
%	\end{tabularx}
%	\begin{adjustwidth}{+\extralength}{0cm}
%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
%	\end{adjustwidth}
%\end{table}
%\finishlandscape

% Example of a figure that spans the whole page width. The same concept works for tables, too.
\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=13.5cm]{Definitions/logo-mdpi}
\end{adjustwidth}
\caption{This is a wide figure.\label{fig2}}
\end{figure}

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}

This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.}

\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}

\informedconsent{Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans.

Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}

\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}. If the study did not report any data, you might add ``Not applicable'' here.}

\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the~results''.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\sampleavailability{Samples of the compounds ... are available from the authors.}

%% Only for journal Encyclopedia
%\entrylink{The Link to this entry published on the encyclopedia platform.}

\abbreviations{Abbreviations}{
The following abbreviations are used in this manuscript:\\

\noindent 
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & Linear dichroism
\end{tabular}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}
\subsection[\appendixname~\thesubsection]{}
The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data. Mathematical proofs of results not central to the paper can be added as an appendix.

\begin{table}[H] 
\caption{This is a table caption.\label{tab5}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCC}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabularx}
\end{table}

\section[\appendixname~\thesection]{}
All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}
%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\bibliography{your_external_BibTeX_file}

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
\bibitem[Author2(year)]{ref-book1}
Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
% Reference 3
\bibitem[Author3(year)]{ref-book2}
Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
% Reference 4
\bibitem[Author4(year)]{ref-unpublish}
Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
% Reference 5
\bibitem[Author5(year)]{ref-communication}
Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
% Reference 6
\bibitem[Author6(year)]{ref-proceeding}
Author 1, A.B.; Author 2, C.D.; Author 3, E.F. Title of presentation. In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
% Reference 7
\bibitem[Author7(year)]{ref-thesis}
Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
% Reference 8
\bibitem[Author8(year)]{ref-url}
Title of Site. Available online: URL (accessed on Day Month Year).
\end{thebibliography}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{adjustwidth}
\end{document}

### 2022-10-05 manuscripts/owp_3dep_2022/owp_3dep.tex
updated state

@@
              postcode={35401}, 
-             state={VA},
+             state={AL},
              country={USA}

### 2022-10-05 manuscripts/owp_3dep_2022/owp_3dep.tex
moved to MDPI

@@
-%% 
-%% Copyright 2007-2020 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for Elsevier's document class `elsarticle'
-%% with harvard style bibliographic references
+%  LaTeX support: latex@mdpi.com 
+%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
 
-\documentclass[preprint,review,12pt]{dependencies/elsarticle}
+%=================================================================
+\documentclass[water,article,submit,pdftex,moreauthors]{dependencies/Definitions/mdpi} 
+% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
 
-%% Use the option review to obtain double line spacing
-%% \documentclass[preprint,review,12pt]{elsarticle}
+%--------------------
+% Class Options:
+%--------------------
+%----------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, alloys, analytica, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, currophthalmol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, ent, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, futureinternet, futureparasites, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, kidneydial, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis
 
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn]{elsarticle}
-%% \documentclass[final,3p,times]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn]{elsarticle}
-%% \documentclass[final,5p,times]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn]{elsarticle}
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by: 
+% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
+% supfile = supplementary materials
 
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
 
-%% The amssymb package provides various useful mathematical symbols
-\usepackage{amssymb}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
 
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-\usepackage{lineno}
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+%=================================================================
+% MDPI internal commands
+\firstpage{1} 
+\makeatletter 
+\setcounter{page}{\@firstpage} 
+\makeatother
+\pubvolume{1}
+\issuenum{1}
+\articlenumber{0}
+\pubyear{2022}
+\copyrightyear{2022}
+%\externaleditor{Academic Editor: Firstname Lastname}
+\datereceived{} 
+%\daterevised{} % Only for the journal Acoustics
+\dateaccepted{} 
+\datepublished{} 
+%\datecorrected{} % Corrected papers include a "Corrected: XXX" date in the original paper.
+%\dateretracted{} % Corrected papers include a "Retracted: XXX" date in the original paper.
+\hreflink{https://doi.org/} % If needed use \linebreak
+%\doinum{}
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
 %
-%% Additional packages
 \usepackage[printonlyused]{acronym}
@@
 \usepackage{url}
+\usepackage{amssymb}
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\journal{Advances in Water Resources}
-%
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{On the Effects of Varying Spatial Resolutions of 3DEP Data on Flood Inundation Extents Produced from Height Above Nearest Drainage}
+
+% MDPI internal command: Title for citation in the left column
+\TitleCitation{Title}
+
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{0000-0003-2525-4712} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0002-5430-6000} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorC}{0000-0003-2939-4114} % Add \orcidC{} behind the author's name
+\newcommand{\orcidauthorD}{0000-0001-9247-821X} % Add \orcidD{} behind the author's name
+\newcommand{\orcidauthorE}{0000-0001-9849-7411} % Add \orcidE{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Fernando Aristizabal $^{1,2,3,}$\orcidA{}*, Taher Chegini $^{4,}$\orcidB, Fernando Salas $^{2,}$\orcidC, Gregory Petrochenkov $^{5,}$\orcidD and Jasmeet Judge $^{3,}$\orcidE}
+
+%\longauthorlist{yes}
+
+% MDPI internal command: Authors, for metadata in PDF
+\AuthorNames{Fernando Aristizabal, Taher Chegini, Gregory Petrochenkov, Fernando Salas, and Jasmeet Judge}
+
+% MDPI internal command: Authors, for citation in the left column
+\AuthorCitation{Aristizabal, F.; Chegini, T.; Petrochenkov, G.; Salas, F.; Judge, J.}
+% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ \quad Lynker, 338 E Market St, Leesburg, VA, 20176, USA  \\
+$^{2}$ \quad National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration, 205 Hackberry Ln, Tuscaloosa, AL, 35401, USA \\
+$^{3}$ \quad Center for Remote Sensing, Agricultural and Biological Engineering, University of Florida, 1741 Museum Rd, Gainesville, FL, 32603, USA \\
+$^{4}$ \quad Civil and Environmental Engineering, University of Houston, 4226 Martin Luther King Boulevard, Houston, TX, 77204, USA \\
+$^{5}$ \quad Hydrologic Applied Innovations Lab, New York Water Science Center, United States Geological Survey, 425 Jordan Rd, Troy, NY, 12180, USA}
+
+% Contact information of the corresponding author
+\corres{Correspondence: fernando.aristizabal@noaa.gov}
+
+% Current address and/or shared authorship
+%\firstnote{Current address: Affiliation 3.} 
+%\secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{A single paragraph of about 200 words maximum.
+For research articles, abstracts should give a pertinent overview of the work.
+We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations.
+The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}
+
+% Keywords
+\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
+
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+\PACS{92.40.Qk}  % surface water, water resources, water supply, floods, lakes, rivers, streamflow,
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% FRONT MATTER
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{frontmatter}
-%
-%% Title, authors and addresses
-%
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%% \fntext[label3]{}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% TITLE 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{comment}
+\setcounter{section}{-1} %% Remove this when starting to work on the template.
+\section{How to Use this Template}
 
-\title{On the Effects of Varying Spatial Resolutions of 3DEP Data on Flood Inundation Extents Produced from Height Above Nearest Drainage}
+The template details the sections that can be used in a manuscript.
+Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section).
+Please check the instructions on the authors' page of the journal to verify the correct order and names.
+For any questions, please contact the editorial office of the journal or support@mdpi.com.
+For LaTeX-related questions please contact latex@mdpi.com.%\endnote{This is an endnote.} % To use endnotes, please un-comment \printendnotes below (before References).
+Only journal Laws uses \footnote.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% AUTHORS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% The order of the section titles is different for some journals. Please refer to the "Instructions for Authors” on the journal homepage.
+\end{comment}
 
-\author[lynk,nwc,uf]{Fernando Aristizabal\corref{fa_corr}}
-\ead{fernando.aristizabal@noaa.gov}
-\ead[url]{https://www.linkedin.com/in/fernando-aristizabal}
-%\fntext[fa_corr]{Please direct correspondence to Fernando Aristizabal.}
-\cortext[fa_corr]{Please direct correspondence to Fernando Aristizabal.}
-
-\author[nwc]{Fernando Salas}
-\author[uh]{Taher Chegini}
-\author[usgs]{Gregory Petrochenkov}
-\author[uf]{Jasmeet Judge}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% AFFILIATIONS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\affiliation[lynk]{
-             organization={Lynker},
-             addressline={338 E Market St}, 
-             city={Leesburg},
-             postcode={20176}, 
-             state={VA},
-             country={USA}
-            }
-
-\affiliation[nwc]{
-             organization={National Water Center, Office of Water Prediction, National Oceanic and Atmospheric Administration},
-             addressline={205 Hackberry Ln}, 
-             city={Tuscaloosa},
-             postcode={35401}, 
-             state={AL},
-             country={USA}
-            }
-%
-\affiliation[uf]{
-             organization={Center for Remote Sensing, Agricultural and Biological Engineering, University of Florida},
-             addressline={1741 Museum Rd}, 
-             city={Gainesville},
-             postcode={32603}, 
-             state={FL},
-             country={USA}
-            }
-%
-\affiliation[uh]{
-             organization={Civil and Environmental Engineering, University of Houston},
-             addressline={4226 Martin Luther King Boulevard}, 
-             city={Houston},
-             postcode={77204}, 
-             state={TX},
-             country={USA}
-            }
-%
-\affiliation[usgs]{
-             organization={Hydrologic Applied Innovations Lab, New York Water Science Center, United States Geological Survey},
-             addressline={425 Jordan Rd}, 
-             city={Troy},
-             postcode={12180}, 
-             state={NY},
-             country={USA}
-            }
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ABSTRACT
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{abstract}
-\end{abstract}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Graphical abstract
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{graphicalabstract}
-%\includegraphics{grabs}
-%\end{graphicalabstract}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Research highlights
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{highlights}
-\item HAND approximates flood inundation for large scale, high resolution, and operational applications.
-\item Utilizing LiDAR derived data from the USGS 3DEP program enhances the skill on flood inundation maps produced from HAND.
-\item Varying the spatial resolution provides no skill enhancement when evaluated at large scales across three study sites and benchmark datasets.
-\end{highlights}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% KEYWORDS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% MAX SIX WORDS %%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\begin{keyword}
-%% PACS codes here, in the form: \PACS code \sep code
-Hydrology, surface water, water resources, water supply, floods, lakes, rivers, streamflow, natural and man-made disasters, weather analysis and prediction, topography
-
-\PACS
-92.40.-t \sep % hydrology, cryosphere
-92.40.Qk \sep % surface water, water resources, water supply, floods, lakes, rivers, streamflow,
-*92.40.Q- \sep % surface water
-*92.40.qp \sep % streamflow, floods
-89.60.Gg \sep % disasters (natural and man-made),
-92.60.Wc \sep % weather analysis and prediction
-91.10.Jf      % topography (earth)
-%
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-%
-\end{keyword}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{frontmatter}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% MAIN TEXT
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% prints line numbers
-\linenumbers
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% INTRODUCTION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 %%%% Motivates and introduces flooding %%%%
-Floods are among the most frequent, damaging, and deadly of natural disasters \citep{doocy2013human,stromberg2007natural,kahn2005death}. 
-The frequency and intensity of flood events as well as the exposure of people and property to them have been increasing in recent times driven by secular changes in climate, infrastructure, and demographics \citep{berz2000flood,mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
-Unfortunately, these trends are expected to continue placing additional pressure on hydrological extremes \citep{kahn2005death,tabari2020climate,milly2002increasing,wing2018estimates}.
-Floods impact mortality and morbidity through drowning or physical trauma at the individual health scale, while increasing the risk of infectious disease at the public health level \citep{jonkman2005global,beinin2012medical,alajo2006cholera,french1983mortality}.
-Flooding disrupts systems providing human needs such as transportation routes, supply chains, water delivery, waste management, communications, and energy grids \citep{wijkman2021natural}.
-These impacts disproportionately affect certain demographics such as the socioeconomically-disadvantaged, youth, and elderly who are more likely to live in vulnerable areas with less access to educational resources, \acp{EWS}, and the capacity or resources to evacuate impacted areas \citep{kahn2005death,smiley2022social,stromberg2007natural,jonkman2005global,tellman2020using,tellman2021satellite}.
-These inequitable impacts further entrench poverty and inequalities \citep{stallings1988conflict,birkmann2010extreme}.
-In political terms, severe disasters, including floods, can reduce social order, strain governance systems, collapse social safety nets, increase the risk of social conflict \citep{drury1998disasters,xu2016natural,zahran2009natural}.
+Floods are among the most frequent, damaging, and deadly of natural disasters \cite{doocy2013human,stromberg2007natural,kahn2005death}. 
+The frequency and intensity of flood events as well as the exposure of people and property to them have been increasing in recent times driven by secular changes in climate, infrastructure, and demographics \cite{berz2000flood,mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
+Unfortunately, these trends are expected to continue placing additional pressure on hydrological extremes \cite{kahn2005death,tabari2020climate,milly2002increasing,wing2018estimates}.
+Floods impact mortality and morbidity through drowning or physical trauma at the individual health scale, while increasing the risk of infectious disease at the public health level \cite{jonkman2005global,beinin2012medical,alajo2006cholera,french1983mortality}.
+Flooding disrupts systems providing human needs such as transportation routes, supply chains, water delivery, waste management, communications, and energy grids \cite{wijkman2021natural}.
+These impacts disproportionately affect certain demographics such as the socioeconomically-disadvantaged, youth, and elderly who are more likely to live in vulnerable areas with less access to educational resources, \acp{EWS}, and the capacity or resources to evacuate impacted areas \cite{kahn2005death,smiley2022social,stromberg2007natural,jonkman2005global,tellman2020using,tellman2021satellite}.
+These inequitable impacts further entrench poverty and inequalities \cite{stallings1988conflict,birkmann2010extreme}.
+In political terms, severe disasters, including floods, can reduce social order, strain governance systems, collapse social safety nets, increase the risk of social conflict \cite{drury1998disasters,xu2016natural,zahran2009natural}.
 These dire consequences motivate adaption and mitigation efforts such as \acp{EWS}, protective infrastructure (e.g.
-storage, defenses, drainage, infiltration), public awareness and education, and zoning regulations \citep{tumbare2000mitigating,tauhid2018mitigating,charlesworth201115}.
+storage, defenses, drainage, infiltration), public awareness and education, and zoning regulations \cite{tumbare2000mitigating,tauhid2018mitigating,charlesworth201115}.
 
 %%%% Motivates and introduces NWM %%%%
-Due to the growing consequences and risks presented by increasing flood impacts, \acp{EWS}, or forecasting systems, can help understand future conditions and provide intelligence to furnish adequate warnings to protect life, prevent damages, and enhance resilience \citep{stromberg2007natural,cools2016lessons,unisdr2015making,baudoin2014early,golnaraghi2012overview,unep2012early,liu2018review}.
+Due to the growing consequences and risks presented by increasing flood impacts, \acp{EWS}, or forecasting systems, can help understand future conditions and provide intelligence to furnish adequate warnings to protect life, prevent damages, and enhance resilience \cite{stromberg2007natural,cools2016lessons,unisdr2015making,baudoin2014early,golnaraghi2012overview,unep2012early,liu2018review}.
 The early warning of flood disasters at national scales requires the use of continental-scale, forecast hydrology models and modeling frameworks that span intranational political boundaries.
 The applications of these models extend beyond \acp{EWS} to provide historical trends for applications in infrastructure planning, public planning, insurance underwriting, and more.
-The \ac{OWP}, an office of the \ac{NOAA} along with partners at the \ac{NCAR}, developed such a continental-scale model known as the \ac{US} \ac{NWM} \citep{salas2018towards,gochis2021wrf,cosgrove2019evolution,cohen2018featured,noaa2016national,water2022nwm}.
-The \ac{NWM} is based on a configuration of the \ac{WRF-Hydro} model that accounts for land surface processes as well as overland and channel routing \citep{gochis2021wrf,salas2018towards,cosgrove2019evolution}.
-Operationally, the \ac{NWM} produces streamflow analysis and forecasts at multiple time horizons depending on location which include the \ac{CONUS}, \ac{PR}, and \ac{HI} \citep{cosgrove2019evolution,noaa2016national,water2022nwm}.
-The \ac{NWM} routes streamflow across the \ac{NWM} \ac{V2.1} stream network, based on the \ac{NHDPlusV2} network, is comprised of more than 5.5 million \acp{km} of lines discretized into more than 2.8 million forecast points \citep{aristizabal2022extending}.
-The \ac{NWM} \ac{V2.1} stream network belongs to the \ac{NWM} ``hydrofabric'' defined as a catalog of geospatial layers relevant to hydrology modeling including stream network lines, catchments, reservoirs, and more \citep{water2022nwm,cosgrove2019evolution}.
+The \ac{OWP}, an office of the \ac{NOAA} along with partners at the \ac{NCAR}, developed such a continental-scale model known as the \ac{US} \ac{NWM} \cite{salas2018towards,gochis2021wrf,cosgrove2019evolution,cohen2018featured,noaa2016national,water2022nwm}.
+The \ac{NWM} is based on a configuration of the \ac{WRF-Hydro} model that accounts for land surface processes as well as overland and channel routing \cite{gochis2021wrf,salas2018towards,cosgrove2019evolution}.
+Operationally, the \ac{NWM} produces streamflow analysis and forecasts at multiple time horizons depending on location which include the \ac{CONUS}, \ac{PR}, and \ac{HI} \cite{cosgrove2019evolution,noaa2016national,water2022nwm}.
+The \ac{NWM} routes streamflow across the \ac{NWM} \ac{V2.1} stream network, based on the \ac{NHDPlusV2} network, is comprised of more than 5.5 million \acp{km} of lines discretized into more than 2.8 million forecast points \cite{aristizabal2022extending}.
+The \ac{NWM} \ac{V2.1} stream network belongs to the \ac{NWM} ``hydrofabric'' defined as a catalog of geospatial layers relevant to hydrology modeling including stream network lines, catchments, reservoirs, and more \cite{water2022nwm,cosgrove2019evolution}.
 While streamflow is an important variable for engineering and scientific applications of fluvial flooding, flood inundation stages, extents and depths are much more tangible variables to the stakeholders flood events directly impact.
@@
 Solving this system in full \ac{2D} form requires numerical methods that can be very cost prohibitive and numerically unstable in an operational setting across continental-scales at high spatial discretizations (10 \ac{m} or higher).
-This use case motivates the implementation of an inundation proxy, also known as a zero-physics or a simplified conceptual model, that is agnostic to the \acp{SWE} while still computing accurate fluvial inundation extents and depths for this problem \citep{teng2015rapid,bates2000simple}.
-\ac{HAND} detrends elevations within \acp{DEM} to compute drainage potentials by normalizing elevations to the nearest, relevant drainage line instead of datums that represent mean sea level \citep{renno2008hand,nobre2011height,nobre2016hand}.
-\ac{HAND} as a terrain index has been used extensively for \ac{FIM} purposes from both modeled or observed stream flows and stages \citep{nobre2016hand,afshari2018comparison,garousi2019terrain,johnson2019integrated,zheng2018geoflood,zheng2018river,zhang2018comparative,teng2015rapid,li2022accounting,li2020evaluation}, as well as for assisting the remote sensing detection of fluvial inundation \citep{aristizabal2020high,shastry2019using,aristizabal2021mapping,huang2017comparison,twele2016sentinel}.
-\ac{HAND} operates as an inundation proxy by thresholding the relative elevation (or \ac{HAND}) values with a singular river stage value for each catchment corresponding to the drainage area of a given river reach \citep{nobre2016hand,garousi2019terrain,johnson2019integrated,zheng2018geoflood,teng2015rapid,li2020evaluation,liu2016cybergis,maidment2017conceptual,liu2018cybergis,liu2020height,liu2018review}.
-When used to generate inundation extents and depths from streamflow, reach-averaged \ac{SRC} sample geometric variables along an entire reach and normalize using the length of the reach to create stage-discharge relationships \citep{zheng2018river,aristizabal2022extending,godbout2019error}.
+This use case motivates the implementation of an inundation proxy, also known as a zero-physics or a simplified conceptual model, that is agnostic to the \acp{SWE} while still computing accurate fluvial inundation extents and depths for this problem \cite{teng2015rapid,bates2000simple}.
+\ac{HAND} detrends elevations within \acp{DEM} to compute drainage potentials by normalizing elevations to the nearest, relevant drainage line instead of datums that represent mean sea level \cite{renno2008hand,nobre2011height,nobre2016hand}.
+\ac{HAND} as a terrain index has been used extensively for \ac{FIM} purposes from both modeled or observed stream flows and stages \cite{nobre2016hand,afshari2018comparison,garousi2019terrain,johnson2019integrated,zheng2018geoflood,zheng2018river,zhang2018comparative,teng2015rapid,li2022accounting,li2020evaluation}, as well as for assisting the remote sensing detection of fluvial inundation \cite{aristizabal2020high,shastry2019using,aristizabal2021mapping,huang2017comparison,twele2016sentinel}.
+\ac{HAND} operates as an inundation proxy by thresholding the relative elevation (or \ac{HAND}) values with a singular river stage value for each catchment corresponding to the drainage area of a given river reach \cite{nobre2016hand,garousi2019terrain,johnson2019integrated,zheng2018geoflood,teng2015rapid,li2020evaluation,liu2016cybergis,maidment2017conceptual,liu2018cybergis,liu2020height,liu2018review}.
+When used to generate inundation extents and depths from streamflow, reach-averaged \ac{SRC} sample geometric variables along an entire reach and normalize using the length of the reach to create stage-discharge relationships \cite{zheng2018river,aristizabal2022extending,godbout2019error}.
 These relationships depend on a friction parameter, Manning's n, and used to convert streamflows to stages for eventual \ac{2D} mapping with \ac{HAND}.
-Numerous investigations have validated the use of \ac{HAND} for flood mapping applications as a suitable alternative to more sophisticated physics-based techniques for large scale and high resolution use cases \citep{johnson2019integrated,li2020evaluation,li2022comparative,aristizabal2022extending,nobre2016hand,godbout2019error,afshari2018comparison,zhang2018comparative,teng2015rapid,teng2017flood,diehl2021improving,hocini2021performance,bates2003optimal}.
+Numerous investigations have validated the use of \ac{HAND} for flood mapping applications as a suitable alternative to more sophisticated physics-based techniques for large scale and high resolution use cases \cite{johnson2019integrated,li2020evaluation,li2022comparative,aristizabal2022extending,nobre2016hand,godbout2019error,afshari2018comparison,zhang2018comparative,teng2015rapid,teng2017flood,diehl2021improving,hocini2021performance,bates2003optimal}.
 
 %%%% Motivates and introduces OWP FIM %%%%
-Several prior and active large-scale \ac{HAND} implementations catered to operational \ac{EWS} applications including the \ac{NFIE} \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis}, GeoFlood \citep{zheng2018geoflood,hocini2021performance,d2022identification,carruthers2021assessment,zheng2022application}, and \ac{PyGFT}, i.e.
-\ac{GIS} \citep{petrochenkov2020pygft,verdin2016software}.
-The \ac{NFIE} was a broad, inter-institutional, and pioneering effort to apply HAND to the initial versions of the \ac{NWM} which leveraged 1/3 arc-second (10 \ac{m}) seamless elevation data available at the time \citep{maidment2017conceptual,liu2016cybergis,liu2018cybergis} from the \ac{USGS}'s \ac{NED} \citep{gesch2002national,gesch2007digital}.
-\citet{zheng2018geoflood} applied HAND for operational applications with 1/27 arc-second (1 \ac{m}) elevation data with a novel least cost, geodesic based stream delineation method \citep{passalacqua2010geometric,passalacqua2012automatic,zheng2018geoflood,zheng2019automatic,carruthers2021assessment,d2022identification,zheng2022application}.
-For applications with the \ac{NWM}, an advanced version of \ac{HAND} coupled with the use of \acp{SRC}, known as \ac{OWP} \ac{FIM}, converts \ac{NWM} analysis, reanalysis, and forecast streamflows to river stages and fluvial inundation depths and extents on an operational basis to \ac{CONUS} while extending the modeling domain to \ac{PR} and \ac{HI} \citep{aristizabal2022extending,inundationMapping2022}.
-\ac{OWP} \ac{FIM} utilizes some of the latest datasets including the \ac{NHDPlusHR} \citep{moore2019user}, \ac{NLD} \citep{engineers2016national}, and the \ac{NWM} \ac{V2.1} hydrofabric \citep{water2022nwm,noaa2016national,nwm2022hydrofabric,gochis2021wrf}.
-These datasets enforce hydrologically relevant features such as levees and the general location of stream lines to facilitate conflation with the forecast stream network \citep{aristizabal2022extending,inundationMapping2022}.
-Additionally, \ac{OWP} \ac{FIM} advanced a fundamental limitation of \ac{HAND} that limits sourcing fluvial inundation only from the nearest, relevant drainage line \citep{mcgehee2016modified,aristizabal2022extending,zhang2018comparative,li2022comparative,zheng2018geoflood,zheng2018river,nobre2016hand}.
-Stream lines of higher Horton-Strahler stream order that could contribute inundation to a given area have no way of extending beyond catchment lines which creates artificial bottlenecks in inundation extents especially along junctions of high order rivers with their lower flow tributaries \citep{aristizabal2022extending,mcgehee2016modified}.
-To resolve this limitation, \ac{OWP} \ac{FIM} disaggregates the \ac{NWM} \ac{V2.1} stream network into segments of effective unit stream order called \acp{LP} \citep{aristizabal2022extending}.
-In terms of terrain data, \ac{OWP} \ac{FIM} utilizes the 10 \ac{m} \ac{DEM} from the \ac{NHDPlusHR} elevation dataset which is the elevation basis, derived in batches from \ac{3DEP}, for additional hydrography products within the \ac{NHDPlusHR} \citep{aristizabal2022extending,moore2019user}.
-The previous advances in \ac{OWP} \ac{FIM} stopped short of accounting for \ac{LiDAR} point elevation observations \citep{aristizabal2022extending} that are now nearing their first collection cycle to form a novel seamless, continental scale \ac{DEM} from \ac{3DEP} \citep{usgs2022status,usgs2022partnerships}.
+Several prior and active large-scale \ac{HAND} implementations catered to operational \ac{EWS} applications including the \ac{NFIE} \cite{maidment2017conceptual,liu2016cybergis,liu2018cybergis}, GeoFlood \cite{zheng2018geoflood,hocini2021performance,d2022identification,carruthers2021assessment,zheng2022application}, and \ac{PyGFT}, i.e.
+\ac{GIS} \cite{petrochenkov2020pygft,verdin2016software}.
+The \ac{NFIE} was a broad, inter-institutional, and pioneering effort to apply HAND to the initial versions of the \ac{NWM} which leveraged 1/3 arc-second (10 \ac{m}) seamless elevation data available at the time \cite{maidment2017conceptual,liu2016cybergis,liu2018cybergis} from the \ac{USGS}'s \ac{NED} \cite{gesch2002national,gesch2007digital}.
+\citet{zheng2018geoflood} applied HAND for operational applications with 1/27 arc-second (1 \ac{m}) elevation data with a novel least cost, geodesic based stream delineation method \cite{passalacqua2010geometric,passalacqua2012automatic,zheng2018geoflood,zheng2019automatic,carruthers2021assessment,d2022identification,zheng2022application}.
+For applications with the \ac{NWM}, an advanced version of \ac{HAND} coupled with the use of \acp{SRC}, known as \ac{OWP} \ac{FIM}, converts \ac{NWM} analysis, reanalysis, and forecast streamflows to river stages and fluvial inundation depths and extents on an operational basis to \ac{CONUS} while extending the modeling domain to \ac{PR} and \ac{HI} \cite{aristizabal2022extending,inundationMapping2022}.
+\ac{OWP} \ac{FIM} utilizes some of the latest datasets including the \ac{NHDPlusHR} \cite{moore2019user}, \ac{NLD} \cite{engineers2016national}, and the \ac{NWM} \ac{V2.1} hydrofabric \cite{water2022nwm,noaa2016national,nwm2022hydrofabric,gochis2021wrf}.
+These datasets enforce hydrologically relevant features such as levees and the general location of stream lines to facilitate conflation with the forecast stream network \cite{aristizabal2022extending,inundationMapping2022}.
+Additionally, \ac{OWP} \ac{FIM} advanced a fundamental limitation of \ac{HAND} that limits sourcing fluvial inundation only from the nearest, relevant drainage line \cite{mcgehee2016modified,aristizabal2022extending,zhang2018comparative,li2022comparative,zheng2018geoflood,zheng2018river,nobre2016hand}.
+Stream lines of higher Horton-Strahler stream order that could contribute inundation to a given area have no way of extending beyond catchment lines which creates artificial bottlenecks in inundation extents especially along junctions of high order rivers with their lower flow tributaries \cite{aristizabal2022extending,mcgehee2016modified}.
+To resolve this limitation, \ac{OWP} \ac{FIM} disaggregates the \ac{NWM} \ac{V2.1} stream network into segments of effective unit stream order called \acp{LP} \cite{aristizabal2022extending}.
+In terms of terrain data, \ac{OWP} \ac{FIM} utilizes the 10 \ac{m} \ac{DEM} from the \ac{NHDPlusHR} elevation dataset which is the elevation basis, derived in batches from \ac{3DEP}, for additional hydrography products within the \ac{NHDPlusHR} \cite{aristizabal2022extending,moore2019user}.
+The previous advances in \ac{OWP} \ac{FIM} stopped short of accounting for \ac{LiDAR} point elevation observations \cite{aristizabal2022extending} that are now nearing their first collection cycle to form a novel seamless, continental scale \ac{DEM} from \ac{3DEP} \cite{usgs2022status,usgs2022partnerships}.
 
 %%%% 3DEP program introduction %%%%
-Broad scale terrain information in the form of \acp{DEM} is fundamental to all \ac{FIM} models and a significant influence on skill \citep{bales2009sources,dobbs2010evaluation,wang2005comparison,merwade2008uncertainty,witt2015evaluation}.
-The \ac{NGP}, under the \ac{USGS}, is the primary authority on collecting, processing, and maintaining terrestrial elevation data within the \ac{US} in collaboration with Federal partners within the \ac{NDEP} \citep{omb2016circularA16,dewberry2011final,national2007elevation,national2009mapping,sugarbaker20143d}.
-The \ac{NED} \citep{gesch2002national,gesch2007digital}, forms the seamless elevation layers of the \ac{TNM} \citep{gesch2009national,archuleta2017national,arundel2015preparing,arundel2018assimilation,kelmelis2003national}.
+Broad scale terrain information in the form of \acp{DEM} is fundamental to all \ac{FIM} models and a significant influence on skill \cite{bales2009sources,dobbs2010evaluation,wang2005comparison,merwade2008uncertainty,witt2015evaluation}.
+The \ac{NGP}, under the \ac{USGS}, is the primary authority on collecting, processing, and maintaining terrestrial elevation data within the \ac{US} in collaboration with Federal partners within the \ac{NDEP} \cite{omb2016circularA16,dewberry2011final,national2007elevation,national2009mapping,sugarbaker20143d}.
+The \ac{NED} \cite{gesch2002national,gesch2007digital}, forms the seamless elevation layers of the \ac{TNM} \cite{gesch2009national,archuleta2017national,arundel2015preparing,arundel2018assimilation,kelmelis2003national}.
 Prior to the introduction of \ac{3DEP}, \ac{TNM} was originally composed of three seamless \acp{DEM} at 1/3 (10 \ac{m}), 1 (30 \ac{m}), and 2 (90 \ac{m}) arc-second resolutions produced from a variety legacy sources including digital photogrammetry, cartographic contours, mapped hydrography, and elevations from \ac{SRTM} .
-\citep{gesch2002national,gesch2007digital,arundel2015preparing}.
-High quality elevations derived from \ac{LiDAR} and \ac{InSAR} have been integrated into \ac{TNM} seamless elevation products as made available prior to and after the introduction of \ac{3DEP} \citep{snyder2013national,gesch2002national,arundel2015preparing}.
-Work by \citet{gesch2014accuracy},\citep{gesch2007digital}, and \citet{dobbs2010evaluation} illustrated that the inclusion of higher quality elevation data sources had significant improvement in the accuracy of \ac{NED} data when compared to the \ac{NGS} \citep{roman2010geodesy}.
+\cite{gesch2002national,gesch2007digital,arundel2015preparing}.
+High quality elevations derived from \ac{LiDAR} and \ac{InSAR} have been integrated into \ac{TNM} seamless elevation products as made available prior to and after the introduction of \ac{3DEP} \cite{snyder2013national,gesch2002national,arundel2015preparing}.
+Work by \citet{gesch2014accuracy},\cite{gesch2007digital}, and \citet{dobbs2010evaluation} illustrated that the inclusion of higher quality elevation data sources had significant improvement in the accuracy of \ac{NED} data when compared to the \ac{NGS} \cite{roman2010geodesy}.
 \citet{gesch2014accuracy} identified that the \ac{NED} 1/3 arc-second \ac{DEM}, as of April 2013, had a mean error of -0.29 \ac{m} with an \ac{RMSE} of 1.55 \ac{m} when compared to over 25 thousand reference points.
-At the time of evaluation, the \ac{NED} was subject to legacy, lower quality data sources dating almost a century in the past \citep{sugarbaker20143d,gesch2014accuracy,gesch2007digital}.
-This reduction in error and its impact on people and commerce \citep{dewberry2011final} motivated action on collection of elevation data from higher quality data sources \citep{sugarbaker20143d}.
+At the time of evaluation, the \ac{NED} was subject to legacy, lower quality data sources dating almost a century in the past \cite{sugarbaker20143d,gesch2014accuracy,gesch2007digital}.
+This reduction in error and its impact on people and commerce \cite{dewberry2011final} motivated action on collection of elevation data from higher quality data sources \cite{sugarbaker20143d}.
 
@@
 %%% NEED LIDAR REFERENCES %%%%
-\ac{3DEP} is a national, multi-organizational effort by the \ac{NDEP} to survey elevations with high quality sensors in response to growing stakeholder needs on a recurring collection cycle of no greater than 8 \acp{year} \citep{dewberry2011final,snyder2013national,sugarbaker20143d}.
+\ac{3DEP} is a national, multi-organizational effort by the \ac{NDEP} to survey elevations with high quality sensors in response to growing stakeholder needs on a recurring collection cycle of no greater than 8 \acp{year} \cite{dewberry2011final,snyder2013national,sugarbaker20143d}.
 \ac{3DEP} leverages two main collection technologies including \ac{LiDAR} for the \ac{CONUS}, \ac{HI}, and \ac{US} territories as well as \ac{InSAR} for \ac{AK}.
@@
 The travel time of the returns, along with knowledge of the speed of light, serve as a relative positioning of the target(s) referenced to a common vertical datum while the intensities serve as indicators of what the target(s) represent.
-Modes within the relationship of return intensities with respect to travel time/distance from the \ac{LiDAR} wave forms can be indicative of vegetation or other \acp{LULC} that reflect signals at varying distances and magnitudes and influence elevation errors \citep{gesch2014accuracy}.
+Modes within the relationship of return intensities with respect to travel time/distance from the \ac{LiDAR} wave forms can be indicative of vegetation or other \acp{LULC} that reflect signals at varying distances and magnitudes and influence elevation errors \cite{gesch2014accuracy}.
 These modes can be discretized into varying \ac{DEM} products representing bare earth, structures, or canopy elevations.
-The horizontal and vertical accuracies and the horizontal resolutions of terrain observations derived from \ac{LiDAR}, and even the consequential economic benefits \citep{dewberry2011final,dewberry2022nation}, are dependent on a variety of sensor, platform, target, and collection specifications and practices such as nominal pulse spacing, nominal pulse density, and \ac{LULC} of the target \citep{heidemann2018lidar,passalacqua2015analyzing,smith2019determining,salach2018accuracy,gesch2014accuracy}.
+The horizontal and vertical accuracies and the horizontal resolutions of terrain observations derived from \ac{LiDAR}, and even the consequential economic benefits \cite{dewberry2011final,dewberry2022nation}, are dependent on a variety of sensor, platform, target, and collection specifications and practices such as nominal pulse spacing, nominal pulse density, and \ac{LULC} of the target \cite{heidemann2018lidar,passalacqua2015analyzing,smith2019determining,salach2018accuracy,gesch2014accuracy}.
 \ac{LiDAR} produces point cloud datasets which are scattered, geo-referenced points representing full wave forms or discretized return intensities.
-Various assessments of the vertical accuracries of \ac{LiDAR} point clouds have yielded satisfactory results in agreement with \ac{3DEP} requirements \citep{stoker2022accuracy,kim2022absolute,callahan2022vertical,kim2022absolute,salach2018accuracy,passalacqua2015analyzing}.
-Point clouds must undergo a series of operations to produce analysis ready, seamless \acp{DEM} \citep{passalacqua2015analyzing}.
+Various assessments of the vertical accuracries of \ac{LiDAR} point clouds have yielded satisfactory results in agreement with \ac{3DEP} requirements \cite{stoker2022accuracy,kim2022absolute,callahan2022vertical,kim2022absolute,salach2018accuracy,passalacqua2015analyzing}.
+Point clouds must undergo a series of operations to produce analysis ready, seamless \acp{DEM} \cite{passalacqua2015analyzing}.
 
 %%%% 3DEP DEMs and Accuracies %%%%
-\ac{3DEP} extends \ac{TNM} to include a 1/27 arc-second (1 \ac{m}), \ac{LiDAR} derived \ac{DEM} product for \ac{CONUS}, \ac{HI}, and \ac{US} territories as well as a 1/2 arc-second (5 \ac{m}) \ac{DEM} for \ac{AK} derived from \ac{InSAR} \citep{sugarbaker20143d,stoker2015usgs}.
-To create bare earth \acp{DEM}, \ac{LiDAR} observations must undergo a series of processes that filter out returns from vegetation, anthropogenic, and other features then grid the observations with resampling methods \citep{passalacqua2015analyzing}.
-The 1 \ac{m} \ac{3DEP} product is a \ac{hydro-flattened}, topographic, and bare-earth raster \ac{DEM} gridded to 1 \ac{km} square shaped tiles with 6 pixels of overlap \citep{arundel20151}.
-Hydro-flattening refers to a process in which hydrologic features such as lakes, reservoirs, streams, rivers, and more are flattened in elevation for bathymetric regions from lower bank to lower bank represented by breaklines \citep{archuleta2017national,maune2018digital}.
-This flattening excludes along gradient directions, parallel to the direction of the breaklines, for hydrologic features that naturally exhibit water conveyance such as streams, rivers, and long reservoirs \citep{arundel20151}.
+\ac{3DEP} extends \ac{TNM} to include a 1/27 arc-second (1 \ac{m}), \ac{LiDAR} derived \ac{DEM} product for \ac{CONUS}, \ac{HI}, and \ac{US} territories as well as a 1/2 arc-second (5 \ac{m}) \ac{DEM} for \ac{AK} derived from \ac{InSAR} \cite{sugarbaker20143d,stoker2015usgs}.
+To create bare earth \acp{DEM}, \ac{LiDAR} observations must undergo a series of processes that filter out returns from vegetation, anthropogenic, and other features then grid the observations with resampling methods \cite{passalacqua2015analyzing}.
+The 1 \ac{m} \ac{3DEP} product is a \ac{hydro-flattened}, topographic, and bare-earth raster \ac{DEM} gridded to 1 \ac{km} square shaped tiles with 6 pixels of overlap \cite{arundel20151}.
+Hydro-flattening refers to a process in which hydrologic features such as lakes, reservoirs, streams, rivers, and more are flattened in elevation for bathymetric regions from lower bank to lower bank represented by breaklines \cite{archuleta2017national,maune2018digital}.
+This flattening excludes along gradient directions, parallel to the direction of the breaklines, for hydrologic features that naturally exhibit water conveyance such as streams, rivers, and long reservoirs \cite{arundel20151}.
 This process includes elevations underneath bridges that are not accurately observed from topographic \ac{LiDAR}.
-According to specifications, the horizontal accuracy of 1 \ac{m} \ac{3DEP} is expected to be within 1 \ac{m} while the vertical accuracies are within 19.6 \ac{cm} and 30 \ac{cm} at the 95\% confidence interval for non-vegetative and vegetative regions, respectively \citep{arundel20151,heidemann2018lidar}.
-Non-vegetative vertical accuracies fall within 10 \ac{cm} \ac{RMSE} \citep{arundel20151,heidemann2018lidar}.
+According to specifications, the horizontal accuracy of 1 \ac{m} \ac{3DEP} is expected to be within 1 \ac{m} while the vertical accuracies are within 19.6 \ac{cm} and 30 \ac{cm} at the 95\% confidence interval for non-vegetative and vegetative regions, respectively \cite{arundel20151,heidemann2018lidar}.
+Non-vegetative vertical accuracies fall within 10 \ac{cm} \ac{RMSE} \cite{arundel20151,heidemann2018lidar}.
  Work by \citet{stoker2022accuracy,callahan2022vertical,kim2022absolute} have verified the vertical accuracies and general quality of the \acp{DEM} for \ac{3DEP} specifications.
@@
 %%%% FIM studies on importance of Lidar and elevation data %%%%
-As previously stated, the features of topographic information including its source, accuracies, resolutions, and processing methods are of primary importance to the quality of \acp{FIM} \citep{national2007elevation,national2009mapping,carswell20183d,bales2009sources}.
+As previously stated, the features of topographic information including its source, accuracies, resolutions, and processing methods are of primary importance to the quality of \acp{FIM} \cite{national2007elevation,national2009mapping,carswell20183d,bales2009sources}.
 Uncertainties in elevations and the processes used to determine them are propogated into \ac{FIM} extents.
-Uncertainty in FIM \citep{merwade2008uncertainty} including data sources, gridding, resampling, resolution, mosaicing, hydrologic conditioning, and more.
+Uncertainty in FIM \cite{merwade2008uncertainty} including data sources, gridding, resampling, resolution, mosaicing, hydrologic conditioning, and more.
 
-\citep{tarolli2014high} good review of channel extraction and other issues surrounding elevations and resolutions.
+\cite{tarolli2014high} good review of channel extraction and other issues surrounding elevations and resolutions.
 
-Flood specific assessments for 3DEP \citep{carswell20183d,bales2009sources,gesch2018best,podhoranyi2015inaccuracy,lamichhane2018effect,tsubaki2013uncertainty,dobbs2010evaluation,arrighi2019effects,zazo2015analysis,bhuyian2018accounting,gesch2012elevation,witt2015evaluation}.
+Flood specific assessments for 3DEP \cite{carswell20183d,bales2009sources,gesch2018best,podhoranyi2015inaccuracy,lamichhane2018effect,tsubaki2013uncertainty,dobbs2010evaluation,arrighi2019effects,zazo2015analysis,bhuyian2018accounting,gesch2012elevation,witt2015evaluation}.
 
@@
 
-DEMs and flooding \citep{casas2006topographic,thomas2016quantifying,savage2016does,passalacqua2010geometric,passalacqua2012automatic,munoth2019effects}
- - Compares Lidar data to NED for flooding using HEC-RAS \citep{wang2005comparison}.
+DEMs and flooding \cite{casas2006topographic,thomas2016quantifying,savage2016does,passalacqua2010geometric,passalacqua2012automatic,munoth2019effects}
+ - Compares Lidar data to NED for flooding using HEC-RAS \cite{wang2005comparison}.
 
-Effects of resolution on hilly areas \citep{dai2019effects}
+Effects of resolution on hilly areas \cite{dai2019effects}
 
-Concluded 10m DEM was good enough for hydro applications \citep{zhang1994digital}
+Concluded 10m DEM was good enough for hydro applications \cite{zhang1994digital}
 
-more FIM and elevations \citep{werner2001impact,omer2003impact,bates2003optimal,tate2002creating,colby2000modeling}.
+more FIM and elevations \cite{werner2001impact,omer2003impact,bates2003optimal,tate2002creating,colby2000modeling}.
 
-Changed the horizontal resolution of DEM for HAND \citep{li2022accounting}, showed that DEM resolution was the most important factor influencing FIM extent skill when varied alongside water depth and drainage threshold.
-\citep{garousi2019terrain} using 3m vs 10m with HAND.
+Changed the horizontal resolution of DEM for HAND \cite{li2022accounting}, showed that DEM resolution was the most important factor influencing FIM extent skill when varied alongside water depth and drainage threshold.
+\cite{garousi2019terrain} using 3m vs 10m with HAND.
 
-Effects of flat areas on FIM quality and overestimation \citep{garousi2019terrain,godbout2019error,jafarzadegan2017based,papaioannou2017probabilistic}
+Effects of flat areas on FIM quality and overestimation \cite{garousi2019terrain,godbout2019error,jafarzadegan2017based,papaioannou2017probabilistic}
 
-\citep{lopez2018influence} how DEM resolution affects hydro connectivity and inundation extents.
+\cite{lopez2018influence} how DEM resolution affects hydro connectivity and inundation extents.
 
@@
 Previous efforts with \ac{OWP} \ac{FIM} avoided use of 1 \ac{m} \ac{3DEP} \ac{DEM} products due to lack of spatial coverage and no seamless data availability.
-As the 1 \ac{m} \ac{3DEP} product rapidly approaches full \ac{CONUS} coverage in 2023, we propose investigating the integration of \ac{3DEP} data into \ac{OWP} \ac{FIM} for continental-scale inundation forecasting abilities \citep{usgs2022status,usgs2022partnerships}.
+As the 1 \ac{m} \ac{3DEP} product rapidly approaches full \ac{CONUS} coverage in 2023, we propose investigating the integration of \ac{3DEP} data into \ac{OWP} \ac{FIM} for continental-scale inundation forecasting abilities \cite{usgs2022status,usgs2022partnerships}.
 We will investigate utilizing \ac{3DEP} data for \ac{HAND} computation to generate the \ac{FIM} hydrofabric.
 Additionally, we investigate the utility of varying spatial resolutions from 1, 3, 5, 10, 15, and 20 \ac{m}.
-\ac{HAND} depends on the drainage assumptions which requires \acp{DEM} to undergo a long series of enforcement processes to ensure monotonically decreasing elevations with hydrologically correct flow directions \citep{garousi2019terrain,nobre2011height,nobre2016hand,aristizabal2022extending}.
+\ac{HAND} depends on the drainage assumptions which requires \acp{DEM} to undergo a long series of enforcement processes to ensure monotonically decreasing elevations with hydrologically correct flow directions \cite{garousi2019terrain,nobre2011height,nobre2016hand,aristizabal2022extending}.
 The resampling of \acp{DEM} into varying spatial resolutions could interact with these \ac{hydro-conditioning} operations thus influencing the \ac{FIM} hydrofabric and the resulting quality of the \acp{FIM} produced.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% MATERIAL AND METHODS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Material and Methods}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Materials and Methods}
 \label{sec:material_and_matheds}
-%
-\ac{HAND} assumes monotonically decreasing elevations with hydrologically relevant flow paths ensuring all cells in a region drain \citep{renno2008hand,nobre2011height,nobre2016hand}.
-This requires the use of extensive \ac{hydro-conditioning} techniques on \acp{DEM} to enforce drainage and agreement with existing hydrography \citep{aristizabal2022extending,maidment2017conceptual,liu2016cybergis,liu2020height}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\ac{3DEP}}
-\label{ssec:3dep}
-%
+Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
+Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
+Please disclose at the submission stage any restrictions on the availability of materials or information.
+New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
+
+Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers.
+If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
+They must be provided prior to publication.
+
+Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
+\begin{quote}
+This is an example of a quote.
+\end{quote}
 
 \begin{comment}
-%% make a table of different 3DEP data sources
-Table \ref{tab:tnm_layer_specifications} illustrates the various elevation layers of \ac{TNM} along with their respective specifications and sources.
-\begin{table}[h!]
- \begin{center}
-  \caption{\acf{TNM} - Elevation layers specifications. These layers are also known as the \acf{NED}}
-  \label{tab:tnm_layer_specifications}
-  \begin{tabular}{c|c|c|c|c|c} % <-- Alignments: 1st column left, 2nd middle and 3rd right, with vertical lines in between
-   \multirow{2}*{Layer} & \multicolumn{2}{|c|}{Spatial Resolution} & \multicolumn{2}{|c|}{Accuracy (\ac{m})} & \multirow{2}*{Sources} \\
-    & arc-second & (\ac{m}) & Horizontal & Vertical &  \\
-   \hline
-   1/3 arc-second seamless & 1/3 & 10 & varies & varies & Many  \\
-   \hline
-  \end{tabular}
- \end{center}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+
+This section may be divided by subheadings.
+It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
+\subsection{Subsection}
+\subsubsection{Subsubsection}
+
+Bulleted lists look like this:
+\begin{itemize}
+\item	First bullet;
+\item	Second bullet;
+\item	Third bullet.
+\end{itemize}
+
+Numbered lists can be added as follows:
+\begin{enumerate}
+\item	First item; 
+\item	Second item;
+\item	Third item.
+\end{enumerate}
+
+The text continues here.
+
+\subsection{Figures, Tables and Schemes}
+
+All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, Table~\ref{tab2}, etc.
+
+\begin{figure}[H]
+\includegraphics[width=10.5 cm]{dependencies/Definitions/logo-mdpi}
+\caption{This is a figure.
+Schemes follow the same formatting.
+If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
+(\textbf{b}) Description of what is contained in the second panel.
+Figures should be placed in the main text near to the first time they are cited.
+A caption on a single line should be centered.\label{fig1}}
+\end{figure}   
+\unskip
+
+\begin{table}[H] 
+\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCC}
+\toprule
+\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+\midrule
+Entry 1		& Data			& Data\\
+Entry 2		& Data			& Data\\
+\bottomrule
+\end{tabularx}
 \end{table}
-\end{comment}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{\ac{3DEP} Data Acquisition and Preparation}
-\label{ssec:py3dep}
-%
-Since the highest resolution \ac{3DEP} \ac{DEM} is at the 1/3 arc-second, we leverage a tool known as \ac{Py3DEP} to handle mosaicing and resampling of best available \acp{DEM} into a singular seamless one.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}
-\label{ssec:evaluation}
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{\ac{BLE}}
-\label{sssec:ble}
-%
+\unskip
 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{\ac{RAS2FIM}}
-\label{sssec:ras2fim}
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{\ac{InSAR}}
-\label{sssec:insar}
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% RESULTS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{sec:results}
-%
+\begin{table}[H]
+\caption{This is a wide table.\label{tab2}}
+	\begin{adjustwidth}{-\extralength}{0cm}
+		\newcolumntype{C}{>{\centering\arraybackslash}X}
+		\begin{tabularx}{\fulllength}{CCCC}
+			\toprule
+			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
+			\midrule
+			Entry 1		& Data			& Data			& Data\\
+			Entry 2		& Data			& Data			& Data \textsuperscript{1}\\
+			\bottomrule
+		\end{tabularx}
+	\end{adjustwidth}
+	\noindent{\footnotesize{\textsuperscript{1} This is a table footnote.}}
+\end{table}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% DISCUSSION
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\begin{listing}[H]
+%\caption{Title of the listing}
+%\rule{\columnwidth}{1pt}
+%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
+%\rule{\columnwidth}{1pt}
+%\end{listing}
+
+Text.
+
+Text.
+
+\subsection{Formatting of Mathematical Components}
+
+This is the example 1 of equation:
+\begin{linenomath}
+\begin{equation}
+a = 1,
+\end{equation}
+\end{linenomath}
+the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
+%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
+This ensures correct linenumbering.
+The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
+
+This is the example 2 of equation:
+\begin{adjustwidth}{-\extralength}{0cm}
+\begin{equation}
+a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
+\end{equation}
+\end{adjustwidth}
+
+% Example of a page in landscape format (with table and table footnote).
+%\startlandscape
+%\begin{table}[H] %% Table in wide page
+%\caption{This is a very wide table.\label{tab3}}
+%	\begin{tabularx}{\textwidth}{CCCC}
+%		\toprule
+%		\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}	& \textbf{Title 4}\\
+%		\midrule
+%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
+%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
+%		\bottomrule
+%	\end{tabularx}
+%	\begin{adjustwidth}{+\extralength}{0cm}
+%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
+%	\end{adjustwidth}
+%\end{table}
+%\finishlandscape
+
+% Example of a figure that spans the whole page width. The same concept works for tables, too.
+\begin{figure}[H]
+\begin{adjustwidth}{-\extralength}{0cm}
+\centering
+\includegraphics[width=13.5cm]{dependencies/Definitions/logo-mdpi}
+\end{adjustwidth}
+\caption{This is a wide figure.\label{fig2}}
+\end{figure}
+
+Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
+%% Example of a theorem:
+\begin{Theorem}
+Example text of a theorem.
+\end{Theorem}
+
+The text continues here. Proofs must be formatted as follows:
+
+%% Example of a proof:
+\begin{proof}[Proof of Theorem 1]
+Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
+\end{proof}
+The text continues here.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
-\label{sec:discussion}
-%
-Future work could employ the use of STAC catalogs to acquire and process data from the cloud from Open Topography (\url{https://stacindex.org/catalogs/opentopography#/})
-Tangentially, we assert the gridding of source elevations from \ac{InSAR} and \ac{LiDAR} point clouds to be out of scope for this analysis elected instead to deal with already gridded elevations.
-Future work can further understand how \ac{HAND} is affect by gridding techniques most specifically that of point cloud conversion.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% CONCLUSIONS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
+The findings and their implications should be discussed in the broadest context possible.
+Future research directions may also be highlighted.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusions}
-\label{sec:conclusions}
-%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Flush acronym usage
 \acresetall
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ACKNOWLEDGMENTS
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgments}
-\label{sec:acknowledgments}
-%
-Collate acknowledgements in a separate section at the end of the article before the references and do not, therefore, include them on the title page, as a footnote to the title or otherwise.
-List here those individuals who provided help during the research (e.g., providing language help, writing assistance or proof reading the article, etc.).
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% FUNDING SOURCES
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Funding Sources}
-\label{sec:funding_sources}
-%
-List funding sources in this standard way to facilitate compliance to funder's requirements:
 
-Funding: This work was supported by the National Institutes of Health [grant numbers xxxx, yyyy]; the Bill \& Melinda Gates Foundation, Seattle, WA [grant number zzzz]; and the United States Institutes of Peace [grant number aaaa].
+This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
 
-It is not necessary to include detailed descriptions on the program or type of grants and awards.
-When funding is from a block grant or other resources available to a university, college, or other research institution, submit the name of the institute or organization that provided the funding.
-If no funding has been provided for the research, it is recommended to include the following sentence:
-This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% APPENDIX
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Patents}
+
+This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}
+
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
+A supporting video article is available at doi: link.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``Conceptualization, X.X.
+and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
+All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+Authorship must be limited to those who have contributed substantially to the work~reported.}
+
+\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
+
+\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
+You might choose to exclude this statement if the study did not require ethical approval.
+Please note that the Editorial Office might ask you for further information.
+Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans.
+OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals.
+OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}
+
+\informedconsent{Any research article describing a study involving humans should contain this statement.
+Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
+You might also choose to exclude this statement if the study did not involve humans.
+
+Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
+Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}
+
+\dataavailability{In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
+Please refer to suggested Data Availability Statements in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.
+If the study did not report any data, you might add ``Not applicable'' here.}
+
+\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+
+\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
+Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section.
+If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the~results''.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+\sampleavailability{Samples of the compounds ... are available from the authors.}
+
+%% Only for journal Encyclopedia
+%\entrylink{The Link to this entry published on the encyclopedia platform.}
+
+\abbreviations{Abbreviations}{
+The following abbreviations are used in this manuscript:\\
+
+\noindent 
+\begin{tabular}{@{}ll}
+MDPI & Multidisciplinary Digital Publishing Institute\\
+DOAJ & Directory of open access journals\\
+TLA & Three letter acronym\\
+LD & Linear dichroism
+\end{tabular}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixstart
 \appendix
+\section[\appendixname~\thesection]{}
+\subsection[\appendixname~\thesubsection]{}
+The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
 
-%% \section{}
-%% \label{}
-%
+\begin{table}[H] 
+\caption{This is a table caption.\label{tab5}}
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCC}
+\toprule
+\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+\midrule
+Entry 1		& Data			& Data\\
+Entry 2		& Data			& Data\\
+\bottomrule
+\end{tabularx}
+\end{table}
+
+\section[\appendixname~\thesection]{}
+All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
+
+\end{comment}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acronyms}
+\section[\appendixname \thesection]{Acronyms}
 \label{sec:acronyms}
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BIBLIOGRAPHY
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{adjustwidth}{-\extralength}{0cm}
+%\printendnotes[custom] % Un-comment to print a list of endnotes
+
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
+\bibliography{bibliography/owp_3dep_2022.bib}
+
+% If authors have biography, please use the format below
+%\section*{Short Biography of Authors}
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{dependencies/Definitions/author1.pdf}}}
+%{\textbf{Firstname Lastname} Biography of first author}
 %
-%% For citations use: 
-%%       \citet{<label>} ==> Jones et al. [21]
-%%       \citep{<label>} ==> [21]
-%%
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{dependencies/Definitions/author2.jpg}}}
+%{\textbf{Firstname Lastname} Biography of second author}
 
-%% If you have bibdatabase file and want bibtex to generate the
-%% bibitems, please use
-%%
-\bibliographystyle{elsarticle-num-names} 
-\bibliography{bibliography/owp_3dep_2022}
+% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% END
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{adjustwidth}
 \end{document}
-\endinput

### 2022-10-05 manuscripts/owp_3dep_2022/Makefile.sh

@@ -1,7 +1,7 @@
 #!/bin/bash -e
 
 # declare files
-mainFile=owp_3dep.tex
+mainFile=template.tex
 bibFile=bibliography/owp_3dep_2022.bib
 #expandedFile=owp_3dep_expanded.tex
 #differenceFile=owp_3dep_expanded_diff.tex

### 2022-10-05 manuscripts/owp_3dep_2022/Makefile.sh
fixed makefile for mdpi

@@ -1,7 +1,7 @@
 #!/bin/bash -e
 
 # declare files
-mainFile=template.tex
+mainFile=owp_3dep.tex
 bibFile=bibliography/owp_3dep_2022.bib
 #expandedFile=owp_3dep_expanded.tex
 #differenceFile=owp_3dep_expanded_diff.tex

### 2023-02-11 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
second review responses and changes submitted to coauthors 20230211

@@
 %
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
+\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{1,4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
 %
@@
 One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
+AHPS utilizes a variety of techniques including distributed modeling \cite{reed2004overall,koren2004hydrology,duan2003global}, ensemble forecasting \cite{day1985extended,seo2000simulation,mullusky2002simplified,herr2002simplified}, hydraulic routing \cite{cajina2002fldview,fread1973technique}, multi-sensor precipitation methods \cite{seo2002real,kondragunta2001outlier,breidenbach1999accounting,bonnin1996noaa}, flood inundation mapping \cite{cajina2002fldview}, and calibration techniques \cite{zhang2003hydrologic,hogue2003multi,duan2003global,gupta2003advances,parada2003multi}.
 AHPS provides forecasting services in the form of ensemble streamflows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
@@
 \centering
-\includegraphics[scale=1.0]{figures/forecast_points.jpg}
+\includegraphics[scale=0.75]{figures/forecast_points.jpg}
 \caption{Forecast points with and without Flood Inundation Maps (FIM) in United States' Advanced Hydrologic Prediction System (AHPS).
@@
 The use of a 2-dimensional (2D) hydrodynamic model across a continental-scale and high spatial resolutions is very cost prohibitive especially in an operational setting.
-The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert 1-dimensional (1D) riverine discharges to stages, and finally to inundation extents and depths.
+The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert 1-dimensional (1D) riverine discharges to stages, and finally to inundation extents.
 %
@@
 \citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as SRCs by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
-This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
+This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative,hocini2020performance}.
 %
@@
 HAND, to our knowledge, has only been applied to natural, inland, and riverine inundation applications thus it is also missing pluvial, coastal, ground water, and dam break components among other possible sources of flooding.
-Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest stream line which is used for catchment allocation and relative elevation calculation \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,maidment2017conceptual,garousi2019terrain,zheng2018river,zheng2018geoflood,johnson2019integrated,renno2008hand}.
+Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest stream line which is used for catchment allocation and relative elevation calculation \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,maidment2017conceptual,garousi2019terrain,zheng2018river,zheng2018geoflood,johnson2019integrated,renno2008hand,hocini2020performance}.
 Stream thalweg networks must also collectively drain to a singular outlet point for a given processing region \cite{nobre2016hand,zheng2018geoflood,renno2008hand}.
@@
 \subsection{Office of Water Prediction Flood Inundation Mapping}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-In order to mitigate the ever increasing threat of flooding to life and property, an operational capability is required to extend NWM streamflow forecasts to river stages, inundation extents, and inundation depths.
+\label{ssec:owp_fim}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+In order to mitigate the ever increasing threat of flooding to life and property, an operational capability is required to extend NWM streamflow forecasts to river stages and inundation extents.
 OWP FIM is introduced here as a continental scale capability that generates these products at high spatial and temporal resolutions.
@@
 Within the operational framework of OWP FIM, we introduce research demonstrating how FIM performance skill with HAND can be improved by discretizing stream networks into units of an effective unit Horton-Strahler stream order \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} for HAND computation contexts.
+
 Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to the stream accumulation threshold which is closely related to stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network into discrete level paths (LPs) of singular, effective stream order, can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
-This capability is introduced progressively as MS (whose network represents about 4\% of FR network) and to a higher degree Generalized Mainstems (GMS) (covers entire FR network) which will be explained later on.
+HAND produced FIMs are limited in only providing inundation sourced from the nearest drainage line, however, depending on flow conditions and topography, a given area may have multiple contributing fluvial sources of surface water inundation.
+The forecasting community, in reviewing HAND, have noted significant negative effects at the confluence of lower flow tributaries with higher flow rivers for which the phrase ``catchment boundary issue'' has been termed.
+In previous studies, FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND which is closely related to the maximum Horton-Strahler stream order of the network \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+This sensitivity is partly in due to the limitation that catchment boundaries place on inundation extents where only the nearest drainage line can source inundation for any particular area.
+
+Figure \ref{fig:catchment_boundaries_issue} illustrates the exact situation our solution proposes to address where two tributaries converge with a higher order stream segment. 
+An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 yr extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
+The higher discharge along the main segment in Figure \ref{fig:catchment_boundaries_issue} of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS.
+This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
+As a parallel problem, there is excess water accumulated along the mainstem that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the mainstem catchments.
+%
+\begin{figure}[H]
+\centering
+\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
+\caption{The figure represents an agreement map between a HAND derived FIM and one produced from the Base Level Engineering (BLE) program for a 100 yr magnitude event at HUC8 12090301.
+Agreement maps are symbolized by false negatives (FN), true negatives (TN), false positives (FP), and true positives (TP) where inundated represents the positive condition (see Section \ref{ssec:evaluation} for more details).
+The streamflows associated with each river segment are shown in CMS while the flow directions are symbolized as red arrows.
+The presence of FNs at the confluence of tributaries (circled in red) with the main segment is associated with lower flow rates in the tributaries that don't account for backwater effects.
+Additionally, the flow of 1900 CMS from the main segment cannot extend to the neighboring catchments belonging to its tributaries shown here.
+Water pools up vertically along the catchment boundaries of the higher order segment distorting rating curve behavior (Section \ref{sec:discussion}).
+Sourcing fluvial inundation from HAND is limited to only its nearest drainage line which is the main issue this study aims to address.
+}
+\label{fig:catchment_boundaries_issue}
+\end{figure}
+%
+
+We seek to resolve this catchment boundary problem or nearest drainage limitation by discretizing the target stream network into stream networks of reduced, unit stream order to avoid the constraining of catchments by those belonging to lower order neighbors.
+By discretizing the network into stream networks of unit stream order (later defined as level paths or LPs), we remove the influence of neighboring catchments that constrain the inundation extent.
+This creates much larger and overlapping catchments that can source fluvial inundation from multiple reaches as required by the given river stage at current flow conditions.
+We present two successive methods, National Weather Service MS (Section \ref{sssec:nws_mainstems}) and Generalized Mainstems (GMS) (Section \ref{sssec:generalized_mainstems}), implemented that reduce the effective Horton-Strahler stream orders of the networks employed and test our presented hypothesis that stream networks of unit stream order enhance quality of FIM extents produced with HAND by expanding the nearest drainage definition to increase potential inundated areas.
+
+Here we demonstrate how reducing a HAND processing unit's stream network into discrete LPs of singular, effective stream order, can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
+This capability is introduced progressively as MS (whose network represents about 4\% of FR network) and to a higher degree GMS (covers entire FR network) which will be explained later on.
 The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM.
@@
 \hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
+InFRM & Flood Inundation Extents & Inundation extents produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
 \hline
@@
 \subsubsection{Stream Network Enforcement} 
-\label{ssec:stream_network_enforcment}
-%
-The location of the stream network is enforced to ensure general agreement with the NWM network which is used for forecasting the streamflow inputs.
+\label{ssec:stream_network_enforcement}
+%
+Both the location and geometry of the stream network are enforced to agree with the NWM stream network and to enforce synthetic bathymetry and generate more hydrologically correct catchments.
 The NHDPlusHR Beta BurnLineEvent layer is used to enforce stream locations in the NHDPlusHR workflow and best agrees with thalweg locations in the DEM used so it is also used here for hydro-enforcement \cite{moore2019user}.
-
-However, to better match the drainage density of the NWM FR V2.1 stream network, which is based on the NHDPlus V2, the BurnlineEvents are pruned utilizing a nearest neighbor search around the NWM flowlines.
-Headwater points are first derived for the NWM FR V2.1.
-For every NWM headwater point, the nearest NHDPlusHR point is selected and placed into a set while those excluded are discarded.
-Only the nearest point on the NHDPlusHR is used so any portion of the NHDPlusHR network upstream of this nearest point is discarded to avoid extending inundation too far above the modeling domain.
-The points in this nearest neighbor set are then traversed downstream.
-Any headwater portion in the NHDPlusHR or any other stream not traversed are pruned away to better match the resolution and spatial locations of the NWM stream network and its headwater points.
-The resulting pruned NHDPlusHR stream network gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHDPlusHR network corresponds to the NHDPlusHR network at or downstream of NWM V2.1 headwater locations only.
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents. 
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{This figure illustrates some of the datasets that result from the pruning of NHDPlusHR Beta BurnlineEvents (dotted black) to the stream density of the NWM FR V2.1 density (blue).
-The stream network used for forecasting, NWM FR V2.1, is of lower stream density than that of the NHDPlusHR which has better agreement with the thalweg locations in the DEM used.
-Thus, we opt to prune the NHDPlusHR network to match the general location and density of the NWM network.
-The nearest neighbor segment in the NHDPlusHR of each NWM headwater locations and the nearest point on that segment is determined to match the closest point to that of the NWM headwater.
-These points are then traversed downstream and any segments not traversed are pruned away.
-The resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR BurnlineEvents.}
-\label{fig:stream_density_pruning}
-\end{figure}
-%
-
-The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
-The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop and buffer distance. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling an inverted triangular notch shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop which were set to fixed values of 70 m, 1000 m, and 10 m, respectively, but available to the user via the parameter file.
-While the values for these parameters are critical to the inundation extents produced, especially for lower flow rates where bathymetric information has more influence, finding their optimal values for OWP FIM was not done since it was out of the main scientific scope of this article.
-Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement techniques especially when computational costs are considered. 
+This network goes through a stream density reduction procedure to better match the stream density of the NWM stream network.
+This reduced density NHDPlusHR network is now used for hydro-enforcement along with the AGREE DEM Surface Reconditioning System to excavate channels into the DEM \cite{hellweger1997agree}.
 Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
 Some of these drawbacks are addressed by additional conditioning techniques applied later on.
+For more details on how we reduce the density of the NHDPlusHR stream network along with technical implementation information on the AGREE DEM procedure, please see \ref{sec:app_stream_network_enforcement}.
 %
@@
 These datasets include the HAND or relative elevation model (REM) raster, reach-level catchments raster/polygons, DEM-derived streamlines, SRCs, and cross-walk table.
+Within the context of this section, we refer to HAND as a grid of relative elevations to detrend elevations away from mean level towards elevations referenced to the nearest drainage line (See Section \ref{sssec:hand}).
+SRCs are considered stage-discharge relationships that are derived synthetically using the Manning's equation (See Section \ref{sssec:synthetic_rating_curve}).
+Lastly, a cross-walk table attempts to conflate NWM reach identifiers to those derived of the FIM stream network (See Section \ref{sssec:cross_walking_networks}).
 The following sub-sections describe how the subset and hydro-enforced geospatial datasets are converted into the FIM hydrofabric.
@@
 %
-To facilitate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D-8 flow directions.
-D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
-The horizontal component of slope is defined as one for the four neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$ by means of the Pythagorean theorem.
-Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression-filled cells must be resolved.
-
+To facilitate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D-8 flow directions \cite{o1984extraction}.
 Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment} where flats are resolved by incrementing elevations iteratively.
-Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
-The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
-The established literature in this niche field of hydrology discusses how prevalent flats can be in given study areas and how difficult the problem is from both computational and hydrologic stand-points \cite{garbrecht1997assignment,tarboton2009generalized,tarboton2005terrain,survila2016scalable,barnes2014efficient,tesfa2011extraction,wallis2009parallel}.
 OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
+For more information on the derivations of flow directions and resolving flats, please see \ref{sec:app_flow_direction_and_flat_resolution}.
 %
@@
 The FIM stream network is of similar drainage density as the NWM V2.1 network and fully converges at all junctions leaving no divergences in the network.
-This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are pruned headwater locations of the NHDPlusHR Beta BurnlineEvents layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
+This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcement} and \ref{sec:app_stream_network_enforcement}).
+These seeds points are headwater locations of the NHDPlusHR Beta BurnlineEvents layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
 Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
@@
 Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve SRC and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
+This parameter was held constant across all versions of HAND that we used which are introduced in Section \ref{ssec:stream_order_reduction}.
 Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretization of the stream network.
@@
 \subsubsection{Height Above Nearest Drainage}
+\label{sssec:hand}
 %
 Once the pixel level catchments are derived, the final relative elevations can be computed.
-The elevation of every thalweg pixel is subtracted from the elevations of the non-thalweg pixels within the same, corresponding pixel-level catchment described in Section \ref{sssec:catchments}.
-The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
-Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling \cite{djokic2019arc}.
+To compute these relative elevations, we utilize the same technique found in previous HAND implementations \cite{zheng2018geoflood,zheng2018river,nobre2011height,nobre2016hand,maidment2017conceptual,garousi2019terrain}.
+For each pixel level catchment described in Section \ref{sssec:catchments}, we subtract the elevations of the non-thalweg pixels from the thalweg pixel of that catchment.
+Contrary to some of the some other previous HAND implementations, the DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning} \cite{djokic2019arc}.
+This DEM utilizes the native elevations in regions outside of the excavated channel from the AGREE DEM method \cite{djokic2019arc}.
 Any negative values resulting from this subtraction with native elevations are replaced by zero.
@@
 We leave this issue open in this study and elaborate on needs with respect to bathymetry and Manning's n values in the Discussion section (Section \ref{sec:discussion}).
+We constrain the equation by selecting 84 stage values (y in Eq.
+\ref{eq:reach_averaged_mannings_equation}) from 0 to 25 meters in depth at a third of a meter increments to calculate the discharge values for each stage value.
+Due to the varying nature of stage (y in Equation \ref{eq:reach_averaged_mannings_equation}) as explained, the terms V(y) and B(y) also update according to previous work with reach averaged SRCs \cite{zheng2018river}.
 
@@
 For now and for the purpose of this study, we examine the SRCs with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values (y in Eq.
-\ref{eq:reach_averaged_mannings_equation}) from 0 to 25 meters in depth at a third of a meter increments to calculate the discharge values for each stage value.
 %
@@
 %
-The DEM based stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with NWM reach identifiers so that a discharge can be converted to stage and later inundation extents and depths.
+The DEM based stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with NWM reach identifiers so that a discharge can be converted to stage and later inundation extents.
 For the FR version of HAND, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
@@
 As previously discussed, HAND based FIMs are subject to many assumptions and limitations in order serve as a suitable inundation proxy for large scale, high resolution domains.
-HAND produced FIMs are limited in only providing inundation sourced from the nearest drainage line, however, depending on flow conditions and topography, a given area may have multiple contributing fluvial sources of surface water inundation.
-The forecasting community, in reviewing HAND, have noted significant negative effects at the confluence of lower flow tributaries with higher flow rivers for which the phrase ``catchment boundary issue'' has been termed.
-In previous studies, FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND which is closely related to the maximum Horton-Strahler stream order of the network \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is partly in due to the limitation that catchment boundaries place on inundation extents where only the nearest drainage line can source inundation for any particular area.
-
-Figure \ref{fig:catchment_boundaries_issue} illustrates the exact situation our solution proposes to address where two tributaries converge with a higher order stream segment. 
-An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 yr extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the main segment in Figure \ref{fig:catchment_boundaries_issue} of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS.
-This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the mainstem that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the mainstem catchments.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{The figure represents an agreement map between a HAND derived FIM and one produced from the Base Level Engineering (BLE) program for a 100 yr magnitude event at HUC8 12090301.
-Agreement maps are symbolized by false negatives (FN), true negatives (TN), false positives (FP), and true positives (TP) where inundated represents the positive condition (see Section \ref{ssec:evaluation} for more details).
-The streamflows associated with each river segment are shown in CMS while the flow directions are symbolized as red arrows.
-The presence of FNs at the confluence of tributaries (circled in red) with the main segment is associated with lower flow rates in the tributaries that don't account for backwater effects.
-Additionally, the flow of 1900 CMS from the main segment cannot extend to the neighboring catchments belonging to its tributaries shown here.
-Water pools up vertically along the catchment boundaries of the higher order segment distorting rating curve behavior (Section \ref{sec:discussion}).
-Sourcing fluvial inundation from HAND is limited to only its nearest drainage line which is the main issue this study aims to address.
-}
-\label{fig:catchment_boundaries_issue}
-\end{figure}
-%
-
-We seek to resolve this catchment boundary problem or nearest drainage limitation by discretizing the target stream network into stream networks of reduced, unit stream order to avoid the constraining of catchments by those belonging to lower order neighbors.
-By discretizing the network into stream networks of unit stream order (later defined as LP), we remove the influence of neighboring catchments that constrain the inundation extent.
-This creates much larger and overlapping catchments that can source fluvial inundation from multiple reaches as required by the given river stage at current flow conditions.
-We present two successive methods, National Weather Service MS (Section \ref{sssec:nws_mainstems}) and GMS (Section \ref{sssec:generalized_mainstems}), implemented that reduce the effective Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND by expanding the nearest drainage definition to increase potential inundation areas.
-
+As introduced in Section \ref{ssec:owp_fim}, HAND based FIMs fail to account for multiple sources of fluvial inundation which leads to flow constrictions at high flow confluences.
+We hypothesized that reducing the scale of HAND computation down to stream networks of unit Horton-Strahler stream order can help account for multiple fluvial sources of inundation.
 To clarify the phrase ``reducing Horton-Strahler stream order'' used extensively in this paper, every FIM used in evaluation contains a flood extent sourced from every NWM forecast point in the given evaluation domain.
@@
 Figure \ref{fig:level_path_methods} illustrates how LPs (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-The figure shows computed arbolate sums and unique LP identifiers on a HUC12 (120903010404) for clarity but were computed at the corresponding HUC8.
+The figure shows computed arbolate sums and unique LP identifiers on a headwater subset of HUC8 (12020002) for clarity but were computed at the corresponding HUC8.
 The mainstem of the figure runs from the red ellipses to the black one which is the outlet.
@@
 At the LP scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the LP in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement by burning these lines and seeding from its headwater points directly instead of going through the NHDPlusHR network as described in Section \ref{ssec:stream_network_enforcment} .
+The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement by burning these lines and seeding from its headwater points directly instead of going through the NHDPlusHR network as described in Section \ref{ssec:stream_network_enforcement} and \ref{sec:app_stream_network_enforcement}.
 This decision was motivated by the difficulty in deriving LPs in the NWM stream network with high agreement with the LPs derived for the NHDPlusHR stream lines.
@@
 \includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates the NWM Full Resolution V2.1 stream network discretized into level paths (LP), symbolized by unique colors, as well as the values of the arbolate sums to the nearest whole km distance.
-The LPs were derived on a HUC8 level (12090301) but only illustrated for a HUC12 (120903010404) for clarity.
-The mainstem of this HUC12 runs from the red ellipses to the outlet denoted by the black ellipses.
+\caption{Illustrates the NWM Full Resolution V2.1 stream network discretized into level paths (LP), symbolized by unique colors, as well as the values of the arbolate sums in km units.
+The LPs were derived on a HUC8 level (12020002) but only illustrated for a small, headwater subset of this HUC8 for clarity purposes.
 Arbolate sums are defined as the cumulative drainage distances of all upstream stream lines.
@@
 For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
+While we do not evaluate FIM depths in this study, we do compute depths first as to threshold them to produce extents.
 For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
@@
 For GMS, all FIMs from all the LPs in a given area are mosaiced together then K is set to this number of LPs.
+Lastly, we apply Equation \ref{eq:hand_fim} to the mosaiced depths for the MS and GMS cases in order to obtain FIM extents which is the subject of our evaluations.
 Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
@@
 %
-Possible benchmark FIM candidates for evaluation purposes include high water marks, remote sensing observation, crowd-sourced information, and modeled extents.
-These sources are all subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to a lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
-While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to as to isolate out other hydrological factors.
-
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
-This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, higher level of sophistication when compared to HAND, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
+OWP FIM is linked with the NWM for operational purposes, utilizing streamflow inputs, to produce FIM at a continental scale.
+We evaluated various validation data sources for operational use, each with its own strengths and weaknesses, including high water marks (HWMs) \cite{musser2017characterization,watson2017characterization,breaker2016characterization}, remote sensing observations \cite{aristizabal2020high,aristizabal2021mapping}, and modeled extents from hyper-resolution, local models \cite{mcenery2005noaa}.
+Although they have advantages, the three data sources mentioned do not have streamflow information, which is required by OWP FIM.
+Using the NWM as a source of streamflow input would be a logical choice due to its operational use, but this would introduce hydro-climatic uncertainties that could impact the results of adding a multi-fluvial source extension to HAND.
+On the other hand, relying on point sources of streamflow or stage, such as gages, would eliminate this limitation, but limit the information available to only a single source of inundation at a specific site.
+Given these limitations, we investigated the use of existing HEC-RAS based models to address the limitations caused by hydro-climatic uncertainties and sparse streamflow inputs.
+
+Our HAND based approach coupled with SRCs requires streamflow as input and is agnostic as to the source of that streamflow whether forecasted, observed, or probablistic.
+Due to this fact, evaluation of our relative elevation CFIM method was conducted by comparison to the HEC-RAS 1D models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
+This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, higher level of sophistication when compared to HAND, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid,bates201630m,wing2017validation,hu2021real,li2022accounting,li2022comprehensive,zheng2018geoflood}.
 We selected 49 available HUC8s, shown in Figure \ref{fig:all_ble_maps}, which span about 185 thousand $km^2$ across nine states.
@@
 We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation sourced from OWP FIM.
+Since the BLE does account for reservoir inundation, some of the BLE reservoir inundation extents extend beyond the NWM reservoir geometries contributing to false negatives (FNs), or under-prediction.
 
@@
 This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with unit Horton-Strahler order due to the lack of representation of these additional headwater streams in the NWM network.
+Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of FNs along headwater streams with unit Horton-Strahler order due to the lack of representation of these additional headwater streams in the NWM network.
 While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
@@
 This can be inferred by close inspection of Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} where changes in POD are significantly higher than that of FAR.
-%This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
+This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
 Upon investigation of the performance of HAND derived FIM, we note a general increase of catchment sizes for the MS and GMS methods when compared to the FR method as they are now delineated independently of any tributaries that would constrain catchment sizes otherwise.
@@
 The TPs are no longer bounded by the catchment lines and allowed to expand to their natural extents.
+Additionally, this study has a limitation in that it did not assess FIM depths, so limited conclusions can be made about the impact of drainage order reduction techniques on the accuracy of FIM depths or water surface elevations.
 
@@
 Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
-While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
+While the channel slope and reach lengths are not exactly the same across methods, their averaged differences were found to be statistically insignificant which only leaves room for deviations in volume and bed area.
+Specifically, we found FR to have an average reach length for the study region of 1354.8 km with a standard deviation 249.6km.
+For GMS, we computed an average reach length of 1398.1 km with a standard deviation of 50.2 km.
 Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius but these associations only hold when reach length, L, is considered.
@@
 The NWM stream network was derived using the NHDPlus V2 dataset which was derived from coarser DEMs than those used here. 
-Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
+Additional conflation errors are realized by cross-walking the stream network used by the BLE maps and those of HAND.
+The methods that intersect HEC-RAS cross-sections with NWM reaches could introduce errors that violate mass conservation.
+Additionally, as previously noted and illustrated in Figure \ref{fig:ble_evaluation_method}, the BLE network is denser than that of the NWM which leads to FNs due to headwater streams that are present in the BLE but not within the NWM.
 Until a singular stream network is used for the NWM, BLE benchmark, and for HAND based FIM, conflation will continue being a source of error.
@@
 Higher flow events tend to cover regions with more floodplain inundation thus less sensitive to errors from bathymetric data quality.
-On a related note, the use of the AGREE DEM method discussed in Section \ref{ssec:stream_network_enforcment} also interacts with the bathymetry issue introducing several artificial geometry parameters that affect SRC shape and quality.
-Due to focus on the nearest drainage problem, we leave future work related to SRC representation including roughness estimation, bathymetric data assimilation, and bathymetry adjustments as opportunities for major enhancements in HAND based FIM.
-
-Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
+On a related note, the use of the AGREE DEM method discussed in Section \ref{ssec:stream_network_enforcement} and \ref{sec:app_stream_network_enforcement} also interacts with the bathymetry issue introducing several artificial geometry parameters that affect SRC shape and quality.
+To reiterate in this discussion, Section \ref{ssec:evaluation} pointed out that the BLE's representation of reservoir inundation extends beyond the NWM's defined reservoir boundaries, which results in an under-prediction in the OWP FIM that is reflected in the computed metrics.
+Due to focus on the nearest drainage problem, we leave future work related to SRC representation including roughness estimation, bathymetric data assimilation and adjustments as well as reservoir inundation as opportunities for major enhancements in HAND based FIM.
+
+Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment, and reservoir representation are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
 For HAND to be derived and thus create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
@@
 Ultimately, the use of more advanced 2D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/rating_curve_comparison.jpg}
-\caption{Illustrates average quantities for the three methods, FR, MS, and GMS, for each stage value (m). 
-The values are (a) Discharge $m^3s^{-1}$, (b) Volume $m^3$, (c) Bed Area $m^2$, (d) a function of Volume and Bed Area, and (e) number of pixels.
-}
-\label{fig:rating_curve_comparison}
-\end{figure}
-%
 %
@@
 %
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale=1.0]{figures/rating_curve_comparison.jpg}
+    \caption{Illustrates average quantities for the three methods, FR, MS, and GMS, for each stage value (m). 
+    The values are (a) Discharge $m^3s^{-1}$, (b) Volume $m^3$, (c) Bed Area $m^2$, (d) a function of Volume and Bed Area, and (e) number of pixels.
+    }
+    \label{fig:rating_curve_comparison}
+    \end{figure}
+    %
+ %
 % Conclusions
@@
 %
-Floods present a significant, under-served, and expanding risk to life, property, and resources.
-Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
-The National Water Model (NWM), developed by the National Oceanic and Atmospheric Administration's (NOAA) Office of Water Prediction (OWP) and the National Center for Atmospheric Research (NCAR), provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
-Additional modeling is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
-Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
-HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
-Scalable, open-source software, known as OWP FIM, was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
-HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
-Finally, we used this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
-FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
-
 The main contribution and conclusion of this work centers around a fundamental limitation in HAND based FIM which is a failure to account for multiple possible sources of fluvial inundation since HAND only considers inundation from the nearest drainage line.
@@
 Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
-We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
+We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden, for his technical expertise.
 Additionally, David Blodgett from the United States Geolgical Survey (USGS) Water Mission Area was instrumental in helping define level paths and other hydrographic work.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %% Optional Appendix goes here
@@
 %
-%\appendix
-%\section{Here is a sample appendix}
-
+\clearpage 
+\appendix
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% APPENDIX
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Stream Network Enforcement}
+\label{sec:app_stream_network_enforcement}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The location of the stream network is enforced to ensure general agreement with the NWM network which is used for forecasting the streamflow inputs.
+The NHDPlusHR Beta BurnLineEvent layer is used to enforce stream locations in the NHDPlusHR workflow and best agrees with thalweg locations in the DEM used so it is also used here for hydro-enforcement \cite{moore2019user}.
+
+However, to better match the drainage density of the NWM FR V2.1 stream network, which is based on the NHDPlus V2, the BurnlineEvents are reduced in density using a nearest neighbor search around the NWM flowlines.
+Headwater points are first derived for the NWM FR V2.1.
+For every NWM headwater point, the nearest NHDPlusHR point is selected and placed into a set while those excluded are discarded.
+Only the nearest point on the NHDPlusHR is used so any portion of the NHDPlusHR network upstream of this nearest point is discarded to avoid extending inundation too far above the modeling domain.
+The points in this nearest neighbor set are then traversed downstream.
+Any headwater portion in the NHDPlusHR or any other stream not traversed are removed to better match the resolution and spatial locations of the NWM stream network and its headwater points.
+The resulting NHDPlusHR stream network of lower drainage density gets hydro-enforced in subsequent operations.
+This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the reduced NHDPlusHR network corresponds to the NHDPlusHR network at or downstream of NWM V2.1 headwater locations only.
+Additionally, the remaining NHDPlusHR headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
+This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents. 
+%
+\begin{figure}[H]
+\centering
+\includegraphics[scale=1.0]{figures/headwaters.jpg}
+\caption{This figure illustrates some of the datasets that result from the pruning of NHDPlusHR Beta BurnlineEvents (dotted black) to the stream density of the NWM FR V2.1 density (blue).
+The stream network used for forecasting, NWM FR V2.1, is of lower stream density than that of the NHDPlusHR which has better agreement with the thalweg locations in the DEM used.
+Thus, we opt to trim the NHDPlusHR network to match the general location and density of the NWM network.
+The nearest neighbor segment in the NHDPlusHR of each NWM headwater locations and the nearest point on that segment is determined to match the closest point to that of the NWM headwater.
+These points are then traversed downstream and any segments not traversed are removed.
+The resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR BurnlineEvents.}
+\label{fig:stream_density_pruning}
+\end{figure}
+%
+
+The reduced density stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
+The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
+The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop and buffer distance. 
+The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling an inverted triangular notch shape with a significantly lower elevation along the thalweg line only.
+In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop which were set to fixed values of 70 m, 1000 m, and 10 m, respectively, but available to the user via the parameter file.
+While the values for these parameters are critical to the inundation extents produced, especially for lower flow rates where bathymetric information has more influence, finding their optimal values for OWP FIM was not done since it was out of the main scientific scope of this article.
+Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
+\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement techniques especially when computational costs are considered. 
+Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
+Some of these drawbacks are addressed by additional conditioning techniques applied in the methodology.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Flow Directions and Flats Resolution}
+\label{sec:app_flow_direction_and_flat_resolution}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+To facilitate the generation of a connected stream network and its associated catchments from the conditioned DEM, the depression-filled DEM is used to derive connectivity in the form of D-8 flow directions.
+D-8 seeks to allocate a drainage direction for every pixel based on the adjacent eight pixel neighborhood with the steepest slope \cite{o1984extraction}.
+The horizontal component of slope is defined as one for the four neighboring pixels in the main cardinal directions while the intercardinal pixels are designated a horizontal component of $\sqrt{2}$ by means of the Pythagorean theorem.
+Flow directions are derived for non-depression filled regions trivially with the above procedure but to define connectivity for every grid cell the remaining flats corresponding to depression-filled cells must be resolved.
+
+Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment} where flats are resolved by incrementing elevations iteratively.
+Software implementations have developed means to partition the problem and resolve flats iteratively with communication across processes \cite{tarboton2009generalized,tesfa2011extraction,wallis2009parallel,tarboton2005terrain}.
+The excessive iteration and communication leads to poor computational performance which motivated further work on how to optimize flat resolution \cite{survila2016scalable,barnes2014efficient}.
+The established literature in this niche field of hydrology discusses how prevalent flats can be in given study areas and how difficult the problem is from both computational and hydrologic stand-points \cite{garbrecht1997assignment,tarboton2009generalized,tarboton2005terrain,survila2016scalable,barnes2014efficient,tesfa2011extraction,wallis2009parallel}.
+OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-02-11 manuscripts/owp_fim4_2022/reviews/second_review/responses_aristizabal_et_al_202302.tex
second review responses and changes submitted to coauthors 20230211

% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from 
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
% 
% Licensed under cc by-sa 3.0 with attribution required.
% See https://creativecommons.org/licenses/by-sa/3.0/
% and https://stackoverflow.blog/2009/06/25/attribution-required/

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{apacite}

% bibstyle
\bibliographystyle{apacite}

% import Eq and Section references from the main manuscript where needed
% \usepackage{xr}
% \externaldocument{manuscript}

% package needed for optional arguments
\usepackage{xifthen}
% define counters for reviewers and their points
\newcounter{reviewer}
\setcounter{reviewer}{0}
\newcounter{point}[reviewer]
\setcounter{point}{0}

% This refines the format of how the reviewer/point reference will appear.
\renewcommand{\thepoint}{P\,\thereviewer.\arabic{point}}

% command declarations for reviewer points and our responses
\newcommand{\reviewersection}{\stepcounter{reviewer} \bigskip \hrule
                  \section*{Reviewer \thereviewer}}

\newenvironment{point}
   {\refstepcounter{point} \bigskip \noindent {\textbf{Reviewer~Point~\thepoint} } ---\ }
   {\par }

\newcommand{\shortpoint}[1]{\refstepcounter{point}  \bigskip \noindent 
	{\textbf{Reviewer~Point~\thepoint} } ---~#1\par }

\newenvironment{reply}
   {\medskip \noindent \begin{sf}\textbf{Reply}:\  }
   {\medskip \end{sf}}

\newcommand{\shortreply}[2][]{\medskip \noindent \begin{sf}\textbf{Reply}:\  #2
	\ifthenelse{\equal{#1}{}}{}{ \hfill \footnotesize (#1)}%
	\medskip \end{sf}}

\begin{document}

\section*{Response to the reviewers}

\subsection*{Editor Response}

\noindent Associate Editor Evaluations: \newline
\noindent Recommendation (Required): Return to author for major revisions \newline
\noindent Accurate Key Points: Yes \newline

\noindent Associate Editor (Remarks to Author): \newline
\newline
\noindent Dear authors,

Your manuscript has been re-assessed by two of the original reviewers and one new reviewer.
The two original reviewer acknowledge the improvements made with respect to the earlier version and point out a few additional areas where further improvements are needed.
The new reviewer stresses that some further major revisions in terms of level of detail and storytelling are needed and requests some improvements with respect to model evaluation.
I would like to give the authors another opportunity to address both the major and minor concerns and am looking forward to reading a revised version of the manuscript. \newline
\noindent Best regards, \newline
\noindent Your AE

\begin{reply}
We thank you, the editor, as well as the reviewers for carefully reading the manuscript and for providing valuable feedback to improve its quality.
We have gone through the reviews and made every effort to address any specific concerns while staying within the scope of the paper.
We hope this revised version of the manuscript meets the standards for publication and addresses the requests from the reviewers.
In instances where we disagreed with the suggestions, we have provided a well-reasoned explanation for why we felt the change was not necessary.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\reviewersection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Introductory Comment}

% Point one description 
\begin{point}
\noindent Recommendation (Required): Return to author for minor revisions \newline
\noindent Significant: Yes, the paper is a significant contribution and worthy of prompt publication. \newline
\noindent Supported: Yes \newline
\noindent Referencing: Yes \newline
\noindent Quality: Yes, it is well-written, logically organized, and the figures and tables are appropriate. \newline
\noindent Data: Yes \newline
\noindent Accurate Key Points: Yes \newline

\noindent Reviewer \#1 (Formal Review for Authors (shown to authors)):

I have read the revised version and want to thank the authors for addressing all questions and concerns I raised in the first round.
As a general note, it would be very helpful if authors could be specific as to where the manuscript has been changed (providing line numbers in the revised version).

The proposed approach (computing HAND based on level paths instead of nearest drainage) is now better clarified in the methodology.
The newly added Fig. 7 helps readers understand the concept and novelty of the developed approach. 
The results and discussion sections are extended, and they comprehensively outline the method's limitations.
Most of my comments were addressed, and I feel the revised manuscript is suitable for publication in WRR.
I recommend, however, that the manuscript be returned to the authors for the following minor edits before publication.
\label{pt:intro_1}
\end{point}

% Our reply
\begin{reply}
Thanks for going through our paper for a second time and providing numerous suggestions for improvement.
We were glad that you found that all of the original questions and concerns were properly addressed.
We hoped to have adequately answered your final minor suggestions properly as well.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Comments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
I believe the last part of the Fig. 9 caption, starting with "The BL stream network is also shown which is denser than ... This creates additional ..." should also be added to the discussion section as one of the sources of uncertainty or errors.
Perhaps somewhere between lines 1025-1039.
\label{pt:comment_1_1}
\end{point}

\begin{reply}
Thank you for identify an important source of error that we have identified and suggesting it go within the Discussion section.
Based on your suggestion, we have added a few sentences that better discusses the conflation errors that could occur by using BLE associated streamflows with the NWM network.
These sentences we placed near the end of the paragraph you suggested the discussion be located (INSERT LINE NUMBERS PRIOR TO SUBMISSION). 
\end{reply}

\begin{point}
The level path approach leads to delineating catchments with larger sizes.
The authors made this point discussed in the discussion section.
However, there is less discussion about the stream lengths in the discussion section.
For example, what are the average, minimum, and maximum sizes for the stream length, L, within these large catchments?
This is important because L impacts the stage-discharge relationship.
As far as I know, previous work conducted some sensitivity analysis to show a proper range for L in HAND calculation (see references below). 
\begin{itemize}
\item Godbout, L. D. 2018. "Error assessment for Height Above the Nearest Drainage Inundation 935 Mapping", Master of Science in Engineering thesis, The University of Texas at Austin, 936 https://repositories.lib.utexas.edu/handle/2152/68235.
\item Zheng, X., Maidment D. R., Tarboton D. G., Liu Y. Y., \& Passalacqua P. (2018a). GeoFlood: 1054 Large-Scale Flood Inundation Mapping Based on High-Resolution Terrain Analysis. 1055 Water Resources Research, 54. https://doi.org/10.1029/2018WR023457
\end{itemize}
\label{pt:comment_1_2}
\end{point}

\begin{reply}
Thank you for pointing out this important parameter in the behaviour of synthetic rating curves (SRCs).
We agree that SRCs are sensitive to a variety of values and one of which as you have pointed out is reach length.
The two citations you provided reviewed or mentioned this issue in detail so we did cite these references throughout the paper.
While cited throughout the paper for various reasons, they were used in the context of reach length on or near lines 150 and 492 of the previous submission (INPUT LINE NUMBERS PRIOR TO SUBMISSION).
On line 490 of our previous submission, we mentioned that the reach length is capped to 1.5 km which is consistent with the findings from the Godbout paper and also consistent with what Zheng used within GeoFlood.
Additionally, we kept this constraint for all versions of HAND that we computed including FR, MS, and GMS, thus we did not observe a notable difference in reach length across the three versions.
Due to this, we did not include it in the discussion of the paper.
Nevertheless, we decided to add a sentence on (INPUT LINE NUMBERS PRIOR TO SUBMISSION) acknowledging that this parameter was held fixed at 1.5 km for all versions of HAND.
This constraint is very important since we did not find a significant difference in the reach lengths.
To demonstrate this we computed the averages and standard deviations for the two primary versions FR and GMS. 
We found GMS to have a mean derived reach length of 1398.1 km with a standard deviation of 50.2 km.
The FR version had a mean of 1354.8 km with a standard deviation of 249.7 km.
We found this difference in reach lengths across FR and GMS to be statistically insignificant. 
In accordance with this, we added these numbers specifically (INPUT LINE NUMBERS PRIOR TO SUBMISSION) to the discussion section to inform readers that reach length, albeit an important variable, does not contribute to variance in SRC values across HAND versions.
\end{reply}

\begin{point}
The first paragraph of the conclusion can be removed as it is more like an introduction of the flood inundation mapping and HAND method. 
\end{point}

\begin{reply}
Based on your suggestion, we completely removed the opening paragraph of the conclusion.
Thanks for providing specific feedback to help reduce some of the superflous content.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Begin a new reviewer section
\clearpage
\reviewersection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Introductory Comment}

\begin{point}
\noindent Reviewer \#2 Evaluations: \newline
\noindent Recommendation (Required): Publish in present form \newline
\noindent Significant: Yes, the paper is a significant contribution and worthy of prompt publication. \newline
\noindent Supported: Yes \newline
\noindent Referencing: Yes \newline
\noindent Quality: The organization of the manuscript and presentation of the data and results need some improvement. \newline
\noindent Data: Yes \newline
\noindent Accurate Key Points: Yes \newline

\noindent Reviewer \#2 (Formal Review for Authors (shown to authors)): \newline

The main concerns from my last round of review have been addressed in this revision.
As I mentioned last time, there is significant value to expose operational research approaches adopted by federal agencies to general academic audiences for broad discussion and rigorous review.
Therefore, I recommend accepting this paper.
Some issues that still need to be fixed are listed below:
\label{pt:intro_2}
\end{point}

\begin{reply}
We appreciate the time and energy dedicated in producing your reviews as they have greatly enhanced the quality of our manuscript.
We have gone through your latest round of suggestions and feel we have addressed them as best as possible.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Comments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
Figure 5 It's better to pick a headwater basin for illustration. 
The current spot is still a little misleading: the upstream purple link with a label of 1667 is assigned as a part of the mainstem LP.
However, in this figure, it's shorter than the yellow branch above with labels like 15 and 7.
If this is the full network, I think that the yellow branch should be chosen as a part of the mainstem LP, instead of the current one. 
\end{point}

\begin{reply}
Based on your advice, we elected to highlight a small headwater subset of the stream network in the HUC8 12020002.
This subset avoids inheritance issues from using nested HUCs.
\end{reply}

\begin{point}
Figure 6 (a) The symbology needs some extra work. 
The two demo LPs should be highlighted with a thicker line.
\end{point}

\begin{reply}
Based on your comment, we increased the line weights of the stream lines and catchments to make them more visible.
We hope this thickness stands out better to the readers.
\end{reply}

\begin{point}
Figure 8 Remove the extra axis stickers on the upper boundary.
\end{point}

\begin{reply}
Thanks for the attention to detail.
The extra tick marks have been removed.
\end{reply}

\begin{point}
Figure 8 9 12 It's better to present results at the same site throughout all the sections, instead of zooming in and out to different parts of your study area.
The authors show cross-sections at one site and then jump to another location or zoom level to show the inundation extent.
There is no intrinsic connection between these plots in this case. 
\end{point}

\begin{reply}
Please note for the response below that Figures 8, 9, and 12 correspond to 7, 8 and 10 now, respectively, as some changes were made in the course of this review.
The original Figure numbers are referenced below.

After giving this some thought, we have elected to maintain the same locations for the three suggested figures.
We do appreciate your concern to maintain consistency across plots which could, in a way, facilitate understanding for readers as their context would stay more fixed.
However, we believe these three figures convey different concepts and ideas, thus, are best represented by utilizing separate sites and scales.
To be specific, Figure 8 is meant to convey the spatial extents of the study domain. 
In order to do this, we must take a very high-level view of an entire region of the US.
In this large scale, the central tenants of Figures 9 and 12 would not be able to be properly conveyed.

Figure 9 is meant to illustrate a cross-walking technique for assigning recurrence flows at two intervals, 100 and 500 year, from HEC-RAS derived BLE cross-sections.
These cross-sections are intersected with the NWM reaches and the intersecting flow values are aggregated utilizing a median filter.
This site was selected because it illustrates two common sources of error with this technique.
One is the lack of certain headwaters represented in the BLE but not the NWM stream network.
This leads to FNs in the HAND inundation when compared to BLE extents.
Another source of error is the fact that cross-sections could span stream segments of varying stream order, thus have large differences in flow rates.
While most of these are addressed by taking the median value, some may suffer from errors which could violate mass conservation assumptions.
We wanted to pick a site that graphically conveyed our method that does reduce hydro-climatic uncertainties but also be true to the limitations of the technique.
We don't feel that representing this in the same map frames as Figures 8 or 12 would convey this message to the readers.

Figure 12 is meant to convey the impact of computing HAND with overlapping, reduced stream order spatial units on the extents of the resulting FIM.
Due to the nature of only accounting for a single fluvial source within HAND, this problem is most evident in areas with very large mainstems with high flows that are directly fed by lower order and flow tributaries.
When this occurs, it creates a significant bottle-necking effect in the inundation because the inundation from the mainstem cannot influence the inundation within the neighboring catchments tributaries.
For this reason, we wanted to select this region due to its high flow mainstem and many lower flow tributaries that impede inundation with the HAND method.
By computing HAND independently for each level path and mosaicing the inundations together, we are able to account for inundation from multiple fluvial sources.
This region conveys that concept well in terms of inundation extents which is the main focus of the paper.
Conveying this idea within the map frame of Figure 8 would not give the reader a close enough look as to what is actually happening.
The mapframe in Figure 9 would, indeed, illustrate a positive change in the spatial extents but due to the lower stream order nature of this region it would not have the same impact as utilizing the higher order region in the current Figure 12 map frame would.

We hope that this explanation is satisfying to you as we are just trying to convey different concepts with each figure and felt that utilizing different regions that best convey the messages would be most appropriate.
We do appreciate that switching across regions could be cognitively expensive for the reader but we hope that this price is rewarded by ensuring our explanations are complete. 
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Begin a new reviewer section
\clearpage
\reviewersection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Introductory Comment}

Did not repeat review.
Skip ahead to reviewer \#4 below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Begin a new reviewer section
\clearpage
\reviewersection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Introductory Comment}

\begin{point}
\noindent Reviewer \#4 Evaluations: \newline
\noindent Recommendation (Required): Return to author for major revisions \newline
\noindent Significant: The paper has some unclear or incomplete reasoning but will likely be a significant contribution with revision and clarification. \newline
\noindent Supported: Yes \newline
\noindent Referencing: Yes \newline
\noindent Quality: The organization of the manuscript and presentation of the data and results need some improvement. \newline
\noindent Data: Yes \newline
\noindent Accurate Key Points: Yes \newline

\noindent Reviewer \#4 (Formal Review for Authors (shown to authors)): \newline

This manuscript proposes a new version of HAND-based FIM that accounts for multiple sources of fluvial inundation and results in an improvement in overall FIM skill.
The topic of this manuscript for improving the FIM skill at such a large scale is of interest to the flood community and adds value to the literature.
However, I have some main concerns mostly regarding how it has been written for a scientific journal and the validation parts.
Since I was not the reviewer in the previous round, I decided to only read the revised version and add my comments.
Later on, I checked the reviewer comments in the previous round and interestingly I found several common concerns that have not been fully addressed yet.
Please see my two major comments and the following minor ones below:
\label{pt:intro_3}
\end{point}

\begin{reply}
Thank you for reviewing the paper and for providing feedback to help improve its content.
We hope we have interpreted your requests correctly and assumed the correct solution to many of the problems you noted.
Every effort has been made to make sure your corrections have been well addressed.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Major Comments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
The language of this manuscript, especially the method section is too detailed and technical, it includes many jargons, and is long for a scientific journal.
More specifically, I would like to point to sections 2.3 and 2.4. 
\end{point}

\begin{reply}
We have gone through your sub-comments below and have made substantial changes to reduce detailed jargon.
Please note that some of these details were added in response to comments from previous reviewers looking for specific, technical clarifications on detailed issues.
We hope these changes streamline the paper while also maintaining the details necessary to communicate the differences of our approach compared to previous efforts.
\end{reply}

\begin{point}
Section 2.3.1 includes too much detailed information that may not be that much beneficial for the majority of the Journal audience who are not experts in this area.
I don't think these detailed technical guidelines are required to be in the main body of the manuscript as it may make the readers reluctant to follow the rest of the manuscript.
In addition, since the main novelty and contribution of this manuscript is section 2.5, it would be more reasonable if the authors focus on this section as the core of the method section and make other parts as concise as possible.
My suggestion is that the authors move additional information to the appendix so those specific audiences who are interested can follow details such as DEM preprocessing, etc.
Similarly, sub-sections 2.4.1, 2.4.2, and 2.4.3 can be joined under a concise section and additional information can be moved to the appendix.
Figure 3 can be also moved to the appendix. 
\end{point}

\begin{reply}
Thank you for suggesting specific changes that could potentially lighten the paper.
As per your suggestion, we have greatly reduced the content of Section 2.3.1 and moved most of the details to the Appendix.
We also felt the same about Section 2.4.1 as most of the details are only of concern to a specific subset of the community that deals with the computational aspects of hydro-conditioning operations.
We greatly condensed this section and moved the remainder to the Appendix.

As per Section 2.4.2, the other reviewers were very keen as to how the stream network was defined and how the reach lengths were computed.
These concerns were rightfully brought up as previous research has shown how reach averaged SRCs are sensitive to sampling which is the reach length parameter in this context.
Much effort was taken to ensure these concerns by these experts were properly addressed and wish to ensure these details are stated in the manuscript.
Several other points in the paper including some in the Discussion section allude to this important parameter that the community has identified.
Due to this reason, we wish to maintain this section in the paper.

Lastly with Section 2.4.3, derivation of the catchments is a simple procedure and not much background is stated here.
Most of the section is dedicated to explaining pieces of information that make the rest of the manuscript understandable.
This includes the fact that two sets of catchments are derived including the pixel scale and reach scale catchments.
Without this, the reader would be lost in other parts of the paper.
This section is also very short so moving a few sentences to the appendix would not make much sense.
For these reasons, we also elected to keep this section in place.
\end{reply}

\begin{point}
Using too much jargon make the paper hard to understand and follow by the majority of readers who are not a technical expert in this area.
I always encourage the authors to avoid these terminologies as they adversely affect the quality of the manuscript.
Some examples are "etch or burn flowlines" (line 210), "BurnlinEvents are pruned" (line 351), and "The elevations found in the NLD are burned onto the DEM" (line 399). 
\end{point}

\begin{reply}
Based on your feedback, we have elected to remove the terms ``etch'' and ``burn'' in favor of the term ``enforce'' which we feel has a broader understanding and less likely to be taken as geospatial jargon.
The term ``pruned'' was either removed or replaced with more generally acceptable words.
Lastly, ``BurnLineEvents'' was maintained as it is a term inherited from the NHD to differentiate it from other stream networks contained within the NHDPlusHR.
We did wish to maintain technical consistency with the source datasets.
\end{reply}

\begin{point}
The amount of text provided under the method section is too much.
It is highly recommended to keep a balance among the lengths of different sections of a manuscript.
As an example, the manuscript has 12 figures while the first 9 figures belong to the method section and only 3 figures have been developed for the result section.
Typically we expect to see more figures in the result section. 
\end{point}

\begin{reply}
Thanks for your observation noting the imbalances in the sections of the paper.
Due to your previous comments suggesting specific changes, we have greatly reduced the amount text and figures within the methods sections.
Due to your suggestions, we have moved two figures out of the methods and into the introduction or the appendix.
With those figures a considerable amount of text has been removed from this section which we believe better balances the paper out.
Our methodology section is now greatly reduced albeit still dense in order to adequately explain the many techniques employed in developing continental scale FIMs, addressing fundamental limitations with HAND, and validating the proposed enhancements.
Thanks for your specific suggestions addressing ways to reduce information within the methods section and the paper overall.
\end{reply}

\begin{point}
The evaluation approach used to validate the effectiveness of the proposed HAND-based technique for FIM is missing some important components:
\end{point}

\begin{reply}
Please see responses to specific points below.
\end{reply}

\begin{point}
The reference data used in this manuscript is FEMA flood extent maps corresponding to synthetic storms (100-year and 500-year floods).
However, the authors claim that the application of their method is for real-time flood inundation mapping.
In general, the mechanisms and approaches used for detecting flood extent corresponding to synthetic storms differ from those used for real-time flood inundation mapping.
If the authors want to show that their approach is beneficial for real-time flood inundation mapping, they need to demonstrate this by simulating real flood events (not synthetic cases).
Following this concern, I found a meaningful portion of the introduction, discussion and conclusion refer to the operational forecasting by NWM streamflow forecasts and AHPS systems.
However, the manuscript has not used any of these forecasts provided by NWM or other operational agencies.
To me, this is an inconsistency between what has been implemented (synthetic storm simulations) and what is mentioned as the goal of this method (forecasting real flood events by NWM).
\end{point}

\begin{reply}
We appreciate your concerns regarding the use and evaluation of the specific FIM model.
We would like to take the opportunity to clearly articulate our reasoning and provide a more comprehensive explanation. We are confident that our clarifications and proposed revisions to the manuscript will effectively address your issues, while preserving the core objectives of the paper.

As outlined in the sub-section titled "OWP FIM" in the Introduction, there are two main objectives of the paper.
The first objective is to introduce, explain, and assess a novel HAND implementation that is being utilized for operational flood forecasting alongside the NWM.
The second objective is to identify a limitation with HAND-based FIMs, present a new solution, and evaluate its effectiveness by quantifying and analyzing the improvements.
Providing a detailed explanation of the first objective provides the reader with the necessary background to understand the motivation and context behind this model's creation.
It is important to note for the hydrology community that this model is considered operational, as they may find it relevant for future work.
However, we acknowledge that referring to the model as operational and then evaluating it using risk-based, synthetic data may cause confusion, and we aim to clarify this.

It is correct to state that the process of creating a synthetic FIM differs from the process of producing a FIM for operational forecasting when the entire hydro-climatic process is taken into account.
The key difference is that HAND operates somewhat independently of this process, as it couples with hydro-climatic processes through catchment scale depths.
By coupling HAND with SRCs, we have the ability to accept streamflows, which serve as the main point of connection with the NWM and any other streamflow sources we may use for retrospective, monitoring, forecasting, or evaluation purposes.
This makes HAND somewhat exclusive of the hydro-climatic processes modeled ``upstream'' and focuses solely on computing inundation depths and extents.
HAND is agnostic to the nature of the streamflow values it receives, be it synthetic or operational, and only concerns itself with the magnitude of the streamflow values.

It's crucial to examine multiple sources of flood extent information for evaluations as each has its own limitations and level of uncertainty.
The sources examined include using high water marks, remote sensing, and hyper-local flood inundation models.
It's important to acknowledge the limitations and advantages of each source.
For instance, agencies like the USGS create FIMs by combining local HWMs with a DEM.
HWMs and FIMs based on a cluster of HWMs can serve as a reliable source of validation data for large-scale, operational models.
RS-based FIMs can also provide valuable validation information, depending on factors such as the type of sensor, the flooded environment, atmospheric conditions, and more.
FIM models are often compared to RS-based inundation extents in academic literature as they provide information where no other source of observed or modeled data could be available.
Additionally, hyper-local flood inundation models, such as those produced by the AHPS mentioned in the paper and by the USGS' Flood Inundation Mapper program, also provide high-quality information that can be used to validate large-scale inundation proxies like the OWP FIM.
All of these methods are potentially suitable for operational evaluation as they reflect real-world flood events but also subject to limitations themselves.

In order to use the three alternative sources of benchmark data mentioned previously in an operational capacity with HAND, one must utilize operational streamflow values.
These streamflow values would come from the NWM if we are looking to evaluate operational forecasts.
If operational observation data were considered, we could have used gage information.
Utilizing the NWM offers a dense availability of streamflow data in the forecasting context that OWP FIM would be intended to be used for in an operational setting.
Using NWM streamflow data would allow us to account for multiple fluvial sources of inundation which would be required to evaluate our technique.
However, the limitation of using NWM streamflows would be introduction of hydro-climactic uncertainties that would add error and variations in our analysis outputs. 
This could modify, dilute, or add noise to the effects of stream order reduction in order to account for multiple fluvial sources with HAND.
Another aspect to take into account is that relying on observational data from gages to obtain streamflows would reduce the number of streamflow values accessible in a specific area, thereby spatially restricting the observations.
A similar problem would arise by using the point stages or streamflows that hyper-local FIMs offer.
These single point samples of streamflow data would lack the network density required to assess the performance of a proposed technique interested in accounting for multiple fluvial sources of inundation with HAND.
This would likely lack sufficient information to properly account for cases presented throughout the paper where low flow tributaries converge with high flow, higher order streams.
These two issues brought about by using NWM streamflows and single point streamflows from modeled or observed gages deterred us from using these methods to evaluate our version of HAND and its proposed improvements.
Even though we found these three sources to be less than optimal for our purpose, we don't advocate ignoring them. 
We intend to use them for on-going and future evaluations of forecasting capabilities as they offer advantages themselves, but for the time being we deemed the BLE to be a good evaluation dataset for HAND based FIMs and the principal modification proposed.

Based on our analysis, we concluded that the BLE, a synthetic dataset, was the most suitable option for our study.
By using the BLE and its accompanying streamflow data, we were able to exclude the hydro-climatic uncertainties that would have been present if we used modeled streamflow values from other sources.
Moreover, the ample supply of streamflow data from the BLE enabled us to accurately evaluate the impact of our suggested modifications for multi-fluvial sourced flooding with HAND.
After conducting a literature review, we discovered several studies that employed synthetic datasets to evaluate models for operational purposes \cite{wing2017validation,hu2021real,li2022accounting,li2022comprehensive,zheng2018geoflood,wing2017validation,afshari2018comparison}.
Consequently, we decided to utilize the BLE for our study, however, it is crucial to acknowledge the requirement for continual evaluations using a range of benchmark datasets across broad geographic areas.
This is a challenging matter that calls for sustained efforts from various viewpoints, and we therefore appreciate and consider your feedback as valuable.

We hope we alleviated any concerns about the suitability of using a synthetic dataset, the BLE, for evaluating a model intended for operational forecasting, OWP HAND based FIM.
We concur that the paper did not provide a logically sound explanation for the conclusion to employ a synthetic evaluation for a model intended for operational use.
It appeared that the appropriate location to offer this explanation was in the first paragraph of the Evaluation section.
The current explanation appeared insufficient, so we significantly strengthened it to ensure that the reader fully understands the benefits of using the BLE and the limitations avoided by not using alternative benchmark datasets.
Hopefully this revised paragraph conveys our justification in a concise manner.

We conducted a thorough review of the manuscript and looked for references to operational forecasting especially in the sections you mentioned which were the Introduction, Discussion, and Conclusions.
The only reference to operational forecasting in the Conclusions section was found in the opening paragraph, which was removed upon request from another reviewer as it was considered redundant in summarizing the paper.
In the Discussion section, we found no explicit references to operational forecasting as the focus of the section was to examine the impact of the primary modification proposed for HAND.
Some room was allocated in the section to address other sources of error and to present future possibilities for improvement in this field.
In the Introduction, two sub-sections, titled ``Operational Forecasting'' and ``National Water Model,''  discuss operational capabilities in detail.  
As previously mentioned, the purpose of this implementation of HAND is for operational forecasting, making it crucial to provide context on past operational systems and the NWM.
We defend the presence of the sub-sections titled ``Operational Forecasting'' and ``National Water Model'' by highlighting the rationale behind them:
\begin{itemize}
\item Introduce legacy forecasting systems including their limitations in the aspects of spatial and temporal coverage.
\item Convey the fact that the NWM fills in these gaps by greatly expanding the forecasting points in both space and time.
\item Acknowledge the value of assimilating legacy forecasts into NWM forecasts.
\item Introduce the term hydrofabric which is the collection of geospatial datasets used for NWM purposes including catchments, stream network, and reservoirs.
\item Mention the use of the two stream networks with varying density including the full resolution (FR) and its subset mainstems (MS) which is defined as all FR streamlines at or downstream of AHPS forecasting points.
\item This network also allowed for computing HAND with a stream network of lower drainage density which demonstrated our multiple fluvial inundation technique for 4\% of the modeling domain.
\end{itemize}   
We acknowledge that some readers would not be interested in learning details about these authoritative forecasting systems but many readers would.
One of your review questions below asked for some details on the AHPS for which we re-included previously removed information.
These sections also included significant background on the hydrofabric which serves as the geospatial backbone of both the NWM and OWP HAND FIM.
We wish to pay respect to these key items that provide the background for the community while hopefully not taking away too much from the main scope of the paper.
In lieu of this, we opted not to remove any more information concerning references to operational forecasting.
We hope we have offered sufficient justification as to why.
\end{reply}

\begin{point}
The authors should clearly explain if their method is only used for flood extent mapping (binary mapping) or if it is useful for flood depth mapping.
According to Eq. 4, this method can be used to estimate the flood depth as well.
Thus, some metrics (such as RMSE) that show the effectiveness of this method for flood depth mapping are required and should be added to the manuscript.
The reference data for this evaluation could be High Water Marks (HWMs).
The current validation approach only tests the effectiveness of the proposed method for flood extent mapping. 
\end{point}

\begin{reply}
According to previous work, HAND is technically able to produce both inundation extents and depths \cite{nobre2016hand,maidment2017conceptual,garousi2019terrain,li2020evaluation,li2022accounting,li2023comparative}. 
We want to acknowledge the capabilities of HAND by pointing out that it can produce both, and by thresholding the depths, we can obtain extents.
However, the OWP is only producing FIM extents as a product with only plans in the future to offer depths as well.
For this reason, FIM extents is of principal interest to OWP and of primary concern in this study.
We acknowledge the importance of flood depths and understand that extents and depths have a nuanced relationship.
We certainly value the need to evaluate depths moving forward but do to constraints we have opted to ensure that the methods proposed here were evaluated for extents.
Additionally, there are a number of studies with HAND and other FIM techniques that are able to produce both spatially distributed extents and depths or elevations but only chose to evaluate extents \cite{garousi2019terrain,li2023comparative,li2022accounting,papaioannou2016flood,mcgehee2016modified,zhang2018comparative,shastry2019using,follum2017autorapid,wing2017validation,aristizabal2021mapping,rahimzadeh2019evaluating,amarnath2015modelling,wood2016calibration,hooker2022spatial,tansar2020flood}.
In the future, we hope to produce more evaluations to cover depths and probabilistic information but for now constrain the scope of this study to only include extents for evaluation purposes.

In terms of the content of the manuscript, we wish to ensure that your concern is properly accounted for.
There is a need to convey to the readers that we are only evaluating our methods on FIM extents and not depths.
It is also important to convey that although related effects on FIM extents don't translate the same to FIM depths.
Evaluating FIM depths would make the study more robust but was simply left out of scope due to the many other topics to cover and for organizational priorities.
To account for these key points, we have removed any superfluous mentions of FIM depths that don't tie into the fact that they are computed prior to achieving extents.
Additionally, we have added some notes in the Discussion section that acknowledge the limitation in our evaluation methodology by not including depths.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Minor Comments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{point}
Lines 84 and 109: What are the hydrologic/hydrodynamic models used by AHPS for streamflow forecasting and flood inundation mapping?
Are AHPS more accurate and reliable than NWM forecasts?
\end{point}

\begin{reply}
We previously had a deeper explanation on the AHPS in Section 1.1 and removed most of the content due to it being only related to the scope of flood inundation mapping with HAND.
This was partly done in response to requests to remove irrelevant content from previous reviewers.
In lieu of your questions, we have elected to maintain the current explanation while adding a shorter sentence to use the references we cited originally.
This should give the reader adequate information to pursue if more information is required on their behalf.
With regards to your second question, yes, the Office of Water Prediction does indeed consider AHPS to be of high quality.
This is alluded to on L108-112 of the original document that mentions that AHPS forecasts are assimilated in the NWM by replacing the NWM at those respective locations and routing that flow downstream.
We termed this assimilation to be incorporating the best available regional scale forecasts.
We hope that this sufficiently answers your questions while also not adding too much more to the paper.
\end{reply}

\begin{point}
Figure 1: AHPS Location and NWM FR Streams provided in the legend is hardly visible and distinguishable in the figure.
Please give a better illustration. 
\end{point}

\begin{reply}
This figure is challenging because it attempts to convey many pieces of fine scale information across very large spatial scales. 
We have iterated on this a few times trying to optimize the illustration to best account for the shear scale and density of information.
Previous reviews have stated that the underlying networks were too thin and needed to be thickened. 
This has possibly drowned some of the points.
We value your feedback and make an attempt to revise the figure to see if we can best account for everything just a little bit better.
In our latest effort which is motivated by your feedback, we have elected to expand the image as to take the maximum advantage of limited space in the horizontal direction.
Hopefully, this gives the reader more scale to operate with.
The NWM MS stream lines were transitioned to green while the AHPS locations without FIM were transitioned to red.
This is likely to contrast with one another more.
Additionally, we amplified the sizes of the legend objects so the readers can clearly identify what the symbols should be on the map.
We hope this representation best communicates all the datasets better given the steep constraints.
\end{reply}

\begin{point}
Lines 148-152: In one of the recent comparisons, Hocini et al., (2020) compared the performance of the HAND method with 1D and 2D hydrodynamic models and showed that HAND performance is weaker in many cases. 
\begin{itemize}
\item Hocini, N., Payrastre, O., Bourgin, F., Gaume, E., Davy, P., Lague, D., Poinsignon, L., Pons, F.,
2020. Performance of automated flood inundation mapping methods in a context of flash
floods: a comparison of three methods based either on the Height Above Nearest Drainage
(HAND) concept, or on 1D/2D shallow water equations. Hydrol Earth Syst Sci Discuss
Httpsdoi Org105194hess-2020-597
\end{itemize}
\end{point}

\begin{reply}
Thank you for furnishing this important and relevant reference.
Given that this method evaluates and compares the use of HAND with models that have higher order physics, we found it necessary to include in this paper.
It is now cited in multiple locations.
\end{reply}

\begin{point}
Lines 444-445: Following my major comment, I think the majority of readers would like to know what HAND, SRCs, and the cross-walk table are. 
However, the main focus is on detailed information about geospatial analysis used to get DEM, streamlines and etc.
\end{point}

\begin{reply}
Due to the ambiguity of these terms as they are stated on the lines you referenced, we have added a few sentences there to help explain these three geospatial objects before they are deep-dived later in the section.
\end{reply}

\begin{point}
Lines 512-515: This is not a straightforward approach to explaining HAND calculation.
There are many better ways to explain it so that everyone can understand it.
Please reword it.
\end{point}

\begin{reply}
We have reworded large portions of this paragraph hoping to more clearly explain how HAND is computed while also staying true to the level of detail requested by previous authors.
Specifically it was asked to elaborate on the use of the native elevations outside of the AGREE conditioned regions.
We did our best here so if you have any specific suggestions those would be appreciated and help us better address your concerns.
\end{reply}

\begin{point}
Lines 539-543: what is known and what is unknown in this equation?
Are we trying to estimate Q or y?
Please add more information.
\end{point}

\begin{reply}
Thank you for asking for clarification on this. 
In practice, we pick discretized stage values and compute the discharge for the synthetic rating curve.
The last sentence of this section describes that we elect 84 samples between 0 and 25 meters in increments of 1/3 m.
In order to avoid ambiguity, we have elected to move this sentence to the end of the paragraph immediately proceeding after equation 1.
We hope this better conveys how we constrain the equation to compute discharge for SRCs.
\end{reply}

\begin{point}
Lines 569-571: I think several parameters of Manning's equations are dependent on the value of stage so depending on any of these 84 stage values we have different parameters.
Please clarify it in the text here.
\end{point}

\begin{reply}
Based on your suggestion, we have elected to add a sentence at the end of this paragraph that explicitly states that the parameters V(y) and B(y) which are functions of the stage, y, also change when the stage changes.
\end{reply}

\begin{point}
Lines 609-612: This is the main novelty and contribution of this manuscript.
I would recommend the authors move this information (texts and figures) to the introduction and specifically mention that this is the problem they are seeking to address.
It is better that transfer motivation and objectives to the beginning of the manuscript and only explain the method in this section. 
\end{point}

\begin{reply}
Due to your recommendation and to better motivate the study, we moved most of the discussion you elected to the last subsection of the Introduction titled ``Office of Water Prediction Flood Inundation Mapping.''
The paragraph you alluded to introducing our scale reduction techniques has been greatly reduced and, in doing so, removed a Figure from the already dense methods section.
\end{reply}

\begin{point}
Line 619: Please specify what unary stream order networks are.
It seems this is a key point.
\end{point}

\begin{reply}
Good catch.
Based on the fact that this is now the only use of the term unary, we removed its use in lieu of using ``stream networks of unit stream order'' which is more generally accepted.
\end{reply}

\begin{point}
Lines 629-631: Please reword this part.
It is not clear whether this method aims to improve the computational demand or improve the accuracy by reducing the negative effects of nearest drainage or both.
\end{point}

\begin{reply}
Thanks for catching this ambiguity.
We have adjusted to specify that we intend these techniques to enhance the quality of FIM extents not improve computational performance.
\end{reply}

\begin{point}
Line 761: Associated streamflow data.... Like what data?
\end{point}

\begin{reply}
We decided to remove the term ``associated'' in favoring the phrase ``corresponding to the flow peak'' which adds specificity to the term ``streamflow data.''
\end{reply}

\begin{point}
Line 931: Figure 12 is referenced before Figure 11 so please replace their order.
\end{point}

\begin{reply}
Another good catch.
We have swapped the order of the figures thus Figure 12 is now Figure 11 and vice versa.
Please note that the original Figure numbers on these have changed.
\end{reply}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% references and end document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage % this clears figures before references
\bibliography{responses}

\end{document}

### 2023-02-11 manuscripts/owp_fim4_2022/sections/appendix.tex
second review responses and changes submitted to coauthors 20230211

--- 2022-09-27 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-%
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements, including customized APA reference formatting.
-%
-% You may copy this file and give it your
-% article name, and enter your text.
-%
-%
-% Step 1: Set the \documentclass
-%
-%
-
-%% To submit your paper:
-\documentclass[draft]{dependencies/agujournal2019}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
-\usepackage[inline]{dependencies/trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-\usepackage{rotating}
-
-%% added
-\usepackage{multirow}
-\usepackage{amsmath}
-\usepackage{float}
-
-\linenumbers
-
-%%%%%%%
-% As of 2018 we recommend use of the TrackChanges package to mark revisions.
-% Add track changes editors
-\addeditor{Fernando Aristizabal}
-
-% The trackchanges package adds five new LaTeX commands:
-%
-%  \note[editor]{The note}
-%  \annote[editor]{Text to annotate}{The note}
-%  \add[editor]{Text to add}
-%  \remove[editor]{Text to remove}
-%  \change[editor]{Text to remove}{Text to add}
-%
-% complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%
-
-\drafttrue
-
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-% ie, \journalname{Water Resources Research}
-
-\journalname{Water Resources Research}
-
-% starts document
-\begin{document}
-
-%% Title and Authors
-%% ------------------------------------------------------------------------ %%
-%  Title
-%
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
-
-% Example: \title{This is a test title}
-%
-%\title{Reducing Horton-Strahler Stream Order Can Enhance Flood Inundation Mapping Skill with Applications for the U.S. National Water Model}
-\title{Extending Height Above Nearest Drainage to Model Multiple Fluvial Sources in Flood Inundation Mapping Applications for the U.S. National Water Model}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
-
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-%
-%
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
-%
-%
-\affiliation{1}{Lynker, Leesburg, VA, USA}
-\affiliation{2}{Center for Remote Sensing, Agricultural and Biological Engineering Department, University of Florida, Gainesville, FL, USA}
-\affiliation{3}{National Water Center, Office of Water Prediction, National Weather Service, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
-\affiliation{4}{Hydrologic Applied Innovations Lab, New York Water Science Center, United States Geological Survey, Troy, NY, USA}
-\affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Maryland, College Park, MD, USA}
-%
-%(repeat as many times as is necessary)
-%
-%% Corresponding Author:
-% Corresponding author mailing address and e-mail address:
-
-% (include name and email addresses of the corresponding author.  More
-% than one corresponding author is allowed in this LaTeX file and for
-% publication; but only one corresponding author is allowed in our
-% editorial system.)
-%
-% Example: \correspondingauthor{First and Last Name}{email@address.edu}
-%
-\correspondingauthor{Fernando Aristizabal}{fernando.aristizabal@noaa.gov}
-
-%% Keypoints
-%% Keypoints, final entry on title page.
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-
-% Example:
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-% \end{keypoints}
-%
-\begin{keypoints}
-\item Flood maps derived from Height Above Nearest Drainage (HAND) are subject to nearest drainage line limitations that affect inundation skill.
-\item A means of resolving this limitation is provided by reducing HAND processing units to level paths with effective unit stream order.
-\item Discretizing the stream network for HAND computation affects the stage-discharge relationship and leads to higher skill inundation.
-\end{keypoints}
-%
-
-%% Abstract
-% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT and PLAIN LANGUAGE SUMMARY
-%
-% A good Abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
-
-% The Plain Language Summary should be written for a broad audience,
-% including journalists and the science-interested public, that will not have 
-% a background in your field.
-%
-% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
-% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
-% see http://sharingscience.agu.org/creating-plain-language-summary/)
-%
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
-
-\begin{abstract}
-Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is a means able of producing flood inundation maps (FIMs) from the National Water Model (NWM) at large scales and high resolutions using reach-averaged synthetic rating curves.
-We highlight here that HAND is limited to producing inundation only when sourced from its nearest drainage line, thus lacks the ability to source inundation from multiple fluvial sources.
-A version of HAND, known as Generalized Mainstems (GMS), is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
-The FIMs associated with each independent LP are then mosaiced together, effectively turning the stream network into discrete groups of homogeneous unit stream order by removing the influence of neighboring tributaries.
-Improvement in mapping skill is observed by significantly reducing false negatives at river junctions when the inundation extents are compared to FIMs from that of benchmarks.
-A more marginal reduction in the false alarm rate is also observed due to a shift introduced in the stage-discharge relationship by increasing the size of the catchments.
-We observe that the improvement of this method applied at 4-5\% of the entire stream network to 100\% of the network is about the same magnitude improvement as going from no drainage order reduction to 4-5\% of the network.
-This novel contribution is framed in a new open-source implementation that utilizes the latest combination of hydro-conditioning techniques to enforce drainage and counter limitations in the input data.
-\end{abstract}
-%
-\section*{Plain Language Summary}
-Flooding is one of the most impactful natural disasters on life and property.
-The United States National Water Model (NWM) provides flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
-In order to convert flow rates from the NWM to flood inundation maps (FIM), a model, known as Height Above Nearest Drainage (HAND), is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model suffers from issues in mapping performance because inundation sourced from rivers is only considered from the nearest river line.
-We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-This is done by splitting the stream network into continuous river segments known as level paths (LPs).
-These LPs have no tributaries, thus are known to be stream lines with a unit stream order indicating no branching.
-HAND is computed independently for each LP and the resulting FIMs are mosaiced together to form one seamless map.
-We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance.
-%
-
-%%% Suggested section heads:
-% \section{Introduction}
-%
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
-
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-%
-% \section{Data} (Or section title might be a descriptive heading about data)
-%
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-%
-% \section{Conclusions}
-
-% Introduction
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Flooding is one of the most significant natural disasters in the United States (US) affecting both the loss of life and property. 
-In 2017 and 2019, river and flash flooding combined represented the leading cause of death and the second leading cause in 2018 among all natural disasters in the US \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}.
-More than an average of 104 deaths per year are attributed to flood events from the 10 year period ending in 2019 \cite{us_department_of_commerce_2020}. 
-With respect to property damages, river and flash flooding have contributed to 60.7, 1.6, and 3.7 billion non-inflation adjusted US dollars in the annual periods of 2017 to 2019, respectively \cite{national_weather_service_2020,national_weather_service_2019,national_weather_service_2018}, with the large spike in 2017 attributed to the Hurricane Harvey event along the Gulf Coast.
-Trends related to flood damages and fatalities have been steadily increasing over recent decades \cite{mallakpour2015changing,downton2005reanalysis,kunkel1999temporal,pielke2000precipitation,corringham2019effect}.
-Some are expecting that the hydrologic cycle will intensify due to climate change which will lead to more extreme precipitation in some areas along with a greater risk of flooding \cite{tabari2020climate,milly2002increasing,wing2018estimates}.
-Increasing trends in frequency and risk are not uniform across spatial regions with work by \citeA{slater2016recent} indicating that trends are increasing across the US Midwest and Great Lakes regions while decreasing in the coastal Southeast, Southwest, and California.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Operational Forecasting}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Operational flood forecasting systems are primary tools in developing accurate forecasts for public awareness prior to life threatening and property damaging events. 
-One of these operational systems is the Advanced Hydrologic Prediction System (AHPS) maintained by the National Oceanic Atmospheric Administration (NOAA) National Weather Service (NWS) with thousands of forecasting points across the US at typically short forecast horizons of 24 or 72 hours \cite{mcenery2005noaa}.
-AHPS provides forecasting services in the form of ensemble streamflows at more than 3,600 locations and flood inundation maps (FIM) at more than 150 of those points shown in Figure \ref{fig:forecast_points}.
-Additionally, two forecasting networks, Full Resolution (FR) and Mainstems (MS) stream networks, relevant to the National Water Model (NWM) (see Section \ref{ssec:national_water_model}) are rendered in Figure \ref{fig:forecast_points}.
-The FR network refers to the entire NWM forecasting domain while MS refers to the subset of the FR network that is at or downstream of AHPS forecasting points (see Section \ref{ssec:national_water_model}).
-On an approximate basis, there is only one forecast point every 1,450 km of river (FR) and one forecast point with FIM every 29,000 km of river (FR).
-Despite the AHPS advances in operational flood forecasting, it lacks sufficient domain coverage, spatial resolution, and long-range forecast horizons to address the increasingly complex water challenges facing the US.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/forecast_points.jpg}
-\caption{Forecast points with and without Flood Inundation Maps (FIM) in United States' Advanced Hydrologic Prediction System (AHPS).
-Note that only a small fraction of the AHPS forecast points have existing FIM.
-Also shown are the National Water Model (NWM) stream networks at the Full Resolution (FR) and Mainstems (MS) resolution.
-The FR network constitutes the entire NWM stream network while the MS resolution network is the FR network at or downstream of the AHPS forecast points shown.
-}
-\label{fig:forecast_points}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{National Water Model}
-\label{ssec:national_water_model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Additional work is required to address the gaps that the AHPS leaves in terms of spatial resolution, spatial coverage, and temporal forecast horizons.
-In response to growing stakeholder demand for enhanced and integrated water resource forecasts, the Office of Water Prediction (OWP) at the National Water Center (NWC) along with its partners at the National Center for Atmospheric Research (NCAR) have developed and implemented operationally the NWM which is a configuration of the Weather Research and Forecasting Hydrologic Model (WRF-Hydro) \cite{salas2018towards,gochis2021wrf,cosgrove2019evolution}.
-The NWM forecasts river discharges at more than 2.7 million forecast points at a variety of time horizons including lookback-range (3-28 hrs), short-range (18 hr), medium-range (10 day) and long-range (30 day) forecast horizons.
-The NWM enhances the spatial and temporal domain of the current AHPS capabilities operated at the 13 River Forecast Centers (RFC) in areas known as `hydro-blind'.
-As a complement to the operational NWM, RFC forecasts from AHPS forecast points are assimilated in the NWM and routed downstream to the next downstream AHPS forecast point where the process iterates again.
-This assimilation into the NWM is used to enhance forecasting skill by leveraging best available regional-scale forecasts.
-The river network upon which this special assimilation technique operates on is herein referred to as the Mainstem (MS) stream network.
-Figure \ref{fig:forecast_points} shows the NWM V2.1 FR stream network as well as the NWM V2.1 MS network.
-The MS network contains roughly 120 thousand forecasting points or roughly 4.4\% of the reaches of the FR stream network.
-
-The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the ``hydrofabric'' in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus,nhdplus2022vectors}.
-The term ``hydrofabric'' is used within the NWM jargon to describe the subset of hydrography composed of the geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data.
-The NWM provides stream forecasts at these hydrofabric segments using the Muskingum-Cunge method to reduce computational requirements of a continental scale model but fails to consider backwater dynamics \cite{bedient2008hydrology,ponce1994variable,gochis2021wrf}.
-The need for high resolution FIM at 10 m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale.
-The use of a 2-dimensional (2D) hydrodynamic model across a continental-scale and high spatial resolutions is very cost prohibitive especially in an operational setting.
-The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert 1-dimensional (1D) riverine discharges to stages, and finally to inundation extents and depths.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Height Above Nearest Drainage}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-HAND normalizes topography along the nearest drainage path and it has been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
-\citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
-The terrain index also provides additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel,aristizabal2021mapping}.
-\citeA{zheng2018river} developed a methodology for determining stage-discharge relationships known as SRCs by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
-This collection of methods, coupling HAND with SRCs, have been experimented with and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{HAND's Assumptions and Limitations}
-\label{sssec:hand_assumptions_limitations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-HAND operates on many underlying assumptions since it can only be used as an inundation proxy or no physics model and thus, not a true hydrodynamic inundation model \cite{nobre2016hand,liu2016cybergis,liu2020height}.
-HAND, to our knowledge, has only been applied to natural, inland, and riverine inundation applications thus it is also missing pluvial, coastal, ground water, and dam break components among other possible sources of flooding.
-Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest stream line which is used for catchment allocation and relative elevation calculation \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,maidment2017conceptual,garousi2019terrain,zheng2018river,zheng2018geoflood,johnson2019integrated,renno2008hand}.
-Stream thalweg networks must also collectively drain to a singular outlet point for a given processing region \cite{nobre2016hand,zheng2018geoflood,renno2008hand}.
-Since elevations don't naturally do this, they must undergo a long series of hydro-conditioning processes to enforce monotonically decreasing elevations across an entire processing unit along with hydrologically correct directions of flow \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,donchyts2016global,renno2008hand}.
-The level of digital elevation map (DEM) manipulation required to enforce this assumption can be substantial depending on the region and can be a significant source of error.
-The drainage enforcing assumption also interacts with an inability to properly account for fluvial inundation in regions of DEM depressions that lack natural drainage to riverine areas \cite{nobre2016hand,renno2008hand}.
-
-When used for FIM applications, HAND assumes only fluvial inundation sourced from its nearest drainage line is accounted for \cite{nobre2016hand,mcgehee2016modified}.
-Catchments are independent of one another for FIM purposes meaning a reaches' stage value is only used to threshold the HAND values within its respective catchment \cite{liu2016cybergis,zheng2018river,zheng2018geoflood}.
-This assumption plays to the ``Nearest Drainage'' term in HAND and creates a significant limitation within HAND for FIM applications \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-At the junction of high stream order and high flow rivers with lower flow tributaries, there can be a lack of inundation extents exhibited which is known colloquially in the forecasting community as the ``catchment boundary problem''.
-The academic community has somewhat referenced this issue before but it has been characterized more as a problem with the stream delineation process that comes from thresholding the drainage accumulation maps \cite{nobre2016hand,li2020evaluation}.
-Later in this study, we will re-introduce this problem and demonstrate how we initialize with a stream network (that of the NWM's) and thus avoid having to threshold accumulations to some arbitrary value to define stream networks.
-We illustrate how computing HAND independently for stream lines of unit stream order can significantly enhance FIM performance by accounting for multiple sources of fluvial inundation that may exist in certain regions and flow scenarios.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{HAND Implementations}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Due to significant advances in high performance computing (HPC) and large scale high resolution DEMs such as the 3D Elevation Program (3DEP) seamless at the 1/3 arc-second (approximately 10 m depending on latitude) scale, HAND has been implemented into software for large-scale, continental computation.
-As part of the OWP's Innovators Program and NWC's Summer Institute, the National Flood Interoperability Experiment (NFIE) generated FIM hydrofabric (will be used interchangeably with the datasets produced by HAND) rapidly on a HPC \cite{maidment2017conceptual,liu2016cybergis}.
-NFIE used open-source dependencies including the Terrain Analysis Using Digital Elevation Models (TauDEM) \cite{tarboton2005terrain} and the Geospatial Data Abstraction Library (GDAL) \cite{warmerdam2008geospatial} to compute HAND for the Continental United States (CONUS) at 331 Hydrologic Unit Code (HUC) 6 processing units in 1.34 central processing unit (CPU) years.
-By allocating 31 nodes at 20 cores per node for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2 terrabyte (TB) of peak memory and 5 TB of total disk space.
-Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (NHDPlusHR) which better agrees with the 10 m DEM from the NHDPlusHR program \cite{liu2020height}.
-The original NFIE dataset was employed by the NWC as an unofficial demonstration to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
-Further work by \citeA{djokic2019arc}, implemented a series of improvements to HAND including equidistant reaches, updates to use with NHDPlusHR hydrography, and AGREE DEM reconditioning \cite{hellweger1997agree} into an ESRI Arc-Hydro workflow with use in ArcGIS.
-More notably the software added the ability to derive HAND on both the NWM FR and MS stream networks to consider multiple sources of fluvial inundation along high impact rivers of primary forecasting concern.
-
-Related to these efforts, the United States Geological Survey (USGS) has invested in relative elevation HAND-like methods via work in the GIS Flood Tool (GFT) that also uses SRCs with cross-sections for stage-discharge relationships \cite{verdin2016software}.
-Additional investment by \citeA{petrochenkov2020pygft} was able to successfully scale this approach by transitioning the method to open-source Python source code (PyGFT) and implementing novel interpolation methods to help address some of the catchment boundary discontinuities discussed more in this paper.
-In addition to the domestic work done in the US, some studies have expanded upon HAND to cover global domains at 30 m resolutions \cite{yamazaki2019merit,donchyts2016global}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Office of Water Prediction Flood Inundation Mapping}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-In order to mitigate the ever increasing threat of flooding to life and property, an operational capability is required to extend NWM streamflow forecasts to river stages, inundation extents, and inundation depths.
-OWP FIM is introduced here as a continental scale capability that generates these products at high spatial and temporal resolutions.
-Here we introduce OWP FIM that utilizes a few of the latest techniques in HAND based FIM oriented for use with the NWM in continental scale operational forecasting settings. 
-Within the operational framework of OWP FIM, we introduce research demonstrating how FIM performance skill with HAND can be improved by discretizing stream networks into units of an effective unit Horton-Strahler stream order \cite{horton1945erosional,strahler1952hypsometric,strahler1952hypsometric} for HAND computation contexts.
-Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to the stream accumulation threshold which is closely related to stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-Here we demonstrate how reducing a HAND processing unit's stream network into discrete level paths (LPs) of singular, effective stream order, can enhance FIM skill by accounting for multiple possible sources of fluvial inundation.
-This capability is introduced progressively as MS (whose network represents about 4\% of FR network) and to a higher degree Generalized Mainstems (GMS) (covers entire FR network) which will be explained later on.
-The following methods and results describe the work in more detail and demonstrate its efficacy in producing enhanced FIM for the NWM.
-%
- %
-% Materials and Methods
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-\label{sec:materials_and_methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-OWP FIM is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
-Figure \ref{fig:methods_overview} gives a high level overview of the methodology used in OWP FIM and in this study.
-Input data from multiple sources are preprocessed (not illustrated) and then subset to processing areas based on the model used, FR, MS, or GMS.
-The standard processing unit of OWP FIM is a HUC8 and the entire NWM FR stream network is used for enforcment.
-Later, we explain how only the NWM MS stream network is used for the MS version of HAND, while for GMS, the FR stream network is discretized into LPs before computing HAND.
-A series of hydro-conditioning steps enforces the location of stream lines, monotonically decreasing elevations, excavated bathymetry, stream thalweg breaching, and levee enforcement.
-After a DEM suitable for HAND's assumptions is conditioned, the FIM hydrofabric is generated including stream network, catchments, HAND, SRCs, and cross-walk table.
-The FIM hydrofabric is defined as the datasets required to make an inundation map from discharges including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains SRC and cross-walk information).
-In operational circumstances, the NWM streamflows are used in conjunction with the FIM hydrofabric to derive forecast FIMs.
-However for evaluation purposes, we use the streamflows from the cross-sections of our benchmark model.
-As later discussed in Section \ref{ssec:inundation_mapping}, independent FIMs from multiple fluvial sources are mosaiced together.
-For the case of MS, two sources are mosaiced together (FR and MS) while for GMS the inundation from every LP is composited together.
-The evaluation FIM extents are compared to the extents of the benchmark model and metrics are computed.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=6.0]{figures/methods_overview.jpg}
-\caption{
-Methodology overview detailing high level steps followed in the study.
-The flow chart begins with the input data organized by source.
-Subsetting the data into processing units depends on which model is being considered.
-FR utilizes the entire NWM stream network processed at HUC8 processing areas.
-MS only computes HAND using the NWM stream at or downstream of legacy forecasting points.
-The resulting inundation from the MS HAND is eventually layered with the FIM from FR HAND to account for high levels of inundation contributed by the mainstem.
-Generalized Mainstem (GMS) discretized NWM streams into level paths (LP) then computes HAND and the FIMs independently only to mosaic them later. 
-This better accounts for multiple possible sources of fluvial inundation.
-The dotted lines denote the use of NWM streamflow forecasts to produce operational FIM but not used in this study.
-All acronyms used in the figure are defined in the paper.
-}
-\label{fig:methods_overview}
-\end{figure}
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Software Dependencies and Architecture}
-\label{ssec:software}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
-Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}.
-To simplify setup and enhance portability across host operating systems, OWP FIM packages all dependencies up in a Docker image \cite{merkel2014docker}. 
-A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub where a user could consult the Readme.md page for more information on how to acquire the datasets and reproduce the pipeline \cite{inundationMapping2022}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Datasets}
-\label{ssec:datasets}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data processed for OWP FIM are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by the Earth Science Information Partners (ESIP) \cite{esipData2022}.
-The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}.
-We utilized a series of data products from the NHDPlusHR including the BurnLineEvents \cite{nhdplus2022vectors}, Value Added Attributes (VAA) \cite{nhdplus2022vectors}, Water Boundaries (WBD) or HUC Layers \cite{nhdplus2022wbd}, and the DEM elevation rasters \cite{nhdplus2022dems}.
-These BurnLines used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk against for forecasting purposes \cite{nwm2022hydrofabric}.
-For enforcing levee data, the USACE NLD is used to burn feature elevations into DEMs \cite{engineers2016national}.
-Since NHDPlusHR datasets extend beyond land borders into sea and Great Lake regions, we used the land-sea border from OpenStreetMap (OSM) \cite{osm2021landsea} and the land-lake border from Great Lakes Hydrography Dataset (GLHD) \cite{GreatLakesHydrographyDataset} to exclude those areas from production of FIMs.
-Additionally, the Base Level Engineering (BLE) datasets within FEMA Region 6 spanning parts of nine states including Colorado, New Mexico, Texas, Oklahoma, Kansas, Arkansas, Louisiana, Missouri and Mississippi at two recurrence intervals, 1\% (100 year or yr) and 0.2\% (500 year or yr), are used for validation in this study and furnished by the Interagency Flood Risk Management (InFRM) consortium \cite{fema2021base,fema2021estimated}.
-These BLE datasets are provided at the watershed scale (HUC8) utilizing best available DEMs and simulations from the Hydrologic Engineering Center's River Analysis System (HEC-RAS) model \cite{us2022hydrologic}.
-The full input datasets presented by source are listed in Table \ref{tab:data}.
-%
-\begin{sidewaystable}
-\caption{Data sources, names, descriptions, and citations.}
-\label{tab:data}
-\centering
-\begin{tabular}{|p{1.75cm}|p{4cm}|p{11cm}|p{5cm}|}
-%\begin{tabular}{l c c}
-\hline
-Source & Name & Description & Citations\\
-\hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlusHR for hydro-enforcement. & \cite{nhdplus2022vectors} \\
-\hline
-USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display. & \cite{nhdplus2022vectors} \\
-\hline
-USGS & NHDPlusHR DEM & DEM used for NHDPlusHR at 1/3 arc-second (10 m) spatial resolution and vertical units in centimeters. & \cite{nhdplus2022dems} \\
-\hline
-USGS & NHDPlusHR WBD & Water Boundaries (WBD) or HUCs used for spatial processing units. & \cite{nhdplus2022wbd} \\
-\hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting adapted from NHDPlus V2 NHDFlowline\_Network feature class. & \cite{nwm2022hydrofabric} \\
-\hline
-NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM adapted from NHDPlus V2 Catchment feature class. & \cite{nwm2022hydrofabric} \\
-\hline
-NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as reservoirs or lakes adapted from NHDPlus V2 NHDWaterbody feature class. & \cite{nwm2022hydrofabric} \\
-\hline
-USACE & NLD & Levee database of locations and elevations. & \cite{engineers2016national} \\
-\hline
-OSM & Land-Sea Border & Border of land and sea. & \cite{osm2021landsea} \\
-\hline
-GLHD & Land-Great Lakes Border & Border of land and Great Lakes. & \cite{GreatLakesHydrographyDataset} \\
-\hline
-InFRM & Cross-Sections & HEC-RAS 1D cross-sections used for modeling in BLE datasets. Includes discharges for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
-\hline
-InFRM & Flood Inundation Extents & Inundation depths produced by InFRM BLE HEC-RAS 1D for 1\% and 0.2\% recurrence interval events. & \cite{fema2021estimated} \\
-\hline
-\end{tabular}
-%\end{table}
-\end{sidewaystable}
-%
-Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain which includes 2,188 HUC8s for the FR and GMS networks and 1,604 HUC8s for the MS method. 
-These methods will be explained in more detail later.
-An enhancement of OWP FIM over previous HAND based FIM versions is the support for Hawaii and Puerto Rico which are expansion domains in the NWM V2.0 and V2.1, respectively.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hydro-conditioning}
-\label{ssec:hydro_conditioning}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The DEM is subject to a series of hydro-conditioning procedures to enhance its suitability for riverine flood inundation mapping with HAND. 
-These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
-HAND inherently requires all areas eligible for inundation to drain to the designated drainage network.
-So to satisfy this requirment, DEMs must undergo significant manipulation.
-In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations to enable eligiblity for flooding.
-Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
-It can also be used to simulate more accurate bathymetry which is not accounted for in the 10 m DEM \cite{gesch2002national}.
-
-Specifically within the context of OWP FIM, the hydro-conditioning operations that take place in sequential order are presented. 
-Prior to any hydro-conditioning, all input datasets must be subset from their original spatial domain scales into the processing units of size HUC8. 
-The subsetting is done by spatial query for the cases of the levees, DEM, and NWM hydrofabric while the NHDPlusHR BurnLineEvents are subset via attribute query for the given reach code's membership in the processing unit.
-Hydro-conditioning raster operations take place on buffered boundary definitions to avoid edge contamination and effects \cite{lindsay2013measuring}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Network Enforcement} 
-\label{ssec:stream_network_enforcment}
-%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% APPENDIX
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Stream Network Enforcement}
+\label{sec:app_stream_network_enforcement}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The location of the stream network is enforced to ensure general agreement with the NWM network which is used for forecasting the streamflow inputs.
@@
 
-However, to better match the drainage density of the NWM FR V2.1 stream network, which is based on the NHDPlus V2, the BurnlineEvents are pruned utilizing a nearest neighbor search around the NWM flowlines.
+However, to better match the drainage density of the NWM FR V2.1 stream network, which is based on the NHDPlus V2, the BurnlineEvents are reduced in density using a nearest neighbor search around the NWM flowlines.
 Headwater points are first derived for the NWM FR V2.1.
@@
 The points in this nearest neighbor set are then traversed downstream.
-Any headwater portion in the NHDPlusHR or any other stream not traversed are pruned away to better match the resolution and spatial locations of the NWM stream network and its headwater points.
-The resulting pruned NHDPlusHR stream network gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHDPlusHR network corresponds to the NHDPlusHR network at or downstream of NWM V2.1 headwater locations only.
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
+Any headwater portion in the NHDPlusHR or any other stream not traversed are removed to better match the resolution and spatial locations of the NWM stream network and its headwater points.
+The resulting NHDPlusHR stream network of lower drainage density gets hydro-enforced in subsequent operations.
+This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the reduced NHDPlusHR network corresponds to the NHDPlusHR network at or downstream of NWM V2.1 headwater locations only.
+Additionally, the remaining NHDPlusHR headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
 This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents. 
@@
 The stream network used for forecasting, NWM FR V2.1, is of lower stream density than that of the NHDPlusHR which has better agreement with the thalweg locations in the DEM used.
-Thus, we opt to prune the NHDPlusHR network to match the general location and density of the NWM network.
+Thus, we opt to trim the NHDPlusHR network to match the general location and density of the NWM network.
 The nearest neighbor segment in the NHDPlusHR of each NWM headwater locations and the nearest point on that segment is determined to match the closest point to that of the NWM headwater.
-These points are then traversed downstream and any segments not traversed are pruned away.
+These points are then traversed downstream and any segments not traversed are removed.
 The resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR BurnlineEvents.}
@@
 
-The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
+The reduced density stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
 The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
@@
 Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
-Some of these drawbacks are addressed by additional conditioning techniques applied later on.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Levee Enforcement}
-%
-Coarse DEM's at 10 m, 30 m, and higher resolutions can lack sufficient representation of fine grain features such as embankments, flood walls, and closure structures \cite{arundel2018assimilation,dobbs2010evaluation,wang2005comparison,sanders2007evaluation}.
-In order to better represent the influences of these features upon hydraulics and inundation extents, the National Levee Database (NLD) published by USACE was used to enforce elevations within the 1/3 arc-second DEM.
-The elevations found in the NLD are burned onto the DEM if those elevations were found to exceed those already in place.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Depression Filling}
-\label{sssec:depression_filling}
-%
-Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes with HAND.
-The partially conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
-Priority-Flood developed by \citeA{barnes2014priority} is an algorithm for filling said depressions and shown to have improved performance over early works in the field by \citeA{jenson1988extracting} implemented in \citeA{tarboton2005terrain} as well as \citeA{planchon2002fast}.
-The depression filling algorithm used in our pipeline is a Priority-Flood variant developed by \cite{zhou2016efficient} with enhanced single-thread performance and a time complexity of O(n log n) for floating point grids.
-This performance was enabled by limiting the processing queue with a region-growing method to exclude many of the slope cells \cite{zhou2016efficient}.
-The depression filling technique employed here does leave the existence of flat regions where pits previously existed thus later requiring the need for resolving these flats.
-The enhanced variant of Priority-Flood is implemented and made available by \citeA{barnes2018richdem} and \citeA{zhou2015filldem}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Stream Thalweg Elevation Conditioning}
-\label{sssec:stream_thalweg_elevation_conditioning}
-%
-Thalweg elevations are critical components of relative elevation based inundation mapping thus much is performed to ensure the best available, monotonically decreasing, elevations are derived prior to the normalizing of elevations.
-Work on the AGREE DEM method from several authors have illustrated that the AGREE DEM method does not prevent situations where the burned thalweg and the thalweg endemic to the DEM run parallel to one another \cite{hellweger1997agree,baker2006comparison,saunders1999preparation,saunders1996gis,quenzer1998gis,saunders1995grid}.
-These works observe that the artificial elevations enforced by the hydrographically based stream network and AGREE DEM disagree with those naturally occurring in the native DEM.
-In order to mitigate this documented issue, the normalized excavation algorithm \cite{saunders1999preparation} is used to seek a zonal (nearest neighbor) elevation minimum on the original, unconditioned DEM for each thalweg pixel.
-Each zone is defined as the thalweg's pixel nearest neighborhood within a maximum distance of 50 m.
-The zonal minimum is computed for each thalweg pixel zone and the minimum is used to replace the existing thalweg elevation value.
-This step essentially enforces an estimate of the native DEM thalweg elevations onto the sharp drop enforced thalweg elevations from the AGREE procedure.
-
-The next step involves conditioning these local minimums along the thalweg to enforce monotonically decreasing thalweg elevations for FIM.
-\citeA{garousi2019terrain} proposed an algorithm that breaches stream thalweg pixel elevations in a depth first manner. 
-This procedure was found to increase the Critical Success Index (CSI) of resulting FIMs from HAND and is employed in OWP FIM to enforce monotonically decreasing elevations with thalweg pixel networks.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Deriving FIM Hydrofabric}
-\label{ssec:deriving_fim_hydrofabric}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
-These datasets include the HAND or relative elevation model (REM) raster, reach-level catchments raster/polygons, DEM-derived streamlines, SRCs, and cross-walk table.
-The following sub-sections describe how the subset and hydro-enforced geospatial datasets are converted into the FIM hydrofabric.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Flow Directions and Flats Resolution}
-\label{ssec:flow_direction_and_flat_resolution}
+Some of these drawbacks are addressed by additional conditioning techniques applied in the methodology.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Flow Directions and Flats Resolution}
+\label{sec:app_flow_direction_and_flat_resolution}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Deriving FIM Stream Network}
-\label{sssec:deriving_fim_stream_network}
-%
-The derivations of relative elevations and catchments from the newly conditioned DEM involves re-deriving a new, DEM based, FIM stream network. 
-The FIM stream network is of similar drainage density as the NWM V2.1 network and fully converges at all junctions leaving no divergences in the network.
-This is accomplished by using the seed points generated from the stream network enforcement process (Section \ref{ssec:stream_network_enforcment}).
-These seeds points are pruned headwater locations of the NHDPlusHR Beta BurnlineEvents layer that spatially correspond to the headwater definitions in the stream network of the NWM V2.1.
-Feeding the seed points and previously computed flow directions into flow accumulation methods \cite{wallis2009parallel,tarboton1997new,tarboton2005terrain} yields a stream link accumulation raster that can be converted to a vector file for further processing.
-
-Each stream link in this derived FIM stream network is split into equidistant reaches of 1.5 km in length which is a user exposed parameter.
-Stream links are defined here as segments of rivers discretized by junctions with other NWM river segments.
-Stream links are then further segmented at NWM lakes and HUC8 boundaries.
-Discretizing at NWM lakes isolates reaches and catchments associated with lakes and reservoirs to avoid mapping them using the Manning's equation and could potentially enable volume based mapping in the future as a feature enhancement.
-Based on previous research, splitting each remaining stream link into equidistant reaches not to exceed a parameterized value of 1.5 km helps improve SRC and mapping skill \cite{garousi2019terrain,godbout2019error,zheng2018geoflood}.
-Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much slope variance into one discretization of the stream network.
-Short stream segments that are introduced as a result of forced network breaks due to reservoir, levee, or HUC boundaries inherit the SRC properties of the upstream or downstream segment, depending on the topology.
-Section \ref{sssec:synthetic_rating_curve} details the derivation of the SRC and the dependence on channel length. 
-Additionally every reach (and later catchment) is assigned a globally unique identifier based on the HUC8 membership.
-This stream network is important since it drives the HAND calculation and derivation of catchments.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Catchments}
-\label{sssec:catchments}
-%
-Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
-Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
-The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing, nearest drainage region for each outlet point.
-Two sets of catchments are derived, one set of catchments denotes the unique drainage region for each thalweg pixel which is used for relative elevation calculation.
-The other catchments are derived for the drainage region for each stream reach as defined in Section \ref{sssec:deriving_fim_stream_network}. 
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Height Above Nearest Drainage}
-%
-Once the pixel level catchments are derived, the final relative elevations can be computed.
-The elevation of every thalweg pixel is subtracted from the elevations of the non-thalweg pixels within the same, corresponding pixel-level catchment described in Section \ref{sssec:catchments}.
-The DEM used for this operation is the DEM resulting from the thalweg conditioning procedures described in Section \ref{sssec:stream_thalweg_elevation_conditioning}.
-Outside of the excavated channel from the AGREE DEM method, the native non-drainage enforced elevations are used to reduce sources of error in relative elevations due to pit filling \cite{djokic2019arc}.
-Any negative values resulting from this subtraction with native elevations are replaced by zero.
-Again, HAND assumes and requires processing areas to drain thus have monotonically decreasing elevations with hydrologically correct flow directions all leading to a singular outlet point.
-While this is required for the generation of DEM-derived catchments and stream lines, it is not necessarily required for the computation of the relative elevations.
-Since the use of hydro-conditioning processes to fit the drainage requirement for HAND can be extensive, we found it more fitting to use the native elevations this final HAND computation thus avoid the use of manipulated values that fit modeling assumptions.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Synthetic Rating Curves}
-\label{sssec:synthetic_rating_curve}
-%
-A method for converting forecast river discharges from the NWM to stages or river depths at the reach scale is necessary for producing FIMs with HAND. 
-For 1D hydrodynamic models such as the routing methods in the NWM, the typical procedure is to establish the stage-discharge relationship by sampling data from the DEM to derive a SRC at discrete cross-sections \cite{quintero2021development,di2011hydraulic}.
-For this application, we utilized the reach averaged approach for developing SRCs \cite{zheng2018river}.
-The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-Previously not reported in literature to our knowledge in this form, the reach averaged Manning's formula is derived to be 
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:reach_averaged_mannings_equation}
-Q(y) = \frac{1}{n} \frac{V(y)^{5/3}S^{1/2}}{L B(y)^{2/3}} 
-\end{equation}
-\end{linenomath*}
-%
-where Q is discharge at stage y, n is the Manning's n roughness coefficient, V is volume at y, S is channel slope, L is the along-flow reach length, and B is wetted bed area at y.
-Q, V, and B are taken at specific y values so are more formally written as $Q = Q(y)$, $V = V(y)$, and $B = B(y)$, respectively.
-All units are international (SI) given the one in the numerator above n.
-The reach averaged method has been compared to rating curves from HEC-RAS and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
-The reach averaged geometry parameters including number of wet cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
-Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to calculate the reach length.
-While the AGREE DEM is subject to hydro-conditioning processes, it does introduce some notion of bathymetry estimation that the native DEMs lack while being sensitive to additional parameters that could yield further errors in the FIM.
-We leave this issue open in this study and elaborate on needs with respect to bathymetry and Manning's n values in the Discussion section (Section \ref{sec:discussion}).
-
-Setting of the Manning's n roughness coefficient has precedent in previous continental-scale FIM (CFIM) studies \cite{maidment2017conceptual,liu2016cybergis,liu2020height,djokic2019arc,garousi2019terrain,zheng2018geoflood} with two noted values of 0.05 and 0.06 for NFIE and \citeA{djokic2019arc} respectively.
-These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better localizing Manning's n according to available data including but not limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
-For now and for the purpose of this study, we examine the SRCs with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
-After all the parameters to the Manning's equation have been determined with either hydrofabric sampling or user parameterization, we select 84 stage values (y in Eq.
-\ref{eq:reach_averaged_mannings_equation}) from 0 to 25 meters in depth at a third of a meter increments to calculate the discharge values for each stage value.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cross-walking with NWM Stream Network}
-\label{sssec:cross_walking_networks}
-%
-The DEM based stream network derived in Section \ref{sssec:deriving_fim_stream_network} must be associated with NWM reach identifiers so that a discharge can be converted to stage and later inundation extents and depths.
-For the FR version of HAND, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent HAND methods, MS and GMS, discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, respectively, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
-Additionally, only relevant catchments from the NWM for the given LP are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
-While these conflation methods are approximate, they can lead to some substantial errors which will be discussed more in Section \ref{sec:discussion}.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stream Order Reduction}
-\label{ssec:stream_order_reduction}
-%
-As previously discussed, HAND based FIMs are subject to many assumptions and limitations in order serve as a suitable inundation proxy for large scale, high resolution domains.
-HAND produced FIMs are limited in only providing inundation sourced from the nearest drainage line, however, depending on flow conditions and topography, a given area may have multiple contributing fluvial sources of surface water inundation.
-The forecasting community, in reviewing HAND, have noted significant negative effects at the confluence of lower flow tributaries with higher flow rivers for which the phrase ``catchment boundary issue'' has been termed.
-In previous studies, FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND which is closely related to the maximum Horton-Strahler stream order of the network \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is partly in due to the limitation that catchment boundaries place on inundation extents where only the nearest drainage line can source inundation for any particular area.
-
-Figure \ref{fig:catchment_boundaries_issue} illustrates the exact situation our solution proposes to address where two tributaries converge with a higher order stream segment. 
-An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 yr extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the main segment in Figure \ref{fig:catchment_boundaries_issue} of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS.
-This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
-As a parallel problem, there is excess water accumulated along the mainstem that cannot extend in either a fluvial or pluvial manner beyond the boundaries of the mainstem catchments.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/catchment_boundaries_issue.jpg}
-\caption{The figure represents an agreement map between a HAND derived FIM and one produced from the Base Level Engineering (BLE) program for a 100 yr magnitude event at HUC8 12090301.
-Agreement maps are symbolized by false negatives (FN), true negatives (TN), false positives (FP), and true positives (TP) where inundated represents the positive condition (see Section \ref{ssec:evaluation} for more details).
-The streamflows associated with each river segment are shown in CMS while the flow directions are symbolized as red arrows.
-The presence of FNs at the confluence of tributaries (circled in red) with the main segment is associated with lower flow rates in the tributaries that don't account for backwater effects.
-Additionally, the flow of 1900 CMS from the main segment cannot extend to the neighboring catchments belonging to its tributaries shown here.
-Water pools up vertically along the catchment boundaries of the higher order segment distorting rating curve behavior (Section \ref{sec:discussion}).
-Sourcing fluvial inundation from HAND is limited to only its nearest drainage line which is the main issue this study aims to address.
-}
-\label{fig:catchment_boundaries_issue}
-\end{figure}
-%
-
-We seek to resolve this catchment boundary problem or nearest drainage limitation by discretizing the target stream network into stream networks of reduced, unit stream order to avoid the constraining of catchments by those belonging to lower order neighbors.
-By discretizing the network into stream networks of unit stream order (later defined as LP), we remove the influence of neighboring catchments that constrain the inundation extent.
-This creates much larger and overlapping catchments that can source fluvial inundation from multiple reaches as required by the given river stage at current flow conditions.
-We present two successive methods, National Weather Service MS (Section \ref{sssec:nws_mainstems}) and GMS (Section \ref{sssec:generalized_mainstems}), implemented that reduce the effective Horton-Strahler stream orders of the networks employed and test our presented hypothesis that unary stream order networks enhance FIM performance skill with HAND by expanding the nearest drainage definition to increase potential inundation areas.
-
-To clarify the phrase ``reducing Horton-Strahler stream order'' used extensively in this paper, every FIM used in evaluation contains a flood extent sourced from every NWM forecast point in the given evaluation domain.
-What we do to reduce stream order is discretize the NWM FR network into different units of size, MS network (\ref{sssec:nws_mainstems}) and GMS LPs (\ref{sssec:generalized_mainstems}), that effectively reduce the HAND computation to independent networks of unit stream order.
-These independent HAND datasets are later used to produce FIM independently and mosaiced together (see Section \ref{ssec:inundation_mapping}).
-The inundation from the MS HAND is mosaiced with the inundation from FR HAND, while the inundation of each individual LP from GMS is mosaiced together.
-The Horton-Strahler stream order is only reduced for HAND computation purposes to reduce the negative effects of the nearest drainage limitation inherent to HAND.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{NWS Mainstems}
-\label{sssec:nws_mainstems}
-%
-The initial attempt at drainage order reduction to solve the catchment boundary issue was to use a stream network relevant to the NWS forecasting community. 
-The Mainstems (MS) network is a subset of the NWM FR network at and downstream of AHPS forecast points as seen in Figure \ref{fig:forecast_points}.
-The MS network comprises about 200 thousand km of stream length which is less than 4\% of the FR total stream length of 5.5 million km.
-It also spans 121,724 reaches across 1,608 HUC8s.
-In this technique, we derive HAND using the FR stream network as well as the MS network which was originally proposed by \citeA{djokic2019arc}.
-Inundation is derived independently from the resulting FR and MS HAND hydrofabrics and are mosaiced together using the technique proposed in Section \ref{ssec:inundation_mapping} to form the MS FIMs. 
-Within each HUC, one might typically only find a MS stream network of uniform stream order but this can vary if more than one AHPS forecasting point is found within or upstream of the HUC in question.
-So while we may refer to the MS network as that of one with unit stream order, we acknowledge there are many cases where additional or converging forecast points create multiple branches within a given processing unit.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Generalized Mainstems}
-\label{sssec:generalized_mainstems}
-%
-Since MS only covers 4\% of the entire FR stream network, we sought to expand drainage order reduction techniques to all reaches within the NWM modeling domain.
-In order to do this, we discretized the NWM network into LPs which when considered individually have unit Horton-Strahler stream orders.
-LPs group flowlines by maximizing the length of each flow path and minimizing the number of LP identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-In order to derive LPs for the NWM FR stream network at the HUC8 processing area, we first compute arbolate sums which are defined as the cumulative drainage distance of all upstream drainage lines.
-Arbolate sum is also inclusive of the current drainage reach as well.
-Arbolate sums are computed by starting at the headwater points and summing up drainage distances as you traverse downstream.
-
-Arbolate sum is critical to discretizing the NWM network into LP identifiers.
-Starting at a HUC8's outlet, a unique LP is propagated upstream. 
-At every confluence, the direction of maximum arbolate sum is sought to propagate the current LP identifier.
-For the remaining parent reaches of the given junction, a new LP identifier is assigned and the process recursively continues with them.
-Figure \ref{fig:level_path_methods} illustrates how LPs (symbolized by unique colors) are propagated upstream by the value of arbolate sum.
-The figure shows computed arbolate sums and unique LP identifiers on a HUC12 (120903010404) for clarity but were computed at the corresponding HUC8.
-The mainstem of the figure runs from the red ellipses to the black one which is the outlet.
-From the figure, we can see how unique colors are propagated in the direction of the maximum arbolate sum.
-
-Each HUC8 is discretized into LPs independently and the relevant inputs as described in Table \ref{tab:data} are assigned to each LP processing unit given a buffer of seven km.
-This buffer was selected to avoid edge contamination \cite{lindsay2013measuring} and to ensure adequate data availability for wide rivers with large catchments in regions with low slope.
-Further work could be dedicated to tune this user exposed parameter to better balance its effect on FIM extents and computational expense since larger buffers create additional floating point calculations and storage requirements.
-For the time being, we designate this issue to be out of scope.
-
-At the LP scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the LP in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement by burning these lines and seeding from its headwater points directly instead of going through the NHDPlusHR network as described in Section \ref{ssec:stream_network_enforcment} .
-This decision was motivated by the difficulty in deriving LPs in the NWM stream network with high agreement with the LPs derived for the NHDPlusHR stream lines.
-We found that the same algorithm to compute arbolate sums and LPs could yield enough disagreements associated with disordered branches or slight differences in arbolate sums that could significantly affect the agreement of the LP identifiers in the NWM and NHDPlusHR networks.
-This yielded enough error to justify the use of the NWM directly for hydro-enforcement operations.
-
-Once the NWM FR stream network is discretized into LPs, we independently compute HAND using each LP as the target stream network to be used.
-To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
-In Figure \ref{fig:gms_methods}a, we uniquely color code the LPs derived for the NWM stream network. 
-For each one of these lines, we derive HAND and its associated datasets including catchments, crosswalks, and rating curves.
-Each LP is buffered to a polygon with a user-exposed, distance parameter of seven km that is used to subset the original DEM for two selected LPs in Figure \ref{fig:gms_methods}b.
-We illustrate two HAND grids for two of the LPs in this HUC8 in Figure \ref{fig:gms_methods}c.
-Once the FIM hydrofabrics for each LP are generated, we can inundate them individually also shown in Figure \ref{fig:gms_methods}d.
-Lastly, these individual FIMs are mosaiced together as explained in Section \ref{ssec:inundation_mapping} and shown in Figure \ref{fig:gms_methods}d.
-
-For a more intimate look at the drainage order reduction procedure GMS, and its effects, we allude to Figure \ref{fig:gms_methods_2} which references the same area (in HUC8 12090301) and set of river junctions as in Figure \ref{fig:catchment_boundaries_issue}.
-The catchments and stream lines for HAND computed at the FR scale are illustrated in Figure \ref{fig:gms_methods_2}a where the respective inundation at the 100 yr magnitude is heavily constrained by the limited catchment extents especially at junctions.
-In subsequent sub-figures, we show the same datasets for the HAND computation problem for this region but discretized into independent LPs for the main LP (b), the eastern tributary (c), and the western tributary (d).
-Notably, inspecting (b), one sees how removing the tributaries creates much larger catchments for the main LP. 
-These catchments include drainage areas that would traditionally be considered nearest to the tributaries thus ineligible to receive inundation sourced from the main LP.
-The inundation extents in (b) overlap those of (c) and (d) and are mosaiced together by methods explained in Section \ref{ssec:inundation_mapping}.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/level_path_methods.jpg}
-\caption{Illustrates the NWM Full Resolution V2.1 stream network discretized into level paths (LP), symbolized by unique colors, as well as the values of the arbolate sums to the nearest whole km distance.
-The LPs were derived on a HUC8 level (12090301) but only illustrated for a HUC12 (120903010404) for clarity.
-The mainstem of this HUC12 runs from the red ellipses to the outlet denoted by the black ellipses.
-Arbolate sums are defined as the cumulative drainage distances of all upstream stream lines.
-Arbolate sums are computed for the NWM network by starting at the headwater points then traversing downstream and adding the distances cumulatively. 
-LPs are derived by starting at an outlet point with a unique identifier (ID).
-The unique LP ID is propagated upstream until a junction is reached where the current LP ID is propagated in the direction of maximum arbolate sum.
-The remaining converging segments at the given junction are each assigned a new unique LP ID and the process is repeated recursively until all reaches have been assigned a LP.
-Thus, LP serve as a proxy means of assigning membership to a given river when presented with a confluence.
-Each individual LP has a unit Horton-Strahler stream order thus serves as a great method for our proposed technique.
-}
-\label{fig:level_path_methods}
-\end{figure}
-%
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/gms_methods.jpg}
-\caption{Overall procedure for GMS HAND at HUC8 12090301.
-In (a), we illustrate all NWM stream lines symbolized by their LP with 372 unique LP IDs in this HUC.
-Meanwhile (b), demonstrates the DEM clipped to a seven km buffer around two selected LPs.
-In (c), we show how HAND can be computed just for each one of these two LPs independently. 
-We also show inundation maps created for these two LPs in (c). 
-In (d), we show all the inundation maps for all the LPs mosaiced together. }
-\label{fig:gms_methods}
-\end{figure}
-%
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.70]{figures/gms_methods_2.jpg}
-\caption{This image, with the same spatial domain as Figure \ref{fig:catchment_boundaries_issue} (HUC 12090301), demonstrates how computing HAND on level path (LP) bases leads to larger, independent catchments and more expansive inundation extents (100 yr flows).
-In (a), the catchments and stream network are shown for HAND computed in Full-Resolution (FR) method which shows constrained inundation extents around the two junctions.
-(b) demonstrates the LP associated with this region's highest order river.
-By delineating catchments at this scale independent of the neighboring tributaries, the drainage areas are allowed to expand thus allowing inundation extents to cover previously restricted areas.
-In (c) and (d), we show the stream lines, catchments, and inundation extents of the two tributaries.
-Later in Section \ref{ssec:inundation_mapping}, we describe how the inundation in (b), (c), and (d) are mosaiced together to form one seamless inundation map.
-This process allows for multiple, possible contributing sources of fluvial inundation to be considered thus enhancing FIM skill.
-}
-\label{fig:gms_methods_2}
-\end{figure}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Inundation Mapping}
-\label{ssec:inundation_mapping}
-%
-The FIM hydrofabric consisting of the relative elevations grid, catchments grid, catchment polygons, rating curve, and cross-walking data are all used to convert forecasts from the NWM into forecasts extents.
-For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
-From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC8 spatial processing units in a rapid, parallel operation.
-The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
-
-Utilizing the stage-discharge relationships in the SRCs, each forecast for each catchment identifier is assigned a stage value. 
-The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
-We use the basic logic already established in previous works to conduct this \cite{nobre2016hand,liu2016cybergis,maidment2017conceptual}.
-Mathematically, the HAND values, $H_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
-For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
-For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim_depth}
-    D_{ij} = S_i - H_{ij}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:hand_fim}
-    F_{ij} = D_{ij} > 0
-\end{equation}
-\end{linenomath*}
-%
-For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast in the form of a single raster file.
-For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
-Consolidating the depths using a maximum function was decided upon based on intuition which we believe to best represent the depth of water in an area with multiple contributing fluvial inundation sources.
-Other aggregation methods could lead to different results but were not investigated here.
-Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i,
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim_depths}
-    D_{ij} = \max_{k=[1,...,K]} D_{ijk}
-\end{equation}
-\end{linenomath*}
-%
-. Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps,
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:comp_fim}
-    F_{ij} = \max_{k=[1,...,K]} F_{ijk}
-\end{equation}
-\end{linenomath*}
-. For the MS and GMS methods, the contributing areas are defined differently.
-For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to two for that case.
-For GMS, all FIMs from all the LPs in a given area are mosaiced together then K is set to this number of LPs.
-Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evaluation}
-\label{ssec:evaluation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Possible benchmark FIM candidates for evaluation purposes include high water marks, remote sensing observation, crowd-sourced information, and modeled extents.
-These sources are all subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to a lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
-While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to as to isolate out other hydrological factors.
-
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
-This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, higher level of sophistication when compared to HAND, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
-We selected 49 available HUC8s, shown in Figure \ref{fig:all_ble_maps}, which span about 185 thousand $km^2$ across nine states.
-The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM as well as the corresponding discharges and mapping extents for evaluation.
-We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation sourced from OWP FIM.
-
-By using the same HEC-RAS derived discharges and FIM extents for creating maps with OWP FIM, we are able to separate out errors introduced by NWM inputs and processes including land surface interactions, groundwater fluxes, atmospheric forcings, hydraulic routing, and others that would have potentially affected our conclusions if we had used NWM forecasted discharges.
-Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
-We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
-To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
-This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
-Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with unit Horton-Strahler order due to the lack of representation of these additional headwater streams in the NWM network.
-While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
-\caption{
-Shows 185 thousand $km^2$ of modeled areas for the Base Level Engineering (BLE) domain of 49 HUC8s across nine states at 0.2\% recurrence magnitude for flow rates.
-BLE maps are produced for two recurrence flows, 1\% (100 yr) and 0.2\% (500 yr), using 1D HEC-RAS models.
-The maps are used as benchmarks for validation purposes of OWP FIM.
-}
-\label{fig:all_ble_maps}
-\end{figure}
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
-\caption{
-Illustrates Base Level Engineering (BLE) cross sections and stream lines at the HUC8 12100203 near the confluences of West Fork Plum Creek and Clear Fork Plum Creek with Plum Creek.
-BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM V2.1 Full Resolution (FR) stream lines.
-Additionally, we illustrate the NWM V2.1 catchments to provide a sense of how many cross-sections may intersect a given NWM flowline.
-The BLE stream network is also shown which is denser than the NWM V2.1 stream lines meaning there are several lower order streams represented in the BLE stream network that are not in the NWM V2.1 stream lines.
-This creates additional inundation areas in the validation data that are not modeled with our HAND based FIMs.
-}
-\label{fig:ble_evaluation_method}
-\end{figure}
-%
-
-The metrics employed in this study to evaluate inundation extents include CSI, Probability of Detection (POD), and False Alarm Ratio (FAR) and are presented in Equations \ref{eq:csi}, \ref{eq:pod}, \ref{eq:far}, respectively.
-To calculate these secondary metrics, one must define three primary metrics starting with true positives (TP) which is predicted wet and wet in the BLE benchmark dataset.
-The two types of errors consist of false positives (FP), or type I errors, which is dry in the benchmark but predicted wet and false negatives (FN), or type II errors, which is wet in the benchmark but predicted dry.
-Lastly, the reader may come across true negatives (TN) which is defined as dry in both the benchmark and predicted datasets.
-Maximizing POD indicates a model's ability to detect the given threat of interest, inundation, while minimizing FAR is sought to indicate a models ability in reducing FN errors.
-In other words, POD is an indicator of model skill in inundated regions while FAR is an indicator of model skill in non-inundated regions.
-Some work by \citeA{gerapetritis2004behavior} denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
-We use all three secondary metrics here to add value to the discussion while avoiding aggregating away the meaning of all four primary metrics.
-
-While these metrics are commonly employed in the evaluation of FIM and binary weather prediction communities in general, they do come with some notable limitations including frequency dependence in the case of CSI and FAR \cite{gerapetritis2004behavior,stephens2014problems,schaefer1990critical,jolliffe2012forecast}.
-Thus, frequency dependent statistics should be used with caution when comparing across sites with varying frequencies. 
-Lastly, approximately six HUC8s do not have NWM MS reaches thus we imputed the metrics for FR for these sites as the best available forecasting capability to compare GMS metrics to.
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:csi}
-CSI = \frac{TP}{TP + FN + FP}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:pod}
-POD = \frac{TP}{TP + FN}
-\end{equation}
-\end{linenomath*}
-%
-\begin{linenomath*}
-\begin{equation}
-\label{eq:far}
-FAR = \frac{FP}{TP + FP}
-\end{equation}
-\end{linenomath*}
-%
- %
-% Results
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{sec:results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Flood Mapping Performance}
-\label{ssec:flood_mapping_performance}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-We produced FIMs for the entire BLE domain within the 49 HUC8 areas across several states in the south central US. 
-The forecasted FIMs using the discharges for the 1\% (100 year) and 0.2\% (500 year) recurrence flows directly from HEC-RAS were used to avoid noise and errors from hydrological processes.
-We computed the statistics, CSI, POD, and FAR, for both 100 and 500 year events for Mannings N set to 0.06 and 0.12.
-These results are presented in Figure \ref{fig:violin_plot} as violin plots and in Table \ref{tab:aggregate_metrics} as aggregated metrics with the results for MS and GMS presented as percentage changes from their respective FR values.
-To be more specific, Table \ref{tab:aggregate_metrics} sums the primary metrics, TP, FP, FN, and TN, across all HUC8s then recomputes the secondary metrics which was done to better account for large variances in HUC8 size.
-The same trends discussed below are consistent across both reporting methods (Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}).
-
-The distribution of these flood extent metrics can be examined in Figure \ref{fig:violin_plot} as violin plots.
-Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), Manning's n value (0.06, 0.12), recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
-For example, let's examine the violin plot for the row marked CSI and column for Manning's n = 0.06.
-This sub-figure shows the CSI distributions across all 49 HUC8s when Manning's n is set to 0.06.
-Each independent, whole violin represents the HUC8 metric value distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100 yr and orange for 500 yr).
-The horizontal dashed and dotted lines represent the 25th, 50th, and 75th percentiles from bottom to top, respectively.
-Additionally, we show trend lines symbolized in green that for each metric and Manning's n combination denotes the best fit line for the three methods (FR, MS, and GMS).
-To avoid having two trend lines per sub-figure, we elected to aggregate the two magnitudes together as they tend to observe similar trends across the three models.
-The slope of each trend line is quantified in the figure by its $\beta_1$ value and the p-value for that statistic which tests the significance of that trend (deviation from a zero sloped trend line).
-
-% Interpretation of metrics
-Both Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} contain a fair amount of information that is central to the objectives of this study.
-As previously stated in Section \ref{ssec:evaluation}, we consider CSI as a general proxy for the skill of the inundation extents with POD denoting skill on inundated areas and FAR indicating skill on non-inundated areas.
-Again, the main objective of the study is to introduce how computing HAND with disaggregated stream networks to those with unit stream order can enhance the fidelity of FIMs by capturing fluvial inundation from multiple sources as opposed to that of just the nearest drainage line.
-As can be seen in Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, CSI generally increases from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
-This increase is primarily driven by an increase in POD thus generally increasing the probability of correctly detecting inundation.
-Also, we note that FAR is somewhat, albeit marginally, decreased from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
-The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}).
-So overall and in other words, the broader distribution of HUC8s improves across the three methods.
-Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit stream order and mosaic the resulting FIMs, the better those FIMs perform.
-We move more details on the relationship between stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).
-
-Additional noteworthy trends in Figure \ref{fig:violin_plot} center around the all-around better performance of FIMs for those of higher Manning's n values and recurrence flows.
-The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value of 0.11 used in the BLE model \cite{fema2021base,fema2021estimated}.
-Most of this improvement is driven by significant increases in POD, but unfortunately, it also leads to a significant amount of over-prediction as observed by the increase in FAR.
-More work can be invested to better regionalize Manning's n values for FIM purposes with HAND.
-We also observe additional trends associated with the magnitude or recurrence interval of the flow rates used with the higher flow rates exhibiting better overall CSI, POD, and FAR values than the lower, 100 yr magnitude.
-We introduce in the Discussion (Section \ref{sec:discussion}) that this skill premium exhibited by higher flow events could be due higher quality elevation data in regions that are not described as bathymetric areas.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.8]{figures/violin_plots.jpg}
-\caption{Shows kernel density estimation of the distributions (sample size = 49) for 1\% (100 year) and 0.2\% (500 year) along with horizontal, dashed lines for the 25th, 50th, and 75th percentiles (in order from bottom to top).
-The sub-figures separate the combination of three metrics (CSI, POD, and FAR) for two settings of Manning's n (0.06 and 0.12).
-Trend lines for each combination of metric and Manning's n are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}
-\label{fig:violin_plot}
-\end{figure}
-%
-\begin{table}[H]
-\caption{ Recomputed CSI, POD, and FAR using the primary metrics, TPs, FPs, and FNs, aggregated to the BLE domain.
-         The values for MS and GMS methods are expressed in percentage change (\%) from their respective values with the same Manning's n, magnitude, and metric combination in the Full Resolution (FR) method columns.
-         The best value across models is highlighted in bold.}
-\label{tab:aggregate_metrics}
-\centering
-%\begin{tabular}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
-\begin{tabular}{|c|c||c|c|c|c|c|c|}
-\hline
-\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{\shortstack{MS\\(\% Change)}} & \multicolumn{2}{|c|}{\shortstack{GMS\\(\% Change)}} \\
-%\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{\shortstack{MS\\(\% Change)}} & \multicolumn{2}{|c|}{\shortstack{GMS\\(\% Change)}} \\
-\cline{3-8}
-  &  & 100 yr & 500 yr & 100 yr & 500 yr & 100 yr & 500 yr \\
-\hline
-\multirow{2}{*}{CSI} & 0.06 & 0.5576 & 0.5839 & 2.53 & 2.59 & \textbf{3.95} & \textbf{4.04} \\
-\cline{2-8}
-  & 0.12 & 0.5915 & 0.6149 & 2.35 & 2.26 & \textbf{4.51} & \textbf{4.65} \\
-\hline
-\multirow{2}{*}{POD} & 0.06 & 0.6354 & 0.6575 & 2.68 & 2.74 & \textbf{4.39} & \textbf{4.38} \\
-\cline{2-8}
-  & 0.12 & 0.7255 & 0.7446 & 2.83 & 2.71 & \textbf{4.84} & \textbf{4.89} \\
-\hline
-\multirow{2}{*}{FAR} & 0.06 & 0.1800 & 0.1609 & -0.72 & \textbf{-1.24} & \textbf{-1.22} & \textbf{-1.24} \\
-\cline{2-8}
-  & 0.12 & 0.2379 & 0.2208 & -0.21 & -0.18 & \textbf{-2.31} & \textbf{-2.72} \\
-\hline
-\end{tabular}
-\end{table}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Computational Performance}
-\label{ssec:compuational_performance}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-The NFIE experiments were able to produce HAND for 331 HUC6's in 1.34 CPU years \cite{liu2016cybergis} and estimates using work from \citeA{djokic2019arc} put producing HAND at the FR NWM at 0.55 CPU years.
-For our work, we were able to produce HAND at the full NWM resolution in 0.13 CPU years which represents a substantial speed-up compared to previous works.
-For the MS resolution, an additional 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
-GMS which generalizes HAND production to the LP scale adds a significant amount of CPU time to the process bringing the estimate total to about 1.17 CPU years.
-%
- %
-% Discussion
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-\label{sec:discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Overall, the main observation of this study was how FIM performance can be improved by reducing the Horton-Strahler stream order of the target stream network used for HAND computation.
-Most of this change is accounted for by substantially increasing POD and inundation extents in some areas thus reducing FNs.
-We believe, as we later argue, that the increase in POD is primarily driven by an increase in the catchment sizes that is inherently enabled by dividing up the stream network into independent stream networks of unit stream order.
-Additionally, we note that reducing drainage order also has some minor influence on reducing inundation extents in other areas and the rate of false alarms.
-We believe that this effect is driven by a change in the stage-discharge relationship where a given streamflow leads to lower river stage values when HAND is computed with target stream networks of unit drainage order.
-We seek to explain that these two effects, catchment boundary enlargement and stage-discharge curve lowering, are highly interrelated and cannot be easily detangled.
-Lastly, we discuss the diminishing effects on performance that the MS and GMS techniques may have and also any additional effects including enhanced cross-walking abilities.
-
-As evident in the results of the study in Section \ref{sec:results}, a sizable amount of the increase in CSI observed by reducing stream order for HAND computation can be attributed to increases in POD.
-This can be inferred by close inspection of Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} where changes in POD are significantly higher than that of FAR.
-%This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
-Upon investigation of the performance of HAND derived FIM, we note a general increase of catchment sizes for the MS and GMS methods when compared to the FR method as they are now delineated independently of any tributaries that would constrain catchment sizes otherwise.
-Additionally, we note significantly less water build up along catchment boundaries especially at the junction of lower order tributaries with lower flow rates and higher order rivers with more flow.
-This allows for inundation extents to expand across regions previously encapsulated by catchments of joining reaches in lower flow tributaries.
-The water built up along the catchment boundaries can be thought of as a column of water in a cylindrical container (catchment) that has exceeded the elevation of the container's rim which does not represent accurate physics.
-
-Large scale HUC8 level evaluations can fail to demonstrate fine grain enhancements as they aggregate away many changes that are only clear at more local scales.
-Future assessments of OWP FIM should consider finer grain evaluation units as well possible impact assessments using asset information such as building footprints to better illustrate fine grain changes in a more relevant manner to stakeholders.
-For now, we provide Figure \ref{fig:gms_enhancement} which best illustrates the improvement offered by multi-source fluvial flooding capabilities in a more local context.
-The figure is comprised of two agreement rasters for two different HAND based FIMs compared to the validation dataset for a given region with a high flow mainstem (500 yr recurrence flow) running horizontally along the region.
-Sub-figure \ref{fig:gms_enhancement}a demonstrates the agreement raster for the FR stream network as well as its respective catchment boundary lines symbolized in white and stream network shown in green.
-Inspection of this sub-figure denotes a clear spatial pattern where TPs or areas correctly inundated are pooled alongside catchment boundary lines. 
-On the other side of the catchment boundary, one can witness large swaths of FNs that should be inundated. 
-The FNs also exhibit a spatial pattern as in they tend to collocate within catchments of the pictured mainstems tributaries.
-This sort of behavior was introduced early in the paper and shown qualitatively in Figure \ref{fig:catchment_boundaries_issue}.
-
-As an enhancement, this paper proposes computing HAND for stream networks comprised of level-paths independently of one another.
-In sub-figure \ref{fig:gms_enhancement}b, the agreement raster for the GMS technique is illustrated as well as the stream network lines in green.
-While the entire mosaiced inundation map from GMS (as described in Section \ref{ssec:inundation_mapping} and Equation \ref{eq:hand_fim_depth}) is used to produce this agreement map, we only show the catchments associated with the mainstem of this region that is shown to follow a clear horizontal path.
-Showing all the catchments for the tributaries that were all derived independently would convolute the image.
-The main message illustrated here is that the catchments associated with the mainstem of this area significantly increase in size.
-Since they are not restricted by the catchments of tributaries that lie in the same drainage areas as those of the mainstem, they extend to include those as well.
-The consequence for inundation extents is a general increase in spatial coverage of the river's water which shows to be in much better agreement with the benchmark map.
-The TPs are no longer bounded by the catchment lines and allowed to expand to their natural extents.
-
-We note here as a contribution of this study that a major inherent, limitation of HAND is the ``nearest drainage'' constraint or the idea that a given river reach only drains or, in HAND's case, inundates its immediate, unique drainage area or catchment.
-In other words, HAND based FIMs are limited to producing fluvial inundation to only their nearest drainage area or catchment.
-However, we know that fluvial inundation can be sourced from several streams nearby that also serve as drainage outlets to the area in question.
-Generally speaking, drainage areas are known to be hierarchical in nature so a given drainage area for a given outlet point can be decomposed into nested drainage areas for outlet points that lie in the original drainage area.
-A perfect example of this are points that lie on tributary reaches closely neighboring a mainstem.
-These points lie in the drainage area of reaches in the mainstem but inundation from the mainstem cannot reach these tributary catchments because of the ``nearest'' assumption in HAND.
-Hence it's important to state that just like there are different sources of flooding such as fluvial, pluvial, groundwater, reservoir, barrier failure (dam/levee/embankment), and coastal, there can also be multiple sources of a riverine flood.
-HAND is only equipped to handle riverine flooding from the nearest drainage line.
-Other relevant drainage lines that produce fluvial flood waters are not considered here especially if the routing model used doesn't consider backwater effects.
-
-The remaining portion of the improvement in CSI was found to come from a marginal yet notable reduction in FAR.
-Upon investigation of the spatial results in the agreement maps, we found some areas of slight reductions in FPs especially where changes in catchment boundaries may have occurred due to the reduction in effective stream order in computing HAND.
-These observations pointed to changes in the SRCs introduced by stream order reduction and catchment definition adjustments.
-Figure \ref{fig:rating_curve_comparison} illustrates the general effect that stream order reduction has on SRCs.
-Sub-figure \ref{fig:rating_curve_comparison}a shows how the average SRCs for all reaches for stage values 0 to 25 meters at one-third meter intervals tend to shift the curve down and to the right with ever increasing stream order reduction (FR to MS to GMS).
-This bias is more pronounced for GMS since that implements stream order reduction down to the unit level for the entire FR network while MS only does so for 4-5\% of the network.
-
-Attempting to diagnose this bias in the SRC leads one to Equation \ref{eq:reach_averaged_mannings_equation} which shows the reach averaged SRC relationship between stage and discharge.
-Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
-While the channel slope and reach lengths are not exactly the same across methods, their averaged differences are very negligible which only leaves room for deviations in volume and bed area.
-Again, volume (V(y) or simply V) is synonymous to reach-averaged cross-sectional area and bed area (B(y) or B) is analogous to reach-averaged hydraulic radius but these associations only hold when reach length, L, is considered.
-Discharge, Q, is directly related to volume and inversely related to bed area and each parameter is weighed according to the magnitude of its exponent which are $\frac{5}{3}$ and $\frac{2}{3}$ respectively (see Equation \ref{eq:reach_averaged_mannings_equation}).
-Figures \ref{fig:rating_curve_comparison}b and \ref{fig:rating_curve_comparison}c show how volume and bed area compare across the three methods with GMS having significantly greater values than MS which has greater values than FR.
-Again the relative discrepancy between FR vs MS and MS vs GMS is explained by the extent of their spatial coverages.
-Both V and B values increase but are weighed differently by their respective exponents and pull Q in different directions.
-We show in Figure \ref{fig:rating_curve_comparison}d the relationship of $\frac{V^{5/3}}{B^{2/3}}$ and plot this ratio against stage, y, to show how these two parameters collectively pull the Q up and changes the SRC accordingly.
-In other words, the magnitude and weight of the volume at each stage level exceeds the influence of the magnitude and weight of the bed area.
-Both parameters are set to increase mainly due to much larger catchments leading to more pixels at each stage level as shown in Figure \ref{fig:rating_curve_comparison}e.
-
-Much of the increase in inundated pixels, volume, and bed area can be explained by much larger catchments that encompass neighboring tributaries.
-These tributaries have a significant amount of bathymetry that is low-lying thus easily included in the geometry for the SRC derivation. 
-They also contribute volume and bed area that is technically not perpendicular to the flux of streamflow being accounted for in the stream in question. 
-Careful examination of Figure \ref{fig:gms_enhancement}b shows how much larger catchments include neighboring tributaries and the geometry associated with those tributaries. 
-This geometry is not perpendicular to the flow that is associated with the main reach thus leading to biases in the SRC.
-We consider this to have a nuanced effect on skill, while reducing the rate of FPs it also can lead to FNs due to biases in the SRC.
-
-We note that reducing stream order does suffer from diminishing returns where the increase in mapping skill for applying stream order reduction to roughly 4-5\% of the stream network is about the same as the increase for applying stream order reduction to the remaining 95-96\% of the stream network.
-This motivates further work in identifying what the optimal coverage of stream order reduction could be and how to parameterize that coverage. 
-One option could be removing lower stream orders (e.g. 1 and 2) from stream order reduction and simply using the inundation from FR from these areas.
-
-Additional analysis of Figure \ref{fig:gms_enhancement}a, reveals that some catchments don't have inundation or significant inundation.
-While the cause of these errors can be varied, we assert here that conflating four networks for use in evaluations leads to significant error.
-Section \ref{sssec:cross_walking_networks} details how reach identifiers are conflated for the FIM network back to that of the NWM. 
-One of the issues with the FR version of HAND occurs when a reach of given stream order accidentally conflates to that of a neighboring tributary that is of lower order which leads to areas of FNs.
-The utilization of MS and GMS only conflates to NWM catchments directly associated with the LP in question which is inherently easy to do with those methods. 
-Thus part of the improvement in MS and GMS methods is due to a slight improvement in cross-walking methodology.
-The NWM stream network was derived using the NHDPlus V2 dataset which was derived from coarser DEMs than those used here. 
-Additional conflation is identified in cross-walking the stream network used by the BLE maps and those of HAND.
-Until a singular stream network is used for the NWM, BLE benchmark, and for HAND based FIM, conflation will continue being a source of error.
-
-Our qualitative analysis suggests that the SRCs offer a significant opportunity for improvement in HAND based FIM for future development.
-The bathymetry of the 10 m DEM from 3DEP is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
-Manning's n which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
-Some adjusting parameter must be introduced to enhance the estimation of the bathymetric representation.
-Lidar DEMs from the USGS at 3 m and 1 m scale could be utilized to derive HAND as well which we conject should show better agreement with higher fidelity FIMs also derived from the same Lidar based DEMs.
-We suspect that a significant amount of the difference in performance between 100 yr and 500 yr magnitude events can be attributed to poor SRC performance due to poor bathymetric representation.
-Lower magnitude events are, logically, more susceptible to poor bathymetric data due a greater proportion of the inundation being attributable to areas that are more typically under normal flow conditions.
-Higher flow events tend to cover regions with more floodplain inundation thus less sensitive to errors from bathymetric data quality.
-On a related note, the use of the AGREE DEM method discussed in Section \ref{ssec:stream_network_enforcment} also interacts with the bathymetry issue introducing several artificial geometry parameters that affect SRC shape and quality.
-Due to focus on the nearest drainage problem, we leave future work related to SRC representation including roughness estimation, bathymetric data assimilation, and bathymetry adjustments as opportunities for major enhancements in HAND based FIM.
-
-Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
-For HAND to be derived and thus create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
-In other words, an entire area eligible for flooding must monotonically decrease in elevation. 
-DEM's naturally don't do this and the dynamics of true flood events don't follow drainage patterns.
-Enforcing this assumption for HAND leads to significant amount of DEM manipulations that introduce basic errors.
-These errors are deep into the assumptions of HAND and thus more difficult to disentangle.
-Ultimately, the use of more advanced 2D hydrodynamic models should be considered for dealing with this limitation of HAND but would come at significant expense at the given high resolution across very large spatial scales and frequent forecast resolutions.
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/rating_curve_comparison.jpg}
-\caption{Illustrates average quantities for the three methods, FR, MS, and GMS, for each stage value (m). 
-The values are (a) Discharge $m^3s^{-1}$, (b) Volume $m^3$, (c) Bed Area $m^2$, (d) a function of Volume and Bed Area, and (e) number of pixels.
-}
-\label{fig:rating_curve_comparison}
-\end{figure}
-%
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/gms_enhancement.jpg}
-\caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105 at the 500 yr recurrence magnitude.
-Catchment boundaries and stream lines are shown in white and dotted green, respectively.
-Sub-figure (a) shows agreement of FR HAND denoting significant areas of under-prediction due to junctions and catchment boundaries.
-Meanwhile, (b) shows the agreement for GMS and much larger catchments leading to much better inundation agreement for this given reach. 
-Overall, this illustrates the benefits of stream order reduction for deriving HAND datasets.
-}
-\label{fig:gms_enhancement}
-\end{figure}
-%
- %
-% Conclusions
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-\label{sec:conclusions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-Floods present a significant, under-served, and expanding risk to life, property, and resources.
-Previous flood forecasting systems lacked the coverage to adequately inform society of these risks.
-The National Water Model (NWM), developed by the National Oceanic and Atmospheric Administration's (NOAA) Office of Water Prediction (OWP) and the National Center for Atmospheric Research (NCAR), provides increased spatial coverage and resolution as well as additional forecast time horizons on mostly hourly intervals.
-Additional modeling is required to convert streamflows from the NWM to river stages and finally to flood inundation maps (FIM).
-Height Above Nearest Drainage (HAND) is a means of detrending digital elevations maps (DEM) by normalizing elevation to the nearest relevant drainage point.
-HAND coupled with the use of reach averaged synthetic rating curves (SRC) provide such a means of creating continental scale FIM capabilities at high resolutions (1/3 arc-second, 10 m) and high temporal frequencies (up to 1 hr).
-Scalable, open-source software, known as OWP FIM, was developed to produce HAND and associated datasets (catchments, SRCs, and cross-walking data) for the NWM forecasting area including Hawaii and Puerto Rico \cite{inundationMapping2022}.
-HAND is produced using the latest hydro-conditioning techniques to enforce monotonically decreasing elevations including stream burning, levee enforcement, pit-filling, stream channel excavation, thalweg breaching, headwater seeding, stream reach resampling, and more.
-Finally, we used this implementation to investigate the skill of the FIMs by varying the scale of the processing units used to derive HAND.
-FIM skill was evaluated over large spatial scales by comparison to HEC-RAS 1D models.
-
-The main contribution and conclusion of this work centers around a fundamental limitation in HAND based FIM which is a failure to account for multiple possible sources of fluvial inundation since HAND only considers inundation from the nearest drainage line.
-We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
-In order to reduce stream order of the NWM stream network for HAND computation, we dissected the NWM network into two simpler units of singular Horton-Strahler stream order and mosaiced the resulting FIMs derived from each.
-The NWM Mainstems (MS) stream network, which covers roughly 4-5\% of the NWM Full Resolution (FR) network, spans all established forecasting points in the Advanced Hydrologic Prediction System (AHPS) and downstream reaches.
-The inundation from MS derived HAND is mosaiced together with the inundation of FR derived HAND.
-Extending order reduction to the entire network, the Generalized Mainstem (GMS) technique discretizes the NWM FR network into level paths (LP) of unit stream order for HAND computation.
-All LP based FIM derived from LP based HAND datasets are mosaiced together to form one seamless FIM.
-Dissecting the stream network into regions of LPs with unit stream order is necessary because HAND has a ``nearest drainage'' limitation meaning it only accounts for riverine inundation sourced from the nearest drainage line.
-In our evaluation of this technique, we observe that HAND based FIM improves in skill as we extend from nearest drainage inundation in FR to multiple drainage support in MS for only 4-5\% of the FR network.
-Extending multiple drainage support to the entire FR network with GMS based HAND improves skill at around the same magnitude that MS improved upon FR.
-Thus we conclude that deriving HAND with independent stream networks of unit Horton-Strahler stream order enhances the skill of FIM but offers diminishing returns as we extend from 4-5\% of the network with MS to 100\% of the network with GMS since deriving HAND and FIMs at these localized scale does add computational expense.
-
-This primary contribution also affects the parameters used to compute stage-discharge relationships shifting discharge higher at given stages which reduced the rate of false positives.
-This shift in SRC behavior is driven by larger catchments that influence reach averaged geometric parameters in the Manning's equation.
-Related to SRCs, we noted better results and more sensitivity to unit stream order networks with the higher Manning's n value of 0.12 when compared to 0.06 for high magnitude events at 1\% (100 year or yr) and 0.2\% (500 year or yr) recurrence intervals.
-Additionally, we noted better performance for more intense 500 yr events which we attribute to a stronger influence of poor quality bathymetric data in 100 yr magnitude inundation extents.
-While the AGREE DEM procedure is meant to add some bathymetry primarily motivated to enhance catchment and stream line delineation, it does introduce three parameters that have major implications in the quality of SRCs and the resulting FIMs.
-Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
-We leave questions related to Manning's n localization as well as bathymetry integration, estimation, and/or calibration open for future research to answer.
-Two other issues left open for improvement include the integration of higher resolution Lidar-based digital elevation maps (DEM) as well as the use of physics-based models for continental scale, high resolution forecasting applications.
-Due to inherent limitations with HAND, scalable, physics-based methods are necessary to consider to provide a better representation of flood extent dynamics in steady and unsteady conditions.
- %
-% Open Research
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Open Research}
-%\openresearch
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs \cite{nwm2022hydrofabric}.
-These are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
-OWP Flood Inundation Mapping (FIM) capabilities rely extensively on the use of the National Hydrography Plus High Resolution (NHDPlusHR) datasets including BurnLineEvents \cite{nhdplus2022vectors}, value-added attributes (VAA) \cite{nhdplus2022vectors}, water boundaries (WBD) or hydrologic unit code (HUC) geometries \cite{nhdplus2022wbd}, and digital elevation maps (DEM) \cite{nhdplus2022dems}.
-Some additional datasets for processing include the National Levee Database (NLD) furnished by the United States Army Core of Engineers (USACE) \cite{engineers2016national}, Land-sea border from the Great Lakes Hydrography Dataset (GLHD) furnished by the Great Lakes Aquatic Habitat Framework (GLAHF) \cite{GreatLakesHydrographyDataset}, and a Land-sea border provided by OpenStreetMap (OSM) \cite{osm2021landsea}.
-Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
-Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
-
-Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022}.
- %
-% Acknowledgments
-\clearpage % this clears figures before references
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Acknowledgments
-\acknowledgments
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-This work was funded by the Office of Water Prediction (OWP) which is part of the National Oceanic and Atmospheric Administration's (NOAA) National Weather Service (NWS).
-Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
-We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
-Additionally, David Blodgett from the United States Geolgical Survey (USGS) Water Mission Area was instrumental in helping define level paths and other hydrographic work.
-More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
-Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
- %
-% Bibliography
-\clearpage % this clears figures before references
-\bibliography{bibliography/owp_fim4_2022}
-%
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-%\section{= enter section title =}
-%Text here ===>>>
-%\section{= enter section title =}
-%Text here ===>>>
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables;  other caption information
-%  should be placed in last line of the table, using
-% \multicolumn2l{$^a$ This is a table note.}
-%
-%----------------
-% EXAMPLE FIGURES
-%
-% \begin{figure}
-% \includegraphics{example.png}
-% \caption{caption}
-% \end{figure}
-%
-% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
-% \begin{figure}
-% \noindent\includegraphics[width=\textwidth]{anothersample.png}
-%\caption{caption}
-%\label{pngfiguresample}
-%\end{figure}
-%
-%
-% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
-This is common when trying to add a PDF figure without pdflatex.
-% \begin{figure}
-% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
-%\caption{caption}
-%\label{pdffiguresample}
-%\end{figure}
-%
-%
-% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
-%
-
-%
-% ---------------
-% EXAMPLE TABLE
-%
-% \begin{table}
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \centering
-% \begin{tabular}{l c}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Footnote text here.}
-% \end{tabular}
-% \end{table}
-
-%% SIDEWAYS FIGURE and TABLE
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{figsamp}
-% \caption{caption here}
-% \label{newfig}
-% \end{sidewaysfigure}
-%
-%  \begin{sidewaystable}
-%  \caption{Caption here}
-% \label{tab:signif_gap_clos}
-%  \begin{tabular}{ccc}
-% one&two&three\\
-% four&five&six
-%  \end{tabular}
-%  \end{sidewaystable}
-
-%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
-%\begin{linenomath*}
-%\begin{equation}
-%y|{f} \sim g(m, \sigma),
-%\end{equation}
-%\end{linenomath*}
-
-%%% End of body of article
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary, Notation or Acronym section goes here:
-%
-%%%%%%%%%%%%%%
-% Glossary is only allowed in Reviews of Geophysics
-%  \begin{glossary}
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \end{glossary}
-
-%
-%%%%%%%%%%%%%%
-% Acronyms
-%   \begin{acronyms}
-%   \acro{Acronym}
-%   Definition here
-%   \acro{EMOS}
-%   Ensemble model output statistics
-%   \acro{ECMWF}
-%   Centre for Medium-Range Weather Forecasts
-%   \end{acronyms}
-
-%
-%%%%%%%%%%%%%%
-% Notation
-%   \begin{notation}
-%   \notation{$a+b$} Notation Definition here
-%   \notation{$e=mc^2$}
-%   Equation in German-born physicist Albert Einstein's theory of special
-%  relativity that showed that the increased relativistic mass ($m$) of a
-%  body comes from the energy of motion of the body—that is, its kinetic
-%  energy ($E$)—divided by the speed of light squared ($c^2$).
-%   \end{notation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-%
-% The acknowledgments must list:
-%
-% >>>>	A statement that indicates to the reader where the data
-% 	supporting the conclusions can be obtained (for example, in the
-% 	references, tables, supporting information, and other databases).
-%
-% 	All funding sources related to this work from all authors
-%
-% 	Any real or perceived financial conflicts of interests for any
-%	author
-%
-% 	Other affiliations for any author that may be perceived as
-% 	having a conflict of interest with respect to the results of this
-% 	paper.
-%
-%
-% It is also the appropriate place to thank colleagues and other contributors.
-% AGU does not normally allow dedications.
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% \bibliography{<name of your .bib file>} don't specify the file extension
-%
-% don't specify bibliography style
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%Reference citation instructions and examples:
-%
-% Please use ONLY \cite and \citeA for reference citations.
-% \cite for parenthetical references
-% ...as shown in recent studies (Simpson et al., 2019)
-% \citeA for in-text citations
-% ...Simpson et al. (2019) have shown...
-%
-%
-%...as shown by \citeA{jskilby}.
-%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
-%...has been shown \cite{jskilbye}.
-%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
-%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
-%...has been shown by \cite <e.g.,>[and others]{lewin76}.
-%
-% apacite uses < > for prenotes and [ ] for postnotes
-% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
-%
-
-\end{document}
-
-%%%% More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-% If you are using line numbers, remember to surround
-% equations with \begin{linenomath*}...\end{linenomath*}
-
-%  To add line numbers to lines in equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}

### 2023-02-11 manuscripts/owp_fim4_2022/sections/discussion.tex
second review responses and changes submitted to coauthors 20230211

@@
 Higher flow events tend to cover regions with more floodplain inundation thus less sensitive to errors from bathymetric data quality.
-On a related note, the use of the AGREE DEM method discussed in Section \ref{ssec:stream_network_enforcment} also interacts with the bathymetry issue introducing several artificial geometry parameters that affect SRC shape and quality.
-Due to focus on the nearest drainage problem, we leave future work related to SRC representation including roughness estimation, bathymetric data assimilation, and bathymetry adjustments as opportunities for major enhancements in HAND based FIM.
+On a related note, the use of the AGREE DEM method discussed in Section \ref{ssec:stream_network_enforcement} and \ref{sec:app_stream_network_enforcement} also interacts with the bathymetry issue introducing several artificial geometry parameters that affect SRC shape and quality.
+To reiterate in this discussion, Section \ref{ssec:evaluation} pointed out that the BLE's representation of reservoir inundation extends beyond the NWM's defined reservoir boundaries, which results in an under-prediction in the OWP FIM that is reflected in the computed metrics.
+Due to focus on the nearest drainage problem, we leave future work related to SRC representation including roughness estimation, bathymetric data assimilation and adjustments as well as reservoir inundation as opportunities for major enhancements in HAND based FIM.
 
-Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
+Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment, and reservoir representation are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
 For HAND to be derived and thus create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
@@
 %
+%
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale=1.0]{figures/rating_curve_comparison.jpg}
+    \caption{Illustrates average quantities for the three methods, FR, MS, and GMS, for each stage value (m). 
+    The values are (a) Discharge $m^3s^{-1}$, (b) Volume $m^3$, (c) Bed Area $m^2$, (d) a function of Volume and Bed Area, and (e) number of pixels.
+    }
+    \label{fig:rating_curve_comparison}
+    \end{figure}
+    %

### 2023-02-11 manuscripts/owp_fim4_2022/sections/introduction.tex
second review responses and changes submitted to coauthors 20230211

@@
 \subsection{Office of Water Prediction Flood Inundation Mapping}
+\label{ssec:owp_fim}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-In order to mitigate the ever increasing threat of flooding to life and property, an operational capability is required to extend NWM streamflow forecasts to river stages, inundation extents, and inundation depths.
+In order to mitigate the ever increasing threat of flooding to life and property, an operational capability is required to extend NWM streamflow forecasts to river stages and inundation extents.
 OWP FIM is introduced here as a continental scale capability that generates these products at high spatial and temporal resolutions.

### 2023-02-11 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
second review responses and changes submitted to coauthors 20230211

@@
 \subsubsection{Stream Network Enforcement} 
-\label{ssec:stream_network_enforcment}
+\label{ssec:stream_network_enforcement}
 %
-The location of the stream network is enforced to ensure general agreement with the NWM network which is used for forecasting the streamflow inputs.
+Both the location and geometry of the stream network are enforced to agree with the NWM stream network and to enforce synthetic bathymetry and generate more hydrologically correct catchments.
 The NHDPlusHR Beta BurnLineEvent layer is used to enforce stream locations in the NHDPlusHR workflow and best agrees with thalweg locations in the DEM used so it is also used here for hydro-enforcement \cite{moore2019user}.
-
-However, to better match the drainage density of the NWM FR V2.1 stream network, which is based on the NHDPlus V2, the BurnlineEvents are pruned utilizing a nearest neighbor search around the NWM flowlines.
-Headwater points are first derived for the NWM FR V2.1.
-For every NWM headwater point, the nearest NHDPlusHR point is selected and placed into a set while those excluded are discarded.
-Only the nearest point on the NHDPlusHR is used so any portion of the NHDPlusHR network upstream of this nearest point is discarded to avoid extending inundation too far above the modeling domain.
-The points in this nearest neighbor set are then traversed downstream.
-Any headwater portion in the NHDPlusHR or any other stream not traversed are pruned away to better match the resolution and spatial locations of the NWM stream network and its headwater points.
-The resulting pruned NHDPlusHR stream network gets hydro-enforced in subsequent operations.
-This procedure is best illustrated in Figure \ref{fig:stream_density_pruning} which shows that the pruned NHDPlusHR network corresponds to the NHDPlusHR network at or downstream of NWM V2.1 headwater locations only.
-Additionally, the NHDPlusHR pruned headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents. 
-%
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{This figure illustrates some of the datasets that result from the pruning of NHDPlusHR Beta BurnlineEvents (dotted black) to the stream density of the NWM FR V2.1 density (blue).
-The stream network used for forecasting, NWM FR V2.1, is of lower stream density than that of the NHDPlusHR which has better agreement with the thalweg locations in the DEM used.
-Thus, we opt to prune the NHDPlusHR network to match the general location and density of the NWM network.
-The nearest neighbor segment in the NHDPlusHR of each NWM headwater locations and the nearest point on that segment is determined to match the closest point to that of the NWM headwater.
-These points are then traversed downstream and any segments not traversed are pruned away.
-The resulting stream network (red) matches the drainage density of NWM V2.1 while corresponding spatially with the NHDPlusHR BurnlineEvents.}
-\label{fig:stream_density_pruning}
-\end{figure}
-%
-
-The pruned stream network is then utilized to hydro-enforce the DEM with a methodology developed by \citeA{hellweger1997agree} known as the AGREE DEM Surface Reconditioning System. 
-The AGREE algorithm seeks to burn artificially deep thalweg elevations by a uniform value known as sharp drop. 
-The modification continues by excavating an area of a given buffer distance from the thalweg by a depth proportional to the distance from the channel given by the smooth drop and buffer distance. 
-The resulting enforcement of the thalweg and general bathymetric region results in a cross-section resembling an inverted triangular notch shape with a significantly lower elevation along the thalweg line only.
-In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop which were set to fixed values of 70 m, 1000 m, and 10 m, respectively, but available to the user via the parameter file.
-While the values for these parameters are critical to the inundation extents produced, especially for lower flow rates where bathymetric information has more influence, finding their optimal values for OWP FIM was not done since it was out of the main scientific scope of this article.
-Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
-\citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement techniques especially when computational costs are considered. 
+This network goes through a stream density reduction procedure to better match the stream density of the NWM stream network.
+This reduced density NHDPlusHR network is now used for hydro-enforcement along with the AGREE DEM Surface Reconditioning System to excavate channels into the DEM \cite{hellweger1997agree}.
 Downsides to the technique include the possibility of exhibiting parallel streams where the burned stream and real stream are both represented \cite{hellweger1997agree,saunders1999preparation} and some distortion of the catchment boundaries can also be observed \cite{saunders1999preparation,saunders1996gis}.
 Some of these drawbacks are addressed by additional conditioning techniques applied later on.
+For more details on how we reduce the density of the NHDPlusHR stream network along with technical implementation information on the AGREE DEM procedure, please see \ref{sec:app_stream_network_enforcement}.
 %
@@
 %
-Possible benchmark FIM candidates for evaluation purposes include high water marks, remote sensing observation, crowd-sourced information, and modeled extents.
-These sources are all subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to a lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
-While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to as to isolate out other hydrological factors.
+OWP FIM is linked with the NWM for operational purposes, utilizing streamflow inputs, to produce FIM at a continental scale.
+We evaluated various validation data sources for operational use, each with its own strengths and weaknesses, including high water marks (HWMs) \cite{musser2017characterization,watson2017characterization,breaker2016characterization}, remote sensing observations \cite{aristizabal2020high,aristizabal2021mapping}, and modeled extents from hyper-resolution, local models \cite{mcenery2005noaa}.
+Although they have advantages, the three data sources mentioned do not have streamflow information, which is required by OWP FIM.
+Using the NWM as a source of streamflow input would be a logical choice due to its operational use, but this would introduce hydro-climatic uncertainties that could impact the results of adding a multi-fluvial source extension to HAND.
+On the other hand, relying on point sources of streamflow or stage, such as gages, would eliminate this limitation, but limit the information available to only a single source of inundation at a specific site.
+Given these limitations, we investigated the use of existing HEC-RAS based models to address the limitations caused by hydro-climatic uncertainties and sparse streamflow inputs.
 
-Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
-This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, higher level of sophistication when compared to HAND, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
+Our HAND based approach coupled with SRCs requires streamflow as input and is agnostic as to the source of that streamflow whether forecasted, observed, or probablistic.
+Due to this fact, evaluation of our relative elevation CFIM method was conducted by comparison to the HEC-RAS 1D models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
+This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, higher level of sophistication when compared to HAND, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid,bates201630m,wing2017validation,hu2021real,li2022accounting,li2022comprehensive,zheng2018geoflood}.
 We selected 49 available HUC8s, shown in Figure \ref{fig:all_ble_maps}, which span about 185 thousand $km^2$ across nine states.

### 2023-02-11 manuscripts/owp_fim4_2022/reviews/second_review/makefile_responses.sh

@@ -1,0 +1,13 @@
+#!/bin/bash -e
+
+#parent_path=$( cd "$(dirname "${BASH_SOURCE[0]}")" ; pwd -P )
+#cd $parent_path
+
+fullFile=responses_aristizabal_et_al_202302.tex
+baseName="${fullFile%%.*}"
+extension="${fullFile#*.}"
+
+pdflatex "$fullFile"
+bibtex "$baseName".aux
+pdflatex "$fullFile"
+pdflatex "$fullFile"

### 2023-04-09 manuscripts/owp_fim4_2022/owp_fim4_expanded.tex
submitted to WRR April 2023

@@
 %
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{1,4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
+\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{1,3}, Trevor Grout\affil{1,3}\thanks{Colorado Basin River Forecast Center, National Weather Service, National Oceanic and Atmospheric Administration, Salt Lake City, UT, USA}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{1,3}, Jasmeet Judge\affil{2}}
 %
@@
 \affiliation{3}{National Water Center, Office of Water Prediction, National Weather Service, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
-\affiliation{4}{Hydrologic Applied Innovations Lab, New York Water Science Center, United States Geological Survey, Troy, NY, USA}
-\affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Maryland, College Park, MD, USA}
 %
@@
 \begin{keypoints}
-\item Flood maps derived from Height Above Nearest Drainage (HAND) are subject to nearest drainage line limitations that affect inundation skill.
+\item Flood maps derived from Height Above Nearest Drainage (HAND) are subject to nearest flowpath limitations that affect inundation skill.
 \item A means of resolving this limitation is provided by reducing HAND processing units to level paths with effective unit stream order.
@@
 Height Above Nearest Drainage (HAND), a drainage normalizing terrain index, is a means able of producing flood inundation maps (FIMs) from the National Water Model (NWM) at large scales and high resolutions using reach-averaged synthetic rating curves.
-We highlight here that HAND is limited to producing inundation only when sourced from its nearest drainage line, thus lacks the ability to source inundation from multiple fluvial sources.
+We highlight here that HAND is limited to producing inundation only when sourced from its nearest flowpath, thus lacks the ability to source inundation from multiple fluvial sources.
 A version of HAND, known as Generalized Mainstems (GMS), is proposed that discretizes a target stream network into segments of unit Horton-Strahler stream order known as level paths (LP).
@@
 Flooding is one of the most impactful natural disasters on life and property. 
-The United States National Water Model (NWM) provides flood forecasts for the entire country so that adequate warnings can be raised to the public to enable safe evacuations and protective measures.
-In order to convert flow rates from the NWM to flood inundation maps (FIM), a model, known as Height Above Nearest Drainage (HAND), is used that converts elevation data from height above mean sea-level to height above the nearest river bottom.
-This model suffers from issues in mapping performance because inundation sourced from rivers is only considered from the nearest river line.
+The United States National Water Model (NWM) provides flood forecasts to adequately warn people for safe evacuations and protective measures across the entire country.
+To convert streamflow from the NWM to flood inundation maps (FIM), a model, Height Above Nearest Drainage (HAND), is used that translates elevation data from height above mean sea-level to height above the nearest river.
+This model suffers from issues in mapping performance because inundation sourced from rivers is only considered from the nearest river.
 We developed a technique that mitigates these errors by removing consideration for neighboring tributaries in the relative elevation computation process.
-This is done by splitting the stream network into continuous river segments known as level paths (LPs).
-These LPs have no tributaries, thus are known to be stream lines with a unit stream order indicating no branching.
+This is done by splitting the stream network into continuous river segments known as level paths (LPs) which removes the effects of tributaries.
 HAND is computed independently for each LP and the resulting FIMs are mosaiced together to form one seamless map.
+By computing HAND and catchments on an LP scale, catchments and FIMs are allowed to overlap which can account for multiple river sources of inundation especially at river confluences.
 We compared these HAND derived FIMs to maps from physically-based models and found improvement in mapping performance.
@@
 %
-HAND normalizes topography along the nearest drainage path and it has been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and groundwater dynamics \cite{renno2008hand,nobre2011height}.
 \citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
@@
 HAND, to our knowledge, has only been applied to natural, inland, and riverine inundation applications thus it is also missing pluvial, coastal, ground water, and dam break components among other possible sources of flooding.
-Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest stream line which is used for catchment allocation and relative elevation calculation \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,maidment2017conceptual,garousi2019terrain,zheng2018river,zheng2018geoflood,johnson2019integrated,renno2008hand,hocini2020performance}.
+Additionally, in order to flood an area, HAND assumes all areas eligible for inundation must drain to some nearest flowpath which is used for catchment allocation and relative elevation calculation \cite{nobre2016hand,nobre2011height,liu2016cybergis,liu2020height,maidment2017conceptual,garousi2019terrain,zheng2018river,zheng2018geoflood,johnson2019integrated,renno2008hand,hocini2020performance}.
 Stream thalweg networks must also collectively drain to a singular outlet point for a given processing region \cite{nobre2016hand,zheng2018geoflood,renno2008hand}.
@@
 
-When used for FIM applications, HAND assumes only fluvial inundation sourced from its nearest drainage line is accounted for \cite{nobre2016hand,mcgehee2016modified}.
+When used for FIM applications, HAND assumes only fluvial inundation sourced from its nearest flowpath is accounted for \cite{nobre2016hand,mcgehee2016modified}.
 Catchments are independent of one another for FIM purposes meaning a reaches' stage value is only used to threshold the HAND values within its respective catchment \cite{liu2016cybergis,zheng2018river,zheng2018geoflood}.
@@
 Later in this study, we will re-introduce this problem and demonstrate how we initialize with a stream network (that of the NWM's) and thus avoid having to threshold accumulations to some arbitrary value to define stream networks.
-We illustrate how computing HAND independently for stream lines of unit stream order can significantly enhance FIM performance by accounting for multiple sources of fluvial inundation that may exist in certain regions and flow scenarios.
+We illustrate how computing HAND independently for flowpaths of unit stream order can significantly enhance FIM performance by accounting for multiple sources of fluvial inundation that may exist in certain regions and flow scenarios.
 %
@@
 By allocating 31 nodes at 20 cores per node for a total of 620 available cores to the overall operation, it enabled the production to finish up in 36 hours consuming 3.2 terrabyte (TB) of peak memory and 5 TB of total disk space.
-Originally, NFIE utilized the NHD Medium Resolution (MR) to etch or burn flowlines prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (NHDPlusHR) which better agrees with the 10 m DEM from the NHDPlusHR program \cite{liu2020height}.
+Originally, NFIE utilized the NHD Medium Resolution (MR) to enforce flowpaths prior to further conditioning but more recent work has advanced this to the more current NHDPlus High Resolution (NHDPlusHR) which better agrees with the 10 m DEM from the NHDPlusHR program \cite{liu2020height}.
 The original NFIE dataset was employed by the NWC as an unofficial demonstration to produce forecast FIM from the NWM for additional guidance in hydro-blind regions.
@@
 Previous authors dating back to the first HAND for FIM work by \citeA{nobre2016hand} have noted a sensitivity of mapping skill to the stream accumulation threshold which is closely related to stream density and the maximum Horton-Strahler stream order (or simply stream order) of the processing unit employed \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation}.
-HAND produced FIMs are limited in only providing inundation sourced from the nearest drainage line, however, depending on flow conditions and topography, a given area may have multiple contributing fluvial sources of surface water inundation.
+HAND produced FIMs are limited in only providing inundation sourced from the nearest flowpath, however, depending on flow conditions and topography, a given area may have multiple contributing fluvial sources of surface water inundation.
 The forecasting community, in reviewing HAND, have noted significant negative effects at the confluence of lower flow tributaries with higher flow rivers for which the phrase ``catchment boundary issue'' has been termed.
-In previous studies, FIM skill has been shown to be sensitive to the drainage density of the stream network employed as the datum for HAND which is closely related to the maximum Horton-Strahler stream order of the network \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
-This sensitivity is partly in due to the limitation that catchment boundaries place on inundation extents where only the nearest drainage line can source inundation for any particular area.
+In previous studies, FIM skill has been shown sensitivity to the drainage density of the stream network employed as the datum for HAND which is closely related to the maximum Horton-Strahler stream order of the network \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
+This sensitivity is partly in due to the limitation that catchment boundaries place on inundation extents where only the nearest flowpath can source inundation for any particular area.
 
@@
 An actual map with OWP FIM is generated using the NWM full-resolution stream network and compared with a FEMA 100 yr extent (see Section \ref{ssec:evaluation} for more details) showing significant under-prediction in inundation extent.
-The higher discharge along the main segment in Figure \ref{fig:catchment_boundaries_issue} of 1,900 cubic meters per second (CMS) does not translate to the lower flow rates along the tributaries of 84 and 195 CMS.
+The higher discharge along the main segment in Figure \ref{fig:catchment_boundaries_issue} of 1,900 cubic meters per second (cumecs) does not translate to the lower flow rates along the tributaries of 84 and 195 cumecs.
 This is due to a lack of representation of backwater conditions in the hydraulic routing techniques used.
@@
 Agreement maps are symbolized by false negatives (FN), true negatives (TN), false positives (FP), and true positives (TP) where inundated represents the positive condition (see Section \ref{ssec:evaluation} for more details).
-The streamflows associated with each river segment are shown in CMS while the flow directions are symbolized as red arrows.
+The streamflows associated with each river segment are shown in cumecs while the flow directions are symbolized as red arrows.
 The presence of FNs at the confluence of tributaries (circled in red) with the main segment is associated with lower flow rates in the tributaries that don't account for backwater effects.
-Additionally, the flow of 1900 CMS from the main segment cannot extend to the neighboring catchments belonging to its tributaries shown here.
+Additionally, the flow of 1900 cumecs from the main segment cannot extend to the neighboring catchments belonging to its tributaries shown here.
 Water pools up vertically along the catchment boundaries of the higher order segment distorting rating curve behavior (Section \ref{sec:discussion}).
-Sourcing fluvial inundation from HAND is limited to only its nearest drainage line which is the main issue this study aims to address.
+Sourcing fluvial inundation from HAND is limited to only its nearest flowpath which is the main issue this study aims to address.
 }
@@
 Later, we explain how only the NWM MS stream network is used for the MS version of HAND, while for GMS, the FR stream network is discretized into LPs before computing HAND.
-A series of hydro-conditioning steps enforces the location of stream lines, monotonically decreasing elevations, excavated bathymetry, stream thalweg breaching, and levee enforcement.
+A series of hydro-conditioning steps enforces the location of flowpaths, monotonically decreasing elevations, excavated bathymetry, stream thalweg breaching, and levee enforcement.
 After a DEM suitable for HAND's assumptions is conditioned, the FIM hydrofabric is generated including stream network, catchments, HAND, SRCs, and cross-walk table.
@@
 \centering
-\includegraphics[scale=6.0]{figures/methods_overview.jpg}
+\includegraphics[scale=0.17]{figures/methods_overview.jpg}
 \caption{
@@
 A user only needs to install Docker on their host machine and build the image from the provided recipe. 
-Source code is made available for this project on GitHub where a user could consult the Readme.md page for more information on how to acquire the datasets and reproduce the pipeline \cite{inundationMapping2022}.
+Source code is made available for this project where a user could consult the Readme.md page for more information on how to acquire the datasets and reproduce the pipeline \cite{inundationMapping2022,imHS2023}.
 %
@@
 Data sources used within OWP FIM are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
-Instructions for accessing data processed for OWP FIM are provided on the project's GitHub page via an Amazon Web Services (AWS) S3 bucket furnished by the Earth Science Information Partners (ESIP) \cite{esipData2022}.
+Instructions for accessing data processed for OWP FIM are provided on the project's GitHub or HydroShare page via an Amazon Web Services (AWS) S3 bucket furnished by the Earth Science Information Partners (ESIP) \cite{esipData2022,inundationMapping2022,imHS2023}.
 The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}.
 We utilized a series of data products from the NHDPlusHR including the BurnLineEvents \cite{nhdplus2022vectors}, Value Added Attributes (VAA) \cite{nhdplus2022vectors}, Water Boundaries (WBD) or HUC Layers \cite{nhdplus2022wbd}, and the DEM elevation rasters \cite{nhdplus2022dems}.
-These BurnLines used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk against for forecasting purposes \cite{nwm2022hydrofabric}.
+These BurnLines used in conjunction with the hydrofabric of the NWM V2.1 to help define flowpaths for OWP FIM while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk against for forecasting purposes \cite{nwm2022hydrofabric}.
 For enforcing levee data, the USACE NLD is used to burn feature elevations into DEMs \cite{engineers2016national}.
@@
 \hline
-USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlusHR for hydro-enforcement. & \cite{nhdplus2022vectors} \\
+USGS & NHDPlusHR BurnLineEvents & Flowpaths used by NHDPlusHR for hydro-enforcement. & \cite{nhdplus2022vectors} \\
 \hline
@@
 \hline
-NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting adapted from NHDPlus V2 NHDFlowline\_Network feature class. & \cite{nwm2022hydrofabric} \\
+NOAA-OWP & NWM Streams & Flowpaths used by NWM for routing and forecasting adapted from NHDPlus V2 NHDFlowline\_Network feature class. & \cite{nwm2022hydrofabric} \\
 \hline
@@
 In other words, all areas within a given processing unit for HAND must have monotonically decreasing elevations to enable eligiblity for flooding.
-Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic.
+Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowpaths, lakes, or drainage divides whether natural or anthropogenic.
 It can also be used to simulate more accurate bathymetry which is not accounted for in the 10 m DEM \cite{gesch2002national}.
@@
 %
-Local depressions are naturally occurring features of a DEM but must be addressed if a connected drainage network with continuous catchments are to be derived for flood modeling purposes with HAND.
+Local depressions are naturally occurring features of a DEM but must be addressed to derive a connected drainage network with continuous catchments for flood modeling purposes with HAND.
 The partially conditioned DEM was removed of depressions by filling areas with pits while preserving the stream and levee information previously enforced.
@@
 The FIM Hydrofabric is defined here as the collection of geospatial datasets that are used for converting NWM discharges into inundation extents.
-These datasets include the HAND or relative elevation model (REM) raster, reach-level catchments raster/polygons, DEM-derived streamlines, SRCs, and cross-walk table.
-Within the context of this section, we refer to HAND as a grid of relative elevations to detrend elevations away from mean level towards elevations referenced to the nearest drainage line (See Section \ref{sssec:hand}).
+These datasets include the HAND or relative elevation model (REM) raster, reach-level catchments raster/polygons, DEM-derived flowpaths, SRCs, and cross-walk table.
+Within the context of this section, we refer to HAND as a grid of relative elevations to detrend elevations away from mean level towards elevations referenced to the nearest flowpath (See Section \ref{sssec:hand}).
 SRCs are considered stage-discharge relationships that are derived synthetically using the Manning's equation (See Section \ref{sssec:synthetic_rating_curve}).
@@
 Flat resolution from flats endemic to the DEM or from depression filled regions is a costly, non-trivial procedure which was originally addressed by \citeA{garbrecht1997assignment} where flats are resolved by incrementing elevations iteratively.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
+OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found efficient and effective \cite{survila2016scalable,cybergis2016}.
 For more information on the derivations of flow directions and resolving flats, please see \ref{sec:app_flow_direction_and_flat_resolution}.
@@
 Catchments were derived using the D8 connectivity established by \citeA{o1984extraction}.
-Outlet points are set at the pixel center points of the delineated stream lines explained in Section \ref{sssec:deriving_fim_stream_network}.
+Outlet points are set at the pixel center points of the delineated flowpaths explained in Section \ref{sssec:deriving_fim_stream_network}.
 The outlets act as root nodes in a tree structure and the connectivity is traversed to derive the contributing, nearest drainage region for each outlet point.
@@
 Again, HAND assumes and requires processing areas to drain thus have monotonically decreasing elevations with hydrologically correct flow directions all leading to a singular outlet point.
-While this is required for the generation of DEM-derived catchments and stream lines, it is not necessarily required for the computation of the relative elevations.
+While this is required for the generation of DEM-derived catchments and flowpaths, it is not necessarily required for the computation of the relative elevations.
 Since the use of hydro-conditioning processes to fit the drainage requirement for HAND can be extensive, we found it more fitting to use the native elevations this final HAND computation thus avoid the use of manipulated values that fit modeling assumptions.
@@
 The reach averaged approach seeks to sample the geometry parameters in the Manning's equation \cite{gauckler1867etudes,manning1890flow} on a reach scale then dividing those by length. 
-Previously not reported in literature to our knowledge in this form, the reach averaged Manning's formula is derived to be 
+Previously not reported in literature to our knowledge in this form, the reach averaged Manning's formula is derived as 
 %
@@
 These values are applied universally to the entire forecasting domain across space, time, and discharge profiles.
-We note significant opportunity to enhance CFIM skill by better localizing Manning's n according to available data including but not limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated}.
+We note significant opportunity to enhance CFIM skill by better localizing Manning's n according to available data including but not limited to land cover, land use, stream order, stream geometry, drainage area, reach length, and discharge percentiles \cite{garousi2019terrain,johnson2019integrated,godbout2019error,zheng2018river}.
 For now and for the purpose of this study, we examine the SRCs with Manning's n set to both 0.06 and 0.12 which we hope will shed some light on the sensitivity of this parameter to HAND based FIMs.
@@
 For the FR version of HAND, we overlap the reach catchments derived in Section \ref{sssec:catchments} with the NWM catchments matching the ID of the NWM catchment that most overlaps the derived catchment for HAND.
-For two subsequent HAND methods, MS and GMS, discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, respectively, we find the mid-point of the derived stream reach line described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
+For two subsequent HAND methods, MS and GMS, discussed in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}, respectively, we find the mid-point of the derived stream reach described in Section \ref{sssec:deriving_fim_stream_network} and find the NWM catchment that contains the mid-point.
 Additionally, only relevant catchments from the NWM for the given LP are selected for cross-walking for methods in Sections \ref{sssec:nws_mainstems} and \ref{sssec:generalized_mainstems}.
@@
 In order to do this, we discretized the NWM network into LPs which when considered individually have unit Horton-Strahler stream orders.
-LPs group flowlines by maximizing the length of each flow path and minimizing the number of LP identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
-In order to derive LPs for the NWM FR stream network at the HUC8 processing area, we first compute arbolate sums which are defined as the cumulative drainage distance of all upstream drainage lines.
+LPs group flowpaths by maximizing the length of each flowpath and minimizing the number of LP identifiers within a given domain \cite{moore2019user,mckay2012nhdplus}. 
+In order to derive LPs for the NWM FR stream network at the HUC8 processing area, we first compute arbolate sums which are defined as the cumulative drainage distance of all upstream flowpaths.
 Arbolate sum is also inclusive of the current drainage reach as well.
@@
 Further work could be dedicated to tune this user exposed parameter to better balance its effect on FIM extents and computational expense since larger buffers create additional floating point calculations and storage requirements.
-For the time being, we designate this issue to be out of scope.
+For the time being, we designate this issue out of scope.
 
 At the LP scale, the methods in Sections \ref{ssec:hydro_conditioning} and \ref{ssec:deriving_fim_hydrofabric} are executed leaving out any tributaries of the LP in question at the time.
-The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement by burning these lines and seeding from its headwater points directly instead of going through the NHDPlusHR network as described in Section \ref{ssec:stream_network_enforcement} and \ref{sec:app_stream_network_enforcement}.
-This decision was motivated by the difficulty in deriving LPs in the NWM stream network with high agreement with the LPs derived for the NHDPlusHR stream lines.
+The only exception to this is the use of the NWM stream network directly for use with hydro-enforcement by burning these flowpaths and seeding from its headwater points directly instead of going through the NHDPlusHR network as described in Section \ref{ssec:stream_network_enforcement} and \ref{sec:app_stream_network_enforcement}.
+This decision was motivated by the difficulty in deriving LPs in the NWM stream network with high agreement with the LPs derived for the NHDPlusHR flowpaths.
 We found that the same algorithm to compute arbolate sums and LPs could yield enough disagreements associated with disordered branches or slight differences in arbolate sums that could significantly affect the agreement of the LP identifiers in the NWM and NHDPlusHR networks.
@@
 
-Once the NWM FR stream network is discretized into LPs, we independently compute HAND using each LP as the target stream network to be used.
+Once the NWM FR stream network is discretized into LPs, we independently compute HAND using each LP as the target stream network used.
 To illustrate the GMS procedure, we reference Figure \ref{fig:gms_methods} to show how deriving HAND and FIMs from GMS works.
@@
 For a more intimate look at the drainage order reduction procedure GMS, and its effects, we allude to Figure \ref{fig:gms_methods_2} which references the same area (in HUC8 12090301) and set of river junctions as in Figure \ref{fig:catchment_boundaries_issue}.
-The catchments and stream lines for HAND computed at the FR scale are illustrated in Figure \ref{fig:gms_methods_2}a where the respective inundation at the 100 yr magnitude is heavily constrained by the limited catchment extents especially at junctions.
+The catchments and flowpaths for HAND computed at the FR scale are illustrated in Figure \ref{fig:gms_methods_2}a where the respective inundation at the 100 yr magnitude is heavily constrained by the limited catchment extents especially at junctions.
 In subsequent sub-figures, we show the same datasets for the HAND computation problem for this region but discretized into independent LPs for the main LP (b), the eastern tributary (c), and the western tributary (d).
@@
 The LPs were derived on a HUC8 level (12020002) but only illustrated for a small, headwater subset of this HUC8 for clarity purposes.
-Arbolate sums are defined as the cumulative drainage distances of all upstream stream lines.
+Arbolate sums are defined as the cumulative drainage distances of all upstream flowpaths.
 Arbolate sums are computed for the NWM network by starting at the headwater points then traversing downstream and adding the distances cumulatively. 
@@
 \caption{Overall procedure for GMS HAND at HUC8 12090301.
-In (a), we illustrate all NWM stream lines symbolized by their LP with 372 unique LP IDs in this HUC.
+In (a), we illustrate all NWM flowpaths symbolized by their LP with 372 unique LP IDs in this HUC.
 Meanwhile (b), demonstrates the DEM clipped to a seven km buffer around two selected LPs.
@@
 By delineating catchments at this scale independent of the neighboring tributaries, the drainage areas are allowed to expand thus allowing inundation extents to cover previously restricted areas.
-In (c) and (d), we show the stream lines, catchments, and inundation extents of the two tributaries.
+In (c) and (d), we show the flowpaths, catchments, and inundation extents of the two tributaries.
 Later in Section \ref{ssec:inundation_mapping}, we describe how the inundation in (b), (c), and (d) are mosaiced together to form one seamless inundation map.
-This process allows for multiple, possible contributing sources of fluvial inundation to be considered thus enhancing FIM skill.
+This process allows to consider for multiple, possible contributing sources of fluvial inundation thus enhancing FIM skill.
 }
@@
 By using the same HEC-RAS derived discharges and FIM extents for creating maps with OWP FIM, we are able to separate out errors introduced by NWM inputs and processes including land surface interactions, groundwater fluxes, atmospheric forcings, hydraulic routing, and others that would have potentially affected our conclusions if we had used NWM forecasted discharges.
-Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
+Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE flowpaths as well as the BLE cross-sections that have recurrence discharges associated with them.
 We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
@@
 \caption{
-Illustrates Base Level Engineering (BLE) cross sections and stream lines at the HUC8 12100203 near the confluences of West Fork Plum Creek and Clear Fork Plum Creek with Plum Creek.
-BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM V2.1 Full Resolution (FR) stream lines.
-Additionally, we illustrate the NWM V2.1 catchments to provide a sense of how many cross-sections may intersect a given NWM flowline.
-The BLE stream network is also shown which is denser than the NWM V2.1 stream lines meaning there are several lower order streams represented in the BLE stream network that are not in the NWM V2.1 stream lines.
+Illustrates Base Level Engineering (BLE) cross sections and flowpaths at the HUC8 12100203 near the confluences of West Fork Plum Creek and Clear Fork Plum Creek with Plum Creek.
+BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM V2.1 Full Resolution (FR) flowpaths.
+Additionally, we illustrate the NWM V2.1 catchments to provide a sense of how many cross-sections may intersect a given NWM flowpaths.
+The BLE stream network is also shown which is denser than the NWM V2.1 flowpaths meaning there are several lower order streams represented in the BLE stream network that are not in the NWM V2.1 flowpaths.
 This creates additional inundation areas in the validation data that are not modeled with our HAND based FIMs.
@@
 In other words, POD is an indicator of model skill in inundated regions while FAR is an indicator of model skill in non-inundated regions.
-Some work by \citeA{gerapetritis2004behavior} denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown to be optimized mathematically when $POD = 1 - FAR$.
+Some work by \citeA{gerapetritis2004behavior} denotes CSI a good proxy for measuring a forecasting system's utility in protecting life and property and has been shown as mathematically optimized when $POD = 1 - FAR$.
 We use all three secondary metrics here to add value to the discussion while avoiding aggregating away the meaning of all four primary metrics.
@@
 These results are presented in Figure \ref{fig:violin_plot} as violin plots and in Table \ref{tab:aggregate_metrics} as aggregated metrics with the results for MS and GMS presented as percentage changes from their respective FR values.
-To be more specific, Table \ref{tab:aggregate_metrics} sums the primary metrics, TP, FP, FN, and TN, across all HUC8s then recomputes the secondary metrics which was done to better account for large variances in HUC8 size.
+More specifically, Table \ref{tab:aggregate_metrics} sums the primary metrics, TP, FP, FN, and TN, across all HUC8s then recomputes the secondary metrics which was done to better account for large variances in HUC8 size.
 The same trends discussed below are consistent across both reporting methods (Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}).
@@
 As previously stated in Section \ref{ssec:evaluation}, we consider CSI as a general proxy for the skill of the inundation extents with POD denoting skill on inundated areas and FAR indicating skill on non-inundated areas.
-Again, the main objective of the study is to introduce how computing HAND with disaggregated stream networks to those with unit stream order can enhance the fidelity of FIMs by capturing fluvial inundation from multiple sources as opposed to that of just the nearest drainage line.
+Again, the main objective of the study is to introduce how computing HAND with disaggregated stream networks to those with unit stream order can enhance the fidelity of FIMs by capturing fluvial inundation from multiple sources as opposed to that of just the nearest flowpath.
 As can be seen in Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, CSI generally increases from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
@@
 This can be inferred by close inspection of Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} where changes in POD are significantly higher than that of FAR.
-This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown to be a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
+This intuition is confirmed by previously mentioned work by \citeA{gerapetritis2004behavior} where CSI was shown as a direct function of both POD and FAR where CSI is maximized when $POD = 1 - FAR$.
 Upon investigation of the performance of HAND derived FIM, we note a general increase of catchment sizes for the MS and GMS methods when compared to the FR method as they are now delineated independently of any tributaries that would constrain catchment sizes otherwise.
@@
 As an enhancement, this paper proposes computing HAND for stream networks comprised of level-paths independently of one another.
-In sub-figure \ref{fig:gms_enhancement}b, the agreement raster for the GMS technique is illustrated as well as the stream network lines in green.
+In sub-figure \ref{fig:gms_enhancement}b, the agreement raster for the GMS technique is illustrated as well as the flowpaths in green.
 While the entire mosaiced inundation map from GMS (as described in Section \ref{ssec:inundation_mapping} and Equation \ref{eq:hand_fim_depth}) is used to produce this agreement map, we only show the catchments associated with the mainstem of this region that is shown to follow a clear horizontal path.
@@
 Since they are not restricted by the catchments of tributaries that lie in the same drainage areas as those of the mainstem, they extend to include those as well.
-The consequence for inundation extents is a general increase in spatial coverage of the river's water which shows to be in much better agreement with the benchmark map.
+The consequence for inundation extents is a general increase in spatial coverage of the river's water which shows better agreement with the benchmark maps.
 The TPs are no longer bounded by the catchment lines and allowed to expand to their natural extents.
@@
 However, we know that fluvial inundation can be sourced from several streams nearby that also serve as drainage outlets to the area in question.
-Generally speaking, drainage areas are known to be hierarchical in nature so a given drainage area for a given outlet point can be decomposed into nested drainage areas for outlet points that lie in the original drainage area.
+Generally speaking, drainage areas are hierarchical in nature so a given drainage area of an outlet point can be decomposed into nested drainage areas for outlet points that lie in the original drainage area.
 A perfect example of this are points that lie on tributary reaches closely neighboring a mainstem.
@@
 Hence it's important to state that just like there are different sources of flooding such as fluvial, pluvial, groundwater, reservoir, barrier failure (dam/levee/embankment), and coastal, there can also be multiple sources of a riverine flood.
-HAND is only equipped to handle riverine flooding from the nearest drainage line.
-Other relevant drainage lines that produce fluvial flood waters are not considered here especially if the routing model used doesn't consider backwater effects.
+HAND is only equipped to handle riverine flooding from the nearest flowpaths.
+Other relevant flowpaths that produce fluvial flood waters are not considered here especially if the routing model used doesn't consider backwater effects.
 
@@
 Across the three methods explored, FR, MS, and GMS, one identifies differences in the inputs and outputs and notes no difference in the stages and Manning's n values.
-While the channel slope and reach lengths are not exactly the same across methods, their averaged differences were found to be statistically insignificant which only leaves room for deviations in volume and bed area.
+While the channel slope and reach lengths are not exactly the same across methods, their averaged differences were statistically insignificant which only leaves room for deviations in volume and bed area.
 Specifically, we found FR to have an average reach length for the study region of 1354.8 km with a standard deviation 249.6km.
@@
 Until a singular stream network is used for the NWM, BLE benchmark, and for HAND based FIM, conflation will continue being a source of error.
+Future work could aim at producing a singular stream network that complies with the Open Geospatial Consortium (OGC) WaterML2.0 HY\_FEATURES data model for standard water feature representation \cite{simons2014waterml,boston2019water,looser2014international,blodgett2018ogc}.
 
 Our qualitative analysis suggests that the SRCs offer a significant opportunity for improvement in HAND based FIM for future development.
-The bathymetry of the 10 m DEM from 3DEP is known to be lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
+The bathymetry of the 10 m DEM from 3DEP is known as lacking proper representation thus leading to inadequate representation of volume and bed area with all three methods employed.
 Manning's n which typically accounts for roughness could be tuned to account for these DEM limitations or could be held fixed to some local value associated with a given flood magnitude.
@@
 Lastly, after errors introduced by conflation, poor roughness estimation, bathymetric/elevation adjustment, and reservoir representation are accounted for, HAND still has another fundamental limitation that is inherently baked into how it works.
-For HAND to be derived and thus create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
+To derive HAND and create a FIM for a given area, that area must entirely drain to the stream network and the stream network must also drain itself.
 In other words, an entire area eligible for flooding must monotonically decrease in elevation. 
@@
 \caption{OWP FIM inundation agreement, TP, FP, FN, and TN, with BLE HEC-RAS maps in HUC 11140105 at the 500 yr recurrence magnitude.
-Catchment boundaries and stream lines are shown in white and dotted green, respectively.
+Catchment boundaries and flowpaths are shown in white and dotted green, respectively.
 Sub-figure (a) shows agreement of FR HAND denoting significant areas of under-prediction due to junctions and catchment boundaries.
@@
 %
-The main contribution and conclusion of this work centers around a fundamental limitation in HAND based FIM which is a failure to account for multiple possible sources of fluvial inundation since HAND only considers inundation from the nearest drainage line.
+The main contribution and conclusion of this work centers around a fundamental limitation in HAND based FIM which is a failure to account for multiple possible sources of fluvial inundation since HAND only considers inundation from the nearest flowpath.
 We illustrate that reducing the Horton-Strahler stream order of a HAND processing unit down to one enhances skill by significantly reducing false negatives at junctions of major streams.
@@
 All LP based FIM derived from LP based HAND datasets are mosaiced together to form one seamless FIM.
-Dissecting the stream network into regions of LPs with unit stream order is necessary because HAND has a ``nearest drainage'' limitation meaning it only accounts for riverine inundation sourced from the nearest drainage line.
+Dissecting the stream network into regions of LPs with unit stream order is necessary because HAND has a ``nearest drainage'' limitation meaning it only accounts for riverine inundation sourced from the nearest flowpath.
 In our evaluation of this technique, we observe that HAND based FIM improves in skill as we extend from nearest drainage inundation in FR to multiple drainage support in MS for only 4-5\% of the FR network.
@@
 Additionally, we noted better performance for more intense 500 yr events which we attribute to a stronger influence of poor quality bathymetric data in 100 yr magnitude inundation extents.
-While the AGREE DEM procedure is meant to add some bathymetry primarily motivated to enhance catchment and stream line delineation, it does introduce three parameters that have major implications in the quality of SRCs and the resulting FIMs.
+While the AGREE DEM procedure is meant to add some bathymetry primarily motivated to enhance catchment and flowpath delineation, it does introduce three parameters that have major implications in the quality of SRCs and the resulting FIMs.
 Utilizing the highest resolution Lidar and bathymetric data should also improve the vertical accuracy of HAND and better account for fine grain features that greatly affect inundation extents.
@@
 %
-National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs \cite{nwm2022hydrofabric}.
+National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, flowpaths, and reservoirs \cite{nwm2022hydrofabric}.
 These are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
@@
 Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
-Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
-
-Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022}.
+Additionally, some FIM hydrofabric data including HAND grids, catchments, flowpaths, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
+
+Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository and a HydroShare resource called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022,imHS2023}.
  %
@@
 We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden, for his technical expertise.
-Additionally, David Blodgett from the United States Geolgical Survey (USGS) Water Mission Area was instrumental in helping define level paths and other hydrographic work.
-More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
+Additionally, David Blodgett from the United States Geological Survey (USGS) Water Mission Area was instrumental in helping define level paths and other hydrographic work.
+More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub and HydroShare \cite{inundationMapping2022,imHS2023}.
 Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.
@@
 
-However, to better match the drainage density of the NWM FR V2.1 stream network, which is based on the NHDPlus V2, the BurnlineEvents are reduced in density using a nearest neighbor search around the NWM flowlines.
+However, to better match the drainage density of the NWM FR V2.1 stream network, which is based on the NHDPlus V2, the BurnlineEvents are reduced in density using a nearest neighbor search around the NWM flowpaths.
 Headwater points are first derived for the NWM FR V2.1.
@@
 Additionally, the remaining NHDPlusHR headwaters are later used for seeding a new FIM drainage network that best agrees with the DEM after all hydro-conditioning takes place.
-This results in a stream network that has the same density as the NWM V2.1 flowline network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents. 
+This results in a stream network that has the same density as the NWM V2.1 flowpath network but utilizes the locations of the NHDPlusHR Beta BurnLineEvents. 
 %
@@
 \includegraphics[scale=1.0]{figures/headwaters.jpg}
-\caption{This figure illustrates some of the datasets that result from the pruning of NHDPlusHR Beta BurnlineEvents (dotted black) to the stream density of the NWM FR V2.1 density (blue).
+\caption{This figure illustrates some of the datasets that result from the drainage density reduction of the NHDPlusHR Beta BurnlineEvents (dotted black) to the stream density of the NWM FR V2.1 density (blue).
 The stream network used for forecasting, NWM FR V2.1, is of lower stream density than that of the NHDPlusHR which has better agreement with the thalweg locations in the DEM used.
@@
 In total, the AGREE algorithm requires three parameters including the buffer distance, smooth drop, and sharp drop which were set to fixed values of 70 m, 1000 m, and 10 m, respectively, but available to the user via the parameter file.
-While the values for these parameters are critical to the inundation extents produced, especially for lower flow rates where bathymetric information has more influence, finding their optimal values for OWP FIM was not done since it was out of the main scientific scope of this article.
-Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison}.
+As for the buffer distance, we conducted analysis varying this parameter from 50 to 90 m in 10 m increments and found that the 70 m value as a good compromise in improvement \cite{grout2020althydrocond}.
+While the values for these parameters are critical to the inundation extents produced, especially for lower flow rates where bathymetric information has more influence, we left further examination of these parameters and their influence on FIM skill out of the scope of this work.
+Using the AGREE method as opposed to simple thalweg burning techniques helps prevent distortions in the delineation of streams as well as the catchment boundaries \cite{saunders1995grid,saunders1996gis,mizgalewicz1996modeling,hellweger1997agree,quenzer1998gis,baker2006comparison,grout2020althydrocond}.
 \citeA{baker2006comparison} noted AGREE produced satisfactory results when compared to other enforcement techniques especially when computational costs are considered. 
@@
 The established literature in this niche field of hydrology discusses how prevalent flats can be in given study areas and how difficult the problem is from both computational and hydrologic stand-points \cite{garbrecht1997assignment,tarboton2009generalized,tarboton2005terrain,survila2016scalable,barnes2014efficient,tesfa2011extraction,wallis2009parallel}.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
+OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found efficient and effective \cite{survila2016scalable,cybergis2016}. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-04-09 manuscripts/owp_fim4_2022/reviews/second_review/responses_aristizabal_et_al_202302.tex
submitted to WRR April 2023

@@
 Based on your suggestion, we have added a few sentences that better discusses the conflation errors that could occur by using BLE associated streamflows with the NWM network.
-These sentences we placed near the end of the paragraph you suggested the discussion be located (INSERT LINE NUMBERS PRIOR TO SUBMISSION). 
+These sentences we placed near the end of the paragraph you suggested the discussion be located around line 1030. 
 \end{reply}
@@
 The two citations you provided reviewed or mentioned this issue in detail so we did cite these references throughout the paper.
-While cited throughout the paper for various reasons, they were used in the context of reach length on or near lines 150 and 492 of the previous submission (INPUT LINE NUMBERS PRIOR TO SUBMISSION).
-On line 490 of our previous submission, we mentioned that the reach length is capped to 1.5 km which is consistent with the findings from the Godbout paper and also consistent with what Zheng used within GeoFlood.
+While cited throughout the paper for various reasons, they were or are now used in the context of reach length on lines 149-150, 485-486, and 565.
+On line 485, we mention that the reach length is capped to 1.5 km which is consistent with the findings from the Godbout paper and also consistent with what Zheng used within GeoFlood.
 Additionally, we kept this constraint for all versions of HAND that we computed including FR, MS, and GMS, thus we did not observe a notable difference in reach length across the three versions.
 Due to this, we did not include it in the discussion of the paper.
-Nevertheless, we decided to add a sentence on (INPUT LINE NUMBERS PRIOR TO SUBMISSION) acknowledging that this parameter was held fixed at 1.5 km for all versions of HAND.
+Nevertheless, we decided to add a sentence on 486 acknowledging that this parameter was held fixed at 1.5 km for all versions of HAND.
 This constraint is very important since we did not find a significant difference in the reach lengths.
@@
 We found this difference in reach lengths across FR and GMS to be statistically insignificant. 
-In accordance with this, we added these numbers specifically (INPUT LINE NUMBERS PRIOR TO SUBMISSION) to the discussion section to inform readers that reach length, albeit an important variable, does not contribute to variance in SRC values across HAND versions.
+In accordance with this, we added these numbers specifically starting on line 983 to the discussion section to inform readers that reach length, albeit an important variable, does not contribute to variance in SRC values across HAND versions.
 \end{reply}

### 2023-04-09 manuscripts/owp_fim4_2022/sections/acknowledgments.tex
second review responses and changes submitted to coauthors 20230211
submitted to WRR April 2023

@@
 Lynker, under contract with OWP, facilitated this work and computational resources used in research and development.
-We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden for his technical expertise.
-Additionally, David Blodgett from the United States Geolgical Survey (USGS) Water Mission Area was instrumental in helping define level paths and other hydrographic work.
-More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub \cite{inundationMapping2022}.
+We would like to thank some notable contributors of this work including Chief Scientist at OWP, Fred Ogden, for his technical expertise.
+Additionally, David Blodgett from the United States Geological Survey (USGS) Water Mission Area was instrumental in helping define level paths and other hydrographic work.
+More information on code availability, usage, and data retrieval for OWP FIM is available on GitHub and HydroShare \cite{inundationMapping2022,imHS2023}.
 Thanks to the Earth and Space Science Informatics Partnership (ESIP) for storing data from this study for public use and dissemination helping to provide transparent datasets for further collaboration with the research community \cite{esipData2022}.

### 2023-04-09 manuscripts/owp_fim4_2022/sections/appendix.tex
submitted to WRR April 2023

@@
 The established literature in this niche field of hydrology discusses how prevalent flats can be in given study areas and how difficult the problem is from both computational and hydrologic stand-points \cite{garbrecht1997assignment,tarboton2009generalized,tarboton2005terrain,survila2016scalable,barnes2014efficient,tesfa2011extraction,wallis2009parallel}.
-OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found to be very efficient and effective \cite{survila2016scalable,cybergis2016}.
+OWP FIM utilized a CyberGIS implementation of the D-8 flow direction algorithm with the accelerated resolution of flats which we found efficient and effective \cite{survila2016scalable,cybergis2016}.

### 2023-04-09 manuscripts/owp_fim4_2022/sections/openresearch.tex
Resubmitted to WRR
submitted to WRR April 2023

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\openresearch
+\section*{Open Research}
+%\openresearch
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
-Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
+Additionally, some FIM hydrofabric data including HAND grids, catchments, flowpaths, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.
 
-Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022}.
+Software used in preprocessing data, producing FIM hydrofrabic, generating FIM, computing metrics, and conducting analysis is available from a publicly available Github repository and a HydroShare resource called ``inundation-mapping'' from the ``NOAA-OWP'' organization \cite{inundationMapping2022,imHS2023}.

### 2023-04-09 manuscripts/owp_fim4_2022/sections/title_authors.tex
submitted to WRR April 2023

--- 2022-08-21 manuscripts/owp_fim4_2022/sections/title_authors.tex
@@
 %
-\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{4}, Trevor Grout\affil{1,3}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{3,5}, Jasmeet Judge\affil{2}}
+\authors{Fernando Aristizabal\affil{1,2,3}, Fernando Salas\affil{3}, Gregory Petrochenkov\affil{1,3}, Trevor Grout\affil{1,3}\thanks{Colorado Basin River Forecast Center, National Weather Service, National Oceanic and Atmospheric Administration, Salt Lake City, UT, USA}, Brian Avant\affil{1,3}, Bradford Bates\affil{1,3}, Ryan Spies\affil{1,3}, Nick Chadwick\affil{3}, Zachary Wills\affil{1,3}, Jasmeet Judge\affil{2}}
 %
@@
 \affiliation{3}{National Water Center, Office of Water Prediction, National Weather Service, National Oceanic and Atmospheric Administration, Tuscaloosa, AL, USA}
-\affiliation{4}{New York Water Science Center, Hydrologic Applied Innovations Lab, United States Geological Survey, Troy, NY, USA}
-\affiliation{5}{Cooperative Institute for Satellite Earth System Studies, University of Maryland, College Park, MD, USA}
 %
